### 2018-08-08 positional_entropy_childes.Rmd
Dan's script

---
title: Positional entropy in childes
author: Joe Klafka
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

```{r setup, include = FALSE}
# load packages
library(knitr)
library(tidyverse)
library(directlabels)
library(childesr) #data
library(SnowballC) #stemmer
library(tidytext)
library(entropy)
library(tidyboot)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
```

```{r}
##functions
compile_child_utterances <- function(age_months) {
	# Given an age in months (integer), returns the child utterances from the Eng-NA corpus
	# for that age.
	d = get_utterances(collection="Eng-NA", age=age_months, role="Target_Child")
	return(d$gloss)
}

compile_adult_utterances <- function(age_months, adult_role) {
	# Given an age in months (integer), returns the adult_role utterances from the Eng-NA corpus
	# for that age (of the child to which the adult is speaking).
	# Valid are "Mother", "Investigator", "Sister", "Nurse", "Doctor" and "Father". 
	d = get_utterances(collection="Eng-NA", age=age_months, role=adult_role)
	return(d$gloss)
}

#IF WE CAN'T USE AN AGE RANGE
compile_ages <- function(age_range) {
	rv <- c()
	for (age in c(age_range[1]:age_range[2])) {
		rv <- c(rv, compile_child_utterances(age))
	}
	return(rv)
}
#IF WE CAN'T USE AN AGE RANGE
```

Get utterances
```{r get_utterances}
prov_utterances <- get_utterances(corpus = "Providence")

brown_utterances <- get_utterances(corpus = "Brown")

NA_utterances <- get_utterances(collection="Eng-NA")
```

compute estimates
```{r calc_pos_entropy}
get_entropies <- function(role, sen_length, utterances) {

  if(role == "Target_Child") {
    sub_utterances <- utterances %>%
      filter(speaker_role == "Target_Child") %>%
      mutate(length = str_count(stem, " ") + 1) # uses stems of words, not original words
  } else {
     sub_utterances <- utterances %>%
      filter(speaker_role != "Target_Child") %>%
      mutate(length = str_count(stem, " ") + 1)
  }

  # 
  # 
  # ggplot(adult_utterances, aes(x = length)) +
  #   geom_histogram(fill = "white", color = "black")

  tokens <- sub_utterances %>%
    filter(length == sen_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, stem) %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n())

  tokens %>%
    group_by(word_order, word) %>%
    summarise(n = n()) %>%
    tidyboot(summary_function = function(x) x %>% 
               summarise(entropy = entropy(n, unit = "log2")),
             statistics_functions = function(x) x %>%
             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
    mutate(role = role, length = sen_length)
}

entropies <- map(2:10, ~get_entropies("not child", .x, prov_utterances)) %>%
  bind_rows()

ggplot(entropies, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

entropies_child <- map(2:10, ~ get_entropies("Target_Child", .x, prov_utterances)) %>%
  bind_rows()

ggplot(entropies_child, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

entropies_all <- bind_rows(entropies, entropies_child)

ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper, color = role)) +
  facet_wrap(~ length) + 
  geom_pointrange(position = position_dodge(.25)) +
  geom_smooth(se = F) +
  ggtitle("Child and Adult Positional Entropy vs. Word Position (Providence corpus)")

```

### 2018-08-16 positional_entropy_childes.Rmd
Added bigram computations

@@
 library(tidyboot)
+library(dplyr)
+library(tokenizers)
+library(tau)
 
@@
 
+new_utterances <- prov_utterances %>% filter(speaker_role=="Target_Child") %>% mutate(len = str_count(stem, " ") + 1)
+#toks <- new_utterances %>%
+
+ggplot(new_utterances, aes(x=len)) +
+  geom_histogram(binwidth = 1)
 ```
+
+```{r ngram_frequencies}
+m_prov <- prov_utterances %>%
+          filter(speaker_role == "Target_Child") %>%
+          mutate(length = str_count(stem, " ") + 1) %>% 
+          filter(length >= 2)
+
+b_prov <- m_prov %>% mutate(bigrams = paste(unlist(textcnt(stem, n = 2, split = " ", method = "string")), collapse = ",")) #get all the counts of bigrams from all of the dataframe
+counts <- b_prov$bigrams[1] %>% strsplit(split=',') %>% unlist() %>% as.numeric()
+fileConn <- file("counts.txt")
+writeLines(b_prov$bigrams[1], fileConn)
+close(fileConn)
+
+b_prov <- m_prov %>% mutate(bigrams = paste(unlist(names(textcnt(stem, n = 2, split = " ", method = "string"))), collapse = ",")) #get all of the bigrams from the entire dataframe
+bigrams <- b_prov$bigrams[1] %>% strsplit(split=',') %>% unlist()
+fileConn <- file("bigram_words.txt")
+writeLines(b_prov$bigrams[1], fileConn)
+close(fileConn)
+
+```

### 2018-08-16 bigrams.py

@@ -1,0 +1,27 @@
+import re
+
+file = open("counts.txt", 'r')
+s = file.read()
+counts = re.findall(r'\d+', s)
+counts = [int(num) for num in counts]
+
+file = open("bigram_words.txt", 'r')
+s = file.read()
+bigrams = re.split(r',', s)
+
+word_dict = {}
+
+for i in range(len(bigrams)):
+    words = re.split(' ', bigrams[i])
+    if words[0] not in word_dict:
+        word_dict[words[0]] = {}
+        word_dict[words[0]]["TOTAL"] = 0
+    if words[1] not in word_dict[words[0]]:
+        word_dict[words[0]][words[1]] = 0
+    word_dict[words[0]][words[1]] += counts[i]
+    word_dict[words[0]]["TOTAL"] += counts[i]
+
+for key in word_dict:
+    div = word_dict[key]["TOTAL"]
+    for val in word_dict[key]:
+        word_dict[key][val] /= div

### 2018-08-16 positional_information_childes.Rmd
Initial information

---
title: Information Content Across Sentence Position in CHILDES
author: Joe Klafka
date: "8/16/2018"
output: html_document
---

```{r setup, include=FALSE}
# load packages
library(knitr)
library(tidyverse)
library(directlabels)
library(childesr) #data
library(SnowballC) #stemmer
library(tidytext)
library(entropy)
library(tidyboot)
library(dplyr)
library(tokenizers)
library(tau)
library(reticulate)
knitr::opts_chunk$set(echo = TRUE)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
```

```{r get_utterances}
prov_utterances <- get_utterances(corpus = "Providence")
```

```{r mutual information}

kid_info <- prov_utterances %>%
          filter(speaker_role == "Target_Child") %>%
          mutate(length = str_count(stem, " ") + 1) %>% 
          filter(length == 5)

```

```{python, engine.path="~/../../usr/local/bin/python3"}

```

### 2018-08-16 positional_entropy_childes.Rmd
mutual information

@@
 ```{r calc_pos_entropy}
-get_entropies <- function(role, sen_length, utterances) {
+get_counts <- function(role, sen_length, utterances) {
 
@@
 
-  tokens <- sub_utterances %>%
+  sub_utterances %>%
     filter(length == sen_length) %>%
@@
     mutate(word_order = 1:n())
+}
 
-  tokens %>%
+unigram_entropy <- function(counts) {
+
+  counts %>%
     group_by(word_order, word) %>%
@@
 
-entropies <- map(2:10, ~get_entropies("not child", .x, prov_utterances)) %>%
+counts <- map(2:6, ~get_counts("not child", .x, prov_utterances))
+
+all_pairwise_mi <- function(df) {
+  max_len <- max(df$word_order)
+
+  combs <- combn(1:max_len, 2, simplify = F)
+
+  map(combs, ~position_mi(df, .x[1], .x[2])) %>%
+    bind_rows()
+
+}
+
+position_mi <- function(df, pos1, pos2) {
+
+  mi <- df %>%
+    filter(word_order %in% c(pos1, pos2)) %>%
+    mutate(word_order = factor(word_order, labels = c("first", "second"))) %>%
+    spread(word_order, word) %>%
+    group_by(first, second) %>%
+    summarise(n = n()) %>%
+    spread(second, n, fill = 0) %>% 
+    ungroup() %>%
+    select(-first) %>%
+    mi.empirical(., unit = "log2")
+
+  data_frame(length = max(df$word_order), pos1 = pos1, pos2 = pos2, mi = mi)
+}
+
+mi <- map(counts, all_pairwise_mi)
+
+ggplot(mi, aes(x = pos2, y = mi, color = as.factor(pos1))) + 
+  facet_wrap(~length) + 
+  geom_point() + 
+  geom_line()
+
+ggplot(mi, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
+  facet_wrap(~length) + 
+  geom_point() + 
+  geom_line()
+
+entropies <- map(2, ~ get_counts("not child", .x, prov_utterances) %>%
+                   bind_rows %>%
+                   unigram_entropy) %>%
   bind_rows()

### 2018-08-16 Unigram Positional Entropy.Rmd
Original positional entropy file, plus Spanish unigrams

--- 2018-08-08 positional_entropy_childes.Rmd
@@
 ---
-title: Positional entropy in childes
+title: Unigram entropy in childes
 author: Joe Klafka
@@
 library(tidyboot)
+library(dplyr)
+library(tokenizers)
+library(tau)
+library(reticulate)
+knitr::opts_chunk$set(echo = TRUE)
 
@@
 
+new_utterances <- prov_utterances %>% filter(speaker_role=="Target_Child") %>% mutate(len = str_count(stem, " ") + 1)
+#toks <- new_utterances %>%
+
+ggplot(new_utterances, aes(x=len)) +
+  geom_histogram(binwidth = 1)
 ```
+
+```{r ngram_frequencies}
+m_prov <- prov_utterances %>%
+          filter(speaker_role == "Target_Child") %>%
+          mutate(length = str_count(stem, " ") + 1) %>% 
+          filter(length >= 2)
+
+b_prov <- m_prov %>% mutate(bigrams = paste(unlist(textcnt(stem, n = 2, split = " ", method = "string")), collapse = ",")) #get all the counts of bigrams from all of the dataframe
+counts <- b_prov$bigrams[1] %>% strsplit(split=',') %>% unlist() %>% as.numeric()
+fileConn <- file("bigram_counts.txt")
+writeLines(b_prov$bigrams[1], fileConn)
+close(fileConn)
+
+b_prov <- m_prov %>% mutate(bigrams = paste(unlist(names(textcnt(stem, n = 2, split = " ", method = "string"))), collapse = ",")) #get all of the bigrams from the entire dataframe
+bigrams <- b_prov$bigrams[1] %>% strsplit(split=',') %>% unlist()
+fileConn <- file("bigram_words.txt")
+writeLines(b_prov$bigrams[1], fileConn)
+close(fileConn)
+```
+
+```{r Spanish unigram frequencies}
+shiro_utterances <- get_utterances(language = "SPA", corpus="Shiro")
+
+get_noneng_entropies <- function(role, sen_length, utterances) {
+
+  if(role == "Target_Child") {
+    sub_utterances <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1) # uses stems of words, not original words
+  } else {
+     sub_utterances <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1)
+  }
+
+  # 
+  # 
+  # ggplot(adult_utterances, aes(x = length)) +
+  #   geom_histogram(fill = "white", color = "black")
+
+  tokens <- sub_utterances %>%
+    filter(length == sen_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, gloss) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n())
+
+  tokens %>%
+    group_by(word_order, word) %>%
+    summarise(n = n()) %>%
+    tidyboot(summary_function = function(x) x %>% 
+               summarise(entropy = entropy(n, unit = "log2")),
+             statistics_functions = function(x) x %>%
+             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
+    mutate(role = role, length = sen_length)
+}
+
+entropies <- map(2:10, ~get_entropies("not child", .x, shiro_utterances)) %>%
+  bind_rows()
+
+ggplot(entropies, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F)
+
+entropies_child <- map(2:10, ~ get_entropies("Target_Child", .x, shiro_utterances)) %>%
+  bind_rows()
+
+ggplot(entropies_child, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F)
+
+entropies_all <- bind_rows(entropies, entropies_child)
+
+ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper, color = role)) +
+  facet_wrap(~ length) + 
+  geom_pointrange(position = position_dodge(.25)) +
+  geom_smooth(se = F) +
+  ggtitle("Child and Adult Positional Entropy vs. Word Position (Spanish Shiro corpus)")
+```

### 2018-08-17 Unigram Positional Entropy.Rmd
Unigram entropy for Spanish, German and Japanese, plus conditional entropy for Brown corpus of English

@@
 shiro_utterances <- get_utterances(language = "SPA", corpus="Shiro") 
+#Venezuelan children narrative (113 children; half 1st graders half 4th graders; half high SES half low SES)
+
+wagner_utterances <- get_utterances(corpus = "Wagner") #German corpus of children during daily routines (age range from 1,5 to 14,10)
+
+okayama_utterances <- get_utterances(corpus = "Okayama") #mother-child conversations in the Osaka area
 
@@
       filter(speaker_role == "Target_Child") %>%
-      mutate(length = str_count(gloss, " ") + 1) # uses stems of words, not original words
+      mutate(length = str_count(gloss, " ") + 1) 
   } else {
@@
     tidyboot(summary_function = function(x) x %>% 
-               summarise(entropy = entropy(n, unit = "log2")),
+               summarise(entropy = entropy.empirical(n, unit = "log2")),
              statistics_functions = function(x) x %>%
@@
 
-entropies <- map(2:10, ~get_entropies("not child", .x, shiro_utterances)) %>%
-  bind_rows()
+entropies <- map(2:10, ~get_noneng_entropies("not child", .x, okayama_utterances)) %>%
+  bind_rows() %>% filter(word_order <= length)
 
@@
 
-entropies_child <- map(2:10, ~ get_entropies("Target_Child", .x, shiro_utterances)) %>%
-  bind_rows()
+entropies_child <- map(2:10, ~ get_noneng_entropies("Target_Child", .x, okayama_utterances)) %>%
+  bind_rows() %>% filter(word_order <= length)
 
@@
   geom_smooth(se = F) +
-  ggtitle("Child and Adult Positional Entropy vs. Word Position (Spanish Shiro corpus)")
+  ggtitle("Child and Adult Positional Entropy vs. Word Position (Japanese Okayama corpus)")
 ```

### 2018-08-17 positional_entropy_childes.Rmd
Unigram entropy for Spanish, German and Japanese, plus conditional entropy for Brown corpus of English

@@
 library(tau)
+library(gtools)
 
@@
 
+```{r conditional_entropy}
+get_counts <- function(role, sen_length, utterances) {
+
+  if(role == "Target_Child") {
+    sub_utterances <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(length = str_count(stem, " ") + 1) # uses stems of words, not original words
+  } else {
+     sub_utterances <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      mutate(length = str_count(stem, " ") + 1)
+  }
+
+  # 
+  # 
+  # ggplot(adult_utterances, aes(x = length)) +
+  #   geom_histogram(fill = "white", color = "black")
+
+  sub_utterances %>%
+    filter(length == sen_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, stem) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n())
+}
+
+unigram_entropy <- function(counts) {
+
+  counts %>%
+    group_by(word_order, word) %>%
+    summarise(n = n()) %>%
+    tidyboot(summary_function = function(x) x %>% 
+               summarise(entropy = entropy(n, unit = "log2")),
+             statistics_functions = function(x) x %>%
+             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
+    mutate(role = role, length = sen_length)
+}
+
+counts <- map(2:6, ~get_counts("not child", .x, prov_utterances))
+
+all_conditional_entropy <- function(df) {
+  max_len <- max(df$word_order)
+  #combs <- combn(1:max_len, 2, simplify = F)
+  perms <- permutations(max_len, 2, 1:max_len, repeats.allowed = F) 
+  combs <-  setNames(split(perms, seq(nrow(perms))), rownames(perms))
+
+  map(combs, ~conditional_entropy(df, .x[1], .x[2])) %>%
+    bind_rows()
+
+}
+
+conditional_entropy <- function(df, X, Y) { #Y is conditioning variable
+
+  d <- df %>%
+    filter(word_order %in% c(X, Y)) %>%
+    mutate(word_order = factor(word_order, labels = c("first", "second"))) %>%
+    spread(word_order, word)
+
+  ent <- d$first %>% table() %>% as.vector() %>% entropy.empirical(unit="log2")
+
+  mi <- d %>% 
+    group_by(first, second) %>%
+    summarise(n = n()) %>%
+    spread(second, n, fill = 0) %>% 
+    ungroup() %>%
+    select(-first) %>%
+    mi.empirical(., unit = "log2")
+
+  data_frame(length = max(df$word_order), X = X, Y = Y, mi = mi, ent = ent, ce = ent - mi)
+}
+
+ce <- map(counts, all_conditional_entropy)
+
+ggplot(mi, aes(x = pos2, y = mi, color = as.factor(pos1))) + 
+  facet_wrap(~ length) + 
+  geom_point() + 
+  geom_line()
+
+ggplot(ce[[5]], aes(x = X, y = ce, color = as.factor(Y))) + 
+  facet_wrap(~ length) + 
+  geom_point() + 
+  geom_line() + 
+  ggtitle("Conditional entropy vs. conditioned random variable (utterances of length 6)")
+
+```
+
 ```{r ngram_frequencies}

### 2018-08-20 Unigram Positional Entropy.Rmd
Updated non-English to include stemming

@@
 
-get_noneng_entropies <- function(role, sen_length, utterances) {
+get_noneng_entropies <- function(role, sen_length, utterances, langue) {
 
@@
       filter(speaker_role == "Target_Child") %>%
-      mutate(length = str_count(gloss, " ") + 1) 
+      mutate(stems = wordStem(gloss, language = langue)) %>% 
+      mutate(length = str_count(stems, " ") + 1) 
   } else {
@@
       filter(speaker_role != "Target_Child") %>%
-      mutate(length = str_count(gloss, " ") + 1)
+      mutate(stems = wordStem(gloss, language = langue)) %>% 
+      mutate(length = str_count(stems, " ") + 1)
   }
@@
     mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss) %>%
+    unnest_tokens(word, stems) %>%
     group_by(utterance_id) %>%
@@
 
-entropies <- map(2:10, ~get_noneng_entropies("not child", .x, okayama_utterances)) %>%
-  bind_rows() %>% filter(word_order <= length)
+entropies <- map(2:10, ~get_noneng_entropies("not child", .x, wagner_utterances, "german")) %>%
+  bind_rows() %>% filter(word_order <= length) #some utterances have ""
 
@@
 
-entropies_child <- map(2:10, ~ get_noneng_entropies("Target_Child", .x, okayama_utterances)) %>%
+entropies_child <- map(2:10, ~ get_noneng_entropies("Target_Child", .x, wagner_utterances, "german")) %>%
   bind_rows() %>% filter(word_order <= length)
@@
   geom_smooth(se = F) +
-  ggtitle("Child and Adult Positional Entropy vs. Word Position (Japanese Okayama corpus)")
+  ggtitle("Child and Adult Positional Entropy vs. Word Position (German Wagner corpus with stemming)")
 ```

### 2018-08-22 Unigram Positional Entropy.Rmd
Thoroughly knitted

--- 2018-08-17 Unigram Positional Entropy.Rmd
@@
 
-```{r}
-##functions
-compile_child_utterances <- function(age_months) {
-	# Given an age in months (integer), returns the child utterances from the Eng-NA corpus
-	# for that age.
-	d = get_utterances(collection="Eng-NA", age=age_months, role="Target_Child")
-	return(d$gloss)
-}
-
-compile_adult_utterances <- function(age_months, adult_role) {
-	# Given an age in months (integer), returns the adult_role utterances from the Eng-NA corpus
-	# for that age (of the child to which the adult is speaking).
-	# Valid are "Mother", "Investigator", "Sister", "Nurse", "Doctor" and "Father". 
-	d = get_utterances(collection="Eng-NA", age=age_months, role=adult_role)
-	return(d$gloss)
-}
-
-#IF WE CAN'T USE AN AGE RANGE
-compile_ages <- function(age_range) {
-	rv <- c()
-	for (age in c(age_range[1]:age_range[2])) {
-		rv <- c(rv, compile_child_utterances(age))
-	}
-	return(rv)
-}
-#IF WE CAN'T USE AN AGE RANGE
-```
-
 Get utterances
@@
 prov_utterances <- get_utterances(corpus = "Providence")
-
-brown_utterances <- get_utterances(corpus = "Brown")
-
-NA_utterances <- get_utterances(collection="Eng-NA")
 ```
 
-compute estimates
+Compute Providence Corpus Estimates
 ```{r calc_pos_entropy}
@@
 
+Compile all bigrams in corpus (save for another time) 
 ```{r ngram_frequencies}
@@
 
+Non-English Unigram frequencies
 ```{r Spanish unigram frequencies}
@@
 
+okayama_utterances <- get_utterances(corpus = "Okayama") #mother-child conversations in the Osaka area
+
+get_noneng_entropies <- function(role, sen_length, utterances, lang) {
+
+  if(role == "Target_Child") {
+    sub_utterances <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(stems = wordStem(gloss, language = lang)) %>% 
+      mutate(length = str_count(stems, " ") + 1) 
+  } else {
+     sub_utterances <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      mutate(stems = wordStem(gloss, language = lang)) %>% 
+      mutate(length = str_count(stems, " ") + 1)
+  }
+
+  # 
+  # 
+  # ggplot(adult_utterances, aes(x = length)) +
+  #   geom_histogram(fill = "white", color = "black")
+
+  tokens <- sub_utterances %>%
+    filter(length == sen_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, stems) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n())
+
+  tokens %>%
+    group_by(word_order, word) %>%
+    summarise(n = n()) %>%
+    tidyboot(summary_function = function(x) x %>% 
+               summarise(entropy = entropy.empirical(n, unit = "log2")),
+             statistics_functions = function(x) x %>%
+             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
+    mutate(role = role, length = sen_length)
+}
+
+entropies <- map(2:10, ~get_noneng_entropies("not child", .x, shiro_utterances, "spanish")) %>%
+  bind_rows() %>% filter(word_order <= length) #some utterances have ""
+
+ggplot(entropies, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F) +
+  ggtitle("Adult Unigram Entropy vs. Word Position (Spanish Shiro Corpus with stemming)")
+
+entropies_child <- map(2:10, ~ get_noneng_entropies("Target_Child", .x, shiro_utterances, "spanish")) %>%
+  bind_rows() %>% filter(word_order <= length)
+
+ggplot(entropies_child, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F) +
+  ggtitle("Child Unigram Entropy vs. Word Position (Spanish Shiro Corpus with stemming)")
+
+entropies_all <- bind_rows(entropies, entropies_child)
+
+ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper, color = role)) +
+  facet_wrap(~ length) + 
+  geom_pointrange(position = position_dodge(.25)) +
+  geom_smooth(se = F) +
+  ggtitle("Adult and Child Unigram Entropy vs. Word Position (Spanish Shiro Corpus with stemming)")
+```
+
+```{r German Unigram Frequencies}
 wagner_utterances <- get_utterances(corpus = "Wagner") #German corpus of children during daily routines (age range from 1,5 to 14,10)
 
-okayama_utterances <- get_utterances(corpus = "Okayama") #mother-child conversations in the Osaka area
+entropies <- map(2:10, ~get_noneng_entropies("not child", .x, wagner_utterances, "german")) %>%
+  bind_rows() %>% filter(word_order <= length) #some utterances have ""
 
-get_noneng_entropies <- function(role, sen_length, utterances) {
+ggplot(entropies, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F) +
+  ggtitle("Adult Unigram Entropy vs. Word Position (German Wagner Corpus with stemming)")
+
+entropies_child <- map(2:10, ~ get_noneng_entropies("Target_Child", .x, wagner_utterances, "german")) %>%
+  bind_rows() %>% filter(word_order <= length)
+
+ggplot(entropies_child, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F) +
+  ggtitle("Child Unigram Entropy vs. Word Position (German Wagner Corpus with stemming)")
+
+entropies_all <- bind_rows(entropies, entropies_child)
+
+ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper, color = role)) +
+  facet_wrap(~ length) + 
+  geom_pointrange(position = position_dodge(.25)) +
+  geom_smooth(se = F) +
+  ggtitle("Adult and Child Unigram Entropy vs. Word Position (German Wagner Corpus with stemming)")
+```
+
+```{r Japanese Unigram Entropy}
+get_japanese_entropies <- function(role, sen_length, utterances) {
 
@@
 
-entropies <- map(2:10, ~get_noneng_entropies("not child", .x, okayama_utterances)) %>%
-  bind_rows() %>% filter(word_order <= length)
+entropies <- map(2:10, ~get_japanese_entropies("not child", .x, okayama_utterances)) %>%
+  bind_rows() %>% filter(word_order <= length) #some utterances have "" counted for word_order
 
@@
   geom_pointrange() +
-  geom_smooth(se = F)
+  geom_smooth(se = F) +
+  ggtitle("Adult Unigram Entropy vs. Word Position (Japanese Okayama Corpus)")
 
-entropies_child <- map(2:10, ~ get_noneng_entropies("Target_Child", .x, okayama_utterances)) %>%
+entropies_child <- map(2:10, ~ get_japanese_entropies("Target_Child", .x, okayama_utterances)) %>%
   bind_rows() %>% filter(word_order <= length)
@@
   geom_pointrange() +
-  geom_smooth(se = F)
+  geom_smooth(se = F) +
+  ggtitle("Child Unigram Entropy vs. Word Position (Japanese Okayama Corpus)")
 
@@
   geom_smooth(se = F) +
-  ggtitle("Child and Adult Positional Entropy vs. Word Position (Japanese Okayama corpus)")
+  ggtitle("Adult and Child Unigram Entropy vs. Word Position (Japanese Okayama Corpus)")
 ```

### 2018-08-23 Unigram Positional Entropy.Rmd
Added Mandarin unigram entropy computations

@@
 ```
+
+```{r Chinese Unigram Entropy}
+zhou_dinner <- get_utterances(corpus = "zhoudinner")
+
+get_chinese_entropies <- function(role, sen_length, utterances) {
+
+  if(role == "Target_Child") {
+    sub_utterances <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      filter(gloss != "") %>% 
+      mutate(length = str_count(stem, " ") + 1) 
+  } else {
+     sub_utterances <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      filter(speaker_role != "") %>% 
+      mutate(length = str_count(stem, " ") + 1)
+  }
+
+  # 
+  # 
+  # ggplot(adult_utterances, aes(x = length)) +
+  #   geom_histogram(fill = "white", color = "black")
+
+  tokens <- sub_utterances %>%
+    filter(length == sen_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, stem) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n())
+
+  tokens %>%
+    group_by(word_order, word) %>%
+    summarise(n = n()) %>%
+    tidyboot(summary_function = function(x) x %>% 
+               summarise(entropy = entropy.empirical(n, unit = "log2")),
+             statistics_functions = function(x) x %>%
+             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
+    mutate(role = role, length = sen_length)
+}
+
+entropies <- map(2:10, ~get_chinese_entropies("not child", .x, zhou_dinner)) %>%
+  bind_rows() %>% filter(word_order <= length) #some utterances have "" counted for word_order
+
+ggplot(entropies, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F) +
+  ggtitle("Adult Unigram Entropy vs. Word Position (Mandarin Zhou Dinner Corpus)")
+
+entropies_child <- map(2:10, ~ get_chinese_entropies("Target_Child", .x, zhou_dinner)) %>%
+  bind_rows() %>% filter(word_order <= length)
+
+ggplot(entropies_child, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F) +
+  ggtitle("Child Unigram Entropy vs. Word Position (Mandarin Zhou Dinner Corpus)")
+
+entropies_all <- bind_rows(entropies, entropies_child)
+
+ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper, color = role)) +
+  facet_wrap(~ length) + 
+  geom_pointrange(position = position_dodge(.25)) +
+  geom_smooth(se = F) +
+  ggtitle("Adult and Child Unigram Entropy vs. Word Position (Mandarin Zhou Dinner Corpus)")
+```

### 2018-08-24 positional_entropy_childes.Rmd
Updated non-English to include stemming
Issues with ggplot2 displaying lists

@@
 
-ggplot(mi, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
+ggplot(mi[[4]], aes(x = pos1, y = mi, color = as.factor(pos2))) + 
   facet_wrap(~length) + 
@@
 ```{r conditional_entropy}
-get_counts <- function(role, sen_length, utterances) {
-
-  if(role == "Target_Child") {
-    sub_utterances <- utterances %>%
-      filter(speaker_role == "Target_Child") %>%
-      mutate(length = str_count(stem, " ") + 1) # uses stems of words, not original words
-  } else {
-     sub_utterances <- utterances %>%
-      filter(speaker_role != "Target_Child") %>%
-      mutate(length = str_count(stem, " ") + 1)
-  }
-
-  # 
-  # 
-  # ggplot(adult_utterances, aes(x = length)) +
-  #   geom_histogram(fill = "white", color = "black")
-
-  sub_utterances %>%
-    filter(length == sen_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, stem) %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n())
-}
-
-unigram_entropy <- function(counts) {
-
-  counts %>%
-    group_by(word_order, word) %>%
-    summarise(n = n()) %>%
-    tidyboot(summary_function = function(x) x %>% 
-               summarise(entropy = entropy(n, unit = "log2")),
-             statistics_functions = function(x) x %>%
-             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
-    mutate(role = role, length = sen_length)
-}
 
@@
 
-ggplot(mi, aes(x = pos2, y = mi, color = as.factor(pos1))) + 
+ggplot(ce, aes(x = X, y = ce, color = as.factor(X))) + 
   facet_wrap(~ length) + 
@@
 
-ggplot(ce[[5]], aes(x = X, y = ce, color = as.factor(Y))) + 
+ggplot(ce[[3]], aes(x = X, y = ce, color = as.factor(Y))) + 
   facet_wrap(~ length) + 
@@
   geom_line() + 
-  ggtitle("Conditional entropy vs. conditioned random variable (utterances of length 6)")
+  ggtitle("Conditional entropy vs. conditioned random variable (utterances of length 5)")
 

### 2018-08-30 letter_entropy.Rmd
CE and MI for non-English, plus English letter entropy (without graphs)

---
title: Letter entropy in childes
author: Joe Klafka
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

```{r setup, include = FALSE}
# load packages
library(knitr)
library(tidyverse)
library(directlabels)
library(childesr) #data
library(SnowballC) #stemmer
library(tidytext)
library(entropy)
library(tidyboot)
library(dplyr)
library(tokenizers)
library(tau)
knitr::opts_chunk$set(echo = TRUE)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
```

Get utterances
```{r get_utterances}
prov_utterances <- get_utterances(corpus = "Providence")
```

Compute Providence Corpus Estimates
```{r calc_pos_entropy}
get_counts <- function(role, sen_length, utterances) {

  if(role == "Target_Child") {
    sub_utterances <- utterances %>%
      filter(speaker_role == "Target_Child") %>%
      mutate(length = str_count(stem, " ") + 1) # uses stems of words, not original words
  } else {
     sub_utterances <- utterances %>%
      filter(speaker_role != "Target_Child") %>%
      mutate(length = str_count(stem, " ") + 1)
  }

  # 
  # 
  # ggplot(adult_utterances, aes(x = length)) +
  #   geom_histogram(fill = "white", color = "black")

  tokens <- sub_utterances %>%
              filter(length == sen_length) %>%
              mutate(utterance_id = 1:n()) %>%
              unnest_tokens(word, stem) %>%
              group_by(utterance_id) %>%
              mutate(word_order = 1:n())
}

get_letters <- function(tokens, word_choice) {

  letters <- tokens %>% 
              filter(word_order == word_choice) %>% 
              unnest_tokens(letter, word, token = "characters") %>% 
              group_by(utterance_id) %>% 
              mutate(letter_order = 1:n())

  letter_counts <- letters %>% 
                    group_by(letter_order, letter) %>% 
                    summarize(n = n())

  map(1:10, ~letter_entropy(letter_counts, .x)) 
}

letter_entropy <- function(letter_counts, position) {
  position_counts <- letter_counts %>% filter(letter_order == position)

  entropie <- position_counts$n %>% entropy.empirical(unit = "log2")
}

get_letter_entropies <- function(role, sen_length, utterances) {
  tokens <- get_counts(role, sen_length, utterances)
  map(1:sen_length, ~get_letters(tokens,.x))
}

#currently have list of lists, where the top list is for a word_order and each bottom list is for a letter_order
```

### 2018-08-30 nonenglish_ce&mi.Rmd
CE and MI for non-English, plus English letter entropy (without graphs)

---
title: "Untitled"
author: "Joe Klafka"
date: "8/28/2018"
output: html_document
---

```{r setup, include=FALSE}
# load packages
library(knitr)
library(tidyverse)
library(directlabels)
library(childesr) #data
library(SnowballC) #stemmer
library(tidytext)
library(entropy)
library(tidyboot)
library(dplyr)
library(tokenizers)
library(tau)
library(gtools)
knitr::opts_chunk$set(echo = TRUE)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
```

```{r get_utterances}
prov_utterances <- get_utterances(corpus = "Providence")

brown_utterances <- get_utterances(corpus = "Brown")

zhou_dinner <- get_utterances(corpus = "zhoudinner")

shiro_utterances <- get_utterances(language = "SPA", corpus="Shiro") 
#Venezuelan children narrative (113 children; half 1st graders half 4th graders; half high SES half low SES)

okayama_utterances <- get_utterances(corpus = "Okayama") #mother-child conversations in the Osaka area

wagner_utterances <- get_utterances(corpus = "Wagner") #German corpus of children during daily routines (age range from 1,5 to 14,10)

miipro_utterances <- get_utterances(corpus = "miipro") #Tokyo mother-child conversations in the family's home

leveille_utterances <- get_utterances(corpus = "leveille") #longitudinal study of a French child of university profs at home

```

```{r presentation}

get_counts <- function(role, sen_length, utterances) {

  if(role == "Target_Child") {
    sub_utterances <- utterances %>%
      filter(speaker_role == "Target_Child") %>%
      mutate(length = str_count(gloss, " ") + 1) # uses stems of words, not original words
  } else {
     sub_utterances <- utterances %>%
      filter(speaker_role != "Target_Child") %>%
      mutate(length = str_count(gloss, " ") + 1)
  }

  sub_utterances %>%
    filter(length == sen_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, gloss) %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n())
}

all_conditional_entropy <- function(df) {
  max_len <- max(df$word_order)
  perms <- permutations(max_len, 2, 1:max_len, repeats.allowed = F) 
  combs <-  setNames(split(perms, seq(nrow(perms))), rownames(perms))

  map(combs, ~conditional_entropy(df, .x[1], .x[2])) %>%
    bind_rows()

}

conditional_entropy <- function(df, X, Y) { #Y is conditioning variable

  d <- df %>%
    filter(word_order %in% c(X, Y)) %>%
    mutate(word_order = factor(word_order, labels = c("first", "second"))) %>%
    spread(word_order, word)

  ent <- d$first %>% table() %>% as.vector() %>% entropy.empirical(unit="log2")

  mi <- d %>% 
    group_by(first, second) %>%
    summarise(n = n()) %>%
    spread(second, n, fill = 0) %>% 
    ungroup() %>%
    select(-first) %>%
    mi.empirical(., unit = "log2")

  data_frame(length = max(df$word_order), X = X, Y = Y, mi = mi, ent = ent, ce = ent - mi)
}

get_ce <- function(utterances, role) {

  ce <- map(2:6, ~get_counts(role, .x, utterances)) %>% 
    map(~filter(.,word_order <= length)) %>%
    map(all_conditional_entropy) %>%
    bind_rows()

  return(ce)
}

all_pairwise_mi <- function(df) {
  max_len <- max(df$word_order)

  combs <- combn(1:max_len, 2, simplify = F)

  map(combs, ~position_mi(df, .x[1], .x[2])) %>%
    bind_rows()

}

position_mi <- function(df, pos1, pos2) {

  mi <- df %>%
    filter(word_order %in% c(pos1, pos2)) %>%
    mutate(word_order = factor(word_order, labels = c("first", "second"))) %>%
    spread(word_order, word) %>%
    group_by(first, second) %>%
    summarise(n = n()) %>%
    spread(second, n, fill = 0) %>% 
    ungroup() %>%
    select(-first) %>%
    mi.empirical(., unit = "log2")

  data_frame(length = max(df$word_order), pos1 = pos1, pos2 = pos2, mi = mi)
}

get_mi <- function(utterances, role) {

  mi <- map(2:6, ~get_counts(role, .x, utterances)) %>% 
    map(~filter(.,word_order <= length)) %>%
    map(all_pairwise_mi) %>%
    bind_rows()

  return(mi)
}
```

```{r work_progress}

shiro_mi_children <- get_mi(shiro_utterances, "Target_Child")
shiro_mi_adults <- get_mi(shiro_utterances, "not child")

ggplot(shiro_mi_children, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
  facet_wrap(~length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Mutual information vs. conditioned position (Shiro children)")

ggplot(shiro_mi_adults, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
  facet_wrap(~length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Mutual information vs. conditioned position (Shiro adults)")

shiro_ce_children <- get_ce(shiro_utterances, "Target_Child")
shiro_ce_adults <- get_ce(shiro_utterances, "not child")

ggplot(shiro_ce_children, aes(x = X, y = ce, color = as.factor(Y))) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Conditional entropy vs. conditioned random variable (Shiro children)")

ggplot(shiro_ce_adults, aes(x = X, y = ce, color = as.factor(Y))) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Conditional entropy vs. conditioned random variable (Shiro adults)")

wagner_ce_children <- get_ce(wagner_utterances, "Target_Child")
wagner_ce_adults <- get_ce(wagner_utterances, "not child")

ggplot(wagner_ce_children, aes(x = X, y = ce, color = as.factor(Y))) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Conditional entropy vs. conditioned random variable (Wagner children)")

ggplot(wagner_ce_adults, aes(x = X, y = ce, color = as.factor(Y))) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Conditional entropy vs. conditioned random variable (Wagner adults)")

okayama_mi_children <- get_mi(okayama_utterances, "Target_Child")
okayama_mi_adults <- get_mi(okayama_utterances, "not child")

ggplot(okayama_mi_children, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
  facet_wrap(~length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Mutual information vs. conditioned position (Okayama children)")

ggplot(okayama_mi_adults, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
  facet_wrap(~length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Mutual information vs. conditioned position (Okayama adults)")

okayama_ce_children <- get_ce(okayama_utterances, "Target_Child")
okayama_ce_adults <- get_ce(okayama_utterances, "not child")

ggplot(okayama_ce_children, aes(x = X, y = ce, color = as.factor(Y))) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Conditional entropy vs. conditioned random variable (Okayama children)")

ggplot(okayama_ce_adults, aes(x = X, y = ce, color = as.factor(Y))) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Conditional entropy vs. conditioned random variable (Okayama adults)")

miipro_mi_children <- get_mi(miipro_utterances, "Target_Child")
miipro_mi_adults <- get_mi(miipro_utterances, "not child")

ggplot(miipro_mi_children, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
  facet_wrap(~length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Mutual information vs. conditioned position (Miipro children)")

ggplot(miipro_mi_adults, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
  facet_wrap(~length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Mutual information vs. conditioned position (Miipro adults)")

miipro_ce_children <- get_ce(miipro_utterances, "Target_Child")
miipro_ce_adults <- get_ce(miipro_utterances, "not child")

ggplot(miipro_ce_children, aes(x = X, y = ce, color = as.factor(Y))) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Conditional entropy vs. conditioned random variable (Miipro children)")

ggplot(miipro_ce_adults, aes(x = X, y = ce, color = as.factor(Y))) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Conditional entropy vs. conditioned random variable (Miipro adults)")

zhou_mi_children <- get_mi(zhou_dinner, "Target_Child")
zhou_mi_adults <- get_mi(zhou_dinner, "not child")

ggplot(zhou_mi_children, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
  facet_wrap(~length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Mutual information vs. conditioned position (Zhou children)")

ggplot(zhou_mi_adults, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
  facet_wrap(~length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Mutual information vs. conditioned position (Zhou adults)")

zhou_ce_children <- get_ce(zhou_dinner, "Target_Child")
zhou_ce_adults <- get_ce(zhou_dinner, "not child")

ggplot(zhou_ce_children, aes(x = X, y = ce, color = as.factor(Y))) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Conditional entropy vs. conditioned random variable (Zhou children)")

ggplot(zhou_ce_adults, aes(x = X, y = ce, color = as.factor(Y))) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Conditional entropy vs. conditioned random variable (Zhou adults)")

leveille_ce_children <- get_ce(leveille_utterances, "Target_Child")
leveille_ce_adults <- get_ce(leveille_utterances, "not child")

ggplot(leveille_ce_children, aes(x = X, y = ce, color = as.factor(Y))) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Conditional entropy vs. conditioned random variable (Leveille children)")

ggplot(leveille_ce_adults, aes(x = X, y = ce, color = as.factor(Y))) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_line() + 
  ggtitle("Conditional entropy vs. conditioned random variable (Leveille adults)")
```

### 2018-08-30 positional_entropy_childes.Rmd
CE and MI for non-English, plus English letter entropy (without graphs)

@@
 
-counts <- map(2:6, ~get_counts("not child", .x, prov_utterances))
+counts <- map(2:6, ~get_counts("not child", .x, brown_utterances))
 
@@
 
-ggplot(ce[[3]], aes(x = X, y = ce, color = as.factor(Y))) + 
+ggplot(ce, aes(x = X, y = ce, color = as.factor(Y))) + 
   facet_wrap(~ length) + 

### 2018-09-10 Unigram Positional Entropy.Rmd
letter entropies and touches

@@
 library(tau)
-library(reticulate)
 knitr::opts_chunk$set(echo = TRUE)

### 2018-09-10 letter_entropy.Rmd
letter entropies and touches

@@
 ---
-title: Letter entropy in childes
+title: By-letter entropy in childes
 author: Joe Klafka
@@
 
-  # 
-  # 
-  # ggplot(adult_utterances, aes(x = length)) +
-  #   geom_histogram(fill = "white", color = "black")
-
   tokens <- sub_utterances %>%
@@
 
-  map(1:10, ~letter_entropy(letter_counts, .x)) 
+  map(1:10, ~letter_entropy(letter_counts, .x)) ##first 10 letters, after this entropy shrunk close to 0
 }
@@
 
+prov_child_letters <- map(2:10, ~get_letter_entropies("Target_Child", .x, prov_utterances)) #list of lists of lists
+
 #currently have list of lists, where the top list is for a word_order and each bottom list is for a letter_order

### 2018-09-10 nonenglish_ce&mi.Rmd
letter entropies and touches

@@
 ---
-title: "Untitled"
-author: "Joe Klafka"
-date: "8/28/2018"
-output: html_document
+title: Non-English Conditional Entropy and Mutual Information
+author: Joe Klafka
+date: "`r Sys.date()`"
+output: 
+  html_document:
+    toc: false
+    number_sections: false
+    theme: lumen
+    toc_float: false
+    code_folding: show 
 ---

### 2018-09-10 positional_entropy_childes.Rmd
letter entropies and touches

--- 2018-08-24 positional_entropy_childes.Rmd
@@
 
-counts <- map(2:6, ~get_counts("not child", .x, prov_utterances))
+counts <- map(2:8, ~get_counts("not child", .x, prov_utterances))
 
@@
 
-ggplot(ce[[3]], aes(x = X, y = ce, color = as.factor(Y))) + 
+ggplot(ce_br, aes(x = X, y = ce, color = as.factor(Y))) + 
   facet_wrap(~ length) + 
@@
   geom_line() + 
-  ggtitle("Conditional entropy vs. conditioned random variable (utterances of length 5)")
+  ggtitle("Conditional entropy vs. conditioned random variable")
 

### 2018-09-14 LDP_entropy.Rmd
cleaned up files, added prelim LDP entropy

---
title: Non-English Conditional Entropy and Mutual Information
author: Joe Klafka
date: "`r Sys.date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

```{r setup, include=FALSE}
# load packages
library(DBI)
library(knitr)
library(tidyverse)
library(directlabels)
library(dplyr)

library(SnowballC) #stemmer
library(tidytext)
library(entropy)
library(tidyboot)
library(tokenizers)
library(tau)
library(gtools)
knitr::opts_chunk$set(echo = TRUE)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
```

```{r open_connection}
connect_to_ldp <- function() {
 DBI::dbConnect(RMySQL::MySQL(),
                host = "mysql.rcc.uchicago.edu",
                dbname = "ldp",
                user = "ldp_user",
                password = "d0aa6b5950e")
}

connection <- connect_to_ldp()

get_table <- function(connection, name) {
 dplyr::tbl(connection, name)
}

utterances <- get_table(connection, "utterances") %>% collect() #to get the "utterances" table as a dataframe named utterances
DBI::dbListTables(connection) #to list the tables in the collection
```

```{r ldp_analysis}
get_unigram_entropies <- function(role, person, sen_length, utterances) {

  if(role == "child") {
    sub_utterances <- utterances %>%
      filter(subject == person) %>% 
      filter(c_utts_orig != "") %>%
      mutate(length = str_count(c_utts_orig, " |\\+|_") + 1)

      colnames(sub_utterances)[colnames(sub_utterances)=="c_utts_orig"] <- "gloss"
  } else {
     sub_utterances <- utterances %>%
      filter(subject == person) %>% 
      filter(p_utts_orig != "") %>%
      mutate(length = str_count(p_utts_orig, " |\\+|_") + 1)

      colnames(sub_utterances)[colnames(sub_utterances)=="p_utts_orig"] <- "gloss"
  }

    tokens <- sub_utterances %>%
    filter(length == sen_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, gloss) %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n()) %>%
    group_by(word_order, word) %>%
    mutate(clean_word = str_replace_all(tolower(word), "-|\\?|\\.|!|,", "")) %>%
    filter(word_order <= length) ##GO BACK AND FIX THIS LATER

  tokens %>%
    summarise(n = n()) %>%
    tidyboot(summary_function = function(x) x %>% 
               summarise(entropy = entropy(n, unit = "log2")),
             statistics_functions = function(x) x %>%
             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
    mutate(role = role, length = sen_length)
}

entropies <- map(2:10, ~get_unigram_entropies("child", , .x, utterances)) %>%
  bind_rows()

ggplot(entropies, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)
```

### 2018-09-14 english_unigram&ce&mi.Rmd
cleaned up files, added prelim LDP entropy

--- 2018-09-10 positional_entropy_childes.Rmd
@@
 ---
-title: Positional entropy in childes
-author: Joe Klafka
+title: Positional entropy, conditional entropy and mutual information in English CHILDES corpora
+author: Josef Klafka and Dan Yurovsky 
 date: "`r Sys.Date()`"
@@
 library(childesr) #data
-library(SnowballC) #stemmer
 library(tidytext)
@@
 library(tokenizers)
-library(tau)
 library(gtools)
@@
 brown_utterances <- get_utterances(corpus = "Brown")
-
-NA_utterances <- get_utterances(collection="Eng-NA")
 ```
 
-compute estimates
+Compute Providence Corpus Estimates
 ```{r calc_pos_entropy}
-get_counts <- function(role, sen_length, utterances) {
+get_entropies <- function(role, sen_length, utterances) {
 
@@
 
+  tokens <- sub_utterances %>%
+    filter(length == sen_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, stem) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(word_order, word)
+
+  tokens %>%
+    summarise(n = n()) %>%
+    tidyboot(summary_function = function(x) x %>% 
+               summarise(entropy = entropy(n, unit = "log2")),
+             statistics_functions = function(x) x %>%
+             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
+    mutate(role = role, length = sen_length)
+}
+
+entropies <- map(2:10, ~get_entropies("not child", .x, brown_utterances)) %>%
+  bind_rows()
+
+ggplot(entropies, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F)
+
+entropies_child <- map(2:10, ~ get_entropies("Target_Child", .x, prov_utterances)) %>%
+  bind_rows()
+
+ggplot(entropies_child, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F)
+
+entropies_all <- bind_rows(entropies, entropies_child)
+
+ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper, color = role)) +
+  facet_wrap(~ length) + 
+  geom_pointrange(position = position_dodge(.25)) +
+  geom_smooth(se = F) +
+  ggtitle("Child and Adult Positional Entropy vs. Word Position (Providence corpus)")
+
+new_utterances <- prov_utterances %>% filter(speaker_role=="Target_Child") %>% mutate(len = str_count(stem, " ") + 1)
+#toks <- new_utterances %>%
+
+ggplot(new_utterances, aes(x=len)) +
+  geom_histogram(binwidth = 1)
+```
+
+compute estimates
+```{r calc_pos_entropy}
+#get counts of each word in each position of utterances of length sen_length
+#from the collection of CHILDES utterances in utterances with speaker_role in role
+get_counts <- function(role, sen_length, utterances) {
+
+  if(role == "Target_Child") {
+    sub_utterances <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(length = str_count(stem, " ") + 1) # uses stems of words, not original words
+  } else {
+     sub_utterances <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      mutate(length = str_count(stem, " ") + 1)
+  }
+
   sub_utterances %>%
@@
 
+#compute the unigram entropy measure from the counts of each word in each position
 unigram_entropy <- function(counts) {
@@
 
-counts <- map(2:6, ~get_counts("not child", .x, prov_utterances))
-
+#compute the pairwise Mutual Information measure between all pairs of word positions
 all_pairwise_mi <- function(df) {
@@
 
+#perform the actual Mutual Information computation
 position_mi <- function(df, pos1, pos2) {
@@
 }
+
+counts <- map(2:6, ~get_counts("not child", .x, prov_utterances))
 
@@
 
+English Conditional Entropy
 ```{r conditional_entropy}
-
-counts <- map(2:8, ~get_counts("not child", .x, prov_utterances))
-
+#compute all Conditional Entropy from all one-way pairs of word positions
 all_conditional_entropy <- function(df) {
@@
 
+#perform the actual Conditional Entropy calculation
 conditional_entropy <- function(df, X, Y) { #Y is conditioning variable
@@
 
+counts <- map(2:8, ~get_counts("not child", .x, prov_utterances))
+
 ce <- map(counts, all_conditional_entropy)
@@
 
+N-Gram frequencies (saved for later)
 ```{r ngram_frequencies}

### 2018-09-14 non-english_entropies.Rmd
cleaned up files, added prelim LDP entropy

--- 2018-09-10 Unigram Positional Entropy.Rmd
@@
 ---
-title: Unigram entropy in childes
-author: Joe Klafka
+title: Unigram positional entropy, conditional entropy and mutual information in non-English CHILDES corpora
+author: Josef Klafka
 date: "`r Sys.Date()`"
@@
 prov_utterances <- get_utterances(corpus = "Providence")
+
+brown_utterances <- get_utterances(corpus = "Brown")
 ```
 
-Compute Providence Corpus Estimates
-```{r calc_pos_entropy}
+Unigram Entropy word length correction
+```{r word_len_corrections}
+get_counts <- function(role, sen_length, utterances) {
+
+  if(role == "Target_Child") {
+    sub_utterances <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1) # uses stems of words, not original words
+  } else {
+     sub_utterances <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1)
+  }
+
+  sub_utterances %>%
+    filter(length == sen_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, gloss) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    filter(word_order <= length)
+}
+
+get_mean_lengths <- function(role, sen_length, utterances) {
+
+  counts <- get_counts(role, sen_length, utterances)
+
+  lens <- counts %>% 
+    group_by(word_order, word) %>% 
+    mutate(word_len = str_length(word))
+
+  lens$word_len %>% 
+    aggregate(list(lens$word_order), mean)
+}
+
 get_entropies <- function(role, sen_length, utterances) {
@@
       filter(speaker_role == "Target_Child") %>%
-      mutate(length = str_count(stem, " ") + 1) # uses stems of words, not original words
+      mutate(length = str_count(gloss, " ") + 1) # uses stems of words, not original words
   } else {
@@
       filter(speaker_role != "Target_Child") %>%
-      mutate(length = str_count(stem, " ") + 1)
+      mutate(length = str_count(gloss, " ") + 1)
   }
-
-  # 
-  # 
-  # ggplot(adult_utterances, aes(x = length)) +
-  #   geom_histogram(fill = "white", color = "black")
 
@@
     mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, stem) %>%
+    unnest_tokens(word, gloss) %>%
     group_by(utterance_id) %>%
-    mutate(word_order = 1:n())
+    mutate(word_order = 1:n()) %>%
+    group_by(word_order, word) %>%
+    filter(word_order <= length)
 
   tokens %>%
-    group_by(word_order, word) %>%
     summarise(n = n()) %>%
@@
 
+get_cor_entropies <- function(role, utterances) {
+
+  mean_lens <- map(2:10, ~get_mean_lengths(role, .x, utterances)) %>% 
+    bind_rows()
+
+  entropies <- map(2:10, ~get_entropies(role, .x, utterances)) %>%
+    bind_rows()
+
+  entropies %>% 
+    mutate(mean_len = mean_lens[, "x"]) %>% 
+    mutate(cor_entropy = empirical_entropy / mean_len)
+}
+
+##
+mean_lens <- map(2:10, ~get_mean_lengths("not child", .x, prov_utterances)) %>% bind_rows()
+
+cor_entropies <- entropies %>% mutate(mean_len = mean_lens[, "x"]) %>% mutate(cor_entropy = empirical_entropy / mean_len)
+
 entropies <- map(2:10, ~get_entropies("not child", .x, prov_utterances)) %>%
   bind_rows()
+##
 
-ggplot(entropies, aes(x = word_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
+cor_entropies <- get_cor_entropies("not child", okayama_utterances)
+
+ggplot(cor_entropies, aes(x = word_order, y = cor_entropy)) +
      facet_wrap(~ length) + 
-  geom_pointrange() +
      geom_smooth(se = F)
-
-entropies_child <- map(2:10, ~ get_entropies("Target_Child", .x, prov_utterances)) %>%
-  bind_rows()
-
-ggplot(entropies_child, aes(x = word_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  geom_pointrange() +
-  geom_smooth(se = F)
-
-entropies_all <- bind_rows(entropies, entropies_child)
-
-ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper, color = role)) +
-  facet_wrap(~ length) + 
-  geom_pointrange(position = position_dodge(.25)) +
-  geom_smooth(se = F) +
-  ggtitle("Child and Adult Positional Entropy vs. Word Position (Providence corpus)")
-
-new_utterances <- prov_utterances %>% filter(speaker_role=="Target_Child") %>% mutate(len = str_count(stem, " ") + 1)
-#toks <- new_utterances %>%
-
-ggplot(new_utterances, aes(x=len)) +
-  geom_histogram(binwidth = 1)
-```
-
-Compile all bigrams in corpus (save for another time) 
-```{r ngram_frequencies}
-m_prov <- prov_utterances %>%
-          filter(speaker_role == "Target_Child") %>%
-          mutate(length = str_count(stem, " ") + 1) %>% 
-          filter(length >= 2)
-
-b_prov <- m_prov %>% mutate(bigrams = paste(unlist(textcnt(stem, n = 2, split = " ", method = "string")), collapse = ",")) #get all the counts of bigrams from all of the dataframe
-counts <- b_prov$bigrams[1] %>% strsplit(split=',') %>% unlist() %>% as.numeric()
-fileConn <- file("bigram_counts.txt")
-writeLines(b_prov$bigrams[1], fileConn)
-close(fileConn)
-
-b_prov <- m_prov %>% mutate(bigrams = paste(unlist(names(textcnt(stem, n = 2, split = " ", method = "string"))), collapse = ",")) #get all of the bigrams from the entire dataframe
-bigrams <- b_prov$bigrams[1] %>% strsplit(split=',') %>% unlist()
-fileConn <- file("bigram_words.txt")
-writeLines(b_prov$bigrams[1], fileConn)
-close(fileConn)
 ```
@@
 #Venezuelan children narrative (113 children; half 1st graders half 4th graders; half high SES half low SES)
-
-okayama_utterances <- get_utterances(corpus = "Okayama") #mother-child conversations in the Osaka area
 
@@
   }
-
-  # 
-  # 
-  # ggplot(adult_utterances, aes(x = length)) +
-  #   geom_histogram(fill = "white", color = "black")
 
@@
 ```{r Japanese Unigram Entropy}
+okayama_utterances <- get_utterances(corpus = "Okayama") #mother-child conversations in the Osaka area
+
 get_japanese_entropies <- function(role, sen_length, utterances) {

### 2018-09-14 bigrams.py

@@ -1,27 +1,0 @@
-import re
-
-file = open("counts.txt", 'r')
-s = file.read()
-counts = re.findall(r'\d+', s)
-counts = [int(num) for num in counts]
-
-file = open("bigram_words.txt", 'r')
-s = file.read()
-bigrams = re.split(r',', s)
-
-word_dict = {}
-
-for i in range(len(bigrams)):
-    words = re.split(' ', bigrams[i])
-    if words[0] not in word_dict:
-        word_dict[words[0]] = {}
-        word_dict[words[0]]["TOTAL"] = 0
-    if words[1] not in word_dict[words[0]]:
-        word_dict[words[0]][words[1]] = 0
-    word_dict[words[0]][words[1]] += counts[i]
-    word_dict[words[0]]["TOTAL"] += counts[i]
-
-for key in word_dict:
-    div = word_dict[key]["TOTAL"]
-    for val in word_dict[key]:
-        word_dict[key][val] /= div

### 2018-09-14 wiki_entropy.Rmd
added Wikipedia script and analyses for Georgian, Kyrgyz, Basque and Danish

---
title: Wikipedia Text Mining and Entropy Analysis
author: Josef Klafka
date:
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

```{r setup, include=FALSE}
# load packages
library(DBI)
library(knitr)
library(tidyverse)
library(directlabels)
library(dplyr)

library(SnowballC) #stemmer
library(tidytext)
library(entropy)
library(tidyboot)
library(tokenizers)
library(tau)
library(gtools)
library(reticulate)
knitr::opts_chunk$set(echo = TRUE)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
use_python("~/../../Library/Frameworks/Python.framework/Versions/3.7/bin/python3.7")
```

```{python}
import regex as re 
import numpy as np
import pandas as pd

def extract_texts(upper_bound):
	rs = ""
	for i in range(upper_bound):
		if i < 10:
			number = "0" + str(i)
		else:
			number = str(i)
		f = open("~/Documents/Labword/Joe/misc/text/AA/wiki_" + number, 'r')
		text = f.read()
		rs += text
	return rs

def get_sen_dict(text):
	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:–'&*#/—“»]|\d+|\(|\)|", "", text) 
	text = re.sub('"', '', text)
	#just the words from the articles
	text = text.replace('\n', " ")
	sens = re.split("[.!?]", text.lower())
	sen_dict = {}
	for sen in sens:
		length = len(re.findall("\w+", sen))
		if length > 0:
			sen_dict[sen] = length
	return sen_dict

def get_sen_df(text):
	sd = get_sen_dict(text)
	s = pd.Series(sd, name = "length")
	s.index.name = "sentence"
	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence 
					 	  #and the setence length respectively
	return df

text = extract_texts(99)
ky_df = get_sen_df(text)
```

```{r analyze_entropies}
get_entropies <- function(sen_length, df) {

  tokens <- df %>%
    filter(length == sen_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, sentence) %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n()) %>%
    group_by(word_order, word) %>%
    filter(word_order <= length)

  tokens %>%
    summarise(n = n()) %>%
    tidyboot(summary_function = function(x) x %>% 
               summarise(entropy = entropy(n, unit = "log2")),
             statistics_functions = function(x) x %>%
             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
    mutate(length = sen_length)
}

entropies <- map(6:12, ~get_entropies(.x, py$ky_df)) %>%
  bind_rows()

ggplot(entropies, aes(x = word_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)
```

### 2018-09-17 wiki_entropy.Rmd
Added Finnish, Korean and Arabic

@@
 			number = str(i)
-		f = open("~/Documents/Labword/Joe/misc/text/AA/wiki_" + number, 'r')
+		f = open("../misc/text/AA/wiki_" + number, 'r')
 		text = f.read()
@@
 
-text = extract_texts(99)
-ky_df = get_sen_df(text)
+text = extract_texts(20)
+ar_df = get_sen_df(text)
 ```
@@
 
-entropies <- map(6:12, ~get_entropies(.x, py$ky_df)) %>%
+entropies <- map(6:12, ~get_entropies(.x, py$ar_df)) %>%
   bind_rows()

### 2018-09-18 wiki_entropy.Rmd
More wikipedia data (Italian, Norwegian, Dutch, Romanian)

@@
 
-```{python}
+```{python, include=FALSE}
 import regex as re 
@@
 
-text = extract_texts(20)
-ar_df = get_sen_df(text)
+text = extract_texts(50)
+ro_df = get_sen_df(text)
 ```
@@
 
-entropies <- map(6:12, ~get_entropies(.x, py$ar_df)) %>%
+entropies <- map(6:12, ~get_entropies(.x, py$ro_df)) %>%
   bind_rows()

### 2018-09-18 english_unigram&ce&mi.Rmd
bootstrapping slope analysis

@@
 
+bootstrap entropy slopes
+```{r }
+lengths <- brown_utterances %>%
+  filter(speaker_role  != "Target_Child") %>%
+  mutate(length = str_count(stem, " ") + 1) %>%
+
+tokens <- lengths %>% 
+  filter(length >= 3) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, stem) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(length, word_order, word) %>%
+    summarise(n = n())
+
+start_slope <- function(df) {
+
+ df %>%
+    filter(word_order %in% c(1,2)) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `2` - `1`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+end_slope <- function(df) {
+
+ df %>%
+    ungroup() %>%
+    mutate(word_order = if_else(word_order == length, 2,
+                                if_else(word_order == length -1 , 1,
+                                        0))) %>%
+    filter(word_order > 0) %>%
+    group_by(length, word_order) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `2` - `1`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+start_time <- Sys.time()
+samples <- replicate(1000, end_slope(tokens))  %>%
+  as_data_frame()
+end_time <- Sys.time()
+
+samples %>%
+  summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+            ci_lower = quantile(value, .025))
+
+```
 Compute Providence Corpus Estimates

### 2018-09-24 english_unigram&ce&mi.Rmd
Deployed all slopes, constructed pipeline for CHILDES corpora

@@
 bootstrap entropy slopes
-```{r }
+```{r slope bootstrapping}
 lengths <- brown_utterances %>%
   filter(speaker_role  != "Target_Child") %>%
-  mutate(length = str_count(stem, " ") + 1) %>%
+  mutate(length = str_count(stem, " ") + 1)
 
 tokens <- lengths %>% 
-  filter(length >= 3) %>%
+  filter(length >= 6) %>%
     mutate(utterance_id = 1:n()) %>%
@@
 
+#extract tokens to be used for slope bootstrapping
+extract_tokens <- function(role, min_length, utterances) {
+if (role == "child") {
+lengths <- utterances %>%
+  filter(speaker_role == "Target_Child") %>%
+  mutate(length = str_count(stem, " ") + 1)
+
+} else {
+  lengths <- utterances %>%
+  filter(speaker_role != "Target_Child") %>%
+  mutate(length = str_count(stem, " ") + 1)
+}
+
+  tokens <- lengths %>% 
+    filter(length >= min_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, stem) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(length, word_order, word) %>%
+    summarise(n = n()) 
+}
+
 start_slope <- function(df) {
@@
     filter(word_order %in% c(1,2)) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `2` - `1`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+second_slope <- function(df) {
+
+ df %>%
+    ungroup() %>%
+    mutate(word_order %in% c(2, 3)) %>%
+    filter(word_order > 0) %>%
+    group_by(length, word_order) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `3` - `2`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+mid_slope <- function(df) {
+
+ df %>%
+    ungroup() %>%
+    mutate(word_order = if_else(word_order == length-2, 2,
+                                if_else(word_order == 3, 1,
+                                        0))) %>%
+    filter(word_order > 0) %>%
+    group_by(length, word_order) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `2` - `1`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+penult_slope <- function(df) {
+
+ df %>%
+    ungroup() %>%
+    mutate(word_order = if_else(word_order == length-1, 2,
+                                if_else(word_order == length -2, 1,
+                                        0))) %>%
+    filter(word_order > 0) %>%
+    group_by(length, word_order) %>%
     sample_frac(1, replace = T) %>%
@@
 start_time <- Sys.time()
-samples <- replicate(1000, end_slope(tokens))  %>%
+samples <- replicate(1000, mid_slope(tokens))  %>%
   as_data_frame()
 end_time <- Sys.time()
+
+end_time - start_time
 

### 2018-09-25 english_unigram&ce&mi.Rmd
Added non-English slope bootstrapping

@@
 ```{r get_utterances}
-prov_utterances <- get_utterances(corpus = "Providence")
+prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
 
-brown_utterances <- get_utterances(corpus = "Brown")
+brown_utterances <- get_utterances(corpus = "Brown") #stem or gloss
+
+shiro_utterances <- get_utterances(corpus = "Shiro") #use gloss
+
+wagner_utterances <- get_utterances(corpus = "Wagner") #use gloss
+
+okayama_utterances <- get_utterances(corpus = "Okayama") #use gloss
+
+zhou_utterances <- get_utterances(corpus = "ZhouDinner") #use stem, NOT gloss
 ```
@@
 ```{r slope bootstrapping}
-lengths <- brown_utterances %>%
-  filter(speaker_role  != "Target_Child") %>%
-  mutate(length = str_count(stem, " ") + 1)
-
-tokens <- lengths %>% 
-  filter(length >= 6) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, stem) %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    group_by(length, word_order, word) %>%
-    summarise(n = n())
 
 #extract tokens to be used for slope bootstrapping
-extract_tokens <- function(role, min_length, utterances) {
+extract_stem_tokens <- function(role, min_length, utterances) {
+
   if (role == "child") {
@@
     group_by(length, word_order, word) %>%
+    summarise(n = n()) 
+}
+
+extract_gloss_tokens <- function(role, min_length, utterances) {
+
+  if (role == "child") {
+    lengths <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1)
+
+} else {
+    lengths <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1)
+}
+
+  tokens <- lengths %>% 
+    filter(length >= min_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, gloss) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(length, word_order, word) %>%
+    filter(word_order <= length) %>% 
     summarise(n = n()) 
@@
 
+#note this measurement only makes sense if utterance length is at least 6
 mid_slope <- function(df) {
@@
 
+get_slopes <- function(role, min_length, utterances, stemgloss) {
+
+  if (stemgloss == "stem") {
+    tokens <- extract_stem_tokens(role, min_length, utterances)  
+  } else {
+    tokens <- extract_gloss_tokens(role, min_length, utterances)
+  }
+
+  start_slopes <- replicate(1000, start_slope(tokens)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+
+  second_slopes <- replicate(1000, second_slope(tokens)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+
+  mid_slopes <- replicate(1000, mid_slope(tokens)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+
+  penult_slopes <- replicate(1000, penult_slope(tokens)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+
+  end_slopes <- replicate(1000, end_slope(tokens)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+
+  rbind(start_slopes, second_slopes, mid_slopes, penult_slopes, end_slopes)
+}
+
+shiro_slopes <- get_slopes("child", 6, shiro_utterances, "gloss") #what we would expect
+shiro_slopes <- get_slopes("not child", 6, shiro_utterances, "gloss")
+wagner_slopes <- get_slopes("child", 6, wagner_utterances, "gloss")
+
 start_time <- Sys.time()
-samples <- replicate(1000, mid_slope(tokens))  %>%
+samples <- replicate(1000, start_slope(tokens))  %>%
   as_data_frame()
@@
 
-samples %>%
+start_slopes <- samples %>%
   summarise(mean = mean(value), ci_upper = quantile(value, .975), 

### 2018-09-26 wiki_entropy.Rmd
changes to wiki_entropy
wiki_entropy slope analysis not function, cannot fix error at present

@@
 	s = pd.Series(sd, name = "length")
-	s.index.name = "sentence"
+	del sd
+	s.index.name = "gloss"
 	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence 
 					 	  #and the setence length respectively
+	df["speaker_role"] = "not child"
 	return df
 
-text = extract_texts(50)
-ro_df = get_sen_df(text)
+text = extract_texts(20)
+xmf_df = get_sen_df(text)
 ```
 
-```{r analyze_entropies}
+```{r slope_code, include=FALSE}
+extract_stem_tokens <- function(role, min_length, utterances) {
+
+  if (role == "child") {
+    lengths <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(length = str_count(stem, " ") + 1)
+
+} else {
+    lengths <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      mutate(length = str_count(stem, " ") + 1)
+}
+
+  tokens <- lengths %>% 
+    filter(length >= min_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, stem) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(length, word_order, word) %>%
+    summarise(n = n()) 
+}
+
+extract_gloss_tokens <- function(role, min_length, utterances) {
+
+  if (role == "child") {
+    lengths <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1)
+
+} else {
+    lengths <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1)
+}
+
+  tokens <- lengths %>% 
+    filter(length >= min_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, gloss) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(length, word_order, word) %>%
+    filter(word_order <= length) %>% 
+    summarise(n = n()) 
+}
+
+start_slope <- function(df) {
+
+ df %>%
+    filter(word_order %in% c(1,2)) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `2` - `1`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+second_slope <- function(df) {
+
+ df %>%
+    ungroup() %>%
+    mutate(word_order %in% c(2, 3)) %>%
+    filter(word_order > 0) %>%
+    group_by(length, word_order) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `3` - `2`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+#note this measurement only makes sense if utterance length is at least 6
+mid_slope <- function(df) {
+
+ df %>%
+    ungroup() %>%
+    mutate(word_order = if_else(word_order == length-2, 2,
+                                if_else(word_order == 3, 1,
+                                        0))) %>%
+    filter(word_order > 0) %>%
+    group_by(length, word_order) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `2` - `1`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+penult_slope <- function(df) {
+
+ df %>%
+    ungroup() %>%
+    mutate(word_order = if_else(word_order == length-1, 2,
+                                if_else(word_order == length -2, 1,
+                                        0))) %>%
+    filter(word_order > 0) %>%
+    group_by(length, word_order) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `2` - `1`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+end_slope <- function(df) {
+
+ df %>%
+    ungroup() %>%
+    mutate(word_order = if_else(word_order == length, 2,
+                                if_else(word_order == length -1 , 1,
+                                        0))) %>%
+    filter(word_order > 0) %>%
+    group_by(length, word_order) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `2` - `1`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+get_slopes <- function(role, min_length, utterances, stemgloss) {
+
+  if (stemgloss == "stem") {
+    tokens <- extract_stem_tokens(role, min_length, utterances)  
+  } else {
+    tokens <- extract_gloss_tokens(role, min_length, utterances)
+  }
+
+  start_slopes <- replicate(1000, start_slope(tokens)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+
+  second_slopes <- replicate(1000, second_slope(tokens)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+
+  mid_slopes <- replicate(1000, mid_slope(tokens)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+
+  penult_slopes <- replicate(1000, penult_slope(tokens)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+
+  end_slopes <- replicate(1000, end_slope(tokens)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+
+  rbind(start_slopes, second_slopes, mid_slopes, penult_slopes, end_slopes)
+}
+
+```
+
+```{r slope_analysis}
+#get_slopes("not child", 6, py$xmf_df, "gloss")
+gtx <- extract_gloss_tokens("not child", 6, py$xmf_df)
+start_slopes <- replicate(1000, start_slope(gtx)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+print(start_slopes)
+```
+
+```{r analyze_entropies, include=FALSE}
 get_entropies <- function(sen_length, df) {
@@
 
-entropies <- map(6:12, ~get_entropies(.x, py$ro_df)) %>%
-  bind_rows()
+#entropies <- map(6:12, ~get_entropies(.x, py$xmf_df)) %>%
+#  bind_rows()
 
-ggplot(entropies, aes(x = word_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  geom_pointrange() +
-  geom_smooth(se = F)
+#ggplot(entropies, aes(x = word_order, y = empirical_entropy,
+#                      ymin = ci_lower, ymax = ci_upper)) +
+#  facet_wrap(~ length) + 
+#  geom_pointrange() +
+#  geom_smooth(se = F)
 ```

### 2018-10-04 english_unigram&ce&mi.Rmd
Fixed entropy pipeline

@@
 #extract tokens to be used for slope bootstrapping
-extract_stem_tokens <- function(role, min_length, utterances) {
+extract_stem_tokens <- function(role, target_length, utterances) {
 
@@
   tokens <- lengths %>% 
-    filter(length >= min_length) %>%
+    filter(length == target_length) %>%
     mutate(utterance_id = 1:n()) %>%
@@
 
-extract_gloss_tokens <- function(role, min_length, utterances) {
+extract_gloss_tokens <- function(role, target_length, utterances) {
 
@@
 
-get_slopes <- function(role, min_length, utterances, stemgloss) {
+get_slopes <- function(role, target_length, utterances, stemgloss) {
 
   if (stemgloss == "stem") {
-    tokens <- extract_stem_tokens(role, min_length, utterances)  
+    tokens <- extract_stem_tokens(role, target_length, utterances)  
   } else {
-    tokens <- extract_gloss_tokens(role, min_length, utterances)
+    tokens <- extract_tokens(role, target_length, target_length, utterances)
   }
@@
 
+```
+
+```{r log_lengths}
+tokens <- extract_tokens("not child", 6, 30, prov_utterances)
+
+tokens %>%
+    filter(word_order %in% c(1,2)) %>%
+    dplyr::group_by(length, word_order) %>%
+    dplyr::summarise(entropy = entropy(n, unit = "log2"), n = n()) %>%
+    spread(word_order, entropy) %>%
+    group_by(length) %>%
+    mutate(diff = `2` - `1`) %>%
+  ggplot(aes(x = length, y = diff)) +
+  geom_point(aes(size = n)) +
+  geom_smooth(se = F) +
+  scale_x_log10() #log length stats
+
+extract_tokens <- function(role, min_length, max_length, utterances) {
+
+  if (role == "child") {
+    lengths <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(length = str_count(stem, " +") +1)
+
+} else {
+    lengths <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+     mutate(length = str_count(stem, " +") +1)
+}
+
+  tokens <- lengths %>% 
+    filter(length >= min_length) %>%
+    filter(length <= max_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, stem, token = stringr::str_split, pattern = " +") %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(length, word_order, word) %>%
+    filter(word_order <= length) %>% 
+    summarise(n = n()) 
+}
+
+shiro_slopes <- map(6:10, ~get_slopes("child", .x, shiro_utterances, "gloss"))
+
 shiro_slopes <- get_slopes("child", 6, shiro_utterances, "gloss") #what we would expect
@@
 wagner_slopes <- get_slopes("child", 6, wagner_utterances, "gloss")
+okayama_slopes <- get_slopes("child", 6, okayama_utterances, "gloss")
+zhou_slopes <- get_slopes("child", 6, zhou_utterances, "stem")
 
@@
 ```
+
+```{r slope_analysis}
+
+new_get_tokens <- function(role, min_length, max_length, utterances) {
+
+  if (role == "child") {
+        lengths <- utterances %>%
+            filter(speaker_role != "Target_Child") %>%
+            mutate(length = str_count(gloss, " ") + 1)
+
+  } else {
+
+        lengths <- utterances %>%
+            filter(speaker_role != "Target_Child") %>%
+            mutate(length = str_count(gloss, " ") + 1)
+  }
+
+  sub_tokens <- lengths %>% 
+     mutate(gloss = str_trim(gloss)) %>%
+     mutate(length = str_count(gloss, " +") +1) %>%
+     filter(length >= min_length) %>%
+     filter(length <= max_length) %>%
+     mutate(utterance_id = 1:n()) %>%
+     unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
+     group_by(utterance_id) %>%
+     mutate(word_order = 1:n()) %>%
+     group_by(length, word_order, word) %>%
+     filter(word_order <= length) %>% 
+     summarise(n = n()) 
+}
+
+min_length <- 6
+max_length <- 20
+
+lengths <- prov_utterances %>%
+       filter(speaker_role != "Target_Child") %>%
+       mutate(length = str_count(gloss, " ") + 1)
+
+sub_tokens <- lengths %>% 
+     mutate(gloss = str_trim(gloss)) %>%
+     mutate(length = str_count(gloss, " +") +1) %>%
+     filter(length >= min_length) %>%
+     filter(length <= max_length) %>%
+     mutate(utterance_id = 1:n())
+
+sub_tokens %>%
+     unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
+     group_by(utterance_id) %>%
+     mutate(word_order = 1:n()) %>%
+     group_by(utterance_id) %>%
+     summarise(n = n()) %>%
+     # filter(n != 6) %>%
+     left_join(sub_tokens) %>%
+     select(gloss, n, length) %>%
+     View()
+
+```
+
 Compute Providence Corpus Estimates

### 2018-10-04 wiki_entropy.Rmd
Fixed entropy pipeline

--- 2018-09-17 wiki_entropy.Rmd
@@
 library(reticulate)
+
+library(feather)
+
 knitr::opts_chunk$set(echo = TRUE)
@@
 
-```{python}
+```{python, include=FALSE, eval = T}
 import regex as re 
@@
 import pandas as pd 
+import csv
 
@@
 	s = pd.Series(sd, name = "length")
-	s.index.name = "sentence"
+	del sd
+	s.index.name = "gloss"
 	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence 
 					 	  #and the setence length respectively
+	df["speaker_role"] = "not child"
 	return df
@@
 text = extract_texts(20)
-ar_df = get_sen_df(text)
+xmf_df = get_sen_df(text)
+
+xmf_df.to_csv("xmf_df.csv")
 ```
 
-```{r analyze_entropies}
+```{r slope_code, include=FALSE, eval = F}
+# extract_stem_tokens <- function(role, min_length, utterances) {
+#   
+#   REPS <- 100
+#   
+#   if (role == "child") {
+#     lengths <- utterances %>%
+#       filter(speaker_role == "Target_Child") %>%
+#       mutate(length = str_count(stem, " ") + 1)
+# 
+#   } else {
+#     lengths <- utterances %>%
+#       filter(speaker_role != "Target_Child") %>%
+#       mutate(length = str_count(stem, " ") + 1)
+#   }
+#   
+#   tokens <- lengths %>% 
+#     filter(length >= min_length) %>%
+#     mutate(utterance_id = 1:n()) %>%
+#     unnest_tokens(word, stem) %>%
+#     group_by(utterance_id) %>%
+#     mutate(word_order = 1:n()) %>%
+#     group_by(length, word_order, word) %>%
+#     summarise(n = n()) 
+# }
+
+extract_wiki_tokens <- function(min_length, max_length, utterances) {
+
+  tokens <- utterances %>% 
+    filter(length >= min_length) %>%
+    filter(length <= max_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, gloss) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(length, word_order, word) %>%
+    filter(word_order <= length) %>% 
+    summarise(n = n()) 
+}
+
+extract_tokens <- function(min_length, max_length, utterances) {
+
+  tokens <- utterances %>% 
+    mutate(lengths = str_count(gloss, " ")) %>%
+    filter(lengths >= min_length) %>%
+    filter(lengths <= max_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(length, word_order, word) %>%
+    filter(word_order <= length) %>% 
+    summarise(n = n()) 
+}
+
+start_slope <- function(df) {
+
+ df %>%
+    filter(word_order %in% c(1,2)) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `2` - `1`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+second_slope <- function(df) {
+
+ df %>%
+    ungroup() %>%
+    mutate(word_order %in% c(2, 3)) %>%
+    filter(word_order > 0) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `3` - `2`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+#note this measurement only makes sense if utterance length is at least 6
+mid_slope <- function(df) {
+
+ df %>%
+    ungroup() %>%
+    mutate(word_order = if_else(word_order == length-2, 2,
+                                if_else(word_order == 3, 1,
+                                        0))) %>%
+    filter(word_order > 0) %>%
+    group_by(length, word_order) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `2` - `1`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+penult_slope <- function(df) {
+
+ df %>%
+    ungroup() %>%
+    mutate(word_order = if_else(word_order == length-1, 2,
+                                if_else(word_order == length -2, 1,
+                                        0))) %>%
+    filter(word_order > 0) %>%
+    group_by(length, word_order) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `2` - `1`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+end_slope <- function(df) {
+
+ df %>%
+    ungroup() %>%
+    mutate(word_order = if_else(word_order == length, 2,
+                                if_else(word_order == length -1 , 1,
+                                        0))) %>%
+    filter(word_order > 0) %>%
+    group_by(length, word_order) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `2` - `1`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+get_wiki_slopes <- function(min_length, max_length, utterances) {
+
+  REPS <- 100
+
+  tokens <- extract_tokens(min_length, max_length, utterances)
+
+  start_slopes <- replicate(REPS, start_slope(tokens)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+
+  second_slopes <- replicate(REPS, second_slope(tokens)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+
+  mid_slopes <- replicate(REPS, mid_slope(tokens)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+
+  penult_slopes <- replicate(REPS, penult_slope(tokens)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+
+  end_slopes <- replicate(REPS, end_slope(tokens)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+
+  rbind(start_slopes, second_slopes, mid_slopes, penult_slopes, end_slopes)
+}
+
+```
+
+```{r slope_analysis, eval = F}
+
+xmf_df <- read_csv("~/Documents/Labwork/Joe/Positional-Information/xmf_df.csv")
+
+#get_slopes("not child", 6, py$xmf_df, "gloss")
+gtx <- extract_tokens(6, 25, xmf_df) %>% filter(word != " ") %>% filter(word != "")
+
+gtx %>%
+    filter(word_order %in% c(1,2)) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2"), n = n()) %>%
+    spread(word_order, entropy) %>%
+    group_by(length) %>%
+    mutate(diff = `2` - `1`) %>%
+  ggplot(aes(x = length, y = diff)) +
+  geom_point(aes(size = n)) +
+  geom_smooth(se = F) +
+  scale_x_log10() #log length stats
+
+get_wiki_slopes(min_length = 6, max_length = 50, utterances=xmf_df)
+
+second_slopes <- replicate(100, second_slope(gtx)) %>% 
+    as_data_frame() %>% View()
+    # summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+    #           ci_lower = quantile(value, .025))
+
+print(second_slopes)
+```
+
+```{r analyze_entropies, include=FALSE, eval = F}
 get_entropies <- function(sen_length, df) {
@@
 
-entropies <- map(6:12, ~get_entropies(.x, py$ar_df)) %>%
-  bind_rows()
+#entropies <- map(6:12, ~get_entropies(.x, py$xmf_df)) %>%
+#  bind_rows()
 
-ggplot(entropies, aes(x = word_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  geom_pointrange() +
-  geom_smooth(se = F)
+#ggplot(entropies, aes(x = word_order, y = empirical_entropy,
+#                      ymin = ci_lower, ymax = ci_upper)) +
+#  facet_wrap(~ length) + 
+#  geom_pointrange() +
+#  geom_smooth(se = F)
 ```

### 2018-10-05 english_unigram&ce&mi.Rmd
Fixing n difference in slope calculation, otherwise calculation works

@@
 
+tkns <- extract_tokens("child", 6, 20, prov_utterances)
+
+new_tkns <- tkns %>% 
+    filter(word_order %in% c(1,2)) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2"), n = n()) %>%
+    spread(word_order, entropy) %>%
+    group_by(length) %>%
+    mutate(diff = `2` - `1`) %>%
+  ggplot(aes(x = length, y = diff)) +
+  geom_point(aes(size = n)) +
+  geom_smooth(se = F) +
+  scale_x_log10()
+
 min_length <- 6

### 2018-10-05 wiki_entropy.Rmd
Fixing n difference in slope calculation, otherwise calculation works

@@
 
-```{r slope_code, include=FALSE, eval = F}
+```{r slope_analysis, include=FALSE, eval = F}
 # extract_stem_tokens <- function(role, min_length, utterances) {
@@
 
+xmf_df <- read_csv("xmf_df.csv")
+
+#get_slopes("not child", 6, py$xmf_df, "gloss")
+
+sub_tokens <- xmf_df %>% 
+     mutate(gloss = str_trim(gloss)) %>%
+     mutate(length = str_count(gloss, " +") +1) %>%
+     filter(length >= min_length) %>%
+     filter(length <= max_length) %>%
+     mutate(utterance_id = 1:n())
+
+sub_tokens %>%
+     unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
+     group_by(utterance_id) %>%
+     mutate(word_order = 1:n()) %>%
+     group_by(utterance_id) %>%
+     summarise(n = n()) %>%
+     # filter(n != 6) %>%
+     left_join(sub_tokens) %>%
+     select(gloss, n, length) %>%
+     View()
+
 extract_wiki_tokens <- function(min_length, max_length, utterances) {
@@
   tokens <- utterances %>% 
+    mutate(gloss = str_trim(gloss)) %>%
+    mutate(length = str_count(gloss, pattern = " +") + 1) %>% 
     filter(length >= min_length) %>%
@@
     mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss) %>%
+    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    # group_by(utterance_id) %>%
+    group_by(length, word_order, word) %>%
+    # filter(word_order <= length) %>% 
+    summarise(n = n()) # %>%
+    # left_join(utterances) %>%
+    # select(gloss, n, length) %>%
+    # View()
+}
+
+gtx <- extract_wiki_tokens(6, 25, xmf_df) %>% filter(word != " ") %>% filter(word != "")
+
+gtx %>%
+    filter(word_order %in% c(1,2)) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2"), n = n()) %>%
+    spread(word_order, entropy) %>%
+    group_by(length) %>%
+    mutate(diff = `2` - `1`) %>%
+  ggplot(aes(x = length, y = diff)) +
+  geom_point(aes(size = n)) +
+  geom_smooth(se = F) +
+  scale_x_log10() #log length stats
+
+```
+
+```{r slope_code}
+extract_stem_tokens <- function(role, min_length, utterances) {
+
+  if (role == "child") {
+    lengths <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(length = str_count(stem, " ") + 1)
+
+} else {
+    lengths <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      mutate(length = str_count(stem, " ") + 1)
+}
+
+  tokens <- lengths %>% 
+    filter(length >= min_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, stem) %>%
     group_by(utterance_id) %>%
@@
     group_by(length, word_order, word) %>%
-    filter(word_order <= length) %>% 
     summarise(n = n()) 
@@
 
-extract_tokens <- function(min_length, max_length, utterances) {
+extract_gloss_tokens <- function(role, min_length, utterances) {
 
-  tokens <- utterances %>% 
-    mutate(lengths = str_count(gloss, " ")) %>%
-    filter(lengths >= min_length) %>%
-    filter(lengths <= max_length) %>%
+  if (role == "child") {
+    lengths <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1)
+
+} else {
+    lengths <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1)
+}
+
+  tokens <- lengths %>% 
+    filter(length >= min_length) %>%
     mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
+    unnest_tokens(word, gloss) %>%
     group_by(utterance_id) %>%
@@
 
-  tokens <- extract_tokens(min_length, max_length, utterances)
+  tokens <- extract_wiki_tokens(min_length, max_length, utterances)
 
@@
 
-xmf_df <- read_csv("~/Documents/Labwork/Joe/Positional-Information/xmf_df.csv")
+xmf_df <- read_csv("xmf_df.csv")
 
 #get_slopes("not child", 6, py$xmf_df, "gloss")
-gtx <- extract_tokens(6, 25, xmf_df) %>% filter(word != " ") %>% filter(word != "")
+gtx <- extract_wiki_tokens(6, 25, xmf_df) %>% filter(word != " ") %>% filter(word != "")
 
-gtx %>%
+tokens %>%
     filter(word_order %in% c(1,2)) %>%
@@
 second_slopes <- replicate(100, second_slope(gtx)) %>% 
-    as_data_frame() %>% View()
-    # summarise(mean = mean(value), ci_upper = quantile(value, .975), 
-    #           ci_lower = quantile(value, .025))
+    as_data_frame() %>% 
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+             ci_lower = quantile(value, .025))
 
@@
 ```{r analyze_entropies, include=FALSE, eval = F}
+```{r slope_analysis}
+#get_slopes("not child", 6, py$xmf_df, "gloss")
+gtx <- extract_gloss_tokens("not child", 6, py$xmf_df)
+start_slopes <- replicate(1000, start_slope(gtx)) %>% 
+    as_data_frame() %>%
+    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
+              ci_lower = quantile(value, .025))
+print(start_slopes)
+```
+
+```{r analyze_entropies, include=FALSE}
 get_entropies <- function(sen_length, df) {

### 2018-10-05 slope_comparison.Rmd
new cleaner file to test slope numbers on

---
title: "Untitled"
author: "Joe Klafka"
date: "10/5/2018"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(directlabels)
library(childesr) #data
library(tidytext)
library(entropy)
library(tidyboot)
library(dplyr)
library(tokenizers)
library(gtools)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
```

```{r data}
prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss

xmf_df <- read_csv("xmf_df.csv")

```

```{r slope_analysis}

extract_childes_tokens <- function(role, min_length, max_length, utterances) {

  if (role == "child") {
        lengths <- utterances %>%
            filter(speaker_role != "Target_Child") %>%
            mutate(length = str_count(gloss, " ") + 1)
  } else {
        lengths <- utterances %>%
            filter(speaker_role != "Target_Child") %>%
            mutate(length = str_count(gloss, " ") + 1)
  }

  sub_tokens <- lengths %>% 
     mutate(gloss = str_trim(gloss)) %>%
     mutate(length = str_count(gloss, " +") +1) %>%
     filter(length >= min_length) %>%
     filter(length <= max_length) %>%
     mutate(utterance_id = 1:n()) %>%
     unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
     group_by(utterance_id) %>%
     mutate(word_order = 1:n()) %>%
     group_by(length, word_order, word) %>%
     filter(word_order <= length) %>% 
     summarise(n = n()) 
}

extract_wiki_tokens <- function(min_length, max_length, utterances) {

  tokens <- utterances %>% 
    mutate(gloss = str_trim(gloss)) %>%
    mutate(length = str_count(gloss, pattern = " +") + 1) %>% 
    filter(length >= min_length) %>%
    filter(length <= max_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n()) %>%
    # group_by(utterance_id) %>%
    group_by(length, word_order, word) %>%
    # filter(word_order <= length) %>% 
    summarise(n = n()) # %>%
    # left_join(utterances) %>%
    # select(gloss, n, length) %>%
    # View()
}

length_comparison <- function(min_length, max_length, utterances) {

  sub_tokens <- utterances %>% 
     mutate(gloss = str_trim(gloss)) %>%
     mutate(length = str_count(gloss, " +") +1) %>%
     filter(length >= min_length) %>%
     filter(length <= max_length) %>%
     mutate(utterance_id = 1:n())

sub_tokens %>%
     unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
     group_by(utterance_id) %>%
     mutate(word_order = 1:n()) %>%
     group_by(utterance_id) %>%
     summarise(n = n()) %>%
     # filter(n != 6) %>%
     left_join(sub_tokens) %>%
     select(gloss, n, length) %>%
     View()
}

n_comparison <- function(tokens) {

  tokens %>%
    filter(word_order %in% c(1,2)) %>%
    dplyr::group_by(length, word_order) %>%
    dplyr::summarise(entropy = entropy(n, unit = "log2"), n = n()) 
}

```

```{r slope_work}
prov_tokens <- extract_childes_tokens("not child", 6, 30, prov_utterances)

xmf_tokens <- extract_wiki_tokens(6, 25, xmf_df) %>% filter(word != " ") %>% filter(word != "")

```

```{r leftover code}
tokens %>%
    filter(word_order %in% c(1,2)) %>%
    dplyr::group_by(length, word_order) %>%
    dplyr::summarise(entropy = entropy(n, unit = "log2"), n = n()) %>%
    spread(word_order, entropy) %>%
    group_by(length) %>%
    mutate(diff = `2` - `1`) %>%
  ggplot(aes(x = length, y = diff)) +
  geom_point(aes(size = n)) +
  geom_smooth(se = F) +
  scale_x_log10() #log length stats

```

### 2018-10-05 wiki_entropy.Rmd
Fixed differing n problem

@@
 #get_slopes("not child", 6, py$xmf_df, "gloss")
-gtx <- extract_wiki_tokens(6, 25, xmf_df) %>% filter(word != " ") %>% filter(word != "")
 
-tokens %>%
+gtx <- extract_wiki_tokens(6, 1000, xmf_df)
+
+gtx %>%
     filter(word_order %in% c(1,2)) %>%
     group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2"), n = n()) %>%
+    summarise(entropy = entropy(n, unit = "log2"), n = sum(n)) %>%
     spread(word_order, entropy) %>%

### 2018-10-08 wiki_entropy.Rmd
started entropy slope analysis on wikipedia data

@@
 # load packages
-library(DBI)
 library(knitr)
@@
 					 	  #and the setence length respectively
-	df["speaker_role"] = "not child"
 	return df
 
-text = extract_texts(20)
-xmf_df = get_sen_df(text)
+text = extract_texts(15)
+am_df = get_sen_df(text)
 
-xmf_df.to_csv("xmf_df.csv")
+am_df.to_csv("am_df.csv")
 ```
 
-```{r slope_analysis, include=FALSE, eval = F}
+```{r slope_prep, include=FALSE, eval = F}
 # extract_stem_tokens <- function(role, min_length, utterances) {
@@
 
-xmf_df <- read_csv("xmf_df.csv")
+extract_wiki_tokens <- function(min_length, max_length, utterances) {
+
+  tokens <- utterances %>% 
+    mutate(gloss = str_trim(gloss)) %>%
+    mutate(length = str_count(gloss, pattern = " +") + 1) %>% 
+    filter(length >= min_length) %>%
+    filter(length <= max_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    # group_by(utterance_id) %>%
+    group_by(length, word_order, word) %>%
+    # filter(word_order <= length) %>% 
+    summarise(n = n()) # %>%
+    # left_join(utterances) %>%
+    # select(gloss, n, length) %>%
+    # View()
+}
+
+df <- read_csv("~/Documents/Labwork/Joe/Positional-Information/ar_df.csv")
+
+wiki_tokens <- extract_wiki_tokens(6, 50, df)
+
+wiki_tokens %>%
+    filter(word_order %in% c(1,2)) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2"), n = sum(n)) %>%
+    spread(word_order, entropy) %>%
+    group_by(length) %>%
+    mutate(diff = `2` - `1`) %>%
+  ggplot(aes(x = length, y = diff)) +
+  geom_point(aes(size = n)) +
+  geom_smooth(se = F) +
+  scale_x_log10() #log length stats
+
+ar_slopes <- get_wiki_slopes(min_length = 6, max_length = 50, utterances=df, 1000)
 
@@
 
-extract_wiki_tokens <- function(min_length, max_length, utterances) {
-
-  tokens <- utterances %>% 
-    mutate(gloss = str_trim(gloss)) %>%
-    mutate(length = str_count(gloss, pattern = " +") + 1) %>% 
-    filter(length >= min_length) %>%
-    filter(length <= max_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    # group_by(utterance_id) %>%
-    group_by(length, word_order, word) %>%
-    # filter(word_order <= length) %>% 
-    summarise(n = n()) # %>%
-    # left_join(utterances) %>%
-    # select(gloss, n, length) %>%
-    # View()
-}
-
-gtx <- extract_wiki_tokens(6, 25, xmf_df) %>% filter(word != " ") %>% filter(word != "")
-
-gtx %>%
-    filter(word_order %in% c(1,2)) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2"), n = n()) %>%
-    spread(word_order, entropy) %>%
-    group_by(length) %>%
-    mutate(diff = `2` - `1`) %>%
-  ggplot(aes(x = length, y = diff)) +
-  geom_point(aes(size = n)) +
-  geom_smooth(se = F) +
-  scale_x_log10() #log length stats
-
 ```
 
-```{r slope_code}
-extract_stem_tokens <- function(role, min_length, utterances) {
-
-  if (role == "child") {
-    lengths <- utterances %>%
-      filter(speaker_role == "Target_Child") %>%
-      mutate(length = str_count(stem, " ") + 1)
-
-} else {
-    lengths <- utterances %>%
-      filter(speaker_role != "Target_Child") %>%
-      mutate(length = str_count(stem, " ") + 1)
-}
-
-  tokens <- lengths %>% 
-    filter(length >= min_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, stem) %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    group_by(length, word_order, word) %>%
-    summarise(n = n()) 
-}
-
-extract_gloss_tokens <- function(role, min_length, utterances) {
-
-  if (role == "child") {
-    lengths <- utterances %>%
-      filter(speaker_role == "Target_Child") %>%
-      mutate(length = str_count(gloss, " ") + 1)
-
-} else {
-    lengths <- utterances %>%
-      filter(speaker_role != "Target_Child") %>%
-      mutate(length = str_count(gloss, " ") + 1)
-}
-
-  tokens <- lengths %>% 
-    filter(length >= min_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss) %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    group_by(length, word_order, word) %>%
-    filter(word_order <= length) %>% 
-    summarise(n = n()) 
-}
+```{r slope_code, include = F, eval=F}
 
@@
 
-get_wiki_slopes <- function(min_length, max_length, utterances) {
-
-  REPS <- 100
+get_wiki_slopes <- function(min_length, max_length, utterances, REPS) {
 
@@
 
-```{r slope_analysis, eval = F}
+```{r slope_analysis, include = F, eval = F}
 
@@
 
-second_slopes <- replicate(100, second_slope(gtx)) %>% 
-    as_data_frame() %>% 
-    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
-             ci_lower = quantile(value, .025))
-
-print(second_slopes)
 ```
@@
 ```{r analyze_entropies, include=FALSE, eval = F}
-```{r slope_analysis}
-#get_slopes("not child", 6, py$xmf_df, "gloss")
-gtx <- extract_gloss_tokens("not child", 6, py$xmf_df)
-start_slopes <- replicate(1000, start_slope(gtx)) %>% 
-    as_data_frame() %>%
-    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
-              ci_lower = quantile(value, .025))
-print(start_slopes)
-```
-
-```{r analyze_entropies, include=FALSE}
 get_entropies <- function(sen_length, df) {
@@
 #  geom_smooth(se = F)
+
+# extract_stem_tokens <- function(role, min_length, utterances) {
+#   
+#   if (role == "child") {
+#     lengths <- utterances %>%
+#       filter(speaker_role == "Target_Child") %>%
+#       mutate(length = str_count(stem, " ") + 1)
+# 
+# } else {
+#     lengths <- utterances %>%
+#       filter(speaker_role != "Target_Child") %>%
+#       mutate(length = str_count(stem, " ") + 1)
+# }
+#   
+#   tokens <- lengths %>% 
+#     filter(length >= min_length) %>%
+#     mutate(utterance_id = 1:n()) %>%
+#     unnest_tokens(word, stem) %>%
+#     group_by(utterance_id) %>%
+#     mutate(word_order = 1:n()) %>%
+#     group_by(length, word_order, word) %>%
+#     summarise(n = n()) 
+# }
+# 
+# extract_gloss_tokens <- function(role, min_length, utterances) {
+#   
+#   if (role == "child") {
+#     lengths <- utterances %>%
+#       filter(speaker_role == "Target_Child") %>%
+#       mutate(length = str_count(gloss, " ") + 1)
+# 
+# } else {
+#     lengths <- utterances %>%
+#       filter(speaker_role != "Target_Child") %>%
+#       mutate(length = str_count(gloss, " ") + 1)
+# }
+#   
+#   tokens <- lengths %>% 
+#     filter(length >= min_length) %>%
+#     mutate(utterance_id = 1:n()) %>%
+#     unnest_tokens(word, gloss) %>%
+#     group_by(utterance_id) %>%
+#     mutate(word_order = 1:n()) %>%
+#     group_by(length, word_order, word) %>%
+#     filter(word_order <= length) %>% 
+#     summarise(n = n()) 
+# }
 ```

### 2018-10-08 wiki_entropy.py

@@ -1,0 +1,50 @@
+import regex as re 
+import numpy as np
+import pandas as pd 
+import csv
+import sys
+
+def extract_texts(upper_bound):
+	rs = ""
+	for i in range(upper_bound):
+		if i < 10:
+			number = "0" + str(i)
+		else:
+			number = str(i)
+		f = open("../misc/text/AA/wiki_" + number, 'r')
+		text = f.read()
+		rs += text
+	return rs
+
+
+def get_sen_dict(text):
+	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:–'&*#/—“»]|\d+|\(|\)|\[|\]", "", text) 
+	text = re.sub('"', '', text)
+	#just the words from the articles
+	text = text.replace('\n', " ")
+	sens = re.split("[.!?]", text.lower())
+	sen_dict = {}
+	for sen in sens:
+		length = len(re.findall("\w+", sen))
+		if length > 0:
+			sen_dict[sen] = length
+	return sen_dict
+
+
+def get_sen_df(text):
+	sd = get_sen_dict(text)
+	s = pd.Series(sd, name = "length")
+	del sd
+	s.index.name = "gloss"
+	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence 
+					 	  #and the setence length respectively
+	return df
+	
+def main(upper_bound, lang_name):
+	text = extract_texts(upper_bound)
+	df = get_sen_df(text)
+	df.to_csv(lang_name + "_df.csv")
+
+
+if __name__ == "__main__":
+	main(int(sys.argv[1]), sys.argv[2])

### 2018-10-09 wiki_entropy/wiki_entropy.Rmd
Created pipeline for wiki_analysis

--- 2018-10-08 wiki_entropy.Rmd
@@
 title: Wikipedia Text Mining and Entropy Analysis
-author: Josef Klafka
+author: Josef Klafka and Dan Yurovsky
 date:
@@
 library(dplyr)
-
-library(SnowballC) #stemmer
 library(tidytext)
@@
 library(tokenizers)
-library(tau)
-library(gtools)
-library(reticulate)
-
-library(feather)
 
@@
 theme_set(theme_classic(base_size = 16))
-use_python("~/../../Library/Frameworks/Python.framework/Versions/3.7/bin/python3.7")
 ```
 
-```{python, include=FALSE, eval = T}
-import regex as re 
-import numpy as np
-import pandas as pd 
-import csv
+```{r python_background}
+## Here's where to find the languages of Wikipedia, and their prefixes https://en.wikipedia.org/wiki/List_of_Wikipedias#Detailed_list
 
-def extract_texts(upper_bound):
-	rs = ""
-	for i in range(upper_bound):
-		if i < 10:
-			number = "0" + str(i)
-		else:
-			number = str(i)
-		f = open("../misc/text/AA/wiki_" + number, 'r')
-		text = f.read()
-		rs += text
-	return rs
+LANGUAGE <- "Alemannic" ##This is the language you want to examine
 
-def get_sen_dict(text):
-	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:–'&*#/—“»]|\d+|\(|\)|", "", text) 
-	text = re.sub('"', '', text)
-	#just the words from the articles
-	text = text.replace('\n', " ")
-	sens = re.split("[.!?]", text.lower())
-	sen_dict = {}
-	for sen in sens:
-		length = len(re.findall("\w+", sen))
-		if length > 0:
-			sen_dict[sen] = length
-	return sen_dict
-
-def get_sen_df(text):
-	sd = get_sen_dict(text)
-	s = pd.Series(sd, name = "length")
-	del sd
-	s.index.name = "gloss"
-	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence 
-					 	  #and the setence length respectively
-	return df
-
-text = extract_texts(15)
-am_df = get_sen_df(text)
-
-am_df.to_csv("am_df.csv")
+system(paste("python3 wiki_entropy.py", LANGUAGE, sep = " "))
 ```
 
-```{r slope_prep, include=FALSE, eval = F}
-# extract_stem_tokens <- function(role, min_length, utterances) {
-#   
-#   REPS <- 100
-#   
-#   if (role == "child") {
-#     lengths <- utterances %>%
-#       filter(speaker_role == "Target_Child") %>%
-#       mutate(length = str_count(stem, " ") + 1)
-# 
-#   } else {
-#     lengths <- utterances %>%
-#       filter(speaker_role != "Target_Child") %>%
-#       mutate(length = str_count(stem, " ") + 1)
-#   }
-#   
-#   tokens <- lengths %>% 
-#     filter(length >= min_length) %>%
-#     mutate(utterance_id = 1:n()) %>%
-#     unnest_tokens(word, stem) %>%
-#     group_by(utterance_id) %>%
-#     mutate(word_order = 1:n()) %>%
-#     group_by(length, word_order, word) %>%
-#     summarise(n = n()) 
-# }
-
+```{r slope_analysis, include=FALSE, eval = F}
 extract_wiki_tokens <- function(min_length, max_length, utterances) {
@@
     mutate(word_order = 1:n()) %>%
-    # group_by(utterance_id) %>%
     group_by(length, word_order, word) %>%
-    # filter(word_order <= length) %>% 
-    summarise(n = n()) # %>%
-    # left_join(utterances) %>%
-    # select(gloss, n, length) %>%
-    # View()
+    summarise(n = n()) 
 }
 
-df <- read_csv("~/Documents/Labwork/Joe/Positional-Information/ar_df.csv")
+df <- read_csv("~/Documents/Labwork/Joe/Positional-Information/wiki_entropy/als_df.csv")
 
@@
 
-ar_slopes <- get_wiki_slopes(min_length = 6, max_length = 50, utterances=df, 1000)
-
-#get_slopes("not child", 6, py$xmf_df, "gloss")
-
-sub_tokens <- xmf_df %>% 
-     mutate(gloss = str_trim(gloss)) %>%
-     mutate(length = str_count(gloss, " +") +1) %>%
-     filter(length >= min_length) %>%
-     filter(length <= max_length) %>%
-     mutate(utterance_id = 1:n())
-
-sub_tokens %>%
-     unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
-     group_by(utterance_id) %>%
-     mutate(word_order = 1:n()) %>%
-     group_by(utterance_id) %>%
-     summarise(n = n()) %>%
-     # filter(n != 6) %>%
-     left_join(sub_tokens) %>%
-     select(gloss, n, length) %>%
-     View()
-
+als_slopes <- get_wiki_slopes(min_length = 6, max_length = 50, utterances=df, 100)
 ```
@@
 ```{r slope_code, include = F, eval=F}
-
 start_slope <- function(df) {
@@
 
-```{r slope_analysis, include = F, eval = F}
+```{r analyze_entropies, include=FALSE, eval = F}
+get_entropies <- function(sen_length, df) {
+
+  tokens <- df %>%
+    filter(length == sen_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, sentence) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(word_order, word) %>%
+    filter(word_order <= length)
+
+  tokens %>%
+    summarise(n = n()) %>%
+    tidyboot(summary_function = function(x) x %>% 
+               summarise(entropy = entropy(n, unit = "log2")),
+             statistics_functions = function(x) x %>%
+             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
+    mutate(length = sen_length)
+}
+```
+
+```{r slope_stuff, include = F, eval = F}
 
@@
 
-```{r analyze_entropies, include=FALSE, eval = F}
-get_entropies <- function(sen_length, df) {
+```{r garbage}
+# extract_stem_tokens <- function(role, min_length, utterances) {
+#   
+#   REPS <- 100
+#   
+#   if (role == "child") {
+#     lengths <- utterances %>%
+#       filter(speaker_role == "Target_Child") %>%
+#       mutate(length = str_count(stem, " ") + 1)
+# 
+#   } else {
+#     lengths <- utterances %>%
+#       filter(speaker_role != "Target_Child") %>%
+#       mutate(length = str_count(stem, " ") + 1)
+#   }
+#   
+#   tokens <- lengths %>% 
+#     filter(length >= min_length) %>%
+#     mutate(utterance_id = 1:n()) %>%
+#     unnest_tokens(word, stem) %>%
+#     group_by(utterance_id) %>%
+#     mutate(word_order = 1:n()) %>%
+#     group_by(length, word_order, word) %>%
+#     summarise(n = n()) 
+# }
+#get_slopes("not child", 6, py$xmf_df, "gloss")
 
-  tokens <- df %>%
-    filter(length == sen_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, sentence) %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    group_by(word_order, word) %>%
-    filter(word_order <= length)
-
-  tokens %>%
-    summarise(n = n()) %>%
-    tidyboot(summary_function = function(x) x %>% 
-               summarise(entropy = entropy(n, unit = "log2")),
-             statistics_functions = function(x) x %>%
-             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
-    mutate(length = sen_length)
-}
+# sub_tokens <- xmf_df %>% 
+#      mutate(gloss = str_trim(gloss)) %>%
+#      mutate(length = str_count(gloss, " +") +1) %>%
+#      filter(length >= min_length) %>%
+#      filter(length <= max_length) %>%
+#      mutate(utterance_id = 1:n())
+# 
+# 
+# sub_tokens %>%
+#      unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
+#      group_by(utterance_id) %>%
+#      mutate(word_order = 1:n()) %>%
+#      group_by(utterance_id) %>%
+#      summarise(n = n()) %>%
+#      filter(n != 6) %>%
+#      left_join(sub_tokens) %>%
+#      select(gloss, n, length) %>%
+#      View()
 

### 2018-10-09 wiki_entropy.py

@@ -1,50 +1,0 @@
-import regex as re 
-import numpy as np
-import pandas as pd 
-import csv
-import sys
-
-def extract_texts(upper_bound):
-	rs = ""
-	for i in range(upper_bound):
-		if i < 10:
-			number = "0" + str(i)
-		else:
-			number = str(i)
-		f = open("../misc/text/AA/wiki_" + number, 'r')
-		text = f.read()
-		rs += text
-	return rs
-
-
-def get_sen_dict(text):
-	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:–'&*#/—“»]|\d+|\(|\)|\[|\]", "", text) 
-	text = re.sub('"', '', text)
-	#just the words from the articles
-	text = text.replace('\n', " ")
-	sens = re.split("[.!?]", text.lower())
-	sen_dict = {}
-	for sen in sens:
-		length = len(re.findall("\w+", sen))
-		if length > 0:
-			sen_dict[sen] = length
-	return sen_dict
-
-
-def get_sen_df(text):
-	sd = get_sen_dict(text)
-	s = pd.Series(sd, name = "length")
-	del sd
-	s.index.name = "gloss"
-	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence 
-					 	  #and the setence length respectively
-	return df
-	
-def main(upper_bound, lang_name):
-	text = extract_texts(upper_bound)
-	df = get_sen_df(text)
-	df.to_csv(lang_name + "_df.csv")
-
-
-if __name__ == "__main__":
-	main(int(sys.argv[1]), sys.argv[2])

### 2018-10-09 wiki_entropy/wiki_entropy.py

@@ -1,0 +1,73 @@
+import regex as re 
+import numpy as np
+import pandas as pd 
+import csv
+import sys
+import json
+import subprocess
+import urllib.request
+
+with open("language_dict.json", 'r') as f:
+	file = f.read()
+LANG_DICT = json.loads(file)
+
+
+def extract_texts(upper_bound):
+	rs = ""
+	for i in range(upper_bound):
+		if i < 10:
+			number = "0" + str(i)
+		else:
+			number = str(i)
+		f = open("text/AA/wiki_" + number, 'r')
+		text = f.read()
+		rs += text
+	return rs
+
+
+def get_sen_dict(text):
+	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:–'&*#/—“»]|\d+|\(|\)|\[|\]", "", text) 
+	text = re.sub('"', '', text)
+	#just the words from the articles
+	text = text.replace('\n', " ")
+	sens = re.split("[.!?]", text.lower())
+	sen_dict = {}
+	for sen in sens:
+		length = len(re.findall("\w+", sen))
+		if length > 0:
+			sen_dict[sen] = length
+	return sen_dict
+
+
+def get_sen_df(text):
+	sd = get_sen_dict(text)
+	s = pd.Series(sd, name = "length")
+	del sd
+	s.index.name = "gloss"
+	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence 
+					 	  #and the setence length respectively
+	return df
+	
+
+def main(lang_name):
+	lang_prefix = LANG_DICT[lang_name]
+	url = "https://dumps.wikimedia.org/other/cirrussearch/current/" \
+		+ lang_prefix + "wiki-20181001-cirrussearch-content.json.gz"	
+	urllib.request.urlretrieve(url, "datafile")
+	
+	subprocess.call(["wikiextractor/cirrus_extract.py", "datafile"])
+
+	directory = subprocess.check_output(["ls", "text/AA"]).decode("utf-8")
+	highest_filenum = int(max(re.findall("\d\d", directory)))
+	upper_bound = min(highest_filenum, 50)
+	# name=$(ls | sort -V | tail -n 1) echo ${name: -2}
+	text = extract_texts(upper_bound)
+	df = get_sen_df(text)
+	df.to_csv(lang_prefix + "_df.csv")
+	del df
+	subprocess.call(["rm", "-r", "text"])
[…]

### 2018-10-10 wiki_entropy/wiki_entropy.py
Added comments, cleaned up code

@@ -7,12 +7,18 @@ import json
 import subprocess
 import urllib.request
 
+
 with open("language_dict.json", 'r') as f:
 	file = f.read()
-LANG_DICT = json.loads(file)
+LANG_DICT = json.loads(file) #maps languages to their wikipedia url prefixes
 
 
 def extract_texts(upper_bound):
+	'''
+	Given an upper_bound on file numbers, extracts the raw text from the files
+	in the text/AA folder. That folder is given as output by the cirrus_extract
+	program. 
+	'''
 	rs = ""
 	for i in range(upper_bound):
 		if i < 10:
@@ -26,7 +32,12 @@ def extract_texts(upper_bound):
 
 
 def get_sen_dict(text):
-	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:–'&*#/—“»]|\d+|\(|\)|\[|\]", "", text) 
+	'''
+	Given the output of extract_texts as a string, constructs a 
+	dictionary mapping each sentence in the string to its length in words.
+	'''
+	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:–'&*#/—“»]|\d+|\(|\)|\[|\]", \
+		"", text) 
 	text = re.sub('"', '', text)
 	#just the words from the articles
 	text = text.replace('\n', " ")
@@ -40,6 +51,10 @@ def get_sen_dict(text):
 
 
 def get_sen_df(text):
+	'''
+	Given the output of extract_texts as a string, returns a Pandas dataframe
+	of the output of get_sen_dict. 
+	'''
 	sd = get_sen_dict(text)
 	s = pd.Series(sd, name = "length")
 	del sd
@@ -50,20 +65,21 @@ def get_sen_df(text):
 	
 
 def main(lang_name):
-	lang_prefix = LANG_DICT[lang_name]
+	lang_prefix = LANG_DICT[lang_name] #gets the wikipedia url prefix for lang_name
 	url = "https://dumps.wikimedia.org/other/cirrussearch/current/" \
 		+ lang_prefix + "wiki-20181001-cirrussearch-content.json.gz"	
-	urllib.request.urlretrieve(url, "datafile")
+	urllib.request.urlretrieve(url, "datafile") #downloads the dump for that language
 	
 	subprocess.call(["wikiextractor/cirrus_extract.py", "datafile"])
 
[…]

### 2018-10-23 english_unigram&ce&mi.Rmd
Created pipeline for wiki_analysis
changes

@@
 
-tkns <- extract_tokens("child", 6, 20, prov_utterances)
+tkns <- new_get_tokens("child", 6, 20, prov_utterances)
 
@@
 
-mi <- map(counts, all_pairwise_mi)
+mi <- map(counts, all_pairwise_mi) %>% bind_rows()
 
@@
 
-ggplot(mi[[4]], aes(x = pos1, y = mi, color = as.factor(pos2))) + 
+ggplot(mi, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
   facet_wrap(~length) + 

### 2018-10-23 wiki_entropy/wiki_entropy.Rmd
touched up wiki_entropy
changes

@@
 ```{r python_background}
-## Here's where to find the languages of Wikipedia, and their prefixes https://en.wikipedia.org/wiki/List_of_Wikipedias#Detailed_list
+# Here's where to find the languages of Wikipedia and their prefixes: https://en.wikipedia.org/wiki/List_of_Wikipedias#Detailed_list
 
-LANGUAGE <- "Alemannic" ##This is the language you want to examine
+LANGUAGE <- "Dutch" # This is the language you want to examine
 
@@
 ```{r slope_analysis, include=FALSE, eval = F}
+df <- read_csv("~/Documents/Labwork/Joe/Positional-Information/wiki_entropy/ru_df.csv")
+
+wiki_tokens <- extract_wiki_tokens(6, 50, df)
+
+wiki_tokens %>%
+    filter(word_order %in% c(1, 2)) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2"), n = sum(n)) %>%
+    spread(word_order, entropy) %>%
+    group_by(length) %>%
+    mutate(diff = `2` - `1`) %>%
+  ggplot(aes(x = length, y = diff)) +
+  geom_point(aes(size = n)) +
+  geom_smooth(se = F) +
+  scale_x_log10() #log length stats
+
+REPS <- 100
+
+ru_slopes <- get_wiki_slopes(min_length = 6, max_length = 50, utterances=df, REPS)
+```
+
+```{r slope_code, include = F, eval=F}
 extract_wiki_tokens <- function(min_length, max_length, utterances) {
@@
 
-df <- read_csv("~/Documents/Labwork/Joe/Positional-Information/wiki_entropy/als_df.csv")
-
-wiki_tokens <- extract_wiki_tokens(6, 50, df)
-
-wiki_tokens %>%
-    filter(word_order %in% c(1,2)) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2"), n = sum(n)) %>%
-    spread(word_order, entropy) %>%
-    group_by(length) %>%
-    mutate(diff = `2` - `1`) %>%
-  ggplot(aes(x = length, y = diff)) +
-  geom_point(aes(size = n)) +
-  geom_smooth(se = F) +
-  scale_x_log10() #log length stats
-
-als_slopes <- get_wiki_slopes(min_length = 6, max_length = 50, utterances=df, 100)
-```
-
-```{r slope_code, include = F, eval=F}
 start_slope <- function(df) {

### 2018-10-23 wiki_entropy/wiki_entropy.py

@@ -67,7 +67,7 @@ def get_sen_df(text):
 def main(lang_name):
 	lang_prefix = LANG_DICT[lang_name] #gets the wikipedia url prefix for lang_name
 	url = "https://dumps.wikimedia.org/other/cirrussearch/current/" \
-		+ lang_prefix + "wiki-20181001-cirrussearch-content.json.gz"	
+		+ lang_prefix + "wiki-20181008-cirrussearch-content.json.gz"	
 	urllib.request.urlretrieve(url, "datafile") #downloads the dump for that language
 	
 	subprocess.call(["wikiextractor/cirrus_extract.py", "datafile"])

### 2018-10-23 wiki_entropy/wiki_entropy.py
Fixed issue with cirrus-extract.py

@@ -60,25 +60,25 @@ def get_sen_df(text):
 	del sd
 	s.index.name = "gloss"
 	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence 
-					 	  #and the setence length respectively
+					 	  #and the sentence length respectively
 	return df
 	
 
 def main(lang_name):
 	lang_prefix = LANG_DICT[lang_name] #gets the wikipedia url prefix for lang_name
 	url = "https://dumps.wikimedia.org/other/cirrussearch/current/" \
-		+ lang_prefix + "wiki-20181008-cirrussearch-content.json.gz"	
+		+ lang_prefix + "wiki-20181015-cirrussearch-content.json.gz"	
 	urllib.request.urlretrieve(url, "datafile") #downloads the dump for that language
 	
-	subprocess.call(["wikiextractor/cirrus_extract.py", "datafile"])
+	subprocess.call(["wiki_extract/cirrus_extract.py", "datafile"])
 
 	directory = subprocess.check_output(["ls", "text/AA"]).decode("utf-8")
-	highest_filenum = int(max(re.findall("\d\d", directory)))
+	highest_filenum = max(int(max(re.findall("\d\d", directory))), 1)
 	upper_bound = min(highest_filenum, 50)
 
 	text = extract_texts(upper_bound)
 	df = get_sen_df(text)
-	df.to_csv(lang_prefix + "_df.csv")
+	df.to_csv(lang_prefix + "_df.csv", index = False)
 
 	del df
 	subprocess.call(["rm", "-r", "text"])

### 2018-10-23 wiki_entropy/wiki_extract/cirrus_extract.py

@@ -1,0 +1,245 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+#
+# =============================================================================
+#  Version: 1.00 (December 15, 2015)
+#  Author: Giuseppe Attardi (attardi@di.unipi.it), University of Pisa
+#
+# =============================================================================
+#  Copyright (c) 2015. Giuseppe Attardi (attardi@di.unipi.it).
+# =============================================================================
+#  This file is part of Tanl.
+#
+#  Tanl is free software; you can redistribute it and/or modify it
+#  under the terms of the GNU General Public License, version 3,
+#  as published by the Free Software Foundation.
+#
+#  Tanl is distributed in the hope that it will be useful,
+#  but WITHOUT ANY WARRANTY; without even the implied warranty of
+#  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+#  GNU General Public License for more details.
+#
+#  You should have received a copy of the GNU General Public License
+#  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+# =============================================================================
+
+"""Wikipedia Cirrus Extractor:
+Extracts and cleans text from a Wikipedia Cirrus dump and stores output in a
+number of files of similar size in a given directory.
+Each file will contain several documents in the format:
+
+    <doc id="" url="" title="">
+        ...
+        </doc>
+
+"""
+
+import sys, os.path, time
+import re
+import json
+import argparse
+import bz2
+import gzip
+import logging
+
+# Program version
+version = '1.00'
+
+urlbase = 'http://it.wikipedia.org/'
+
+# ----------------------------------------------------------------------
+
+class NextFile(object):
+    """
+    Synchronous generation of next available file name.
+    """
+
+    filesPerDir = 100
+
+    def __init__(self, path_name):
+        self.path_name = path_name
+        self.dir_index = -1
+        self.file_index = -1
+
+    def next(self):
+        self.file_index = (self.file_index + 1) % NextFile.filesPerDir
+        if self.file_index == 0:
+            self.dir_index += 1
+        dirname = self._dirname()
+        if not os.path.isdir(dirname):
+            os.makedirs(dirname)
+        return self._filepath()
+
+    def _dirname(self):
+        char1 = self.dir_index % 26
+        char2 = self.dir_index / 26 % 26
+        return os.path.join(self.path_name, '%c%c' % (ord('A') + int(char2), ord('A') + int(char1)))
+
+    def _filepath(self):
[…]

### 2018-10-29 childes/childes_ce_and_mi.Rmd
Reorganized files, deleted redundant and obsfucating code

--- 2018-09-10 nonenglish_ce&mi.Rmd
@@
 
-```{r presentation}
+```{r ce_and_mi_functions}
 
@@
 
-all_conditional_entropy <- function(df) {
-  max_len <- max(df$word_order)
-  perms <- permutations(max_len, 2, 1:max_len, repeats.allowed = F) 
-  combs <-  setNames(split(perms, seq(nrow(perms))), rownames(perms))
-
-  map(combs, ~conditional_entropy(df, .x[1], .x[2])) %>%
-    bind_rows()
-
-}
-
-conditional_entropy <- function(df, X, Y) { #Y is conditioning variable
+position_ce <- function(df, X, Y) { #Y is conditioning variable
 
@@
 
+all_pairwise_ce <- function(df) {
+  max_len <- max(df$word_order)
+  perms <- permutations(max_len, 2, 1:max_len, repeats.allowed = F) 
+  combs <-  setNames(split(perms, seq(nrow(perms))), rownames(perms))
+
+  map(combs, ~position_ce(df, .x[1], .x[2])) %>%
+    bind_rows()
+
+}
+
 get_ce <- function(utterances, role) {
@@
     map(~filter(.,word_order <= length)) %>%
-    map(all_conditional_entropy) %>%
+    map(all_pairwise_ce) %>%
     bind_rows()
@@
   return(ce)
-}
-
-all_pairwise_mi <- function(df) {
-  max_len <- max(df$word_order)
-
-  combs <- combn(1:max_len, 2, simplify = F)
-
-  map(combs, ~position_mi(df, .x[1], .x[2])) %>%
-    bind_rows()
-
 }
@@
 
+all_pairwise_mi <- function(df) {
+  max_len <- max(df$word_order)
+
+  combs <- combn(1:max_len, 2, simplify = F)
+
+  map(combs, ~position_mi(df, .x[1], .x[2])) %>%
+    bind_rows()
+
+}
+
 get_mi <- function(utterances, role) {
@@
 
-```{r work_progress}
+```{r example_run}
 
@@
 
-wagner_ce_children <- get_ce(wagner_utterances, "Target_Child")
-wagner_ce_adults <- get_ce(wagner_utterances, "not child")
-
-ggplot(wagner_ce_children, aes(x = X, y = ce, color = as.factor(Y))) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Conditional entropy vs. conditioned random variable (Wagner children)")
-
-ggplot(wagner_ce_adults, aes(x = X, y = ce, color = as.factor(Y))) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Conditional entropy vs. conditioned random variable (Wagner adults)")
-
-okayama_mi_children <- get_mi(okayama_utterances, "Target_Child")
-okayama_mi_adults <- get_mi(okayama_utterances, "not child")
-
-ggplot(okayama_mi_children, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
-  facet_wrap(~length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Mutual information vs. conditioned position (Okayama children)")
-
-ggplot(okayama_mi_adults, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
-  facet_wrap(~length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Mutual information vs. conditioned position (Okayama adults)")
-
-okayama_ce_children <- get_ce(okayama_utterances, "Target_Child")
-okayama_ce_adults <- get_ce(okayama_utterances, "not child")
-
-ggplot(okayama_ce_children, aes(x = X, y = ce, color = as.factor(Y))) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Conditional entropy vs. conditioned random variable (Okayama children)")
-
-ggplot(okayama_ce_adults, aes(x = X, y = ce, color = as.factor(Y))) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Conditional entropy vs. conditioned random variable (Okayama adults)")
-
-miipro_mi_children <- get_mi(miipro_utterances, "Target_Child")
-miipro_mi_adults <- get_mi(miipro_utterances, "not child")
-
-ggplot(miipro_mi_children, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
-  facet_wrap(~length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Mutual information vs. conditioned position (Miipro children)")
-
-ggplot(miipro_mi_adults, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
-  facet_wrap(~length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Mutual information vs. conditioned position (Miipro adults)")
-
-miipro_ce_children <- get_ce(miipro_utterances, "Target_Child")
-miipro_ce_adults <- get_ce(miipro_utterances, "not child")
-
-ggplot(miipro_ce_children, aes(x = X, y = ce, color = as.factor(Y))) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Conditional entropy vs. conditioned random variable (Miipro children)")
-
-ggplot(miipro_ce_adults, aes(x = X, y = ce, color = as.factor(Y))) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Conditional entropy vs. conditioned random variable (Miipro adults)")
-
-zhou_mi_children <- get_mi(zhou_dinner, "Target_Child")
-zhou_mi_adults <- get_mi(zhou_dinner, "not child")
-
-ggplot(zhou_mi_children, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
-  facet_wrap(~length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Mutual information vs. conditioned position (Zhou children)")
-
-ggplot(zhou_mi_adults, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
-  facet_wrap(~length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Mutual information vs. conditioned position (Zhou adults)")
-
-zhou_ce_children <- get_ce(zhou_dinner, "Target_Child")
-zhou_ce_adults <- get_ce(zhou_dinner, "not child")
-
-ggplot(zhou_ce_children, aes(x = X, y = ce, color = as.factor(Y))) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Conditional entropy vs. conditioned random variable (Zhou children)")
-
-ggplot(zhou_ce_adults, aes(x = X, y = ce, color = as.factor(Y))) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Conditional entropy vs. conditioned random variable (Zhou adults)")
-
-leveille_ce_children <- get_ce(leveille_utterances, "Target_Child")
-leveille_ce_adults <- get_ce(leveille_utterances, "not child")
-
-ggplot(leveille_ce_children, aes(x = X, y = ce, color = as.factor(Y))) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Conditional entropy vs. conditioned random variable (Leveille children)")
-
-ggplot(leveille_ce_adults, aes(x = X, y = ce, color = as.factor(Y))) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Conditional entropy vs. conditioned random variable (Leveille adults)")
 ```

### 2018-10-29 childes/childes_entropy.Rmd
Reorganized files, deleted redundant and obsfucating code

--- 2018-10-04 english_unigram&ce&mi.Rmd
@@
   tokens <- lengths %>% 
-    filter(length == target_length) %>%
+    filter(length >= min_length) %>%
+    filter(length <= max_length) %>%
     mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, stem) %>%
+    unnest_tokens(word, stem, token = stringr::str_split, pattern = " +") %>%
     group_by(utterance_id) %>%
@@
     group_by(length, word_order, word) %>%
+    filter(word_order <= length) %>% 
     summarise(n = n()) 
@@
   tokens <- lengths %>% 
-    filter(length == target_length) %>%
+    filter(length >= min_length) %>%
+    filter(length <= max_length) %>%
     mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss) %>%
+    unnest_tokens(word, stem, token = stringr::str_split, pattern = " +") %>%
     group_by(utterance_id) %>%
@@
     summarise(n = n()) 
+}
+
+#compute the unigram entropy measure from the counts of each word in each position
+unigram_entropy <- function(tokens) {
+
+  tokens %>%
+    group_by(word_order, word) %>%
+    summarise(n = n()) %>%
+    tidyboot(summary_function = function(x) x %>% 
+               summarise(entropy = entropy(n, unit = "log2")),
+             statistics_functions = function(x) x %>%
+             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
+    mutate(role = role, length = sen_length)
 }
@@
 }
-
 ```
 
-```{r log_lengths}
-tokens <- extract_tokens("not child", 6, 30, prov_utterances)
+Compute Providence Corpus Estimates
+```{r calc_pos_entropy}
+tokens <- extract_stem_tokens("not child", 6, 30, prov_utterances)
 
@@
   scale_x_log10() #log length stats
-
-extract_tokens <- function(role, min_length, max_length, utterances) {
-
-  if (role == "child") {
-    lengths <- utterances %>%
-      filter(speaker_role == "Target_Child") %>%
-      mutate(length = str_count(stem, " +") +1)
-
-} else {
-    lengths <- utterances %>%
-      filter(speaker_role != "Target_Child") %>%
-     mutate(length = str_count(stem, " +") +1)
-}
-
-  tokens <- lengths %>% 
-    filter(length >= min_length) %>%
-    filter(length <= max_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, stem, token = stringr::str_split, pattern = " +") %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    group_by(length, word_order, word) %>%
-    filter(word_order <= length) %>% 
-    summarise(n = n()) 
-}
-
-shiro_slopes <- map(6:10, ~get_slopes("child", .x, shiro_utterances, "gloss"))
-
-shiro_slopes <- get_slopes("child", 6, shiro_utterances, "gloss") #what we would expect
-shiro_slopes <- get_slopes("not child", 6, shiro_utterances, "gloss")
-wagner_slopes <- get_slopes("child", 6, wagner_utterances, "gloss")
-okayama_slopes <- get_slopes("child", 6, okayama_utterances, "gloss")
-zhou_slopes <- get_slopes("child", 6, zhou_utterances, "stem")
-
-start_time <- Sys.time()
-samples <- replicate(1000, start_slope(tokens))  %>%
-  as_data_frame()
-end_time <- Sys.time()
-
-end_time - start_time
-
-start_slopes <- samples %>%
-  summarise(mean = mean(value), ci_upper = quantile(value, .975), 
-            ci_lower = quantile(value, .025))
-
-```
-
-```{r slope_analysis}
-
-new_get_tokens <- function(role, min_length, max_length, utterances) {
-
-  if (role == "child") {
-        lengths <- utterances %>%
-            filter(speaker_role != "Target_Child") %>%
-            mutate(length = str_count(gloss, " ") + 1)
-
-  } else {
-
-        lengths <- utterances %>%
-            filter(speaker_role != "Target_Child") %>%
-            mutate(length = str_count(gloss, " ") + 1)
-  }
-
-  sub_tokens <- lengths %>% 
-     mutate(gloss = str_trim(gloss)) %>%
-     mutate(length = str_count(gloss, " +") +1) %>%
-     filter(length >= min_length) %>%
-     filter(length <= max_length) %>%
-     mutate(utterance_id = 1:n()) %>%
-     unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
-     group_by(utterance_id) %>%
-     mutate(word_order = 1:n()) %>%
-     group_by(length, word_order, word) %>%
-     filter(word_order <= length) %>% 
-     summarise(n = n()) 
-}
-
-min_length <- 6
-max_length <- 20
-
-lengths <- prov_utterances %>%
-       filter(speaker_role != "Target_Child") %>%
-       mutate(length = str_count(gloss, " ") + 1)
-
-sub_tokens <- lengths %>% 
-     mutate(gloss = str_trim(gloss)) %>%
-     mutate(length = str_count(gloss, " +") +1) %>%
-     filter(length >= min_length) %>%
-     filter(length <= max_length) %>%
-     mutate(utterance_id = 1:n())
-
-sub_tokens %>%
-     unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
-     group_by(utterance_id) %>%
-     mutate(word_order = 1:n()) %>%
-     group_by(utterance_id) %>%
-     summarise(n = n()) %>%
-     # filter(n != 6) %>%
-     left_join(sub_tokens) %>%
-     select(gloss, n, length) %>%
-     View()
-
-```
-
-Compute Providence Corpus Estimates
-```{r calc_pos_entropy}
-get_entropies <- function(role, sen_length, utterances) {
-
-  if(role == "Target_Child") {
-    sub_utterances <- utterances %>%
-      filter(speaker_role == "Target_Child") %>%
-      mutate(length = str_count(stem, " ") + 1) # uses stems of words, not original words
-  } else {
-     sub_utterances <- utterances %>%
-      filter(speaker_role != "Target_Child") %>%
-      mutate(length = str_count(stem, " ") + 1)
-  }
-
-  # 
-  # 
-  # ggplot(adult_utterances, aes(x = length)) +
-  #   geom_histogram(fill = "white", color = "black")
-
-  tokens <- sub_utterances %>%
-    filter(length == sen_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, stem) %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    group_by(word_order, word)
-
-  tokens %>%
-    summarise(n = n()) %>%
-    tidyboot(summary_function = function(x) x %>% 
-               summarise(entropy = entropy(n, unit = "log2")),
-             statistics_functions = function(x) x %>%
-             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
-    mutate(role = role, length = sen_length)
-}
 
@@
 
-new_utterances <- prov_utterances %>% filter(speaker_role=="Target_Child") %>% mutate(len = str_count(stem, " ") + 1)
-#toks <- new_utterances %>%
-
-ggplot(new_utterances, aes(x=len)) +
-  geom_histogram(binwidth = 1)
 ```
 
-compute estimates
-```{r calc_pos_entropy}
-#get counts of each word in each position of utterances of length sen_length
-#from the collection of CHILDES utterances in utterances with speaker_role in role
-get_counts <- function(role, sen_length, utterances) {
-
-  if(role == "Target_Child") {
-    sub_utterances <- utterances %>%
-      filter(speaker_role == "Target_Child") %>%
-      mutate(length = str_count(stem, " ") + 1) # uses stems of words, not original words
-  } else {
-     sub_utterances <- utterances %>%
-      filter(speaker_role != "Target_Child") %>%
-      mutate(length = str_count(stem, " ") + 1)
-  }
-
-  sub_utterances %>%
-    filter(length == sen_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, stem) %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n())
-}
-
-#compute the unigram entropy measure from the counts of each word in each position
-unigram_entropy <- function(counts) {
-
-  counts %>%
-    group_by(word_order, word) %>%
-    summarise(n = n()) %>%
-    tidyboot(summary_function = function(x) x %>% 
-               summarise(entropy = entropy(n, unit = "log2")),
-             statistics_functions = function(x) x %>%
-             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
-    mutate(role = role, length = sen_length)
-}
+English Mutual Information
+```{r mutual_information}
 
@@
 
-entropies <- map(2, ~ get_counts("not child", .x, prov_utterances) %>%
-                   bind_rows %>%
-                   unigram_entropy) %>%
-  bind_rows()
-
-ggplot(entropies, aes(x = word_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  geom_pointrange() +
-  geom_smooth(se = F)
-
-entropies_child <- map(2:10, ~ get_entropies("Target_Child", .x, prov_utterances)) %>%
-  bind_rows()
-
-ggplot(entropies_child, aes(x = word_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  geom_pointrange() +
-  geom_smooth(se = F)
-
-entropies_all <- bind_rows(entropies, entropies_child)
-
-ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper, color = role)) +
-  facet_wrap(~ length) + 
-  geom_pointrange(position = position_dodge(.25)) +
-  geom_smooth(se = F) +
-  ggtitle("Child and Adult Positional Entropy vs. Word Position (Providence corpus)")
-
-new_utterances <- prov_utterances %>% filter(speaker_role=="Target_Child") %>% mutate(len = str_count(stem, " ") + 1)
-#toks <- new_utterances %>%
-
-ggplot(new_utterances, aes(x=len)) +
-  geom_histogram(binwidth = 1)
 ```
@@
 #compute all Conditional Entropy from all one-way pairs of word positions
-all_conditional_entropy <- function(df) {
+all_pairwise_ce <- function(df) {
   max_len <- max(df$word_order)
-  #combs <- combn(1:max_len, 2, simplify = F)
+
   perms <- permutations(max_len, 2, 1:max_len, repeats.allowed = F) 
@@
 #perform the actual Conditional Entropy calculation
-conditional_entropy <- function(df, X, Y) { #Y is conditioning variable
+get_ce <- function(df, X, Y) { #Y is conditioning variable
 
@@
 
-ce <- map(counts, all_conditional_entropy)
+ce <- map(counts, all_conditional_entropy) %>% bind_rows()
 
@@
 
-ggplot(ce_br, aes(x = X, y = ce, color = as.factor(Y))) + 
+ggplot(ce, aes(x = X, y = ce, color = as.factor(Y))) + 
   facet_wrap(~ length) + 
@@
 ```
-
-N-Gram frequencies (saved for later)
-```{r ngram_frequencies}
-m_prov <- prov_utterances %>%
-          filter(speaker_role == "Target_Child") %>%
-          mutate(length = str_count(stem, " ") + 1) %>% 
-          filter(length >= 2)
-
-b_prov <- m_prov %>% mutate(bigrams = paste(unlist(textcnt(stem, n = 2, split = " ", method = "string")), collapse = ",")) #get all the counts of bigrams from all of the dataframe
-counts <- b_prov$bigrams[1] %>% strsplit(split=',') %>% unlist() %>% as.numeric()
-fileConn <- file("counts.txt")
-writeLines(b_prov$bigrams[1], fileConn)
-close(fileConn)
-
-b_prov <- m_prov %>% mutate(bigrams = paste(unlist(names(textcnt(stem, n = 2, split = " ", method = "string"))), collapse = ",")) #get all of the bigrams from the entire dataframe
-bigrams <- b_prov$bigrams[1] %>% strsplit(split=',') %>% unlist()
-fileConn <- file("bigram_words.txt")
-writeLines(b_prov$bigrams[1], fileConn)
-close(fileConn)
-
-```

### 2018-10-29 ldp/LDP_entropy.Rmd
Reorganized files, deleted redundant and obsfucating code

--- 2018-09-14 LDP_entropy.Rmd
@@
 title: Non-English Conditional Entropy and Mutual Information
-author: Joe Klafka
+author: Josef Klafka
 date: "`r Sys.date()`"

### 2018-10-29 wiki_entropy/wiki_entropy.Rmd
Reorganized files, deleted redundant and obsfucating code

@@
 author: Josef Klafka and Dan Yurovsky
-date:
+date: "`r Sys.Date()`"
 output: 
@@
 
-system(paste("python3 wiki_entropy.py", LANGUAGE, sep = " "))
+system(paste("python3 get_wiki_df.py", LANGUAGE, sep = " "))
 ```
@@
 ```{r slope_analysis, include=FALSE, eval = F}
-df <- read_csv("~/Documents/Labwork/Joe/Positional-Information/wiki_entropy/ru_df.csv")
+df <- read_csv("wiki_df.csv")
 
@@
 
+# plot the entropy slope for sentences of length min_length to max_length
 wiki_tokens %>%
@@
 
-  tokens <- df %>%
-    filter(length == sen_length) %>%
+  tokens <- utterances %>% 
+    mutate(gloss = str_trim(gloss)) %>%
+    mutate(length = str_count(gloss, pattern = " +") + 1) %>% 
+    filter(length >= min_length) %>%
+    filter(length <= max_length) %>%
     mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, sentence) %>%
+    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
     group_by(utterance_id) %>%
     mutate(word_order = 1:n()) %>%
-    group_by(word_order, word) %>%
-    filter(word_order <= length)
+    group_by(length, word_order, word) %>%
+    summarise(n = n())
 
@@
 ```
-
-```{r slope_stuff, include = F, eval = F}
-
-xmf_df <- read_csv("xmf_df.csv")
-
-#get_slopes("not child", 6, py$xmf_df, "gloss")
-
-gtx <- extract_wiki_tokens(6, 1000, xmf_df)
-
-gtx %>%
-    filter(word_order %in% c(1,2)) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2"), n = sum(n)) %>%
-    spread(word_order, entropy) %>%
-    group_by(length) %>%
-    mutate(diff = `2` - `1`) %>%
-  ggplot(aes(x = length, y = diff)) +
-  geom_point(aes(size = n)) +
-  geom_smooth(se = F) +
-  scale_x_log10() #log length stats
-
-get_wiki_slopes(min_length = 6, max_length = 50, utterances=xmf_df)
-
-```
-
-```{r garbage}
-# extract_stem_tokens <- function(role, min_length, utterances) {
-#   
-#   REPS <- 100
-#   
-#   if (role == "child") {
-#     lengths <- utterances %>%
-#       filter(speaker_role == "Target_Child") %>%
-#       mutate(length = str_count(stem, " ") + 1)
-# 
-#   } else {
-#     lengths <- utterances %>%
-#       filter(speaker_role != "Target_Child") %>%
-#       mutate(length = str_count(stem, " ") + 1)
-#   }
-#   
-#   tokens <- lengths %>% 
-#     filter(length >= min_length) %>%
-#     mutate(utterance_id = 1:n()) %>%
-#     unnest_tokens(word, stem) %>%
-#     group_by(utterance_id) %>%
-#     mutate(word_order = 1:n()) %>%
-#     group_by(length, word_order, word) %>%
-#     summarise(n = n()) 
-# }
-#get_slopes("not child", 6, py$xmf_df, "gloss")
-
-# sub_tokens <- xmf_df %>% 
-#      mutate(gloss = str_trim(gloss)) %>%
-#      mutate(length = str_count(gloss, " +") +1) %>%
-#      filter(length >= min_length) %>%
-#      filter(length <= max_length) %>%
-#      mutate(utterance_id = 1:n())
-# 
-# 
-# sub_tokens %>%
-#      unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
-#      group_by(utterance_id) %>%
-#      mutate(word_order = 1:n()) %>%
-#      group_by(utterance_id) %>%
-#      summarise(n = n()) %>%
-#      filter(n != 6) %>%
-#      left_join(sub_tokens) %>%
-#      select(gloss, n, length) %>%
-#      View()
-
-#entropies <- map(6:12, ~get_entropies(.x, py$xmf_df)) %>%
-#  bind_rows()
-
-#ggplot(entropies, aes(x = word_order, y = empirical_entropy,
-#                      ymin = ci_lower, ymax = ci_upper)) +
-#  facet_wrap(~ length) + 
-#  geom_pointrange() +
-#  geom_smooth(se = F)
-
-# extract_stem_tokens <- function(role, min_length, utterances) {
-#   
-#   if (role == "child") {
-#     lengths <- utterances %>%
-#       filter(speaker_role == "Target_Child") %>%
-#       mutate(length = str_count(stem, " ") + 1)
-# 
-# } else {
-#     lengths <- utterances %>%
-#       filter(speaker_role != "Target_Child") %>%
-#       mutate(length = str_count(stem, " ") + 1)
-# }
-#   
-#   tokens <- lengths %>% 
-#     filter(length >= min_length) %>%
-#     mutate(utterance_id = 1:n()) %>%
-#     unnest_tokens(word, stem) %>%
-#     group_by(utterance_id) %>%
-#     mutate(word_order = 1:n()) %>%
-#     group_by(length, word_order, word) %>%
-#     summarise(n = n()) 
-# }
-# 
-# extract_gloss_tokens <- function(role, min_length, utterances) {
-#   
-#   if (role == "child") {
-#     lengths <- utterances %>%
-#       filter(speaker_role == "Target_Child") %>%
-#       mutate(length = str_count(gloss, " ") + 1)
-# 
-# } else {
-#     lengths <- utterances %>%
-#       filter(speaker_role != "Target_Child") %>%
-#       mutate(length = str_count(gloss, " ") + 1)
-# }
-#   
-#   tokens <- lengths %>% 
-#     filter(length >= min_length) %>%
-#     mutate(utterance_id = 1:n()) %>%
-#     unnest_tokens(word, gloss) %>%
-#     group_by(utterance_id) %>%
-#     mutate(word_order = 1:n()) %>%
-#     group_by(length, word_order, word) %>%
-#     filter(word_order <= length) %>% 
-#     summarise(n = n()) 
-# }
-```

### 2018-10-29 wiki_entropy/get_wiki_df.py

@@ -16,8 +16,8 @@ LANG_DICT = json.loads(file) #maps languages to their wikipedia url prefixes
 def extract_texts(upper_bound):
 	'''
 	Given an upper_bound on file numbers, extracts the raw text from the files
-	in the text/AA folder. That folder is given as output by the cirrus_extract
-	program. 
+	in the text/AA folder. That folder is given as output by the 
+	cirrus_extract program. 
 	'''
 	rs = ""
 	for i in range(upper_bound):
@@ -65,26 +65,33 @@ def get_sen_df(text):
 	
 
 def main(lang_name):
-	lang_prefix = LANG_DICT[lang_name] #gets the wikipedia url prefix for lang_name
-	url = "https://dumps.wikimedia.org/other/cirrussearch/current/" \
-		+ lang_prefix + "wiki-20181015-cirrussearch-content.json.gz"	
-	urllib.request.urlretrieve(url, "datafile") #downloads the dump for that language
-	
+	'''
+	Given the English name of a language, constructs a dataframe with the 
+	sentences of a random subset of articles in that language from Wikipedia, 
+	and stores the dataframe in a csv in the working directory. 
+	'''
+	lang_prefix = LANG_DICT[lang_name] 
+	url = "https://dumps.wikimedia.org/other/cirrussearch/20181022/" \
+		+ lang_prefix + "wiki-20181022-cirrussearch-content.json.gz"	
+	urllib.request.urlretrieve(url, "datafile") 
 	subprocess.call(["wiki_extract/cirrus_extract.py", "datafile"])
 
+	# find the number of article files we want to combine
 	directory = subprocess.check_output(["ls", "text/AA"]).decode("utf-8")
 	highest_filenum = max(int(max(re.findall("\d\d", directory))), 1)
 	upper_bound = min(highest_filenum, 50)
 
+	# construct the df
 	text = extract_texts(upper_bound)
 	df = get_sen_df(text)
-	df.to_csv(lang_prefix + "_df.csv", index = False)
+	df.to_csv("wiki_df.csv", index = False)
 
-	del df
-	subprocess.call(["rm", "-r", "text"])
-	subprocess.call(["rm", "datafile"])
+	# delete the text files and datafile 
+	# may be several gigabytes large
+	subprocess.call(["rm", "-r", "text"]) 
+	subprocess.call(["rm", "datafile"]) 
 
 
 if __name__ == "__main__":
-	main(sys.argv[1])
+	main(sys.argv[1]) # takes the English name of a language as sole argument
 	

### 2018-11-01 wiki_entropy/wiki_entropy.Rmd
normal approximation

@@
 
-LANGUAGE <- "Dutch" # This is the language you want to examine
+LANGUAGE <- "Alemannic" # This is the language you want to examine
 
-system(paste("python3 get_wiki_df.py", LANGUAGE, sep = " "))
+system(paste("python get_wiki_df.py", LANGUAGE, sep = " "))
 ```
 
-```{r slope_analysis, include=FALSE, eval = F}
-df <- read_csv("wiki_df.csv")
-
-wiki_tokens <- extract_wiki_tokens(6, 50, df)
-
-# plot the entropy slope for sentences of length min_length to max_length
-wiki_tokens %>%
-    filter(word_order %in% c(1, 2)) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2"), n = sum(n)) %>%
-    spread(word_order, entropy) %>%
-    group_by(length) %>%
-    mutate(diff = `2` - `1`) %>%
-  ggplot(aes(x = length, y = diff)) +
-  geom_point(aes(size = n)) +
-  geom_smooth(se = F) +
-  scale_x_log10() #log length stats
-
-REPS <- 100
-
-ru_slopes <- get_wiki_slopes(min_length = 6, max_length = 50, utterances=df, REPS)
-```
-
-```{r slope_code, include = F, eval=F}
+Extract tokens
+```{r extract_tokens}
 extract_wiki_tokens <- function(min_length, max_length, utterances) {
@@
 }
+```
 
+Slope helpers
+```{r slope_helpers, include = F, eval=F}
 start_slope <- function(df) {
@@
     pull(diff)
+
+}
+
+estimate_slope_normal <- function(df, first_pos, second_pos) {
+
+   mutate(word_order = if_else(word_order == length-1, 2,
+                                if_else(word_order == length -2, 1,
+                                        0))) %>%
+
+  df %>%
+    ungroup() %>%
+    filter(word_order >= first_pos, word_order <= second_pos) %>%
+    mutate(word_order = word_order - max(word_order)) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    lm(entropy ~ length + I(length^2) + word_order, data  = .) %>%
+    tidy() %>%
+    filter(term == "word_order")
 
@@
 }
+```
 
+```{r get_slopes}
 get_wiki_slopes <- function(min_length, max_length, utterances, REPS) {
@@
 
+```{r slope_analysis, include=FALSE, eval = F}
+df <- read_csv("wiki_df.csv")
+
+wiki_tokens <- extract_wiki_tokens(6, 50, df)
+
+# plot the entropy slope for sentences of length min_length to max_length
+wiki_tokens %>%
+    filter(word_order %in% c(1, 2)) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2"), n = sum(n)) %>%
+    spread(word_order, entropy) %>%
+    group_by(length) %>%
+    mutate(diff = `2` - `1`) %>%
+  ggplot(aes(x = length, y = diff)) +
+  geom_point(aes(size = n)) +
+  geom_smooth(se = F) +
+  scale_x_log10() #log length stats
+
+REPS <- 100
+
+ru_slopes <- get_wiki_slopes(min_length = 6, max_length = 50, utterances=df, REPS)
+```
+
 ```{r analyze_entropies, include=FALSE, eval = F}

### 2018-11-18 ldp/LDP_entropy.Rmd
ldp ses analysis code

@@
 utterances <- get_table(connection, "utterances") %>% collect() #to get the "utterances" table as a dataframe named utterances
+home_visits <- get_table(connection, "home_visits")  %>% collect() #information about SES, work, education and other demographic stuff
 DBI::dbListTables(connection) #to list the tables in the collection
@@
 
+```{r filter_utterances}
+get_ses_utterances <- function(status, utterances, home_visits) {
+  income <- home_visits %>% filter(income_category %in% status)
+
+  ids <- unique(income$id) %>% unlist()
+
+  utterances %>% filter(subject %in% ids)
+}
+
+ses_utts <- get_ses_utterances(c(6), utterances, home_visits)
+```
+
 ```{r ldp_analysis}
-get_unigram_entropies <- function(role, person, sen_length, utterances) {
+get_unigram_entropies <- function(role, sen_length, utterances) {
 
@@
     sub_utterances <- utterances %>%
-      filter(subject == person) %>% 
       filter(c_utts_orig != "") %>%
@@
      sub_utterances <- utterances %>%
-      filter(subject == person) %>% 
       filter(p_utts_orig != "") %>%
@@
 
-entropies <- map(2:10, ~get_unigram_entropies("child", , .x, utterances)) %>%
+entropies <- map(2:10, ~get_unigram_entropies("child", .x, ses_utts)) %>%
   bind_rows()

### 2018-11-18 wiki_entropy/get_wiki_df.py

@@ -1,11 +1,12 @@
-import regex as re 
 import numpy as np
+import subprocess
+import regex as re 
 import pandas as pd 
 import csv
 import sys
 import json
-import subprocess
 import urllib.request
+import os
 
 
 with open("language_dict.json", 'r') as f:
@@ -36,7 +37,7 @@ def get_sen_dict(text):
 	Given the output of extract_texts as a string, constructs a 
 	dictionary mapping each sentence in the string to its length in words.
 	'''
-	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:–'&*#/—“»]|\d+|\(|\)|\[|\]", \
+	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:&]|\d+|\(|\)|\[|\]", \
 		"", text) 
 	text = re.sub('"', '', text)
 	#just the words from the articles
@@ -71,14 +72,15 @@ def main(lang_name):
 	and stores the dataframe in a csv in the working directory. 
 	'''
 	lang_prefix = LANG_DICT[lang_name] 
-	url = "https://dumps.wikimedia.org/other/cirrussearch/20181022/" \
-		+ lang_prefix + "wiki-20181022-cirrussearch-content.json.gz"	
+	url = "https://dumps.wikimedia.org/other/cirrussearch/20181112/" \
+		+ lang_prefix + "wiki-20181112-cirrussearch-content.json.gz"	
 	urllib.request.urlretrieve(url, "datafile") 
-	subprocess.call(["wiki_extract/cirrus_extract.py", "datafile"])
+	subprocess.call(["python", "wiki_extract/cirrus_extract.py", "datafile"])
 
 	# find the number of article files we want to combine
-	directory = subprocess.check_output(["ls", "text/AA"]).decode("utf-8")
-	highest_filenum = max(int(max(re.findall("\d\d", directory))), 1)
+	print(os.listdir("text/AA"))
+	highest_filenum = int(sorted(os.listdir("text/AA"))[-1][-2:])
+	#highest_filenum = max(int(max(re.findall("\d\d", directory))), 1)
 	upper_bound = min(highest_filenum, 50)
 
 	# construct the df

### 2018-11-18 wiki_entropy/wiki_extract/cirrus_extract.py

@@ -179,7 +179,7 @@ def process_dump(input_file, out_file, file_size, file_compress):
             url = urlbase + 'wiki?curid=' + id
             header = '<doc id="%s" url="%s" title="%s">\n' % (id, url, title)
             page = header + title + '\n\n' + text + '\n</doc>\n'
-            output.write(page)
+            output.write(page.encode('utf-8'))
 
 # ----------------------------------------------------------------------
 

### 2018-12-06 wiki_entropy/wiki_entropy.Rmd
ldp ses analysis code
relative slope analysis

@@
 
-Slope helpers
-```{r slope_helpers, include = F, eval=F}
-start_slope <- function(df) {
-
- df %>%
-    filter(word_order %in% c(1,2)) %>%
-    sample_frac(1, replace = T) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    spread(word_order, entropy) %>%
-    ungroup() %>%
-    mutate(diff = `2` - `1`) %>%
-    summarise(diff = mean(diff)) %>%
-    pull(diff)
-
-}
-
+normal estimation
+```{r}
 estimate_slope_normal <- function(df, first_pos, second_pos) {
@@
     filter(term == "word_order")
+
+}
+```
+
+relative estimation
+```{r}
+
+df <- read_csv("wiki_df.csv")
+
+df_tokens <- extract_wiki_tokens(6, 50, df)
+
+NUM_SECTIONS = 5
+
+get_quantiles <- function(df, num_sections) {
+  quantile(df$word_order, probs = seq(0, 1, 1/num_sections)) %>% 
+    round() %>%
+    as_data_frame() %>%
+    rename(word_order = value)
+}
+
+pos_list <- df_tokens %>%
+  group_by(length) %>%
+  distinct(word_order) %>%
+  split(.$length) %>%
+  map(~get_quantiles(.x, NUM_SECTIONS)) %>%
+  bind_rows(.id = "length") %>%
+  mutate(length = as.numeric(length))
+
+relative_slopes <- function(df, pos_list) {
+
+  get_slope <- function(x) {
+
+    selected_pos <- pos_list %>%
+      group_by(length) %>%
+      slice(x:(x+1)) %>%
+      group_by(length) %>%
+      summarise(min = min(word_order),
+                max = max(word_order)) %>%
+      split(.$length) %>%
+      map(~seq(.x$min, .x$max, 1) %>% as_data_frame) %>%
+      bind_rows(.id = "length") %>%
+      rename(word_order = value) %>%
+      mutate(length = as.numeric(length))
+
+    df %>%
+      inner_join(selected_pos, by = c("length", "word_order")) %>% 
+      group_by(length, word_order) %>%
+      summarise(entropy = entropy(n, unit = "log2")) %>%
+      lm(entropy ~ length + I(length^2) ++ word_order, data  = .) %>%
+      tidy() %>%
+      filter(term == "word_order")
+  }
+
+  divides <- pos_list %>%
+    group_by(length) %>%
+    summarise(n = n()) %>%
+    summarise(n = mean(n)) %>%
+    pull()
+
+  map(1:((divides)-1), get_slope) %>%
+    bind_rows(.id = "cut")
+
+}
+
+relative_slopes(df_tokens, pos_list)
+
+```
+
+Slope helpers
+```{r slope_helpers, include = F, eval=F}
+start_slope <- function(df) {
+
+ df %>%
+    filter(word_order %in% c(1,2)) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `2` - `1`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
 

### 2018-12-06 wiki_entropy/wiki_entropy.Rmd
typo fix in relative slope

@@
       summarise(entropy = entropy(n, unit = "log2")) %>%
-      lm(entropy ~ length + I(length^2) ++ word_order, data  = .) %>%
+      lm(entropy ~ length + I(length^2) + word_order, data  = .) %>%
       tidy() %>%

### 2018-12-07 wiki_entropy/wiki_clustering.Rmd
initial clustering on absolute data

--- 2018-11-01 wiki_entropy/wiki_entropy.Rmd
@@
 ```{r setup, include=FALSE}
-# load packages
-library(knitr)
 library(tidyverse)
-library(directlabels)
-library(dplyr)
-library(tidytext)
-library(entropy)
-library(tidyboot)
-library(tokenizers)
+library(googlesheets)
 
@@
 
-```{r python_background}
-# Here's where to find the languages of Wikipedia and their prefixes: https://en.wikipedia.org/wiki/List_of_Wikipedias#Detailed_list
+```{r clustering}
+##read in googlesheet
+wiki_gs <- gs_title("Wikipedia_absolute")
+wkpd <- gs_read(wiki_gs) %>%
+  as.data.frame() %>%
+  column_to_rownames(var = "language")
 
-LANGUAGE <- "Alemannic" # This is the language you want to examine
-
-system(paste("python get_wiki_df.py", LANGUAGE, sep = " "))
-```
-
-Extract tokens
-```{r extract_tokens}
-extract_wiki_tokens <- function(min_length, max_length, utterances) {
-
-  tokens <- utterances %>% 
-    mutate(gloss = str_trim(gloss)) %>%
-    mutate(length = str_count(gloss, pattern = " +") + 1) %>% 
-    filter(length >= min_length) %>%
-    filter(length <= max_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    group_by(length, word_order, word) %>%
-    summarise(n = n()) 
-}
-```
-
-Slope helpers
-```{r slope_helpers, include = F, eval=F}
-start_slope <- function(df) {
-
- df %>%
-    filter(word_order %in% c(1,2)) %>%
-    sample_frac(1, replace = T) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    spread(word_order, entropy) %>%
-    ungroup() %>%
-    mutate(diff = `2` - `1`) %>%
-    summarise(diff = mean(diff)) %>%
-    pull(diff)
-
-}
-
-estimate_slope_normal <- function(df, first_pos, second_pos) {
-
-   mutate(word_order = if_else(word_order == length-1, 2,
-                                if_else(word_order == length -2, 1,
-                                        0))) %>%
-
-  df %>%
-    ungroup() %>%
-    filter(word_order >= first_pos, word_order <= second_pos) %>%
-    mutate(word_order = word_order - max(word_order)) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    lm(entropy ~ length + I(length^2) + word_order, data  = .) %>%
-    tidy() %>%
-    filter(term == "word_order")
-
-}
-
-second_slope <- function(df) {
-
- df %>%
-    ungroup() %>%
-    mutate(word_order %in% c(2, 3)) %>%
-    filter(word_order > 0) %>%
-    sample_frac(1, replace = T) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    spread(word_order, entropy) %>%
-    ungroup() %>%
-    mutate(diff = `3` - `2`) %>%
-    summarise(diff = mean(diff)) %>%
-    pull(diff)
-
-}
-
-#note this measurement only makes sense if utterance length is at least 6
-mid_slope <- function(df) {
-
- df %>%
-    ungroup() %>%
-    mutate(word_order = if_else(word_order == length-2, 2,
-                                if_else(word_order == 3, 1,
-                                        0))) %>%
-    filter(word_order > 0) %>%
-    group_by(length, word_order) %>%
-    sample_frac(1, replace = T) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    spread(word_order, entropy) %>%
-    ungroup() %>%
-    mutate(diff = `2` - `1`) %>%
-    summarise(diff = mean(diff)) %>%
-    pull(diff)
-
-}
-
-penult_slope <- function(df) {
-
- df %>%
-    ungroup() %>%
-    mutate(word_order = if_else(word_order == length-1, 2,
-                                if_else(word_order == length -2, 1,
-                                        0))) %>%
-    filter(word_order > 0) %>%
-    group_by(length, word_order) %>%
-    sample_frac(1, replace = T) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    spread(word_order, entropy) %>%
-    ungroup() %>%
-    mutate(diff = `2` - `1`) %>%
-    summarise(diff = mean(diff)) %>%
-    pull(diff)
-
-}
-
-end_slope <- function(df) {
-
- df %>%
-    ungroup() %>%
-    mutate(word_order = if_else(word_order == length, 2,
-                                if_else(word_order == length -1 , 1,
-                                        0))) %>%
-    filter(word_order > 0) %>%
-    group_by(length, word_order) %>%
-    sample_frac(1, replace = T) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    spread(word_order, entropy) %>%
-    ungroup() %>%
-    mutate(diff = `2` - `1`) %>%
-    summarise(diff = mean(diff)) %>%
-    pull(diff)
-
-}
-```
-
-```{r get_slopes}
-get_wiki_slopes <- function(min_length, max_length, utterances, REPS) {
-
-  tokens <- extract_wiki_tokens(min_length, max_length, utterances)
-
-  start_slopes <- replicate(REPS, start_slope(tokens)) %>% 
-    as_data_frame() %>%
-    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
-              ci_lower = quantile(value, .025))
-
-  second_slopes <- replicate(REPS, second_slope(tokens)) %>% 
-    as_data_frame() %>%
-    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
-              ci_lower = quantile(value, .025))
-
-  mid_slopes <- replicate(REPS, mid_slope(tokens)) %>% 
-    as_data_frame() %>%
-    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
-              ci_lower = quantile(value, .025))
-
-  penult_slopes <- replicate(REPS, penult_slope(tokens)) %>% 
-    as_data_frame() %>%
-    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
-              ci_lower = quantile(value, .025))
-
-  end_slopes <- replicate(REPS, end_slope(tokens)) %>% 
-    as_data_frame() %>%
-    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
-              ci_lower = quantile(value, .025))
-
-  rbind(start_slopes, second_slopes, mid_slopes, penult_slopes, end_slopes)
-}
+##do clustering
+wk_clust <- wkpd %>% 
+              dist() %>%
+              hclust("ward.D2") %>%
+              plot(hang = -1)
 
 ```
-
-```{r slope_analysis, include=FALSE, eval = F}
-df <- read_csv("wiki_df.csv")
-
-wiki_tokens <- extract_wiki_tokens(6, 50, df)
-
-# plot the entropy slope for sentences of length min_length to max_length
-wiki_tokens %>%
-    filter(word_order %in% c(1, 2)) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2"), n = sum(n)) %>%
-    spread(word_order, entropy) %>%
-    group_by(length) %>%
-    mutate(diff = `2` - `1`) %>%
-  ggplot(aes(x = length, y = diff)) +
-  geom_point(aes(size = n)) +
-  geom_smooth(se = F) +
-  scale_x_log10() #log length stats
-
-REPS <- 100
-
-ru_slopes <- get_wiki_slopes(min_length = 6, max_length = 50, utterances=df, REPS)
-```
-
-```{r analyze_entropies, include=FALSE, eval = F}
-get_entropies <- function(sen_length, df) {
-
-  tokens <- utterances %>% 
-    mutate(gloss = str_trim(gloss)) %>%
-    mutate(length = str_count(gloss, pattern = " +") + 1) %>% 
-    filter(length >= min_length) %>%
-    filter(length <= max_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    group_by(length, word_order, word) %>%
-    summarise(n = n())
-
-  tokens %>%
-    summarise(n = n()) %>%
-    tidyboot(summary_function = function(x) x %>% 
-               summarise(entropy = entropy(n, unit = "log2")),
-             statistics_functions = function(x) x %>%
-             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
-    mutate(length = sen_length)
-}
-```

### 2018-12-14 cogsci/position-info.Rmd
cogsci abstract draft, clustering code and eye-tracking work added

---
title: "A Language's Unigram Entropy Distribution Predicts Self-Paced Reading Times"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Josef Klafka} \\ \texttt{jklafka@uchicago.edu} \\ Department of Psychology \\ University of Chicago
    \And {\large \bf Daniel Yurovsky} \\ \texttt{yurovsky@uchicago.edu} \\ Department of Psychology \\ University of Chicago}

abstract: 
    "We aim to provide evidence against the popular Uniform Information Density 
    (Jaeger, 2010) and wrap-up effect (Kuperman et al., 2010) hypotheses using cross-linguistic 
    evidence from speech, text and eye-tracking corpora."

keywords:
    "Entropy; self-paced reading; information theory; language processing"

output: cogsci2016::cogsci_paper
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
```

# Uniform Information Density

The average distribution of information in English sentences-what is it? Genzel & Charniak (2002) looked at the log-probability of words occurring in sentences as you proceed through a paragraph or article and found that less common/less likely words occurred more frequently as you moved through the text. The relationship they found was roughly an affine linear function, leading them to propose the *constant entropy rate* (CER) principle: the entropy of words in a sentence, i.e. the uncertainty about what words will appear in a sentence, will increase at a constant, linear rate through a paragraph. This idea was used by Aylett & Turk (2004), Levy & Jaeger (2007) and Frank & Jaeger (2008) among others. The CER principle was adapted by Jaeger (2010) into the *uniform information density* (UID) principle: uniformation is evenly spread throughout a sentence. From a theoretical standpoint, this makes sense as the optimal distribution for information throughout a sentence: if you miss any word I’ve said, the rest of the information in the sentence is intact, and you lose no more than if you miss any other word I’d have said.

The UID hypothesis

## UID challenges

More recent work has stood out in contrast to the traditional UID perspective. Zhan and Levy (2018) study Mandarin Chinese classifier use, and find that the use of specific classifier, versus general classifiers, appears more often in cases where the production of the corresponding noun is more difficult than when the production is easier, as would be predicted by UID. The UID perspective is challenged in Yu et al. (2016), by performing an analysis of entropy by position in the text portion of the British National Corpus. They use the following formula for each word position X of sentences of fixed length k from the corpus, where each i is a word occurring in position X and pi is the number of times word i occurs in position X divided by the number of total words that occur in position X i.e. the number of sentences of length k.

$$H(X) = \sum\limits_w p(w)\log\big(p(w)\big)$$
Yu et al. (2016) refer to their distribution for English as a ‘three-step distribution’: relatively low entropy at the beginning of a sentence, then a jump, then flat entropy in the middle, a dip before the final position and a jump with the final word. View the figure below for a visual demonstration.

```{r yunigram, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "English entropy distribution for three sentence lengths from BNC"}
img <- png::readPNG("figs/yunigram.png")
grid::grid.raster(img)
```

This diverges from the UID account, which would predict a simple affine function. 
Our replication with CHILDES. Conditional entropy and mutual information. Link with Thiessen & Onnis (to appear). Mention Ferrer-i-Cancho (2017), which provides a theoretical basis for the end of the sentence being the center of information. 
Cross-linguistic variation. Our Wikipedia analysis
Link with Kuperman et al. (2010). Talk about challenges to the wrap-up account. Consequences for eye-tracking. Talk about Zhan and Levy (2018) and how current work in eye-tracking and language processing disputes the UID account.

# Formalities, Footnotes, and Floats

Use standard APA citation format. Citations within the text should
include the author's last name and year. If the authors' names are
included in the sentence, place only the year in parentheses, as in
[-@NewellSimon1972a], but otherwise place the entire reference in
parentheses with the authors and year separated by a comma
[@NewellSimon1972a]. List multiple references alphabetically and
separate them by semicolons [@ChalnickBillman1988a; @NewellSimon1972a]. 
Use the et. al. construction only after listing all the authors to a
publication in an earlier reference and for citations with four or
more authors.

For more information on citations in RMarkdown, see **[here](http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html#citations).**

## Footnotes

Indicate footnotes with a number\footnote{Sample of the first
footnote.} in the text. Place the footnotes in 9 point type at the
bottom of the page on which they appear. Precede the footnote with a
horizontal rule.\footnote{Sample of the second footnote.} You can also use 
markdown formatting to include footnotes using this syntax [^1].

[^1]: Sample of a markdown footnote.

## Figures

All artwork must be very dark for purposes of reproduction and should
not be hand drawn. Number figures sequentially, placing the figure
number and caption, in 10 point, after the figure with one line space
above the caption and one line space below it. If necessary, leave extra white space at
the bottom of the page to avoid splitting the figure and figure
caption. You may float figures to the top or bottom of a column, or
set wide figures across both columns.

## Two-column images

You can read local images using png package for example and plot 
it like a regular plot using grid.raster from the grid package. 
With this method you have full control of the size of your image. **Note: Image must be in .png file format for the readPNG function to work.**

You might want to display a wide figure across both columns. To do this, you change the `fig.env` chunk option to `figure*`. To align the image in the center of the page, set `fig.align` option to `center`. To format the width of your caption text, you set the `num.cols.cap` option to `2`.

```{r 2-col-image, fig.env = "figure*", fig.pos = "h", fig.width=4, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns. And the caption text is limited to 0.8 of the width of the document."}
img <- png::readPNG("figs/walrus.png")
grid::grid.raster(img)
```

## One-column images

Single column is the default option, but if you want set it explicitly, set `fig.env` to `figure`. Notice that the `num.cols` option for the caption width is set to `1`.

```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
img <- png::readPNG("figs/lab_logo_stanford.png")
grid::grid.raster(img)
```

## R Plots

You can use R chunks directly to plot graphs. And you can use latex floats in the
fig.pos chunk option to have more control over the location of your plot on the page. For more information on latex placement specifiers see **[here](https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions)**

```{r plot, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2, fig.height=2, fig.cap = "R plot" }
x <- 0:100
y <- 2 * (x + rnorm(length(x), sd = 3) + 3)

ggplot2::ggplot(data = data.frame(x, y), 
       aes(x = x, y = y)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

## Tables

Number tables consecutively; place the table number and title (in
10 point) above the table with one line space above the caption and
one line space below it, as in Table 1. You may float
tables to the top or bottom of a column, set wide tables across both
columns.

You can use the xtable function in the xtable package.

```{r xtable, results="asis"}
n <- 100
x <- rnorm(n)
y <- 2*x + rnorm(n)
out <- lm(y ~ x)

tab1 <- xtable::xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
                      caption = "This table prints across one column.")

print(tab1, type="latex", comment = F, table.placement = "H")
```

# Acknowledgements

Place acknowledgments (including funding information) in a section at
the end of the paper.

# References

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent

### 2018-12-14 cogsci/position-info.tex
cogsci abstract draft, clustering code and eye-tracking work added

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{A Language's Unigram Entropy Distribution Predicts Self-Paced Reading
Times}

\author{{\large \bf Josef Klafka} \\ \texttt{jklafka@uchicago.edu} \\ Department of Psychology \\ University of Chicago \And {\large \bf Daniel Yurovsky} \\ \texttt{yurovsky@uchicago.edu} \\ Department of Psychology \\ University of Chicago}

\begin{document}

\maketitle

\begin{abstract}
The abstract should be one paragraph, indented 1/8 inch on both sides,
in 9 point font with single spacing. The heading Abstract should be 10
point, bold, centered, with one line space below it. This one-paragraph
abstract section is required only for standard spoken papers and
standard posters (i.e., those presentations that will be represented by
six page papers in the Proceedings).

\textbf{Keywords:}
Entropy; self-paced reading; information theory; language processing
\end{abstract}

\section{Uniform Information Density}\label{uniform-information-density}

The average distribution of information in English sentences-what is it?
Genzel \& Charniak (2002) looked at the log-probability of words
occurring in sentences as you proceed through a paragraph or article and
found that less common/less likely words occurred more frequently as you
moved through the text. The relationship they found was roughly an
affine linear function, leading them to propose the \emph{constant
entropy rate} (CER) principle: the entropy of words in a sentence,
i.e.~the uncertainty about what words will appear in a sentence, will
increase at a constant, linear rate through a paragraph. This idea was
used by Aylett \& Turk (2004), Levy \& Jaeger (2007) and Frank \& Jaeger
(2008) among others. The CER principle was adapted by Jaeger (2008) into
the \emph{uniform information density} (UID) principle: uniformation is
evenly spread throughout a sentence. From a theoretical standpoint, this
makes sense as the optimal distribution for information throughout a
sentence: if you miss any word I've said, the rest of the information in
the sentence is intact, and you lose no more than if you miss any other
word I'd have said. Description of the UID perspective and what it
predicts for entropy distributions and what that means.

\subsection{Yu challenge}\label{yu-challenge}

The UID perspective is challenged in Yu et al. (2016), by performing an
analysis of entropy by position in the text portion of the British
National Corpus. They use the following formula for each word position X
of sentences of fixed length k from the corpus, where each i is a word
occurring in position X and pi is the number of times word i occurs in
position X divided by the number of total words that occur in position X
i.e.~the number of sentences of length k.

\[H(X) = \sum\limits_w p(w)\log\big(p(w)\big)\] Yu et al. (2016) refer
to their distribution for English as a `three-step distribution':
relatively low entropy at the beginning of a sentence, then a jump, then
flat entropy in the middle, a dip before the final position and a jump
with the final word. View the figure below for a visual demonstration.

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/yunigram-1}

}

\caption[English entropy distribution for three sentence lengths from BNC]{English entropy distribution for three sentence lengths from BNC}\label{fig:yunigram}
\end{figure*}
\end{CodeChunk}

This diverges from the UID account, which would predict a simple affine
function. Our replication with CHILDES. Conditional entropy and mutual
information. Link with Thiessen \& Onnis (to appear). Mention
Ferrer-i-Cancho (2017), which provides a theoretical basis for the end
of the sentence being the center of information. Cross-linguistic
variation. Our Wikipedia analysis Link with Kuperman et al. (2010).
Consequences for eye-tracking. Talk about Zhan and Levy (2018) and how
current work in eye-tracking and language processing disputes the UID
account. Our eye-tracking experiment.

\section{Formalities, Footnotes, and
Floats}\label{formalities-footnotes-and-floats}

Use standard APA citation format. Citations within the text should
include the author's last name and year. If the authors' names are
included in the sentence, place only the year in parentheses, as in
(1972), but otherwise place the entire reference in parentheses with the
authors and year separated by a comma (Newell \& Simon, 1972). List
multiple references alphabetically and separate them by semicolons
(Chalnick \& Billman, 1988; Newell \& Simon, 1972). Use the et. al.
construction only after listing all the authors to a publication in an
earlier reference and for citations with four or more authors.

For more information on citations in RMarkdown, see
\textbf{\href{http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html\#citations}{here}.}

\subsection{Footnotes}\label{footnotes}

Indicate footnotes with a number\footnote{Sample of the first
footnote.} in the text. Place the footnotes in 9 point type at the
bottom of the page on which they appear. Precede the footnote with a
horizontal rule.\footnote{Sample of the second footnote.} You can also
use markdown formatting to include footnotes using this
syntax.\footnote{Sample of a markdown footnote.}

\subsection{Figures}\label{figures}

All artwork must be very dark for purposes of reproduction and should
not be hand drawn. Number figures sequentially, placing the figure
number and caption, in 10 point, after the figure with one line space
above the caption and one line space below it. If necessary, leave extra
white space at the bottom of the page to avoid splitting the figure and
figure caption. You may float figures to the top or bottom of a column,
or set wide figures across both columns.

\subsection{Two-column images}\label{two-column-images}

You can read local images using png package for example and plot it like
a regular plot using grid.raster from the grid package. With this method
you have full control of the size of your image. \textbf{Note: Image
must be in .png file format for the readPNG function to work.}

You might want to display a wide figure across both columns. To do this,
you change the \texttt{fig.env} chunk option to \texttt{figure*}. To
align the image in the center of the page, set \texttt{fig.align} option
to \texttt{center}. To format the width of your caption text, you set
the \texttt{num.cols.cap} option to \texttt{2}.

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/2-col-image-1}

}

\caption[This image spans both columns]{This image spans both columns. And the caption text is limited to 0.8 of the width of the document.}\label{fig:2-col-image}
\end{figure*}
\end{CodeChunk}

\subsection{One-column images}\label{one-column-images}

Single column is the default option, but if you want set it explicitly,
set \texttt{fig.env} to \texttt{figure}. Notice that the
\texttt{num.cols} option for the caption width is set to \texttt{1}.

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/image-1}

}

\caption[One column image]{One column image.}\label{fig:image}
\end{figure}
\end{CodeChunk}

\subsection{R Plots}\label{r-plots}

You can use R chunks directly to plot graphs. And you can use latex
floats in the fig.pos chunk option to have more control over the
location of your plot on the page. For more information on latex
placement specifiers see
\textbf{\href{https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions}{here}}

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/plot-1}

}

\caption[R plot]{R plot}\label{fig:plot}
\end{figure}
\end{CodeChunk}

\subsection{Tables}\label{tables}

Number tables consecutively; place the table number and title (in 10
point) above the table with one line space above the caption and one
line space below it, as in Table 1. You may float tables to the top or
bottom of a column, set wide tables across both columns.

You can use the xtable function in the xtable package.

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
(Intercept) & -0.02 & 0.10 & -0.2 & 0.82 \\ 
  x & 1.95 & 0.10 & 19.2 & 0.00 \\ 
   \hline
\end{tabular}
\caption{This table prints across one column.} 
\end{table}

\section{Acknowledgements}\label{acknowledgements}

Place acknowledgments (including funding information) in a section at
the end of the paper.

\section{References}\label{references}

\setlength{\parindent}{-0.1in} \setlength{\leftskip}{0.125in} \noindent

\hypertarget{refs}{}
\hypertarget{ref-ChalnickBillman1988a}{}
Chalnick, A., \& Billman, D. (1988). Unsupervised learning of
correlational structure. In \emph{Proceedings of the tenth annual
conference of the cognitive science society} (pp. 510--516). Hillsdale,
NJ: Lawrence Erlbaum Associates.

\hypertarget{ref-NewellSimon1972a}{}
Newell, A., \& Simon, H. A. (1972). \emph{Human problem solving}.
Englewood Cliffs, NJ: Prentice-Hall.

\end{document}

### 2018-12-14 wiki_entropy/wiki_clustering.Rmd
cogsci abstract draft, clustering code and eye-tracking work added

@@
 library(googlesheets)
+library(lingtypology)
+library(knitr)
+library(klaR)
 
@@
 ##read in googlesheet
-wiki_gs <- gs_title("Wikipedia_absolute")
-wkpd <- gs_read(wiki_gs) %>%
+wiki_gs <- gs_title("Wikipedia_relative10")
+wkpd <- gs_read(wiki_gs) 
+LANGUAGES <- wkpd$Language
+
+wkpd <- wkpd %>% 
   as.data.frame() %>%
@@
               dist() %>%
-              hclust("ward.D2") %>%
-              plot(hang = -1)
+              hclust("ward.D2")
+
+wk_clust %>% plot(hang = -1)
+```
+
+```{r WALS_features}
+FEATURES <- c(1:144)
+for (i in c(1:144)) {
+  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
+} #get all WALS features in correct syntax
+
+NUM_CLUSTERS <- 4
+
+wf <- wals.feature(FEATURES)
+wf %>% View()
+wf <- wf[, -(1:3)]
+wf <- wf[, -(ncol(wf) - 1)]
+wf <- wf %>% 
+  filter(language %in% LANGUAGES)
+wf <- wf[, complete.cases(wf)] #this significantly trims down the dataset but is necessary for k-modes
+rv <- kmodes(wf[, -(ncol(wf))], NUM_CLUSTERS, iter.max=10, weighted = FALSE)
 

### 2018-12-17 wiki_entropy/wiki_clustering.Rmd
working on imputing missing WALS data

@@
 ##read in googlesheet
-wiki_gs <- gs_title("Wikipedia_relative10")
+wiki_gs <- gs_title("Wikipedia_absolute")
 wkpd <- gs_read(wiki_gs) 
@@
   as.data.frame() %>%
-  column_to_rownames(var = "language")
+  column_to_rownames(var = "Language")
 
@@
 
-```{r WALS_features}
+```{r WALS_imputation}
+#taken from https://www.tutorialspoint.com/r/r_mean_median_mode.htm
+getmode <- function(v) {
+   uniqv <- unique(v)
+   uniqv[which.max(tabulate(match(v, uniqv)))]
+}
+
+## imputation based on Gelman & Hill Chapter 25
+# http://www.stat.columbia.edu/~gelman/arm/missing.pdf
+random.imp <- function(a) {
+  missed <- is.na(a)
+  num_missed <- sum(missed)
+  a.obs <- a[!missed]
+  imputed <- a
+  imputed[missed] <- base::sample(a.obs, num_missed, replace=TRUE)
+  return(imputed)
+}
+
+impute <- function(a, a.impute) {
+  ifelse(is.na(a), a.impute, a)
+}
+
+wf.sub <- wals.feature(c("1A", "2A", "3A", "4A"))
+
+a1.mode <- getmode(wf.sub$`1A`[!is.na(wf.sub$`1A`)]) #gets the most common non-na value of column 1A
+
+n.sims <- 10
+for (s in 1:n.sims) {
+  a1 <- wf.sub$`1A`
+  a1[is.na(a1)] <- getmode(a1[!is.na(a1)])
+  lm.1 <- glm(a1 ~ unlist(a2.imp, use.names = FALSE) + unlist(a2.imp, use.names = TRUE) + unlist(a2.imp, use.names = TRUE), family = binomial)
+  pred.1 <- rnorm(n, predict(lm.1), sigma.hat(lm.1))
+  a1.imp <- impute(wf.sub[c("1A")], pred.1)
+}
+
+```
+
+```{r WALS_clustering}
 FEATURES <- c(1:144)

### 2018-12-18 wiki_entropy/wiki_clustering.Rmd
working on output of glm with factors

@@
 
-```{r WALS_imputation}
+```{r Gelman}
+## from http://www.stat.columbia.edu/~gelman/bda/regression.R
+
+sigma.hat <- function (object){
+  object.class <- class(object)[[1]]
+  if (object.class=="lm"){
+    sigma <- summary(object)$sigma
+    return (sigma)
+  }
+  else if (object.class=="glm"){
+    sigma <- summary(object, correlation=TRUE)$sigma
+    return (sigma)
+  }
+  else if (object.class=="lmer"){
+    object <- summary (object)
+    fcoef <- .Call("mer_fixef", object, PACKAGE = "lme4")
+    useScale <- object@devComp[8]
+    ngrps <- lapply(object@flist, function(x) length(levels(x)))
+    n.groupings <- length (ngrps)
+    varc <- VarCorr (object, useScale=useScale)
+    sc <- attr(varc, "sc")
+    recorr <- lapply(varc, function(el) el@factors$correlation)
+    reStdDev <- c(lapply(recorr, slot, "sd"), list(Residual = sc))
+    sigmas <- as.list (rep (NA, n.groupings+1))
+    sigmas[1] <- ifelse (useScale, sc, NA)
+    cors <- as.list (rep (NA, n.groupings+1))
+    names (sigmas) <- names (cors) <- c ("data", names (varc))
+    for (k in 1:n.groupings){
+      sigmas[[k+1]] <- reStdDev[[k]]
+      cors[[k+1]] <- as.matrix (recorr[[k]])
+      if (length (cors[[k+1]]) == 1) cors[[k+1]] <- NA
+    }
+    return (list (sigma=sigmas, cors=cors))
+  }
+}
+```
+
+```{r WALS_imputation_functions}
 #taken from https://www.tutorialspoint.com/r/r_mean_median_mode.htm
@@
 ## imputation based on Gelman & Hill Chapter 25
-# http://www.stat.columbia.edu/~gelman/arm/missing.pdf
 random.imp <- function(a) {
@@
 
-wf.sub <- wals.feature(c("1A", "2A", "3A", "4A"))
+get_factor <- function(cname) {
+  a.imp <- wf.sub[c(cname)] %>%
+    random.imp() %>%
+    unlist(use.names = FALSE)
+}
+```
+
+```{r WALS_imputation_code}
+FEATURES <- c(1:144)
+for (i in c(1:144)) {
+  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
+} #get all WALS features in correct syntax
+
+NUM_CLUSTERS <- 4
+
+wf <- wals.feature(FEATURES)
+wf.sub <- wf[c("1A", "2A", "3A", "4A", "language")]
+wf.sub <- wf.sub %>% 
+  filter(language %in% LANGUAGES) %>%
+  unique()
+
+for (i in 1:4) {
+  wf.sub[, i] <- factor(wf.sub[, i])
+}
+
+a1.imp <- get_factor("1A")
+a2.imp <- get_factor("2A")
+a3.imp <- get_factor("3A")
+a4.imp <- get_factor("4A")
 
@@
   a1[is.na(a1)] <- getmode(a1[!is.na(a1)])
-  lm.1 <- glm(a1 ~ unlist(a2.imp, use.names = FALSE) + unlist(a2.imp, use.names = TRUE) + unlist(a2.imp, use.names = TRUE), family = binomial)
-  pred.1 <- rnorm(n, predict(lm.1), sigma.hat(lm.1))
+  lm.1 <- glm(a1 ~ a2.imp + a3.imp + a4.imp, family = "binomial")
+  pred.1 <- rnorm(n, predict(lm.1), sigma.hat(lm.1)) ####
   a1.imp <- impute(wf.sub[c("1A")], pred.1)

### 2018-12-21 cogsci/position-info.Rmd
updated background and our analysis in Cogsci abstract

@@
 abstract: 
-    "We aim to provide evidence against the popular Uniform Information Density 
-    (Jaeger, 2010) and wrap-up effect (Kuperman et al., 2010) hypotheses using cross-linguistic 
-    evidence from speech, text and eye-tracking corpora."
+    "We provide evidence against the Uniform Information Density (Jaeger, 2010) 
+    and wrap-up effect hypotheses."
 
@@
 
-The average distribution of information in English sentences-what is it?
-Genzel & Charniak (2002) looked at the log-probability of words occurring in sentences as you proceed through a paragraph or article and found that less common/less likely words occurred more frequently as you moved through the text.
-The relationship they found was roughly an affine linear function, leading them to propose the *constant entropy rate* (CER) principle: the entropy of words in a sentence, i.e.
-the uncertainty about what words will appear in a sentence, will increase at a constant, linear rate through a paragraph.
-This idea was used by Aylett & Turk (2004), Levy & Jaeger (2007) and Frank & Jaeger (2008) among others.
-The CER principle was adapted by Jaeger (2010) into the *uniform information density* (UID) principle: uniformation is evenly spread throughout a sentence.
-From a theoretical standpoint, this makes sense as the optimal distribution for information throughout a sentence: if you miss any word I’ve said, the rest of the information in the sentence is intact, and you lose no more than if you miss any other word I’d have said.
+A major line of research in the current century has asked how speakers and writers distribute information in the linguistic material they produce.
+This begins with Claude Shannon in the years after World War II.
+Shannon (1948) defined "information" as a "reduction in uncertainty".
+Shannon followed by quantifying uncertainty in his measure of *entropy*: the amount of uncertainty on the outcome of a random variable.
+Shannon proposed that the most efficient method of sending information through a noisy channel is at a constant rate.
+Genzel & Charniak (2002) apply this principle to human communication: if people communicate information through speech and text optimally, then we should transmit information to one another at a constant rate.
+Genzel & Charniak created a distribution of estimates for sentence-level entropy, each sentence taken without context, for newspaper articles from the Penn Treebank corpus.
+Obtaining a roughly monotonically increasing linear distribution, Genzel & Charniak proposed the *Constant Entropy Rate* (CER) principle: the entropy rate in speech and text with context should be constant as sentence number increases.
+Aylett & Turk (2004) proposed a similar principle for the phonological level.
+
+Jaeger (2010) extends the CER principle to all levels of human communication: the *Uniform Information Density* (UID) hypothesis.
+Jaeger argues that speakers will try to distribute information more evenly over the course of an utterance, to be as close to channel capacity (in a Shannon sense) as possible.
+In particular, Jaeger singles out the production of "that" at the beginning of relative clauses in English and the use of contractions such as "he's" and "you're" versus "he is" and "you are" as instances where higher information density in the clause elicits the production of more material by speakers.
+
+The *wrap-up effect* is a popular hypothesis within the eye-tracking community.
+The hypothesis states that, in written text, sentence-final words are processed more slowly on average then sentence-medial or sentence-initial words, due to readers integrating information from the entire sentence in order to form a final, coherent thought expressed by the sentence, among other reasons.
+With these two ideas, information production distribution in spoken and written communication, and information integration in reading written texts, we move on to discuss cross-linguistic differences in the distribution of information across sentences.
 
@@
 
-```{r yunigram, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "English entropy distribution for three sentence lengths from BNC"}
-img <- png::readPNG("figs/yunigram.png")
+# Our Methods
+
+We used the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations.
+We first used the Brown and Providence English corpora from CHILDES.
+The Brown corpus contains conversations between three young children (over 1.5 years old and under 6 years old) and their families in the home.
+The Providence corpus contains transcriptions of audio/video files which recorded interactions between children between 1 and 3 years old and their parents in the home.
+We divided each corpus by speaker (child and non-child) and each sentence length within those divisions and computed the unigram entropy measure for each position within those sentence lengths.
+Plots for the entropy distributions are below.
+The adult and child unigram entropy distributions track one another extremely well, with the child distribution for each sentence length having relatively lower entropy.
+
+```{r brown_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Brown corpus entropy"}
+img <- png::readPNG("figs/positional_entropy_brown.png")
 grid::grid.raster(img)
@@
 
-This diverges from the UID account, which would predict a simple affine function. 
-Our replication with CHILDES.
-Conditional entropy and mutual information.
-Link with Thiessen & Onnis (to appear).
-Mention Ferrer-i-Cancho (2017), which provides a theoretical basis for the end of the sentence being the center of information.
-Cross-linguistic variation. Our Wikipedia analysis
-Link with Kuperman et al.
-(2010).
-Talk about challenges to the wrap-up account.
-Consequences for eye-tracking.
-Talk about Zhan and Levy (2018) and how current work in eye-tracking and language processing disputes the UID account.
+```{r providence_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Providence corpus unigram entropy"}
+img <- png::readPNG("figs/Prov_corpus_result.png")
+grid::grid.raster(img)
+```
+
+Unlike a UID account
+
+We found a similar distribution for Spanish and German corpora from CHILDES. The Spanish corpus we used was the XXX and the German corpus we used was the XXX.
 

### 2018-12-21 cogsci/position-info.tex
updated background and our analysis in Cogsci abstract

@@
 \begin{abstract}
-The abstract should be one paragraph, indented 1/8 inch on both sides,
-in 9 point font with single spacing. The heading Abstract should be 10
-point, bold, centered, with one line space below it. This one-paragraph
-abstract section is required only for standard spoken papers and
-standard posters (i.e., those presentations that will be represented by
-six page papers in the Proceedings).
+We provide evidence against the Uniform Information Density (Jaeger,
+2010) and wrap-up effect hypotheses.
 
@@
 
-The average distribution of information in English sentences-what is it?
-Genzel \& Charniak (2002) looked at the log-probability of words
-occurring in sentences as you proceed through a paragraph or article and
-found that less common/less likely words occurred more frequently as you
-moved through the text. The relationship they found was roughly an
-affine linear function, leading them to propose the \emph{constant
-entropy rate} (CER) principle: the entropy of words in a sentence,
-i.e.~the uncertainty about what words will appear in a sentence, will
-increase at a constant, linear rate through a paragraph. This idea was
-used by Aylett \& Turk (2004), Levy \& Jaeger (2007) and Frank \& Jaeger
-(2008) among others. The CER principle was adapted by Jaeger (2008) into
-the \emph{uniform information density} (UID) principle: uniformation is
-evenly spread throughout a sentence. From a theoretical standpoint, this
-makes sense as the optimal distribution for information throughout a
-sentence: if you miss any word I've said, the rest of the information in
-the sentence is intact, and you lose no more than if you miss any other
-word I'd have said. Description of the UID perspective and what it
-predicts for entropy distributions and what that means.
+A major line of research in the current century has asked how speakers
+and writers distribute information in the linguistic material they
+produce. This begins with Claude Shannon in the years after World War
+II. Shannon (1948) defined ``information'' as a ``reduction in
+uncertainty''. Shannon followed by quantifying uncertainty in his
+measure of \emph{entropy}: the amount of uncertainty on the outcome of a
+random variable. Shannon proposed that the most efficient method of
+sending information through a noisy channel is at a constant rate.
+Genzel \& Charniak (2002) apply this principle to human communication:
+if people communicate information through speech and text optimally,
+then we should transmit information to one another at a constant rate.
+Genzel \& Charniak created a distribution of estimates for
+sentence-level entropy, each sentence taken without context, for
+newspaper articles from the Penn Treebank corpus. Obtaining a roughly
+monotonically increasing linear distribution, Genzel \& Charniak
+proposed the \emph{Constant Entropy Rate} (CER) principle: the entropy
+rate in speech and text with context should be constant as sentence
+number increases. Aylett \& Turk (2004) proposed a similar principle for
+the phonological level.
 
-\subsection{Yu challenge}\label{yu-challenge}
+Jaeger (2010) extends the CER principle to all levels of human
+communication: the \emph{Uniform Information Density} (UID) hypothesis.
+Jaeger argues that speakers will try to distribute information more
+evenly over the course of an utterance, to be as close to channel
+capacity (in a Shannon sense) as possible. In particular, Jaeger singles
+out the production of ``that'' at the beginning of relative clauses in
+English and the use of contractions such as ``he's'' and ``you're''
+versus ``he is'' and ``you are'' as instances where higher information
+density in the clause elicits the production of more material by
+speakers.
 
-The UID perspective is challenged in Yu et al. (2016), by performing an
-analysis of entropy by position in the text portion of the British
-National Corpus. They use the following formula for each word position X
-of sentences of fixed length k from the corpus, where each i is a word
-occurring in position X and pi is the number of times word i occurs in
-position X divided by the number of total words that occur in position X
-i.e.~the number of sentences of length k.
+The \emph{wrap-up effect} is a popular hypothesis within the
+eye-tracking community. The hypothesis states that, in written text,
+sentence-final words are processed more slowly on average then
+sentence-medial or sentence-initial words, due to readers integrating
+information from the entire sentence in order to form a final, coherent
+thought expressed by the sentence, among other reasons. With these two
+ideas, information production distribution in spoken and written
+communication, and information integration in reading written texts, we
+move on to discuss cross-linguistic differences in the distribution of
+information across sentences.
+
+The UID hypothesis
+
+\subsection{UID challenges}\label{uid-challenges}
+
+More recent work has stood out in contrast to the traditional UID
+perspective. Zhan and Levy (2018) study Mandarin Chinese classifier use,
+and find that the use of specific classifier, versus general
+classifiers, appears more often in cases where the production of the
+corresponding noun is more difficult than when the production is easier,
+as would be predicted by UID. The UID perspective is challenged in Yu et
+al. (2016), by performing an analysis of entropy by position in the text
+portion of the British National Corpus. They use the following formula
+for each word position X of sentences of fixed length k from the corpus,
+where each i is a word occurring in position X and pi is the number of
+times word i occurs in position X divided by the number of total words
+that occur in position X i.e.~the number of sentences of length k.
 
@@
 
-This diverges from the UID account, which would predict a simple affine
-function. Our replication with CHILDES. Conditional entropy and mutual
-information. Link with Thiessen \& Onnis (to appear). Mention
-Ferrer-i-Cancho (2017), which provides a theoretical basis for the end
-of the sentence being the center of information. Cross-linguistic
-variation. Our Wikipedia analysis Link with Kuperman et al. (2010).
-Consequences for eye-tracking. Talk about Zhan and Levy (2018) and how
-current work in eye-tracking and language processing disputes the UID
-account. Our eye-tracking experiment.
+\section{Our Methods}\label{our-methods}
+
+We used the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora
+database of spoken adult-child conversations. We first used the Brown
+and Providence English corpora from CHILDES. The Brown corpus contains
+conversations between three young children (over 1.5 years old and under
+6 years old) and their families in the home. The Providence corpus
+contains transcriptions of audio/video files which recorded interactions
+between children between 1 and 3 years old and their parents in the
+home.
+
+\begin{CodeChunk}
+\begin{figure*}[h]
+
+{\centering \includegraphics{figs/brown_PE-1}
+
+}
+
+\caption[Brown corpus entropy]{Brown corpus entropy}\label{fig:brown_PE}
+\end{figure*}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{figure*}[h]
+
+{\centering \includegraphics{figs/providence_PE-1}
+
+}
+
+\caption[Providence corpus unigram entropy]{Providence corpus unigram entropy}\label{fig:providence_PE}
+\end{figure*}
+\end{CodeChunk}
+
+We found a similar distribution for Spanish and German corpora from
+CHILDES. The Spanish corpus we used was the XXX and the German corpus we
+used was the XXX.
 
@@
   \hline
-(Intercept) & -0.02 & 0.10 & -0.2 & 0.82 \\ 
-  x & 1.95 & 0.10 & 19.2 & 0.00 \\ 
+(Intercept) & -0.10 & 0.11 & -0.9 & 0.37 \\ 
+  x & 1.97 & 0.12 & 16.6 & 0.00 \\ 
    \hline

### 2018-12-22 wiki_entropy/wiki_clustering.Rmd
MCA_imputation section added

@@
 library(klaR)
+
+library(missMDA)
 
@@
 wk_clust %>% plot(hang = -1)
+```
+
+```{r MCA_imputation}
+wiki_gs <- gs_title("Wikipedia_absolute")
+wkpd <- gs_read(wiki_gs) 
+LANGUAGES <- wkpd$Language
+
+FEATURES <- c(1:144)
+for (i in c(1:144)) {
+  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
+} #get all WALS features in correct syntax
+
+wf <- wals.feature(FEATURES)
+wf.sub <- wf[c("1A", "2A", "3A", "4A", "language")]
+wf.sub <- wf.sub %>% 
+  filter(language %in% LANGUAGES) %>%
+  unique()
+
+for (i in 1:4) {
+  wf.sub[, i] <- factor(wf.sub[, i])
+}
+
+wf.sub <- wf.sub[rowSums(is.na(wf.sub)) != 4, ]
+rownames(wf.sub) <- wf.sub$language
+wf.sub$language <- NULL
+rv <- MIMCA(wf.sub, ncp = 4)
+
+imputed_wf <- as.data.frame(rv[[1]][[99]])
 ```

### 2019-01-07 childes/childes_entropy.Rmd
working on cogsci

@@
 
+get_counts <- function(role, sen_length, utterances) {
+
+  if(role == "Target_Child") {
+    sub_utterances <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1) # uses stems of words, not original words
+  } else {
+     sub_utterances <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1)
+  }
+
+  sub_utterances %>%
+    filter(length == sen_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, gloss) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    filter(word_order <= length)
+}
+
+get_mean_lengths <- function(role, sen_length, utterances) {
+
+  counts <- get_counts(role, sen_length, utterances)
+
+  lens <- counts %>% 
+    group_by(word_order, word) %>% 
+    mutate(word_len = str_length(word))
+
+  lens$word_len %>% 
+    aggregate(list(lens$word_order), mean)
+}
+
+get_entropies <- function(role, sen_length, utterances) {
+
+  if(role == "Target_Child") {
+    sub_utterances <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1) # uses stems of words, not original words
+  } else {
+     sub_utterances <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1)
+  }
+
+  tokens <- sub_utterances %>%
+    filter(length == sen_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, gloss) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(word_order, word) %>%
+    filter(word_order <= length)
+
+  tokens %>%
+    summarise(n = n()) %>%
+    tidyboot(summary_function = function(x) x %>% 
+               summarise(entropy = entropy(n, unit = "log2")),
+             statistics_functions = function(x) x %>%
+             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
+    mutate(role = role, length = sen_length)
+}
+
 tokens %>%

### 2019-01-07 cogsci/position-info.Rmd
working on cogsci

@@
 # Uniform Information Density
+
+How do people convey information in speech and text?
+Is there a regular distribution to how information is conveyed across speech and text?
+Does speech, with its phonological properties, different from writing, with its semiotic properties, in how information is conveyed?
 
@@
 
-# Our Methods
+# Methods
 
@@
 The Providence corpus contains transcriptions of audio/video files which recorded interactions between children between 1 and 3 years old and their parents in the home.
-We divided each corpus by speaker (child and non-child) and each sentence length within those divisions and computed the unigram entropy measure for each position within those sentence lengths.
+We divided each corpus by speaker into child and non-child categories.
+We further divided the corpora by utterance length, so that all sentences of length $k$ (e.g.
+$6$) were grouped together.
+Finally, within each utterance length, we computed the unigram entropy measure for each position.
+
+The unigram entropy measure does not directly take into account the context of the utterance.
+
+Figure for the method.
+
 Plots for the entropy distributions are below.
@@
 
-Unlike a UID account
+We also ran this analysis on Spanish, German, Mandarin and Japanese corpora from CHILDES.
+The Spanish corpus we used was the XXX and the German corpus we used was the XXX.
+The Mandarin corpus we used was the XXX and the Japanese corpus we used was the XXX.
+For Mandarin, we used pinyin transliterations of the utterances in the corpus with demarcated word boundaries, and for Japanese we used romanji transliterations of words in the corpus.
+Japanese Hiragana, Katakana and Kanji writing systems, and the Chinese characters used for writing Mandarin do not normally demarcate word boundaries.
+JUSTIFICATION FOR WHY THE TRANSLITERATIONS ARE OKAY.
 
-We found a similar distribution for Spanish and German corpora from CHILDES. The Spanish corpus we used was the XXX and the German corpus we used was the XXX.
+# Results
+We found a distinct three-step distribution for English, Spanish and German CHILDES corpora, with a slight dip in the penultimate position of each sentence.
+The Mandarin and Japanese data, by comparison,
 
-# Formalities, Footnotes, and Floats
+# Analysis
 
-Use standard APA citation format. Citations within the text should
-include the author's last name and year. If the authors' names are
-included in the sentence, place only the year in parentheses, as in
-[-@NewellSimon1972a], but otherwise place the entire reference in
-parentheses with the authors and year separated by a comma
-[@NewellSimon1972a]. List multiple references alphabetically and
-separate them by semicolons [@ChalnickBillman1988a; @NewellSimon1972a]. 
-Use the et. al. construction only after listing all the authors to a
-publication in an earlier reference and for citations with four or
-more authors.
-
-For more information on citations in RMarkdown, see **[here](http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html#citations).**
-
-## Footnotes
-
-Indicate footnotes with a number\footnote{Sample of the first
-footnote.} in the text. Place the footnotes in 9 point type at the
-bottom of the page on which they appear. Precede the footnote with a
-horizontal rule.\footnote{Sample of the second footnote.} You can also use 
-markdown formatting to include footnotes using this syntax [^1].
-
-[^1]: Sample of a markdown footnote.
-
-## Figures
-
-All artwork must be very dark for purposes of reproduction and should
-not be hand drawn. Number figures sequentially, placing the figure
-number and caption, in 10 point, after the figure with one line space
-above the caption and one line space below it. If necessary, leave extra white space at
-the bottom of the page to avoid splitting the figure and figure
-caption. You may float figures to the top or bottom of a column, or
-set wide figures across both columns.
-
-## Two-column images
-
-You can read local images using png package for example and plot 
-it like a regular plot using grid.raster from the grid package. 
-With this method you have full control of the size of your image. **Note: Image must be in .png file format for the readPNG function to work.**
-
-You might want to display a wide figure across both columns.
-To do this, you change the `fig.env` chunk option to `figure*`.
-To align the image in the center of the page, set `fig.align` option to `center`.
-To format the width of your caption text, you set the `num.cols.cap` option to `2`.
-
-```{r 2-col-image, fig.env = "figure*", fig.pos = "h", fig.width=4, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns.
-And the caption text is limited to 0.8 of the width of the document."}
-img <- png::readPNG("figs/walrus.png")
-grid::grid.raster(img)
-```
-
-## One-column images
-
-Single column is the default option, but if you want set it explicitly, set `fig.env` to `figure`. Notice that the `num.cols` option for the caption width is set to `1`.
-
-```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
-img <- png::readPNG("figs/lab_logo_stanford.png")
-grid::grid.raster(img)
-```
-
-## R Plots
-
-You can use R chunks directly to plot graphs. And you can use latex floats in the
-fig.pos chunk option to have more control over the location of your plot on the page.
-For more information on latex placement specifiers see **[here](https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions)**
-
-```{r plot, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2, fig.height=2, fig.cap = "R plot" }
-x <- 0:100
-y <- 2 * (x + rnorm(length(x), sd = 3) + 3)
-
-ggplot2::ggplot(data = data.frame(x, y), 
-       aes(x = x, y = y)) + 
-  geom_point() + 
-  geom_smooth(method = "lm")
-```
-
-## Tables
-
-Number tables consecutively; place the table number and title (in
-10 point) above the table with one line space above the caption and
-one line space below it, as in Table 1. You may float
-tables to the top or bottom of a column, set wide tables across both
-columns.
-
-You can use the xtable function in the xtable package.
-
-```{r xtable, results="asis"}
-n <- 100
-x <- rnorm(n)
-y <- 2*x + rnorm(n)
-out <- lm(y ~ x)
-
-tab1 <- xtable::xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
-                      caption = "This table prints across one column.")
-
-print(tab1, type="latex", comment = F, table.placement = "H")
-```
+# Wikipedia Data
+We also wanted to apply this analysis to large-scale text data, to compare with the parent-child speech data results from using CHILDES.
+Using Giuseppe Attardi's Wikiextractor tool \footnote{https://github.com/attardi/wikiextractor}, we extracted the text corpora for $165$ languages from Wikipedia.
+Each language corpus was cleaned and limited to sentences between $6$ and $50$ words.
+Similar to the data from CHILDES, we divided each corpus by sentence length, and then computed the unigram entropy measure on each word position within each sentence length.
+HOW DID WE JOIN TOGETHER SENTENCE LENGTHS
 

### 2019-01-07 wiki_entropy/wiki_clustering.Rmd
working on cogsci

@@
 library(klaR)
-
 library(missMDA)
+library(clusteval)
+library(factoextra)
+library(cluster)
 

### 2019-01-09 wiki_entropy/wiki_clustering.Rmd
initial clusteval code
initial clusteval code - with edits

@@
 ```{r clustering}
+NUM_CLUSTERS <- 10
+
 ##read in googlesheet
@@
 ##do clustering
-wk_clust <- wkpd %>% 
+wk_clust_tree <- wkpd %>% 
               dist() %>%
@@
 
-wk_clust %>% plot(hang = -1)
+wk_clust_tree %>% plot(hang = -1)
+wk_clusters <- wk_clust_tree %>%
+              cutree(k = NUM_CLUSTERS)
 ```
@@
 imputed_wf <- as.data.frame(rv[[1]][[99]])
+
+wals_clust_full <- kmodes(imputed_wf, NUM_CLUSTERS, iter.max=10, weighted = FALSE)
+```
+
+```{r evaluate_clusters}
+wals_clusters <- wals_clust["cluster"] %>% unlist(use.names= FALSE)
 ```
@@
 
-NUM_CLUSTERS <- 4
-
 wf <- wals.feature(FEATURES)

### 2019-01-09 wiki_entropy/wiki_clustering.Rmd

@@
 library(cluster)
+library(widyr)
 
@@
               cutree(k = NUM_CLUSTERS)
+
+wk_clusters %>% 
+  t() %>%
+  as_data_frame() %>% 
+  t() %>%
+  as_data_frame(rownames = "language") %>%
+  rename(cluster = V1) %>%
+  arrange(cluster) %>% 
+  View()
 ```
@@
 wf <- wals.feature(FEATURES)
-wf.sub <- wf[c("1A", "2A", "3A", "4A", "language")]
-wf.sub <- wf.sub %>% 
+
+wf_df <- wf %>%
+  as_data_frame() %>%
   filter(language %in% LANGUAGES) %>%
-  unique()
+  group_by(language) %>%
+  slice(1) %>%
+  gather(feature, value, `1A`:`144A`) %>%
+  group_by(feature) %>%
+  mutate(value = factor(value))
 
@@
 ```
+
+```{r}
+
+wf_many_langs <- wf_df %>%
+  group_by(feature) %>%
+  summarise(n = sum(!is.na(value))) %>%
+  filter(n > 60)
+
+wf_sub_langs <- wf_df %>%
+  filter(feature %in% wf_many_langs$feature) %>%
+  group_by(language) %>%
+  summarise(keep = sum(is.na(value)) == 0) %>%
+  filter(keep)
+
+wf_sub <- wf_df %>%
+  filter(feature %in% wf_many_langs$feature,
+         language %in% wf_sub_langs$language) %>%
+  select(language, feature, value) %>%
+  group_by(feature) %>%
+  mutate(value = as.numeric(as.factor(value))) %>%
+  group_by(feature) %>% # for each pair of language, does this feature assin them the same value
+  nest() %>%
+  mutate(cors = map(data, my_pairwise_cor))
+  select(-data) %>%
+  unnest()
+
+```

### 2019-01-11 cogsci/position-info.Rmd
worked on cogsci

@@
 
-The *wrap-up effect* is a popular hypothesis within the eye-tracking community.
-The hypothesis states that, in written text, sentence-final words are processed more slowly on average then sentence-medial or sentence-initial words, due to readers integrating information from the entire sentence in order to form a final, coherent thought expressed by the sentence, among other reasons.
-With these two ideas, information production distribution in spoken and written communication, and information integration in reading written texts, we move on to discuss cross-linguistic differences in the distribution of information across sentences.
-
 The UID hypothesis
@@
 
-The unigram entropy measure does not directly take into account the context of the utterance.
+The unigram entropy measure works as
 
-Figure for the method.
+Figure for the method?
 
@@
 For Mandarin, we used pinyin transliterations of the utterances in the corpus with demarcated word boundaries, and for Japanese we used romanji transliterations of words in the corpus.
-Japanese Hiragana, Katakana and Kanji writing systems, and the Chinese characters used for writing Mandarin do not normally demarcate word boundaries.
-JUSTIFICATION FOR WHY THE TRANSLITERATIONS ARE OKAY.
+Japanese Hiragana, Katakana and Kanji writing systems, and the Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Japanese and Chinese writing including spaces between word boundaries can have a negative effect on reading times (Sainio et al, 2007; Bai et al, 2008).
 
@@
 We also wanted to apply this analysis to large-scale text data, to compare with the parent-child speech data results from using CHILDES.
-Using Giuseppe Attardi's Wikiextractor tool \footnote{https://github.com/attardi/wikiextractor}, we extracted the text corpora for $165$ languages from Wikipedia.
+Using Giuseppe Attardi's Wikiextractor tool \footnote{https://github.com/attardi/wikiextractor}, we extracted the text corpora for $165$ languages from Wikipedia by downloading the data dump of Wikipedia.
 Each language corpus was cleaned and limited to sentences between $6$ and $50$ words.
@@
 HOW DID WE JOIN TOGETHER SENTENCE LENGTHS
+
+## Clustering
+
+# Conclustions and future research
+The *wrap-up effect* is a popular hypothesis within the eye-tracking community.
+The hypothesis states that, in written text, sentence-final words are processed more slowly on average then sentence-medial or sentence-initial words, due to readers integrating information from the entire sentence in order to form a final, coherent thought expressed by the sentence, among other reasons.
+With these two ideas, information production distribution in spoken and written communication, and information integration in reading written texts, we move on to discuss cross-linguistic differences in the distribution of information across sentences.
 

### 2019-01-11 cogsci/position-info.tex
worked on cogsci

@@
 % cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
+\usepackage{apacite}
+
+% KM added 1/4/18 to allow control of blind submission
 
@@
 
+How do people convey information in speech and text? Is there a regular
+distribution to how information is conveyed across speech and text? Does
+speech, with its phonological properties, different from writing, with
+its semiotic properties, in how information is conveyed?
+
 A major line of research in the current century has asked how speakers
@@
 
-The \emph{wrap-up effect} is a popular hypothesis within the
-eye-tracking community. The hypothesis states that, in written text,
-sentence-final words are processed more slowly on average then
-sentence-medial or sentence-initial words, due to readers integrating
-information from the entire sentence in order to form a final, coherent
-thought expressed by the sentence, among other reasons. With these two
-ideas, information production distribution in spoken and written
-communication, and information integration in reading written texts, we
-move on to discuss cross-linguistic differences in the distribution of
-information across sentences.
-
 The UID hypothesis
@@
 
-\begin{CodeChunk}
-\begin{figure*}[h]
-
-{\centering \includegraphics{figs/yunigram-1}
-
-}
-
-\caption[English entropy distribution for three sentence lengths from BNC]{English entropy distribution for three sentence lengths from BNC}\label{fig:yunigram}
-\end{figure*}
-\end{CodeChunk}
-
-\section{Our Methods}\label{our-methods}
+\section{Methods}\label{methods}
 
@@
 between children between 1 and 3 years old and their parents in the
-home.
+home. We divided each corpus by speaker into child and non-child
+categories. We further divided the corpora by utterance length, so that
+all sentences of length \(k\) (e.g. \(6\)) were grouped together.
+Finally, within each utterance length, we computed the unigram entropy
+measure for each position.
+
+The unigram entropy measure works as
+
+Figure for the method?
+
+Plots for the entropy distributions are below. The adult and child
+unigram entropy distributions track one another extremely well, with the
+child distribution for each sentence length having relatively lower
+entropy.
 
@@
 
-We found a similar distribution for Spanish and German corpora from
-CHILDES. The Spanish corpus we used was the XXX and the German corpus we
-used was the XXX.
+We also ran this analysis on Spanish, German, Mandarin and Japanese
+corpora from CHILDES. The Spanish corpus we used was the XXX and the
+German corpus we used was the XXX. The Mandarin corpus we used was the
+XXX and the Japanese corpus we used was the XXX. For Mandarin, we used
+pinyin transliterations of the utterances in the corpus with demarcated
+word boundaries, and for Japanese we used romanji transliterations of
+words in the corpus. Japanese Hiragana, Katakana and Kanji writing
+systems, and the Chinese characters used for writing Mandarin do not
+normally demarcate word boundaries by spacing words apart, and for
+normal Japanese and Chinese writing including spaces between word
+boundaries can have a negative effect on reading times (Sainio et al,
+2007; Bai et al, 2008).
 
-\section{Formalities, Footnotes, and
-Floats}\label{formalities-footnotes-and-floats}
+\section{Results}\label{results}
 
-Use standard APA citation format. Citations within the text should
-include the author's last name and year. If the authors' names are
-included in the sentence, place only the year in parentheses, as in
-(1972), but otherwise place the entire reference in parentheses with the
-authors and year separated by a comma (Newell \& Simon, 1972). List
-multiple references alphabetically and separate them by semicolons
-(Chalnick \& Billman, 1988; Newell \& Simon, 1972). Use the et. al.
-construction only after listing all the authors to a publication in an
-earlier reference and for citations with four or more authors.
+We found a distinct three-step distribution for English, Spanish and
+German CHILDES corpora, with a slight dip in the penultimate position of
+each sentence. The Mandarin and Japanese data, by comparison,
 
-For more information on citations in RMarkdown, see
-\textbf{\href{http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html\#citations}{here}.}
+\section{Analysis}\label{analysis}
 
-\subsection{Footnotes}\label{footnotes}
+\section{Wikipedia Data}\label{wikipedia-data}
 
-Indicate footnotes with a number\footnote{Sample of the first
-footnote.} in the text. Place the footnotes in 9 point type at the
-bottom of the page on which they appear. Precede the footnote with a
-horizontal rule.\footnote{Sample of the second footnote.} You can also
-use markdown formatting to include footnotes using this
-syntax.\footnote{Sample of a markdown footnote.}
+We also wanted to apply this analysis to large-scale text data, to
+compare with the parent-child speech data results from using CHILDES.
+Using Giuseppe Attardi's Wikiextractor tool
+\footnote{https://github.com/attardi/wikiextractor}, we extracted the
+text corpora for \(165\) languages from Wikipedia by downloading the
+data dump of Wikipedia. Each language corpus was cleaned and limited to
+sentences between \(6\) and \(50\) words. Similar to the data from
+CHILDES, we divided each corpus by sentence length, and then computed
+the unigram entropy measure on each word position within each sentence
+length. HOW DID WE JOIN TOGETHER SENTENCE LENGTHS
 
-\subsection{Figures}\label{figures}
+\subsection{Clustering}\label{clustering}
 
-All artwork must be very dark for purposes of reproduction and should
-not be hand drawn. Number figures sequentially, placing the figure
-number and caption, in 10 point, after the figure with one line space
-above the caption and one line space below it. If necessary, leave extra
-white space at the bottom of the page to avoid splitting the figure and
-figure caption. You may float figures to the top or bottom of a column,
-or set wide figures across both columns.
+\section{Conclustions and future
+research}\label{conclustions-and-future-research}
 
-\subsection{Two-column images}\label{two-column-images}
-
-You can read local images using png package for example and plot it like
-a regular plot using grid.raster from the grid package. With this method
-you have full control of the size of your image. \textbf{Note: Image
-must be in .png file format for the readPNG function to work.}
-
-You might want to display a wide figure across both columns. To do this,
-you change the \texttt{fig.env} chunk option to \texttt{figure*}. To
-align the image in the center of the page, set \texttt{fig.align} option
-to \texttt{center}. To format the width of your caption text, you set
-the \texttt{num.cols.cap} option to \texttt{2}.
-
-\begin{CodeChunk}
-\begin{figure*}[h]
-
-{\centering \includegraphics{figs/2-col-image-1}
-
-}
-
-\caption[This image spans both columns]{This image spans both columns. And the caption text is limited to 0.8 of the width of the document.}\label{fig:2-col-image}
-\end{figure*}
-\end{CodeChunk}
-
-\subsection{One-column images}\label{one-column-images}
-
-Single column is the default option, but if you want set it explicitly,
-set \texttt{fig.env} to \texttt{figure}. Notice that the
-\texttt{num.cols} option for the caption width is set to \texttt{1}.
-
-\begin{CodeChunk}
-\begin{figure}[H]
-
-{\centering \includegraphics{figs/image-1}
-
-}
-
-\caption[One column image]{One column image.}\label{fig:image}
-\end{figure}
-\end{CodeChunk}
-
-\subsection{R Plots}\label{r-plots}
-
-You can use R chunks directly to plot graphs. And you can use latex
-floats in the fig.pos chunk option to have more control over the
-location of your plot on the page. For more information on latex
-placement specifiers see
-\textbf{\href{https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions}{here}}
-
-\begin{CodeChunk}
-\begin{figure}[H]
-
-{\centering \includegraphics{figs/plot-1}
-
-}
-
-\caption[R plot]{R plot}\label{fig:plot}
-\end{figure}
-\end{CodeChunk}
-
-\subsection{Tables}\label{tables}
-
-Number tables consecutively; place the table number and title (in 10
-point) above the table with one line space above the caption and one
-line space below it, as in Table 1. You may float tables to the top or
-bottom of a column, set wide tables across both columns.
-
-You can use the xtable function in the xtable package.
-
-\begin{table}[H]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
-  \hline
-(Intercept) & -0.10 & 0.11 & -0.9 & 0.37 \\ 
-  x & 1.97 & 0.12 & 16.6 & 0.00 \\ 
-   \hline
-\end{tabular}
-\caption{This table prints across one column.} 
-\end{table}
+The \emph{wrap-up effect} is a popular hypothesis within the
+eye-tracking community. The hypothesis states that, in written text,
+sentence-final words are processed more slowly on average then
+sentence-medial or sentence-initial words, due to readers integrating
+information from the entire sentence in order to form a final, coherent
+thought expressed by the sentence, among other reasons. With these two
+ideas, information production distribution in spoken and written
+communication, and information integration in reading written texts, we
+move on to discuss cross-linguistic differences in the distribution of
+information across sentences.
 
@@
 
-\hypertarget{refs}{}
-\hypertarget{ref-ChalnickBillman1988a}{}
-Chalnick, A., \& Billman, D. (1988). Unsupervised learning of
-correlational structure. In \emph{Proceedings of the tenth annual
-conference of the cognitive science society} (pp. 510--516). Hillsdale,
-NJ: Lawrence Erlbaum Associates.
-
-\hypertarget{ref-NewellSimon1972a}{}
-Newell, A., \& Simon, H. A. (1972). \emph{Human problem solving}.
-Englewood Cliffs, NJ: Prentice-Hall.
+\bibliographystyle{apacite}
 

### 2019-01-11 cogsci/position-info.Rmd
cogsci first draft

@@
 abstract: 
-    "We provide evidence against the Uniform Information Density (Jaeger, 2010) 
-    and wrap-up effect hypotheses."
+    "A major line of research in the cognitive sciences has asked how speakers and writers distribute information in the linguistic material they produce.
+We provide evidence against the Uniform Information Density hypothesis (Jaeger, 2010) which proposes that information is transmitted at a constant rate close to channel capacity in human communication.
+We instead argue that the information structure and transmission rate in human communication varies predictably from language to language based on phonological and syntactic features.
+We draw evidence from written and spoken corpora, including corpora of parent and child communication at various ages and large-scale online corpora from Wikipedia in a typologically diverse collection of languages."
 
 keywords:
-    "Entropy; self-paced reading; information theory; language processing"
+    "Entropy; information; information theory; communication"
 
@@
 
-# Uniform Information Density
+# Background
 
@@
 
-The UID hypothesis
-
-## UID challenges
-
 More recent work has stood out in contrast to the traditional UID perspective.
-Zhan and Levy (2018) study Mandarin Chinese classifier use, and find that the use of specific classifier, versus general classifiers, appears more often in cases where the production of the corresponding noun is more difficult than when the production is easier, as would be predicted by UID.
+Zhan and Levy (2018) study Mandarin Chinese classifier use, and find that the use of specific classifiers, versus general classifiers, appears more often in cases where the production of the corresponding noun is more difficult than when the production is easier, as would be predicted by UID.
 The UID perspective is challenged in Yu et al.
-(2016), by performing an analysis of entropy by position in the text portion of the British National Corpus.
-They use the following formula for each word position X of sentences of fixed length k from the corpus, where each i is a word occurring in position X and pi is the number of times word i occurs in position X divided by the number of total words that occur in position X i.e.
-the number of sentences of length k.
-
-$$H(X) = \sum\limits_w p(w)\log\big(p(w)\big)$$
-Yu et al.
-(2016) refer to their distribution for English as a ‘three-step distribution’: relatively low entropy at the beginning of a sentence, then a jump, then flat entropy in the middle, a dip before the final position and a jump with the final word.
-View the figure below for a visual demonstration.
+(2016), by performing an analysis of entropy by position within sentence in the text portion of the British National Corpus.
+This work serves as the basis for ours.
 
@@
 
-We used the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations.
+To explore information transmission rate in speech, we used the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations.
 We first used the Brown and Providence English corpora from CHILDES.
-The Brown corpus contains conversations between three young children (over 1.5 years old and under 6 years old) and their families in the home.
-The Providence corpus contains transcriptions of audio/video files which recorded interactions between children between 1 and 3 years old and their parents in the home.
+The Brown corpus contains individual recordings of conversations between three children between 1.5 and 6-years-old and their families in their homes.
+The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home.
 We divided each corpus by speaker into child and non-child categories.
@@
 Finally, within each utterance length, we computed the unigram entropy measure for each position.
+We follow Yu et al.
+(2016) in using the following formula for each word position X of sentences of fixed length k from the corpus, where each $w$ is a word occurring in position $X$ and $p(w)$ is the number of times word $w$ occurs in position X divided by the number of total words that occur in position $X$ i.e.
+the number of sentences of length $k$.
 
-The unigram entropy measure works as
+$$H(X) = \sum\limits_w p(w)\log\big(p(w)\big)$$
 
-Figure for the method?
+The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances in the English CHILDES corpora.
+The entropy curves capture individual variation across positions in utterances of the same length.
+This allows us to directly observe and judge the amount of variation in words that appear in an individual position of a sentence.
+For speech data, which for the corpora in the CHILDES database consists of short and often disconnected utterances across hours of recordings, the unigram entropy measure is unaffected by context or lack thereof in utterances by the adults and children in the corpora.
+We can directly compare any two positions within utterances to determine the amount of uncertainty, and therefore information, on average contained by words within that position of utterances.
+We are applying the same approach as in Genzel and Charniak (2003), but within sentences instead of across sentences.
 
 Plots for the entropy distributions are below.
-The adult and child unigram entropy distributions track one another extremely well, with the child distribution for each sentence length having relatively lower entropy.
 
-```{r brown_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Brown corpus entropy"}
+```{r brown_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Brown corpus entropy"}
 img <- png::readPNG("figs/positional_entropy_brown.png")
@@
 
-```{r providence_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Providence corpus unigram entropy"}
+```{r providence_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Providence corpus unigram entropy"}
 img <- png::readPNG("figs/Prov_corpus_result.png")
@@
 We also ran this analysis on Spanish, German, Mandarin and Japanese corpora from CHILDES.
-The Spanish corpus we used was the XXX and the German corpus we used was the XXX.
-The Mandarin corpus we used was the XXX and the Japanese corpus we used was the XXX.
-For Mandarin, we used pinyin transliterations of the utterances in the corpus with demarcated word boundaries, and for Japanese we used romanji transliterations of words in the corpus.
+We used the Shiro corpus for Spanish (Shiro, 2000), which contains prompted narratives individually collected from over a hundred Venezualan schoolchildren, half from high SES backgrounds and half from low SES backgrounds.
+The German corpus we used was the Wagner corpus of spontaneous parent-child speech in the home, collected in single sessions for children between 1 and 14 years of age (Wagner, 1985).
+We used the Zhou dinner corpus for Mandarin Chinese (Li & Zhou, 2015), which contains dinner conversations between 5 to 6-year-old children and their parents collected in Shanghai.
+We used the Okayama corpus for Japanese (Okayama, 1973), which consists of conversations between 2 to 4-year-old children and their mothers in single sessions.
+
+For each corpus, we accessed transcripts of the corpus provided through the TalkBank system and computed over Roman alphabet transcriptions or transliterations of the original transcriptions.
+For Mandarin, we used pinyin transliterations of the utterances in the corpus with demarcated word boundaries, and for Japanese we used romanji (Roman alphabet) transliterations of words in the corpus.
 Japanese Hiragana, Katakana and Kanji writing systems, and the Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Japanese and Chinese writing including spaces between word boundaries can have a negative effect on reading times (Sainio et al, 2007; Bai et al, 2008).
 
-# Results
+```{r shiro_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Shiro corpus entropy"}
+img <- png::readPNG("figs/German_unigram_entropies_with_stemming.png")
+grid::grid.raster(img)
+```
+
+```{r wagner_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Wagner corpus unigram entropy"}
+img <- png::readPNG("figs/Spanish_unigram_entropies_with_stemming.png")
+grid::grid.raster(img)
+```
+
+```{r zhou_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Zhou Dinner corpus entropy"}
+img <- png::readPNG("figs/Mandarin_unigram_entropies.png")
+grid::grid.raster(img)
+```
+
+```{r okayama_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Okayama corpus unigram entropy"}
+img <- png::readPNG("figs/Japanese_okayama_entropy.png")
+grid::grid.raster(img)
+```
+
+# Results & Analysis
+The Unigram Information Density account would predict that children have a lower channel capacity for communication and would instead communicate with a less unigram entropy curve.
+The adult and child entropy curves track one another almost identically.
+This not only indicates a robustness in the unigram entropy curve across speakers, but also across ages and addressees.
+We also observed a robustness across corpora for the same languages, and a robustness across utterance lengths within the same corpus's entropy curve.
+This shows that the concept of an "entropy curve" for a specific language is well-founded when considering speech data.
+
 We found a distinct three-step distribution for English, Spanish and German CHILDES corpora, with a slight dip in the penultimate position of each sentence.
-The Mandarin and Japanese data, by comparison,
+We attribute the penultimate dip in these unigram entropy curves to the fact that most of the utterances in the CHILDES English, Spanish and German corpora we examined had a determiner such as "the" or "a" in the second-to-last position of utterances.
+The beginnings of utterances in the English, German and Spanish CHILDES data were usually pronouns or grammatical subjects, while the final words were grammatical objects and had a great deal of variation in the exact word that appeared in the utterance-final position.
+The final position of utterances in child-directed speech is known to be important, dating back to Aslin (1993).
+The Mandarin and Japanese data, by comparison, have noticeably lower positional entropy values in utterance-final positions than in utterance-penultimate positions.
+The Japanese entropy curve begins high and then drops, unlike all of the other entropy curves.
 
-# Analysis
+```{r MI_wagner, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Mutual Information measures on utterances of length 6 in the Wagner corpus"}
+img <- png::readPNG("figs/wagner_mi_6.png")
+grid::grid.raster(img)
+```
 
-# Wikipedia Data
+```{r MI_okayama, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Mutual Information measures on utterances of length 6 in the Okayama corpus"}
+img <- png::readPNG("figs/okayama_mi_6.png")
+grid::grid.raster(img)
+```
+
+The mutual information statistic is a symmetric measure of the information gained about one random variable by learning about another, a bigram measure and indicator of the effect than the context of a word can have on its unigram entropy value.
+As seen in the images below, the final word in utterances of length $6$ in German has a significant mutual information value with all other words in the utterance on average.
+This means that the words appearing in positions $1$ through $5$ all play a role in determining the word that appears in position $6$ and vice versa.
+By contrast, the words that appear in positions $1$ through $4$ have little influence in determining the word that appears in position $5$, the penultimate word that is often a determiner or function word, and instead the final word has the greatest influence in determining the penultimate word.
+For Japanese, the final word in an utterance has little to no influence on determining the rest of the words and vice versa, whereas the other words all play in role in determining what will go in each other's positions.
+
+# Large-scale text data
 We also wanted to apply this analysis to large-scale text data, to compare with the parent-child speech data results from using CHILDES.
-Using Giuseppe Attardi's Wikiextractor tool \footnote{https://github.com/attardi/wikiextractor}, we extracted the text corpora for $165$ languages from Wikipedia by downloading the data dump of Wikipedia.
+Using Giuseppe Attardi's Wikiextractor tool \footnote{https://github.com/attardi/wikiextractor}, we extracted the text corpora for $165$ languages from Wikipedia by downloading a stored collection of Wikipedia entries in each langauge and randomly selecting several thousand articles from each Wikipedia language.
 Each language corpus was cleaned and limited to sentences between $6$ and $50$ words.
 Similar to the data from CHILDES, we divided each corpus by sentence length, and then computed the unigram entropy measure on each word position within each sentence length.
-HOW DID WE JOIN TOGETHER SENTENCE LENGTHS
+
+To classify the unigram entropy curves of the different distributions, we computed three slope treatments of each curve.
+In the *absolute* treatment, with sentence length denoted as $n$, we computed the slope between positions $1$ and $2$, positions $2$ and $3$, positions $3$ and $n-2$, positions $n-2$ and $n-1$ and positions $n-1$ and $n$.
+For the short utterances appearing the CHILDES speech corpora we examined, these appeared to be important junctions in the distributions, with a seeming plateau in the middle of the unigram entropy curve for each of the language corpora we examined in CHILDES.
+
+However, because the portion of sentences of length greater than $10$ in the Wikipedia corpora were significantly larger than the CHILDES corpora, then we also computed relative slope treatments.
+In the *relative 5* treatment, we computed the slopes between $0\%$ and $20\%$, $20\%$ and $40\%$, $40\%$ and $60\%$, $60\%$ and $80\%$ and $80\%$ and $100\%$ of the relative word positions in each sentence.
+When one of these percentages was not a whole number, then the closest whole number position was used instead for slope calculation.
+In the *relative 10* treatment, we computed the slopes between every $10\%$ of the relative word positions in each sentence.
+Each comparative slope within each treatment was averaged together between different sentence lengths, for example in the *relative 5* treatment then all of the $0\%$ to $20\%$ slopes were averaged together.
+This created three treatments for the entropy curve for each langauge in the Wikipedia database.
+
+For the German and Japanese entropy curves, we observed the same shape for the text Wikipedia corpora entropy curves and the speech CHILDES corpora entropy curves, supporting the argument for robustness across text and speech.
 
 ## Clustering
+In the *absolute* and *relative 5* treatments, each language is cast as a point in $5$-dimensional space.
+In the *relative 10* treatment, each language is cast as a point in $10$-dimensional space.
+This allows for direct comparison using cosine similarity and moreover for unsupervised clustering analysis to compare the results of the Wikipedia entropy analysis to known typological features within the langauges in the Wikipedia dataset.
+Using the R hclust package, we performed a hierarchical clustering of the Wikipedia slope data within each treatment.
+A subset of the results at a glance for the *absolute* treatment are below.
 
-# Conclustions and future research
+```{r absolute_clusters, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Absolute clustering of Wikipedia corpora"}
+img <- png::readPNG("figs/absolute_clusters.png")
+grid::grid.raster(img)
+```
+
+## Language Features
+
+To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we looked at the list of $144$ linguistic features in World Atlas of Language Structures (Dryer & Haspelmath, 2013).
+We limited the languages in the WALS database to only those in our Wikipedia dataset and performed missing-value imputation to obtain the features not coded in WALS for the languages from Wikipedia.
+We then performed k-modes clustering and compared the results with the hierarchical clustering from Wikipedia.
+[Still working on this]
+
+We also checked the effects of individual features on the embeddings of languages in the different treatments. [Still working on this]
+
+# Conclusions 
+In this paper we have explored the differences in rate of information transmission among spoken child-adult conversational corpora in several languages and text corpora pulled from Wikipedia for well over a hundred other corpora.
+We have seen that the entropy curve, a representation of information at each position of an utterance without regard to context, varies from language to language and is predicted by a subset of syntactic, phonological and morphological features in a language.
+We have provided evidence against the Uniform Information Density hypothesis and provided an alternate explanation for the rate at which information is encoded and transmitted in human communication, namely one that is robust across corpora, language dependent and arises from the features of the language.
+
+A large volume of research has indicated the effects of surprisal on fixation duration during eye-tracking studies: higher surprisal of a word corresponds to higher fixation duration.
 The *wrap-up effect* is a popular hypothesis within the eye-tracking community.
 The hypothesis states that, in written text, sentence-final words are processed more slowly on average then sentence-medial or sentence-initial words, due to readers integrating information from the entire sentence in order to form a final, coherent thought expressed by the sentence, among other reasons.
-With these two ideas, information production distribution in spoken and written communication, and information integration in reading written texts, we move on to discuss cross-linguistic differences in the distribution of information across sentences.
-
-# Acknowledgements
-
-Place acknowledgments (including funding information) in a section at
-the end of the paper.
-
-# References
+The wrap-up effect is drawn from evidence in English, German and Dutch, all of which are languages with a large final increase in their entropy curve from our study.
+We aim to see in the future if as a consquence of our work in this paper, the wrap-up effect falls out from the final jump in the entropy curves in some languages.
 

### 2019-01-11 cogsci/position-info.tex
cogsci first draft

@@
 \begin{abstract}
-We provide evidence against the Uniform Information Density (Jaeger,
-2010) and wrap-up effect hypotheses.
+A major line of research in the cognitive sciences has asked how
+speakers and writers distribute information in the linguistic material
+they produce. We provide evidence against the Uniform Information
+Density hypothesis (Jaeger, 2010) which proposes that information is
+transmitted at a constant rate close to channel capacity in human
+communication. We instead argue that the information structure and
+transmission rate in human communication varies predictably from
+language to language based on phonological and syntactic features. We
+draw evidence from written and spoken corpora, including corpora of
+parent and child communication at various ages and large-scale online
+corpora from Wikipedia in a typologically diverse collection of
+languages.
 
 \textbf{Keywords:}
-Entropy; self-paced reading; information theory; language processing
+Entropy; information; information theory; communication
 \end{abstract}
 
-\section{Uniform Information Density}\label{uniform-information-density}
+\section{Background}\label{background}
 
@@
 
-The UID hypothesis
-
-\subsection{UID challenges}\label{uid-challenges}
-
 More recent work has stood out in contrast to the traditional UID
 perspective. Zhan and Levy (2018) study Mandarin Chinese classifier use,
-and find that the use of specific classifier, versus general
+and find that the use of specific classifiers, versus general
 classifiers, appears more often in cases where the production of the
@@
 as would be predicted by UID. The UID perspective is challenged in Yu et
-al. (2016), by performing an analysis of entropy by position in the text
-portion of the British National Corpus. They use the following formula
-for each word position X of sentences of fixed length k from the corpus,
-where each i is a word occurring in position X and pi is the number of
-times word i occurs in position X divided by the number of total words
-that occur in position X i.e.~the number of sentences of length k.
-
-\[H(X) = \sum\limits_w p(w)\log\big(p(w)\big)\] Yu et al. (2016) refer
-to their distribution for English as a `three-step distribution':
-relatively low entropy at the beginning of a sentence, then a jump, then
-flat entropy in the middle, a dip before the final position and a jump
-with the final word. View the figure below for a visual demonstration.
+al. (2016), by performing an analysis of entropy by position within
+sentence in the text portion of the British National Corpus. This work
+serves as the basis for ours.
 
@@
 
-We used the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora
-database of spoken adult-child conversations. We first used the Brown
-and Providence English corpora from CHILDES. The Brown corpus contains
-conversations between three young children (over 1.5 years old and under
-6 years old) and their families in the home. The Providence corpus
-contains transcriptions of audio/video files which recorded interactions
-between children between 1 and 3 years old and their parents in the
-home. We divided each corpus by speaker into child and non-child
-categories. We further divided the corpora by utterance length, so that
-all sentences of length \(k\) (e.g. \(6\)) were grouped together.
-Finally, within each utterance length, we computed the unigram entropy
-measure for each position.
+To explore information transmission rate in speech, we used the CHILDES
+TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken
+adult-child conversations. We first used the Brown and Providence
+English corpora from CHILDES. The Brown corpus contains individual
+recordings of conversations between three children between 1.5 and
+6-years-old and their families in their homes. The Providence corpus
+recorded interactions between children between 1 and 3 years old and
+their parents in the home. We divided each corpus by speaker into child
+and non-child categories. We further divided the corpora by utterance
+length, so that all sentences of length \(k\) (e.g. \(6\)) were grouped
+together. Finally, within each utterance length, we computed the unigram
+entropy measure for each position. We follow Yu et al. (2016) in using
+the following formula for each word position X of sentences of fixed
+length k from the corpus, where each \(w\) is a word occurring in
+position \(X\) and \(p(w)\) is the number of times word \(w\) occurs in
+position X divided by the number of total words that occur in position
+\(X\) i.e.~the number of sentences of length \(k\).
 
-The unigram entropy measure works as
+\[H(X) = \sum\limits_w p(w)\log\big(p(w)\big)\]
 
-Figure for the method?
+The result of this method can be plotted for each utterance length as an
+\emph{entropy curve}, which can be visually compared across utterance
+length to observe the how the unigram entropy changes across absolute
+positions in each of the utterances in the English CHILDES corpora. The
+entropy curves capture individual variation across positions in
+utterances of the same length. This allows us to directly observe and
+judge the amount of variation in words that appear in an individual
+position of a sentence. For speech data, which for the corpora in the
+CHILDES database consists of short and often disconnected utterances
+across hours of recordings, the unigram entropy measure is unaffected by
+context or lack thereof in utterances by the adults and children in the
+corpora. We can directly compare any two positions within utterances to
+determine the amount of uncertainty, and therefore information, on
+average contained by words within that position of utterances. We are
+applying the same approach as in Genzel and Charniak (2003), but within
+sentences instead of across sentences.
 
-Plots for the entropy distributions are below. The adult and child
-unigram entropy distributions track one another extremely well, with the
-child distribution for each sentence length having relatively lower
-entropy.
+Plots for the entropy distributions are below.
 
@@
 We also ran this analysis on Spanish, German, Mandarin and Japanese
-corpora from CHILDES. The Spanish corpus we used was the XXX and the
-German corpus we used was the XXX. The Mandarin corpus we used was the
-XXX and the Japanese corpus we used was the XXX. For Mandarin, we used
+corpora from CHILDES. We used the Shiro corpus for Spanish (Shiro,
+2000), which contains prompted narratives individually collected from
+over a hundred Venezualan schoolchildren, half from high SES backgrounds
+and half from low SES backgrounds. The German corpus we used was the
+Wagner corpus of spontaneous parent-child speech in the home, collected
+in single sessions for children between 1 and 14 years of age (Wagner,
+1985). We used the Zhou dinner corpus for Mandarin Chinese (Li \& Zhou,
+2015), which contains dinner conversations between 5 to 6-year-old
+children and their parents collected in Shanghai. We used the Okayama
+corpus for Japanese (Okayama, 1973), which consists of conversations
+between 2 to 4-year-old children and their mothers in single sessions.
+
+For each corpus, we accessed transcripts of the corpus provided through
+the TalkBank system and computed over Roman alphabet transcriptions or
+transliterations of the original transcriptions. For Mandarin, we used
 pinyin transliterations of the utterances in the corpus with demarcated
-word boundaries, and for Japanese we used romanji transliterations of
-words in the corpus. Japanese Hiragana, Katakana and Kanji writing
-systems, and the Chinese characters used for writing Mandarin do not
-normally demarcate word boundaries by spacing words apart, and for
-normal Japanese and Chinese writing including spaces between word
-boundaries can have a negative effect on reading times (Sainio et al,
-2007; Bai et al, 2008).
+word boundaries, and for Japanese we used romanji (Roman alphabet)
+transliterations of words in the corpus. Japanese Hiragana, Katakana and
+Kanji writing systems, and the Chinese characters used for writing
+Mandarin do not normally demarcate word boundaries by spacing words
+apart, and for normal Japanese and Chinese writing including spaces
+between word boundaries can have a negative effect on reading times
+(Sainio et al, 2007; Bai et al, 2008).
 
-\section{Results}\label{results}
+\begin{CodeChunk}
+\begin{figure*}[h]
+
+{\centering \includegraphics{figs/shiro_PE-1}
+
+}
+
+\caption[Shiro corpus entropy]{Shiro corpus entropy}\label{fig:shiro_PE}
+\end{figure*}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{figure*}[h]
+
+{\centering \includegraphics{figs/wagner_PE-1}
+
+}
+
+\caption[Wagner corpus unigram entropy]{Wagner corpus unigram entropy}\label{fig:wagner_PE}
+\end{figure*}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{figure*}[h]
+
+{\centering \includegraphics{figs/zhou_PE-1}
+
+}
+
+\caption[Zhou Dinner corpus entropy]{Zhou Dinner corpus entropy}\label{fig:zhou_PE}
+\end{figure*}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{figure*}[h]
+
+{\centering \includegraphics{figs/okayama_PE-1}
+
+}
+
+\caption[Okayama corpus unigram entropy]{Okayama corpus unigram entropy}\label{fig:okayama_PE}
+\end{figure*}
+\end{CodeChunk}
+
+\section{Results \& Analysis}\label{results-analysis}
+
+The Unigram Information Density account would predict that children have
+a lower channel capacity for communication and would instead communicate
+with a less unigram entropy curve. The adult and child entropy curves
+track one another almost identically. This not only indicates a
+robustness in the unigram entropy curve across speakers, but also across
+ages and addressees. We also observed a robustness across corpora for
+the same languages, and a robustness across utterance lengths within the
+same corpus's entropy curve. This shows that the concept of an ``entropy
+curve'' for a specific language is well-founded when considering speech
+data.
 
@@
 German CHILDES corpora, with a slight dip in the penultimate position of
-each sentence. The Mandarin and Japanese data, by comparison,
+each sentence. We attribute the penultimate dip in these unigram entropy
+curves to the fact that most of the utterances in the CHILDES English,
+Spanish and German corpora we examined had a determiner such as ``the''
+or ``a'' in the second-to-last position of utterances. The beginnings of
+utterances in the English, German and Spanish CHILDES data were usually
+pronouns or grammatical subjects, while the final words were grammatical
+objects and had a great deal of variation in the exact word that
+appeared in the utterance-final position. The final position of
+utterances in child-directed speech is known to be important, dating
+back to Aslin (1993). The Mandarin and Japanese data, by comparison,
+have noticeably lower positional entropy values in utterance-final
+positions than in utterance-penultimate positions. The Japanese entropy
+curve begins high and then drops, unlike all of the other entropy
+curves.
 
-\section{Analysis}\label{analysis}
+\begin{CodeChunk}
+\begin{figure*}[h]
 
-\section{Wikipedia Data}\label{wikipedia-data}
+{\centering \includegraphics{figs/MI_wagner-1}
+
+}
+
+\caption[Mutual Information measures on utterances of length 6 in the Wagner corpus]{Mutual Information measures on utterances of length 6 in the Wagner corpus}\label{fig:MI_wagner}
+\end{figure*}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{figure*}[h]
+
+{\centering \includegraphics{figs/MI_okayama-1}
+
+}
+
+\caption[Mutual Information measures on utterances of length 6 in the Okayama corpus]{Mutual Information measures on utterances of length 6 in the Okayama corpus}\label{fig:MI_okayama}
+\end{figure*}
+\end{CodeChunk}
+
+The mutual information statistic is a symmetric measure of the
+information gained about one random variable by learning about another,
+a bigram measure and indicator of the effect than the context of a word
+can have on its unigram entropy value. As seen in the images below, the
+final word in utterances of length \(6\) in German has a significant
+mutual information value with all other words in the utterance on
+average. This means that the words appearing in positions \(1\) through
+\(5\) all play a role in determining the word that appears in position
+\(6\) and vice versa. By contrast, the words that appear in positions
+\(1\) through \(4\) have little influence in determining the word that
+appears in position \(5\), the penultimate word that is often a
+determiner or function word, and instead the final word has the greatest
+influence in determining the penultimate word. For Japanese, the final
+word in an utterance has little to no influence on determining the rest
+of the words and vice versa, whereas the other words all play in role in
+determining what will go in each other's positions.
+
+\section{Large-scale text data}\label{large-scale-text-data}
 
@@
 \footnote{https://github.com/attardi/wikiextractor}, we extracted the
-text corpora for \(165\) languages from Wikipedia by downloading the
-data dump of Wikipedia. Each language corpus was cleaned and limited to
-sentences between \(6\) and \(50\) words. Similar to the data from
-CHILDES, we divided each corpus by sentence length, and then computed
-the unigram entropy measure on each word position within each sentence
-length. HOW DID WE JOIN TOGETHER SENTENCE LENGTHS
+text corpora for \(165\) languages from Wikipedia by downloading a
+stored collection of Wikipedia entries in each langauge and randomly
+selecting several thousand articles from each Wikipedia language. Each
+language corpus was cleaned and limited to sentences between \(6\) and
+\(50\) words. Similar to the data from CHILDES, we divided each corpus
+by sentence length, and then computed the unigram entropy measure on
+each word position within each sentence length.
+
+To classify the unigram entropy curves of the different distributions,
+we computed three slope treatments of each curve. In the \emph{absolute}
+treatment, with sentence length denoted as \(n\), we computed the slope
+between positions \(1\) and \(2\), positions \(2\) and \(3\), positions
+\(3\) and \(n-2\), positions \(n-2\) and \(n-1\) and positions \(n-1\)
+and \(n\). For the short utterances appearing the CHILDES speech corpora
+we examined, these appeared to be important junctions in the
+distributions, with a seeming plateau in the middle of the unigram
+entropy curve for each of the language corpora we examined in CHILDES.
+
+However, because the portion of sentences of length greater than \(10\)
+in the Wikipedia corpora were significantly larger than the CHILDES
+corpora, then we also computed relative slope treatments. In the
+\emph{relative 5} treatment, we computed the slopes between \(0\%\) and
+\(20\%\), \(20\%\) and \(40\%\), \(40\%\) and \(60\%\), \(60\%\) and
+\(80\%\) and \(80\%\) and \(100\%\) of the relative word positions in
+each sentence. When one of these percentages was not a whole number,
+then the closest whole number position was used instead for slope
+calculation. In the \emph{relative 10} treatment, we computed the slopes
+between every \(10\%\) of the relative word positions in each sentence.
+Each comparative slope within each treatment was averaged together
+between different sentence lengths, for example in the \emph{relative 5}
+treatment then all of the \(0\%\) to \(20\%\) slopes were averaged
+together. This created three treatments for the entropy curve for each
+langauge in the Wikipedia database.
+
+For the German and Japanese entropy curves, we observed the same shape
+for the text Wikipedia corpora entropy curves and the speech CHILDES
+corpora entropy curves, supporting the argument for robustness across
+text and speech.
 
@@
 
-\section{Conclustions and future
-research}\label{conclustions-and-future-research}
+In the \emph{absolute} and \emph{relative 5} treatments, each language
+is cast as a point in \(5\)-dimensional space. In the \emph{relative 10}
+treatment, each language is cast as a point in \(10\)-dimensional space.
+This allows for direct comparison using cosine similarity and moreover
+for unsupervised clustering analysis to compare the results of the
+Wikipedia entropy analysis to known typological features within the
+langauges in the Wikipedia dataset. Using the R hclust package, we
+performed a hierarchical clustering of the Wikipedia slope data within
+each treatment. A subset of the results at a glance for the
+\emph{absolute} treatment are below.
 
-The \emph{wrap-up effect} is a popular hypothesis within the
-eye-tracking community. The hypothesis states that, in written text,
-sentence-final words are processed more slowly on average then
-sentence-medial or sentence-initial words, due to readers integrating
-information from the entire sentence in order to form a final, coherent
-thought expressed by the sentence, among other reasons. With these two
-ideas, information production distribution in spoken and written
-communication, and information integration in reading written texts, we
-move on to discuss cross-linguistic differences in the distribution of
-information across sentences.
+\begin{CodeChunk}
+\begin{figure*}[h]
 
-\section{Acknowledgements}\label{acknowledgements}
+{\centering \includegraphics{figs/absolute_clusters-1}
 
-Place acknowledgments (including funding information) in a section at
-the end of the paper.
+}
 
-\section{References}\label{references}
+\caption[Absolute clustering of Wikipedia corpora]{Absolute clustering of Wikipedia corpora}\label{fig:absolute_clusters}
+\end{figure*}
+\end{CodeChunk}
+
+\subsection{Language Features}\label{language-features}
+
+To determine which phonological, morphological and syntactic features
+affected the embedding of a language in the Wikipedia dataset, we looked
+at the list of \(144\) linguistic features in World Atlas of Language
+Structures (Dryer \& Haspelmath, 2013). We limited the languages in the
+WALS database to only those in our Wikipedia dataset and performed
+missing-value imputation to obtain the features not coded in WALS for
+the languages from Wikipedia. We then performed k-modes clustering and
+compared the results with the hierarchical clustering from Wikipedia.
+{[}Still working on this{]}
+
+We also checked the effects of individual features on the embeddings of
+languages in the different treatments. {[}Still working on this{]}
+
+\section{Conclusions}\label{conclusions}
+
+In this paper we have explored the differences in rate of information
+transmission among spoken child-adult conversational corpora in several
+languages and text corpora pulled from Wikipedia for well over a hundred
+other corpora. We have seen that the entropy curve, a representation of
+information at each position of an utterance without regard to context,
+varies from language to language and is predicted by a subset of
+syntactic, phonological and morphological features in a language. We
+have provided evidence against the Uniform Information Density
+hypothesis and provided an alternate explanation for the rate at which
+information is encoded and transmitted in human communication, namely
+one that is robust across corpora, language dependent and arises from
+the features of the language.
+
+A large volume of research has indicated the effects of surprisal on
+fixation duration during eye-tracking studies: higher surprisal of a
+word corresponds to higher fixation duration. The \emph{wrap-up effect}
+is a popular hypothesis within the eye-tracking community. The
+hypothesis states that, in written text, sentence-final words are
+processed more slowly on average then sentence-medial or
+sentence-initial words, due to readers integrating information from the
+entire sentence in order to form a final, coherent thought expressed by
+the sentence, among other reasons. The wrap-up effect is drawn from
+evidence in English, German and Dutch, all of which are languages with a
+large final increase in their entropy curve from our study. We aim to
+see in the future if as a consquence of our work in this paper, the
+wrap-up effect falls out from the final jump in the entropy curves in
+some languages.
 

### 2019-01-15 cogsci/position-info.Rmd
Cogsci edits to-do list

@@
 ---
-title: "A Language's Unigram Entropy Distribution Predicts Self-Paced Reading Times"
+title: "A Language's Unigram Entropy Distribution Predicts Self-Paced Reading Times [Rewrite title]"
 bibliography: library.bib
@@
 abstract: 
-    "A major line of research in the cognitive sciences has asked how speakers and writers distribute information in the linguistic material they produce.
-We provide evidence against the Uniform Information Density hypothesis (Jaeger, 2010) which proposes that information is transmitted at a constant rate close to channel capacity in human communication.
+    "A major line of research in the cognitive sciences has asked how speakers and writers distribute information in the linguistic material they produce [include citations here or rewrite this paragraph].
+We provide evidence against the Uniform Information Density hypothesis (Jaeger, 2010) which proposes that information is transmitted at a constant rate close to channel capacity in human communication [Why does it matter that this is the hypothesis we are arguing against].
 We instead argue that the information structure and transmission rate in human communication varies predictably from language to language based on phonological and syntactic features.
-We draw evidence from written and spoken corpora, including corpora of parent and child communication at various ages and large-scale online corpora from Wikipedia in a typologically diverse collection of languages."
+We draw evidence from written and spoken corpora, including corpora of parent and child communication at various ages and large-scale online corpora from Wikipedia in a typologically diverse collection of languages [Rewrite final sentence for clarity]."
 
@@
 Does speech, with its phonological properties, different from writing, with its semiotic properties, in how information is conveyed?
+[Are we pitting speaking against writing?]
 
@@
 Jaeger argues that speakers will try to distribute information more evenly over the course of an utterance, to be as close to channel capacity (in a Shannon sense) as possible.
-In particular, Jaeger singles out the production of "that" at the beginning of relative clauses in English and the use of contractions such as "he's" and "you're" versus "he is" and "you are" as instances where higher information density in the clause elicits the production of more material by speakers.
+In particular, Jaeger singles out the production of "that" at the beginning of relative clauses in English and the use of contractions such as "he's" and "you're" versus "he is" and "you are" as instances where higher information density in the clause elicits the production of more material by speakers [Explain the Jaeger data more clearly and succintly, connect with other examples in the literature more broadly].
 
@@
 
-To explore information transmission rate in speech, we used the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations.
+To explore information transmission rate in speech, we used the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations [Why does it matter to examine child vs adult data].
 We first used the Brown and Providence English corpora from CHILDES.
@@
 the number of sentences of length $k$.
+[Give the equation its own paragraph]
 
@@
 The Unigram Information Density account would predict that children have a lower channel capacity for communication and would instead communicate with a less unigram entropy curve.
-The adult and child entropy curves track one another almost identically.
+The adult and child entropy curves track one another almost identically [Make this contrast with the prediction stronger, introduce .
 This not only indicates a robustness in the unigram entropy curve across speakers, but also across ages and addressees.
@@
 
-# Large-scale text data
+# Large-scale text data [Why does this comparison matter?]
 We also wanted to apply this analysis to large-scale text data, to compare with the parent-child speech data results from using CHILDES.
@@
 
+## Flora's feedback: 
+1. Extend the background section with more information about the literature (what are some of the consequence of the unigram entropy, where has this been applied). 
+2. Talk about the speech and text distinction in the literature and how that affects your work. 
+3. Talk about how your work looks at multiple languages (has this been done before). 
+4. Talk more about child versus adult distinction. 
+6. Talk more about entropy in general, and then why this measure applies to this work (Genzel and Charniak?). 
+7. Have a method and results section (i.e. clearer division) with both the CHILDES and the Wikipedia data.
+## Self feedback:
+1. Motivating sections and why you chose the corpora you did
+2. Clean up and rerun Wikipedia code
+3. Write WALS features section. Why do we care about the typological features? 
+4. Write clustering section and section about what features matter explored using statistical modeling. 
+5.
+Spinal column: Why does information structure through entropy matter?
+How has it impacted the literature (Genzel & Charniak, Hale, Levy, Piantadosi, etc.)?
+Why does UID matter?
+What does the stuff we do here show?
+Why do we choose the particular corpora, entropy methods and analyses we do?
+How does this potentially connect with eye-tracking and sentence processing in general?
+
 ```{r}

### 2019-01-22 cogsci/position-info.Rmd
updated CogSci intro, working on clustering comparison

@@
 ---
-title: "A Language's Unigram Entropy Distribution Predicts Self-Paced Reading Times [Rewrite title]"
+title: "Rate of Information "
 bibliography: library.bib
@@
 
-How do people convey information in speech and text?
-Is there a regular distribution to how information is conveyed across speech and text?
-Does speech, with its phonological properties, different from writing, with its semiotic properties, in how information is conveyed?
-[Are we pitting speaking against writing?]
+When people communicate in speaking or writing, they can have a variety of different goals.
+Perhaps the most important of these goals is to communicate information.
+If we take this as the primary goal of communication, then how information is transmitted should have regular properties.
+There is a rich literature on information, dating back to Claude Shannon’s theory of information (1948).
+We know from Shannon’s work that rate of information transmission should be constant and as close as possible to the capacity of the communication channel.
+Levy & Jaeger (2007) introduce the *uniform information density* hypothesis, arguing that speakers structure utterances so as to optimize information density.
+Viewing human communication as a noisy channel, the authors argue that speakers and writers should attempt to communicate information as evenly as possible and as close to the channel capacity as possible.
+Their evidence comes in the form of the use or lack thereof of “that”-complementizers in introducing English relative clauses.
+When the word preceding the relative clause or the initial word of the relative clause has high information content, then the authors argue that speakers are more likely to use the complementizer.
+The authors use adult-adult conversations from the Penn Treebank corpus (CITATION).
 
-A major line of research in the current century has asked how speakers and writers distribute information in the linguistic material they produce.
-This begins with Claude Shannon in the years after World War II.
-Shannon (1948) defined "information" as a "reduction in uncertainty".
-Shannon followed by quantifying uncertainty in his measure of *entropy*: the amount of uncertainty on the outcome of a random variable.
-Shannon proposed that the most efficient method of sending information through a noisy channel is at a constant rate.
-Genzel & Charniak (2002) apply this principle to human communication: if people communicate information through speech and text optimally, then we should transmit information to one another at a constant rate.
-Genzel & Charniak created a distribution of estimates for sentence-level entropy, each sentence taken without context, for newspaper articles from the Penn Treebank corpus.
-Obtaining a roughly monotonically increasing linear distribution, Genzel & Charniak proposed the *Constant Entropy Rate* (CER) principle: the entropy rate in speech and text with context should be constant as sentence number increases.
-Aylett & Turk (2004) proposed a similar principle for the phonological level.
+Zhan and Levy (2018) study Mandarin Chinese classifier use, and find that the use of specific classifiers, versus general classifiers, appears more often in cases where the production of the corresponding noun is more difficult than when the production is easier, as would be predicted by the UID perspective.
 
-Jaeger (2010) extends the CER principle to all levels of human communication: the *Uniform Information Density* (UID) hypothesis.
-Jaeger argues that speakers will try to distribute information more evenly over the course of an utterance, to be as close to channel capacity (in a Shannon sense) as possible.
-In particular, Jaeger singles out the production of "that" at the beginning of relative clauses in English and the use of contractions such as "he's" and "you're" versus "he is" and "you are" as instances where higher information density in the clause elicits the production of more material by speakers [Explain the Jaeger data more clearly and succintly, connect with other examples in the literature more broadly].
+Let us assume that the UID perspective is correct.
+Then entropy should monotonically and uniformly increase at all word positions across utterances on average.
+We will find a contradiction with this claim through examining the average entropy at each position of a sentence for both speech and text data.
 
-More recent work has stood out in contrast to the traditional UID perspective.
-Zhan and Levy (2018) study Mandarin Chinese classifier use, and find that the use of specific classifiers, versus general classifiers, appears more often in cases where the production of the corresponding noun is more difficult than when the production is easier, as would be predicted by UID.
-The UID perspective is challenged in Yu et al.
-(2016), by performing an analysis of entropy by position within sentence in the text portion of the British National Corpus.
-This work serves as the basis for ours.
+In Yu et al.
+(2016), the authors challenge the UID perspective through computing the unigram positional entropy within the text portion of the British National Corpus (BNC).
+We follow Yu et al.
+(2016) in using the following formula for each word position X of sentences of fixed length k from the corpus, where each $w$ is a word occurring in position $X$ and $p(w)$ is the number of times word $w$ occurs in position X divided by the number of total words that occur in position $X$ i.e.
+the number of sentences of length $k$.
+[Give the equation its own paragraph]
+
+$$H(X) = \sum\limits_w p(w)\log\big(p(w)\big)$$
+
+The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances
 
@@
 
-To explore information transmission rate in speech, we used the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations [Why does it matter to examine child vs adult data].
+For spoken data, we choose to examine child-adult conversations in the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations in multiple languages.
+We expect based on the UID account that the entropy curves of the children will be significantly lower than those of the adults, due to the children's comparatively lower channel capacity in information transmission [more support].
 We first used the Brown and Providence English corpora from CHILDES.
@@
 Finally, within each utterance length, we computed the unigram entropy measure for each position.
-We follow Yu et al.
-(2016) in using the following formula for each word position X of sentences of fixed length k from the corpus, where each $w$ is a word occurring in position $X$ and $p(w)$ is the number of times word $w$ occurs in position X divided by the number of total words that occur in position $X$ i.e.
-the number of sentences of length $k$.
-[Give the equation its own paragraph]
 
-$$H(X) = \sum\limits_w p(w)\log\big(p(w)\big)$$
-
-The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances in the English CHILDES corpora.
+In the English CHILDES corpora.
 The entropy curves capture individual variation across positions in utterances of the same length.
@@
 # Results & Analysis
-The Unigram Information Density account would predict that children have a lower channel capacity for communication and would instead communicate with a less unigram entropy curve.
 The adult and child entropy curves track one another almost identically [Make this contrast with the prediction stronger, introduce .

### 2019-01-22 wiki_entropy/wiki_clustering.Rmd
updated CogSci intro, working on clustering comparison

@@
 library(knitr)
-library(klaR)
 library(missMDA)
-library(clusteval)
-library(factoextra)
 library(cluster)
@@
 
-```{r evaluate_clusters}
-wals_clusters <- wals_clust["cluster"] %>% unlist(use.names= FALSE)
-```
-
-```{r Gelman}
-## from http://www.stat.columbia.edu/~gelman/bda/regression.R
-
-sigma.hat <- function (object){
-  object.class <- class(object)[[1]]
-  if (object.class=="lm"){
-    sigma <- summary(object)$sigma
-    return (sigma)
-  }
-  else if (object.class=="glm"){
-    sigma <- summary(object, correlation=TRUE)$sigma
-    return (sigma)
-  }
-  else if (object.class=="lmer"){
-    object <- summary (object)
-    fcoef <- .Call("mer_fixef", object, PACKAGE = "lme4")
-    useScale <- object@devComp[8]
-    ngrps <- lapply(object@flist, function(x) length(levels(x)))
-    n.groupings <- length (ngrps)
-    varc <- VarCorr (object, useScale=useScale)
-    sc <- attr(varc, "sc")
-    recorr <- lapply(varc, function(el) el@factors$correlation)
-    reStdDev <- c(lapply(recorr, slot, "sd"), list(Residual = sc))
-    sigmas <- as.list (rep (NA, n.groupings+1))
-    sigmas[1] <- ifelse (useScale, sc, NA)
-    cors <- as.list (rep (NA, n.groupings+1))
-    names (sigmas) <- names (cors) <- c ("data", names (varc))
-    for (k in 1:n.groupings){
-      sigmas[[k+1]] <- reStdDev[[k]]
-      cors[[k+1]] <- as.matrix (recorr[[k]])
-      if (length (cors[[k+1]]) == 1) cors[[k+1]] <- NA
-    }
-    return (list (sigma=sigmas, cors=cors))
-  }
-}
-```
-
-```{r WALS_imputation_functions}
-#taken from https://www.tutorialspoint.com/r/r_mean_median_mode.htm
-getmode <- function(v) {
-   uniqv <- unique(v)
-   uniqv[which.max(tabulate(match(v, uniqv)))]
-}
-
-## imputation based on Gelman & Hill Chapter 25
-random.imp <- function(a) {
-  missed <- is.na(a)
-  num_missed <- sum(missed)
-  a.obs <- a[!missed]
-  imputed <- a
-  imputed[missed] <- base::sample(a.obs, num_missed, replace=TRUE)
-  return(imputed)
-}
-
-impute <- function(a, a.impute) {
-  ifelse(is.na(a), a.impute, a)
-}
-
-get_factor <- function(cname) {
-  a.imp <- wf.sub[c(cname)] %>%
-    random.imp() %>%
-    unlist(use.names = FALSE)
-}
-```
-
-```{r WALS_imputation_code}
-FEATURES <- c(1:144)
-for (i in c(1:144)) {
-  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
-} #get all WALS features in correct syntax
-
-NUM_CLUSTERS <- 4
-
-wf <- wals.feature(FEATURES)
-wf.sub <- wf[c("1A", "2A", "3A", "4A", "language")]
-wf.sub <- wf.sub %>% 
-  filter(language %in% LANGUAGES) %>%
-  unique()
-
-for (i in 1:4) {
-  wf.sub[, i] <- factor(wf.sub[, i])
-}
-
-a1.imp <- get_factor("1A")
-a2.imp <- get_factor("2A")
-a3.imp <- get_factor("3A")
-a4.imp <- get_factor("4A")
-
-a1.mode <- getmode(wf.sub$`1A`[!is.na(wf.sub$`1A`)]) #gets the most common non-na value of column 1A
-
-n.sims <- 10
-for (s in 1:n.sims) {
-  a1 <- wf.sub$`1A`
-  a1[is.na(a1)] <- getmode(a1[!is.na(a1)])
-  lm.1 <- glm(a1 ~ a2.imp + a3.imp + a4.imp, family = "binomial")
-  pred.1 <- rnorm(n, predict(lm.1), sigma.hat(lm.1)) ####
-  a1.imp <- impute(wf.sub[c("1A")], pred.1)
-}
-
-```
-
-```{r WALS_clustering}
-FEATURES <- c(1:144)
-for (i in c(1:144)) {
-  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
-} #get all WALS features in correct syntax
-
-wf <- wals.feature(FEATURES)
-wf %>% View()
-wf <- wf[, -(1:3)]
-wf <- wf[, -(ncol(wf) - 1)]
-wf <- wf %>% 
-  filter(language %in% LANGUAGES)
-wf <- wf[, complete.cases(wf)] #this significantly trims down the dataset but is necessary for k-modes
-rv <- kmodes(wf[, -(ncol(wf))], NUM_CLUSTERS, iter.max=10, weighted = FALSE)
-
-```
-
 ```{r}
@@
          language %in% wf_sub_langs$language) %>%
-  select(language, feature, value) %>%
+  dplyr::select(language, feature, value) %>%
   group_by(feature) %>%
@@
   group_by(feature) %>% # for each pair of language, does this feature assin them the same value
-  nest() %>%
+  nest() %>% #maps features to a table mapping languages to numerical factor feature values
   mutate(cors = map(data, my_pairwise_cor))
@@
 
+wf_comp <- wf_sub %$%
+  data %>%
+  do.call(cbind, args=.) 
+wf_comp <- wf_comp[, -seq(3, length(wf_comp), 2)] #get values and one column of language
+
+langs_to_compare <- wf_comp$language
+wkpd_comp <- wkpd %>%
+    filter(Language %in% langs_to_compare)
+
 ```
+
+```{r evaluate_clusters}
+wals_clusters <- wals_clust["cluster"] %>% unlist(use.names= FALSE)
+```

### 2019-01-23 wiki_entropy/wiki_clustering.Rmd
nearly there with cosine comparison

@@
 library(widyr)
+library(lsa)
+library(magrittr)
 
@@
   group_by(feature) %>% # for each pair of language, does this feature assin them the same value
-  nest() %>% #maps features to a table mapping languages to numerical factor feature values
-  mutate(cors = map(data, my_pairwise_cor))
-  select(-data) %>%
-  unnest()
+  nest() #maps features to a table mapping languages to numerical factor feature values
+  # mutate(cors = map(data, my_pairwise_cor))
+  # select(-data) %>%
+  # unnest()
 
@@
   do.call(cbind, args=.)
-wf_comp <- wf_comp[, -seq(3, length(wf_comp), 2)] #get values and one column of language
 
+wf_comp <- wf_comp[, -seq(3, length(wf_comp), 2)] 
 langs_to_compare <- wf_comp$language
+wf_comp %<>% 
+  as.data.frame(row.names=wf_comp$language)
+
 wkpd_comp <- wkpd %>%
     filter(Language %in% langs_to_compare) 
+wkpd_comp %<>% as.data.frame(row.names=wkpd_comp$Language)
+
+lang_pairs <- langs_to_compare %>%
+  combn(2, simplify = FALSE)
+
+K <- length(lang_pairs)
+
+lang_x <- lang_pairs %>%
+    unlist() %>%
+    matrix(nrow = K, byrow=TRUE) %>%
+    as.data.frame() %>%
+    mutate(wiki = lsa::cosine(wkpd_comp[V1, ], wkpd_comp[V2, ]), 
+           wals = lsa::cosine(wf_comp[V1, ], wf_comp[V2, ]))
 

### 2019-01-23 cogsci/position-info.Rmd
working on rewriting intro

@@
 If we take this as the primary goal of communication, then how information is transmitted should have regular properties.
-There is a rich literature on information, dating back to Claude Shannon’s theory of information (1948).
-We know from Shannon’s work that rate of information transmission should be constant and as close as possible to the capacity of the communication channel.
-Levy & Jaeger (2007) introduce the *uniform information density* hypothesis, arguing that speakers structure utterances so as to optimize information density.
+From Shannon (1948), we know that when information is being communicated, the rate of information transmission should be constant and as close as possible to the capacity of the communication channel to optimally transmit the information.
+Levy & Jaeger (2007) introduce the *uniform information density* (UID) hypothesis, arguing that speakers structure utterances so as to optimize information density.
 Viewing human communication as a noisy channel, the authors argue that speakers and writers should attempt to communicate information as evenly as possible and as close to the channel capacity as possible.
-Their evidence comes in the form of the use or lack thereof of “that”-complementizers in introducing English relative clauses.
+Their evidence concerns the use or lack of use of “that”-complementizers in introducing English relative clauses.
 When the word preceding the relative clause or the initial word of the relative clause has high information content, then the authors argue that speakers are more likely to use the complementizer.
-The authors use adult-adult conversations from the Penn Treebank corpus (CITATION).
+The authors use adult-adult conversations from the Penn Treebank corpus.
 
-Zhan and Levy (2018) study Mandarin Chinese classifier use, and find that the use of specific classifiers, versus general classifiers, appears more often in cases where the production of the corresponding noun is more difficult than when the production is easier, as would be predicted by the UID perspective.
-
-Let us assume that the UID perspective is correct.
-Then entropy should monotonically and uniformly increase at all word positions across utterances on average.
-We will find a contradiction with this claim through examining the average entropy at each position of a sentence for both speech and text data.
+However, some recent work has contradicted the UID hypothesis.
+Zhan and Levy (2018) study Mandarin Chinese classifier use for specific and general Mandarin classifiers.
+Specific classifiers are specific to certain nouns, and UID would predict that specific classifiers would appear when the production of the corresponding noun is easier, to spread the information contained by the classifier and noun less densely over time.
+General classifiers can be applied to any noun and contain no special information content.
+Zhan and Levy find that speakers use specific classifiers  more often in cases where the production of the corresponding noun is more difficult than when the production is easier; speakers do not avoid unexpected peaks in information content for their listeners, but instead maximize their ease of production.
 
@@
 the number of sentences of length $k$.
-[Give the equation its own paragraph]
 
@@
 The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances
+
+Let us assume that the UID perspective is correct.
+Then entropy should monotonically and uniformly increase at all word positions across utterances on average.
+We will find a contradiction with this claim through examining the average entropy at each position of a sentence for both speech and text data.
 

### 2019-01-23 cogsci/position-info.Rmd
working on yu et al section of intro

@@
 
-In Yu et al.
-(2016), the authors challenge the UID perspective through computing the unigram positional entropy within the text portion of the British National Corpus (BNC).
-We follow Yu et al.
-(2016) in using the following formula for each word position X of sentences of fixed length k from the corpus, where each $w$ is a word occurring in position $X$ and $p(w)$ is the number of times word $w$ occurs in position X divided by the number of total words that occur in position $X$ i.e.
+Yu et al.
+(2016) challenges the UID hypothesis through examining the distribution of uncertainty across sentence positions within the text portion of the English-language British National Corpus (BNC).
+They partition the corpus by sentence length in number of words, and proceed for each sentence length $k$ within the corpus.
+for each word position X of sentences of fixed length k from the corpus, where each $w$ is a word occurring in position $X$ and $p(w)$ is the number of times word $w$ occurs in position X divided by the number of total words that occur in position $X$ i.e.
 the number of sentences of length $k$.
@@
 $$H(X) = \sum\limits_w p(w)\log\big(p(w)\big)$$
-
+With this measure, Yu et al.
+compute the unigram entropy at each position of sentences of each length within the corpus.
 The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances

### 2019-01-23 wiki_entropy/wiki_clustering.Rmd
regression framework

@@
 library(magrittr)
+library(broom)
 
@@
 ```{r}
+cosines <- wkpd %>%
+  t() %>%
+  cosine() %>%
+  as_data_frame(rownames = "language1") %>%
+  gather(language2, cosine, -language1)
+
+```
+
+```{r}
 
@@
          language %in% wf_sub_langs$language) %>%
-  dplyr::select(language, feature, value) %>%
+  select(language, feature, value) %>%
+  mutate(value = as.numeric(as.factor(value)))
+
+wf_pairwise <- expand.grid(language1 = wf_sub$language,
+                           language2 = wf_sub$language) %>%
+  left_join(wf_sub, by = c("language1" = "language")) %>%
+  rename(value1 = value) %>%
+  left_join(wf_sub, by = c("language2" = "language", "feature")) %>%
+  rename(value2 = value) %>%
+  mutate(same = value1 == value2) %>%
+  select(-value1, -value2) %>%
+  left_join(cosines, by = c("language1", "language2"))
+
+models <- wf_pairwise %>%
   group_by(feature) %>%
-  mutate(value = as.numeric(as.factor(value))) %>%
-  group_by(feature) %>% # for each pair of language, does this feature assin them the same value
-  nest() #maps features to a table mapping languages to numerical factor feature values
-  # mutate(cors = map(data, my_pairwise_cor))
-  # select(-data) %>%
-  # unnest()
+  nest() %>%
+  mutate(model = map(data, ~glm(same ~ cosine, 
+                                family = "binomial", data = .)))
+```
+
+```{r}
 

### 2019-01-24 cogsci/position-info.Rmd
rewrites

@@
 ---
-title: "Rate of Information "
+title: "Information is Not Communicated Uniformly: Evidence from Spoken and Written Corpora"
 bibliography: library.bib
@@
 abstract: 
-    "A major line of research in the cognitive sciences has asked how speakers and writers distribute information in the linguistic material they produce [include citations here or rewrite this paragraph].
-We provide evidence against the Uniform Information Density hypothesis (Jaeger, 2010) which proposes that information is transmitted at a constant rate close to channel capacity in human communication [Why does it matter that this is the hypothesis we are arguing against].
-We instead argue that the information structure and transmission rate in human communication varies predictably from language to language based on phonological and syntactic features.
-We draw evidence from written and spoken corpora, including corpora of parent and child communication at various ages and large-scale online corpora from Wikipedia in a typologically diverse collection of languages [Rewrite final sentence for clarity]."
+    "We provide evidence against the popular Uniform Information Density hypothesis (Levy & Jaeger, 2007) which proposes that information is transmitted at a constant rate close to channel capacity in human communication.
+Using a method based on the original Genzel & Charniak (2002) entropy measure, we construct a word-level model for entropy applicable to both spoken and written corpora.
+We apply this model to corpora from Wikipedia in well over a hundred languages.
+We find that not only is the Uniform Information Density hypothesis prediction wrong, but also that the by-word entropy distribution of a language is related to the typological features of the language.
+We use this evidence to suggest that people do not communicate information at a uniform rate, but that information distribution varies from language to language based on phonogical, morphological and syntactic features."
 
@@
 
+What Zhan and Levy (2018) and most work in favor of UID have in common is they focus on information distribution at particular points in a sentence.
+By contrast, Jain et al.
+(2018) examines word order across sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on word order.
+For this paper, we focus on the sentence level instead of specific points, following a method described in Yu et al.
+(2016).
+
 Yu et al.
@@
 They partition the corpus by sentence length in number of words, and proceed for each sentence length $k$ within the corpus.
-for each word position X of sentences of fixed length k from the corpus, where each $w$ is a word occurring in position $X$ and $p(w)$ is the number of times word $w$ occurs in position X divided by the number of total words that occur in position $X$ i.e.
+for each word position $X$ of sentences of length $X$, let each word $w$ be a unique word occurring in position $X$ and $p(w)$ is the number of times word $w$ occurs in position $X$ divided by the number of total words that occur in position $X$ i.e.
 the number of sentences of length $k$.
+Then $p(w)$ is the probability of randomly obtaining word $w$ by choosing a word in position $X$ in sentences of length $k$.
 
@@
 compute the unigram entropy at each position of sentences of each length within the corpus.
-The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances
+The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances.
+Genzel & Charniak (2002) similarly examine a unigram entropy measure on sentences, and found that entropy at the sentence level increases linearly with sentence index within a corpus.
+UID applies this uniformity of entropy rate in sentences to all levels of speech, and so the Yu et al.
+method, which examines text at the word level, should find an affine function at the word level.
 
 Let us assume that the UID perspective is correct.
-Then entropy should monotonically and uniformly increase at all word positions across utterances on average.
+Then we expect that the entropy curve for every language should be close to an affine, linear function with some noise.
+Entropy should monotonically and uniformly increase at all word positions across utterances on average.
 We will find a contradiction with this claim through examining the average entropy at each position of a sentence for both speech and text data.
 
-# Methods
+# Spoken data
 
-For spoken data, we choose to examine child-adult conversations in the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations in multiple languages.
-We expect based on the UID account that the entropy curves of the children will be significantly lower than those of the adults, due to the children's comparatively lower channel capacity in information transmission [more support].
+## Methods
+
+We begin with the same source as Levy & Jaeger (2007) and Jaeger (2010), the Switchboard corpus of telephone conversations [CITATION].
+
+We also examine child-adult conversations in the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations in multiple languages.
+Children are not fully developed speakers, and so we expect to observe a different shape of the distribution.
 We first used the Brown and Providence English corpora from CHILDES.
@@
 
-We also ran this analysis on Spanish, German, Mandarin and Japanese corpora from CHILDES.
+We also ran this analysis on Spanish and Mandarin corpora from CHILDES.
 We used the Shiro corpus for Spanish (Shiro, 2000), which contains prompted narratives individually collected from over a hundred Venezualan schoolchildren, half from high SES backgrounds and half from low SES backgrounds.
-The German corpus we used was the Wagner corpus of spontaneous parent-child speech in the home, collected in single sessions for children between 1 and 14 years of age (Wagner, 1985).
 We used the Zhou dinner corpus for Mandarin Chinese (Li & Zhou, 2015), which contains dinner conversations between 5 to 6-year-old children and their parents collected in Shanghai.
-We used the Okayama corpus for Japanese (Okayama, 1973), which consists of conversations between 2 to 4-year-old children and their mothers in single sessions.
 
@@
 For Mandarin, we used pinyin transliterations of the utterances in the corpus with demarcated word boundaries, and for Japanese we used romanji (Roman alphabet) transliterations of words in the corpus.
-Japanese Hiragana, Katakana and Kanji writing systems, and the Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Japanese and Chinese writing including spaces between word boundaries can have a negative effect on reading times (Sainio et al, 2007; Bai et al, 2008).
+The Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Chinese writing including spaces between word boundaries can have a negative effect on reading times (Bai et al, 2008).
 
@@
 img <- png::readPNG("figs/German_unigram_entropies_with_stemming.png")
-grid::grid.raster(img)
-```
-
-```{r wagner_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Wagner corpus unigram entropy"}
-img <- png::readPNG("figs/Spanish_unigram_entropies_with_stemming.png")
 grid::grid.raster(img)
@@
 
-```{r okayama_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Okayama corpus unigram entropy"}
-img <- png::readPNG("figs/Japanese_okayama_entropy.png")
-grid::grid.raster(img)
-```
-
-# Results & Analysis
+## Results & Analysis
 The adult and child entropy curves track one another almost identically [Make this contrast with the prediction stronger, introduce .
@@
 
-We found a distinct three-step distribution for English, Spanish and German CHILDES corpora, with a slight dip in the penultimate position of each sentence.
-We attribute the penultimate dip in these unigram entropy curves to the fact that most of the utterances in the CHILDES English, Spanish and German corpora we examined had a determiner such as "the" or "a" in the second-to-last position of utterances.
-The beginnings of utterances in the English, German and Spanish CHILDES data were usually pronouns or grammatical subjects, while the final words were grammatical objects and had a great deal of variation in the exact word that appeared in the utterance-final position.
+We found a distinct three-step distribution for English and Spanish CHILDES corpora, with a slight dip in the penultimate position of each sentence.
+We attribute the penultimate dip in these unigram entropy curves to the fact that most of the utterances in the CHILDES English and Spanish corpora we examined had a determiner such as "the" or "a" in the second-to-last position of utterances.
+The beginnings of utterances in the English and Spanish CHILDES data were usually pronouns or grammatical subjects, while the final words were grammatical objects and had a great deal of variation in the exact word that appeared in the utterance-final position.
 The final position of utterances in child-directed speech is known to be important, dating back to Aslin (1993).
-The Mandarin and Japanese data, by comparison, have noticeably lower positional entropy values in utterance-final positions than in utterance-penultimate positions.
-The Japanese entropy curve begins high and then drops, unlike all of the other entropy curves.
+The Mandarin corpus entropy curve, by comparison, has a noticeably lower positional entropy values in utterance-final positions than in utterance-penultimate positions.
 
-```{r MI_wagner, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Mutual Information measures on utterances of length 6 in the Wagner corpus"}
-img <- png::readPNG("figs/wagner_mi_6.png")
-grid::grid.raster(img)
-```
+# Written data
+We also wanted to apply this analysis to large-scale text data, to compare with the parent-child speech data results from using CHILDES. WHY ARE WE DOING THIS.
 
-```{r MI_okayama, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Mutual Information measures on utterances of length 6 in the Okayama corpus"}
-img <- png::readPNG("figs/okayama_mi_6.png")
-grid::grid.raster(img)
-```
+## Methods
 
-The mutual information statistic is a symmetric measure of the information gained about one random variable by learning about another, a bigram measure and indicator of the effect than the context of a word can have on its unigram entropy value.
-As seen in the images below, the final word in utterances of length $6$ in German has a significant mutual information value with all other words in the utterance on average.
-This means that the words appearing in positions $1$ through $5$ all play a role in determining the word that appears in position $6$ and vice versa.
-By contrast, the words that appear in positions $1$ through $4$ have little influence in determining the word that appears in position $5$, the penultimate word that is often a determiner or function word, and instead the final word has the greatest influence in determining the penultimate word.
-For Japanese, the final word in an utterance has little to no influence on determining the rest of the words and vice versa, whereas the other words all play in role in determining what will go in each other's positions.
-
-# Large-scale text data [Why does this comparison matter?]
-We also wanted to apply this analysis to large-scale text data, to compare with the parent-child speech data results from using CHILDES.
 Using Giuseppe Attardi's Wikiextractor tool \footnote{https://github.com/attardi/wikiextractor}, we extracted the text corpora for $165$ languages from Wikipedia by downloading a stored collection of Wikipedia entries in each langauge and randomly selecting several thousand articles from each Wikipedia language.
@@
 
-## Clustering
+## Results and Analysis
+
 In the *absolute* and *relative 5* treatments, each language is cast as a point in $5$-dimensional space.
@@
 
-```{r absolute_clusters, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Absolute clustering of Wikipedia corpora"}
-img <- png::readPNG("figs/absolute_clusters.png")
-grid::grid.raster(img)
-```
-
-## Language Features
-
 To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we looked at the list of $144$ linguistic features in World Atlas of Language Structures (Dryer & Haspelmath, 2013).
 We limited the languages in the WALS database to only those in our Wikipedia dataset and performed missing-value imputation to obtain the features not coded in WALS for the languages from Wikipedia.
-We then performed k-modes clustering and compared the results with the hierarchical clustering from Wikipedia.
-[Still working on this]
+[Miss MDA or do another kind of missing value imputation, description of method]
 
-We also checked the effects of individual features on the embeddings of languages in the different treatments. [Still working on this]
+One approach is to cluster the languages used in the Wikipedia analysis on the basis of WALS features and then directly compare the WALS clustering results with the results of the hierarchical clustering on the Wikipedia slope data for each of the different treatments using clustering similarity evaluation methods such as the Rand index (Rand, 1971).
+However, the problem then arises of which combination of WALS features and how many features to include in the clustering analysis.
+This is a computationally intractable operation.
+Additionallly, deciding how many clusters to use in an unsupervised clustering analysis is an unsolved problem in machine learning.
+
+We instead check the effects of individual features on the embeddings of languages in the different treatments.
+We computed pairwise cosine similarity between each pair of language vectors within a treatment.
+For a subset of WALS features which had values entered for most of the languages we obtained from Wikipedia, we used a generalized linear model to see whether the cosine similarity between languages mattered in predicting if the languages shared the same value for a WALS feature.
+We found this to be true.
+
+RESULTS OF GLM BELOW.
 

### 2019-01-24 wiki_entropy/wiki_clustering.Rmd
rewrites

@@
 ```{r}
+wiki_gs <- gs_title("Wikipedia_absolute")
+wkpd <- gs_read(wiki_gs) 
+LANGUAGES <- wkpd$Language
+
+FEATURES <- c(1:144)
+for (i in c(1:144)) {
+  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
+} #get all WALS features in correct syntax
+
+wf <- wals.feature(FEATURES)
+
+wf_df <- wf %>%
+  as_data_frame() %>%
+  filter(language %in% LANGUAGES) %>%
+  group_by(language) %>%
+  slice(1) %>%
+  gather(feature, value, `1A`:`144A`) %>%
+  group_by(feature) %>%
+  mutate(value = factor(value))
+
 cosines <- wkpd %>%
+  column_to_rownames("Language") %>%
   t() %>%
@@
 
-```{r}
-
-wf_comp <- wf_sub %$%
-  data %>%
-  do.call(cbind, args=.)
-
-wf_comp <- wf_comp[, -seq(3, length(wf_comp), 2)] 
-langs_to_compare <- wf_comp$language
-wf_comp %<>% 
-  as.data.frame(row.names=wf_comp$language)
-
-wkpd_comp <- wkpd %>%
-    filter(Language %in% langs_to_compare) 
-wkpd_comp %<>% as.data.frame(row.names=wkpd_comp$Language)
-
-lang_pairs <- langs_to_compare %>%
-  combn(2, simplify = FALSE)
-
-K <- length(lang_pairs)
-
-lang_x <- lang_pairs %>%
-    unlist() %>%
-    matrix(nrow = K, byrow=TRUE) %>%
-    as.data.frame() %>%
-    mutate(wiki = lsa::cosine(wkpd_comp[V1, ], wkpd_comp[V2, ]), 
-           wals = lsa::cosine(wf_comp[V1, ], wf_comp[V2, ]))
-
-```
-
 ```{r evaluate_clusters}

### 2019-01-25 cogsci/position-info.Rmd
almost done with second cogsci draft, working on imputation and glm

@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/',
-                      echo=F, warning=F, cache=F, message=F, sanitize = T)
+knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/', echo=F, warning=F, cache=F, message=F, sanitize = T)
 ```
@@
 The authors use adult-adult conversations from the Penn Treebank corpus.
+
+UID has been applied broadly over the past decade, to determining whether linguistic alignment takes place (Jaeger & Snider, 2013), Zipfian word length distributions (Piantadosi et al., 2011), communication efficiency (Mahowald et al., 2013), dialogue and turn-taking (Xu & Reitter, 2018) and the significance of ambiguity in language (Piantadosi et al., 2012), among other research.
 
@@
 
+The UID hypothesis predicts that information transmission rate will tend towards uniformity in both speech and text.
+We begin by examining speech to determine if we obtain the entropy curve shape we expect for spoken communication.
+We begin with the same source as Levy & Jaeger (2007) and Jaeger (2010), the Switchboard corpus of adult telephone conversations [CITATION].
+We also examine child-adult conversations in the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations in multiple languages.
+Children are not fully developed speakers, so we want to compare the entropy curve we obtain by computing over the utterances in the CHILDES corpora to the utterances in the adult-adult Switchboard corpus.
+
 ## Methods
 
-We begin with the same source as Levy & Jaeger (2007) and Jaeger (2010), the Switchboard corpus of telephone conversations [CITATION].
+Switchboard
 
-We also examine child-adult conversations in the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations in multiple languages.
-Children are not fully developed speakers, and so we expect to observe a different shape of the distribution.
 We first used the Brown and Providence English corpora from CHILDES.
@@
 ## Results & Analysis
-The adult and child entropy curves track one another almost identically [Make this contrast with the prediction stronger, introduce .
+The adult and child entropy curves track one another almost identically.
+This is surprising because UID would predict that the young children in the corpora who are not fully developed speakers would have more noise in their distribution of information rates.
 This not only indicates a robustness in the unigram entropy curve across speakers, but also across ages and addressees.
@@
 We found a distinct three-step distribution for English and Spanish CHILDES corpora, with a slight dip in the penultimate position of each sentence.
-We attribute the penultimate dip in these unigram entropy curves to the fact that most of the utterances in the CHILDES English and Spanish corpora we examined had a determiner such as "the" or "a" in the second-to-last position of utterances.
-The beginnings of utterances in the English and Spanish CHILDES data were usually pronouns or grammatical subjects, while the final words were grammatical objects and had a great deal of variation in the exact word that appeared in the utterance-final position.
 The final position of utterances in child-directed speech is known to be important, dating back to Aslin (1993).
@@
 
+We attribute the penultimate dip in the English and Spanish entropy curves to the fact that most of the utterances in the CHILDES English and Spanish corpora we examined had a determiner such as "the" or "a" in the second-to-last position of utterances.
+The beginnings of utterances in the English and Spanish CHILDES data were usually pronouns or grammatical subjects, while the final words were grammatical objects and had a great deal of variation in the exact word that appeared in the utterance-final position.
+
+These entropy curves are not what would be expected from UID.
+The robustness of the three-step distribution for English and its replication in Spanish do not resemble the affine function we would expect from UID.
+The Mandarin entropy curve, which does not at all resemble either the English/Spanish distribution or the predictions of UID, suggests that the entropy curve can vary from language to language.
+UID predicts that each language should have a similar distribution.
+
 # Written data
-We also wanted to apply this analysis to large-scale text data, to compare with the parent-child speech data results from using CHILDES. WHY ARE WE DOING THIS.
+
+The UID hypothesis also applies to written communication: we expect people to communicate information at a uniform rate through writing as well.
+We use Wikipedia as a source for written data, which provides two advantages.
+One, the quantity of data in Wikipedia is large for each language and two, there are hundreds of languages with Wikipedias.
+This allows us to perform the entropy analysis on a much greater scale and to directly compare the results of the entropy analysis on each language to one another, and ultimately to predict what typological features of a language help determine its entropy curve, if any.
+We will describe our method of harvesting and distilling text data from each Wikipedia as well as how we compare the entropy curves from each language to one another.
 
@@
 We limited the languages in the WALS database to only those in our Wikipedia dataset and performed missing-value imputation to obtain the features not coded in WALS for the languages from Wikipedia.
-[Miss MDA or do another kind of missing value imputation, description of method]
 
@@
 
-## Flora's feedback: 
-1. Extend the background section with more information about the literature (what are some of the consequence of the unigram entropy, where has this been applied). 
-2. Talk about the speech and text distinction in the literature and how that affects your work. 
-3. Talk about how your work looks at multiple languages (has this been done before). 
-4. Talk more about child versus adult distinction. 
-6. Talk more about entropy in general, and then why this measure applies to this work (Genzel and Charniak?). 
-7. Have a method and results section (i.e. clearer division) with both the CHILDES and the Wikipedia data.
-## Self feedback:
-1. Motivating sections and why you chose the corpora you did
-2. Clean up and rerun Wikipedia code
-3. Write WALS features section. Why do we care about the typological features? 
-4. Write clustering section and section about what features matter explored using statistical modeling. 
-5.
-Spinal column: Why does information structure through entropy matter?
-How has it impacted the literature (Genzel & Charniak, Hale, Levy, Piantadosi, etc.)?
-Why does UID matter?
-What does the stuff we do here show?
-Why do we choose the particular corpora, entropy methods and analyses we do?
-How does this potentially connect with eye-tracking and sentence processing in general?
-
 ```{r}

### 2019-01-25 wiki_entropy/wiki_clustering.Rmd
almost done with second cogsci draft, working on imputation and glm

@@
 
-```{r clustering}
+```{r hierarchical clustering}
 NUM_CLUSTERS <- 10
@@
   View()
-```
-
-```{r MCA_imputation}
-wiki_gs <- gs_title("Wikipedia_absolute")
-wkpd <- gs_read(wiki_gs) 
-LANGUAGES <- wkpd$Language
-
-FEATURES <- c(1:144)
-for (i in c(1:144)) {
-  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
-} #get all WALS features in correct syntax
-
-wf <- wals.feature(FEATURES)
-
-wf_df <- wf %>%
-  as_data_frame() %>%
-  filter(language %in% LANGUAGES) %>%
-  group_by(language) %>%
-  slice(1) %>%
-  gather(feature, value, `1A`:`144A`) %>%
-  group_by(feature) %>%
-  mutate(value = factor(value))
-
-for (i in 1:4) {
-  wf.sub[, i] <- factor(wf.sub[, i])
-}
-
-wf.sub <- wf.sub[rowSums(is.na(wf.sub)) != 4, ]
-rownames(wf.sub) <- wf.sub$language
-wf.sub$language <- NULL
-rv <- MIMCA(wf.sub, ncp = 4)
-
-imputed_wf <- as.data.frame(rv[[1]][[99]])
-
-wals_clust_full <- kmodes(imputed_wf, NUM_CLUSTERS, iter.max=10, weighted = FALSE)
 ```
@@
 
+```{r imputation for glm}
+imputed_result <- wf_df %>%
+  select(language, feature, value) %>%
+  ungroup() %>% 
+  spread(feature, value) %>% 
+  column_to_rownames("language") %>%
+  sapply(as.factor) %>%
+  as_data_frame() %>%
+  MIMCA(ncp = 7, maxiter = 999)
+
+imputed_wf <- as.data.frame(imputed_result[[1]][[99]])
+```
+
 ```{r evaluate_clusters}

### 2019-01-25 wiki_entropy/wiki_clustering.Rmd
rank models

--- 2019-01-23 wiki_entropy/wiki_clustering.Rmd
@@
 cosines <- wkpd %>%
-  t() %>%
+  column_to_rownames("Language") %>%
+  t()%>%
   cosine() %>%
@@
                                 family = "binomial", data = .)))
+
+models %>%
+  mutate(coeffs = map(model, tidy)) %>%
+  select(-data, -model) %>%
+  unnest() %>%
+  filter(term == "cosine") %>%
+  arrange(desc(statistic))
 ```

### 2019-01-25 wiki_entropy/wiki_clustering.Rmd

--- 2019-01-25 wiki_entropy/wiki_clustering.Rmd
@@
 cosines <- wkpd %>%
-  column_to_rownames("Language") %>%
   t() %>%

### 2019-01-25 cogsci/position-info.Rmd
working on Cogsci submission: bibtex drafted, clustering table here

@@
 library(xtable)
+library(knitr)
 ```
@@
 
-When people communicate in speaking or writing, they can have a variety of different goals.
-Perhaps the most important of these goals is to communicate information.
-If we take this as the primary goal of communication, then how information is transmitted should have regular properties.
+Perhaps the most important reason people talk and write to one another is to give information.
 From Shannon (1948), we know that when information is being communicated, the rate of information transmission should be constant and as close as possible to the capacity of the communication channel to optimally transmit the information.
-Levy & Jaeger (2007) introduce the *uniform information density* (UID) hypothesis, arguing that speakers structure utterances so as to optimize information density.
-Viewing human communication as a noisy channel, the authors argue that speakers and writers should attempt to communicate information as evenly as possible and as close to the channel capacity as possible.
+Levy & Jaeger (2007) introduce the *uniform information density* (UID) hypothesis, arguing that people unconsciously structure their speech and writing to communicate information at a constant, optimal rate.
+Levy & Jaeger (2007) argue that speakers and writers should attempt to communicate information as evenly as possible and as close to the channel capacity as possible.
 Their evidence concerns the use or lack of use of “that”-complementizers in introducing English relative clauses.
@@
 UID has been applied broadly over the past decade, to determining whether linguistic alignment takes place (Jaeger & Snider, 2013), Zipfian word length distributions (Piantadosi et al., 2011), communication efficiency (Mahowald et al., 2013), dialogue and turn-taking (Xu & Reitter, 2018) and the significance of ambiguity in language (Piantadosi et al., 2012), among other research.
-
-However, some recent work has contradicted the UID hypothesis.
+However, other recent work has contradicted the UID hypothesis.
 Zhan and Levy (2018) study Mandarin Chinese classifier use for specific and general Mandarin classifiers.
@@
 
-What Zhan and Levy (2018) and most work in favor of UID have in common is they focus on information distribution at particular points in a sentence.
+Similar to the original Levy & Jaeger (2007), Zhan and Levy (2018) focuses on information distribution at particular points in a sentence.
 By contrast, Jain et al.
-(2018) examines word order across sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on word order.
+(2018) examine word order across spoken sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on word order.
 For this paper, we focus on the sentence level instead of specific points, following a method described in Yu et al.
@@
 Yu et al.
-(2016) challenges the UID hypothesis through examining the distribution of uncertainty across sentence positions within the text portion of the English-language British National Corpus (BNC).
-They partition the corpus by sentence length in number of words, and proceed for each sentence length $k$ within the corpus.
-for each word position $X$ of sentences of length $X$, let each word $w$ be a unique word occurring in position $X$ and $p(w)$ is the number of times word $w$ occurs in position $X$ divided by the number of total words that occur in position $X$ i.e.
+(2016) challenges the UID hypothesis through examining an entropy measure across sentence positions within the text portion of the English-language British National Corpus (BNC).
+They partition the corpus by sentence length in number of words.
+For each word position $X$ of sentences of length $k$, they define $w$ as a unique word occurring in position $X$.
+They define $p(w)$ as the number of times word $w$ occurs in position $X$ divided by the number of total words that occur in position $X$ i.e.
 the number of sentences of length $k$.
-Then $p(w)$ is the probability of randomly obtaining word $w$ by choosing a word in position $X$ in sentences of length $k$.
+Then $p(w)$ is the probability of obtaining word $w$ by choosing a word at random in position $X$ in sentences of length $k$.
 
@@
 We begin with the same source as Levy & Jaeger (2007) and Jaeger (2010), the Switchboard corpus of adult telephone conversations [CITATION].
-We also examine child-adult conversations in the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations in multiple languages.
+We also examine child-adult conversations in the CHILDES TalkBank (Brown 1973; MacWhinney 2014) corpora database of spoken adult-child conversations in multiple languages.
 Children are not fully developed speakers, so we want to compare the entropy curve we obtain by computing over the utterances in the CHILDES corpora to the utterances in the adult-adult Switchboard corpus.
@@
 
-Switchboard
+Switchboard [To be put in]
 
-We first used the Brown and Providence English corpora from CHILDES.
+We used the Brown and Providence English corpora from CHILDES.
 The Brown corpus contains individual recordings of conversations between three children between 1.5 and 6-years-old and their families in their homes.
@@
 We can directly compare any two positions within utterances to determine the amount of uncertainty, and therefore information, on average contained by words within that position of utterances.
-We are applying the same approach as in Genzel and Charniak (2003), but within sentences instead of across sentences.
+We are applying the same approach as in Genzel and Charniak (2002), but within sentences instead of across sentences.
 
@@
 
-RESULTS OF GLM BELOW.
+```{r tables}
+glm_model <- read.csv("glm.csv")
+knitr::kable(glm_model)
+```
 
@@
 # The following code is some latex to format the bibliography. Do not remove it.
+
 ```

### 2019-01-25 cogsci/position-info.tex
working on Cogsci submission: bibtex drafted, clustering table here

@@
 
-\title{A Language's Unigram Entropy Distribution Predicts Self-Paced Reading
-Times}
+\title{Information is Not Communicated Uniformly: Evidence from Spoken and
+Written Corpora}
 
@@
 \begin{abstract}
-A major line of research in the cognitive sciences has asked how
-speakers and writers distribute information in the linguistic material
-they produce. We provide evidence against the Uniform Information
-Density hypothesis (Jaeger, 2010) which proposes that information is
+We provide evidence against the popular Uniform Information Density
+hypothesis (Levy \& Jaeger, 2007) which proposes that information is
 transmitted at a constant rate close to channel capacity in human
-communication. We instead argue that the information structure and
-transmission rate in human communication varies predictably from
-language to language based on phonological and syntactic features. We
-draw evidence from written and spoken corpora, including corpora of
-parent and child communication at various ages and large-scale online
-corpora from Wikipedia in a typologically diverse collection of
-languages.
+communication. Using a method based on the original Genzel \& Charniak
+(2002) entropy measure, we construct a word-level model for entropy
+applicable to both spoken and written corpora. We apply this model to
+corpora from Wikipedia in well over a hundred languages. We find that
+not only is the Uniform Information Density hypothesis prediction wrong,
+but also that the by-word entropy distribution of a language is related
+to the typological features of the language. We use this evidence to
+suggest that people do not communicate information at a uniform rate,
+but that information distribution varies from language to language based
+on phonogical, morphological and syntactic features.
 
@@
 
-How do people convey information in speech and text? Is there a regular
-distribution to how information is conveyed across speech and text? Does
-speech, with its phonological properties, different from writing, with
-its semiotic properties, in how information is conveyed?
+Perhaps the most important reason people talk and write to one another
+is to give information. From Shannon (1948), we know that when
+information is being communicated, the rate of information transmission
+should be constant and as close as possible to the capacity of the
+communication channel to optimally transmit the information. Levy \&
+Jaeger (2007) introduce the \emph{uniform information density} (UID)
+hypothesis, arguing that people unconsciously structure their speech and
+writing to communicate information at a constant, optimal rate. Levy \&
+Jaeger (2007) argue that speakers and writers should attempt to
+communicate information as evenly as possible and as close to the
+channel capacity as possible. Their evidence concerns the use or lack of
+use of ``that''-complementizers in introducing English relative clauses.
+When the word preceding the relative clause or the initial word of the
+relative clause has high information content, then the authors argue
+that speakers are more likely to use the complementizer. The authors use
+adult-adult conversations from the Penn Treebank corpus.
 
-A major line of research in the current century has asked how speakers
-and writers distribute information in the linguistic material they
-produce. This begins with Claude Shannon in the years after World War
-II. Shannon (1948) defined ``information'' as a ``reduction in
-uncertainty''. Shannon followed by quantifying uncertainty in his
-measure of \emph{entropy}: the amount of uncertainty on the outcome of a
-random variable. Shannon proposed that the most efficient method of
-sending information through a noisy channel is at a constant rate.
-Genzel \& Charniak (2002) apply this principle to human communication:
-if people communicate information through speech and text optimally,
-then we should transmit information to one another at a constant rate.
-Genzel \& Charniak created a distribution of estimates for
-sentence-level entropy, each sentence taken without context, for
-newspaper articles from the Penn Treebank corpus. Obtaining a roughly
-monotonically increasing linear distribution, Genzel \& Charniak
-proposed the \emph{Constant Entropy Rate} (CER) principle: the entropy
-rate in speech and text with context should be constant as sentence
-number increases. Aylett \& Turk (2004) proposed a similar principle for
-the phonological level.
+UID has been applied broadly over the past decade, to determining
+whether linguistic alignment takes place (Jaeger \& Snider, 2013),
+Zipfian word length distributions (Piantadosi et al., 2011),
+communication efficiency (Mahowald et al., 2013), dialogue and
+turn-taking (Xu \& Reitter, 2018) and the significance of ambiguity in
+language (Piantadosi et al., 2012), among other research. However, other
+recent work has contradicted the UID hypothesis. Zhan and Levy (2018)
+study Mandarin Chinese classifier use for specific and general Mandarin
+classifiers. Specific classifiers are specific to certain nouns, and UID
+would predict that specific classifiers would appear when the production
+of the corresponding noun is easier, to spread the information contained
+by the classifier and noun less densely over time. General classifiers
+can be applied to any noun and contain no special information content.
+Zhan and Levy find that speakers use specific classifiers more often in
+cases where the production of the corresponding noun is more difficult
+than when the production is easier; speakers do not avoid unexpected
+peaks in information content for their listeners, but instead maximize
+their ease of production.
 
-Jaeger (2010) extends the CER principle to all levels of human
-communication: the \emph{Uniform Information Density} (UID) hypothesis.
-Jaeger argues that speakers will try to distribute information more
-evenly over the course of an utterance, to be as close to channel
-capacity (in a Shannon sense) as possible. In particular, Jaeger singles
-out the production of ``that'' at the beginning of relative clauses in
-English and the use of contractions such as ``he's'' and ``you're''
-versus ``he is'' and ``you are'' as instances where higher information
-density in the clause elicits the production of more material by
-speakers.
+Similar to the original Levy \& Jaeger (2007), Zhan and Levy (2018)
+focuses on information distribution at particular points in a sentence.
+By contrast, Jain et al. (2018) examine word order across spoken
+sentences in Hindi, a freer word order language than English, and find
+that information density has no significant effect on word order. For
+this paper, we focus on the sentence level instead of specific points,
+following a method described in Yu et al. (2016).
 
-More recent work has stood out in contrast to the traditional UID
-perspective. Zhan and Levy (2018) study Mandarin Chinese classifier use,
-and find that the use of specific classifiers, versus general
-classifiers, appears more often in cases where the production of the
-corresponding noun is more difficult than when the production is easier,
-as would be predicted by UID. The UID perspective is challenged in Yu et
-al. (2016), by performing an analysis of entropy by position within
-sentence in the text portion of the British National Corpus. This work
-serves as the basis for ours.
+Yu et al. (2016) challenges the UID hypothesis through examining an
+entropy measure across sentence positions within the text portion of the
+English-language British National Corpus (BNC). They partition the
+corpus by sentence length in number of words. For each word position
+\(X\) of sentences of length \(k\), they define \(w\) as a unique word
+occurring in position \(X\). They define \(p(w)\) as the number of times
+word \(w\) occurs in position \(X\) divided by the number of total words
+that occur in position \(X\) i.e.~the number of sentences of length
+\(k\). Then \(p(w)\) is the probability of obtaining word \(w\) by
+choosing a word at random in position \(X\) in sentences of length
+\(k\).
 
-\section{Methods}\label{methods}
+\[H(X) = \sum\limits_w p(w)\log\big(p(w)\big)\] With this measure, Yu et
+al. compute the unigram entropy at each position of sentences of each
+length within the corpus. The result of this method can be plotted for
+each utterance length as an \emph{entropy curve}, which can be visually
+compared across utterance length to observe the how the unigram entropy
+changes across absolute positions in each of the utterances. Genzel \&
+Charniak (2002) similarly examine a unigram entropy measure on
+sentences, and found that entropy at the sentence level increases
+linearly with sentence index within a corpus. UID applies this
+uniformity of entropy rate in sentences to all levels of speech, and so
+the Yu et al. method, which examines text at the word level, should find
+an affine function at the word level.
 
-To explore information transmission rate in speech, we used the CHILDES
-TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken
-adult-child conversations. We first used the Brown and Providence
-English corpora from CHILDES. The Brown corpus contains individual
-recordings of conversations between three children between 1.5 and
-6-years-old and their families in their homes. The Providence corpus
-recorded interactions between children between 1 and 3 years old and
-their parents in the home. We divided each corpus by speaker into child
-and non-child categories. We further divided the corpora by utterance
-length, so that all sentences of length \(k\) (e.g. \(6\)) were grouped
-together. Finally, within each utterance length, we computed the unigram
-entropy measure for each position. We follow Yu et al. (2016) in using
-the following formula for each word position X of sentences of fixed
-length k from the corpus, where each \(w\) is a word occurring in
-position \(X\) and \(p(w)\) is the number of times word \(w\) occurs in
-position X divided by the number of total words that occur in position
-\(X\) i.e.~the number of sentences of length \(k\).
+Let us assume that the UID perspective is correct. Then we expect that
+the entropy curve for every language should be close to an affine,
+linear function with some noise. Entropy should monotonically and
+uniformly increase at all word positions across utterances on average.
+We will find a contradiction with this claim through examining the
+average entropy at each position of a sentence for both speech and text
+data.
 
-\[H(X) = \sum\limits_w p(w)\log\big(p(w)\big)\]
+\section{Spoken data}\label{spoken-data}
 
-The result of this method can be plotted for each utterance length as an
-\emph{entropy curve}, which can be visually compared across utterance
-length to observe the how the unigram entropy changes across absolute
-positions in each of the utterances in the English CHILDES corpora. The
-entropy curves capture individual variation across positions in
-utterances of the same length. This allows us to directly observe and
-judge the amount of variation in words that appear in an individual
-position of a sentence. For speech data, which for the corpora in the
-CHILDES database consists of short and often disconnected utterances
-across hours of recordings, the unigram entropy measure is unaffected by
-context or lack thereof in utterances by the adults and children in the
-corpora. We can directly compare any two positions within utterances to
-determine the amount of uncertainty, and therefore information, on
-average contained by words within that position of utterances. We are
-applying the same approach as in Genzel and Charniak (2003), but within
-sentences instead of across sentences.
+The UID hypothesis predicts that information transmission rate will tend
+towards uniformity in both speech and text. We begin by examining speech
+to determine if we obtain the entropy curve shape we expect for spoken
+communication. We begin with the same source as Levy \& Jaeger (2007)
+and Jaeger (2010), the Switchboard corpus of adult telephone
+conversations {[}CITATION{]}. We also examine child-adult conversations
+in the CHILDES TalkBank (Brown 1973; MacWhinney 2014) corpora database
+of spoken adult-child conversations in multiple languages. Children are
+not fully developed speakers, so we want to compare the entropy curve we
+obtain by computing over the utterances in the CHILDES corpora to the
+utterances in the adult-adult Switchboard corpus.
+
+\subsection{Methods}\label{methods}
+
+Switchboard {[}To be put in{]}
+
+We used the Brown and Providence English corpora from CHILDES. The Brown
+corpus contains individual recordings of conversations between three
+children between 1.5 and 6-years-old and their families in their homes.
+The Providence corpus recorded interactions between children between 1
+and 3 years old and their parents in the home. We divided each corpus by
+speaker into child and non-child categories. We further divided the
+corpora by utterance length, so that all sentences of length \(k\) (e.g.
+\(6\)) were grouped together. Finally, within each utterance length, we
+computed the unigram entropy measure for each position.
+
+In the English CHILDES corpora. The entropy curves capture individual
+variation across positions in utterances of the same length. This allows
+us to directly observe and judge the amount of variation in words that
+appear in an individual position of a sentence. For speech data, which
+for the corpora in the CHILDES database consists of short and often
+disconnected utterances across hours of recordings, the unigram entropy
+measure is unaffected by context or lack thereof in utterances by the
+adults and children in the corpora. We can directly compare any two
+positions within utterances to determine the amount of uncertainty, and
+therefore information, on average contained by words within that
+position of utterances. We are applying the same approach as in Genzel
+and Charniak (2002), but within sentences instead of across sentences.
 
@@
 
-We also ran this analysis on Spanish, German, Mandarin and Japanese
-corpora from CHILDES. We used the Shiro corpus for Spanish (Shiro,
-2000), which contains prompted narratives individually collected from
-over a hundred Venezualan schoolchildren, half from high SES backgrounds
-and half from low SES backgrounds. The German corpus we used was the
-Wagner corpus of spontaneous parent-child speech in the home, collected
-in single sessions for children between 1 and 14 years of age (Wagner,
-1985). We used the Zhou dinner corpus for Mandarin Chinese (Li \& Zhou,
-2015), which contains dinner conversations between 5 to 6-year-old
-children and their parents collected in Shanghai. We used the Okayama
-corpus for Japanese (Okayama, 1973), which consists of conversations
-between 2 to 4-year-old children and their mothers in single sessions.
+We also ran this analysis on Spanish and Mandarin corpora from CHILDES.
+We used the Shiro corpus for Spanish (Shiro, 2000), which contains
+prompted narratives individually collected from over a hundred
+Venezualan schoolchildren, half from high SES backgrounds and half from
+low SES backgrounds. We used the Zhou dinner corpus for Mandarin Chinese
+(Li \& Zhou, 2015), which contains dinner conversations between 5 to
+6-year-old children and their parents collected in Shanghai.
 
@@
 word boundaries, and for Japanese we used romanji (Roman alphabet)
-transliterations of words in the corpus. Japanese Hiragana, Katakana and
-Kanji writing systems, and the Chinese characters used for writing
-Mandarin do not normally demarcate word boundaries by spacing words
-apart, and for normal Japanese and Chinese writing including spaces
-between word boundaries can have a negative effect on reading times
-(Sainio et al, 2007; Bai et al, 2008).
+transliterations of words in the corpus. The Chinese characters used for
+writing Mandarin do not normally demarcate word boundaries by spacing
+words apart, and for normal Chinese writing including spaces between
+word boundaries can have a negative effect on reading times (Bai et al,
+2008).
 
@@
 
-{\centering \includegraphics{figs/wagner_PE-1}
-
-}
-
-\caption[Wagner corpus unigram entropy]{Wagner corpus unigram entropy}\label{fig:wagner_PE}
-\end{figure*}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{figure*}[h]
-
 {\centering \includegraphics{figs/zhou_PE-1}
@@
 
-\begin{CodeChunk}
-\begin{figure*}[h]
+\subsection{Results \& Analysis}\label{results-analysis}
 
-{\centering \includegraphics{figs/okayama_PE-1}
-
-}
-
-\caption[Okayama corpus unigram entropy]{Okayama corpus unigram entropy}\label{fig:okayama_PE}
-\end{figure*}
-\end{CodeChunk}
-
-\section{Results \& Analysis}\label{results-analysis}
-
-The Unigram Information Density account would predict that children have
-a lower channel capacity for communication and would instead communicate
-with a less unigram entropy curve. The adult and child entropy curves
-track one another almost identically. This not only indicates a
+The adult and child entropy curves track one another almost identically.
+This is surprising because UID would predict that the young children in
+the corpora who are not fully developed speakers would have more noise
+in their distribution of information rates. This not only indicates a
 robustness in the unigram entropy curve across speakers, but also across
@@
 
-We found a distinct three-step distribution for English, Spanish and
-German CHILDES corpora, with a slight dip in the penultimate position of
-each sentence. We attribute the penultimate dip in these unigram entropy
-curves to the fact that most of the utterances in the CHILDES English,
-Spanish and German corpora we examined had a determiner such as ``the''
-or ``a'' in the second-to-last position of utterances. The beginnings of
-utterances in the English, German and Spanish CHILDES data were usually
-pronouns or grammatical subjects, while the final words were grammatical
-objects and had a great deal of variation in the exact word that
-appeared in the utterance-final position. The final position of
-utterances in child-directed speech is known to be important, dating
-back to Aslin (1993). The Mandarin and Japanese data, by comparison,
-have noticeably lower positional entropy values in utterance-final
-positions than in utterance-penultimate positions. The Japanese entropy
-curve begins high and then drops, unlike all of the other entropy
-curves.
+We found a distinct three-step distribution for English and Spanish
+CHILDES corpora, with a slight dip in the penultimate position of each
+sentence. The final position of utterances in child-directed speech is
+known to be important, dating back to Aslin (1993). The Mandarin corpus
+entropy curve, by comparison, has a noticeably lower positional entropy
+values in utterance-final positions than in utterance-penultimate
+positions.
 
-\begin{CodeChunk}
-\begin{figure*}[h]
+We attribute the penultimate dip in the English and Spanish entropy
+curves to the fact that most of the utterances in the CHILDES English
+and Spanish corpora we examined had a determiner such as ``the'' or
+``a'' in the second-to-last position of utterances. The beginnings of
+utterances in the English and Spanish CHILDES data were usually pronouns
+or grammatical subjects, while the final words were grammatical objects
+and had a great deal of variation in the exact word that appeared in the
+utterance-final position.
 
-{\centering \includegraphics{figs/MI_wagner-1}
+These entropy curves are not what would be expected from UID. The
+robustness of the three-step distribution for English and its
+replication in Spanish do not resemble the affine function we would
+expect from UID. The Mandarin entropy curve, which does not at all
+resemble either the English/Spanish distribution or the predictions of
+UID, suggests that the entropy curve can vary from language to language.
+UID predicts that each language should have a similar distribution.
 
-}
+\section{Written data}\label{written-data}
 
-\caption[Mutual Information measures on utterances of length 6 in the Wagner corpus]{Mutual Information measures on utterances of length 6 in the Wagner corpus}\label{fig:MI_wagner}
-\end{figure*}
-\end{CodeChunk}
+The UID hypothesis also applies to written communication: we expect
+people to communicate information at a uniform rate through writing as
+well. We use Wikipedia as a source for written data, which provides two
+advantages. One, the quantity of data in Wikipedia is large for each
+language and two, there are hundreds of languages with Wikipedias. This
+allows us to perform the entropy analysis on a much greater scale and to
+directly compare the results of the entropy analysis on each language to
+one another, and ultimately to predict what typological features of a
+language help determine its entropy curve, if any. We will describe our
+method of harvesting and distilling text data from each Wikipedia as
+well as how we compare the entropy curves from each language to one
+another.
 
-\begin{CodeChunk}
-\begin{figure*}[h]
+\subsection{Methods}\label{methods-1}
 
-{\centering \includegraphics{figs/MI_okayama-1}
-
-}
-
-\caption[Mutual Information measures on utterances of length 6 in the Okayama corpus]{Mutual Information measures on utterances of length 6 in the Okayama corpus}\label{fig:MI_okayama}
-\end{figure*}
-\end{CodeChunk}
-
-The mutual information statistic is a symmetric measure of the
-information gained about one random variable by learning about another,
-a bigram measure and indicator of the effect than the context of a word
-can have on its unigram entropy value. As seen in the images below, the
-final word in utterances of length \(6\) in German has a significant
-mutual information value with all other words in the utterance on
-average. This means that the words appearing in positions \(1\) through
-\(5\) all play a role in determining the word that appears in position
-\(6\) and vice versa. By contrast, the words that appear in positions
-\(1\) through \(4\) have little influence in determining the word that
-appears in position \(5\), the penultimate word that is often a
-determiner or function word, and instead the final word has the greatest
-influence in determining the penultimate word. For Japanese, the final
-word in an utterance has little to no influence on determining the rest
-of the words and vice versa, whereas the other words all play in role in
-determining what will go in each other's positions.
-
-\section{Large-scale text data}\label{large-scale-text-data}
-
-We also wanted to apply this analysis to large-scale text data, to
-compare with the parent-child speech data results from using CHILDES.
 Using Giuseppe Attardi's Wikiextractor tool
@@
 
-\subsection{Clustering}\label{clustering}
+\subsection{Results and Analysis}\label{results-and-analysis}
 
@@
 
-\begin{CodeChunk}
-\begin{figure*}[h]
-
-{\centering \includegraphics{figs/absolute_clusters-1}
-
-}
-
-\caption[Absolute clustering of Wikipedia corpora]{Absolute clustering of Wikipedia corpora}\label{fig:absolute_clusters}
-\end{figure*}
-\end{CodeChunk}
-
-\subsection{Language Features}\label{language-features}
-
 To determine which phonological, morphological and syntactic features
@@
 missing-value imputation to obtain the features not coded in WALS for
-the languages from Wikipedia. We then performed k-modes clustering and
-compared the results with the hierarchical clustering from Wikipedia.
-{[}Still working on this{]}
+the languages from Wikipedia.
 
-We also checked the effects of individual features on the embeddings of
-languages in the different treatments. {[}Still working on this{]}
+One approach is to cluster the languages used in the Wikipedia analysis
+on the basis of WALS features and then directly compare the WALS
+clustering results with the results of the hierarchical clustering on
+the Wikipedia slope data for each of the different treatments using
+clustering similarity evaluation methods such as the Rand index (Rand,
+1971). However, the problem then arises of which combination of WALS
+features and how many features to include in the clustering analysis.
+This is a computationally intractable operation. Additionallly, deciding
+how many clusters to use in an unsupervised clustering analysis is an
+unsolved problem in machine learning.
+
+We instead check the effects of individual features on the embeddings of
+languages in the different treatments. We computed pairwise cosine
+similarity between each pair of language vectors within a treatment. For
+a subset of WALS features which had values entered for most of the
+languages we obtained from Wikipedia, we used a generalized linear model
+to see whether the cosine similarity between languages mattered in
+predicting if the languages shared the same value for a WALS feature. We
+found this to be true.
+
+\begin{CodeChunk}
+
+  X  feature   term       estimate   std.error    statistic     p.value
+---  --------  -------  ----------  ----------  -----------  ----------
+  1  83A       cosine    1.8386502   0.0103168   178.219026   0.0000000
+  2  95A       cosine    1.8971831   0.0110958   170.981718   0.0000000
+  3  81A       cosine    1.5149072   0.0098565   153.696819   0.0000000
+  4  97A       cosine    1.5767216   0.0115245   136.814432   0.0000000
+  5  144A      cosine    0.8782219   0.0117991    74.431427   0.0000000
+  6  138A      cosine    0.3951762   0.0085684    46.120042   0.0000000
+  7  87A       cosine    0.3334803   0.0083557    39.910595   0.0000000
+  8  143A      cosine    0.3736337   0.0094014    39.742361   0.0000000
+  9  82A       cosine    0.0298738   0.0092598     3.226171   0.0012546
+
+\end{CodeChunk}
 

### 2019-01-25 cogsci/position-info.Rmd
cogsci

@@
 library(xtable)
-library(knitr)
 ```
@@
 We begin by examining speech to determine if we obtain the entropy curve shape we expect for spoken communication.
-We begin with the same source as Levy & Jaeger (2007) and Jaeger (2010), the Switchboard corpus of adult telephone conversations [CITATION].
+We begin with the same source as Levy & Jaeger (2007) and Jaeger (2010), the Switchboard corpus of adult telephone conversations.
 We also examine child-adult conversations in the CHILDES TalkBank (Brown 1973; MacWhinney 2014) corpora database of spoken adult-child conversations in multiple languages.
@@
 Each language corpus was cleaned and limited to sentences between $6$ and $50$ words.
-Similar to the data from CHILDES, we divided each corpus by sentence length, and then computed the unigram entropy measure on each word position within each sentence length.
-
-To classify the unigram entropy curves of the different distributions, we computed three slope treatments of each curve.
+Similar to our process for the Switchboard and CHILDES spoken corpora, we divided each corpus by sentence length, and then computed the unigram entropy measure on each word position within each sentence length.
+We wanted to directly compare and classify the unigram entropy curves of the languages from Wikipedia.
+We computed three slope treatments of each curve.
 In the *absolute* treatment, with sentence length denoted as $n$, we computed the slope between positions $1$ and $2$, positions $2$ and $3$, positions $3$ and $n-2$, positions $n-2$ and $n-1$ and positions $n-1$ and $n$.
@@
 
-For the German and Japanese entropy curves, we observed the same shape for the text Wikipedia corpora entropy curves and the speech CHILDES corpora entropy curves, supporting the argument for robustness across text and speech.
-
 ## Results and Analysis
 
-In the *absolute* and *relative 5* treatments, each language is cast as a point in $5$-dimensional space.
-In the *relative 10* treatment, each language is cast as a point in $10$-dimensional space.
-This allows for direct comparison using cosine similarity and moreover for unsupervised clustering analysis to compare the results of the Wikipedia entropy analysis to known typological features within the langauges in the Wikipedia dataset.
-Using the R hclust package, we performed a hierarchical clustering of the Wikipedia slope data within each treatment.
-A subset of the results at a glance for the *absolute* treatment are below.
-
+In the *absolute* and *relative 5* treatments, each language is embedded in $5$-dimensional space.
+In the *relative 10* treatment, each language is embedded in $10$-dimensional space.
+This allows for direct comparison between languages on Wikipedia using cosine similarity and moreover for unsupervised clustering analysis to compare the results of the Wikipedia entropy analysis to known typological features within the langauges in the Wikipedia dataset.
 To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we looked at the list of $144$ linguistic features in World Atlas of Language Structures (Dryer & Haspelmath, 2013).
@@
 For a subset of WALS features which had values entered for most of the languages we obtained from Wikipedia, we used a generalized linear model to see whether the cosine similarity between languages mattered in predicting if the languages shared the same value for a WALS feature.
-We found this to be true.
+For eight features and $45$ languages, we found this to be true.
 
@@
 glm_model <- read.csv("glm.csv")
-knitr::kable(glm_model)
+tab1 <- xtable::xtable(glm_model)
+print(tab1, type="latex", comment = F, table.placement = "H")
 ```
@@
 # Conclusions 
-In this paper we have explored the differences in rate of information transmission among spoken child-adult conversational corpora in several languages and text corpora pulled from Wikipedia for well over a hundred other corpora.
-We have seen that the entropy curve, a representation of information at each position of an utterance without regard to context, varies from language to language and is predicted by a subset of syntactic, phonological and morphological features in a language.
-We have provided evidence against the Uniform Information Density hypothesis and provided an alternate explanation for the rate at which information is encoded and transmitted in human communication, namely one that is robust across corpora, language dependent and arises from the features of the language.
+In this paper, we have characterized and applied a model based on Genzel & Charniak (2002) for distinguishing In doing so, we have provided evidence against the Uniform Information Density hypothesis.
+Our work indicates that information is not distributed uniformly throughout an utterance regardless of languages, but that languages possess a characteristic information distribution determined by their syntactic, morphological and phonological features.
 
-A large volume of research has indicated the effects of surprisal on fixation duration during eye-tracking studies: higher surprisal of a word corresponds to higher fixation duration.
+Our work has focused on words within sentences in both speech and text across typologically diverse languages.
+Other researchers have explored information rate cross-linguistically as a single number characterizing each language, most notably including Aylett & Turk (2004).
+That work characterizes information rate as information per unit time, and is complementary to the work we accomplish here.
+
+A large volume of research has indicated the effects of surprisal on fixation duration during eye-tracking studies (Demberg & Keller, 2008) (Boston et al., 2008) (Smith & Lev, 2013).
+Higher surprisal of a word corresponds to higher fixation duration.
 The *wrap-up effect* is a popular hypothesis within the eye-tracking community.
@@
 # The following code is some latex to format the bibliography. Do not remove it.
-
 ```

### 2019-01-25 cogsci/position-info.tex
cogsci

@@
 \begin{CodeChunk}
-
-  X  feature   term       estimate   std.error    statistic     p.value
----  --------  -------  ----------  ----------  -----------  ----------
-  1  83A       cosine    1.8386502   0.0103168   178.219026   0.0000000
-  2  95A       cosine    1.8971831   0.0110958   170.981718   0.0000000
-  3  81A       cosine    1.5149072   0.0098565   153.696819   0.0000000
-  4  97A       cosine    1.5767216   0.0115245   136.814432   0.0000000
-  5  144A      cosine    0.8782219   0.0117991    74.431427   0.0000000
-  6  138A      cosine    0.3951762   0.0085684    46.120042   0.0000000
-  7  87A       cosine    0.3334803   0.0083557    39.910595   0.0000000
-  8  143A      cosine    0.3736337   0.0094014    39.742361   0.0000000
-  9  82A       cosine    0.0298738   0.0092598     3.226171   0.0012546
-
+\begin{CodeOutput}
+\begin{table}[H]
+\centering
+\begin{tabular}{rrllrrrr}
+  \hline
+ & X & feature & term & estimate & std.error & statistic & p.value \\ 
+  \hline
+1 &   1 & 83A & cosine & 1.84 & 0.01 & 178.22 & 0.00 \\ 
+  2 &   2 & 95A & cosine & 1.90 & 0.01 & 170.98 & 0.00 \\ 
+  3 &   3 & 81A & cosine & 1.51 & 0.01 & 153.70 & 0.00 \\ 
+  4 &   4 & 97A & cosine & 1.58 & 0.01 & 136.81 & 0.00 \\ 
+  5 &   5 & 144A & cosine & 0.88 & 0.01 & 74.43 & 0.00 \\ 
+  6 &   6 & 138A & cosine & 0.40 & 0.01 & 46.12 & 0.00 \\ 
+  7 &   7 & 87A & cosine & 0.33 & 0.01 & 39.91 & 0.00 \\ 
+  8 &   8 & 143A & cosine & 0.37 & 0.01 & 39.74 & 0.00 \\ 
+  9 &   9 & 82A & cosine & 0.03 & 0.01 & 3.23 & 0.00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+\end{CodeOutput}
 \end{CodeChunk}

### 2019-01-26 cogsci/position-info.Rmd
second cogsci draft ready

@@
 
+The entropy curve is a consequence of the information-theoretic rate of information transfer in a language, where information is defined as a reduction in uncertainty.
+Thus the more entropy at a word position in a sentence (Yu et al., 2016) or a sentence index in a text (Genzel & Charniak, 2002), the more information is on average contained around that position.
 Let us assume that the UID perspective is correct.
@@
 We will find a contradiction with this claim through examining the average entropy at each position of a sentence for both speech and text data.
+
+[Put in an illustration of what UID predicts?]
 
@@
 
+[Put in an illustration of the slope sections and treatments here?]
+
 ## Results and Analysis
@@
 For eight features and $45$ languages, we found this to be true.
+The table below shows the results for the generalized linear model we computed using WALS features and the absolute treatment.
 
@@
 
+Cosine distance played some role in the feature determination.
+The top features were $83$, $81$, $95$ and $97$, which all characterize word order.
+This makes intuitive sense, and indicates that for this small subset of features and subset of languages that the embeddings in the slope space are related to the WALS features.
+Therefore typological features play a role in determining the positional entropy values for a language.
+This means that the entropy curves are in some way structured by the syntactic, morphological and phonological features of a language.
+
 # Conclusions 
-In this paper, we have characterized and applied a model based on Genzel & Charniak (2002) for distinguishing In doing so, we have provided evidence against the Uniform Information Density hypothesis.
+In this paper, we have characterized and applied a model based on Genzel & Charniak (2002) for distinguishing entropy at each word position within sentences.
+In doing so, we have provided evidence against the Uniform Information Density hypothesis.
 Our work indicates that information is not distributed uniformly throughout an utterance regardless of languages, but that languages possess a characteristic information distribution determined by their syntactic, morphological and phonological features.
@@
 That work characterizes information rate as information per unit time, and is complementary to the work we accomplish here.
+Languages may possess both an overall information transfer rate, showing the average amount of semantic information transmitted per syllable or word, and an entropy distribution indicative of the language's information distribution.
 
-A large volume of research has indicated the effects of surprisal on fixation duration during eye-tracking studies (Demberg & Keller, 2008) (Boston et al., 2008) (Smith & Lev, 2013).
-Higher surprisal of a word corresponds to higher fixation duration.
-The *wrap-up effect* is a popular hypothesis within the eye-tracking community.
-The hypothesis states that, in written text, sentence-final words are processed more slowly on average then sentence-medial or sentence-initial words, due to readers integrating information from the entire sentence in order to form a final, coherent thought expressed by the sentence, among other reasons.
+Research has indicated the effects of surprisal on fixation duration during eye-tracking studies (Demberg & Keller, 2008) (Boston et al., 2008) (Smith & Levy, 2013).
+Higher surprisal of a word is on average predictive of higher fixation duration.
+The *wrap-up effect* states that when a person reads written text, he or she will process sentence-final words more slowly on average then sentence-medial or sentence-initial words, due to integrating information from the entire sentence to form a final understanding of the sentence's meaning (Stowe et al., 2018) (Kuperman et al., 2010).
 The wrap-up effect is drawn from evidence in English, German and Dutch, all of which are languages with a large final increase in their entropy curve from our study.
-We aim to see in the future if as a consquence of our work in this paper, the wrap-up effect falls out from the final jump in the entropy curves in some languages.
+We aim to see in the future if as a consequence of our work in this paper, the wrap-up effect falls out from the final jump in the entropy curves in some languages.
+We would analyze eye-tracking data from languages with low positional entropy in sentence-final positions.
+Since the lexical surprisal metric in eye-tracking and our positional entropy metric are both measures of log-probability surprisal, then in languages such as Hindi where the sentence-final positional entropy values are low we expect to find no wrap-up effect.
+
+# Acknowledgements
+
+[Not here yet.]
+
+# References
 

### 2019-01-26 cogsci/position-info.tex
second cogsci draft ready

@@
 
-Let us assume that the UID perspective is correct. Then we expect that
-the entropy curve for every language should be close to an affine,
-linear function with some noise. Entropy should monotonically and
-uniformly increase at all word positions across utterances on average.
-We will find a contradiction with this claim through examining the
-average entropy at each position of a sentence for both speech and text
-data.
+The entropy curve is a consequence of the information-theoretic rate of
+information transfer in a language, where information is defined as a
+reduction in uncertainty. Thus the more entropy at a word position in a
+sentence (Yu et al., 2016) or a sentence index in a text (Genzel \&
+Charniak, 2002), the more information is on average contained around
+that position. Let us assume that the UID perspective is correct. Then
+we expect that the entropy curve for every language should be close to
+an affine, linear function with some noise. Entropy should monotonically
+and uniformly increase at all word positions across utterances on
+average. We will find a contradiction with this claim through examining
+the average entropy at each position of a sentence for both speech and
+text data.
+
+{[}Put in an illustration of what UID predicts?{]}
 
@@
 and Jaeger (2010), the Switchboard corpus of adult telephone
-conversations {[}CITATION{]}. We also examine child-adult conversations
-in the CHILDES TalkBank (Brown 1973; MacWhinney 2014) corpora database
-of spoken adult-child conversations in multiple languages. Children are
-not fully developed speakers, so we want to compare the entropy curve we
-obtain by computing over the utterances in the CHILDES corpora to the
-utterances in the adult-adult Switchboard corpus.
+conversations. We also examine child-adult conversations in the CHILDES
+TalkBank (Brown 1973; MacWhinney 2014) corpora database of spoken
+adult-child conversations in multiple languages. Children are not fully
+developed speakers, so we want to compare the entropy curve we obtain by
+computing over the utterances in the CHILDES corpora to the utterances
+in the adult-adult Switchboard corpus.
 
@@
 language corpus was cleaned and limited to sentences between \(6\) and
-\(50\) words. Similar to the data from CHILDES, we divided each corpus
-by sentence length, and then computed the unigram entropy measure on
-each word position within each sentence length.
-
-To classify the unigram entropy curves of the different distributions,
-we computed three slope treatments of each curve. In the \emph{absolute}
-treatment, with sentence length denoted as \(n\), we computed the slope
-between positions \(1\) and \(2\), positions \(2\) and \(3\), positions
-\(3\) and \(n-2\), positions \(n-2\) and \(n-1\) and positions \(n-1\)
-and \(n\). For the short utterances appearing the CHILDES speech corpora
-we examined, these appeared to be important junctions in the
-distributions, with a seeming plateau in the middle of the unigram
-entropy curve for each of the language corpora we examined in CHILDES.
+\(50\) words. Similar to our process for the Switchboard and CHILDES
+spoken corpora, we divided each corpus by sentence length, and then
+computed the unigram entropy measure on each word position within each
+sentence length. We wanted to directly compare and classify the unigram
+entropy curves of the languages from Wikipedia. We computed three slope
+treatments of each curve. In the \emph{absolute} treatment, with
+sentence length denoted as \(n\), we computed the slope between
+positions \(1\) and \(2\), positions \(2\) and \(3\), positions \(3\)
+and \(n-2\), positions \(n-2\) and \(n-1\) and positions \(n-1\) and
+\(n\). For the short utterances appearing the CHILDES speech corpora we
+examined, these appeared to be important junctions in the distributions,
+with a seeming plateau in the middle of the unigram entropy curve for
+each of the language corpora we examined in CHILDES.
 
@@
 
-For the German and Japanese entropy curves, we observed the same shape
-for the text Wikipedia corpora entropy curves and the speech CHILDES
-corpora entropy curves, supporting the argument for robustness across
-text and speech.
+{[}Put in an illustration of the slope sections and treatments here?{]}
 
@@
 In the \emph{absolute} and \emph{relative 5} treatments, each language
-is cast as a point in \(5\)-dimensional space. In the \emph{relative 10}
-treatment, each language is cast as a point in \(10\)-dimensional space.
-This allows for direct comparison using cosine similarity and moreover
-for unsupervised clustering analysis to compare the results of the
-Wikipedia entropy analysis to known typological features within the
-langauges in the Wikipedia dataset. Using the R hclust package, we
-performed a hierarchical clustering of the Wikipedia slope data within
-each treatment. A subset of the results at a glance for the
-\emph{absolute} treatment are below.
-
-To determine which phonological, morphological and syntactic features
-affected the embedding of a language in the Wikipedia dataset, we looked
-at the list of \(144\) linguistic features in World Atlas of Language
-Structures (Dryer \& Haspelmath, 2013). We limited the languages in the
-WALS database to only those in our Wikipedia dataset and performed
+is embedded in \(5\)-dimensional space. In the \emph{relative 10}
+treatment, each language is embedded in \(10\)-dimensional space. This
+allows for direct comparison between languages on Wikipedia using cosine
+similarity and moreover for unsupervised clustering analysis to compare
+the results of the Wikipedia entropy analysis to known typological
+features within the langauges in the Wikipedia dataset. To determine
+which phonological, morphological and syntactic features affected the
+embedding of a language in the Wikipedia dataset, we looked at the list
+of \(144\) linguistic features in World Atlas of Language Structures
+(Dryer \& Haspelmath, 2013). We limited the languages in the WALS
+database to only those in our Wikipedia dataset and performed
 missing-value imputation to obtain the features not coded in WALS for
@@
 to see whether the cosine similarity between languages mattered in
-predicting if the languages shared the same value for a WALS feature. We
-found this to be true.
+predicting if the languages shared the same value for a WALS feature.
+For eight features and \(45\) languages, we found this to be true. The
+table below shows the results for the generalized linear model we
+computed using WALS features and the absolute treatment.
 
@@
 
+Cosine distance played some role in the feature determination. The top
+features were \(83\), \(81\), \(95\) and \(97\), which all characterize
+word order. This makes intuitive sense, and indicates that for this
+small subset of features and subset of languages that the embeddings in
+the slope space are related to the WALS features. Therefore typological
+features play a role in determining the positional entropy values for a
+language. This means that the entropy curves are in some way structured
+by the syntactic, morphological and phonological features of a language.
+
 \section{Conclusions}\label{conclusions}
 
-In this paper we have explored the differences in rate of information
-transmission among spoken child-adult conversational corpora in several
-languages and text corpora pulled from Wikipedia for well over a hundred
-other corpora. We have seen that the entropy curve, a representation of
-information at each position of an utterance without regard to context,
-varies from language to language and is predicted by a subset of
-syntactic, phonological and morphological features in a language. We
-have provided evidence against the Uniform Information Density
-hypothesis and provided an alternate explanation for the rate at which
-information is encoded and transmitted in human communication, namely
-one that is robust across corpora, language dependent and arises from
-the features of the language.
+In this paper, we have characterized and applied a model based on Genzel
+\& Charniak (2002) for distinguishing entropy at each word position
+within sentences. In doing so, we have provided evidence against the
+Uniform Information Density hypothesis. Our work indicates that
+information is not distributed uniformly throughout an utterance
+regardless of languages, but that languages possess a characteristic
+information distribution determined by their syntactic, morphological
+and phonological features.
 
-A large volume of research has indicated the effects of surprisal on
-fixation duration during eye-tracking studies: higher surprisal of a
-word corresponds to higher fixation duration. The \emph{wrap-up effect}
-is a popular hypothesis within the eye-tracking community. The
-hypothesis states that, in written text, sentence-final words are
-processed more slowly on average then sentence-medial or
-sentence-initial words, due to readers integrating information from the
-entire sentence in order to form a final, coherent thought expressed by
-the sentence, among other reasons. The wrap-up effect is drawn from
+Our work has focused on words within sentences in both speech and text
+across typologically diverse languages. Other researchers have explored
+information rate cross-linguistically as a single number characterizing
+each language, most notably including Aylett \& Turk (2004). That work
+characterizes information rate as information per unit time, and is
+complementary to the work we accomplish here. Languages may possess both
+an overall information transfer rate, showing the average amount of
+semantic information transmitted per syllable or word, and an entropy
+distribution indicative of the language's information distribution.
+
+Research has indicated the effects of surprisal on fixation duration
+during eye-tracking studies (Demberg \& Keller, 2008) (Boston et al.,
+2008) (Smith \& Levy, 2013). Higher surprisal of a word is on average
+predictive of higher fixation duration. The \emph{wrap-up effect} states
+that when a person reads written text, he or she will process
+sentence-final words more slowly on average then sentence-medial or
+sentence-initial words, due to integrating information from the entire
+sentence to form a final understanding of the sentence's meaning (Stowe
+et al., 2018) (Kuperman et al., 2010). The wrap-up effect is drawn from
 evidence in English, German and Dutch, all of which are languages with a
 large final increase in their entropy curve from our study. We aim to
-see in the future if as a consquence of our work in this paper, the
+see in the future if as a consequence of our work in this paper, the
 wrap-up effect falls out from the final jump in the entropy curves in
-some languages.
+some languages. We would analyze eye-tracking data from languages with
+low positional entropy in sentence-final positions. Since the lexical
+surprisal metric in eye-tracking and our positional entropy metric are
+both measures of log-probability surprisal, then in languages such as
+Hindi where the sentence-final positional entropy values are low we
+expect to find no wrap-up effect.
+
+\section{Acknowledgements}\label{acknowledgements}
+
+{[}Not here yet.{]}
+
+\section{References}\label{references}
 

### 2019-01-26 cogsci/position-info.Rmd
updates to Rmd

@@
 author-information: > 
-    \author{{\large \bf Josef Klafka} \\ \texttt{jklafka@uchicago.edu} \\ Department of Psychology \\ University of Chicago
-    \And {\large \bf Daniel Yurovsky} \\ \texttt{yurovsky@uchicago.edu} \\ Department of Psychology \\ University of Chicago}
+ \author{Josef Klafka \and Daniel Yurovsky \\
+         \texttt{\{jklafka, yurovsky\}@uchicago.edu} \\
+        Department of Psychology \\ University of Chicago}
 
-abstract: 
-    "We provide evidence against the popular Uniform Information Density hypothesis (Levy & Jaeger, 2007) which proposes that information is transmitted at a constant rate close to channel capacity in human communication.
+abstract: >
+    We provide evidence against the popular Uniform Information Density hypothesis (Levy & Jaeger, 2007) which proposes that information is transmitted at a constant rate close to channel capacity in human communication.
 Using a method based on the original Genzel & Charniak (2002) entropy measure, we construct a word-level model for entropy applicable to both spoken and written corpora.
@@
 We find that not only is the Uniform Information Density hypothesis prediction wrong, but also that the by-word entropy distribution of a language is related to the typological features of the language.
-We use this evidence to suggest that people do not communicate information at a uniform rate, but that information distribution varies from language to language based on phonogical, morphological and syntactic features."
+We use this evidence to suggest that people do not communicate information at a uniform rate, but that information distribution varies from language to language based on phonogical, morphological and syntactic features.
 
-keywords:
-    "Entropy; information; information theory; communication"
+keywords: >
+    Entropy; information; information theory; communication
 
 output: cogsci2016::cogsci_paper
+#final-submission: \cogscifinalcopy
 ---
@@
 library(grid)
-library(ggplot2)
+library(tidyverse)
 library(xtable)
@@
 
-# Background
-
 Perhaps the most important reason people talk and write to one another is to give information.
-From Shannon (1948), we know that when information is being communicated, the rate of information transmission should be constant and as close as possible to the capacity of the communication channel to optimally transmit the information.
-Levy & Jaeger (2007) introduce the *uniform information density* (UID) hypothesis, arguing that people unconsciously structure their speech and writing to communicate information at a constant, optimal rate.
-Levy & Jaeger (2007) argue that speakers and writers should attempt to communicate information as evenly as possible and as close to the channel capacity as possible.
+From @shannon1948, we know that when information is being communicated, the rate of information transmission should be constant and as close as possible to the capacity of the communication channel to optimally transmit the information.
+@levy2007 introduce the *uniform information density* (UID) hypothesis, arguing that people unconsciously structure their speech and writing to communicate information at a constant, optimal rate.
+@levy2007 argue that speakers and writers should attempt to communicate information as evenly as possible and as close to the channel capacity as possible.
 Their evidence concerns the use or lack of use of “that”-complementizers in introducing English relative clauses.
@@
 
-UID has been applied broadly over the past decade, to determining whether linguistic alignment takes place (Jaeger & Snider, 2013), Zipfian word length distributions (Piantadosi et al., 2011), communication efficiency (Mahowald et al., 2013), dialogue and turn-taking (Xu & Reitter, 2018) and the significance of ambiguity in language (Piantadosi et al., 2012), among other research.
+UID has been applied broadly over the past decade, to determining whether linguistic alignment takes place [@jaeger2013], Zipfian word length distributions [@piantadosi2011], communication efficiency [@mahowald2013], dialogue and turn-taking [@xu2018] and the significance of ambiguity in language [@piantadosi2012], among other research.
 However, other recent work has contradicted the UID hypothesis.
-Zhan and Levy (2018) study Mandarin Chinese classifier use for specific and general Mandarin classifiers.
+@zhan2018 study Mandarin Chinese classifier use for specific and general Mandarin classifiers.
 Specific classifiers are specific to certain nouns, and UID would predict that specific classifiers would appear when the production of the corresponding noun is easier, to spread the information contained by the classifier and noun less densely over time.
@@
 
-Similar to the original Levy & Jaeger (2007), Zhan and Levy (2018) focuses on information distribution at particular points in a sentence.
-By contrast, Jain et al.
-(2018) examine word order across spoken sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on word order.
-For this paper, we focus on the sentence level instead of specific points, following a method described in Yu et al.
-(2016).
+Similar to the original @levy2007, @zhan2018 focuses on information distribution at particular points in a sentence.
+By contrast, @jain2018 examine word order across spoken sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on word order.
+For this paper, we focus on the sentence level instead of specific points, following a method described in @yu2016.
 
-Yu et al.
-(2016) challenges the UID hypothesis through examining an entropy measure across sentence positions within the text portion of the English-language British National Corpus (BNC).
+@yu2016 challenges the UID hypothesis through examining an entropy measure across sentence positions within the text portion of the English-language British National Corpus (BNC, CITE).
 They partition the corpus by sentence length in number of words.
@@
 $$H(X) = \sum\limits_w p(w)\log\big(p(w)\big)$$
-With this measure, Yu et al.
-compute the unigram entropy at each position of sentences of each length within the corpus.
+With this measure, @yu2016  compute the unigram entropy at each position of sentences of each length within the corpus.
 The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances.
-Genzel & Charniak (2002) similarly examine a unigram entropy measure on sentences, and found that entropy at the sentence level increases linearly with sentence index within a corpus.
-UID applies this uniformity of entropy rate in sentences to all levels of speech, and so the Yu et al.
-method, which examines text at the word level, should find an affine function at the word level.
+@genzel2002 similarly examine a unigram entropy measure on sentences, and found that entropy at the sentence level increases linearly with sentence index within a corpus.
+UID applies this uniformity of entropy rate in sentences to all levels of speech, and so the @yu2016, which examines text at the word level, should find an affine function at the word level.
 
 The entropy curve is a consequence of the information-theoretic rate of information transfer in a language, where information is defined as a reduction in uncertainty.
-Thus the more entropy at a word position in a sentence (Yu et al., 2016) or a sentence index in a text (Genzel & Charniak, 2002), the more information is on average contained around that position.
+Thus the more entropy at a word position in a sentence [@yu2016] or a sentence index in a text [@genzel2002], the more information is on average contained around that position.
 Let us assume that the UID perspective is correct.
@@
 We begin by examining speech to determine if we obtain the entropy curve shape we expect for spoken communication.
-We begin with the same source as Levy & Jaeger (2007) and Jaeger (2010), the Switchboard corpus of adult telephone conversations.
-We also examine child-adult conversations in the CHILDES TalkBank (Brown 1973; MacWhinney 2014) corpora database of spoken adult-child conversations in multiple languages.
+We begin with the same source as @levy2007 and @jaeger2010, the Switchboard corpus of adult telephone conversations.
+We also examine child-adult conversations in the CHILDES TalkBank [@brown1973; @macwhinney2014] corpora database of spoken adult-child conversations in multiple languages.
 Children are not fully developed speakers, so we want to compare the entropy curve we obtain by computing over the utterances in the CHILDES corpora to the utterances in the adult-adult Switchboard corpus.
@@
 We can directly compare any two positions within utterances to determine the amount of uncertainty, and therefore information, on average contained by words within that position of utterances.
-We are applying the same approach as in Genzel and Charniak (2002), but within sentences instead of across sentences.
+We are applying the same approach as in @genzel2002, but within sentences instead of across sentences.
 
@@
 
-```{r tables}
-glm_model <- read.csv("glm.csv")
+```{r tables, results="asis", tab.env = "table"}
+glm_model <- read_csv("glm.csv") %>%
+    select(-X1)
+
 tab1 <- xtable::xtable(glm_model)
-print(tab1, type="latex", comment = F, table.placement = "H")
+print(tab1, type="latex", comment = F, table.placement = "tb",
+      include.rownames = F)
 ```

### 2019-01-26 cogsci/position-info.tex
updates to Rmd

@@
 
-\author{{\large \bf Josef Klafka} \\ \texttt{jklafka@uchicago.edu} \\ Department of Psychology \\ University of Chicago \And {\large \bf Daniel Yurovsky} \\ \texttt{yurovsky@uchicago.edu} \\ Department of Psychology \\ University of Chicago}
+\author{Josef Klafka \and Daniel Yurovsky \\
+        \texttt{\{jklafka, yurovsky\}@uchicago.edu} \\
+       Department of Psychology \\ University of Chicago}
 
@@
 
-\section{Background}\label{background}
-
 Perhaps the most important reason people talk and write to one another
@@
 should be constant and as close as possible to the capacity of the
-communication channel to optimally transmit the information. Levy \&
-Jaeger (2007) introduce the \emph{uniform information density} (UID)
+communication channel to optimally transmit the information.
+({\textbf{???}}) introduce the \emph{uniform information density} (UID)
 hypothesis, arguing that people unconsciously structure their speech and
-writing to communicate information at a constant, optimal rate. Levy \&
-Jaeger (2007) argue that speakers and writers should attempt to
+writing to communicate information at a constant, optimal rate.
+({\textbf{???}}) argue that speakers and writers should attempt to
 communicate information as evenly as possible and as close to the
@@
 whether linguistic alignment takes place (Jaeger \& Snider, 2013),
-Zipfian word length distributions (Piantadosi et al., 2011),
-communication efficiency (Mahowald et al., 2013), dialogue and
-turn-taking (Xu \& Reitter, 2018) and the significance of ambiguity in
-language (Piantadosi et al., 2012), among other research. However, other
-recent work has contradicted the UID hypothesis. Zhan and Levy (2018)
-study Mandarin Chinese classifier use for specific and general Mandarin
-classifiers. Specific classifiers are specific to certain nouns, and UID
-would predict that specific classifiers would appear when the production
-of the corresponding noun is easier, to spread the information contained
-by the classifier and noun less densely over time. General classifiers
-can be applied to any noun and contain no special information content.
-Zhan and Levy find that speakers use specific classifiers more often in
-cases where the production of the corresponding noun is more difficult
-than when the production is easier; speakers do not avoid unexpected
-peaks in information content for their listeners, but instead maximize
-their ease of production.
+Zipfian word length distributions (Piantadosi, Tily, \& Gibson, 2011),
+communication efficiency (Mahowald, Fedorenko, Piantadosi, \& Gibson,
+2013), dialogue and turn-taking (Xu \& Reitter, 2018) and the
+significance of ambiguity in language (Piantadosi, Tily, \& Gibson,
+2012), among other research. However, other recent work has contradicted
+the UID hypothesis. Zhan \& Levy (2018) study Mandarin Chinese
+classifier use for specific and general Mandarin classifiers. Specific
+classifiers are specific to certain nouns, and UID would predict that
+specific classifiers would appear when the production of the
+corresponding noun is easier, to spread the information contained by the
+classifier and noun less densely over time. General classifiers can be
+applied to any noun and contain no special information content. Zhan and
+Levy find that speakers use specific classifiers more often in cases
+where the production of the corresponding noun is more difficult than
+when the production is easier; speakers do not avoid unexpected peaks in
+information content for their listeners, but instead maximize their ease
+of production.
 
-Similar to the original Levy \& Jaeger (2007), Zhan and Levy (2018)
-focuses on information distribution at particular points in a sentence.
-By contrast, Jain et al. (2018) examine word order across spoken
-sentences in Hindi, a freer word order language than English, and find
-that information density has no significant effect on word order. For
-this paper, we focus on the sentence level instead of specific points,
-following a method described in Yu et al. (2016).
+Similar to the original ({\textbf{???}}), Zhan \& Levy (2018) focuses on
+information distribution at particular points in a sentence. By
+contrast, Jain, Singh, Ranjan, Rajkumar, \& Agarwal (2018) examine word
+order across spoken sentences in Hindi, a freer word order language than
+English, and find that information density has no significant effect on
+word order. For this paper, we focus on the sentence level instead of
+specific points, following a method described in Yu, Cong, Liang, \& Liu
+(2016).
 
@@
 entropy measure across sentence positions within the text portion of the
-English-language British National Corpus (BNC). They partition the
+English-language British National Corpus (BNC, CITE). They partition the
 corpus by sentence length in number of words. For each word position
@@
 \[H(X) = \sum\limits_w p(w)\log\big(p(w)\big)\] With this measure, Yu et
-al. compute the unigram entropy at each position of sentences of each
-length within the corpus. The result of this method can be plotted for
-each utterance length as an \emph{entropy curve}, which can be visually
-compared across utterance length to observe the how the unigram entropy
-changes across absolute positions in each of the utterances. Genzel \&
-Charniak (2002) similarly examine a unigram entropy measure on
+al. (2016) compute the unigram entropy at each position of sentences of
+each length within the corpus. The result of this method can be plotted
+for each utterance length as an \emph{entropy curve}, which can be
+visually compared across utterance length to observe the how the unigram
+entropy changes across absolute positions in each of the utterances.
+Genzel \& Charniak (2002) similarly examine a unigram entropy measure on
 sentences, and found that entropy at the sentence level increases
@@
 uniformity of entropy rate in sentences to all levels of speech, and so
-the Yu et al. method, which examines text at the word level, should find
+the Yu et al. (2016), which examines text at the word level, should find
 an affine function at the word level.
@@
 
-\section{Spoken data}\label{spoken-data}
+\hypertarget{spoken-data}{%
+\section{Spoken data}\label{spoken-data}}
 
@@
 to determine if we obtain the entropy curve shape we expect for spoken
-communication. We begin with the same source as Levy \& Jaeger (2007)
-and Jaeger (2010), the Switchboard corpus of adult telephone
-conversations. We also examine child-adult conversations in the CHILDES
-TalkBank (Brown 1973; MacWhinney 2014) corpora database of spoken
-adult-child conversations in multiple languages. Children are not fully
-developed speakers, so we want to compare the entropy curve we obtain by
-computing over the utterances in the CHILDES corpora to the utterances
-in the adult-adult Switchboard corpus.
+communication. We begin with the same source as ({\textbf{???}}) and
+Jaeger (2010), the Switchboard corpus of adult telephone conversations.
+We also examine child-adult conversations in the CHILDES TalkBank
+(Brown, 1973; MacWhinney, 2014) corpora database of spoken adult-child
+conversations in multiple languages. Children are not fully developed
+speakers, so we want to compare the entropy curve we obtain by computing
+over the utterances in the CHILDES corpora to the utterances in the
+adult-adult Switchboard corpus.
 
-\subsection{Methods}\label{methods}
+\hypertarget{methods}{%
+\subsection{Methods}\label{methods}}
 
@@
 position of utterances. We are applying the same approach as in Genzel
-and Charniak (2002), but within sentences instead of across sentences.
+\& Charniak (2002), but within sentences instead of across sentences.
 
@@
 
-\subsection{Results \& Analysis}\label{results-analysis}
+\hypertarget{results-analysis}{%
+\subsection{Results \& Analysis}\label{results-analysis}}
 
@@
 
-\section{Written data}\label{written-data}
+\hypertarget{written-data}{%
+\section{Written data}\label{written-data}}
 
@@
 
-\subsection{Methods}\label{methods-1}
+\hypertarget{methods-1}{%
+\subsection{Methods}\label{methods-1}}
 
@@
 
-\subsection{Results and Analysis}\label{results-and-analysis}
+\hypertarget{results-and-analysis}{%
+\subsection{Results and Analysis}\label{results-and-analysis}}
 
@@
 
-\begin{CodeChunk}
-\begin{CodeOutput}
-\begin{table}[H]
+\begin{table}[tb]
 \centering
-\begin{tabular}{rrllrrrr}
+\begin{tabular}{llrrrr}
   \hline
- & X & feature & term & estimate & std.error & statistic & p.value \\ 
+feature & term & estimate & std.error & statistic & p.value \\ 
   \hline
-1 &   1 & 83A & cosine & 1.84 & 0.01 & 178.22 & 0.00 \\ 
-  2 &   2 & 95A & cosine & 1.90 & 0.01 & 170.98 & 0.00 \\ 
-  3 &   3 & 81A & cosine & 1.51 & 0.01 & 153.70 & 0.00 \\ 
-  4 &   4 & 97A & cosine & 1.58 & 0.01 & 136.81 & 0.00 \\ 
-  5 &   5 & 144A & cosine & 0.88 & 0.01 & 74.43 & 0.00 \\ 
-  6 &   6 & 138A & cosine & 0.40 & 0.01 & 46.12 & 0.00 \\ 
-  7 &   7 & 87A & cosine & 0.33 & 0.01 & 39.91 & 0.00 \\ 
-  8 &   8 & 143A & cosine & 0.37 & 0.01 & 39.74 & 0.00 \\ 
-  9 &   9 & 82A & cosine & 0.03 & 0.01 & 3.23 & 0.00 \\ 
+83A & cosine & 1.84 & 0.01 & 178.22 & 0.00 \\ 
+  95A & cosine & 1.90 & 0.01 & 170.98 & 0.00 \\ 
+  81A & cosine & 1.51 & 0.01 & 153.70 & 0.00 \\ 
+  97A & cosine & 1.58 & 0.01 & 136.81 & 0.00 \\ 
+  144A & cosine & 0.88 & 0.01 & 74.43 & 0.00 \\ 
+  138A & cosine & 0.40 & 0.01 & 46.12 & 0.00 \\ 
+  87A & cosine & 0.33 & 0.01 & 39.91 & 0.00 \\ 
+  143A & cosine & 0.37 & 0.01 & 39.74 & 0.00 \\ 
+  82A & cosine & 0.03 & 0.01 & 3.23 & 0.00 \\ 
    \hline
@@
 \end{table}
-\end{CodeOutput}
-\end{CodeChunk}
 
@@
 
-\section{Conclusions}\label{conclusions}
+\hypertarget{conclusions}{%
+\section{Conclusions}\label{conclusions}}
 
@@
 
-\section{Acknowledgements}\label{acknowledgements}
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
 
@@
 
-\section{References}\label{references}
+\hypertarget{references}{%
+\section{References}\label{references}}
 
-\setlength{\parindent}{-0.1in} \setlength{\leftskip}{0.125in} \noindent
+\setlength{\parindent}{-0.1in} 
+\setlength{\leftskip}{0.125in}
+
+\noindent
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-brown1973}{}%
+Brown, R. (1973). \emph{A first language: The early stages.} Harvard U.
+Press.
+
+\leavevmode\hypertarget{ref-genzel2002}{}%
+Genzel, D., \& Charniak, E. (2002). Entropy rate constancy in text. In
+\emph{Proceedings of the 40th annual meeting on association for
+computational linguistics} (pp. 199--206). Association for Computational
+Linguistics.
+
+\leavevmode\hypertarget{ref-jaeger2010}{}%
+Jaeger, T. F. (2010). Redundancy and reduction: Speakers manage
+syntactic information density. \emph{Cognitive Psychology},
+\emph{61}(1), 23--62.
+
+\leavevmode\hypertarget{ref-jaeger2013}{}%
+Jaeger, T. F., \& Snider, N. E. (2013). Alignment as a consequence of
+expectation adaptation: Syntactic priming is affected by the prime's
+prediction error given both prior and recent experience.
+\emph{Cognition}, \emph{127}(1), 57--83.
+
+\leavevmode\hypertarget{ref-jain2018}{}%
+Jain, A., Singh, V., Ranjan, S., Rajkumar, R., \& Agarwal, S. (2018).
+Uniform information density effects on syntactic choice in hindi. In
+\emph{Proceedings of the workshop on linguistic complexity and natural
+language processing} (pp. 38--48).
+
+\leavevmode\hypertarget{ref-macwhinney2014}{}%
+MacWhinney, B. (2014). \emph{The childes project: Tools for analyzing
+talk, volume ii: The database}. Psychology Press.
+
+\leavevmode\hypertarget{ref-mahowald2013}{}%
+Mahowald, K., Fedorenko, E., Piantadosi, S. T., \& Gibson, E. (2013).
+Info/information theory: Speakers choose shorter words in predictive
+contexts. \emph{Cognition}, \emph{126}(2), 313--318.
+
+\leavevmode\hypertarget{ref-piantadosi2011}{}%
+Piantadosi, S. T., Tily, H., \& Gibson, E. (2011). Word lengths are
+optimized for efficient communication. \emph{Proceedings of the National
+Academy of Sciences}, \emph{108}(9), 3526--3529.
+
+\leavevmode\hypertarget{ref-piantadosi2012}{}%
+Piantadosi, S. T., Tily, H., \& Gibson, E. (2012). The communicative
+function of ambiguity in language. \emph{Cognition}, \emph{122}(3),
+280--291.
+
+\leavevmode\hypertarget{ref-shannon1948}{}%
+Shannon, C. E. (1948). A mathematical theory of communication.
+\emph{Bell System Technical Journal}, \emph{27}(3), 379--423.
+
+\leavevmode\hypertarget{ref-xu2018}{}%
+Xu, Y., \& Reitter, D. (2018). Information density converges in
+dialogue: Towards an information-theoretic model. \emph{Cognition},
+\emph{170}, 147--163.
+
+\leavevmode\hypertarget{ref-yu2016}{}%
+Yu, S., Cong, J., Liang, J., \& Liu, H. (2016). The distribution of
+information content in english sentences. \emph{arXiv Preprint
+arXiv:1609.07681}.
+
+\leavevmode\hypertarget{ref-zhan2018}{}%
+Zhan, M., \& Levy, R. (2018). Comparing theories of speaker choice using
+a model of classifier production in mandarin chinese. In
+\emph{Proceedings of the 2018 conference of the north american chapter
+of the association for computational linguistics: Human language
+technologies, volume 1 (long papers)} (Vol. 1, pp. 1997--2005).
 

### 2019-01-26 cogsci/position-info.Rmd
some intro cleanup

@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/', echo=F, warning=F, cache=F, message=F, sanitize = T)
+knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/', echo=F, warning=F, cache=T, message=F, sanitize = T)
 ```
@@
 
-Perhaps the most important reason people talk and write to one another is to give information.
-From @shannon1948, we know that when information is being communicated, the rate of information transmission should be constant and as close as possible to the capacity of the communication channel to optimally transmit the information.
-@levy2007 introduce the *uniform information density* (UID) hypothesis, arguing that people unconsciously structure their speech and writing to communicate information at a constant, optimal rate.
-@levy2007 argue that speakers and writers should attempt to communicate information as evenly as possible and as close to the channel capacity as possible.
+Over 7,000 languages are spoken around the modern world [@simons2018].
+These language vary along many dimensions, but all share a core goal: communicating information.
+If speakers and writers of these languages act near-optimally to achieve their communicative goals, regularities of use across these diverse languages can be explained by a rational theory of communication [@anderson1991].
+Information theory, a mathematical framework developed by @shannon1948 to describe the transmission and decoding of signals, has been a unifying language for the recent development of such theories in human and machine language processing [@jelinek1976;@levy2007].
+
+These theories model the process of communication as transmission of information over a noisy channel.
+The producer begins with an intended meaning, packages this meaning into a linguistic format, and then sends it to the receiver over a communcative channel.
+The receiver must then decode from the signal they receive on their end of the chanel the producer's intended meaning.
+The problem is that the channel is noisy, and sometimes the signal can get corrupted (e.g.
+the producer can misspeak, or the receiver can mishear).
+In order to maximize the probability that the correct meaning is transmitted, these theories predict that producers should choose linguistic messages that keep the rate of information across words constant.
+The intuition is that if the receiver misperceives a word, and that word contains most of the information in the sentence, then the communication will have failed.
+Because producers cannot predict which word a speaker will mishear, their best strategy is spread the information evenly across all of the words in a sentence, i.e.
+maintain *uniform information density* [@genzel2002; @levy2007].
+
+**Evidence for UID here. Maybe see if you can present this work at a slightly higher level**
 Their evidence concerns the use or lack of use of “that”-complementizers in introducing English relative clauses.
@@
 UID has been applied broadly over the past decade, to determining whether linguistic alignment takes place [@jaeger2013], Zipfian word length distributions [@piantadosi2011], communication efficiency [@mahowald2013], dialogue and turn-taking [@xu2018] and the significance of ambiguity in language [@piantadosi2012], among other research.
+
+**Evidence against UID here. Again, can you present this at a higher level?**
 However, other recent work has contradicted the UID hypothesis.
@@
 Zhan and Levy find that speakers use specific classifiers more often in cases where the production of the corresponding noun is more difficult than when the production is easier; speakers do not avoid unexpected peaks in information content for their listeners, but instead maximize their ease of production.
-
 Similar to the original @levy2007, @zhan2018 focuses on information distribution at particular points in a sentence.
 By contrast, @jain2018 examine word order across spoken sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on word order.
-For this paper, we focus on the sentence level instead of specific points, following a method described in @yu2016.
+
+Recently, @yu2016 developed a more direct test of the Uniform Information Density hypothesis, applying the logic used by @genzel2002 to look at the distribution of information *within* individual sentences.
+**SOME HIGHER LEVEL DESCRIPTION HERE, WE'll GET INTO THE DETAILS BELOW**
+
+Because people process language incrementally--using the previous words in a sentence to predict the words that will come next, the amount of information that a word contains when seen in isolation should increase over the course of a sentence.
+Analyzing a large corpus of written English, they find a different pattern: Entropy increases over the first few words of an utterance and then remains constant until the final word where it again jumps up (see Figure \ref{fig:our_replication}).
+@yu2016 conclude that the Uniform Information Density hypothesis must not hold for medial words in a sentence.
+
+We extend and generalize @yu2016 in two ways: we confirm that this same pattern is found in spoken English--in both adult and child speakers.
+We then examine entropy curves cross-linguistically, finding a diversity of shapes across the world's languages.
+Finally, we show that entropy curves are predictable from the structure of individual languages, namely word order.
+Taken together, our results suggest a refinement of the Uniform information density hypothesis: Speakers may choose to order words in order to preserve uniform uniformation, but they must do so under prior constraints on word order imposed by the language they speak.
+
+The UID hypothesis predicts that information transmission rate will tend towards uniformity in both speech and text.
+We begin by examining speech to determine if we obtain the entropy curve shape we expect for spoken communication.
+We begin with the same source as @levy2007 and @jaeger2010, the Switchboard corpus of adult telephone conversations.
+We also examine child-adult conversations in the CHILDES TalkBank [@brown1973; @macwhinney2014] corpora database of spoken adult-child conversations in multiple languages.
+Children are not fully developed speakers, so we want to compare the entropy curve we obtain by computing over the utterances in the CHILDES corpora to the utterances in the adult-adult Switchboard corpus.
+
+## Methods
 
@@
 UID applies this uniformity of entropy rate in sentences to all levels of speech, and so the @yu2016, which examines text at the word level, should find an affine function at the word level.
-
-The entropy curve is a consequence of the information-theoretic rate of information transfer in a language, where information is defined as a reduction in uncertainty.
-Thus the more entropy at a word position in a sentence [@yu2016] or a sentence index in a text [@genzel2002], the more information is on average contained around that position.
-Let us assume that the UID perspective is correct.
-Then we expect that the entropy curve for every language should be close to an affine, linear function with some noise.
-Entropy should monotonically and uniformly increase at all word positions across utterances on average.
-We will find a contradiction with this claim through examining the average entropy at each position of a sentence for both speech and text data.
-
-[Put in an illustration of what UID predicts?]
-
-# Spoken data
-
-The UID hypothesis predicts that information transmission rate will tend towards uniformity in both speech and text.
-We begin by examining speech to determine if we obtain the entropy curve shape we expect for spoken communication.
-We begin with the same source as @levy2007 and @jaeger2010, the Switchboard corpus of adult telephone conversations.
-We also examine child-adult conversations in the CHILDES TalkBank [@brown1973; @macwhinney2014] corpora database of spoken adult-child conversations in multiple languages.
-Children are not fully developed speakers, so we want to compare the entropy curve we obtain by computing over the utterances in the CHILDES corpora to the utterances in the adult-adult Switchboard corpus.
-
-## Methods
 

### 2019-01-29 cogsci/position-info.Rmd
cogsci edits

@@
 We apply this model to corpora from Wikipedia in well over a hundred languages.
-We find that not only is the Uniform Information Density hypothesis prediction wrong, but also that the by-word entropy distribution of a language is related to the typological features of the language.
+We find that the Uniform Information Density hypothesis does not predict the outcome of our word-level entropy model, but also that the by-word entropy distribution of a language is related to the typological features of the language.
 We use this evidence to suggest that people do not communicate information at a uniform rate, but that information distribution varies from language to language based on phonogical, morphological and syntactic features.
@@
 
+# Introduction
+
 Over 7,000 languages are spoken around the modern world [@simons2018].
@@
 
-**Evidence for UID here. Maybe see if you can present this work at a slightly higher level**
-Their evidence concerns the use or lack of use of “that”-complementizers in introducing English relative clauses.
-When the word preceding the relative clause or the initial word of the relative clause has high information content, then the authors argue that speakers are more likely to use the complementizer.
-The authors use adult-adult conversations from the Penn Treebank corpus.
+<!-- **Evidence for UID here. Maybe see if you can present this work at a slightly higher level** -->
+The original evidence in @levy2007 finds that the insertion of complementizers in relative clauses in English corresponds to where neighboring words have high information content.
+Similarly, @frank2008speaking argues that contradictions in English such as "you're" do not occur when neighboring words are highly informative.
+The evidence in favor of UID largely been situation-specific and English-language driven, while the hypothesis itself has been applied broadly over the past decade.
+Applications include determining whether linguistic alignment takes place [@jaeger2013], Zipfian word length distributions [@piantadosi2011], communication efficiency [@mahowald2013], dialogue and turn-taking [@xu2018] and the significance of ambiguity in language [@piantadosi2012], among other research.
 
-UID has been applied broadly over the past decade, to determining whether linguistic alignment takes place [@jaeger2013], Zipfian word length distributions [@piantadosi2011], communication efficiency [@mahowald2013], dialogue and turn-taking [@xu2018] and the significance of ambiguity in language [@piantadosi2012], among other research.
-
-**Evidence against UID here. Again, can you present this at a higher level?**
+<!-- **Evidence against UID here. Again, can you present this at a higher level?** -->
 However, other recent work has contradicted the UID hypothesis.
-@zhan2018 study Mandarin Chinese classifier use for specific and general Mandarin classifiers.
-Specific classifiers are specific to certain nouns, and UID would predict that specific classifiers would appear when the production of the corresponding noun is easier, to spread the information contained by the classifier and noun less densely over time.
-General classifiers can be applied to any noun and contain no special information content.
-Zhan and Levy find that speakers use specific classifiers more often in cases where the production of the corresponding noun is more difficult than when the production is easier; speakers do not avoid unexpected peaks in information content for their listeners, but instead maximize their ease of production.
-Similar to the original @levy2007, @zhan2018 focuses on information distribution at particular points in a sentence.
-By contrast, @jain2018 examine word order across spoken sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on word order.
+Similar to the original @levy2007, @zhan2018 focuses on information distribution at particular points in sentences.
+@zhan2018 finds that more information-rich classifiers in Mandarin Chinese are produced when production of the neighboring noun is difficult, not when the information content is high.
+@jain2018 examine word order across spoken sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on word order.
 
 Recently, @yu2016 developed a more direct test of the Uniform Information Density hypothesis, applying the logic used by @genzel2002 to look at the distribution of information *within* individual sentences.
-**SOME HIGHER LEVEL DESCRIPTION HERE, WE'll GET INTO THE DETAILS BELOW**
-
-Because people process language incrementally--using the previous words in a sentence to predict the words that will come next, the amount of information that a word contains when seen in isolation should increase over the course of a sentence.
+<!-- **SOME HIGHER LEVEL DESCRIPTION HERE, WE'll GET INTO THE DETAILS BELOW** -->Because people process language incrementally--using the previous words in a sentence to predict the words that will come next, the amount of information that a word contains when seen in isolation should increase over the course of a sentence.
 Analyzing a large corpus of written English, they find a different pattern: Entropy increases over the first few words of an utterance and then remains constant until the final word where it again jumps up (see Figure \ref{fig:our_replication}).
@@
 
-We extend and generalize @yu2016 in two ways: we confirm that this same pattern is found in spoken English--in both adult and child speakers.
+We extend and generalize @yu2016 in three ways: we confirm that this same pattern is found in spoken English--in both adult and child speakers.
 We then examine entropy curves cross-linguistically, finding a diversity of shapes across the world's languages.
 Finally, we show that entropy curves are predictable from the structure of individual languages, namely word order.
-Taken together, our results suggest a refinement of the Uniform information density hypothesis: Speakers may choose to order words in order to preserve uniform uniformation, but they must do so under prior constraints on word order imposed by the language they speak.
+Taken together, our results suggest a refinement of the Uniform Information Density hypothesis: Speakers may choose to order words in order to preserve uniform information, but they must do so under prior constraints on word order imposed by the language they speak.
 
@@
 
-@yu2016 challenges the UID hypothesis through examining an entropy measure across sentence positions within the text portion of the English-language British National Corpus (BNC, CITE).
+@yu2016 challenges the UID hypothesis through examining an entropy measure across sentence positions within the text portion of the English-language British National Corpus (BNC) @clear1993british.
 They partition the corpus by sentence length in number of words.
@@
 
-We used the Brown and Providence English corpora from CHILDES.
-The Brown corpus contains individual recordings of conversations between three children between 1.5 and 6-years-old and their families in their homes.
+We used the Providence English corpds from CHILDES.
 The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home.
-We divided each corpus by speaker into child and non-child categories.
-We further divided the corpora by utterance length, so that all sentences of length $k$ (e.g.
+We divided the corpus by speaker into child and non-child categories.
+We further divided the corpus by utterance length, so that all sentences of length $k$ (e.g.
 $6$) were grouped together.
@@
 
-In the English CHILDES corpora.
 The entropy curves capture individual variation across positions in utterances of the same length.
@@
 We are applying the same approach as in @genzel2002, but within sentences instead of across sentences.
-
-Plots for the entropy distributions are below.
-
-```{r brown_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Brown corpus entropy"}
-img <- png::readPNG("figs/positional_entropy_brown.png")
-grid::grid.raster(img)
-```
 
@@
 ```{r shiro_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Shiro corpus entropy"}
-img <- png::readPNG("figs/German_unigram_entropies_with_stemming.png")
+img <- png::readPNG("figs/Spanish_unigram_entropies_with_stemming.png")
 grid::grid.raster(img)

### 2019-01-29 cogsci/position-info.tex
cogsci edits

@@
 corpora from Wikipedia in well over a hundred languages. We find that
-not only is the Uniform Information Density hypothesis prediction wrong,
-but also that the by-word entropy distribution of a language is related
-to the typological features of the language. We use this evidence to
-suggest that people do not communicate information at a uniform rate,
-but that information distribution varies from language to language based
-on phonogical, morphological and syntactic features.
+the Uniform Information Density hypothesis does not predict the outcome
+of our word-level entropy model, but also that the by-word entropy
+distribution of a language is related to the typological features of the
+language. We use this evidence to suggest that people do not communicate
+information at a uniform rate, but that information distribution varies
+from language to language based on phonogical, morphological and
+syntactic features.
 
@@
 
-Perhaps the most important reason people talk and write to one another
-is to give information. From Shannon (1948), we know that when
-information is being communicated, the rate of information transmission
-should be constant and as close as possible to the capacity of the
-communication channel to optimally transmit the information.
-({\textbf{???}}) introduce the \emph{uniform information density} (UID)
-hypothesis, arguing that people unconsciously structure their speech and
-writing to communicate information at a constant, optimal rate.
-({\textbf{???}}) argue that speakers and writers should attempt to
-communicate information as evenly as possible and as close to the
-channel capacity as possible. Their evidence concerns the use or lack of
-use of ``that''-complementizers in introducing English relative clauses.
-When the word preceding the relative clause or the initial word of the
-relative clause has high information content, then the authors argue
-that speakers are more likely to use the complementizer. The authors use
-adult-adult conversations from the Penn Treebank corpus.
+\section{Introduction}\label{introduction}
 
-UID has been applied broadly over the past decade, to determining
-whether linguistic alignment takes place (Jaeger \& Snider, 2013),
-Zipfian word length distributions (Piantadosi, Tily, \& Gibson, 2011),
-communication efficiency (Mahowald, Fedorenko, Piantadosi, \& Gibson,
-2013), dialogue and turn-taking (Xu \& Reitter, 2018) and the
-significance of ambiguity in language (Piantadosi, Tily, \& Gibson,
-2012), among other research. However, other recent work has contradicted
-the UID hypothesis. Zhan \& Levy (2018) study Mandarin Chinese
-classifier use for specific and general Mandarin classifiers. Specific
-classifiers are specific to certain nouns, and UID would predict that
-specific classifiers would appear when the production of the
-corresponding noun is easier, to spread the information contained by the
-classifier and noun less densely over time. General classifiers can be
-applied to any noun and contain no special information content. Zhan and
-Levy find that speakers use specific classifiers more often in cases
-where the production of the corresponding noun is more difficult than
-when the production is easier; speakers do not avoid unexpected peaks in
-information content for their listeners, but instead maximize their ease
-of production.
+Over 7,000 languages are spoken around the modern world (Simons \&
+Fenning, 2018). These language vary along many dimensions, but all share
+a core goal: communicating information. If speakers and writers of these
+languages act near-optimally to achieve their communicative goals,
+regularities of use across these diverse languages can be explained by a
+rational theory of communication (Anderson, 1991). Information theory, a
+mathematical framework developed by Shannon (1948) to describe the
+transmission and decoding of signals, has been a unifying language for
+the recent development of such theories in human and machine language
+processing (Jelinek, 1976; R. P. Levy \& Jaeger, 2007).
 
-Similar to the original ({\textbf{???}}), Zhan \& Levy (2018) focuses on
-information distribution at particular points in a sentence. By
-contrast, Jain, Singh, Ranjan, Rajkumar, \& Agarwal (2018) examine word
+These theories model the process of communication as transmission of
+information over a noisy channel. The producer begins with an intended
+meaning, packages this meaning into a linguistic format, and then sends
+it to the receiver over a communcative channel. The receiver must then
+decode from the signal they receive on their end of the chanel the
+producer's intended meaning. The problem is that the channel is noisy,
+and sometimes the signal can get corrupted (e.g.~the producer can
+misspeak, or the receiver can mishear). In order to maximize the
+probability that the correct meaning is transmitted, these theories
+predict that producers should choose linguistic messages that keep the
+rate of information across words constant. The intuition is that if the
+receiver misperceives a word, and that word contains most of the
+information in the sentence, then the communication will have failed.
+Because producers cannot predict which word a speaker will mishear,
+their best strategy is spread the information evenly across all of the
+words in a sentence, i.e.~maintain \emph{uniform information density}
+(Genzel \& Charniak, 2002; R. P. Levy \& Jaeger, 2007).
+
+The original evidence in R. P. Levy \& Jaeger (2007) finds that the
+insertion of complementizers in relative clauses in English corresponds
+to where neighboring words have high information content. Similarly,
+Frank \& Jaeger (2008) argues that contradictions in English such as
+``you're'' do not occur when neighboring words are highly informative.
+The evidence in favor of UID largely been situation-specific and
+English-language driven, while the hypothesis itself has been applied
+broadly over the past decade. Applications include determining whether
+linguistic alignment takes place (Jaeger \& Snider, 2013), Zipfian word
+length distributions (Piantadosi, Tily, \& Gibson, 2011), communication
+efficiency (Mahowald, Fedorenko, Piantadosi, \& Gibson, 2013), dialogue
+and turn-taking (Xu \& Reitter, 2018) and the significance of ambiguity
+in language (Piantadosi, Tily, \& Gibson, 2012), among other research.
+
+However, other recent work has contradicted the UID hypothesis. Similar
+to the original R. P. Levy \& Jaeger (2007), Zhan \& Levy (2018) focuses
+on information distribution at particular points in sentences. Zhan \&
+Levy (2018) finds that more information-rich classifiers in Mandarin
+Chinese are produced when production of the neighboring noun is
+difficult, not when the information content is high. On a sentential
+level, Jain, Singh, Ranjan, Rajkumar, \& Agarwal (2018) examine word
 order across spoken sentences in Hindi, a freer word order language than
 English, and find that information density has no significant effect on
-word order. For this paper, we focus on the sentence level instead of
-specific points, following a method described in Yu, Cong, Liang, \& Liu
-(2016).
+word order.
+
+Recently, Yu, Cong, Liang, \& Liu (2016) developed a more direct test of
+the Uniform Information Density hypothesis, applying the logic used by
+Genzel \& Charniak (2002) to look at the distribution of information
+\emph{within} individual sentences. Because people process language
+incrementally--using the previous words in a sentence to predict the
+words that will come next, the amount of information that a word
+contains when seen in isolation should increase over the course of a
+sentence. Analyzing a large corpus of written English, they find a
+different pattern: Entropy increases over the first few words of an
+utterance and then remains constant until the final word where it again
+jumps up (see Figure \ref{fig:our_replication}). Yu et al. (2016)
+conclude that the Uniform Information Density hypothesis must not hold
+for medial words in a sentence.
+
+We extend and generalize Yu et al. (2016) in three ways: we confirm that
+this same pattern is found in spoken English--in both adult and child
+speakers. We then examine entropy curves cross-linguistically, finding a
+diversity of shapes across the world's languages. Finally, we show that
+entropy curves are predictable from the structure of individual
+languages, namely word order. Taken together, our results suggest a
+refinement of the Uniform Information Density hypothesis: Speakers may
+choose to order words in order to preserve uniform information, but they
+must do so under prior constraints on word order imposed by the language
+they speak.
+
+The UID hypothesis predicts that information transmission rate will tend
+towards uniformity in both speech and text. We begin by examining speech
+to determine if we obtain the entropy curve shape we expect for spoken
+communication. We begin with the same source as R. P. Levy \& Jaeger
+(2007) and Jaeger (2010), the Switchboard corpus of adult telephone
+conversations. We also examine child-adult conversations in the CHILDES
+TalkBank (Brown, 1973; MacWhinney, 2014) corpora database of spoken
+adult-child conversations in multiple languages. Children are not fully
+developed speakers, so we want to compare the entropy curve we obtain by
+computing over the utterances in the CHILDES corpora to the utterances
+in the adult-adult Switchboard corpus.
+
+\subsection{Methods}\label{methods}
 
@@
 entropy measure across sentence positions within the text portion of the
-English-language British National Corpus (BNC, CITE). They partition the
-corpus by sentence length in number of words. For each word position
-\(X\) of sentences of length \(k\), they define \(w\) as a unique word
-occurring in position \(X\). They define \(p(w)\) as the number of times
-word \(w\) occurs in position \(X\) divided by the number of total words
-that occur in position \(X\) i.e.~the number of sentences of length
-\(k\). Then \(p(w)\) is the probability of obtaining word \(w\) by
-choosing a word at random in position \(X\) in sentences of length
-\(k\).
+English-language British National Corpus (BNC) Clear (1993). They
+partition the corpus by sentence length in number of words. For each
+word position \(X\) of sentences of length \(k\), they define \(w\) as a
+unique word occurring in position \(X\). They define \(p(w)\) as the
+number of times word \(w\) occurs in position \(X\) divided by the
+number of total words that occur in position \(X\) i.e.~the number of
+sentences of length \(k\). Then \(p(w)\) is the probability of obtaining
+word \(w\) by choosing a word at random in position \(X\) in sentences
+of length \(k\).
 
@@
 
-The entropy curve is a consequence of the information-theoretic rate of
-information transfer in a language, where information is defined as a
-reduction in uncertainty. Thus the more entropy at a word position in a
-sentence (Yu et al., 2016) or a sentence index in a text (Genzel \&
-Charniak, 2002), the more information is on average contained around
-that position. Let us assume that the UID perspective is correct. Then
-we expect that the entropy curve for every language should be close to
-an affine, linear function with some noise. Entropy should monotonically
-and uniformly increase at all word positions across utterances on
-average. We will find a contradiction with this claim through examining
-the average entropy at each position of a sentence for both speech and
-text data.
-
-{[}Put in an illustration of what UID predicts?{]}
-
-\hypertarget{spoken-data}{%
-\section{Spoken data}\label{spoken-data}}
-
-The UID hypothesis predicts that information transmission rate will tend
-towards uniformity in both speech and text. We begin by examining speech
-to determine if we obtain the entropy curve shape we expect for spoken
-communication. We begin with the same source as ({\textbf{???}}) and
-Jaeger (2010), the Switchboard corpus of adult telephone conversations.
-We also examine child-adult conversations in the CHILDES TalkBank
-(Brown, 1973; MacWhinney, 2014) corpora database of spoken adult-child
-conversations in multiple languages. Children are not fully developed
-speakers, so we want to compare the entropy curve we obtain by computing
-over the utterances in the CHILDES corpora to the utterances in the
-adult-adult Switchboard corpus.
-
-\hypertarget{methods}{%
-\subsection{Methods}\label{methods}}
-
 Switchboard {[}To be put in{]}
 
-We used the Brown and Providence English corpora from CHILDES. The Brown
-corpus contains individual recordings of conversations between three
-children between 1.5 and 6-years-old and their families in their homes.
-The Providence corpus recorded interactions between children between 1
-and 3 years old and their parents in the home. We divided each corpus by
-speaker into child and non-child categories. We further divided the
-corpora by utterance length, so that all sentences of length \(k\) (e.g.
-\(6\)) were grouped together. Finally, within each utterance length, we
+We used the Providence English corpds from CHILDES. The Providence
+corpus recorded interactions between children between 1 and 3 years old
+and their parents in the home. We divided the corpus by speaker into
+child and non-child categories. We further divided the corpus by
+utterance length, so that all sentences of length \(k\) (e.g. \(6\))
+were grouped together. Finally, within each utterance length, we
 computed the unigram entropy measure for each position.
 
-In the English CHILDES corpora. The entropy curves capture individual
-variation across positions in utterances of the same length. This allows
-us to directly observe and judge the amount of variation in words that
-appear in an individual position of a sentence. For speech data, which
-for the corpora in the CHILDES database consists of short and often
-disconnected utterances across hours of recordings, the unigram entropy
-measure is unaffected by context or lack thereof in utterances by the
-adults and children in the corpora. We can directly compare any two
-positions within utterances to determine the amount of uncertainty, and
-therefore information, on average contained by words within that
-position of utterances. We are applying the same approach as in Genzel
-\& Charniak (2002), but within sentences instead of across sentences.
-
-Plots for the entropy distributions are below.
-
-\begin{CodeChunk}
-\begin{figure*}[h]
-
-{\centering \includegraphics{figs/brown_PE-1}
-
-}
-
-\caption[Brown corpus entropy]{Brown corpus entropy}\label{fig:brown_PE}
-\end{figure*}
-\end{CodeChunk}
+The entropy curves capture individual variation across positions in
+utterances of the same length. This allows us to directly observe and
+judge the amount of variation in words that appear in an individual
+position of a sentence. For speech data, which for the corpora in the
+CHILDES database consists of short and often disconnected utterances
+across hours of recordings, the unigram entropy measure is unaffected by
+context or lack thereof in utterances by the adults and children in the
+corpora. We can directly compare any two positions within utterances to
+determine the amount of uncertainty, and therefore information, on
+average contained by words within that position of utterances. We are
+applying the same approach as in Genzel \& Charniak (2002), but within
+sentences instead of across sentences.
 
@@
 
-\hypertarget{results-analysis}{%
-\subsection{Results \& Analysis}\label{results-analysis}}
+\subsection{Results \& Analysis}\label{results-analysis}
 
@@
 
-\hypertarget{written-data}{%
-\section{Written data}\label{written-data}}
+\section{Written data}\label{written-data}
 
@@
 
-\hypertarget{methods-1}{%
-\subsection{Methods}\label{methods-1}}
+\subsection{Methods}\label{methods-1}
 
@@
 
-\hypertarget{results-and-analysis}{%
-\subsection{Results and Analysis}\label{results-and-analysis}}
+\subsection{Results and Analysis}\label{results-and-analysis}
 
@@
 
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
+\section{Conclusions}\label{conclusions}
 
@@
 
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
+\section{Acknowledgements}\label{acknowledgements}
 
@@
 
-\hypertarget{references}{%
-\section{References}\label{references}}
+\section{References}\label{references}
 
-\setlength{\parindent}{-0.1in} 
-\setlength{\leftskip}{0.125in}
-
-\noindent
+\setlength{\parindent}{-0.1in} \setlength{\leftskip}{0.125in} \noindent
 
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-brown1973}{}%
+\hypertarget{ref-anderson1991}{}
+Anderson, J. R. (1991). The adaptive nature of human categorization.
+\emph{Psychological Review}, \emph{98}(3), 409.
+
+\hypertarget{ref-brown1973}{}
 Brown, R. (1973). \emph{A first language: The early stages.} Harvard U.
@@
 
-\leavevmode\hypertarget{ref-genzel2002}{}%
+\hypertarget{ref-clear1993british}{}
+Clear, J. H. (1993). The british national corpus. \emph{The Digital
+World}, 163--187.
+
+\hypertarget{ref-frank2008speaking}{}
+Frank, A. F., \& Jaeger, T. F. (2008). Speaking rationally: Uniform
+information density as an optimal strategy for language production. In
+\emph{Proceedings of the annual meeting of the cognitive science
+society} (Vol. 30).
+
+\hypertarget{ref-genzel2002}{}
 Genzel, D., \& Charniak, E. (2002). Entropy rate constancy in text. In
@@
 
-\leavevmode\hypertarget{ref-jaeger2010}{}%
+\hypertarget{ref-jaeger2010}{}
 Jaeger, T. F. (2010). Redundancy and reduction: Speakers manage
@@
 
-\leavevmode\hypertarget{ref-jaeger2013}{}%
+\hypertarget{ref-jaeger2013}{}
 Jaeger, T. F., \& Snider, N. E. (2013). Alignment as a consequence of
@@
 
-\leavevmode\hypertarget{ref-jain2018}{}%
+\hypertarget{ref-jain2018}{}
 Jain, A., Singh, V., Ranjan, S., Rajkumar, R., \& Agarwal, S. (2018).
@@
 
-\leavevmode\hypertarget{ref-macwhinney2014}{}%
+\hypertarget{ref-jelinek1976}{}
+Jelinek, F. (1976). Continuous speech recognition by statistical
+methods. \emph{Proceedings of the IEEE}, \emph{64}(4), 532--556.
+
+\hypertarget{ref-levy2007}{}
+Levy, R. P., \& Jaeger, T. F. (2007). Speakers optimize information
+density through syntactic reduction. In \emph{Advances in neural
+information processing systems} (pp. 849--856).
+
+\hypertarget{ref-macwhinney2014}{}
 MacWhinney, B. (2014). \emph{The childes project: Tools for analyzing
@@
 
-\leavevmode\hypertarget{ref-mahowald2013}{}%
+\hypertarget{ref-mahowald2013}{}
 Mahowald, K., Fedorenko, E., Piantadosi, S. T., \& Gibson, E. (2013).
@@
 
-\leavevmode\hypertarget{ref-piantadosi2011}{}%
+\hypertarget{ref-piantadosi2011}{}
 Piantadosi, S. T., Tily, H., \& Gibson, E. (2011). Word lengths are
@@
 
-\leavevmode\hypertarget{ref-piantadosi2012}{}%
+\hypertarget{ref-piantadosi2012}{}
 Piantadosi, S. T., Tily, H., \& Gibson, E. (2012). The communicative
@@
 
-\leavevmode\hypertarget{ref-shannon1948}{}%
+\hypertarget{ref-shannon1948}{}
 Shannon, C. E. (1948). A mathematical theory of communication.
@@
 
-\leavevmode\hypertarget{ref-xu2018}{}%
+\hypertarget{ref-simons2018}{}
+Simons, G. F., \& Fenning, C. D. (Eds.). (2018). \emph{Ethnologue:
+Languages of the world, 21st edition}. Dallas, Texas: SIL International.
+
+\hypertarget{ref-xu2018}{}
 Xu, Y., \& Reitter, D. (2018). Information density converges in
@@
 
-\leavevmode\hypertarget{ref-yu2016}{}%
+\hypertarget{ref-yu2016}{}
 Yu, S., Cong, J., Liang, J., \& Liu, H. (2016). The distribution of
@@
 
-\leavevmode\hypertarget{ref-zhan2018}{}%
+\hypertarget{ref-zhan2018}{}
 Zhan, M., \& Levy, R. (2018). Comparing theories of speaker choice using

### 2019-01-29 cogsci/position-info.Rmd
incorporating switchboard code

@@
 ---
-title: "Information is Not Communicated Uniformly: Evidence from Spoken and Written Corpora"
+title: "Information Distribution Depends on a Language's Grammar and Sounds"
 bibliography: library.bib
@@
 These theories model the process of communication as transmission of information over a noisy channel.
-The producer begins with an intended meaning, packages this meaning into a linguistic format, and then sends it to the receiver over a communcative channel.
-The receiver must then decode from the signal they receive on their end of the chanel the producer's intended meaning.
+The producer begins with an intended meaning, packages this meaning into language, and then sends the meaning to their intended receiver over a communcative channel.
+The receiver must then decode from the signal they receive on their end of the channel the producer's intended meaning.
 The problem is that the channel is noisy, and sometimes the signal can get corrupted (e.g.
@@
 Recently, @yu2016 developed a more direct test of the Uniform Information Density hypothesis, applying the logic used by @genzel2002 to look at the distribution of information *within* individual sentences.
-<!-- **SOME HIGHER LEVEL DESCRIPTION HERE, WE'll GET INTO THE DETAILS BELOW** -->Because people process language incrementally--using the previous words in a sentence to predict the words that will come next, the amount of information that a word contains when seen in isolation should increase over the course of a sentence.
+<!-- **SOME HIGHER LEVEL DESCRIPTION HERE, WE'll GET INTO THE DETAILS BELOW** -->Because people process language incrementally--using the previous words in a sentence to predict the words that will come next--the amount of information that a word contains when seen in isolation should increase over the course of a sentence.
 Analyzing a large corpus of written English, they find a different pattern: Entropy increases over the first few words of an utterance and then remains constant until the final word where it again jumps up (see Figure \ref{fig:our_replication}).
@@
 Finally, we show that entropy curves are predictable from the structure of individual languages, namely word order.
-Taken together, our results suggest a refinement of the Uniform Information Density hypothesis: Speakers may choose to order words in order to preserve uniform information, but they must do so under prior constraints on word order imposed by the language they speak.
+Taken together, our results suggest a refinement of the Uniform Information Density hypothesis: speakers may structure their utterances to optimize information density, but they must do so under the constraints of the language they use to communicate.
 
@@
 
-Switchboard [To be put in]
+```{r switchboard, include=FALSE}
 
-We used the Providence English corpds from CHILDES.
+process_switchboard <- function() {
+
+}
+
+unigram_entropy <- function(tokens) {
+
+  tokens %>%
+    group_by(word_order, word) %>%
+    summarise(n = n()) %>%
+    tidyboot(summary_function = function(x) x %>% 
+               summarise(entropy = entropy(n, unit = "log2")),
+             statistics_functions = function(x) x %>%
+             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
+    mutate(role = role, length = sen_length)
+}
+
+tokens <- process_switchboard()
+
+entropies <- map(2:10, ~unigram_entropy("", .x, brown_utterances)) %>%
+  bind_rows()
+
+ggplot(entropies, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F)
+```
+
+We used the Providence English corpus from CHILDES.
 The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home.
@@
 
-[Put in an illustration of the slope sections and treatments here?]
-
 ## Results and Analysis
@@
 
-# Conclusions 
+# Discussion 
 In this paper, we have characterized and applied a model based on Genzel & Charniak (2002) for distinguishing entropy at each word position within sentences.

### 2019-01-29 cogsci/position-info.Rmd
new cogsci abstract & title

@@
 ---
-title: "Information Distribution Depends on a Language's Grammar and Sounds"
+title: "Information Distribution Depends on Language-Specific Features"
 bibliography: library.bib
@@
 abstract: >
-    We provide evidence against the popular Uniform Information Density hypothesis (Levy & Jaeger, 2007) which proposes that information is transmitted at a constant rate close to channel capacity in human communication.
-Using a method based on the original Genzel & Charniak (2002) entropy measure, we construct a word-level model for entropy applicable to both spoken and written corpora.
-We apply this model to corpora from Wikipedia in well over a hundred languages.
-We find that the Uniform Information Density hypothesis does not predict the outcome of our word-level entropy model, but also that the by-word entropy distribution of a language is related to the typological features of the language.
-We use this evidence to suggest that people do not communicate information at a uniform rate, but that information distribution varies from language to language based on phonogical, morphological and syntactic features.
+    We argue for a reformulation of the Uniform Information Density hypothesis @levy2007 in which the grammatical, phonological and morphological features of a language determine how information is distributed throughout utterances.
+We expand on a method from @yu2016 to compute entropy based on word position within utterances as a derivation of information distribution.
+We analyze adult and child spoken corpora as well as text corpora drawn from a variety of languages on Wikipedia.
+We find that each language has a robust, characteristic entropy curve across word position, and present evidence that these entropy curves are in part determined by typological features.
+We conclude that speakers may tend towards communicating information uniformly, but the distribution of information in their utterances is conditioned on the language they communicate in.
 
@@
 We also ran this analysis on Spanish and Mandarin corpora from CHILDES.
-We used the Shiro corpus for Spanish (Shiro, 2000), which contains prompted narratives individually collected from over a hundred Venezualan schoolchildren, half from high SES backgrounds and half from low SES backgrounds.
+We used the Shiro corpus for Spanish @shiro2000, which contains prompted narratives individually collected from over a hundred Venezualan schoolchildren, half from high SES backgrounds and half from low SES backgrounds.
 We used the Zhou dinner corpus for Mandarin Chinese (Li & Zhou, 2015), which contains dinner conversations between 5 to 6-year-old children and their parents collected in Shanghai.
 
-For each corpus, we accessed transcripts of the corpus provided through the TalkBank system and computed over Roman alphabet transcriptions or transliterations of the original transcriptions.
-For Mandarin, we used pinyin transliterations of the utterances in the corpus with demarcated word boundaries, and for Japanese we used romanji (Roman alphabet) transliterations of words in the corpus.
-The Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Chinese writing including spaces between word boundaries can have a negative effect on reading times (Bai et al, 2008).
+For each corpus, we accessed transcripts of the corpus provided through the TalkBank system and computed over the Latin alphabet transcriptions or transliterations of the original transcriptions.
+For Mandarin, we used pinyin transliterations of the utterances in the corpus with demarcated word boundaries.
+The Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Chinese writing including spaces between word boundaries can have a negative effect on reading times @bai2008.
 
@@
 
-Using Giuseppe Attardi's Wikiextractor tool \footnote{https://github.com/attardi/wikiextractor}, we extracted the text corpora for $165$ languages from Wikipedia by downloading a stored collection of Wikipedia entries in each langauge and randomly selecting several thousand articles from each Wikipedia language.
+Using Giuseppe Attardi's Wikiextractor tool \footnote{https://github.com/attardi/wikiextractor}, we extract text corpora from Wikipedia by downloading a stored collection of Wikipedia entries in each langauge and randomly selecting several thousand articles from each Wikipedia language.
 Each language corpus was cleaned and limited to sentences between $6$ and $50$ words.
@@
 In the *relative 10* treatment, each language is embedded in $10$-dimensional space.
-This allows for direct comparison between languages on Wikipedia using cosine similarity and moreover for unsupervised clustering analysis to compare the results of the Wikipedia entropy analysis to known typological features within the langauges in the Wikipedia dataset.
-To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we looked at the list of $144$ linguistic features in World Atlas of Language Structures (Dryer & Haspelmath, 2013).
-We limited the languages in the WALS database to only those in our Wikipedia dataset and performed missing-value imputation to obtain the features not coded in WALS for the languages from Wikipedia.
+This allows for direct comparison between languages on Wikipedia using cosine similarity and moreover for unsupervised clustering analysis to compare the results of our position-wise analysis on datasets from Wikipedia to known typological features within the languages.
+To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we use the linguistic features in World Atlas of Language Structures (Dryer & Haspelmath, 2013).
 
-One approach is to cluster the languages used in the Wikipedia analysis on the basis of WALS features and then directly compare the WALS clustering results with the results of the hierarchical clustering on the Wikipedia slope data for each of the different treatments using clustering similarity evaluation methods such as the Rand index (Rand, 1971).
+[Ideally do everything, but missing values in WALS, and how to compare every combination of features] An ideal approach using all of the corpora we can download from Wikipedia and all $144$ features from the WALS database.
+Cluster the languages used in the Wikipedia analysis on the basis of WALS features and then directly compare the WALS clustering results with the results of the hierarchical clustering on the Wikipedia slope data for each of the different treatments using clustering similarity evaluation methods such as the Rand index (Rand, 1971).
 However, the problem then arises of which combination of WALS features and how many features to include in the clustering analysis.
@@
 Cosine distance played some role in the feature determination.
-The top features were $83$, $81$, $95$ and $97$, which all characterize word order.
-This makes intuitive sense, and indicates that for this small subset of features and subset of languages that the embeddings in the slope space are related to the WALS features.
+The top four features all characterize word order, which indicates that for this subset of features and languages that the embeddings in the slope space are related to the WALS features.
 Therefore typological features play a role in determining the positional entropy values for a language.
@@
 # Discussion 
-In this paper, we have characterized and applied a model based on Genzel & Charniak (2002) for distinguishing entropy at each word position within sentences.
-In doing so, we have provided evidence against the Uniform Information Density hypothesis.
-Our work indicates that information is not distributed uniformly throughout an utterance regardless of languages, but that languages possess a characteristic information distribution determined by their syntactic, morphological and phonological features.
+In this paper, we have extended and applied a model from @yu2016 and derived from @genzel2002 for distinguishing entropy at each word position within sentences.
+We have argued that the outcomes of this model are derived from the information distribution of a language.
+As this model varies cross-linguistically in terms of
 
-Our work has focused on words within sentences in both speech and text across typologically diverse languages.
-Other researchers have explored information rate cross-linguistically as a single number characterizing each language, most notably including Aylett & Turk (2004).
-That work characterizes information rate as information per unit time, and is complementary to the work we accomplish here.
-Languages may possess both an overall information transfer rate, showing the average amount of semantic information transmitted per syllable or word, and an entropy distribution indicative of the language's information distribution.
+Follow-up with WALS features using missing-value imputation. However here are some of the problems with that.
 
-Research has indicated the effects of surprisal on fixation duration during eye-tracking studies (Demberg & Keller, 2008) (Boston et al., 2008) (Smith & Levy, 2013).
-Higher surprisal of a word is on average predictive of higher fixation duration.
-The *wrap-up effect* states that when a person reads written text, he or she will process sentence-final words more slowly on average then sentence-medial or sentence-initial words, due to integrating information from the entire sentence to form a final understanding of the sentence's meaning (Stowe et al., 2018) (Kuperman et al., 2010).
-The wrap-up effect is drawn from evidence in English, German and Dutch, all of which are languages with a large final increase in their entropy curve from our study.
-We aim to see in the future if as a consequence of our work in this paper, the wrap-up effect falls out from the final jump in the entropy curves in some languages.
-We would analyze eye-tracking data from languages with low positional entropy in sentence-final positions.
-Since the lexical surprisal metric in eye-tracking and our positional entropy metric are both measures of log-probability surprisal, then in languages such as Hindi where the sentence-final positional entropy values are low we expect to find no wrap-up effect.
+Our work complements the approach of studies such as @aylett2004, where languages are characterized by a rate of semantic information per syllable.
+This body of work attests to languages possessing different rates of information transfer based on typological features as well.
+
+We also propose follow-up within linguistic processing using evidence from cross-linguistic eye-tracking.
+Research has indicated the effects of surprisal on fixation duration during eye-tracking studies [@demberg2008; @boston2008; @smith2013]; higher surprisal of a word is on average predictive of higher fixation duration.
+The *wrap-up effect* states that when a person reads written text, he or she will process sentence-final words more slowly on average then sentence-medial or sentence-initial words, due to integrating information from the entire sentence to form a final understanding of the sentence's meaning [@stowe2018; @kuperman2010].
+The wrap-up effect is drawn from evidence in languages with a large final increase in their entropy curve from our study, so perhaps the wrap-up effect derives from the sentence-final increase in entropy curves observed in English and other Germanic languages.
 

### 2019-01-29 cogsci/position-info.Rmd
switchboard in cogsci draft

@@
 ```{r switchboard, include=FALSE}
+library(feather)
 
-process_switchboard <- function() {
+sb_tokens <- read_feather("../switchboard/switchboard.feather") %>%
+  mutate(length = str_count(value, " ") + 1) %>%
+  mutate(utterance_id = 1:n()) %>%
+  unnest_tokens(word, value) %>%
+  group_by(utterance_id) %>%
+  mutate(word_order = 1:n())
 
-}
-
-unigram_entropy <- function(tokens) {
-
+unigram_entropy <- function(sen_length, tokens) {
   tokens %>%
+    filter(length == sen_length) %>%
     group_by(word_order, word) %>%
@@
              summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
-    mutate(role = role, length = sen_length)
+    mutate(length = sen_length)
 }
 
-tokens <- process_switchboard()
-
-entropies <- map(2:10, ~unigram_entropy("", .x, brown_utterances)) %>%
+entropies <- map(2:10, ~unigram_entropy(.x, sb_tokens)) %>%
   bind_rows()

### 2019-01-30 cogsci/position-info.Rmd
fixing switchboard

@@
 library(xtable)
+library(tidytext)
+library(tidyboot)
+library(entropy)
 ```
@@
 sb_tokens <- read_feather("../switchboard/switchboard.feather") %>%
-  mutate(length = str_count(value, " ") + 1) %>%
+  mutate(length = str_count(value, pattern = " +") + 1) %>% 
   mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, value) %>%
+  unnest_tokens(word, value, token = stringr::str_split, pattern = " +") %>%
   group_by(utterance_id) %>%

### 2019-01-30 cogsci/position-info.Rmd
bnc

--- 2019-01-29 cogsci/position-info.Rmd
@@
 library(xtable)
+library(tidytext)
+library(tidyboot)
+library(entropy)
+library(here)
+library(feather)
 ```
@@
 
-```{r switchboard, include=FALSE}
-library(feather)
-
-sb_tokens <- read_feather("../switchboard/switchboard.feather") %>%
-  mutate(length = str_count(value, " ") + 1) %>%
+```{r bnc_and_switchboard, eval = F}
+sb_tokens <- read_feather(here("../switchboard/switchboard.feather")) %>%
+  mutate(length = str_count(value, pattern = " +") + 1) %>% 
   mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, value) %>%
+  unnest_tokens(word, value, token = stringr::str_split, pattern = " +") %>%
   group_by(utterance_id) %>%
   mutate(word_order = 1:n())
+
+bnc10_tokens <- read_feather(here("../bnc/bnc10.feather")) %>%
+  mutate(word = as.numeric(as.factor(word)))
+bnc20_tokens <- read_feather(here("../bnc/bnc20.feather")) %>%
+  mutate(word = as.numeric(as.factor(word)))
+bnc30_tokens <- read_feather(here("../bnc/bnc30.feather")) %>%
+  mutate(word = as.numeric(as.factor(word)))
 
@@
 
-entropies <- map(2:10, ~unigram_entropy(.x, sb_tokens)) %>%
-  bind_rows()
+switchboard_entropies <- map_df(2:10, ~unigram_entropy(.x, sb_tokens))
 
-ggplot(entropies, aes(x = word_order, y = empirical_entropy,
+bnc10_entropies <- unigram_entropy(10, bnc10_tokens)
+bnc20_entropies <- unigram_entropy(20, bnc20_tokens)
+bnc30_entropies <- unigram_entropy(30, bnc30_tokens)
+
+bnc_entropies <- bind_rows(bnc10_entropies, bnc20_entropies, bnc30_entropies)
+
+write_csv(bnc_entropies, here("data/bnc_entropies.csv"))
+write_csv(switchboard_entropies, here("data/switchboard_entropies.csv"))
+```
+
+```{r read_bnc_and_switchboard}
+bnc_entropies <- read_csv(here("data/bnc_entropies.csv"))
+switchboard_entropies <- read_csv(here("data/switchboard_entropies.csv"))
+```
+
+```{r plot_bnc_and_switchboard}
+ggplot(switchboard_entropies, aes(x = word_order, y = empirical_entropy,
                       ymin = ci_lower, ymax = ci_upper)) +
@@
   geom_smooth(se = F)
+
+ggplot(bnc_entropies, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_line()
 ```

### 2019-01-30 cogsci/position-info.Rmd
cogsci reorganization

--- 2019-01-30 cogsci/position-info.Rmd
@@
 
+```{r our_replication, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Providence corpus unigram entropy"}
+
+img <- png::readPNG("figs/Prov_corpus_result.png")
+grid::grid.raster(img)
+```
+
 We extend and generalize @yu2016 in three ways: we confirm that this same pattern is found in spoken English--in both adult and child speakers.
@@
 Children are not fully developed speakers, so we want to compare the entropy curve we obtain by computing over the utterances in the CHILDES corpora to the utterances in the adult-adult Switchboard corpus.
+Finally, we extend our cross-linguistic findings on a small number of languages from the CHILDES experiment to a direct comparison of several dozen languages in Wikipedia.
 
-## Methods
+## General methods
 
@@
 $$H(X) = \sum\limits_w p(w)\log\big(p(w)\big)$$
+
 With this measure, @yu2016 compute the unigram entropy at each position of sentences of each length within the corpus.
@@
 UID applies this uniformity of entropy rate in sentences to all levels of speech, and so the @yu2016, which examines text at the word level, should find an affine function at the word level.
+
+The entropy curves capture individual variation across positions in utterances of the same length.
+This allows us to directly observe and judge the amount of variation in words that appear in an individual position of a sentence.
+We can directly compare any two positions within utterances to determine the amount of uncertainty, and therefore information, on average contained by words within that position of utterances.
+We are applying the same approach as in @genzel2002, but within sentences instead of across sentences.
+
+# Experiment 1
+
+We first used the same source for data as @levy2007, the Switchboard corpus of American telephone conversations @godfrey1992.
 
@@
 
-entropies <- map(2:10, ~unigram_entropy(.x, sb_tokens)) %>%
+entropies <- map(c(5, 7, 9), ~unigram_entropy(.x, sb_tokens)) %>%
   bind_rows()
@@
   facet_wrap(~ length) + 
+  ggtitle("Representative Switchboard Entropy Curves Faceted on Utterance Length") + 
+  xlab("Word Order within Utterance") +
+  ylab("Positional Entropy") +
+  scale_x_discrete(limits = "1":"9") + 
   geom_pointrange() +
@@
 
+# Experiment 2
+
 We used the Providence English corpus from CHILDES.
 The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home.
-We divided the corpus by speaker into child and non-child categories.
+For speech data, which for the corpora in the CHILDES database consists of short and often disconnected utterances across hours of recordings, the unigram entropy measure is unaffected by context or lack thereof in utterances by the adults and children in the corpora.We divided the corpus by speaker into child and non-child categories.
 We further divided the corpus by utterance length, so that all sentences of length $k$ (e.g.
@@
 Finally, within each utterance length, we computed the unigram entropy measure for each position.
-
-The entropy curves capture individual variation across positions in utterances of the same length.
-This allows us to directly observe and judge the amount of variation in words that appear in an individual position of a sentence.
-For speech data, which for the corpora in the CHILDES database consists of short and often disconnected utterances across hours of recordings, the unigram entropy measure is unaffected by context or lack thereof in utterances by the adults and children in the corpora.
-We can directly compare any two positions within utterances to determine the amount of uncertainty, and therefore information, on average contained by words within that position of utterances.
-We are applying the same approach as in @genzel2002, but within sentences instead of across sentences.
+Computation of positional entropy was identical to the Switchboard method after dividing for utterance length.
 
@@
 
-# Written data
+# Experiment 3
 

### 2019-01-30 cogsci/position-info.tex
cogsci reorganization

@@
 
-\title{Information is Not Communicated Uniformly: Evidence from Spoken and
-Written Corpora}
+\title{Information Distribution Depends on Language-Specific Features}
 
@@
 \begin{abstract}
-We provide evidence against the popular Uniform Information Density
-hypothesis (Levy \& Jaeger, 2007) which proposes that information is
-transmitted at a constant rate close to channel capacity in human
-communication. Using a method based on the original Genzel \& Charniak
-(2002) entropy measure, we construct a word-level model for entropy
-applicable to both spoken and written corpora. We apply this model to
-corpora from Wikipedia in well over a hundred languages. We find that
-the Uniform Information Density hypothesis does not predict the outcome
-of our word-level entropy model, but also that the by-word entropy
-distribution of a language is related to the typological features of the
-language. We use this evidence to suggest that people do not communicate
-information at a uniform rate, but that information distribution varies
-from language to language based on phonogical, morphological and
-syntactic features.
+We argue for a reformulation of the Uniform Information Density
+hypothesis R. P. Levy \& Jaeger (2007) in which the grammatical,
+phonological and morphological features of a language determine how
+information is distributed throughout utterances. We expand on a method
+from Yu, Cong, Liang, \& Liu (2016) to compute entropy based on word
+position within utterances as a derivation of information distribution.
+We analyze adult and child spoken corpora as well as text corpora drawn
+from a variety of languages on Wikipedia. We find that each language has
+a robust, characteristic entropy curve across word position, and present
+evidence that these entropy curves are in part determined by typological
+features. We conclude that speakers may tend towards communicating
+information uniformly, but the distribution of information in their
+utterances is conditioned on the language they communicate in.
 
@@
 information over a noisy channel. The producer begins with an intended
-meaning, packages this meaning into a linguistic format, and then sends
-it to the receiver over a communcative channel. The receiver must then
-decode from the signal they receive on their end of the chanel the
-producer's intended meaning. The problem is that the channel is noisy,
-and sometimes the signal can get corrupted (e.g.~the producer can
-misspeak, or the receiver can mishear). In order to maximize the
+meaning, packages this meaning into language, and then sends the meaning
+to their intended receiver over a communcative channel. The receiver
+must then decode from the signal they receive on their end of the
+channel the producer's intended meaning. The problem is that the channel
+is noisy, and sometimes the signal can get corrupted (e.g.~the producer
+can misspeak, or the receiver can mishear). In order to maximize the
 probability that the correct meaning is transmitted, these theories
@@
 Chinese are produced when production of the neighboring noun is
-difficult, not when the information content is high. On a sentential
-level, Jain, Singh, Ranjan, Rajkumar, \& Agarwal (2018) examine word
-order across spoken sentences in Hindi, a freer word order language than
-English, and find that information density has no significant effect on
-word order.
+difficult, not when the information content is high. Jain, Singh,
+Ranjan, Rajkumar, \& Agarwal (2018) examine word order across spoken
+sentences in Hindi, a freer word order language than English, and find
+that information density has no significant effect on word order.
 
-Recently, Yu, Cong, Liang, \& Liu (2016) developed a more direct test of
-the Uniform Information Density hypothesis, applying the logic used by
-Genzel \& Charniak (2002) to look at the distribution of information
-\emph{within} individual sentences. Because people process language
+Recently, Yu et al. (2016) developed a more direct test of the Uniform
+Information Density hypothesis, applying the logic used by Genzel \&
+Charniak (2002) to look at the distribution of information \emph{within}
+individual sentences. Because people process language
 incrementally--using the previous words in a sentence to predict the
-words that will come next, the amount of information that a word
+words that will come next--the amount of information that a word
 contains when seen in isolation should increase over the course of a
@@
 
+\begin{CodeChunk}
+\begin{figure*}[h]
+
+{\centering \includegraphics{figs/our_replication-1}
+
+}
+
+\caption[Providence corpus unigram entropy]{Providence corpus unigram entropy}\label{fig:our_replication}
+\end{figure*}
+\end{CodeChunk}
+
 We extend and generalize Yu et al. (2016) in three ways: we confirm that
@@
 languages, namely word order. Taken together, our results suggest a
-refinement of the Uniform Information Density hypothesis: Speakers may
-choose to order words in order to preserve uniform information, but they
-must do so under prior constraints on word order imposed by the language
-they speak.
+refinement of the Uniform Information Density hypothesis: speakers may
+structure their utterances to optimize information density, but they
+must do so under the constraints of the language they use to
+communicate.
 
@@
 
-\subsection{Methods}\label{methods}
+\section{Experiment 1}\label{experiment-1}
 
@@
 
-Switchboard {[}To be put in{]}
+\section{Experiment 2}\label{experiment-2}
 
-We used the Providence English corpds from CHILDES. The Providence
+We used the Providence English corpus from CHILDES. The Providence
 corpus recorded interactions between children between 1 and 3 years old
@@
 We also ran this analysis on Spanish and Mandarin corpora from CHILDES.
-We used the Shiro corpus for Spanish (Shiro, 2000), which contains
+We used the Shiro corpus for Spanish Shiro (2000), which contains
 prompted narratives individually collected from over a hundred
@@
 For each corpus, we accessed transcripts of the corpus provided through
-the TalkBank system and computed over Roman alphabet transcriptions or
-transliterations of the original transcriptions. For Mandarin, we used
-pinyin transliterations of the utterances in the corpus with demarcated
-word boundaries, and for Japanese we used romanji (Roman alphabet)
-transliterations of words in the corpus. The Chinese characters used for
-writing Mandarin do not normally demarcate word boundaries by spacing
-words apart, and for normal Chinese writing including spaces between
-word boundaries can have a negative effect on reading times (Bai et al,
-2008).
+the TalkBank system and computed over the Latin alphabet transcriptions
+or transliterations of the original transcriptions. For Mandarin, we
+used pinyin transliterations of the utterances in the corpus with
+demarcated word boundaries. The Chinese characters used for writing
+Mandarin do not normally demarcate word boundaries by spacing words
+apart, and for normal Chinese writing including spaces between word
+boundaries can have a negative effect on reading times Bai, Yan,
+Liversedge, Zang, \& Rayner (2008).
 
@@
 
-\section{Written data}\label{written-data}
+\section{Experiment 3}\label{experiment-3}
 
@@
 
-\subsection{Methods}\label{methods-1}
+\subsection{Methods}\label{methods}
 
 Using Giuseppe Attardi's Wikiextractor tool
-\footnote{https://github.com/attardi/wikiextractor}, we extracted the
-text corpora for \(165\) languages from Wikipedia by downloading a
-stored collection of Wikipedia entries in each langauge and randomly
-selecting several thousand articles from each Wikipedia language. Each
-language corpus was cleaned and limited to sentences between \(6\) and
-\(50\) words. Similar to our process for the Switchboard and CHILDES
-spoken corpora, we divided each corpus by sentence length, and then
-computed the unigram entropy measure on each word position within each
-sentence length. We wanted to directly compare and classify the unigram
-entropy curves of the languages from Wikipedia. We computed three slope
-treatments of each curve. In the \emph{absolute} treatment, with
-sentence length denoted as \(n\), we computed the slope between
-positions \(1\) and \(2\), positions \(2\) and \(3\), positions \(3\)
-and \(n-2\), positions \(n-2\) and \(n-1\) and positions \(n-1\) and
-\(n\). For the short utterances appearing the CHILDES speech corpora we
-examined, these appeared to be important junctions in the distributions,
-with a seeming plateau in the middle of the unigram entropy curve for
-each of the language corpora we examined in CHILDES.
+\footnote{https://github.com/attardi/wikiextractor}, we extract text
+corpora from Wikipedia by downloading a stored collection of Wikipedia
+entries in each langauge and randomly selecting several thousand
+articles from each Wikipedia language. Each language corpus was cleaned
+and limited to sentences between \(6\) and \(50\) words. Similar to our
+process for the Switchboard and CHILDES spoken corpora, we divided each
+corpus by sentence length, and then computed the unigram entropy measure
+on each word position within each sentence length. We wanted to directly
+compare and classify the unigram entropy curves of the languages from
+Wikipedia. We computed three slope treatments of each curve. In the
+\emph{absolute} treatment, with sentence length denoted as \(n\), we
+computed the slope between positions \(1\) and \(2\), positions \(2\)
+and \(3\), positions \(3\) and \(n-2\), positions \(n-2\) and \(n-1\)
+and positions \(n-1\) and \(n\). For the short utterances appearing the
+CHILDES speech corpora we examined, these appeared to be important
+junctions in the distributions, with a seeming plateau in the middle of
+the unigram entropy curve for each of the language corpora we examined
+in CHILDES.
 
@@
 
-{[}Put in an illustration of the slope sections and treatments here?{]}
-
 \subsection{Results and Analysis}\label{results-and-analysis}
@@
 similarity and moreover for unsupervised clustering analysis to compare
-the results of the Wikipedia entropy analysis to known typological
-features within the langauges in the Wikipedia dataset. To determine
-which phonological, morphological and syntactic features affected the
-embedding of a language in the Wikipedia dataset, we looked at the list
-of \(144\) linguistic features in World Atlas of Language Structures
-(Dryer \& Haspelmath, 2013). We limited the languages in the WALS
-database to only those in our Wikipedia dataset and performed
-missing-value imputation to obtain the features not coded in WALS for
-the languages from Wikipedia.
+the results of our position-wise analysis on datasets from Wikipedia to
+known typological features within the languages. To determine which
+phonological, morphological and syntactic features affected the
+embedding of a language in the Wikipedia dataset, we use the linguistic
+features in World Atlas of Language Structures (Dryer \& Haspelmath,
+2013).
 
-One approach is to cluster the languages used in the Wikipedia analysis
-on the basis of WALS features and then directly compare the WALS
-clustering results with the results of the hierarchical clustering on
-the Wikipedia slope data for each of the different treatments using
-clustering similarity evaluation methods such as the Rand index (Rand,
-1971). However, the problem then arises of which combination of WALS
-features and how many features to include in the clustering analysis.
-This is a computationally intractable operation. Additionallly, deciding
-how many clusters to use in an unsupervised clustering analysis is an
-unsolved problem in machine learning.
+{[}Ideally do everything, but missing values in WALS, and how to compare
+every combination of features{]} An ideal approach using all of the
+corpora we can download from Wikipedia and all \(144\) features from the
+WALS database. Cluster the languages used in the Wikipedia analysis on
+the basis of WALS features and then directly compare the WALS clustering
+results with the results of the hierarchical clustering on the Wikipedia
+slope data for each of the different treatments using clustering
+similarity evaluation methods such as the Rand index (Rand, 1971).
+However, the problem then arises of which combination of WALS features
+and how many features to include in the clustering analysis. This is a
+computationally intractable operation. Additionallly, deciding how many
+clusters to use in an unsupervised clustering analysis is an unsolved
+problem in machine learning.
 
@@
 Cosine distance played some role in the feature determination. The top
-features were \(83\), \(81\), \(95\) and \(97\), which all characterize
-word order. This makes intuitive sense, and indicates that for this
-small subset of features and subset of languages that the embeddings in
-the slope space are related to the WALS features. Therefore typological
-features play a role in determining the positional entropy values for a
-language. This means that the entropy curves are in some way structured
-by the syntactic, morphological and phonological features of a language.
+four features all characterize word order, which indicates that for this
+subset of features and languages that the embeddings in the slope space
+are related to the WALS features. Therefore typological features play a
+role in determining the positional entropy values for a language. This
+means that the entropy curves are in some way structured by the
+syntactic, morphological and phonological features of a language.
 
-\section{Conclusions}\label{conclusions}
+\section{Discussion}\label{discussion}
 
-In this paper, we have characterized and applied a model based on Genzel
-\& Charniak (2002) for distinguishing entropy at each word position
-within sentences. In doing so, we have provided evidence against the
-Uniform Information Density hypothesis. Our work indicates that
-information is not distributed uniformly throughout an utterance
-regardless of languages, but that languages possess a characteristic
-information distribution determined by their syntactic, morphological
-and phonological features.
+In this paper, we have extended and applied a model from Yu et al.
+(2016) and derived from Genzel \& Charniak (2002) for distinguishing
+entropy at each word position within sentences. We have argued that the
+outcomes of this model are derived from the information distribution of
+a language. As this model varies cross-linguistically in terms of
 
-Our work has focused on words within sentences in both speech and text
-across typologically diverse languages. Other researchers have explored
-information rate cross-linguistically as a single number characterizing
-each language, most notably including Aylett \& Turk (2004). That work
-characterizes information rate as information per unit time, and is
-complementary to the work we accomplish here. Languages may possess both
-an overall information transfer rate, showing the average amount of
-semantic information transmitted per syllable or word, and an entropy
-distribution indicative of the language's information distribution.
+Follow-up with WALS features using missing-value imputation. However
+here are some of the problems with that.
 
-Research has indicated the effects of surprisal on fixation duration
-during eye-tracking studies (Demberg \& Keller, 2008) (Boston et al.,
-2008) (Smith \& Levy, 2013). Higher surprisal of a word is on average
-predictive of higher fixation duration. The \emph{wrap-up effect} states
-that when a person reads written text, he or she will process
-sentence-final words more slowly on average then sentence-medial or
-sentence-initial words, due to integrating information from the entire
-sentence to form a final understanding of the sentence's meaning (Stowe
-et al., 2018) (Kuperman et al., 2010). The wrap-up effect is drawn from
-evidence in English, German and Dutch, all of which are languages with a
-large final increase in their entropy curve from our study. We aim to
-see in the future if as a consequence of our work in this paper, the
-wrap-up effect falls out from the final jump in the entropy curves in
-some languages. We would analyze eye-tracking data from languages with
-low positional entropy in sentence-final positions. Since the lexical
-surprisal metric in eye-tracking and our positional entropy metric are
-both measures of log-probability surprisal, then in languages such as
-Hindi where the sentence-final positional entropy values are low we
-expect to find no wrap-up effect.
+Our work complements the approach of studies such as Aylett \& Turk
+(2004), where languages are characterized by a rate of semantic
+information per syllable. This body of work attests to languages
+possessing different rates of information transfer based on typological
+features as well.
+
+We also propose follow-up within linguistic processing using evidence
+from cross-linguistic eye-tracking. Research has indicated the effects
+of surprisal on fixation duration during eye-tracking studies (Boston,
+Hale, Kliegl, Patil, \& Vasishth, 2008; Demberg \& Keller, 2008; Smith
+\& Levy, 2013); higher surprisal of a word is on average predictive of
+higher fixation duration. The \emph{wrap-up effect} states that when a
+person reads written text, he or she will process sentence-final words
+more slowly on average then sentence-medial or sentence-initial words,
+due to integrating information from the entire sentence to form a final
+understanding of the sentence's meaning (Kuperman, Dambacher, Nuthmann,
+\& Kliegl, 2010; Stowe, Kaan, Sabourin, \& Taylor, 2018). The wrap-up
+effect is drawn from evidence in languages with a large final increase
+in their entropy curve from our study, so perhaps the wrap-up effect
+derives from the sentence-final increase in entropy curves observed in
+English and other Germanic languages.
 
@@
 
+\hypertarget{ref-aylett2004}{}
+Aylett, M., \& Turk, A. (2004). The smooth signal redundancy hypothesis:
+A functional explanation for relationships between redundancy, prosodic
+prominence, and duration in spontaneous speech. \emph{Language and
+Speech}, \emph{47}(1), 31--56.
+
+\hypertarget{ref-bai2008}{}
+Bai, X., Yan, G., Liversedge, S. P., Zang, C., \& Rayner, K. (2008).
+Reading spaced and unspaced chinese text: Evidence from eye movements.
+\emph{Journal of Experimental Psychology: Human Perception and
+Performance}, \emph{34}(5), 1277.
+
+\hypertarget{ref-boston2008}{}
+Boston, M. F., Hale, J., Kliegl, R., Patil, U., \& Vasishth, S. (2008).
+Parsing costs as predictors of reading difficulty: An evaluation using
+the potsdam sentence corpus. \emph{Journal of Eye Movement Research},
+\emph{2}(1).
+
 \hypertarget{ref-brown1973}{}
@@
 World}, 163--187.
+
+\hypertarget{ref-demberg2008}{}
+Demberg, V., \& Keller, F. (2008). Data from eye-tracking corpora as
+evidence for theories of syntactic processing complexity.
+\emph{Cognition}, \emph{109}(2), 193--210.
 
@@
 
+\hypertarget{ref-kuperman2010}{}
+Kuperman, V., Dambacher, M., Nuthmann, A., \& Kliegl, R. (2010). The
+effect of word position on eye-movements in sentence and paragraph
+reading. \emph{The Quarterly Journal of Experimental Psychology},
+\emph{63}(9), 1838--1857.
+
 \hypertarget{ref-levy2007}{}
@@
 
+\hypertarget{ref-shiro2000}{}
+Shiro, M. (2000). Diferencias sociales en la construcción del`` yo'' y
+del`` otro'': Expresiones evaluativas en la narrativa de niños
+caraqueños en edad escolar. In \emph{Lengua, discurso, texto: I simposio
+internacional de análisis del discurso} (pp. 1303--1318). Visor.
+
 \hypertarget{ref-simons2018}{}
@@
 Languages of the world, 21st edition}. Dallas, Texas: SIL International.
+
+\hypertarget{ref-smith2013}{}
+Smith, N. J., \& Levy, R. (2013). The effect of word predictability on
+reading time is logarithmic. \emph{Cognition}, \emph{128}(3), 302--319.
+
+\hypertarget{ref-stowe2018}{}
+Stowe, L. A., Kaan, E., Sabourin, L., \& Taylor, R. C. (2018). The
+sentence wrap-up dogma. \emph{Cognition}, \emph{176}, 232--247.
 

### 2019-01-30 process_bnc.Rmd

---
title: "process_bnc"
output: html_document
---

```{r load_libraries}
library(tidyverse)
library(stringr)
library(feather)
library(here)

NEW_DIR = here("BNC_parsed")
```

```{r parse_bnc, eval = F}  
files <- list.files(here("BNC/Texts"), "*.xml", full.names = T, recursive = T)

if(!dir.exists(NEW_DIR))
  dir.create(NEW_DIR)

parse_BNC <- function(file) {

  filename <- str_split(file, "/") %>% 
    unlist() %>% 
    last() %>% 
    str_split(., "\\.") %>%
    unlist() %>%
    first() %>%
    paste0(.,".txt")

  newname = paste0(NEW_DIR, "/", filename)

  print(file)

  system(paste0("xsltproc BNC/XML/Scripts/justTheWords.xsl ", file, " > ", newname))
}

map(files, parse_BNC)

write_feather(txt, "switchboard/switchboard.feather")
```

```{r read_parsed_bnc}
files <- list.files(NEW_DIR,  "*.txt", full.names = T)

txt <- map(files, read_lines) %>%
  unlist() %>%
  enframe(name = NULL) %>%
  filter(value != "") %>%
  mutate(value = str_trim(gsub('[[:punct:] ]+',' ', value)))

write_feather(txt, "bnc/bnc.feather")
```

```{r post_process_bnc}
bnc_tokens <- read_feather(here("../bnc/bnc.feather")) %>%
  mutate(length = str_count(value, pattern = " +") + 1) %>% 
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, value, token = stringr::str_split, pattern = " +") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n())

bnc_tokens %>%
  filter(length == 10) %>%
  write_feather(here("../bnc/bnc10.feather"))

bnc_tokens %>%
  filter(length == 20) %>%
  write_feather(here("../bnc/bnc20.feather"))

bnc_tokens %>%
  filter(length == 30) %>%
  write_feather(here("../bnc/bnc30.feather"))

```

### 2019-01-30 cogsci/position-info.Rmd
editing cogsci paper in progress

--- 2019-01-30 cogsci/position-info.Rmd
@@
 library(entropy)
+library(here)
+library(feather)
+library(magrittr)
 ```
@@
 
-```{r our_replication, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Providence corpus unigram entropy"}
-
-img <- png::readPNG("figs/Prov_corpus_result.png")
-grid::grid.raster(img)
-```
-
 We extend and generalize @yu2016 in three ways: we confirm that this same pattern is found in spoken English--in both adult and child speakers.
@@
 
-@yu2016 challenges the UID hypothesis through examining an entropy measure across sentence positions within the text portion of the English-language British National Corpus (BNC) @clear1993british.
-They partition the corpus by sentence length in number of words.
-For each word position $X$ of sentences of length $k$, they define $w$ as a unique word occurring in position $X$.
-They define $p(w)$ as the number of times word $w$ occurs in position $X$ divided by the number of total words that occur in position $X$ i.e.
+Here we will describe our adaptation of the by-word entropy method in @yu2016, which uses the text portion of the British National Corpus @clear1993british.
+We replicate their results at the end of this subsection.
+Given a text or speech corpus divided into individual utterances, we partition the corpus by utterance length in number of words.
+For each word position $X$ of utterances of length $k$, we define $w$ as a unique word occurring in position $X$.
+We further define $p(w)$ as the number of times word $w$ occurs in position $X$, divided by the number of total words that occur in position $X$ i.e.
 the number of sentences of length $k$.
-Then $p(w)$ is the probability of obtaining word $w$ by choosing a word at random in position $X$ in sentences of length $k$.
+This creates a probability distribution over the words occurring in position $X$, and computing the Shannon entropy $H(X)$ @shannon1948 of this probability distribution gives the positional entropy of position $X$ in utterances of length $k$.
 
@@
 
-With this measure, @yu2016 compute the unigram entropy at each position of sentences of each length within the corpus.
+Our replication of the @yu2016 results
+
+```{r read_and_plot_bnc}
+bnc_entropies <- read_csv(here("data/bnc_entropies.csv"))
+
+ggplot(bnc_entropies, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  ggtitle("Representative BNC Entropy Curves Faceted on Utterance Length") + 
+  xlab("Word Order within Utterance") +
+  ylab("Positional Entropy") +
+  geom_pointrange() +
+  geom_line()
+```
+
+With this measure, we compute the unigram entropy at each position of sentences of each length within the corpus.
 The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances.
 @genzel2002 similarly examine a unigram entropy measure on sentences, and found that entropy at the sentence level increases linearly with sentence index within a corpus.
-UID applies this uniformity of entropy rate in sentences to all levels of speech, and so the @yu2016, which examines text at the word level, should find an affine function at the word level.
+UID applies this uniformity of entropy rate in sentences to all levels of speech, and so our method obtained from @yu2016, which examines text at the word level, should find an affine function at the word level.
 
@@
 We first used the same source for data as @levy2007, the Switchboard corpus of American telephone conversations @godfrey1992.
+The Switchboard corpus contains telephone conversations of American English between strangers, with individual utterances ranging from one word to several dozen words.
+We divided the Switchboard corpus by utterance length and computed the positional entropy for each position within each sentence length, then plotted the result as an entropy curve.
 
-```{r switchboard, include=FALSE}
-library(feather)
+## Results and Analysis
 
-sb_tokens <- read_feather("../switchboard/switchboard.feather") %>%
-  mutate(length = str_count(value, pattern = " +") + 1) %>% 
-  mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, value, token = stringr::str_split, pattern = " +") %>%
-  group_by(utterance_id) %>%
-  mutate(word_order = 1:n())
+Our plots for three representative utterance lengths are below.
 
-unigram_entropy <- function(sen_length, tokens) {
-  tokens %>%
-    filter(length == sen_length) %>%
-    group_by(word_order, word) %>%
-    summarise(n = n()) %>%
-    tidyboot(summary_function = function(x) x %>% 
-               summarise(entropy = entropy(n, unit = "log2")),
-             statistics_functions = function(x) x %>%
-             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
-    mutate(length = sen_length)
-}
+```{r read_and_plot_switchboard}
+switchboard_entropies <- read_csv(here("data/switchboard_entropies.csv")) %>%
+  filter(length %in% c(5, 7, 9))
 
-entropies <- map(c(5, 7, 9), ~unigram_entropy(.x, sb_tokens)) %>%
-  bind_rows()
-
-ggplot(entropies, aes(x = word_order, y = empirical_entropy,
+ggplot(switchboard_entropies, aes(x = word_order, y = empirical_entropy,
                       ymin = ci_lower, ymax = ci_upper)) +
@@
 
-We used the Providence English corpus from CHILDES.
+After replicating the @yu2016 results using the spoken conversation corpus from @levy2007, we want to determine whether the entropy curves will remain robust over child-adult conversational corpora.
+
+The corpora in the CHILDES database consists of short and often disconnected utterances across hours of recordings, the unigram entropy measure is unaffected by context or lack thereof in utterances by the adults and children in the corpora.
+For our first corpus, we used the Providence English corpus from CHILDES @macwhinney2014.
 The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home.
-For speech data, which for the corpora in the CHILDES database consists of short and often disconnected utterances across hours of recordings, the unigram entropy measure is unaffected by context or lack thereof in utterances by the adults and children in the corpora.We divided the corpus by speaker into child and non-child categories.
+We divided the corpus by speaker into child and non-child categories.
 We further divided the corpus by utterance length, so that all sentences of length $k$ (e.g.
@@
 Finally, within each utterance length, we computed the unigram entropy measure for each position.
-Computation of positional entropy was identical to the Switchboard method after dividing for utterance length.
-
-```{r providence_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Providence corpus unigram entropy"}
-img <- png::readPNG("figs/Prov_corpus_result.png")
-grid::grid.raster(img)
-```
+Computation of positional entropy was identical to our general method after dividing for utterance length.
 
@@
 The Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Chinese writing including spaces between word boundaries can have a negative effect on reading times @bai2008.
+
+## Results and Analysis
+```{r providence_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Providence corpus unigram entropy"}
+img <- png::readPNG("figs/Prov_corpus_result.png")
+grid::grid.raster(img)
+```
 
@@
 
-## Results & Analysis
 The adult and child entropy curves track one another almost identically.
@@
 We use Wikipedia as a source for written data, which provides two advantages.
-One, the quantity of data in Wikipedia is large for each language and two, there are hundreds of languages with Wikipedias.
-This allows us to perform the entropy analysis on a much greater scale and to directly compare the results of the entropy analysis on each language to one another, and ultimately to predict what typological features of a language help determine its entropy curve, if any.
-We will describe our method of harvesting and distilling text data from each Wikipedia as well as how we compare the entropy curves from each language to one another.
+One, the quantity of data in Wikipedia is very large for each language, at least several megabytes, and two, there are hundreds of languages with their own Wikipedia corpora.
+We can therefore run our entropy analysis on a much greater scale than using small spoken CHILDES corpora and directly compare the results of the entropy analysis on each language to one another.
+We expect to find linguistic features that determine the shape of a language's entropy curve.
+We treated sentences in the Wikipedia text corpora as equivalent to utterances in the Switchboard and CHILDES corpora.
 
@@
 Using Giuseppe Attardi's Wikiextractor tool \footnote{https://github.com/attardi/wikiextractor}, we extract text corpora from Wikipedia by downloading a stored collection of Wikipedia entries in each langauge and randomly selecting several thousand articles from each Wikipedia language.
-Each language corpus was cleaned and limited to sentences between $6$ and $50$ words.
-Similar to our process for the Switchboard and CHILDES spoken corpora, we divided each corpus by sentence length, and then computed the unigram entropy measure on each word position within each sentence length.
-We wanted to directly compare and classify the unigram entropy curves of the languages from Wikipedia.
-We computed three slope treatments of each curve.
-In the *absolute* treatment, with sentence length denoted as $n$, we computed the slope between positions $1$ and $2$, positions $2$ and $3$, positions $3$ and $n-2$, positions $n-2$ and $n-1$ and positions $n-1$ and $n$.
-For the short utterances appearing the CHILDES speech corpora we examined, these appeared to be important junctions in the distributions, with a seeming plateau in the middle of the unigram entropy curve for each of the language corpora we examined in CHILDES.
+Each language corpus was cleaned and limited to sentences between six and $50$ words.
+Similar to our process for the Switchboard and CHILDES spoken corpora, we divided each corpus by sentence length, and then computed our positional entropy measure on each word position within each sentence length.
 
-However, because the portion of sentences of length greater than $10$ in the Wikipedia corpora were significantly larger than the CHILDES corpora, then we also computed relative slope treatments.
+We computed two slope treatments of each curve.
+In the *absolute* treatment, with sentence length denoted as $k$, we computed the slope between positions $1$ and $2$, positions $2$ and $3$, positions $3$ and $k-2$, positions $k-2$ and $k-1$ and positions $k-1$ and $k$.
+For the short utterances appearing the CHILDES speech corpora, the slopes between the first and second slopes on either end of the distributions appeared to be more characteristic of the distribution as a whole, with a plateau in the middle of the entropy curve for each of the language corpora we examined in CHILDES.
+
+However, sentences of length greater than ten in the Wikipedia corpora were significantly more common than in the CHILDES.
+Sentences on Wikipedia are commonly , then we also computed relative slope treatments.
 In the *relative 5* treatment, we computed the slopes between $0\%$ and $20\%$, $20\%$ and $40\%$, $40\%$ and $60\%$, $60\%$ and $80\%$ and $80\%$ and $100\%$ of the relative word positions in each sentence.
 When one of these percentages was not a whole number, then the closest whole number position was used instead for slope calculation.
-In the *relative 10* treatment, we computed the slopes between every $10\%$ of the relative word positions in each sentence.
-Each comparative slope within each treatment was averaged together between different sentence lengths, for example in the *relative 5* treatment then all of the $0\%$ to $20\%$ slopes were averaged together.
+Each comparative slope within each treatment was averaged together between different sentence lengths, for example in the relative $5$ treatment then all of the $0\%$ to $20\%$ slopes were averaged together.
 This created three treatments for the entropy curve for each langauge in the Wikipedia database.
+
+In both the absolute and relative $5$ treatments, each language is embedded in $5$-dimensional space.
+This allows for direct comparison between languages on Wikipedia using cosine similarity and moreover for unsupervised clustering analysis to compare the results of our position-wise analysis on datasets from Wikipedia to known typological features within the languages.
+To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we use the linguistic features in World Atlas of Language Structures (Dryer & Haspelmath, 2013).
+
+We check the effects of individual features on the embeddings of languages in the different treatments.
+We computed pairwise cosine similarity between each pair of language vectors within a treatment.
+For a subset of WALS features which had values entered for most of the languages we obtained from Wikipedia, we used a generalized linear model to see whether the cosine similarity between languages mattered in predicting if the languages shared the same value for a WALS feature.
 
@@
 
-In the *absolute* and *relative 5* treatments, each language is embedded in $5$-dimensional space.
-In the *relative 10* treatment, each language is embedded in $10$-dimensional space.
-This allows for direct comparison between languages on Wikipedia using cosine similarity and moreover for unsupervised clustering analysis to compare the results of our position-wise analysis on datasets from Wikipedia to known typological features within the languages.
-To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we use the linguistic features in World Atlas of Language Structures (Dryer & Haspelmath, 2013).
-
-[Ideally do everything, but missing values in WALS, and how to compare every combination of features] An ideal approach using all of the corpora we can download from Wikipedia and all $144$ features from the WALS database.
-Cluster the languages used in the Wikipedia analysis on the basis of WALS features and then directly compare the WALS clustering results with the results of the hierarchical clustering on the Wikipedia slope data for each of the different treatments using clustering similarity evaluation methods such as the Rand index (Rand, 1971).
-However, the problem then arises of which combination of WALS features and how many features to include in the clustering analysis.
-This is a computationally intractable operation.
-Additionallly, deciding how many clusters to use in an unsupervised clustering analysis is an unsolved problem in machine learning.
-
-We instead check the effects of individual features on the embeddings of languages in the different treatments.
-We computed pairwise cosine similarity between each pair of language vectors within a treatment.
-For a subset of WALS features which had values entered for most of the languages we obtained from Wikipedia, we used a generalized linear model to see whether the cosine similarity between languages mattered in predicting if the languages shared the same value for a WALS feature.
-For eight features and $45$ languages, we found this to be true.
-The table below shows the results for the generalized linear model we computed using WALS features and the absolute treatment.
+For eight features and $45$ languages, we found this to be true. The table below shows the results for the generalized linear model we computed using WALS features and the absolute treatment.
 
@@
 
+We also used the linear model approach to evaluate the effects of cosine distance in the relative $5$ treatment on the values of the same WALS features and found similar outcomes.
+
 Cosine distance played some role in the feature determination.
@@
 In this paper, we have extended and applied a model from @yu2016 and derived from @genzel2002 for distinguishing entropy at each word position within sentences.
-We have argued that the outcomes of this model are derived from the information distribution of a language.
-As this model varies cross-linguistically in terms of
+We have argued that the outcomes of this model are derived from the distribution of information transmission in human communication.
+Language is not a uniform noisy channel of communication, but each language individually represents a different noisy channel of communication characterized in part by typological features.
 
-Follow-up with WALS features using missing-value imputation. However here are some of the problems with that.
+Our work complements the approach of studies such as @aylett2004, where languages are characterized by a single number, representing the rate of semantic information transfer per syllable.
+This body of work attests to languages possessing different rates of information transfer based on typological features as well.
+Within the information transfer rate literature, the rate is an overall feature of language that does not vary corpus to corpus.
+Similarly, in our work we have obtained a characteristic entropy distribution for each language.
+Both information rate and entropy distribution are characterized by typological features.
+This indicates that the structure and rate of how people communicate information varies cross-linguistically based on typological features.
 
-Our work complements the approach of studies such as @aylett2004, where languages are characterized by a rate of semantic information per syllable.
-This body of work attests to languages possessing different rates of information transfer based on typological features as well.
+An ideal approach would use all of the corpora we can download from Wikipedia and all $144$ features from the WALS database.
+Cluster the languages used in the Wikipedia analysis on the basis of WALS features and then directly compare the WALS clustering results with the results of the hierarchical clustering on the Wikipedia slope data for each of the different treatments using clustering similarity evaluation methods such as the Rand index (Rand, 1971).
+However, the problem then arises of which combination of WALS features and how many features to include in the clustering analysis.
+This is a computationally intractable operation.
+Additionallly, deciding how many clusters to use in an unsupervised clustering analysis is an unsolved problem in machine learning.
 
-We also propose follow-up within linguistic processing using evidence from cross-linguistic eye-tracking.
-Research has indicated the effects of surprisal on fixation duration during eye-tracking studies [@demberg2008; @boston2008; @smith2013]; higher surprisal of a word is on average predictive of higher fixation duration.
+On a more mechanical level, we expect that our entropy model will have implications for how long people take to read individual words at different positions of a sentence.
+Eye-tracking research has indicated the effects of surprisal on fixation duration during eye-tracking studies [@demberg2008; @boston2008; @smith2013]; higher surprisal of a word is on average predictive of higher fixation duration.
 The *wrap-up effect* states that when a person reads written text, he or she will process sentence-final words more slowly on average then sentence-medial or sentence-initial words, due to integrating information from the entire sentence to form a final understanding of the sentence's meaning [@stowe2018; @kuperman2010].

### 2019-01-30 cogsci/position-info.tex
editing cogsci paper in progress

@@
 
-\begin{CodeChunk}
-\begin{figure*}[h]
-
-{\centering \includegraphics{figs/our_replication-1}
-
-}
-
-\caption[Providence corpus unigram entropy]{Providence corpus unigram entropy}\label{fig:our_replication}
-\end{figure*}
-\end{CodeChunk}
-
 We extend and generalize Yu et al. (2016) in three ways: we confirm that
@@
 computing over the utterances in the CHILDES corpora to the utterances
-in the adult-adult Switchboard corpus.
+in the adult-adult Switchboard corpus. Finally, we extend our
+cross-linguistic findings on a small number of languages from the
+CHILDES experiment to a direct comparison of several dozen languages in
+Wikipedia.
+
+\subsection{General methods}\label{general-methods}
+
+Here we will describe our adaptation of the by-word entropy method in Yu
+et al. (2016), which uses the text portion of the British National
+Corpus Clear (1993). We replicate their results at the end of this
+subsection. Given a text or speech corpus divided into individual
+utterances, we partition the corpus by utterance length in number of
+words. For each word position \(X\) of utterances of length \(k\), we
+define \(w\) as a unique word occurring in position \(X\). We further
+define \(p(w)\) as the number of times word \(w\) occurs in position
+\(X\), divided by the number of total words that occur in position \(X\)
+i.e.~the number of sentences of length \(k\). This creates a probability
+distribution over the words occurring in position \(X\), and computing
+the Shannon entropy \(H(X)\) Shannon (1948) of this probability
+distribution gives the positional entropy of position \(X\) in
+utterances of length \(k\).
+
+\[H(X) = \sum\limits_w p(w)\log\big(p(w)\big)\]
+
+Our replication of the Yu et al. (2016) results
+
+\begin{CodeChunk}
+
+\includegraphics{figs/read_and_plot_bnc-1} \end{CodeChunk}
+
+With this measure, we compute the unigram entropy at each position of
+sentences of each length within the corpus. The result of this method
+can be plotted for each utterance length as an \emph{entropy curve},
+which can be visually compared across utterance length to observe the
+how the unigram entropy changes across absolute positions in each of the
+utterances. Genzel \& Charniak (2002) similarly examine a unigram
+entropy measure on sentences, and found that entropy at the sentence
+level increases linearly with sentence index within a corpus. UID
+applies this uniformity of entropy rate in sentences to all levels of
+speech, and so our method obtained from Yu et al. (2016), which examines
+text at the word level, should find an affine function at the word
+level.
+
+The entropy curves capture individual variation across positions in
+utterances of the same length. This allows us to directly observe and
+judge the amount of variation in words that appear in an individual
+position of a sentence. We can directly compare any two positions within
+utterances to determine the amount of uncertainty, and therefore
+information, on average contained by words within that position of
+utterances. We are applying the same approach as in Genzel \& Charniak
+(2002), but within sentences instead of across sentences.
 
@@
 
-Yu et al. (2016) challenges the UID hypothesis through examining an
-entropy measure across sentence positions within the text portion of the
-English-language British National Corpus (BNC) Clear (1993). They
-partition the corpus by sentence length in number of words. For each
-word position \(X\) of sentences of length \(k\), they define \(w\) as a
-unique word occurring in position \(X\). They define \(p(w)\) as the
-number of times word \(w\) occurs in position \(X\) divided by the
-number of total words that occur in position \(X\) i.e.~the number of
-sentences of length \(k\). Then \(p(w)\) is the probability of obtaining
-word \(w\) by choosing a word at random in position \(X\) in sentences
-of length \(k\).
+We first used the same source for data as R. P. Levy \& Jaeger (2007),
+the Switchboard corpus of American telephone conversations Godfrey,
+Holliman, \& McDaniel (1992). The Switchboard corpus contains telephone
+conversations of American English between strangers, with individual
+utterances ranging from one word to several dozen words. We divided the
+Switchboard corpus by utterance length and
 
-\[H(X) = \sum\limits_w p(w)\log\big(p(w)\big)\] With this measure, Yu et
-al. (2016) compute the unigram entropy at each position of sentences of
-each length within the corpus. The result of this method can be plotted
-for each utterance length as an \emph{entropy curve}, which can be
-visually compared across utterance length to observe the how the unigram
-entropy changes across absolute positions in each of the utterances.
-Genzel \& Charniak (2002) similarly examine a unigram entropy measure on
-sentences, and found that entropy at the sentence level increases
-linearly with sentence index within a corpus. UID applies this
-uniformity of entropy rate in sentences to all levels of speech, and so
-the Yu et al. (2016), which examines text at the word level, should find
-an affine function at the word level.
+\subsection{Results and Analysis}\label{results-and-analysis}
+
+\begin{CodeChunk}
+
+\includegraphics{figs/read_and_plot_switchboard-1} \end{CodeChunk}
 
@@
 
-We used the Providence English corpus from CHILDES. The Providence
+After replicating the Yu et al. (2016) results using the spoken
+conversation corpus from R. P. Levy \& Jaeger (2007), we want to
+determine whether the entropy curves will remain robust over child-adult
+conversational corpora.
+
+The corpora in the CHILDES database consists of short and often
+disconnected utterances across hours of recordings, the unigram entropy
+measure is unaffected by context or lack thereof in utterances by the
+adults and children in the corpora. For our first corpus, we used the
+Providence English corpus from CHILDES MacWhinney (2014). The Providence
 corpus recorded interactions between children between 1 and 3 years old
@@
 were grouped together. Finally, within each utterance length, we
-computed the unigram entropy measure for each position.
-
-The entropy curves capture individual variation across positions in
-utterances of the same length. This allows us to directly observe and
-judge the amount of variation in words that appear in an individual
-position of a sentence. For speech data, which for the corpora in the
-CHILDES database consists of short and often disconnected utterances
-across hours of recordings, the unigram entropy measure is unaffected by
-context or lack thereof in utterances by the adults and children in the
-corpora. We can directly compare any two positions within utterances to
-determine the amount of uncertainty, and therefore information, on
-average contained by words within that position of utterances. We are
-applying the same approach as in Genzel \& Charniak (2002), but within
-sentences instead of across sentences.
-
-\begin{CodeChunk}
-\begin{figure*}[h]
-
-{\centering \includegraphics{figs/providence_PE-1}
-
-}
-
-\caption[Providence corpus unigram entropy]{Providence corpus unigram entropy}\label{fig:providence_PE}
-\end{figure*}
-\end{CodeChunk}
+computed the unigram entropy measure for each position. Computation of
+positional entropy was identical to our general method after dividing
+for utterance length.
 
@@
 Liversedge, Zang, \& Rayner (2008).
+
+\subsection{Results and Analysis}\label{results-and-analysis-1}
+
+\begin{CodeChunk}
+\begin{figure*}[h]
+
+{\centering \includegraphics{figs/providence_PE-1}
+
+}
+
+\caption[Providence corpus unigram entropy]{Providence corpus unigram entropy}\label{fig:providence_PE}
+\end{figure*}
+\end{CodeChunk}
 
@@
 \end{CodeChunk}
-
-\subsection{Results \& Analysis}\label{results-analysis}
 
@@
 then the closest whole number position was used instead for slope
-calculation. In the \emph{relative 10} treatment, we computed the slopes
-between every \(10\%\) of the relative word positions in each sentence.
-Each comparative slope within each treatment was averaged together
-between different sentence lengths, for example in the \emph{relative 5}
-treatment then all of the \(0\%\) to \(20\%\) slopes were averaged
-together. This created three treatments for the entropy curve for each
-langauge in the Wikipedia database.
+calculation. Each comparative slope within each treatment was averaged
+together between different sentence lengths, for example in the
+\emph{relative 5} treatment then all of the \(0\%\) to \(20\%\) slopes
+were averaged together. This created three treatments for the entropy
+curve for each langauge in the Wikipedia database.
 
-\subsection{Results and Analysis}\label{results-and-analysis}
-
-In the \emph{absolute} and \emph{relative 5} treatments, each language
-is embedded in \(5\)-dimensional space. In the \emph{relative 10}
-treatment, each language is embedded in \(10\)-dimensional space. This
-allows for direct comparison between languages on Wikipedia using cosine
-similarity and moreover for unsupervised clustering analysis to compare
-the results of our position-wise analysis on datasets from Wikipedia to
-known typological features within the languages. To determine which
+In both the \emph{absolute} and \emph{relative 5} treatments, each
+language is embedded in \(5\)-dimensional space. This allows for direct
+comparison between languages on Wikipedia using cosine similarity and
+moreover for unsupervised clustering analysis to compare the results of
+our position-wise analysis on datasets from Wikipedia to known
+typological features within the languages. To determine which
 phonological, morphological and syntactic features affected the
@@
 
-{[}Ideally do everything, but missing values in WALS, and how to compare
-every combination of features{]} An ideal approach using all of the
-corpora we can download from Wikipedia and all \(144\) features from the
-WALS database. Cluster the languages used in the Wikipedia analysis on
-the basis of WALS features and then directly compare the WALS clustering
-results with the results of the hierarchical clustering on the Wikipedia
-slope data for each of the different treatments using clustering
-similarity evaluation methods such as the Rand index (Rand, 1971).
-However, the problem then arises of which combination of WALS features
-and how many features to include in the clustering analysis. This is a
-computationally intractable operation. Additionallly, deciding how many
-clusters to use in an unsupervised clustering analysis is an unsolved
-problem in machine learning.
+An ideal approach would use all of the corpora we can download from
+Wikipedia and all \(144\) features from the WALS database. Cluster the
+languages used in the Wikipedia analysis on the basis of WALS features
+and then directly compare the WALS clustering results with the results
+of the hierarchical clustering on the Wikipedia slope data for each of
+the different treatments using clustering similarity evaluation methods
+such as the Rand index (Rand, 1971). However, the problem then arises of
+which combination of WALS features and how many features to include in
+the clustering analysis. This is a computationally intractable
+operation. Additionallly, deciding how many clusters to use in an
+unsupervised clustering analysis is an unsolved problem in machine
+learning.
 
@@
 predicting if the languages shared the same value for a WALS feature.
+
+\subsection{Results and Analysis}\label{results-and-analysis-2}
+
 For eight features and \(45\) languages, we found this to be true. The
@@
 
+\hypertarget{ref-godfrey1992}{}
+Godfrey, J. J., Holliman, E. C., \& McDaniel, J. (1992). SWITCHBOARD:
+Telephone speech corpus for research and development. In
+\emph{Acoustics, speech, and signal processing, 1992. icassp-92., 1992
+ieee international conference on} (Vol. 1, pp. 517--520). IEEE.
+
 \hypertarget{ref-jaeger2010}{}

### 2019-01-30 wiki_entropy/wiki_clustering.Rmd
working on Cogsci submission: bibtex drafted, clustering table here
editing cogsci paper in progress

@@
 ```{r}
-wiki_gs <- gs_title("Wikipedia_absolute")
+wiki_gs <- gs_title("Wikipedia_relative5")
 wkpd <- gs_read(wiki_gs) 
@@
 cosines <- wkpd %>%
+  column_to_rownames("Language") %>%
   t() %>%
@@
 ```
-
-```{r evaluate_clusters}
-wals_clusters <- wals_clust["cluster"] %>% unlist(use.names= FALSE)
-```

### 2019-01-30 cogsci/position-info.Rmd
cogsci full draft minus Yurovsky edits

@@
 
-Our replication of the @yu2016 results
+With this measure, we compute the unigram entropy at each position of sentences of each length within the corpus.
+The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances.
+@genzel2002 similarly examine a unigram entropy measure on sentences, and found that entropy at the sentence level increases linearly with sentence index within a corpus.
+UID applies this uniformity of entropy rate in sentences to all levels of speech, and so our method obtained from @yu2016, which examines text at the word level, should find an affine function at the word level.
+
+The entropy curves capture individual variation across positions in utterances of the same length.
+This allows us to directly observe and judge the amount of variation in words that appear in an individual position of a sentence.
+We can directly compare any two positions within utterances to determine the amount of uncertainty, and therefore information, on average contained by words within that position of utterances.
+We are applying the same approach as in @genzel2002, but within sentences instead of across sentences.
+
+We replicated the @yu2016 results and found the same characteristic three-step distribution as they did for the BNC entropy curves.
 
@@
 
-With this measure, we compute the unigram entropy at each position of sentences of each length within the corpus.
-The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances.
-@genzel2002 similarly examine a unigram entropy measure on sentences, and found that entropy at the sentence level increases linearly with sentence index within a corpus.
-UID applies this uniformity of entropy rate in sentences to all levels of speech, and so our method obtained from @yu2016, which examines text at the word level, should find an affine function at the word level.
-
-The entropy curves capture individual variation across positions in utterances of the same length.
-This allows us to directly observe and judge the amount of variation in words that appear in an individual position of a sentence.
-We can directly compare any two positions within utterances to determine the amount of uncertainty, and therefore information, on average contained by words within that position of utterances.
-We are applying the same approach as in @genzel2002, but within sentences instead of across sentences.
+The positional entropy at the beginning of sentences is low, then rises and plateaus for sentence-medial word positions before dropping slightly in the second-to-last position and rising again.
+However, this curve could simply be a result of the written words in the BNC.
+We will expand this analysis and replicate the result with a spoken American English corpus.
 
@@
 Our plots for three representative utterance lengths are below.
+We found the same characteristic shape for the American English Switchboard corpus as in our replication of the @yu2016 results for the BNC.
 
@@
 
+The shape we find in these two English data is notable for a number of reasons, but what we find most interesting is one, the shape of the distribution does not follow our predictions from UID and two, the distribution is robust across an unrelated pair of spoken and written English corpora.
+This suggests that the entropy curve is characteristic of the English language at least, and is not a specific effect of computing positional entropies in the written British English texts in the BNC.
+For our next step, we want to expand the positional entropy analysis to child-adult conversational corpora and languages beyond English to see if the same three-step distribution appears.
+
 # Experiment 2
@@
 After replicating the @yu2016 results using the spoken conversation corpus from @levy2007, we want to determine whether the entropy curves will remain robust over child-adult conversational corpora.
+We also wanted to see whether the curves would show the same characteristic three-step shape across languages.
+To accomplish both of these goals, we used data from the CHILDES Talkbank system @macwhinney2014.
 
 The corpora in the CHILDES database consists of short and often disconnected utterances across hours of recordings, the unigram entropy measure is unaffected by context or lack thereof in utterances by the adults and children in the corpora.
-For our first corpus, we used the Providence English corpus from CHILDES @macwhinney2014.
+For our first corpus, we used the Providence English corpus @demuth2006word.
 The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home.
@@
 We also ran this analysis on Spanish and Mandarin corpora from CHILDES.
+Spanish is an Indo-European language like English, possessing similar grammar, word order and numerous cognate word.
+Mandarin Chinese is typologically completed unrelated to English.
 We used the Shiro corpus for Spanish @shiro2000, which contains prompted narratives individually collected from over a hundred Venezualan schoolchildren, half from high SES backgrounds and half from low SES backgrounds.
-We used the Zhou dinner corpus for Mandarin Chinese (Li & Zhou, 2015), which contains dinner conversations between 5 to 6-year-old children and their parents collected in Shanghai.
+We used the Zhou dinner corpus for Mandarin Chinese @zhou2015dinner, which contains dinner conversations between 5 to 6-year-old children and their parents collected in Shanghai.
 
@@
 The Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Chinese writing including spaces between word boundaries can have a negative effect on reading times @bai2008.
+
+We expect that our entropy analysis on the Providence corpus will produce an identical entropy curve shape to the BNC and Switchboard curves.
+We expect similar results to English in the Spanish Shiro entropy curve, while we expect that our Mandarin corpus will produce a different entropy curve shape.
 
@@
 We found a distinct three-step distribution for English and Spanish CHILDES corpora, with a slight dip in the penultimate position of each sentence.
-The final position of utterances in child-directed speech is known to be important, dating back to Aslin (1993).
 The Mandarin corpus entropy curve, by comparison, has a noticeably lower positional entropy values in utterance-final positions than in utterance-penultimate positions.
-
 We attribute the penultimate dip in the English and Spanish entropy curves to the fact that most of the utterances in the CHILDES English and Spanish corpora we examined had a determiner such as "the" or "a" in the second-to-last position of utterances.
@@
 The Mandarin entropy curve, which does not at all resemble either the English/Spanish distribution or the predictions of UID, suggests that the entropy curve can vary from language to language.
-UID predicts that each language should have a similar distribution.
+UID predicts that each language should have a similar distribution, which should approximate an affine function.
+
+We found that all of the English corpora followed the same three-step entropy curve shape, regardless of whether the corpus contained spoken or written data.
+The Spanish corpus we used also followed this same shape, which we expected.
+However, Mandarin did not have the same shape, indicating that the shape of the entropy curves may arise from cross-linguistic variation.
+Similar grammatical and other features between two languages, such as Spanish and English, may produce a similar entropy curve.
 
@@
 However, sentences of length greater than ten in the Wikipedia corpora were significantly more common than in the CHILDES.
-Sentences on Wikipedia are commonly , then we also computed relative slope treatments.
-In the *relative 5* treatment, we computed the slopes between $0\%$ and $20\%$, $20\%$ and $40\%$, $40\%$ and $60\%$, $60\%$ and $80\%$ and $80\%$ and $100\%$ of the relative word positions in each sentence.
-When one of these percentages was not a whole number, then the closest whole number position was used instead for slope calculation.
-Each comparative slope within each treatment was averaged together between different sentence lengths, for example in the relative $5$ treatment then all of the $0\%$ to $20\%$ slopes were averaged together.
-This created three treatments for the entropy curve for each langauge in the Wikipedia database.
+Therefore we also computed relative slope treatments.
+In the *relative 5* treatment, we computed the slopes between every $20%$ of the relative word positions in each sentence length, e.g.
+the slope between $0\%$ and $20\%$ of sentences of length $10$ is the slope between the position entropies at the sentence-initial word position and the position representing the third word.
+When computing the word position index to make those cuts, if the slope was not a whole number, then the closest whole number position was used instead for slope calculation.
+Each comparative slope within each treatment was averaged together between different sentence lengths, for example all of the $0\%$ to $20\%$ slopes over all sentence lengths were averaged together.
 
 In both the absolute and relative $5$ treatments, each language is embedded in $5$-dimensional space.
-This allows for direct comparison between languages on Wikipedia using cosine similarity and moreover for unsupervised clustering analysis to compare the results of our position-wise analysis on datasets from Wikipedia to known typological features within the languages.
-To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we use the linguistic features in World Atlas of Language Structures (Dryer & Haspelmath, 2013).
-
-We check the effects of individual features on the embeddings of languages in the different treatments.
-We computed pairwise cosine similarity between each pair of language vectors within a treatment.
-For a subset of WALS features which had values entered for most of the languages we obtained from Wikipedia, we used a generalized linear model to see whether the cosine similarity between languages mattered in predicting if the languages shared the same value for a WALS feature.
+This permitted us to use cosine similarity for direct comparison between the languages we pulled from Wikipedia.
+To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we use the linguistic features in World Atlas of Language Structures @dryer2013wals.
+We checked the effects of individual features on the embeddings of languages in the different treatments.
+We computed pairwise cosine similarity between each pair of language vectors within each treatment.
+For a subset of eight WALS features, we used a generalized linear model to see whether the cosine similarity between languages mattered in predicting if the languages shared the same value for a WALS feature.
 
@@
 
-For eight features and $45$ languages, we found this to be true. The table below shows the results for the generalized linear model we computed using WALS features and the absolute treatment.
+For eight features and $45$ languages we pulled from Wikipedia, we found that the cosine similarity between WALS features.
+The table below shows the results for the generalized linear model we computed using WALS features and the absolute treatment.
 
@@
 
-Cosine distance played some role in the feature determination.
-The top four features all characterize word order, which indicates that for this subset of features and languages that the embeddings in the slope space are related to the WALS features.
-Therefore typological features play a role in determining the positional entropy values for a language.
-This means that the entropy curves are in some way structured by the syntactic, morphological and phonological features of a language.
+From these results, we argue that cosine distance played some role in the feature determination.
+The top four features in determining the WALS feature value all characterize word order, which indicates that for this subset of features and languages that the embeddings in the slope space are related to the WALS features.
+We argue therefore that typological features play a role in determining the positional entropy values for a language.
+This indicates that the entropy curves are in some way structured by the syntactic, morphological and phonological features of a language.
 
@@
 
-An ideal approach would use all of the corpora we can download from Wikipedia and all $144$ features from the WALS database.
-Cluster the languages used in the Wikipedia analysis on the basis of WALS features and then directly compare the WALS clustering results with the results of the hierarchical clustering on the Wikipedia slope data for each of the different treatments using clustering similarity evaluation methods such as the Rand index (Rand, 1971).
-However, the problem then arises of which combination of WALS features and how many features to include in the clustering analysis.
-This is a computationally intractable operation.
-Additionallly, deciding how many clusters to use in an unsupervised clustering analysis is an unsolved problem in machine learning.
+We initially wanted to use several hundred language corpora pulled from Wikipedia and all $144$ linguistic features from WALS.
+We considered an computing unsupervized clusterings of the $5$-dimensional vectors of our Wikipedia entropy curves, and comparing those clusterings to clusterings of the WALS features directly using a cluster similarity measure such as the Rand index @rand1971objective.
+However, the problem then arises of which combination of WALS features and how many features to include in the clustering analysis, with $144!$ different combinations.
+Two additional problems presented themselves: one, deciding how many clusters to use in an unsupervised clustering analysis is an unsolved problem in machine learning; and two, not all of the languages on Wikipedia have values inputted for their WALS features.
+As a follow-up, we are considering missing-data imputation on the WALS features in order to run the generalized linear model analysis on a large dataset of languages.
 
@@
 Eye-tracking research has indicated the effects of surprisal on fixation duration during eye-tracking studies [@demberg2008; @boston2008; @smith2013]; higher surprisal of a word is on average predictive of higher fixation duration.
-The *wrap-up effect* states that when a person reads written text, he or she will process sentence-final words more slowly on average then sentence-medial or sentence-initial words, due to integrating information from the entire sentence to form a final understanding of the sentence's meaning [@stowe2018; @kuperman2010].
-The wrap-up effect is drawn from evidence in languages with a large final increase in their entropy curve from our study, so perhaps the wrap-up effect derives from the sentence-final increase in entropy curves observed in English and other Germanic languages.
+The *wrap-up effect* states that people process sentence-final words more slowly on average then sentence-medial or sentence-initial words when reading, due to integrating information to form a final understanding of the sentence's meaning [@stowe2018; @kuperman2010].
+The wrap-up effect is drawn from evidence in languages with a large final increase in their entropy curve from our study, so we hypothesize that the supposed wrap-up effect derives from the same source sentence-final increase in entropy curves observed in English and other Germanic languages.
 

### 2019-01-30 cogsci/position-info.Rmd
entropy code in cogsci rmd for plots

@@
 ## Results and Analysis
+
+```{r entropy generation code}
+prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
+shiro_utterances <- get_utterances(corpus = "Shiro") #use gloss
+zhou_utterances <- get_utterances(corpus = "ZhouDinner") #use stem, NOT gloss
+
+get_entropies <- function(role, sen_length, utterances, stemgloss) {
+
+  if(role == "Target_Child") {
+    sub_utterances <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(length = str_count(stemgloss, pattern = " +") + 1) # uses stems of words, not original words
+  } else {
+     sub_utterances <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      mutate(length = str_count(stemgloss, pattern = " +") + 1)
+  }
+
+  tokens <- sub_utterances %>%
+    filter(length == sen_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, gloss) %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(word_order, word) %>%
+    filter(word_order <= length)
+
+  tokens %>%
+    summarise(n = n()) %>%
+    tidyboot(summary_function = function(x) x %>% 
+               summarise(entropy = entropy(n, unit = "log2")),
+             statistics_functions = function(x) x %>%
+             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
+    mutate(role = role, length = sen_length)
+}
+```
+
 ```{r providence_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Providence corpus unigram entropy"}

### 2019-01-31 cogsci/position-info.Rmd
childes plot no labels yet

--- 2019-01-30 cogsci/position-info.Rmd
@@
 library(magrittr)
+library(childesr)
+library(ggthemes)
+library(gridExtra)
+library(directlabels)
+
+theme_set(theme_classic(base_size = 10))
 ```
@@
 The original evidence in @levy2007 finds that the insertion of complementizers in relative clauses in English corresponds to where neighboring words have high information content.
-Similarly, @frank2008speaking argues that contradictions in English such as "you're" do not occur when neighboring words are highly informative.
+Similarly, @frank2008 argues that contradictions in English such as "you're" do not occur when neighboring words are highly informative.
 The evidence in favor of UID largely been situation-specific and English-language driven, while the hypothesis itself has been applied broadly over the past decade.
@@
 We begin with the same source as @levy2007 and @jaeger2010, the Switchboard corpus of adult telephone conversations.
-We also examine child-adult conversations in the CHILDES TalkBank [@brown1973; @macwhinney2014] corpora database of spoken adult-child conversations in multiple languages.
+We also examine child-adult conversations in the CHILDES TalkBank [@demuth2006; @macwhinney2014] corpora database of spoken adult-child conversations in multiple languages.
 Children are not fully developed speakers, so we want to compare the entropy curve we obtain by computing over the utterances in the CHILDES corpora to the utterances in the adult-adult Switchboard corpus.
@@
 
-Here we will describe our adaptation of the by-word entropy method in @yu2016, which uses the text portion of the British National Corpus @clear1993british.
+Here we will describe our adaptation of the by-word entropy method in @yu2016, which uses the text portion of the British National Corpus [@clear1993].
 We replicate their results at the end of this subsection.
@@
 
-We replicated the @yu2016 results and found the same characteristic three-step distribution as they did for the BNC entropy curves.
-
-```{r read_and_plot_bnc}
-bnc_entropies <- read_csv(here("data/bnc_entropies.csv"))
-
-ggplot(bnc_entropies, aes(x = word_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  ggtitle("Representative BNC Entropy Curves Faceted on Utterance Length") + 
-  xlab("Word Order within Utterance") +
-  ylab("Positional Entropy") +
-  geom_pointrange() +
-  geom_line()
-```
-
 The positional entropy at the beginning of sentences is low, then rises and plateaus for sentence-medial word positions before dropping slightly in the second-to-last position and rising again.
@@
 # Experiment 1
+
+```{r entropy_func}
+unigram_entropy <- function(sen_length, tokens) {
+  tokens %>%
+    filter(length == sen_length) %>%
+    group_by(word_order, word) %>%
+    summarise(n = n()) %>%
+    tidyboot(summary_function = function(x) x %>% 
+               summarise(entropy = entropy(n, unit = "log2")),
+             statistics_functions = function(x) x %>%
+             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
+    mutate(length = sen_length)
+}
+```
+
+```{r bnc_and_switchboard, eval = F}
+
+sb_tokens <- read_feather(here("../switchboard/switchboard.feather")) %>%
+  mutate(length = str_count(value, pattern = " +") + 1) %>% 
+  mutate(utterance_id = 1:n()) %>%
+  unnest_tokens(word, value, token = stringr::str_split, pattern = " +") %>%
+  group_by(utterance_id) %>%
+  mutate(word_order = 1:n())
+
+bnc10_tokens <- read_feather(here("../bnc/bnc10.feather")) %>%
+  mutate(word = as.numeric(as.factor(word)))
+bnc20_tokens <- read_feather(here("../bnc/bnc20.feather")) %>%
+  mutate(word = as.numeric(as.factor(word)))
+bnc30_tokens <- read_feather(here("../bnc/bnc30.feather")) %>%
+  mutate(word = as.numeric(as.factor(word)))
+
+switchboard_entropies <- map_df(2:10, ~unigram_entropy(.x, sb_tokens))
+
+bnc10_entropies <- unigram_entropy(10, bnc10_tokens)
+bnc20_entropies <- unigram_entropy(20, bnc20_tokens)
+bnc30_entropies <- unigram_entropy(30, bnc30_tokens)
+bnc_entropies <- bind_rows(bnc10_entropies, bnc20_entropies, bnc30_entropies)
+
+write_csv(bnc_entropies, here("data/bnc_entropies.csv"))
+write_csv(switchboard_entropies, here("data/switchboard_entropies.csv"))
+```
+
+We replicated the @yu2016 results and found the same characteristic three-step distribution as they did for the BNC entropy curves.
 
@@
 
-```{r read_and_plot_switchboard}
+```{r read_and_plot_switchboard, fig.env = "figure*", fig.width=7, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom).
+Points average entropies, error bars show 95\\% confidence intervals computed by non-parametric bootstrap."}
+
+bnc_entropies <- read_csv(here("data/bnc_entropies.csv"))
+
 switchboard_entropies <- read_csv(here("data/switchboard_entropies.csv")) %>%
@@
 
-ggplot(switchboard_entropies, aes(x = word_order, y = empirical_entropy,
+bnc_plot <- ggplot(bnc_entropies, aes(x = word_order, y = empirical_entropy,
                       ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  ggtitle("Representative Switchboard Entropy Curves Faceted on Utterance Length") + 
-  xlab("Word Order within Utterance") +
-  ylab("Positional Entropy") +
-  scale_x_discrete(limits = "1":"9") + 
-  geom_pointrange() +
-  geom_smooth(se = F)
+  facet_wrap(~ length, scales = "free_x") + 
+  xlab("British National Corpus utterance position") +
+  ylab("Positional entropy") +
+  geom_pointrange(size = .25) +
+  geom_line(size = .25) +
+  theme_classic(base_size = 10)
+
+sw_plot <- ggplot(switchboard_entropies, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length, scales = "free_x") + 
+  xlab("Switchboard utterance position") +
+  ylab("Positional entropy") +
+  geom_pointrange(size = .25) +
+  geom_line(size = .25) +
+  theme_classic(base_size = 10)
+
+grid.arrange(bnc_plot, sw_plot, nrow = 2)
 ```
@@
 The corpora in the CHILDES database consists of short and often disconnected utterances across hours of recordings, the unigram entropy measure is unaffected by context or lack thereof in utterances by the adults and children in the corpora.
-For our first corpus, we used the Providence English corpus @demuth2006word.
+For our first corpus, we used the Providence English corpus @demuth2006.
 The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home.
@@
 We used the Shiro corpus for Spanish @shiro2000, which contains prompted narratives individually collected from over a hundred Venezualan schoolchildren, half from high SES backgrounds and half from low SES backgrounds.
-We used the Zhou dinner corpus for Mandarin Chinese @zhou2015dinner, which contains dinner conversations between 5 to 6-year-old children and their parents collected in Shanghai.
+We used the Zhou dinner corpus for Mandarin Chinese @li2015, which contains dinner conversations between 5 to 6-year-old children and their parents collected in Shanghai.
 
@@
 ## Results and Analysis
-```{r providence_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Providence corpus unigram entropy"}
-img <- png::readPNG("figs/Prov_corpus_result.png")
-grid::grid.raster(img)
+
+```{r childes_entropy, eval = F}
+
+extract_stem_tokens <- function(role, min_length, max_length, utterances) {
+
+  if (role == "child") {
+    lengths <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(length = str_count(stem, " ") + 1)
+
+} else {
+    lengths <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      mutate(length = str_count(stem, " ") + 1)
+}
+
+  tokens <- lengths %>% 
+    filter(length >= min_length) %>%
+    filter(length <= max_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, stem, token = stringr::str_split, pattern = " +") %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(length, word_order, word) %>%
+    filter(word_order <= length) %>% 
+    summarise(n = n()) 
+}
+
+extract_gloss_tokens <- function(role,  min_length, max_length,  utterances) {
+
+  if (role == "child") {
+    lengths <- utterances %>%
+      filter(speaker_role == "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1)
+
+} else {
+    lengths <- utterances %>%
+      filter(speaker_role != "Target_Child") %>%
+      mutate(length = str_count(gloss, " ") + 1)
+}
+
+  tokens <- lengths %>% 
+    filter(length >= min_length) %>%
+    filter(length <= max_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(length, word_order, word) %>%
+    filter(word_order <= length) %>% 
+    summarise(n = n()) 
+}
+
+prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
+shiro_utterances <- get_utterances(corpus = "Shiro") #use gloss
+zhou_utterances <- get_utterances(corpus = "ZhouDinner") #use stem, NOT gloss
+
+prov_child_tokens <- extract_stem_tokens("child", 5, 9, prov_utterances) 
+prov_adult_tokens <- extract_stem_tokens("not child",5, 9, prov_utterances)
+
+prov_child_entropies <-  map_df(c(5,7,9), ~unigram_entropy(.x, prov_child_tokens)) %>%
+  mutate(language = "English", person = "Child")
+prov_adult_entropies <-  map_df(c(5,7,9), ~unigram_entropy(.x, prov_adult_tokens)) %>%
+  mutate(language = "English", person = "Adult")
+
+shiro_child_tokens <- extract_gloss_tokens("child", 5, 9, shiro_utterances) 
+shiro_adult_tokens <- extract_gloss_tokens("not child",5, 9, shiro_utterances)
+
+shiro_child_entropies <-  map_df(c(5,7,9), ~unigram_entropy(.x, shiro_child_tokens)) %>%
+  mutate(language = "Spanish", person = "Child")
+shiro_adult_entropies <-  map_df(c(5,7,9), ~unigram_entropy(.x, shiro_adult_tokens)) %>%
+  mutate(language = "Spanish", person = "Adult")
+
+zhou_child_tokens <- extract_stem_tokens("child", 5, 9, zhou_utterances) 
+zhou_adult_tokens <- extract_stem_tokens("not child",5, 9, zhou_utterances)
+
+zhou_child_entropies <-  map_df(c(5,7,9), ~unigram_entropy(.x, zhou_child_tokens)) %>%
+  mutate(language = "Mandarin", person = "Child")
+zhou_adult_entropies <-  map_df(c(5,7,9), ~unigram_entropy(.x, zhou_adult_tokens)) %>%
+  mutate(language = "Mandarin", person = "Adult")
+
+childes_entropies <- bind_rows(prov_child_entropies, prov_adult_entropies, 
+                               shiro_child_entropies, shiro_adult_entropies,
+                               zhou_child_entropies, zhou_adult_entropies)
+
+write_csv(childes_entropies, here("data/childes_entropies.csv"))
 ```
 
-```{r shiro_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Shiro corpus entropy"}
-img <- png::readPNG("figs/Spanish_unigram_entropies_with_stemming.png")
-grid::grid.raster(img)
-```
+```{r plot_childes, fig.env = "figure*", fig.width=7, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Representative Entropy Curves for Three Childes Corpora in English, Mandarin, and Spanish.
+Points average entropies, error bars show 95\\% confidence intervals computed by non-parametric bootstrap."}
+childes_entropies <- read_csv(here("data/childes_entropies.csv"))
 
-```{r zhou_PE, fig.env = "figure*", fig.pos = "h", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Zhou Dinner corpus entropy"}
-img <- png::readPNG("figs/Mandarin_unigram_entropies.png")
-grid::grid.raster(img)
+ggplot(childes_entropies, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper, color = person,
+                      label = person)) +
+  facet_grid(language ~ length, scales = "free") + 
+  xlab("Utterance position") +
+  ylab("Positional entropy") +
+  geom_pointrange(size = .25) +
+  geom_line(size = .25) +
+  theme_classic(base_size = 10) + 
+  scale_color_ptol() + 
+  theme(legend.position = "none") + 
+  geom_dl(method = list(dl.trans(x=x +2.2), "last.qp", cex=.7)) 
 ```
@@
 Therefore we also computed relative slope treatments.
-In the *relative 5* treatment, we computed the slopes between every $20%$ of the relative word positions in each sentence length, e.g.
+In the *relative 5* treatment, we computed the slopes between every $20\%$ of the relative word positions in each sentence length, e.g.
 the slope between $0\%$ and $20\%$ of sentences of length $10$ is the slope between the position entropies at the sentence-initial word position and the position representing the third word.
@@
 This permitted us to use cosine similarity for direct comparison between the languages we pulled from Wikipedia.
-To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we use the linguistic features in World Atlas of Language Structures @dryer2013wals.
+To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we use the linguistic features in World Atlas of Language Structures @dryer2013.
 We checked the effects of individual features on the embeddings of languages in the different treatments.
@@
 We initially wanted to use several hundred language corpora pulled from Wikipedia and all $144$ linguistic features from WALS.
-We considered an computing unsupervized clusterings of the $5$-dimensional vectors of our Wikipedia entropy curves, and comparing those clusterings to clusterings of the WALS features directly using a cluster similarity measure such as the Rand index @rand1971objective.
+We considered an computing unsupervized clusterings of the $5$-dimensional vectors of our Wikipedia entropy curves, and comparing those clusterings to clusterings of the WALS features directly using a cluster similarity measure such as the Rand index @rand1971.
 However, the problem then arises of which combination of WALS features and how many features to include in the clustering analysis, with $144!$ different combinations.

### 2019-01-31 cogsci/position-info.tex
childes plot no labels yet

@@
 We argue for a reformulation of the Uniform Information Density
-hypothesis R. P. Levy \& Jaeger (2007) in which the grammatical,
-phonological and morphological features of a language determine how
-information is distributed throughout utterances. We expand on a method
-from Yu, Cong, Liang, \& Liu (2016) to compute entropy based on word
-position within utterances as a derivation of information distribution.
-We analyze adult and child spoken corpora as well as text corpora drawn
-from a variety of languages on Wikipedia. We find that each language has
-a robust, characteristic entropy curve across word position, and present
-evidence that these entropy curves are in part determined by typological
+hypothesis Levy \& Jaeger (2007) in which the grammatical, phonological
+and morphological features of a language determine how information is
+distributed throughout utterances. We expand on a method from Yu, Cong,
+Liang, \& Liu (2016) to compute entropy based on word position within
+utterances as a derivation of information distribution. We analyze adult
+and child spoken corpora as well as text corpora drawn from a variety of
+languages on Wikipedia. We find that each language has a robust,
+characteristic entropy curve across word position, and present evidence
+that these entropy curves are in part determined by typological
 features. We conclude that speakers may tend towards communicating
@@
 
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 the recent development of such theories in human and machine language
-processing (Jelinek, 1976; R. P. Levy \& Jaeger, 2007).
+processing (Jelinek, 1976; Levy \& Jaeger, 2007).
 
@@
 words in a sentence, i.e.~maintain \emph{uniform information density}
-(Genzel \& Charniak, 2002; R. P. Levy \& Jaeger, 2007).
+(Genzel \& Charniak, 2002; Levy \& Jaeger, 2007).
 
-The original evidence in R. P. Levy \& Jaeger (2007) finds that the
-insertion of complementizers in relative clauses in English corresponds
-to where neighboring words have high information content. Similarly,
-Frank \& Jaeger (2008) argues that contradictions in English such as
-``you're'' do not occur when neighboring words are highly informative.
-The evidence in favor of UID largely been situation-specific and
-English-language driven, while the hypothesis itself has been applied
-broadly over the past decade. Applications include determining whether
-linguistic alignment takes place (Jaeger \& Snider, 2013), Zipfian word
-length distributions (Piantadosi, Tily, \& Gibson, 2011), communication
+The original evidence in Levy \& Jaeger (2007) finds that the insertion
+of complementizers in relative clauses in English corresponds to where
+neighboring words have high information content. Similarly, Frank \&
+Jaeger (2008) argues that contradictions in English such as ``you're''
+do not occur when neighboring words are highly informative. The evidence
+in favor of UID largely been situation-specific and English-language
+driven, while the hypothesis itself has been applied broadly over the
+past decade. Applications include determining whether linguistic
+alignment takes place (Jaeger \& Snider, 2013), Zipfian word length
+distributions (Piantadosi, Tily, \& Gibson, 2011), communication
 efficiency (Mahowald, Fedorenko, Piantadosi, \& Gibson, 2013), dialogue
@@
 However, other recent work has contradicted the UID hypothesis. Similar
-to the original R. P. Levy \& Jaeger (2007), Zhan \& Levy (2018) focuses
-on information distribution at particular points in sentences. Zhan \&
-Levy (2018) finds that more information-rich classifiers in Mandarin
-Chinese are produced when production of the neighboring noun is
-difficult, not when the information content is high. Jain, Singh,
-Ranjan, Rajkumar, \& Agarwal (2018) examine word order across spoken
-sentences in Hindi, a freer word order language than English, and find
-that information density has no significant effect on word order.
+to the original Levy \& Jaeger (2007), Zhan \& Levy (2018) focuses on
+information distribution at particular points in sentences. Zhan \& Levy
+(2018) finds that more information-rich classifiers in Mandarin Chinese
+are produced when production of the neighboring noun is difficult, not
+when the information content is high. Jain, Singh, Ranjan, Rajkumar, \&
+Agarwal (2018) examine word order across spoken sentences in Hindi, a
+freer word order language than English, and find that information
+density has no significant effect on word order.
 
@@
 to determine if we obtain the entropy curve shape we expect for spoken
-communication. We begin with the same source as R. P. Levy \& Jaeger
-(2007) and Jaeger (2010), the Switchboard corpus of adult telephone
+communication. We begin with the same source as Levy \& Jaeger (2007)
+and Jaeger (2010), the Switchboard corpus of adult telephone
 conversations. We also examine child-adult conversations in the CHILDES
-TalkBank (Brown, 1973; MacWhinney, 2014) corpora database of spoken
-adult-child conversations in multiple languages. Children are not fully
-developed speakers, so we want to compare the entropy curve we obtain by
-computing over the utterances in the CHILDES corpora to the utterances
-in the adult-adult Switchboard corpus. Finally, we extend our
-cross-linguistic findings on a small number of languages from the
-CHILDES experiment to a direct comparison of several dozen languages in
-Wikipedia.
+TalkBank (Demuth, Culbertson, \& Alter, 2006; MacWhinney, 2014) corpora
+database of spoken adult-child conversations in multiple languages.
+Children are not fully developed speakers, so we want to compare the
+entropy curve we obtain by computing over the utterances in the CHILDES
+corpora to the utterances in the adult-adult Switchboard corpus.
+Finally, we extend our cross-linguistic findings on a small number of
+languages from the CHILDES experiment to a direct comparison of several
+dozen languages in Wikipedia.
 
-\subsection{General methods}\label{general-methods}
+\hypertarget{general-methods}{%
+\subsection{General methods}\label{general-methods}}
 
@@
 et al. (2016), which uses the text portion of the British National
-Corpus Clear (1993). We replicate their results at the end of this
+Corpus (Clear, 1993). We replicate their results at the end of this
 subsection. Given a text or speech corpus divided into individual
@@
 \[H(X) = \sum\limits_w p(w)\log\big(p(w)\big)\]
-
-Our replication of the Yu et al. (2016) results
-
-\begin{CodeChunk}
-
-\includegraphics{figs/read_and_plot_bnc-1} \end{CodeChunk}
 
@@
 
-\section{Experiment 1}\label{experiment-1}
+The positional entropy at the beginning of sentences is low, then rises
+and plateaus for sentence-medial word positions before dropping slightly
+in the second-to-last position and rising again. However, this curve
+could simply be a result of the written words in the BNC. We will expand
+this analysis and replicate the result with a spoken American English
+corpus.
 
-We first used the same source for data as R. P. Levy \& Jaeger (2007),
-the Switchboard corpus of American telephone conversations Godfrey,
+\hypertarget{experiment-1}{%
+\section{Experiment 1}\label{experiment-1}}
+
+We replicated the Yu et al. (2016) results and found the same
+characteristic three-step distribution as they did for the BNC entropy
+curves.
+
+We first used the same source for data as Levy \& Jaeger (2007), the
+Switchboard corpus of American telephone conversations Godfrey,
 Holliman, \& McDaniel (1992). The Switchboard corpus contains telephone
@@
 utterances ranging from one word to several dozen words. We divided the
-Switchboard corpus by utterance length and
+Switchboard corpus by utterance length and computed the positional
+entropy for each position within each sentence length, then plotted the
+result as an entropy curve.
 
-\subsection{Results and Analysis}\label{results-and-analysis}
+\hypertarget{results-and-analysis}{%
+\subsection{Results and Analysis}\label{results-and-analysis}}
+
+Our plots for three representative utterance lengths are below. We found
+the same characteristic shape for the American English Switchboard
+corpus as in our replication of the Yu et al. (2016) results for the
+BNC.
 
 \begin{CodeChunk}
+\begin{figure*}[tb]
 
-\includegraphics{figs/read_and_plot_switchboard-1} \end{CodeChunk}
+{\centering \includegraphics{figs/read_and_plot_switchboard-1}
 
-\section{Experiment 2}\label{experiment-2}
+}
+
+\caption[Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom)]{Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom).
+Points average entropies, error bars show 95\% confidence intervals computed by non-parametric bootstrap.}\label{fig:read_and_plot_switchboard}
+\end{figure*}
+\end{CodeChunk}
+
+The shape we find in these two English data is notable for a number of
+reasons, but what we find most interesting is one, the shape of the
+distribution does not follow our predictions from UID and two, the
+distribution is robust across an unrelated pair of spoken and written
+English corpora. This suggests that the entropy curve is characteristic
+of the English language at least, and is not a specific effect of
+computing positional entropies in the written British English texts in
+the BNC. For our next step, we want to expand the positional entropy
+analysis to child-adult conversational corpora and languages beyond
+English to see if the same three-step distribution appears.
+
+\hypertarget{experiment-2}{%
+\section{Experiment 2}\label{experiment-2}}
 
 After replicating the Yu et al. (2016) results using the spoken
-conversation corpus from R. P. Levy \& Jaeger (2007), we want to
-determine whether the entropy curves will remain robust over child-adult
-conversational corpora.
+conversation corpus from Levy \& Jaeger (2007), we want to determine
+whether the entropy curves will remain robust over child-adult
+conversational corpora. We also wanted to see whether the curves would
+show the same characteristic three-step shape across languages. To
+accomplish both of these goals, we used data from the CHILDES Talkbank
+system MacWhinney (2014).
 
@@
 adults and children in the corpora. For our first corpus, we used the
-Providence English corpus from CHILDES MacWhinney (2014). The Providence
-corpus recorded interactions between children between 1 and 3 years old
-and their parents in the home. We divided the corpus by speaker into
-child and non-child categories. We further divided the corpus by
-utterance length, so that all sentences of length \(k\) (e.g. \(6\))
-were grouped together. Finally, within each utterance length, we
-computed the unigram entropy measure for each position. Computation of
-positional entropy was identical to our general method after dividing
-for utterance length.
+Providence English corpus Demuth et al. (2006). The Providence corpus
+recorded interactions between children between 1 and 3 years old and
+their parents in the home. We divided the corpus by speaker into child
+and non-child categories. We further divided the corpus by utterance
+length, so that all sentences of length \(k\) (e.g. \(6\)) were grouped
+together. Finally, within each utterance length, we computed the unigram
+entropy measure for each position. Computation of positional entropy was
+identical to our general method after dividing for utterance length.
 
 We also ran this analysis on Spanish and Mandarin corpora from CHILDES.
-We used the Shiro corpus for Spanish Shiro (2000), which contains
-prompted narratives individually collected from over a hundred
-Venezualan schoolchildren, half from high SES backgrounds and half from
-low SES backgrounds. We used the Zhou dinner corpus for Mandarin Chinese
-(Li \& Zhou, 2015), which contains dinner conversations between 5 to
-6-year-old children and their parents collected in Shanghai.
+Spanish is an Indo-European language like English, possessing similar
+grammar, word order and numerous cognate word. Mandarin Chinese is
+typologically completed unrelated to English. We used the Shiro corpus
+for Spanish Shiro (2000), which contains prompted narratives
+individually collected from over a hundred Venezualan schoolchildren,
+half from high SES backgrounds and half from low SES backgrounds. We
+used the Zhou dinner corpus for Mandarin Chinese Li \& Zhou (2015),
+which contains dinner conversations between 5 to 6-year-old children and
+their parents collected in Shanghai.
 
@@
 
-\subsection{Results and Analysis}\label{results-and-analysis-1}
+We expect that our entropy analysis on the Providence corpus will
+produce an identical entropy curve shape to the BNC and Switchboard
+curves. We expect similar results to English in the Spanish Shiro
+entropy curve, while we expect that our Mandarin corpus will produce a
+different entropy curve shape.
+
+\hypertarget{results-and-analysis-1}{%
+\subsection{Results and Analysis}\label{results-and-analysis-1}}
 
 \begin{CodeChunk}
-\begin{figure*}[h]
+\begin{figure*}[tb]
 
-{\centering \includegraphics{figs/providence_PE-1}
+{\centering \includegraphics{figs/plot_childes-1}
 
@@
 
-\caption[Providence corpus unigram entropy]{Providence corpus unigram entropy}\label{fig:providence_PE}
-\end{figure*}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{figure*}[h]
-
-{\centering \includegraphics{figs/shiro_PE-1}
-
-}
-
-\caption[Shiro corpus entropy]{Shiro corpus entropy}\label{fig:shiro_PE}
-\end{figure*}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{figure*}[h]
-
-{\centering \includegraphics{figs/zhou_PE-1}
-
-}
-
-\caption[Zhou Dinner corpus entropy]{Zhou Dinner corpus entropy}\label{fig:zhou_PE}
+\caption[Representative Entropy Curves for Three Childes Corpora in English, Mandarin, and Spanish]{Representative Entropy Curves for Three Childes Corpora in English, Mandarin, and Spanish.
+Points average entropies, error bars show 95\% confidence intervals computed by non-parametric bootstrap.}\label{fig:plot_childes}
 \end{figure*}
@@
 CHILDES corpora, with a slight dip in the penultimate position of each
-sentence. The final position of utterances in child-directed speech is
-known to be important, dating back to Aslin (1993). The Mandarin corpus
-entropy curve, by comparison, has a noticeably lower positional entropy
-values in utterance-final positions than in utterance-penultimate
-positions.
-
-We attribute the penultimate dip in the English and Spanish entropy
-curves to the fact that most of the utterances in the CHILDES English
-and Spanish corpora we examined had a determiner such as ``the'' or
-``a'' in the second-to-last position of utterances. The beginnings of
-utterances in the English and Spanish CHILDES data were usually pronouns
-or grammatical subjects, while the final words were grammatical objects
-and had a great deal of variation in the exact word that appeared in the
-utterance-final position.
+sentence. The Mandarin corpus entropy curve, by comparison, has a
+noticeably lower positional entropy values in utterance-final positions
+than in utterance-penultimate positions. We attribute the penultimate
+dip in the English and Spanish entropy curves to the fact that most of
+the utterances in the CHILDES English and Spanish corpora we examined
+had a determiner such as ``the'' or ``a'' in the second-to-last position
+of utterances. The beginnings of utterances in the English and Spanish
+CHILDES data were usually pronouns or grammatical subjects, while the
+final words were grammatical objects and had a great deal of variation
+in the exact word that appeared in the utterance-final position.
 
@@
 UID, suggests that the entropy curve can vary from language to language.
-UID predicts that each language should have a similar distribution.
+UID predicts that each language should have a similar distribution,
+which should approximate an affine function.
 
-\section{Experiment 3}\label{experiment-3}
+We found that all of the English corpora followed the same three-step
+entropy curve shape, regardless of whether the corpus contained spoken
+or written data. The Spanish corpus we used also followed this same
+shape, which we expected. However, Mandarin did not have the same shape,
+indicating that the shape of the entropy curves may arise from
+cross-linguistic variation. Similar grammatical and other features
+between two languages, such as Spanish and English, may produce a
+similar entropy curve.
+
+\hypertarget{experiment-3}{%
+\section{Experiment 3}\label{experiment-3}}
 
@@
 well. We use Wikipedia as a source for written data, which provides two
-advantages. One, the quantity of data in Wikipedia is large for each
-language and two, there are hundreds of languages with Wikipedias. This
-allows us to perform the entropy analysis on a much greater scale and to
-directly compare the results of the entropy analysis on each language to
-one another, and ultimately to predict what typological features of a
-language help determine its entropy curve, if any. We will describe our
-method of harvesting and distilling text data from each Wikipedia as
-well as how we compare the entropy curves from each language to one
-another.
+advantages. One, the quantity of data in Wikipedia is very large for
+each language, at least several megabytes, and two, there are hundreds
+of languages with their own Wikipedia corpora. We can therefore run our
+entropy analysis on a much greater scale than using small spoken CHILDES
+corpora and directly compare the results of the entropy analysis on each
+language to one another. We expect to find linguistic features that
+determine the shape of a language's entropy curve. We treated sentences
+in the Wikipedia text corpora as equivalent to utterances in the
+Switchboard and CHILDES corpora.
 
-\subsection{Methods}\label{methods}
+\hypertarget{methods}{%
+\subsection{Methods}\label{methods}}
 
@@
 articles from each Wikipedia language. Each language corpus was cleaned
-and limited to sentences between \(6\) and \(50\) words. Similar to our
+and limited to sentences between six and \(50\) words. Similar to our
 process for the Switchboard and CHILDES spoken corpora, we divided each
-corpus by sentence length, and then computed the unigram entropy measure
-on each word position within each sentence length. We wanted to directly
-compare and classify the unigram entropy curves of the languages from
-Wikipedia. We computed three slope treatments of each curve. In the
-\emph{absolute} treatment, with sentence length denoted as \(n\), we
-computed the slope between positions \(1\) and \(2\), positions \(2\)
-and \(3\), positions \(3\) and \(n-2\), positions \(n-2\) and \(n-1\)
-and positions \(n-1\) and \(n\). For the short utterances appearing the
-CHILDES speech corpora we examined, these appeared to be important
-junctions in the distributions, with a seeming plateau in the middle of
-the unigram entropy curve for each of the language corpora we examined
-in CHILDES.
+corpus by sentence length, and then computed our positional entropy
+measure on each word position within each sentence length.
 
-However, because the portion of sentences of length greater than \(10\)
-in the Wikipedia corpora were significantly larger than the CHILDES
-corpora, then we also computed relative slope treatments. In the
-\emph{relative 5} treatment, we computed the slopes between \(0\%\) and
-\(20\%\), \(20\%\) and \(40\%\), \(40\%\) and \(60\%\), \(60\%\) and
-\(80\%\) and \(80\%\) and \(100\%\) of the relative word positions in
-each sentence. When one of these percentages was not a whole number,
-then the closest whole number position was used instead for slope
-calculation. Each comparative slope within each treatment was averaged
-together between different sentence lengths, for example in the
-\emph{relative 5} treatment then all of the \(0\%\) to \(20\%\) slopes
-were averaged together. This created three treatments for the entropy
-curve for each langauge in the Wikipedia database.
+We computed two slope treatments of each curve. In the \emph{absolute}
+treatment, with sentence length denoted as \(k\), we computed the slope
+between positions \(1\) and \(2\), positions \(2\) and \(3\), positions
+\(3\) and \(k-2\), positions \(k-2\) and \(k-1\) and positions \(k-1\)
+and \(k\). For the short utterances appearing the CHILDES speech
+corpora, the slopes between the first and second slopes on either end of
+the distributions appeared to be more characteristic of the distribution
+as a whole, with a plateau in the middle of the entropy curve for each
+of the language corpora we examined in CHILDES.
 
-In both the \emph{absolute} and \emph{relative 5} treatments, each
-language is embedded in \(5\)-dimensional space. This allows for direct
-comparison between languages on Wikipedia using cosine similarity and
-moreover for unsupervised clustering analysis to compare the results of
-our position-wise analysis on datasets from Wikipedia to known
-typological features within the languages. To determine which
-phonological, morphological and syntactic features affected the
-embedding of a language in the Wikipedia dataset, we use the linguistic
-features in World Atlas of Language Structures (Dryer \& Haspelmath,
-2013).
+However, sentences of length greater than ten in the Wikipedia corpora
+were significantly more common than in the CHILDES. Therefore we also
+computed relative slope treatments. In the \emph{relative 5} treatment,
+we computed the slopes between every \(20\%\) of the relative word
+positions in each sentence length, e.g.~the slope between \(0\%\) and
+\(20\%\) of sentences of length \(10\) is the slope between the position
+entropies at the sentence-initial word position and the position
+representing the third word. When computing the word position index to
+make those cuts, if the slope was not a whole number, then the closest
+whole number position was used instead for slope calculation. Each
+comparative slope within each treatment was averaged together between
+different sentence lengths, for example all of the \(0\%\) to \(20\%\)
+slopes over all sentence lengths were averaged together.
 
-An ideal approach would use all of the corpora we can download from
-Wikipedia and all \(144\) features from the WALS database. Cluster the
-languages used in the Wikipedia analysis on the basis of WALS features
-and then directly compare the WALS clustering results with the results
-of the hierarchical clustering on the Wikipedia slope data for each of
-the different treatments using clustering similarity evaluation methods
-such as the Rand index (Rand, 1971). However, the problem then arises of
-which combination of WALS features and how many features to include in
-the clustering analysis. This is a computationally intractable
-operation. Additionallly, deciding how many clusters to use in an
-unsupervised clustering analysis is an unsolved problem in machine
-learning.
+In both the absolute and relative \(5\) treatments, each language is
+embedded in \(5\)-dimensional space. This permitted us to use cosine
+similarity for direct comparison between the languages we pulled from
+Wikipedia. To determine which phonological, morphological and syntactic
+features affected the embedding of a language in the Wikipedia dataset,
+we use the linguistic features in World Atlas of Language Structures
+Dryer \& Haspelmath (2013). We checked the effects of individual
+features on the embeddings of languages in the different treatments. We
+computed pairwise cosine similarity between each pair of language
+vectors within each treatment. For a subset of eight WALS features, we
+used a generalized linear model to see whether the cosine similarity
+between languages mattered in predicting if the languages shared the
+same value for a WALS feature.
 
-We instead check the effects of individual features on the embeddings of
-languages in the different treatments. We computed pairwise cosine
-similarity between each pair of language vectors within a treatment. For
-a subset of WALS features which had values entered for most of the
-languages we obtained from Wikipedia, we used a generalized linear model
-to see whether the cosine similarity between languages mattered in
-predicting if the languages shared the same value for a WALS feature.
+\hypertarget{results-and-analysis-2}{%
+\subsection{Results and Analysis}\label{results-and-analysis-2}}
 
-\subsection{Results and Analysis}\label{results-and-analysis-2}
-
-For eight features and \(45\) languages, we found this to be true. The
-table below shows the results for the generalized linear model we
-computed using WALS features and the absolute treatment.
+For eight features and \(45\) languages we pulled from Wikipedia, we
+found that the cosine similarity between WALS features. The table below
+shows the results for the generalized linear model we computed using
+WALS features and the absolute treatment.
 
@@
 
-Cosine distance played some role in the feature determination. The top
-four features all characterize word order, which indicates that for this
+We also used the linear model approach to evaluate the effects of cosine
+distance in the relative \(5\) treatment on the values of the same WALS
+features and found similar outcomes.
+
+From these results, we argue that cosine distance played some role in
+the feature determination. The top four features in determining the WALS
+feature value all characterize word order, which indicates that for this
 subset of features and languages that the embeddings in the slope space
-are related to the WALS features. Therefore typological features play a
-role in determining the positional entropy values for a language. This
-means that the entropy curves are in some way structured by the
-syntactic, morphological and phonological features of a language.
+are related to the WALS features. We argue therefore that typological
+features play a role in determining the positional entropy values for a
+language. This indicates that the entropy curves are in some way
+structured by the syntactic, morphological and phonological features of
+a language.
 
-\section{Discussion}\label{discussion}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 entropy at each word position within sentences. We have argued that the
-outcomes of this model are derived from the information distribution of
-a language. As this model varies cross-linguistically in terms of
-
-Follow-up with WALS features using missing-value imputation. However
-here are some of the problems with that.
+outcomes of this model are derived from the distribution of information
+transmission in human communication. Language is not a uniform noisy
+channel of communication, but each language individually represents a
+different noisy channel of communication characterized in part by
+typological features.
 
 Our work complements the approach of studies such as Aylett \& Turk
-(2004), where languages are characterized by a rate of semantic
-information per syllable. This body of work attests to languages
-possessing different rates of information transfer based on typological
-features as well.
+(2004), where languages are characterized by a single number,
+representing the rate of semantic information transfer per syllable.
+This body of work attests to languages possessing different rates of
+information transfer based on typological features as well. Within the
+information transfer rate literature, the rate is an overall feature of
+language that does not vary corpus to corpus. Similarly, in our work we
+have obtained a characteristic entropy distribution for each language.
+Both information rate and entropy distribution are characterized by
+typological features. This indicates that the structure and rate of how
+people communicate information varies cross-linguistically based on
+typological features.
 
-We also propose follow-up within linguistic processing using evidence
-from cross-linguistic eye-tracking. Research has indicated the effects
-of surprisal on fixation duration during eye-tracking studies (Boston,
-Hale, Kliegl, Patil, \& Vasishth, 2008; Demberg \& Keller, 2008; Smith
-\& Levy, 2013); higher surprisal of a word is on average predictive of
-higher fixation duration. The \emph{wrap-up effect} states that when a
-person reads written text, he or she will process sentence-final words
-more slowly on average then sentence-medial or sentence-initial words,
-due to integrating information from the entire sentence to form a final
-understanding of the sentence's meaning (Kuperman, Dambacher, Nuthmann,
-\& Kliegl, 2010; Stowe, Kaan, Sabourin, \& Taylor, 2018). The wrap-up
-effect is drawn from evidence in languages with a large final increase
-in their entropy curve from our study, so perhaps the wrap-up effect
-derives from the sentence-final increase in entropy curves observed in
-English and other Germanic languages.
+We initially wanted to use several hundred language corpora pulled from
+Wikipedia and all \(144\) linguistic features from WALS. We considered
+an computing unsupervized clusterings of the \(5\)-dimensional vectors
+of our Wikipedia entropy curves, and comparing those clusterings to
+clusterings of the WALS features directly using a cluster similarity
+measure such as the Rand index Rand (1971). However, the problem then
+arises of which combination of WALS features and how many features to
+include in the clustering analysis, with \(144!\) different
+combinations. Two additional problems presented themselves: one,
+deciding how many clusters to use in an unsupervised clustering analysis
+is an unsolved problem in machine learning; and two, not all of the
+languages on Wikipedia have values inputted for their WALS features. As
+a follow-up, we are considering missing-data imputation on the WALS
+features in order to run the generalized linear model analysis on a
+large dataset of languages.
 
-\section{Acknowledgements}\label{acknowledgements}
+On a more mechanical level, we expect that our entropy model will have
+implications for how long people take to read individual words at
+different positions of a sentence. Eye-tracking research has indicated
+the effects of surprisal on fixation duration during eye-tracking
+studies (Boston, Hale, Kliegl, Patil, \& Vasishth, 2008; Demberg \&
+Keller, 2008; Smith \& Levy, 2013); higher surprisal of a word is on
+average predictive of higher fixation duration. The \emph{wrap-up
+effect} states that people process sentence-final words more slowly on
+average then sentence-medial or sentence-initial words when reading, due
+to integrating information to form a final understanding of the
+sentence's meaning (Kuperman, Dambacher, Nuthmann, \& Kliegl, 2010;
+Stowe, Kaan, Sabourin, \& Taylor, 2018). The wrap-up effect is drawn
+from evidence in languages with a large final increase in their entropy
+curve from our study, so we hypothesize that the supposed wrap-up effect
+derives from the same source sentence-final increase in entropy curves
+observed in English and other Germanic languages.
+
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
 
@@
 
-\section{References}\label{references}
+\hypertarget{references}{%
+\section{References}\label{references}}
 
-\setlength{\parindent}{-0.1in} \setlength{\leftskip}{0.125in} \noindent
+\setlength{\parindent}{-0.1in} 
+\setlength{\leftskip}{0.125in}
+
+\noindent
 
 \hypertarget{refs}{}
-\hypertarget{ref-anderson1991}{}
+\leavevmode\hypertarget{ref-anderson1991}{}%
 Anderson, J. R. (1991). The adaptive nature of human categorization.
@@
 
-\hypertarget{ref-aylett2004}{}
+\leavevmode\hypertarget{ref-aylett2004}{}%
 Aylett, M., \& Turk, A. (2004). The smooth signal redundancy hypothesis:
@@
 
-\hypertarget{ref-bai2008}{}
+\leavevmode\hypertarget{ref-bai2008}{}%
 Bai, X., Yan, G., Liversedge, S. P., Zang, C., \& Rayner, K. (2008).
@@
 
-\hypertarget{ref-boston2008}{}
+\leavevmode\hypertarget{ref-boston2008}{}%
 Boston, M. F., Hale, J., Kliegl, R., Patil, U., \& Vasishth, S. (2008).
@@
 
-\hypertarget{ref-brown1973}{}
-Brown, R. (1973). \emph{A first language: The early stages.} Harvard U.
-Press.
-
-\hypertarget{ref-clear1993british}{}
+\leavevmode\hypertarget{ref-clear1993}{}%
 Clear, J. H. (1993). The british national corpus. \emph{The Digital
@@
 
-\hypertarget{ref-demberg2008}{}
+\leavevmode\hypertarget{ref-demberg2008}{}%
 Demberg, V., \& Keller, F. (2008). Data from eye-tracking corpora as
@@
 
-\hypertarget{ref-frank2008speaking}{}
+\leavevmode\hypertarget{ref-demuth2006}{}%
+Demuth, K., Culbertson, J., \& Alter, J. (2006). Word-minimality,
+epenthesis and coda licensing in the early acquisition of english.
+\emph{Language and Speech}, \emph{49}(2), 137--173.
+
+\leavevmode\hypertarget{ref-dryer2013}{}%
+Dryer, M. S., \& Haspelmath, M. (2013). \emph{The world atlas of
+language structures online}. Max Planck Institute for Evolutionary
+Anthropology.
+
+\leavevmode\hypertarget{ref-frank2008}{}%
 Frank, A. F., \& Jaeger, T. F. (2008). Speaking rationally: Uniform
@@
 
-\hypertarget{ref-genzel2002}{}
+\leavevmode\hypertarget{ref-genzel2002}{}%
 Genzel, D., \& Charniak, E. (2002). Entropy rate constancy in text. In
@@
 
-\hypertarget{ref-godfrey1992}{}
+\leavevmode\hypertarget{ref-godfrey1992}{}%
 Godfrey, J. J., Holliman, E. C., \& McDaniel, J. (1992). SWITCHBOARD:
 Telephone speech corpus for research and development. In
-\emph{Acoustics, speech, and signal processing, 1992. icassp-92., 1992
+\emph{Acoustics, speech, and signal processing, 1992. ICASSP-92., 1992
 ieee international conference on} (Vol. 1, pp. 517--520). IEEE.
 
-\hypertarget{ref-jaeger2010}{}
+\leavevmode\hypertarget{ref-jaeger2010}{}%
 Jaeger, T. F. (2010). Redundancy and reduction: Speakers manage
@@
 
-\hypertarget{ref-jaeger2013}{}
+\leavevmode\hypertarget{ref-jaeger2013}{}%
 Jaeger, T. F., \& Snider, N. E. (2013). Alignment as a consequence of
@@
 
-\hypertarget{ref-jain2018}{}
+\leavevmode\hypertarget{ref-jain2018}{}%
 Jain, A., Singh, V., Ranjan, S., Rajkumar, R., \& Agarwal, S. (2018).
@@
 
-\hypertarget{ref-jelinek1976}{}
+\leavevmode\hypertarget{ref-jelinek1976}{}%
 Jelinek, F. (1976). Continuous speech recognition by statistical
@@
 
-\hypertarget{ref-kuperman2010}{}
+\leavevmode\hypertarget{ref-kuperman2010}{}%
 Kuperman, V., Dambacher, M., Nuthmann, A., \& Kliegl, R. (2010). The
@@
 
-\hypertarget{ref-levy2007}{}
+\leavevmode\hypertarget{ref-levy2007}{}%
 Levy, R. P., \& Jaeger, T. F. (2007). Speakers optimize information
@@
 
-\hypertarget{ref-macwhinney2014}{}
+\leavevmode\hypertarget{ref-li2015}{}%
+Li, H., \& Zhou, J. (2015). \emph{Study on dinner table talk of
+preschool children family in shanghai} (Master's thesis). East China
+Normal University.
+
+\leavevmode\hypertarget{ref-macwhinney2014}{}%
 MacWhinney, B. (2014). \emph{The childes project: Tools for analyzing
@@
 
-\hypertarget{ref-mahowald2013}{}
+\leavevmode\hypertarget{ref-mahowald2013}{}%
 Mahowald, K., Fedorenko, E., Piantadosi, S. T., \& Gibson, E. (2013).
@@
 
-\hypertarget{ref-piantadosi2011}{}
+\leavevmode\hypertarget{ref-piantadosi2011}{}%
 Piantadosi, S. T., Tily, H., \& Gibson, E. (2011). Word lengths are
@@
 
-\hypertarget{ref-piantadosi2012}{}
+\leavevmode\hypertarget{ref-piantadosi2012}{}%
 Piantadosi, S. T., Tily, H., \& Gibson, E. (2012). The communicative
@@
 
-\hypertarget{ref-shannon1948}{}
+\leavevmode\hypertarget{ref-rand1971}{}%
+Rand, W. M. (1971). Objective criteria for the evaluation of clustering
+methods. \emph{Journal of the American Statistical Association},
+\emph{66}(336), 846--850.
+
+\leavevmode\hypertarget{ref-shannon1948}{}%
 Shannon, C. E. (1948). A mathematical theory of communication.
@@
 
-\hypertarget{ref-shiro2000}{}
-Shiro, M. (2000). Diferencias sociales en la construcción del`` yo'' y
-del`` otro'': Expresiones evaluativas en la narrativa de niños
-caraqueños en edad escolar. In \emph{Lengua, discurso, texto: I simposio
+\leavevmode\hypertarget{ref-shiro2000}{}%
+Shiro, M. (2000). Diferencias sociales en la construcción del" yo" y
+del" otro": Expresiones evaluativas en la narrativa de niños caraqueños
+en edad escolar. In \emph{Lengua, discurso, texto: I simposio
 internacional de análisis del discurso} (pp. 1303--1318). Visor.
 
-\hypertarget{ref-simons2018}{}
+\leavevmode\hypertarget{ref-simons2018}{}%
 Simons, G. F., \& Fenning, C. D. (Eds.). (2018). \emph{Ethnologue:
@@
 
-\hypertarget{ref-smith2013}{}
+\leavevmode\hypertarget{ref-smith2013}{}%
 Smith, N. J., \& Levy, R. (2013). The effect of word predictability on
@@
 
-\hypertarget{ref-stowe2018}{}
+\leavevmode\hypertarget{ref-stowe2018}{}%
 Stowe, L. A., Kaan, E., Sabourin, L., \& Taylor, R. C. (2018). The
@@
 
-\hypertarget{ref-xu2018}{}
+\leavevmode\hypertarget{ref-xu2018}{}%
 Xu, Y., \& Reitter, D. (2018). Information density converges in
@@
 
-\hypertarget{ref-yu2016}{}
+\leavevmode\hypertarget{ref-yu2016}{}%
 Yu, S., Cong, J., Liang, J., \& Liu, H. (2016). The distribution of
@@
 
-\hypertarget{ref-zhan2018}{}
+\leavevmode\hypertarget{ref-zhan2018}{}%
 Zhan, M., \& Levy, R. (2018). Comparing theories of speaker choice using

### 2019-01-31 cogsci/position-info.Rmd
annotated

@@
 
+label_data <- data_frame(length = 9, language = "Spanish", word_order = 6,
+                         person = c("Adult", "Child"), empirical_entropy = c(7,9.5))
+
 ggplot(childes_entropies, aes(x = word_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper, color = person,
+                      color = person,
                       label = person)) +
@@
   ylab("Positional entropy") +
-  geom_pointrange(size = .25) +
+  geom_pointrange(size = .25, aes(ymin = ci_lower, ymax = ci_upper)) +
   geom_line(size = .25) +
@@
   scale_color_ptol() +
-  theme(legend.position = "none") + 
-  geom_dl(method = list(dl.trans(x=x +2.2), "last.qp", cex=.7)) 
+  theme(legend.position = "nones") + 
+  geom_text(data = label_data)
 ```

### 2019-02-01 cogsci/position-info.Rmd
through E3

@@
 abstract: >
-    We argue for a reformulation of the Uniform Information Density hypothesis @levy2007 in which the grammatical, phonological and morphological features of a language determine how information is distributed throughout utterances.
-We expand on a method from @yu2016 to compute entropy based on word position within utterances as a derivation of information distribution.
-We analyze adult and child spoken corpora as well as text corpora drawn from a variety of languages on Wikipedia.
-We find that each language has a robust, characteristic entropy curve across word position, and present evidence that these entropy curves are in part determined by typological features.
-We conclude that speakers may tend towards communicating information uniformly, but the distribution of information in their utterances is conditioned on the language they communicate in.
+    Although languages vary widely in their structure, all are vehicles for the transmission of information.
+Consequently, aspects of speakers' word choice can be understood through the lens models of optimal communication.
+One prediction of these models is that speakers should keep the density of information constant over the duration of an utterance [@levy2007].
+However, different languages have different coding models that constraint the space of speaker's choices (e.g.
+canonical word order).
+We  build on a mdethod developed by @yu2016 to analye the entropy curves of natural language productions across a diverse set of languages and in diverse written and spoken contexts.
+We show languages have characteristic constraints that they impose on speaker's choices that predict deviations from Uniform Information Density, and that cross-linguistic variability in these deviations is predictable in part from syntactic properties of those languages.
 
 keywords: >
-    Entropy; information; information theory; communication
+    Uniform information density; language structure; corpus analysis
 
@@
 
-# Introduction
-
 Over 7,000 languages are spoken around the modern world [@simons2018].
@@
 <!-- **Evidence for UID here. Maybe see if you can present this work at a slightly higher level** -->
-The original evidence in @levy2007 finds that the insertion of complementizers in relative clauses in English corresponds to where neighboring words have high information content.
+The original evidence in @levy2007 finds that the insertion of complementizers (e.g.
+"that") in relative clauses in English corresponds to where neighboring words have high information content.
 Similarly, @frank2008 argues that contradictions in English such as "you're" do not occur when neighboring words are highly informative.
-The evidence in favor of UID largely been situation-specific and English-language driven, while the hypothesis itself has been applied broadly over the past decade.
+The evidence in favor of Uniform Information Density largely been situation-specific and English-language driven, while the hypothesis itself has been applied broadly over the past decade.
 Applications include determining whether linguistic alignment takes place [@jaeger2013], Zipfian word length distributions [@piantadosi2011], communication efficiency [@mahowald2013], dialogue and turn-taking [@xu2018] and the significance of ambiguity in language [@piantadosi2012], among other research.
@@
 <!-- **Evidence against UID here. Again, can you present this at a higher level?** -->
-However, other recent work has contradicted the UID hypothesis.
+However, other recent work has contradicted the Uniform Information Density hypothesis.
 Similar to the original @levy2007, @zhan2018 focuses on information distribution at particular points in sentences.
@@
 Recently, @yu2016 developed a more direct test of the Uniform Information Density hypothesis, applying the logic used by @genzel2002 to look at the distribution of information *within* individual sentences.
-<!-- **SOME HIGHER LEVEL DESCRIPTION HERE, WE'll GET INTO THE DETAILS BELOW** -->Because people process language incrementally--using the previous words in a sentence to predict the words that will come next--the amount of information that a word contains when seen in isolation should increase over the course of a sentence.
-Analyzing a large corpus of written English, they find a different pattern: Entropy increases over the first few words of an utterance and then remains constant until the final word where it again jumps up (see Figure \ref{fig:our_replication}).
+Because people process language incrementally--using the previous words in a sentence to predict the words that will come next--the amount of information that a word contains when seen in isolation should increase over the course of a sentence [@ferrer-i-cancho2013].
+Analyzing a large corpus of written English, they find a different pattern: Entropy increases over the first few words of an utterance and then remains constant until the final word where it again jumps up (see top of Figure \ref{fig:read_and_plot_exp1}).
 @yu2016 conclude that the Uniform Information Density hypothesis must not hold for medial words in a sentence.
 
-We extend and generalize @yu2016 in three ways: we confirm that this same pattern is found in spoken English--in both adult and child speakers.
-We then examine entropy curves cross-linguistically, finding a diversity of shapes across the world's languages.
-Finally, we show that entropy curves are predictable from the structure of individual languages, namely word order.
-Taken together, our results suggest a refinement of the Uniform Information Density hypothesis: speakers may structure their utterances to optimize information density, but they must do so under the constraints of the language they use to communicate.
+We extend and generalize @yu2016 in three ways: We confirm that this same pattern is found in spoken English--both in written language, and in conversational speech between adults and between parents and their children.
+Thus, this entropy curve is a robust feature of English productions.
+We then examine entropy curves cross-linguistically, and show that characteristic curves vary across langauge the world's languages.
+Finally, we show that this variation is predictable in part from the structure of individual languages (i.e.
+word order).
+Taken together, our results suggest a refinement of the Uniform Information Density hypothesis: speakers may structure their utterances to optimize information density, but they must do so under the predictable constraints of their language.
 
-The UID hypothesis predicts that information transmission rate will tend towards uniformity in both speech and text.
-We begin by examining speech to determine if we obtain the entropy curve shape we expect for spoken communication.
-We begin with the same source as @levy2007 and @jaeger2010, the Switchboard corpus of adult telephone conversations.
-We also examine child-adult conversations in the CHILDES TalkBank [@demuth2006; @macwhinney2014] corpora database of spoken adult-child conversations in multiple languages.
-Children are not fully developed speakers, so we want to compare the entropy curve we obtain by computing over the utterances in the CHILDES corpora to the utterances in the adult-adult Switchboard corpus.
-Finally, we extend our cross-linguistic findings on a small number of languages from the CHILDES experiment to a direct comparison of several dozen languages in Wikipedia.
+## Calculating entropy curves
 
-## General methods
-
-Here we will describe our adaptation of the by-word entropy method in @yu2016, which uses the text portion of the British National Corpus [@clear1993].
-We replicate their results at the end of this subsection.
+In all oll of our studies, we used an adaptation of the by-word entropy method developed by  @yu2016.
 Given a text or speech corpus divided into individual utterances, we partition the corpus by utterance length in number of words.
@@
 the number of sentences of length $k$.
-This creates a probability distribution over the words occurring in position $X$, and computing the Shannon entropy $H(X)$ @shannon1948 of this probability distribution gives the positional entropy of position $X$ in utterances of length $k$.
+This creates a probability distribution over the words occurring in position $X$, and computing the @shannon1948 entropy $H(X)$ of this probability distribution gives the positional entropy of position $X$ in utterances of length $k$.
 
@@
 @genzel2002 similarly examine a unigram entropy measure on sentences, and found that entropy at the sentence level increases linearly with sentence index within a corpus.
-UID applies this uniformity of entropy rate in sentences to all levels of speech, and so our method obtained from @yu2016, which examines text at the word level, should find an affine function at the word level.
+Uniform Information Density applies this uniformity of entropy rate in sentences to all levels of speech, and so our method obtained from @yu2016, which examines text at the word level, should find a monotonically increasing  at the word level.
 
@@
 We can directly compare any two positions within utterances to determine the amount of uncertainty, and therefore information, on average contained by words within that position of utterances.
-We are applying the same approach as in @genzel2002, but within sentences instead of across sentences.
+This method is thus identical in logic to @genzel2002, but within sentences instead of across sentences.
 
-The positional entropy at the beginning of sentences is low, then rises and plateaus for sentence-medial word positions before dropping slightly in the second-to-last position and rising again.
-However, this curve could simply be a result of the written words in the BNC.
-We will expand this analysis and replicate the result with a spoken American English corpus.
+# Study 1
 
-# Experiment 1
+We began by replicating Yu et al.'s [-@yu2016] analysis, computing entropy curves on the British National Corpus--a collection of predominantly written English [@clear1993].
+We then applied the same method to the Switchboard corpus--a collection of spoken language [@godfrey1992].
+This allows us to ask whether the characteristic function identified by @yu2016 is a general feature of English, or instead a function of written language.
+
+# Data and Analysis
+
+The British National Corpus consists of predominantly (90%) written language documents collected in the early 1980s and 1990s across a variety of genres (scientific articles, newspapers, fiction, etc).
+It also contains a small collection of spoken language.
+All together, it contains $\sim$ 100 million tokens.
+The Switchboard corpus is a collection of $\sim$ 2,400 telephone conversations between unacquainted adults prompted with subjects of conversation.
+Switchboard is the corpus used in Levy and Jaeger's [-@levy2007] original demonstration of the Uniform Information Density Hypothesis.
+
+For each corpus, we computed entropy curves using the method described above for all sentences from length 4 to 30.
 
@@
 
-We replicated the @yu2016 results and found the same characteristic three-step distribution as they did for the BNC entropy curves.
+# Results
 
-We first used the same source for data as @levy2007, the Switchboard corpus of American telephone conversations @godfrey1992.
-The Switchboard corpus contains telephone conversations of American English between strangers, with individual utterances ranging from one word to several dozen words.
-We divided the Switchboard corpus by utterance length and computed the positional entropy for each position within each sentence length, then plotted the result as an entropy curve.
+The entropy curves computed in both the British National Corpus and Switchboard were remarkably consistent both across corpora and across the range of sentence lenghts we analyzed (Figure \ref{fig:read_and_plot_exp1}).
+Confirming Yu et al.'s [-@yu2016] findings, we find that
+The positional entropy at the beginning of sentences is low, then rises and plateaus for sentence-medial word positions before dropping slightly in the second-to-last position and rising again.
 
-## Results and Analysis
-
-Our plots for three representative utterance lengths are below.
-We found the same characteristic shape for the American English Switchboard corpus as in our replication of the @yu2016 results for the BNC.
-
-```{r read_and_plot_switchboard, fig.env = "figure*", fig.width=7, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom).
+```{r read_and_plot_exp1, fig.env = "figure*", fig.width=7, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom).
 Points average entropies, error bars show 95\\% confidence intervals computed by non-parametric bootstrap."}
@@
 
-The shape we find in these two English data is notable for a number of reasons, but what we find most interesting is one, the shape of the distribution does not follow our predictions from UID and two, the distribution is robust across an unrelated pair of spoken and written English corpora.
-This suggests that the entropy curve is characteristic of the English language at least, and is not a specific effect of computing positional entropies in the written British English texts in the BNC.
-For our next step, we want to expand the positional entropy analysis to child-adult conversational corpora and languages beyond English to see if the same three-step distribution appears.
+The shape of positional entropy curves we find in these two corpora is notable for two reasons: (1) it does not follow our predictions from Uniform Information Density, and (2) the distribution is robust across written and spoken Engilsh.
+This suggests that the entropy curve is characteristic of the English language.
+We next asked whether this shape is a feature even of children's speech, and also whether it varies cross-linguistically.
 
-# Experiment 2
+# Study 2
 
-After replicating the @yu2016 results using the spoken conversation corpus from @levy2007, we want to determine whether the entropy curves will remain robust over child-adult conversational corpora.
-We also wanted to see whether the curves would show the same characteristic three-step shape across languages.
-To accomplish both of these goals, we used data from the CHILDES Talkbank system @macwhinney2014.
+To understand how robust this entropy curve is, we turned to conversational speech from parent-child interactions.
+If we find the same shape even in children's productions, we have even stronger evidence that the 3-step curve is a charcateristic feature of English.
+We thus turned to the Child Language Data Exchange System (CHILDES), a collection of transcripts of parent-child interactions [@macwhinney2014].
 
-The corpora in the CHILDES database consists of short and often disconnected utterances across hours of recordings, the unigram entropy measure is unaffected by context or lack thereof in utterances by the adults and children in the corpora.
-For our first corpus, we used the Providence English corpus @demuth2006.
-The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home.
-We divided the corpus by speaker into child and non-child categories.
-We further divided the corpus by utterance length, so that all sentences of length $k$ (e.g.
-$6$) were grouped together.
-Finally, within each utterance length, we computed the unigram entropy measure for each position.
-Computation of positional entropy was identical to our general method after dividing for utterance length.
+## Data and Analysis
 
-We also ran this analysis on Spanish and Mandarin corpora from CHILDES.
+We analyzed three seperate corpora in CHILDES: The Providence Corpus [@demuth2006], The Shiro corpus [@shiro2000], and the Zhou Dinner Corpus [@li2015].
+The Providence corpus consists of conversations between six 1--3-year-old American English speaking children and their parents recorded in their homes.
+Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home.
+The Shiro Corpus consists of prompted Spanish-language narratives individually collected from over a hundred Venezualan schoolchildren, half from high SES backgrounds and half from low SES backgrounds.
+The Zhou Dinner Corpus contains dinner conversations between 5 to 6-year-old Mandarin speaking children and their parents collected in Shanghai.
 Spanish is an Indo-European language like English, possessing similar grammar, word order and numerous cognate word.
-Mandarin Chinese is typologically completed unrelated to English.
-We used the Shiro corpus for Spanish @shiro2000, which contains prompted narratives individually collected from over a hundred Venezualan schoolchildren, half from high SES backgrounds and half from low SES backgrounds.
-We used the Zhou dinner corpus for Mandarin Chinese @li2015, which contains dinner conversations between 5 to 6-year-old children and their parents collected in Shanghai.
+In contrast, Mandarin Chinese is typologically completed unrelated to English.
 
-For each corpus, we accessed transcripts of the corpus provided through the TalkBank system and computed over the Latin alphabet transcriptions or transliterations of the original transcriptions.
+We accessed the transcripts from each corpus using \texttt{childesr}, an R-interface to a database formatted version of CHILDES [@sanchez2019].
+We divided all utterances from each transcript into those produced by target child, and those produced by all other speakers.
+We then applied the same entropy curve method described above.
 For Mandarin, we used pinyin transliterations of the utterances in the corpus with demarcated word boundaries.
-The Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Chinese writing including spaces between word boundaries can have a negative effect on reading times @bai2008.
-
-We expect that our entropy analysis on the Providence corpus will produce an identical entropy curve shape to the BNC and Switchboard curves.
-We expect similar results to English in the Spanish Shiro entropy curve, while we expect that our Mandarin corpus will produce a different entropy curve shape.
+The Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Chinese writing including spaces between word boundaries can have a negative effect on reading times [@bai2008].
 
@@
 
-```{r plot_childes, fig.env = "figure*", fig.width=7, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Representative Entropy Curves for Three Childes Corpora in English, Mandarin, and Spanish.
+```{r plot_childes, fig.env = "figure*", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Representative Entropy Curves for Three Childes Corpora in English, Mandarin, and Spanish.
 Points average entropies, error bars show 95\\% confidence intervals computed by non-parametric bootstrap."}
@@
 
-The adult and child entropy curves track one another almost identically.
-This is surprising because UID would predict that the young children in the corpora who are not fully developed speakers would have more noise in their distribution of information rates.
-This not only indicates a robustness in the unigram entropy curve across speakers, but also across ages and addressees.
-We also observed a robustness across corpora for the same languages, and a robustness across utterance lengths within the same corpus's entropy curve.
-This shows that the concept of an "entropy curve" for a specific language is well-founded when considering speech data.
-
+Across corpora, we found similar entropy curve shapes for adults and children, but distinct shapes for each language.
+Figure \ref{fig:plot_childes} shows representative curves across corpora, but shapes were robust across the full range utterances we analyzed.
 We found a distinct three-step distribution for English and Spanish CHILDES corpora, with a slight dip in the penultimate position of each sentence.
 The Mandarin corpus entropy curve, by comparison, has a noticeably lower positional entropy values in utterance-final positions than in utterance-penultimate positions.
-We attribute the penultimate dip in the English and Spanish entropy curves to the fact that most of the utterances in the CHILDES English and Spanish corpora we examined had a determiner such as "the" or "a" in the second-to-last position of utterances.
-The beginnings of utterances in the English and Spanish CHILDES data were usually pronouns or grammatical subjects, while the final words were grammatical objects and had a great deal of variation in the exact word that appeared in the utterance-final position.
 
-These entropy curves are not what would be expected from UID.
-The robustness of the three-step distribution for English and its replication in Spanish do not resemble the affine function we would expect from UID.
-The Mandarin entropy curve, which does not at all resemble either the English/Spanish distribution or the predictions of UID, suggests that the entropy curve can vary from language to language.
-UID predicts that each language should have a similar distribution, which should approximate an affine function.
+These results present two important pieces of information.
+First, our analysis of the Providence corpus broadly replicates the shape we found in both the British National Corpus and in Switchboard.
+Thus, the entropy curve of English appears not just in adult-adult conversation, but even in speech produced by parents to their children, and speech produced by very young children to their parents.
+This suggests that it is a highly robust feature of English langauge, as it structures the productions of even pre-schooled aged children.
 
-We found that all of the English corpora followed the same three-step entropy curve shape, regardless of whether the corpus contained spoken or written data.
-The Spanish corpus we used also followed this same shape, which we expected.
-However, Mandarin did not have the same shape, indicating that the shape of the entropy curves may arise from cross-linguistic variation.
-Similar grammatical and other features between two languages, such as Spanish and English, may produce a similar entropy curve.
+Second, the Entropy curves of English, Spanish, and Mandarin were not identical.
+None of these shapes resembled the monotonically increasing function predicted naively from Uniform Information Density, but also Mandarin was quite different from both English and Spanish.
+This suggests that the entropy curve can vary from language to language.
+One possibility is that this variation arises from typological features of languages, such as characteristic word order.
+We noticed, for instance, a relatively high density of determiners in the penultimate position of English and Spanish utterances, which could account for the penultimate dip in those languages.
+In our final analysis, we explored this possibility directly, analyzing a large set of Wikipedia corpora from diverse languages, and asking whether variability in their entropy curves was related to variability in their syntactic structure.
 

### 2019-02-01 cogsci/position-info.tex
through E3

@@
 \begin{abstract}
-We argue for a reformulation of the Uniform Information Density
-hypothesis Levy \& Jaeger (2007) in which the grammatical, phonological
-and morphological features of a language determine how information is
-distributed throughout utterances. We expand on a method from Yu, Cong,
-Liang, \& Liu (2016) to compute entropy based on word position within
-utterances as a derivation of information distribution. We analyze adult
-and child spoken corpora as well as text corpora drawn from a variety of
-languages on Wikipedia. We find that each language has a robust,
-characteristic entropy curve across word position, and present evidence
-that these entropy curves are in part determined by typological
-features. We conclude that speakers may tend towards communicating
-information uniformly, but the distribution of information in their
-utterances is conditioned on the language they communicate in.
+Although languages vary widely in their structure, all are vehicles for
+the transmission of information. Consequently, aspects of speakers' word
+choice can be understood through the lens models of optimal
+communication. One prediction of these models is that speakers should
+keep the density of information constant over the duration of an
+utterance (Levy \& Jaeger, 2007). However, different languages have
+different coding models that constraint the space of speaker's choices
+(e.g.~canonical word order). We build on a mdethod developed by Yu,
+Cong, Liang, \& Liu (2016) to analye the entropy curves of natural
+language productions across a diverse set of languages and in diverse
+written and spoken contexts. We show languages have characteristic
+constraints that they impose on speaker's choices that predict
+deviations from Uniform Information Density, and that cross-linguistic
+variability in these deviations is predictable in part from syntactic
+properties of those languages.
 
 \textbf{Keywords:}
-Entropy; information; information theory; communication
+Uniform information density; language structure; corpus analysis
 \end{abstract}
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
 
@@
 The original evidence in Levy \& Jaeger (2007) finds that the insertion
-of complementizers in relative clauses in English corresponds to where
-neighboring words have high information content. Similarly, Frank \&
-Jaeger (2008) argues that contradictions in English such as ``you're''
-do not occur when neighboring words are highly informative. The evidence
-in favor of UID largely been situation-specific and English-language
-driven, while the hypothesis itself has been applied broadly over the
-past decade. Applications include determining whether linguistic
-alignment takes place (Jaeger \& Snider, 2013), Zipfian word length
-distributions (Piantadosi, Tily, \& Gibson, 2011), communication
-efficiency (Mahowald, Fedorenko, Piantadosi, \& Gibson, 2013), dialogue
-and turn-taking (Xu \& Reitter, 2018) and the significance of ambiguity
-in language (Piantadosi, Tily, \& Gibson, 2012), among other research.
+of complementizers (e.g. ``that'') in relative clauses in English
+corresponds to where neighboring words have high information content.
+Similarly, Frank \& Jaeger (2008) argues that contradictions in English
+such as ``you're'' do not occur when neighboring words are highly
+informative. The evidence in favor of Uniform Information Density
+largely been situation-specific and English-language driven, while the
+hypothesis itself has been applied broadly over the past decade.
+Applications include determining whether linguistic alignment takes
+place (Jaeger \& Snider, 2013), Zipfian word length distributions
+(Piantadosi, Tily, \& Gibson, 2011), communication efficiency (Mahowald,
+Fedorenko, Piantadosi, \& Gibson, 2013), dialogue and turn-taking (Xu \&
+Reitter, 2018) and the significance of ambiguity in language
+(Piantadosi, Tily, \& Gibson, 2012), among other research.
 
-However, other recent work has contradicted the UID hypothesis. Similar
-to the original Levy \& Jaeger (2007), Zhan \& Levy (2018) focuses on
-information distribution at particular points in sentences. Zhan \& Levy
-(2018) finds that more information-rich classifiers in Mandarin Chinese
-are produced when production of the neighboring noun is difficult, not
-when the information content is high. Jain, Singh, Ranjan, Rajkumar, \&
-Agarwal (2018) examine word order across spoken sentences in Hindi, a
-freer word order language than English, and find that information
-density has no significant effect on word order.
+However, other recent work has contradicted the Uniform Information
+Density hypothesis. Similar to the original Levy \& Jaeger (2007), Zhan
+\& Levy (2018) focuses on information distribution at particular points
+in sentences. Zhan \& Levy (2018) finds that more information-rich
+classifiers in Mandarin Chinese are produced when production of the
+neighboring noun is difficult, not when the information content is high.
+Jain, Singh, Ranjan, Rajkumar, \& Agarwal (2018) examine word order
+across spoken sentences in Hindi, a freer word order language than
+English, and find that information density has no significant effect on
+word order.
 
@@
 contains when seen in isolation should increase over the course of a
-sentence. Analyzing a large corpus of written English, they find a
-different pattern: Entropy increases over the first few words of an
-utterance and then remains constant until the final word where it again
-jumps up (see Figure \ref{fig:our_replication}). Yu et al. (2016)
-conclude that the Uniform Information Density hypothesis must not hold
-for medial words in a sentence.
+sentence (Ferrer-i-Cancho, Debowski, \& Prado Martin, 2013). Analyzing a
+large corpus of written English, they find a different pattern: Entropy
+increases over the first few words of an utterance and then remains
+constant until the final word where it again jumps up (see top of Figure
+\ref{fig:read_and_plot_exp1}). Yu et al. (2016) conclude that the
+Uniform Information Density hypothesis must not hold for medial words in
+a sentence.
 
-We extend and generalize Yu et al. (2016) in three ways: we confirm that
-this same pattern is found in spoken English--in both adult and child
-speakers. We then examine entropy curves cross-linguistically, finding a
-diversity of shapes across the world's languages. Finally, we show that
-entropy curves are predictable from the structure of individual
-languages, namely word order. Taken together, our results suggest a
-refinement of the Uniform Information Density hypothesis: speakers may
-structure their utterances to optimize information density, but they
-must do so under the constraints of the language they use to
-communicate.
+We extend and generalize Yu et al. (2016) in three ways: We confirm that
+this same pattern is found in spoken English--both in written language,
+and in conversational speech between adults and between parents and
+their children. Thus, this entropy curve is a robust feature of English
+productions. We then examine entropy curves cross-linguistically, and
+show that characteristic curves vary across langauge the world's
+languages. Finally, we show that this variation is predictable in part
+from the structure of individual languages (i.e.~word order). Taken
+together, our results suggest a refinement of the Uniform Information
+Density hypothesis: speakers may structure their utterances to optimize
+information density, but they must do so under the predictable
+constraints of their language.
 
-The UID hypothesis predicts that information transmission rate will tend
-towards uniformity in both speech and text. We begin by examining speech
-to determine if we obtain the entropy curve shape we expect for spoken
-communication. We begin with the same source as Levy \& Jaeger (2007)
-and Jaeger (2010), the Switchboard corpus of adult telephone
-conversations. We also examine child-adult conversations in the CHILDES
-TalkBank (Demuth, Culbertson, \& Alter, 2006; MacWhinney, 2014) corpora
-database of spoken adult-child conversations in multiple languages.
-Children are not fully developed speakers, so we want to compare the
-entropy curve we obtain by computing over the utterances in the CHILDES
-corpora to the utterances in the adult-adult Switchboard corpus.
-Finally, we extend our cross-linguistic findings on a small number of
-languages from the CHILDES experiment to a direct comparison of several
-dozen languages in Wikipedia.
+\hypertarget{calculating-entropy-curves}{%
+\subsection{Calculating entropy
+curves}\label{calculating-entropy-curves}}
 
-\hypertarget{general-methods}{%
-\subsection{General methods}\label{general-methods}}
-
-Here we will describe our adaptation of the by-word entropy method in Yu
-et al. (2016), which uses the text portion of the British National
-Corpus (Clear, 1993). We replicate their results at the end of this
-subsection. Given a text or speech corpus divided into individual
-utterances, we partition the corpus by utterance length in number of
-words. For each word position \(X\) of utterances of length \(k\), we
-define \(w\) as a unique word occurring in position \(X\). We further
-define \(p(w)\) as the number of times word \(w\) occurs in position
-\(X\), divided by the number of total words that occur in position \(X\)
-i.e.~the number of sentences of length \(k\). This creates a probability
-distribution over the words occurring in position \(X\), and computing
-the Shannon entropy \(H(X)\) Shannon (1948) of this probability
-distribution gives the positional entropy of position \(X\) in
-utterances of length \(k\).
+In all oll of our studies, we used an adaptation of the by-word entropy
+method developed by Yu et al. (2016). Given a text or speech corpus
+divided into individual utterances, we partition the corpus by utterance
+length in number of words. For each word position \(X\) of utterances of
+length \(k\), we define \(w\) as a unique word occurring in position
+\(X\). We further define \(p(w)\) as the number of times word \(w\)
+occurs in position \(X\), divided by the number of total words that
+occur in position \(X\) i.e.~the number of sentences of length \(k\).
+This creates a probability distribution over the words occurring in
+position \(X\), and computing the Shannon (1948) entropy \(H(X)\) of
+this probability distribution gives the positional entropy of position
+\(X\) in utterances of length \(k\).
 
@@
 entropy measure on sentences, and found that entropy at the sentence
-level increases linearly with sentence index within a corpus. UID
-applies this uniformity of entropy rate in sentences to all levels of
-speech, and so our method obtained from Yu et al. (2016), which examines
-text at the word level, should find an affine function at the word
-level.
+level increases linearly with sentence index within a corpus. Uniform
+Information Density applies this uniformity of entropy rate in sentences
+to all levels of speech, and so our method obtained from Yu et al.
+(2016), which examines text at the word level, should find a
+monotonically increasing at the word level.
 
@@
 information, on average contained by words within that position of
-utterances. We are applying the same approach as in Genzel \& Charniak
+utterances. This method is thus identical in logic to Genzel \& Charniak
 (2002), but within sentences instead of across sentences.
 
-The positional entropy at the beginning of sentences is low, then rises
-and plateaus for sentence-medial word positions before dropping slightly
-in the second-to-last position and rising again. However, this curve
-could simply be a result of the written words in the BNC. We will expand
-this analysis and replicate the result with a spoken American English
-corpus.
+\hypertarget{study-1}{%
+\section{Study 1}\label{study-1}}
 
-\hypertarget{experiment-1}{%
-\section{Experiment 1}\label{experiment-1}}
+We began by replicating Yu et al.'s (2016) analysis, computing entropy
+curves on the British National Corpus--a collection of predominantly
+written English (Clear, 1993). We then applied the same method to the
+Switchboard corpus--a collection of spoken language (Godfrey, Holliman,
+\& McDaniel, 1992). This allows us to ask whether the characteristic
+function identified by Yu et al. (2016) is a general feature of English,
+or instead a function of written language.
 
-We replicated the Yu et al. (2016) results and found the same
-characteristic three-step distribution as they did for the BNC entropy
-curves.
+\hypertarget{data-and-analysis}{%
+\section{Data and Analysis}\label{data-and-analysis}}
 
-We first used the same source for data as Levy \& Jaeger (2007), the
-Switchboard corpus of American telephone conversations Godfrey,
-Holliman, \& McDaniel (1992). The Switchboard corpus contains telephone
-conversations of American English between strangers, with individual
-utterances ranging from one word to several dozen words. We divided the
-Switchboard corpus by utterance length and computed the positional
-entropy for each position within each sentence length, then plotted the
-result as an entropy curve.
+The British National Corpus consists of predominantly (90\%) written
+language documents collected in the early 1980s and 1990s across a
+variety of genres (scientific articles, newspapers, fiction, etc). It
+also contains a small collection of spoken language. All together, it
+contains \(\sim\) 100 million tokens. The Switchboard corpus is a
+collection of \(\sim\) 2,400 telephone conversations between
+unacquainted adults prompted with subjects of conversation. Switchboard
+is the corpus used in Levy and Jaeger's (2007) original demonstration of
+the Uniform Information Density Hypothesis.
 
-\hypertarget{results-and-analysis}{%
-\subsection{Results and Analysis}\label{results-and-analysis}}
+For each corpus, we computed entropy curves using the method described
+above for all sentences from length 4 to 30.
 
-Our plots for three representative utterance lengths are below. We found
-the same characteristic shape for the American English Switchboard
-corpus as in our replication of the Yu et al. (2016) results for the
-BNC.
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+The entropy curves computed in both the British National Corpus and
+Switchboard were remarkably consistent both across corpora and across
+the range of sentence lenghts we analyzed (Figure
+\ref{fig:read_and_plot_exp1}). Confirming Yu et al.'s (2016) findings,
+we find that The positional entropy at the beginning of sentences is
+low, then rises and plateaus for sentence-medial word positions before
+dropping slightly in the second-to-last position and rising again.
 
@@
 
-{\centering \includegraphics{figs/read_and_plot_switchboard-1}
+{\centering \includegraphics{figs/read_and_plot_exp1-1}
 
@@
 \caption[Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom)]{Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom).
-Points average entropies, error bars show 95\% confidence intervals computed by non-parametric bootstrap.}\label{fig:read_and_plot_switchboard}
+Points average entropies, error bars show 95\% confidence intervals computed by non-parametric bootstrap.}\label{fig:read_and_plot_exp1}
 \end{figure*}
@@
 
-The shape we find in these two English data is notable for a number of
-reasons, but what we find most interesting is one, the shape of the
-distribution does not follow our predictions from UID and two, the
-distribution is robust across an unrelated pair of spoken and written
-English corpora. This suggests that the entropy curve is characteristic
-of the English language at least, and is not a specific effect of
-computing positional entropies in the written British English texts in
-the BNC. For our next step, we want to expand the positional entropy
-analysis to child-adult conversational corpora and languages beyond
-English to see if the same three-step distribution appears.
+The shape of positional entropy curves we find in these two corpora is
+notable for two reasons: (1) it does not follow our predictions from
+Uniform Information Density, and (2) the distribution is robust across
+written and spoken Engilsh. This suggests that the entropy curve is
+characteristic of the English language. We next asked whether this shape
+is a feature even of children's speech, and also whether it varies
+cross-linguistically.
 
-\hypertarget{experiment-2}{%
-\section{Experiment 2}\label{experiment-2}}
+\hypertarget{study-2}{%
+\section{Study 2}\label{study-2}}
 
-After replicating the Yu et al. (2016) results using the spoken
-conversation corpus from Levy \& Jaeger (2007), we want to determine
-whether the entropy curves will remain robust over child-adult
-conversational corpora. We also wanted to see whether the curves would
-show the same characteristic three-step shape across languages. To
-accomplish both of these goals, we used data from the CHILDES Talkbank
-system MacWhinney (2014).
+To understand how robust this entropy curve is, we turned to
+conversational speech from parent-child interactions. If we find the
+same shape even in children's productions, we have even stronger
+evidence that the 3-step curve is a charcateristic feature of English.
+We thus turned to the Child Language Data Exchange System (CHILDES), a
+collection of transcripts of parent-child interactions (MacWhinney,
+2014).
 
-The corpora in the CHILDES database consists of short and often
-disconnected utterances across hours of recordings, the unigram entropy
-measure is unaffected by context or lack thereof in utterances by the
-adults and children in the corpora. For our first corpus, we used the
-Providence English corpus Demuth et al. (2006). The Providence corpus
-recorded interactions between children between 1 and 3 years old and
-their parents in the home. We divided the corpus by speaker into child
-and non-child categories. We further divided the corpus by utterance
-length, so that all sentences of length \(k\) (e.g. \(6\)) were grouped
-together. Finally, within each utterance length, we computed the unigram
-entropy measure for each position. Computation of positional entropy was
-identical to our general method after dividing for utterance length.
+\hypertarget{data-and-analysis-1}{%
+\subsection{Data and Analysis}\label{data-and-analysis-1}}
 
-We also ran this analysis on Spanish and Mandarin corpora from CHILDES.
-Spanish is an Indo-European language like English, possessing similar
-grammar, word order and numerous cognate word. Mandarin Chinese is
-typologically completed unrelated to English. We used the Shiro corpus
-for Spanish Shiro (2000), which contains prompted narratives
-individually collected from over a hundred Venezualan schoolchildren,
-half from high SES backgrounds and half from low SES backgrounds. We
-used the Zhou dinner corpus for Mandarin Chinese Li \& Zhou (2015),
-which contains dinner conversations between 5 to 6-year-old children and
-their parents collected in Shanghai.
+We analyzed three seperate corpora in CHILDES: The Providence Corpus
+(Demuth, Culbertson, \& Alter, 2006), The Shiro corpus (Shiro, 2000),
+and the Zhou Dinner Corpus (Li \& Zhou, 2015). The Providence corpus
+consists of conversations between six 1--3-year-old American English
+speaking children and their parents recorded in their homes. Providence
+corpus recorded interactions between children between 1 and 3 years old
+and their parents in the home. The Shiro Corpus consists of prompted
+Spanish-language narratives individually collected from over a hundred
+Venezualan schoolchildren, half from high SES backgrounds and half from
+low SES backgrounds. The Zhou Dinner Corpus contains dinner
+conversations between 5 to 6-year-old Mandarin speaking children and
+their parents collected in Shanghai. Spanish is an Indo-European
+language like English, possessing similar grammar, word order and
+numerous cognate word. In contrast, Mandarin Chinese is typologically
+completed unrelated to English.
 
-For each corpus, we accessed transcripts of the corpus provided through
-the TalkBank system and computed over the Latin alphabet transcriptions
-or transliterations of the original transcriptions. For Mandarin, we
-used pinyin transliterations of the utterances in the corpus with
-demarcated word boundaries. The Chinese characters used for writing
-Mandarin do not normally demarcate word boundaries by spacing words
-apart, and for normal Chinese writing including spaces between word
-boundaries can have a negative effect on reading times Bai, Yan,
-Liversedge, Zang, \& Rayner (2008).
+We accessed the transcripts from each corpus using \texttt{childesr}, an
+R-interface to a database formatted version of CHILDES (Sanchez et al.,
+in press). We divided all utterances from each transcript into those
+produced by target child, and those produced by all other speakers. We
+then applied the same entropy curve method described above. For
+Mandarin, we used pinyin transliterations of the utterances in the
+corpus with demarcated word boundaries. The Chinese characters used for
+writing Mandarin do not normally demarcate word boundaries by spacing
+words apart, and for normal Chinese writing including spaces between
+word boundaries can have a negative effect on reading times (Bai, Yan,
+Liversedge, Zang, \& Rayner, 2008).
 
-We expect that our entropy analysis on the Providence corpus will
-produce an identical entropy curve shape to the BNC and Switchboard
-curves. We expect similar results to English in the Spanish Shiro
-entropy curve, while we expect that our Mandarin corpus will produce a
-different entropy curve shape.
-
-\hypertarget{results-and-analysis-1}{%
-\subsection{Results and Analysis}\label{results-and-analysis-1}}
+\hypertarget{results-and-analysis}{%
+\subsection{Results and Analysis}\label{results-and-analysis}}
 
@@
 
-The adult and child entropy curves track one another almost identically.
-This is surprising because UID would predict that the young children in
-the corpora who are not fully developed speakers would have more noise
-in their distribution of information rates. This not only indicates a
-robustness in the unigram entropy curve across speakers, but also across
-ages and addressees. We also observed a robustness across corpora for
-the same languages, and a robustness across utterance lengths within the
-same corpus's entropy curve. This shows that the concept of an ``entropy
-curve'' for a specific language is well-founded when considering speech
-data.
+Across corpora, we found similar entropy curve shapes for adults and
+children, but distinct shapes for each language. Figure
+\ref{fig:plot_childes} shows representative curves across corpora, but
+shapes were robust across the full range utterances we analyzed. We
+found a distinct three-step distribution for English and Spanish CHILDES
+corpora, with a slight dip in the penultimate position of each sentence.
+The Mandarin corpus entropy curve, by comparison, has a noticeably lower
+positional entropy values in utterance-final positions than in
+utterance-penultimate positions.
 
-We found a distinct three-step distribution for English and Spanish
-CHILDES corpora, with a slight dip in the penultimate position of each
-sentence. The Mandarin corpus entropy curve, by comparison, has a
-noticeably lower positional entropy values in utterance-final positions
-than in utterance-penultimate positions. We attribute the penultimate
-dip in the English and Spanish entropy curves to the fact that most of
-the utterances in the CHILDES English and Spanish corpora we examined
-had a determiner such as ``the'' or ``a'' in the second-to-last position
-of utterances. The beginnings of utterances in the English and Spanish
-CHILDES data were usually pronouns or grammatical subjects, while the
-final words were grammatical objects and had a great deal of variation
-in the exact word that appeared in the utterance-final position.
+These results present two important pieces of information. First, our
+analysis of the Providence corpus broadly replicates the shape we found
+in both the British National Corpus and in Switchboard. Thus, the
+entropy curve of English appears not just in adult-adult conversation,
+but even in speech produced by parents to their children, and speech
+produced by very young children to their parents. This suggests that it
+is a highly robust feature of English langauge, as it structures the
+productions of even pre-schooled aged children.
 
-These entropy curves are not what would be expected from UID. The
-robustness of the three-step distribution for English and its
-replication in Spanish do not resemble the affine function we would
-expect from UID. The Mandarin entropy curve, which does not at all
-resemble either the English/Spanish distribution or the predictions of
-UID, suggests that the entropy curve can vary from language to language.
-UID predicts that each language should have a similar distribution,
-which should approximate an affine function.
-
-We found that all of the English corpora followed the same three-step
-entropy curve shape, regardless of whether the corpus contained spoken
-or written data. The Spanish corpus we used also followed this same
-shape, which we expected. However, Mandarin did not have the same shape,
-indicating that the shape of the entropy curves may arise from
-cross-linguistic variation. Similar grammatical and other features
-between two languages, such as Spanish and English, may produce a
-similar entropy curve.
+Second, the Entropy curves of English, Spanish, and Mandarin were not
+identical. None of these shapes resembled the monotonically increasing
+function predicted naively from Uniform Information Density, but also
+Mandarin was quite different from both English and Spanish. This
+suggests that the entropy curve can vary from language to language. One
+possibility is that this variation arises from typological features of
+languages, such as characteristic word order. We noticed, for instance,
+a relatively high density of determiners in the penultimate position of
+English and Spanish utterances, which could account for the penultimate
+dip in those languages. In our final analysis, we explored this
+possibility directly, analyzing a large set of Wikipedia corpora from
+diverse languages, and asking whether variability in their entropy
+curves was related to variability in their syntactic structure.
 
@@
 
-\hypertarget{results-and-analysis-2}{%
-\subsection{Results and Analysis}\label{results-and-analysis-2}}
+\hypertarget{results-and-analysis-1}{%
+\subsection{Results and Analysis}\label{results-and-analysis-1}}
 
@@
 
+\leavevmode\hypertarget{ref-ferrer-i-cancho2013}{}%
+Ferrer-i-Cancho, R., Debowski, L., \& Prado Martin, F. M. del. (2013).
+Constant conditional entropy and related hypotheses. \emph{Journal of
+Statistical Mechanics: Theory and Experiment}, \emph{2013}(07), L07001.
+
 \leavevmode\hypertarget{ref-frank2008}{}%
@@
 ieee international conference on} (Vol. 1, pp. 517--520). IEEE.
-
-\leavevmode\hypertarget{ref-jaeger2010}{}%
-Jaeger, T. F. (2010). Redundancy and reduction: Speakers manage
-syntactic information density. \emph{Cognitive Psychology},
-\emph{61}(1), 23--62.
 
@@
 
+\leavevmode\hypertarget{ref-sanchez2019}{}%
+Sanchez, A., Meylan, S. C., Braginsky, M., MacDonald, K. E., Yurovsky,
+D., \& Frank, M. (in press). Childes-db: A flexible and reproducible
+interface to the child language data exchange system. \emph{Behavior
+Research Methods}.
+
 \leavevmode\hypertarget{ref-shannon1948}{}%

### 2019-02-01 childes/childes_entropy.Rmd
dendrogram and slope plots

@@
   if (role == "child") {
-    lengths <- utterances %>%
+    sub_utterances <- utterances %>%
       filter(speaker_role == "Target_Child") %>%
-      mutate(length = str_count(stem, " ") + 1)
+      mutate(length = str_count(stem, pattern = " +") + 1)
 
 } else {
-    lengths <- utterances %>%
+    sub_utterances <- utterances %>%
       filter(speaker_role != "Target_Child") %>%
-      mutate(length = str_count(stem, " ") + 1)
+      mutate(length = str_count(stem, pattern = " +") + 1)
 }
 
-  tokens <- lengths %>% 
-    filter(length >= min_length) %>%
-    filter(length <= max_length) %>%
+  tokens <- sub_utterances %>%
+    filter(length == target_length) %>%
     mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, stem, token = stringr::str_split, pattern = " +") %>%
+    unnest_tokens(word, stem) %>%
     group_by(utterance_id) %>%
     mutate(word_order = 1:n()) %>%
-    group_by(length, word_order, word) %>%
-    filter(word_order <= length) %>% 
-    summarise(n = n()) 
+    group_by(word_order, word) %>%
+    filter(word_order <= length)
+
+  tokens %>%
+    summarise(n = n()) %>%
+    tidyboot(summary_function = function(x) x %>% 
+               summarise(entropy = entropy(n, unit = "log2")),
+             statistics_functions = function(x) x %>%
+             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
+    mutate(role = role, length = target_length)
 }
@@
   if (role == "child") {
-    lengths <- utterances %>%
+    sub_utterances <- utterances %>%
       filter(speaker_role == "Target_Child") %>%
-      mutate(length = str_count(gloss, " ") + 1)
+      mutate(length = str_count(gloss, pattern = " +") + 1)
 
 } else {
-    lengths <- utterances %>%
+    sub_utterances <- utterances %>%
       filter(speaker_role != "Target_Child") %>%
-      mutate(length = str_count(gloss, " ") + 1)
+      mutate(length = str_count(gloss, pattern = " +") + 1)
 }
 
-  tokens <- lengths %>% 
-    filter(length >= min_length) %>%
-    filter(length <= max_length) %>%
+  tokens <- sub_utterances %>%
+    filter(length == target_length) %>%
     mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, stem, token = stringr::str_split, pattern = " +") %>%
+    unnest_tokens(word, gloss) %>%
     group_by(utterance_id) %>%
     mutate(word_order = 1:n()) %>%
-    group_by(length, word_order, word) %>%
-    filter(word_order <= length) %>% 
-    summarise(n = n()) 
+    group_by(word_order, word) %>%
+    filter(word_order <= length)
+
+  tokens %>%
+    summarise(n = n()) %>%
+    tidyboot(summary_function = function(x) x %>% 
+               summarise(entropy = entropy(n, unit = "log2")),
+             statistics_functions = function(x) x %>%
+             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
+    mutate(role = role, length = target_length)
 }

### 2019-02-01 wiki_entropy/wiki_clustering.Rmd
dendrogram and slope plots

@@
 ```{r imputation for glm}
+dendro_langs <- wf_sub_langs$language
+
+wk_clust_tree <- wkpd %>% 
+              filter(Language %in% dendro_langs) %>%
+              column_to_rownames(var = "Language") %>%
+              dist() %>%
+              hclust("ward.D2")
+
+wk_clust_tree %>% plot(hang = -1)
+
 imputed_result <- wf_df %>%

### 2019-02-01 cogsci/position-info.Rmd
line level edits of everything before study 3

@@
     Although languages vary widely in their structure, all are vehicles for the transmission of information.
-Consequently, aspects of speakers' word choice can be understood through the lens models of optimal communication.
+Consequently, aspects of speakers' word choice can be understood through the models of optimal communication.
 One prediction of these models is that speakers should keep the density of information constant over the duration of an utterance [@levy2007].
-However, different languages have different coding models that constraint the space of speaker's choices (e.g.
+However, different languages have different coding models that constrain the space of a speaker's choices (e.g.
 canonical word order).
-We  build on a mdethod developed by @yu2016 to analye the entropy curves of natural language productions across a diverse set of languages and in diverse written and spoken contexts.
-We show languages have characteristic constraints that they impose on speaker's choices that predict deviations from Uniform Information Density, and that cross-linguistic variability in these deviations is predictable in part from syntactic properties of those languages.
+We build on a method developed by @yu2016 to analyze the word-level entropy curves of natural language productions across a diverse set of languages and in diverse written and spoken contexts.
+We show languages impose characteristic constraints on speaker's choices that predict deviations from Uniform Information Density, and that cross-linguistic variability in these deviations is predictable in part from syntactic properties of those languages.
 
@@
 These theories model the process of communication as transmission of information over a noisy channel.
-The producer begins with an intended meaning, packages this meaning into language, and then sends the meaning to their intended receiver over a communcative channel.
-The receiver must then decode from the signal they receive on their end of the channel the producer's intended meaning.
+The producer begins with an intended meaning, packages this meaning into language, and then sends the encoded meaning to their intended receiver over a communicative channel.
+The receiver must then decode the producer's intended meaning from the signal they receive on their end of the channel.
 The problem is that the channel is noisy, and sometimes the signal can get corrupted (e.g.
 the producer can misspeak, or the receiver can mishear).
-In order to maximize the probability that the correct meaning is transmitted, these theories predict that producers should choose linguistic messages that keep the rate of information across words constant.
+In order to maximize the probability that the correct meaning is transmitted, these theories predict that producers should choose linguistic messages that keep the rate of information constant across words.
 The intuition is that if the receiver misperceives a word, and that word contains most of the information in the sentence, then the communication will have failed.
 Because producers cannot predict which word a speaker will mishear, their best strategy is spread the information evenly across all of the words in a sentence, i.e.
-maintain *uniform information density* [@genzel2002; @levy2007].
+maintain *Uniform Information Density* [@genzel2002; @levy2007].
 
 <!-- **Evidence for UID here. Maybe see if you can present this work at a slightly higher level** -->
-The original evidence in @levy2007 finds that the insertion of complementizers (e.g.
+The evidence in favor of Uniform Information Density largely been situation-specific and English-language driven, while the hypothesis itself has been applied broadly over the past decade.
+The original evidence in favor of Uniform Information Density in @levy2007 finds that the insertion of complementizers (e.g.
 "that") in relative clauses in English corresponds to where neighboring words have high information content.
-Similarly, @frank2008 argues that contradictions in English such as "you're" do not occur when neighboring words are highly informative.
-The evidence in favor of Uniform Information Density largely been situation-specific and English-language driven, while the hypothesis itself has been applied broadly over the past decade.
-Applications include determining whether linguistic alignment takes place [@jaeger2013], Zipfian word length distributions [@piantadosi2011], communication efficiency [@mahowald2013], dialogue and turn-taking [@xu2018] and the significance of ambiguity in language [@piantadosi2012], among other research.
+Similarly, @frank2008 argues that contractions in English such as "you're" do not occur when neighboring words are highly informative.
+Applications of Uniform Information Density include determining whether linguistic alignment takes place [@jaeger2013], Zipfian word length distributions [@piantadosi2011], communication efficiency [@mahowald2013], dialogue and turn-taking [@xu2018] and the significance of ambiguity in language [@piantadosi2012], among other research.
 
@@
 @zhan2018 finds that more information-rich classifiers in Mandarin Chinese are produced when production of the neighboring noun is difficult, not when the information content is high.
-@jain2018 examine word order across spoken sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on word order.
+@jain2018 examine word order across spoken sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on determining a Hindi's speaker's word order.
 
@@
 
-We extend and generalize @yu2016 in three ways: We confirm that this same pattern is found in spoken English--both in written language, and in conversational speech between adults and between parents and their children.
+We extend and generalize @yu2016 in three ways: We confirm that this same pattern is found in spoken English--in written language, and in conversational speech between adults and between parents and their children.
 Thus, this entropy curve is a robust feature of English productions.
-We then examine entropy curves cross-linguistically, and show that characteristic curves vary across langauge the world's languages.
+We then examine entropy curves cross-linguistically, and show that characteristic curves vary across the world's languages.
 Finally, we show that this variation is predictable in part from the structure of individual languages (i.e.
@@
 
-In all oll of our studies, we used an adaptation of the by-word entropy method developed by  @yu2016.
-Given a text or speech corpus divided into individual utterances, we partition the corpus by utterance length in number of words.
-For each word position $X$ of utterances of length $k$, we define $w$ as a unique word occurring in position $X$.
+In all of our studies, we used an adaptation of the by-word entropy method developed by @yu2016.
+Given a text or speech corpus divided into individual sentences, we partition the corpus by sentence length in number of words.
+For each word position $X$ of sentences of length $k$, we define $w$ as a unique word occurring in position $X$.
 We further define $p(w)$ as the number of times word $w$ occurs in position $X$, divided by the number of total words that occur in position $X$ i.e.
 the number of sentences of length $k$.
-This creates a probability distribution over the words occurring in position $X$, and computing the @shannon1948 entropy $H(X)$ of this probability distribution gives the positional entropy of position $X$ in utterances of length $k$.
+This creates a probability distribution over the words occurring in position $X$, and computing the @shannon1948 entropy $H(X)$ of this probability distribution gives the positional entropy of position $X$ in sentences of length $k$.
 
@@
 
-With this measure, we compute the unigram entropy at each position of sentences of each length within the corpus.
-The result of this method can be plotted for each utterance length as an *entropy curve*, which can be visually compared across utterance length to observe the how the unigram entropy changes across absolute positions in each of the utterances.
+With this measure, we compute the unigram entropy at each word position of sentences of each length within the corpus.
+The result of this method can be plotted for each length as an *entropy curve*, which can be visually compared across sentence lengths to observe the how the unigram entropy changes across absolute positions in each of the sentences.
 @genzel2002 similarly examine a unigram entropy measure on sentences, and found that entropy at the sentence level increases linearly with sentence index within a corpus.
-Uniform Information Density applies this uniformity of entropy rate in sentences to all levels of speech, and so our method obtained from @yu2016, which examines text at the word level, should find a monotonically increasing  at the word level.
+Uniform Information Density applies this uniformity of entropy rate in sentences to all levels of speech, and so our method obtained from @yu2016, which examines text at the word level, should find a monotonically increasing affine function at the word level.
 
@@
 The British National Corpus consists of predominantly (90%) written language documents collected in the early 1980s and 1990s across a variety of genres (scientific articles, newspapers, fiction, etc).
-It also contains a small collection of spoken language.
+It also contains a small collection of corpora from spoken language.
 All together, it contains $\sim$ 100 million tokens.
@@
 
-The entropy curves computed in both the British National Corpus and Switchboard were remarkably consistent both across corpora and across the range of sentence lenghts we analyzed (Figure \ref{fig:read_and_plot_exp1}).
+The entropy curves computed in both the British National Corpus and Switchboard were remarkably consistent both across corpora and across the range of sentence lengths we analyzed (Figure \ref{fig:read_and_plot_exp1}).
 Confirming Yu et al.'s [-@yu2016] findings, we find that
-The positional entropy at the beginning of sentences is low, then rises and plateaus for sentence-medial word positions before dropping slightly in the second-to-last position and rising again.
+the positional entropy at the beginning of sentences is low, then rises and plateaus for sentence-medial word positions before dropping slightly in the second-to-last position and rising again.
 
@@
 We analyzed three seperate corpora in CHILDES: The Providence Corpus [@demuth2006], The Shiro corpus [@shiro2000], and the Zhou Dinner Corpus [@li2015].
-The Providence corpus consists of conversations between six 1--3-year-old American English speaking children and their parents recorded in their homes.
-Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home.
+The Providence corpus consists of conversations between six 1--3-year-old American English-speaking children and their parents recorded in their homes.
 The Shiro Corpus consists of prompted Spanish-language narratives individually collected from over a hundred Venezualan schoolchildren, half from high SES backgrounds and half from low SES backgrounds.
-The Zhou Dinner Corpus contains dinner conversations between 5 to 6-year-old Mandarin speaking children and their parents collected in Shanghai.
-Spanish is an Indo-European language like English, possessing similar grammar, word order and numerous cognate word.
-In contrast, Mandarin Chinese is typologically completed unrelated to English.
+The Zhou Dinner Corpus contains dinner conversations between 5--6-year-old Mandarin-speaking children and their parents collected in Shanghai.
+Spanish is an Indo-European language like English, possessing similar grammar, word order and numerous cognate words.
+In contrast, Mandarin Chinese is typologically unrelated to English.
 
@@
 For Mandarin, we used pinyin transliterations of the utterances in the corpus with demarcated word boundaries.
-The Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Chinese writing including spaces between word boundaries can have a negative effect on reading times [@bai2008].
+The Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Chinese writing including spaces between word boundaries is unnatural and can have a negative effect on reading times [@bai2008].
 
-## Results and Analysis
+## Results
 
@@
 
-Second, the Entropy curves of English, Spanish, and Mandarin were not identical.
-None of these shapes resembled the monotonically increasing function predicted naively from Uniform Information Density, but also Mandarin was quite different from both English and Spanish.
+Second, the entropy curves of English, Spanish, and Mandarin were not identical.
+None of these shapes resembled the affine function predicted naively from Uniform Information Density, but also Mandarin was quite different from both English and Spanish.
 This suggests that the entropy curve can vary from language to language.
 One possibility is that this variation arises from typological features of languages, such as characteristic word order.
-We noticed, for instance, a relatively high density of determiners in the penultimate position of English and Spanish utterances, which could account for the penultimate dip in those languages.
+We noticed, for instance, a relatively high density of determiners in the penultimate position of English and Spanish utterances, which could account for the penultimate dip in the entropy curves of those two languages.
 In our final analysis, we explored this possibility directly, analyzing a large set of Wikipedia corpora from diverse languages, and asking whether variability in their entropy curves was related to variability in their syntactic structure.

### 2019-02-01 cogsci/position-info.Rmd
moved clustering code

@@
 
+```{r }
+library(tidyverse)
+library(googlesheets)
+library(lingtypology)
+library(knitr)
+library(missMDA)
+library(cluster)
+library(widyr)
+library(lsa)
+library(magrittr)
+library(broom)
+
+wiki_gs <- gs_title("Wikipedia_relative5")
+wkpd <- gs_read(wiki_gs) 
+LANGUAGES <- wkpd$Language
+
+FEATURES <- c(1:144)
+for (i in c(1:144)) {
+  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
+} #get all WALS features in correct syntax
+
+wf <- wals.feature(FEATURES)
+
+wf_df <- wf %>%
+  as_data_frame() %>%
+  filter(language %in% LANGUAGES) %>%
+  group_by(language) %>%
+  slice(1) %>%
+  gather(feature, value, `1A`:`144A`) %>%
+  group_by(feature) %>%
+  mutate(value = factor(value))
+
+cosines <- wkpd %>%
+  column_to_rownames("Language") %>%
+  t() %>%
+  cosine() %>%
+  as_data_frame(rownames = "language1") %>%
+  gather(language2, cosine, -language1)
+
+wf_many_langs <- wf_df %>%
+  group_by(feature) %>%
+  summarise(n = sum(!is.na(value))) %>%
+  filter(n > 60)
+
+wf_sub_langs <- wf_df %>%
+  filter(feature %in% wf_many_langs$feature) %>%
+  group_by(language) %>%
+  summarise(keep = sum(is.na(value)) == 0) %>%
+  filter(keep)
+
+wf_sub <- wf_df %>%
+  filter(feature %in% wf_many_langs$feature,
+         language %in% wf_sub_langs$language) %>%
+  select(language, feature, value) %>%
+  mutate(value = as.numeric(as.factor(value)))
+
+wf_pairwise <- expand.grid(language1 = wf_sub$language,
+                           language2 = wf_sub$language) %>%
+  left_join(wf_sub, by = c("language1" = "language")) %>%
+  rename(value1 = value) %>%
+  left_join(wf_sub, by = c("language2" = "language", "feature")) %>%
+  rename(value2 = value) %>%
+  mutate(same = value1 == value2) %>%
+  select(-value1, -value2) %>%
+  left_join(cosines, by = c("language1", "language2"))
+
+models <- wf_pairwise %>%
+  group_by(feature) %>%
+  nest() %>%
+  mutate(model = map(data, ~glm(same ~ cosine, 
+                                family = "binomial", data = .)))
+
+models %>%
+  mutate(coeffs = map(model, tidy)) %>%
+  select(-data, -model) %>%
+  unnest() %>%
+  filter(term == "cosine") %>%
+  arrange(desc(statistic))
+
+### generate dendrogram plot
+dendro_langs <- wf_sub_langs$language
+
+wk_clust_tree <- wkpd %>% 
+              filter(Language %in% dendro_langs) %>%
+              column_to_rownames(var = "Language") %>%
+              dist() %>%
+              hclust("ward.D2")
+
+wk_clust_tree %>% plot(hang = -1)
+```
+
 ```{r tables, results="asis", tab.env = "table"}

### 2019-02-02 cogsci/position-info.Rmd
everything in

@@
 library(directlabels)
+library(ggdendro)
 
@@
 
-```{r read_and_plot_exp1, fig.env = "figure*", fig.width=7, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom).
+```{r read_and_plot_exp1, fig.env = "figure*", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom).
 Points average entropies, error bars show 95\\% confidence intervals computed by non-parametric bootstrap."}
@@
 These results present two important pieces of information.
-First, our analysis of the Providence corpus broadly replicates the shape we found in both the British National Corpus and in Switchboard.
+First, the Providence corpus entropy curve broadly replicates the curves found in both the British National Corpus and in Switchboard.
 Thus, the entropy curve of English appears not just in adult-adult conversation, but even in speech produced by parents to their children, and speech produced by very young children to their parents.
-This suggests that it is a highly robust feature of English langauge, as it structures the productions of even pre-schooled aged children.
+This suggests that it is a highly robust feature of English language, as it structures the productions of even pre-schooled aged children.
 
 Second, the entropy curves of English, Spanish, and Mandarin were not identical.
-None of these shapes resembled the affine function predicted naively from Uniform Information Density, but also Mandarin was quite different from both English and Spanish.
+No shape resembled the affine function predicted naively from Uniform Information Density, but also Mandarin was quite different from both English and Spanish.
 This suggests that the entropy curve can vary from language to language.
 One possibility is that this variation arises from typological features of languages, such as characteristic word order.
-We noticed, for instance, a relatively high density of determiners in the penultimate position of English and Spanish utterances, which could account for the penultimate dip in the entropy curves of those two languages.
+We noticed, for instance, a relatively high density of determiners in the penultimate position of English and Spanish utterances, which could account their characteristic penultimate dip.
 In our final analysis, we explored this possibility directly, analyzing a large set of Wikipedia corpora from diverse languages, and asking whether variability in their entropy curves was related to variability in their syntactic structure.
 
-# Experiment 3
+# Study 3
 
-The UID hypothesis also applies to written communication: we expect people to communicate information at a uniform rate through writing as well.
-We use Wikipedia as a source for written data, which provides two advantages.
-One, the quantity of data in Wikipedia is very large for each language, at least several megabytes, and two, there are hundreds of languages with their own Wikipedia corpora.
-We can therefore run our entropy analysis on a much greater scale than using small spoken CHILDES corpora and directly compare the results of the entropy analysis on each language to one another.
-We expect to find linguistic features that determine the shape of a language's entropy curve.
-We treated sentences in the Wikipedia text corpora as equivalent to utterances in the Switchboard and CHILDES corpora.
+In order to understand how entropy curves are related to linguistic structure, we turned to Wikipedia for a larger and more diverse of set of languages.
 
-## Methods
+## Method
 
-Using Giuseppe Attardi's Wikiextractor tool \footnote{https://github.com/attardi/wikiextractor}, we extract text corpora from Wikipedia by downloading a stored collection of Wikipedia entries in each langauge and randomly selecting several thousand articles from each Wikipedia language.
-Each language corpus was cleaned and limited to sentences between six and $50$ words.
-Similar to our process for the Switchboard and CHILDES spoken corpora, we divided each corpus by sentence length, and then computed our positional entropy measure on each word position within each sentence length.
+### Data and Analysis
 
-We computed two slope treatments of each curve.
-In the *absolute* treatment, with sentence length denoted as $k$, we computed the slope between positions $1$ and $2$, positions $2$ and $3$, positions $3$ and $k-2$, positions $k-2$ and $k-1$ and positions $k-1$ and $k$.
-For the short utterances appearing the CHILDES speech corpora, the slopes between the first and second slopes on either end of the distributions appeared to be more characteristic of the distribution as a whole, with a plateau in the middle of the entropy curve for each of the language corpora we examined in CHILDES.
+Wikipedia has two primary advantages: (1) Individual language corpora are generally large, and (2) hundreds of languages are represented.
+To understand how variability in entropy curves is related to the structure of language, we used features from the World Atlas of Language Structures [WALS, @dryer2013].
+WALS is The World Atlas of Language Structures is a curated collection of typological features across the world's languages.
+However, as the collection includes contributions from many authors studying a heterogeneous set of topic and languages, most features are missing for most languages.
+For our sample, we sought to balance diversity of languages with coverage in WALS.
+We selected $45$ languages for which we $8$ WALS features were coded (Figure \ref{fig:clust_tree}).
+Unfortunately, these features all
+reflected the order of words in a sentence in each langauge (e.g.
+whether objects come before verbs, whether adjectives come before nouns).
+A consistent set of results across these features would be consistent with an effect of order on entropy curves, but unfortunately no other feature types are available for comparison.
 
-However, sentences of length greater than ten in the Wikipedia corpora were significantly more common than in the CHILDES.
-Therefore we also computed relative slope treatments.
-In the *relative 5* treatment, we computed the slopes between every $20\%$ of the relative word positions in each sentence length, e.g.
-the slope between $0\%$ and $20\%$ of sentences of length $10$ is the slope between the position entropies at the sentence-initial word position and the position representing the third word.
-When computing the word position index to make those cuts, if the slope was not a whole number, then the closest whole number position was used instead for slope calculation.
-Each comparative slope within each treatment was averaged together between different sentence lengths, for example all of the $0\%$ to $20\%$ slopes over all sentence lengths were averaged together.
-
-In both the absolute and relative $5$ treatments, each language is embedded in $5$-dimensional space.
-This permitted us to use cosine similarity for direct comparison between the languages we pulled from Wikipedia.
-To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we use the linguistic features in World Atlas of Language Structures @dryer2013.
-We checked the effects of individual features on the embeddings of languages in the different treatments.
-We computed pairwise cosine similarity between each pair of language vectors within each treatment.
-For a subset of eight WALS features, we used a generalized linear model to see whether the cosine similarity between languages mattered in predicting if the languages shared the same value for a WALS feature.
+We computed entropy curves separately for each language as before.
+In order to understand how these curves varied across languages, we developed a method for compressing the curves into a few representative values.
+We noticed that curves tended to vary in predictable utterance positions: The majority were flat in the middle of utterances, but had either rising or falling slopes at the beginnings and ends.
+For each language, for each utterance length $k$ we computed 5 slopes: The change between the 1st and 2nd position, the change between the 2nd and 3rd position, the slope between the 3rd position and the 3rd-to-last position ($k-1$), the slope between the 3rd-to-last and the 2nd-to-last position ($k-1$), and the slope between the 2nd-to-last and final positions.
+We then averaged these slopes across utterance lengths within a language to estimate a language-typical signature.
+We performed an additional analysis dividing each utterance into 6ths evenly and computing the slopes between these 6ths.
+Results from this analysis were qualitatively similar.
 
@@
 
-For eight features and $45$ languages we pulled from Wikipedia, we found that the cosine similarity between WALS features.
-The table below shows the results for the generalized linear model we computed using WALS features and the absolute treatment.
+We treated these 5 slopes as a vector, allowing us to compute the pairwise similarity between languages with cosine--a standard measure of vector similarity [@landauer1997].
+Intuitively, vectors with with a smaller angle between them are more similar.
+Figure \ref{fig:clust_tree} shows a hierarchical clustering of these 5 dimensional vectors, demonstrating that entropy curve similarity appears to capture typological relatedly, e.g.
+Slavic languages are similar to each-other.
 
-```{r }
+```{r load_features}
+features <- read_csv(here("data/wals_features.csv"))
+
+glm_model <- read_csv(here("data/glm.csv")) %>%
+  select(-X1) %>%
+  left_join(features) %>%
+  filter(meaning != "Tea") %>%
+  select(-term, -std.error) %>%
+  mutate(r2 = (statistic^2) / (statistic^2 + 164024))
+
+max_r2 <- glm_model %>% 
+  mutate(r2 = r2 * 100) %>% 
+  pull(r2) %>% 
+  max() %>% 
+  round(2)
+```
+
+We then used asked whether these pairwise similarities are related to overlap in linguistic features.
+For each of the 8 features we measured, we fit a logistic regression predicting whether two languages would have the same value for it from their pairwise entropy curve similarity.
+Overlap on all 8 features was reliably predicted by entropy curve similarity (largest $p = .001$), with the amount of variance in feature overlap predicted varying across predictors \ref{tab:tab_features}.
+Order of object and verb ($83A$) was the strongest predictor, explaining `r max_r2`% of the variance in pairwise entropy curve similarity.
+
+```{r, eval = F}
 library(tidyverse)
@@
 
-```{r tables, results="asis", tab.env = "table"}
-glm_model <- read_csv("glm.csv") %>%
-    select(-X1)
+```{r clust_tree, fig.width = 3.5, fig.align='center', set.cap.width=T, fig.cap = "A dendrogram estimated from hierarchically clustering the slopes of our 45 languages", fig.show='hold', cache = T}
+wk_clust_tree <- readRDS("data/clust_tree.RDS")
 
-tab1 <- xtable::xtable(glm_model)
+ggdendrogram(wk_clust_tree, rotate = F, size = 1) + 
+ theme(axis.text.y = element_blank(),
+       axis.text.x = element_text(size = 7))
+
+```
+
+```{r tab_features, results="asis", tab.env = "table"}
+tab1 <- glm_model %>%
+  rename(`t-value` = statistic, `p-value` = p.value,
+         `#` = feature) %>%
+  mutate(r2 = round(r2,3)) %>%
+  select(-estimate,-meaning)  %>%
+  xtable(caption = "Capptoin here",
+         label = "tab:tab_features")
+
 print(tab1, type="latex", comment = F, table.placement = "tb",
@@
 
-We also used the linear model approach to evaluate the effects of cosine distance in the relative $5$ treatment on the values of the same WALS features and found similar outcomes.
+# General Discussion
 
-From these results, we argue that cosine distance played some role in the feature determination.
-The top four features in determining the WALS feature value all characterize word order, which indicates that for this subset of features and languages that the embeddings in the slope space are related to the WALS features.
-We argue therefore that typological features play a role in determining the positional entropy values for a language.
-This indicates that the entropy curves are in some way structured by the syntactic, morphological and phonological features of a language.
-
-# Discussion 
-In this paper, we have extended and applied a model from @yu2016 and derived from @genzel2002 for distinguishing entropy at each word position within sentences.
-We have argued that the outcomes of this model are derived from the distribution of information transmission in human communication.
-Language is not a uniform noisy channel of communication, but each language individually represents a different noisy channel of communication characterized in part by typological features.
-
-Our work complements the approach of studies such as @aylett2004, where languages are characterized by a single number, representing the rate of semantic information transfer per syllable.
-This body of work attests to languages possessing different rates of information transfer based on typological features as well.
-Within the information transfer rate literature, the rate is an overall feature of language that does not vary corpus to corpus.
-Similarly, in our work we have obtained a characteristic entropy distribution for each language.
-Both information rate and entropy distribution are characterized by typological features.
-This indicates that the structure and rate of how people communicate information varies cross-linguistically based on typological features.
-
-We initially wanted to use several hundred language corpora pulled from Wikipedia and all $144$ linguistic features from WALS.
-We considered an computing unsupervized clusterings of the $5$-dimensional vectors of our Wikipedia entropy curves, and comparing those clusterings to clusterings of the WALS features directly using a cluster similarity measure such as the Rand index @rand1971.
-However, the problem then arises of which combination of WALS features and how many features to include in the clustering analysis, with $144!$ different combinations.
-Two additional problems presented themselves: one, deciding how many clusters to use in an unsupervised clustering analysis is an unsolved problem in machine learning; and two, not all of the languages on Wikipedia have values inputted for their WALS features.
-As a follow-up, we are considering missing-data imputation on the WALS features in order to run the generalized linear model analysis on a large dataset of languages.
-
-On a more mechanical level, we expect that our entropy model will have implications for how long people take to read individual words at different positions of a sentence.
-Eye-tracking research has indicated the effects of surprisal on fixation duration during eye-tracking studies [@demberg2008; @boston2008; @smith2013]; higher surprisal of a word is on average predictive of higher fixation duration.
-The *wrap-up effect* states that people process sentence-final words more slowly on average then sentence-medial or sentence-initial words when reading, due to integrating information to form a final understanding of the sentence's meaning [@stowe2018; @kuperman2010].
-The wrap-up effect is drawn from evidence in languages with a large final increase in their entropy curve from our study, so we hypothesize that the supposed wrap-up effect derives from the same source sentence-final increase in entropy curves observed in English and other Germanic languages.
-
-# Acknowledgements
-
-[Not here yet.]
+Taken together, the results of our studies yield 3 main findings.
+First, languages have characteristic entropy curves that can be found in their productions whether written, spoken, or even spoken by children.
+These curves diverge in reliabile ways from the monotonically increasing functions predicting by a naive Uniform Information Density hypothesis.
+Second, these characteristic curves vary across languages--some languages like English have increasing curves, but others like Mandarin are characterized by decreasing entropy of the course of an utterance.
+Finally, these characteristic curves are related to structural properties of those languages, for instance their word orders.
+This work presents an important rejoinder to models of linguistic productions as optimizing communicative success.
+While speakers may do their best to structure the information in their productions to minimize spikes in entropy, they operate under coding models they inherit from their native language, and these coding models constrain the choices they make.
 

### 2019-02-02 cogsci/position-info.tex
everything in

@@
 the transmission of information. Consequently, aspects of speakers' word
-choice can be understood through the lens models of optimal
-communication. One prediction of these models is that speakers should
-keep the density of information constant over the duration of an
-utterance (Levy \& Jaeger, 2007). However, different languages have
-different coding models that constraint the space of speaker's choices
-(e.g.~canonical word order). We build on a mdethod developed by Yu,
-Cong, Liang, \& Liu (2016) to analye the entropy curves of natural
-language productions across a diverse set of languages and in diverse
-written and spoken contexts. We show languages have characteristic
-constraints that they impose on speaker's choices that predict
-deviations from Uniform Information Density, and that cross-linguistic
-variability in these deviations is predictable in part from syntactic
-properties of those languages.
+choice can be understood through the models of optimal communication.
+One prediction of these models is that speakers should keep the density
+of information constant over the duration of an utterance (Levy \&
+Jaeger, 2007). However, different languages have different coding models
+that constrain the space of a speaker's choices (e.g.~canonical word
+order). We build on a method developed by Yu, Cong, Liang, \& Liu (2016)
+to analyze the word-level entropy curves of natural language productions
+across a diverse set of languages and in diverse written and spoken
+contexts. We show languages impose characteristic constraints on
+speaker's choices that predict deviations from Uniform Information
+Density, and that cross-linguistic variability in these deviations is
+predictable in part from syntactic properties of those languages.
 
@@
 information over a noisy channel. The producer begins with an intended
-meaning, packages this meaning into language, and then sends the meaning
-to their intended receiver over a communcative channel. The receiver
-must then decode from the signal they receive on their end of the
-channel the producer's intended meaning. The problem is that the channel
-is noisy, and sometimes the signal can get corrupted (e.g.~the producer
-can misspeak, or the receiver can mishear). In order to maximize the
-probability that the correct meaning is transmitted, these theories
-predict that producers should choose linguistic messages that keep the
-rate of information across words constant. The intuition is that if the
-receiver misperceives a word, and that word contains most of the
-information in the sentence, then the communication will have failed.
-Because producers cannot predict which word a speaker will mishear,
-their best strategy is spread the information evenly across all of the
-words in a sentence, i.e.~maintain \emph{uniform information density}
-(Genzel \& Charniak, 2002; Levy \& Jaeger, 2007).
+meaning, packages this meaning into language, and then sends the encoded
+meaning to their intended receiver over a communicative channel. The
+receiver must then decode the producer's intended meaning from the
+signal they receive on their end of the channel. The problem is that the
+channel is noisy, and sometimes the signal can get corrupted (e.g.~the
+producer can misspeak, or the receiver can mishear). In order to
+maximize the probability that the correct meaning is transmitted, these
+theories predict that producers should choose linguistic messages that
+keep the rate of information constant across words. The intuition is
+that if the receiver misperceives a word, and that word contains most of
+the information in the sentence, then the communication will have
+failed. Because producers cannot predict which word a speaker will
+mishear, their best strategy is spread the information evenly across all
+of the words in a sentence, i.e.~maintain \emph{Uniform Information
+Density} (Genzel \& Charniak, 2002; Levy \& Jaeger, 2007).
 
-The original evidence in Levy \& Jaeger (2007) finds that the insertion
-of complementizers (e.g. ``that'') in relative clauses in English
-corresponds to where neighboring words have high information content.
-Similarly, Frank \& Jaeger (2008) argues that contradictions in English
-such as ``you're'' do not occur when neighboring words are highly
-informative. The evidence in favor of Uniform Information Density
-largely been situation-specific and English-language driven, while the
-hypothesis itself has been applied broadly over the past decade.
-Applications include determining whether linguistic alignment takes
-place (Jaeger \& Snider, 2013), Zipfian word length distributions
-(Piantadosi, Tily, \& Gibson, 2011), communication efficiency (Mahowald,
-Fedorenko, Piantadosi, \& Gibson, 2013), dialogue and turn-taking (Xu \&
-Reitter, 2018) and the significance of ambiguity in language
-(Piantadosi, Tily, \& Gibson, 2012), among other research.
+The evidence in favor of Uniform Information Density largely been
+situation-specific and English-language driven, while the hypothesis
+itself has been applied broadly over the past decade. The original
+evidence in favor of Uniform Information Density in Levy \& Jaeger
+(2007) finds that the insertion of complementizers (e.g. ``that'') in
+relative clauses in English corresponds to where neighboring words have
+high information content. Similarly, Frank \& Jaeger (2008) argues that
+contractions in English such as ``you're'' do not occur when neighboring
+words are highly informative. Applications of Uniform Information
+Density include determining whether linguistic alignment takes place
+(Jaeger \& Snider, 2013), Zipfian word length distributions (Piantadosi,
+Tily, \& Gibson, 2011), communication efficiency (Mahowald, Fedorenko,
+Piantadosi, \& Gibson, 2013), dialogue and turn-taking (Xu \& Reitter,
+2018) and the significance of ambiguity in language (Piantadosi, Tily,
+\& Gibson, 2012), among other research.
 
@@
 English, and find that information density has no significant effect on
-word order.
+determining a Hindi's speaker's word order.
 
@@
 We extend and generalize Yu et al. (2016) in three ways: We confirm that
-this same pattern is found in spoken English--both in written language,
-and in conversational speech between adults and between parents and
-their children. Thus, this entropy curve is a robust feature of English
+this same pattern is found in spoken English--in written language, and
+in conversational speech between adults and between parents and their
+children. Thus, this entropy curve is a robust feature of English
 productions. We then examine entropy curves cross-linguistically, and
-show that characteristic curves vary across langauge the world's
-languages. Finally, we show that this variation is predictable in part
-from the structure of individual languages (i.e.~word order). Taken
-together, our results suggest a refinement of the Uniform Information
-Density hypothesis: speakers may structure their utterances to optimize
+show that characteristic curves vary across the world's languages.
+Finally, we show that this variation is predictable in part from the
+structure of individual languages (i.e.~word order). Taken together, our
+results suggest a refinement of the Uniform Information Density
+hypothesis: speakers may structure their utterances to optimize
 information density, but they must do so under the predictable
@@
 
-In all oll of our studies, we used an adaptation of the by-word entropy
+In all of our studies, we used an adaptation of the by-word entropy
 method developed by Yu et al. (2016). Given a text or speech corpus
-divided into individual utterances, we partition the corpus by utterance
-length in number of words. For each word position \(X\) of utterances of
+divided into individual sentences, we partition the corpus by sentence
+length in number of words. For each word position \(X\) of sentences of
 length \(k\), we define \(w\) as a unique word occurring in position
@@
 this probability distribution gives the positional entropy of position
-\(X\) in utterances of length \(k\).
+\(X\) in sentences of length \(k\).
 
@@
 
-With this measure, we compute the unigram entropy at each position of
-sentences of each length within the corpus. The result of this method
-can be plotted for each utterance length as an \emph{entropy curve},
-which can be visually compared across utterance length to observe the
-how the unigram entropy changes across absolute positions in each of the
-utterances. Genzel \& Charniak (2002) similarly examine a unigram
-entropy measure on sentences, and found that entropy at the sentence
-level increases linearly with sentence index within a corpus. Uniform
-Information Density applies this uniformity of entropy rate in sentences
-to all levels of speech, and so our method obtained from Yu et al.
-(2016), which examines text at the word level, should find a
-monotonically increasing at the word level.
+With this measure, we compute the unigram entropy at each word position
+of sentences of each length within the corpus. The result of this method
+can be plotted for each length as an \emph{entropy curve}, which can be
+visually compared across sentence lengths to observe the how the unigram
+entropy changes across absolute positions in each of the sentences.
+Genzel \& Charniak (2002) similarly examine a unigram entropy measure on
+sentences, and found that entropy at the sentence level increases
+linearly with sentence index within a corpus. Uniform Information
+Density applies this uniformity of entropy rate in sentences to all
+levels of speech, and so our method obtained from Yu et al. (2016),
+which examines text at the word level, should find a monotonically
+increasing affine function at the word level.
 
@@
 variety of genres (scientific articles, newspapers, fiction, etc). It
-also contains a small collection of spoken language. All together, it
-contains \(\sim\) 100 million tokens. The Switchboard corpus is a
-collection of \(\sim\) 2,400 telephone conversations between
+also contains a small collection of corpora from spoken language. All
+together, it contains \(\sim\) 100 million tokens. The Switchboard
+corpus is a collection of \(\sim\) 2,400 telephone conversations between
 unacquainted adults prompted with subjects of conversation. Switchboard
@@
 Switchboard were remarkably consistent both across corpora and across
-the range of sentence lenghts we analyzed (Figure
+the range of sentence lengths we analyzed (Figure
 \ref{fig:read_and_plot_exp1}). Confirming Yu et al.'s (2016) findings,
-we find that The positional entropy at the beginning of sentences is
+we find that the positional entropy at the beginning of sentences is
 low, then rises and plateaus for sentence-medial word positions before
@@
 and the Zhou Dinner Corpus (Li \& Zhou, 2015). The Providence corpus
-consists of conversations between six 1--3-year-old American English
-speaking children and their parents recorded in their homes. Providence
-corpus recorded interactions between children between 1 and 3 years old
-and their parents in the home. The Shiro Corpus consists of prompted
-Spanish-language narratives individually collected from over a hundred
-Venezualan schoolchildren, half from high SES backgrounds and half from
-low SES backgrounds. The Zhou Dinner Corpus contains dinner
-conversations between 5 to 6-year-old Mandarin speaking children and
-their parents collected in Shanghai. Spanish is an Indo-European
-language like English, possessing similar grammar, word order and
-numerous cognate word. In contrast, Mandarin Chinese is typologically
-completed unrelated to English.
+consists of conversations between six 1--3-year-old American
+English-speaking children and their parents recorded in their homes. The
+Shiro Corpus consists of prompted Spanish-language narratives
+individually collected from over a hundred Venezualan schoolchildren,
+half from high SES backgrounds and half from low SES backgrounds. The
+Zhou Dinner Corpus contains dinner conversations between 5--6-year-old
+Mandarin-speaking children and their parents collected in Shanghai.
+Spanish is an Indo-European language like English, possessing similar
+grammar, word order and numerous cognate words. In contrast, Mandarin
+Chinese is typologically unrelated to English.
 
@@
 words apart, and for normal Chinese writing including spaces between
-word boundaries can have a negative effect on reading times (Bai, Yan,
-Liversedge, Zang, \& Rayner, 2008).
+word boundaries is unnatural and can have a negative effect on reading
+times (Bai, Yan, Liversedge, Zang, \& Rayner, 2008).
 
-\hypertarget{results-and-analysis}{%
-\subsection{Results and Analysis}\label{results-and-analysis}}
+\hypertarget{results-1}{%
+\subsection{Results}\label{results-1}}
 
@@
 
-These results present two important pieces of information. First, our
-analysis of the Providence corpus broadly replicates the shape we found
-in both the British National Corpus and in Switchboard. Thus, the
-entropy curve of English appears not just in adult-adult conversation,
-but even in speech produced by parents to their children, and speech
-produced by very young children to their parents. This suggests that it
-is a highly robust feature of English langauge, as it structures the
-productions of even pre-schooled aged children.
+These results present two important pieces of information. First, the
+Providence corpus entropy curve broadly replicates the curves found in
+both the British National Corpus and in Switchboard. Thus, the entropy
+curve of English appears not just in adult-adult conversation, but even
+in speech produced by parents to their children, and speech produced by
+very young children to their parents. This suggests that it is a highly
+robust feature of English language, as it structures the productions of
+even pre-schooled aged children.
 
-Second, the Entropy curves of English, Spanish, and Mandarin were not
-identical. None of these shapes resembled the monotonically increasing
-function predicted naively from Uniform Information Density, but also
-Mandarin was quite different from both English and Spanish. This
-suggests that the entropy curve can vary from language to language. One
-possibility is that this variation arises from typological features of
-languages, such as characteristic word order. We noticed, for instance,
-a relatively high density of determiners in the penultimate position of
-English and Spanish utterances, which could account for the penultimate
-dip in those languages. In our final analysis, we explored this
-possibility directly, analyzing a large set of Wikipedia corpora from
-diverse languages, and asking whether variability in their entropy
-curves was related to variability in their syntactic structure.
+Second, the entropy curves of English, Spanish, and Mandarin were not
+identical. No shape resembled the affine function predicted naively from
+Uniform Information Density, but also Mandarin was quite different from
+both English and Spanish. This suggests that the entropy curve can vary
+from language to language. One possibility is that this variation arises
+from typological features of languages, such as characteristic word
+order. We noticed, for instance, a relatively high density of
+determiners in the penultimate position of English and Spanish
+utterances, which could account their characteristic penultimate dip. In
+our final analysis, we explored this possibility directly, analyzing a
+large set of Wikipedia corpora from diverse languages, and asking
+whether variability in their entropy curves was related to variability
+in their syntactic structure.
 
-\hypertarget{experiment-3}{%
-\section{Experiment 3}\label{experiment-3}}
+\hypertarget{study-3}{%
+\section{Study 3}\label{study-3}}
 
-The UID hypothesis also applies to written communication: we expect
-people to communicate information at a uniform rate through writing as
-well. We use Wikipedia as a source for written data, which provides two
-advantages. One, the quantity of data in Wikipedia is very large for
-each language, at least several megabytes, and two, there are hundreds
-of languages with their own Wikipedia corpora. We can therefore run our
-entropy analysis on a much greater scale than using small spoken CHILDES
-corpora and directly compare the results of the entropy analysis on each
-language to one another. We expect to find linguistic features that
-determine the shape of a language's entropy curve. We treated sentences
-in the Wikipedia text corpora as equivalent to utterances in the
-Switchboard and CHILDES corpora.
+In order to understand how entropy curves are related to linguistic
+structure, we turned to Wikipedia for a larger and more diverse of set
+of languages.
 
-\hypertarget{methods}{%
-\subsection{Methods}\label{methods}}
+\hypertarget{method}{%
+\subsection{Method}\label{method}}
 
-Using Giuseppe Attardi's Wikiextractor tool
-\footnote{https://github.com/attardi/wikiextractor}, we extract text
-corpora from Wikipedia by downloading a stored collection of Wikipedia
-entries in each langauge and randomly selecting several thousand
-articles from each Wikipedia language. Each language corpus was cleaned
-and limited to sentences between six and \(50\) words. Similar to our
-process for the Switchboard and CHILDES spoken corpora, we divided each
-corpus by sentence length, and then computed our positional entropy
-measure on each word position within each sentence length.
+\hypertarget{data-and-analysis-2}{%
+\subsubsection{Data and Analysis}\label{data-and-analysis-2}}
 
-We computed two slope treatments of each curve. In the \emph{absolute}
-treatment, with sentence length denoted as \(k\), we computed the slope
-between positions \(1\) and \(2\), positions \(2\) and \(3\), positions
-\(3\) and \(k-2\), positions \(k-2\) and \(k-1\) and positions \(k-1\)
-and \(k\). For the short utterances appearing the CHILDES speech
-corpora, the slopes between the first and second slopes on either end of
-the distributions appeared to be more characteristic of the distribution
-as a whole, with a plateau in the middle of the entropy curve for each
-of the language corpora we examined in CHILDES.
+Wikipedia has two primary advantages: (1) Individual language corpora
+are generally large, and (2) hundreds of languages are represented. To
+understand how variability in entropy curves is related to the structure
+of language, we used features from the World Atlas of Language
+Structures (WALS, Dryer \& Haspelmath, 2013). WALS is The World Atlas of
+Language Structures is a curated collection of typological features
+across the world's languages. However, as the collection includes
+contributions from many authors studying a heterogeneous set of topic
+and languages, most features are missing for most languages. For our
+sample, we sought to balance diversity of languages with coverage in
+WALS. We selected \(45\) languages for which we \(8\) WALS features were
+coded (Figure \ref{fig:clust_tree}). Unfortunately, these features all
+reflected the order of words in a sentence in each langauge
+(e.g.~whether objects come before verbs, whether adjectives come before
+nouns). A consistent set of results across these features would be
+consistent with an effect of order on entropy curves, but unfortunately
+no other feature types are available for comparison.
 
-However, sentences of length greater than ten in the Wikipedia corpora
-were significantly more common than in the CHILDES. Therefore we also
-computed relative slope treatments. In the \emph{relative 5} treatment,
-we computed the slopes between every \(20\%\) of the relative word
-positions in each sentence length, e.g.~the slope between \(0\%\) and
-\(20\%\) of sentences of length \(10\) is the slope between the position
-entropies at the sentence-initial word position and the position
-representing the third word. When computing the word position index to
-make those cuts, if the slope was not a whole number, then the closest
-whole number position was used instead for slope calculation. Each
-comparative slope within each treatment was averaged together between
-different sentence lengths, for example all of the \(0\%\) to \(20\%\)
-slopes over all sentence lengths were averaged together.
+We computed entropy curves separately for each language as before. In
+order to understand how these curves varied across languages, we
+developed a method for compressing the curves into a few representative
+values. We noticed that curves tended to vary in predictable utterance
+positions: The majority were flat in the middle of utterances, but had
+either rising or falling slopes at the beginnings and ends. For each
+language, for each utterance length \(k\) we computed 5 slopes: The
+change between the 1st and 2nd position, the change between the 2nd and
+3rd position, the slope between the 3rd position and the 3rd-to-last
+position (\(k-1\)), the slope between the 3rd-to-last and the
+2nd-to-last position (\(k-1\)), and the slope between the 2nd-to-last
+and final positions. We then averaged these slopes across utterance
+lengths within a language to estimate a language-typical signature. We
+performed an additional analysis dividing each utterance into 6ths
+evenly and computing the slopes between these 6ths. Results from this
+analysis were qualitatively similar.
 
-In both the absolute and relative \(5\) treatments, each language is
-embedded in \(5\)-dimensional space. This permitted us to use cosine
-similarity for direct comparison between the languages we pulled from
-Wikipedia. To determine which phonological, morphological and syntactic
-features affected the embedding of a language in the Wikipedia dataset,
-we use the linguistic features in World Atlas of Language Structures
-Dryer \& Haspelmath (2013). We checked the effects of individual
-features on the embeddings of languages in the different treatments. We
-computed pairwise cosine similarity between each pair of language
-vectors within each treatment. For a subset of eight WALS features, we
-used a generalized linear model to see whether the cosine similarity
-between languages mattered in predicting if the languages shared the
-same value for a WALS feature.
+\hypertarget{results-and-analysis}{%
+\subsection{Results and Analysis}\label{results-and-analysis}}
 
-\hypertarget{results-and-analysis-1}{%
-\subsection{Results and Analysis}\label{results-and-analysis-1}}
+We treated these 5 slopes as a vector, allowing us to compute the
+pairwise similarity between languages with cosine--a standard measure of
+vector similarity (Landauer \& Dumais, 1997). Intuitively, vectors with
+with a smaller angle between them are more similar. Figure
+\ref{fig:clust_tree} shows a hierarchical clustering of these 5
+dimensional vectors, demonstrating that entropy curve similarity appears
+to capture typological relatedly, e.g.~Slavic languages are similar to
+each-other.
 
-For eight features and \(45\) languages we pulled from Wikipedia, we
-found that the cosine similarity between WALS features. The table below
-shows the results for the generalized linear model we computed using
-WALS features and the absolute treatment.
+We then used asked whether these pairwise similarities are related to
+overlap in linguistic features. For each of the 8 features we measured,
+we fit a logistic regression predicting whether two languages would have
+the same value for it from their pairwise entropy curve similarity.
+Overlap on all 8 features was reliably predicted by entropy curve
+similarity (largest \(p = .001\)), with the amount of variance in
+feature overlap predicted varying across predictors
+\ref{tab:tab_features}. Order of object and verb (\(83A\)) was the
+strongest predictor, explaining 16.22\% of the variance in pairwise
+entropy curve similarity.
+
+\begin{CodeChunk}
+\begin{figure}[tb]
+
+{\centering \includegraphics{figs/clust_tree-1}
+
+}
+
+\caption[A dendrogram estimated from hierarchically clustering the slopes of our 45 languages]{A dendrogram estimated from hierarchically clustering the slopes of our 45 languages}\label{fig:clust_tree}
+\end{figure}
+\end{CodeChunk}
 
@@
 \centering
-\begin{tabular}{llrrrr}
+\begin{tabular}{lrrr}
   \hline
-feature & term & estimate & std.error & statistic & p.value \\ 
+\# & t-value & p-value & r2 \\ 
   \hline
-83A & cosine & 1.84 & 0.01 & 178.22 & 0.00 \\ 
-  95A & cosine & 1.90 & 0.01 & 170.98 & 0.00 \\ 
-  81A & cosine & 1.51 & 0.01 & 153.70 & 0.00 \\ 
-  97A & cosine & 1.58 & 0.01 & 136.81 & 0.00 \\ 
-  144A & cosine & 0.88 & 0.01 & 74.43 & 0.00 \\ 
-  138A & cosine & 0.40 & 0.01 & 46.12 & 0.00 \\ 
-  87A & cosine & 0.33 & 0.01 & 39.91 & 0.00 \\ 
-  143A & cosine & 0.37 & 0.01 & 39.74 & 0.00 \\ 
-  82A & cosine & 0.03 & 0.01 & 3.23 & 0.00 \\ 
+83A & 178.22 & 0.00 & 0.16 \\ 
+  95A & 170.98 & 0.00 & 0.15 \\ 
+  81A & 153.70 & 0.00 & 0.13 \\ 
+  97A & 136.81 & 0.00 & 0.10 \\ 
+  144A & 74.43 & 0.00 & 0.03 \\ 
+  87A & 39.91 & 0.00 & 0.01 \\ 
+  143A & 39.74 & 0.00 & 0.01 \\ 
+  82A & 3.23 & 0.00 & 0.00 \\ 
    \hline
 \end{tabular}
+\caption{Capptoin here} 
+\label{tab:tab_features}
 \end{table}
 
-We also used the linear model approach to evaluate the effects of cosine
-distance in the relative \(5\) treatment on the values of the same WALS
-features and found similar outcomes.
+\hypertarget{general-discussion}{%
+\section{General Discussion}\label{general-discussion}}
 
-From these results, we argue that cosine distance played some role in
-the feature determination. The top four features in determining the WALS
-feature value all characterize word order, which indicates that for this
-subset of features and languages that the embeddings in the slope space
-are related to the WALS features. We argue therefore that typological
-features play a role in determining the positional entropy values for a
-language. This indicates that the entropy curves are in some way
-structured by the syntactic, morphological and phonological features of
-a language.
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-In this paper, we have extended and applied a model from Yu et al.
-(2016) and derived from Genzel \& Charniak (2002) for distinguishing
-entropy at each word position within sentences. We have argued that the
-outcomes of this model are derived from the distribution of information
-transmission in human communication. Language is not a uniform noisy
-channel of communication, but each language individually represents a
-different noisy channel of communication characterized in part by
-typological features.
-
-Our work complements the approach of studies such as Aylett \& Turk
-(2004), where languages are characterized by a single number,
-representing the rate of semantic information transfer per syllable.
-This body of work attests to languages possessing different rates of
-information transfer based on typological features as well. Within the
-information transfer rate literature, the rate is an overall feature of
-language that does not vary corpus to corpus. Similarly, in our work we
-have obtained a characteristic entropy distribution for each language.
-Both information rate and entropy distribution are characterized by
-typological features. This indicates that the structure and rate of how
-people communicate information varies cross-linguistically based on
-typological features.
-
-We initially wanted to use several hundred language corpora pulled from
-Wikipedia and all \(144\) linguistic features from WALS. We considered
-an computing unsupervized clusterings of the \(5\)-dimensional vectors
-of our Wikipedia entropy curves, and comparing those clusterings to
-clusterings of the WALS features directly using a cluster similarity
-measure such as the Rand index Rand (1971). However, the problem then
-arises of which combination of WALS features and how many features to
-include in the clustering analysis, with \(144!\) different
-combinations. Two additional problems presented themselves: one,
-deciding how many clusters to use in an unsupervised clustering analysis
-is an unsolved problem in machine learning; and two, not all of the
-languages on Wikipedia have values inputted for their WALS features. As
-a follow-up, we are considering missing-data imputation on the WALS
-features in order to run the generalized linear model analysis on a
-large dataset of languages.
-
-On a more mechanical level, we expect that our entropy model will have
-implications for how long people take to read individual words at
-different positions of a sentence. Eye-tracking research has indicated
-the effects of surprisal on fixation duration during eye-tracking
-studies (Boston, Hale, Kliegl, Patil, \& Vasishth, 2008; Demberg \&
-Keller, 2008; Smith \& Levy, 2013); higher surprisal of a word is on
-average predictive of higher fixation duration. The \emph{wrap-up
-effect} states that people process sentence-final words more slowly on
-average then sentence-medial or sentence-initial words when reading, due
-to integrating information to form a final understanding of the
-sentence's meaning (Kuperman, Dambacher, Nuthmann, \& Kliegl, 2010;
-Stowe, Kaan, Sabourin, \& Taylor, 2018). The wrap-up effect is drawn
-from evidence in languages with a large final increase in their entropy
-curve from our study, so we hypothesize that the supposed wrap-up effect
-derives from the same source sentence-final increase in entropy curves
-observed in English and other Germanic languages.
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-{[}Not here yet.{]}
+Taken together, the results of our studies yield 3 main findings. First,
+languages have characteristic entropy curves that can be found in their
+productions whether written, spoken, or even spoken by children. These
+curves diverge in reliabile ways from the monotonically increasing
+functions predicting by a naive Uniform Information Density hypothesis.
+Second, these characteristic curves vary across languages--some
+languages like English have increasing curves, but others like Mandarin
+are characterized by decreasing entropy of the course of an utterance.
+Finally, these characteristic curves are related to structural
+properties of those languages, for instance their word orders. This work
+presents an important rejoinder to models of linguistic productions as
+optimizing communicative success. While speakers may do their best to
+structure the information in their productions to minimize spikes in
+entropy, they operate under coding models they inherit from their native
+language, and these coding models constrain the choices they make.
 
@@
 
-\leavevmode\hypertarget{ref-aylett2004}{}%
-Aylett, M., \& Turk, A. (2004). The smooth signal redundancy hypothesis:
-A functional explanation for relationships between redundancy, prosodic
-prominence, and duration in spontaneous speech. \emph{Language and
-Speech}, \emph{47}(1), 31--56.
-
 \leavevmode\hypertarget{ref-bai2008}{}%
@@
 
-\leavevmode\hypertarget{ref-boston2008}{}%
-Boston, M. F., Hale, J., Kliegl, R., Patil, U., \& Vasishth, S. (2008).
-Parsing costs as predictors of reading difficulty: An evaluation using
-the potsdam sentence corpus. \emph{Journal of Eye Movement Research},
-\emph{2}(1).
-
 \leavevmode\hypertarget{ref-clear1993}{}%
@@
 World}, 163--187.
-
-\leavevmode\hypertarget{ref-demberg2008}{}%
-Demberg, V., \& Keller, F. (2008). Data from eye-tracking corpora as
-evidence for theories of syntactic processing complexity.
-\emph{Cognition}, \emph{109}(2), 193--210.
 
@@
 
-\leavevmode\hypertarget{ref-kuperman2010}{}%
-Kuperman, V., Dambacher, M., Nuthmann, A., \& Kliegl, R. (2010). The
-effect of word position on eye-movements in sentence and paragraph
-reading. \emph{The Quarterly Journal of Experimental Psychology},
-\emph{63}(9), 1838--1857.
+\leavevmode\hypertarget{ref-landauer1997}{}%
+Landauer, T. K., \& Dumais, S. T. (1997). A solution to plato's problem:
+The latent semantic analysis theory of acquisition, induction, and
+representation of knowledge. \emph{Psychological Review}, \emph{104}(2),
+211.
 
@@
 
-\leavevmode\hypertarget{ref-rand1971}{}%
-Rand, W. M. (1971). Objective criteria for the evaluation of clustering
-methods. \emph{Journal of the American Statistical Association},
-\emph{66}(336), 846--850.
-
 \leavevmode\hypertarget{ref-sanchez2019}{}%
@@
 Languages of the world, 21st edition}. Dallas, Texas: SIL International.
-
-\leavevmode\hypertarget{ref-smith2013}{}%
-Smith, N. J., \& Levy, R. (2013). The effect of word predictability on
-reading time is logarithmic. \emph{Cognition}, \emph{128}(3), 302--319.
-
-\leavevmode\hypertarget{ref-stowe2018}{}%
-Stowe, L. A., Kaan, E., Sabourin, L., \& Taylor, R. C. (2018). The
-sentence wrap-up dogma. \emph{Cognition}, \emph{176}, 232--247.
 

### 2019-02-02 wiki_entropy/wiki_clustering.Rmd
everything in

@@
 models %>%
-  mutate(coeffs = map(model, tidy)) %>%
+  mutate(coeffs = map(model, tidy),
+         rs = map(model, )) %>%
   select(-data, -model) %>%

### 2019-02-02 cogsci/position-info.Rmd
figs bigger

@@
 
-```{r read_and_plot_exp1, fig.env = "figure*", fig.width=6, fig.height=3, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom).
+```{r read_and_plot_exp1, fig.env = "figure*", fig.width=6, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom).
 Points average entropies, error bars show 95\\% confidence intervals computed by non-parametric bootstrap."}

### 2019-02-02 cogsci/position-info.Rmd
cogsci revisions, rewrote final discussion

@@
     Although languages vary widely in their structure, all are vehicles for the transmission of information.
-Consequently, aspects of speakers' word choice can be understood through the models of optimal communication.
+Consequently, aspects of speakers' word choice can be understood through models of optimal communication.
 One prediction of these models is that speakers should keep the density of information constant over the duration of an utterance [@levy2007].
-However, different languages have different coding models that constrain the space of a speaker's choices (e.g.
+However, different languages have different structural properties that constrain the space of a speaker's choices (e.g.
 canonical word order).
 We build on a method developed by @yu2016 to analyze the word-level entropy curves of natural language productions across a diverse set of languages and in diverse written and spoken contexts.
-We show languages impose characteristic constraints on speaker's choices that predict deviations from Uniform Information Density, and that cross-linguistic variability in these deviations is predictable in part from syntactic properties of those languages.
+We show that languages impose characteristic constraints on speaker's choices that predict deviations from Uniform Information Density, and that cross-linguistic variability in these deviations is predictable in part from syntactic properties of those languages.
 
@@
 If speakers and writers of these languages act near-optimally to achieve their communicative goals, regularities of use across these diverse languages can be explained by a rational theory of communication [@anderson1991].
-Information theory, a mathematical framework developed by @shannon1948 to describe the transmission and decoding of signals, has been a unifying language for the recent development of such theories in human and machine language processing [@jelinek1976;@levy2007].
+Information theory, a mathematical framework developed by @shannon1948 to describe the transmission and decoding of signals, has been a unifying approach for the recent development of theories of communication in human and machine language processing [@jelinek1976;@levy2007].
 
@@
 The receiver must then decode the producer's intended meaning from the signal they receive on their end of the channel.
-The problem is that the channel is noisy, and sometimes the signal can get corrupted (e.g.
+The problem is that this channel is noisy, and sometimes the signal can be corrupted (e.g.
 the producer can misspeak, or the receiver can mishear).
@@
 
-<!-- **Evidence for UID here. Maybe see if you can present this work at a slightly higher level** -->
 The evidence in favor of Uniform Information Density largely been situation-specific and English-language driven, while the hypothesis itself has been applied broadly over the past decade.
@@
 
-<!-- **Evidence against UID here. Again, can you present this at a higher level?** -->
 However, other recent work has contradicted the Uniform Information Density hypothesis.
@@
 @zhan2018 finds that more information-rich classifiers in Mandarin Chinese are produced when production of the neighboring noun is difficult, not when the information content is high.
-@jain2018 examine word order across spoken sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on determining a Hindi's speaker's word order.
+@jain2018 examine word order across spoken sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on determining a Hindi speaker's word order.
 
@@
 Because people process language incrementally--using the previous words in a sentence to predict the words that will come next--the amount of information that a word contains when seen in isolation should increase over the course of a sentence [@ferrer-i-cancho2013].
-Analyzing a large corpus of written English, they find a different pattern: Entropy increases over the first few words of an utterance and then remains constant until the final word where it again jumps up (see top of Figure \ref{fig:read_and_plot_exp1}).
+Analyzing a large corpus of written English, @yu2016 find a different pattern: Unigram entropy increases over the first few words of an utterance and then remains constant until the final word where it again jumps up (see top of Figure \ref{fig:read_and_plot_exp1}).
 @yu2016 conclude that the Uniform Information Density hypothesis must not hold for medial words in a sentence.
 
-We extend and generalize @yu2016 in three ways: We confirm that this same pattern is found in spoken English--in written language, and in conversational speech between adults and between parents and their children.
-Thus, this entropy curve is a robust feature of English productions.
+We extend and generalize @yu2016 in three ways: We confirm that this same pattern is found in English--in written language, and in conversational speech between adults and between parents and their children.
+Thus, this entropy curve is a robust feature of English linguistic productions.
 We then examine entropy curves cross-linguistically, and show that characteristic curves vary across the world's languages.
@@
 
-With this measure, we compute the unigram entropy at each word position of sentences of each length within the corpus.
-The result of this method can be plotted for each length as an *entropy curve*, which can be visually compared across sentence lengths to observe the how the unigram entropy changes across absolute positions in each of the sentences.
+With this measure, we can compute the unigram entropy at each word position of sentences of each length within a corpus.
+The result of this method can be plotted for each sentence length as an *entropy curve*, which can be visually compared across sentence lengths to observe how the unigram entropy changes across absolute positions in each of the sentences.
 @genzel2002 similarly examine a unigram entropy measure on sentences, and found that entropy at the sentence level increases linearly with sentence index within a corpus.
-Uniform Information Density applies this uniformity of entropy rate in sentences to all levels of speech, and so our method obtained from @yu2016, which examines text at the word level, should find a monotonically increasing affine function at the word level.
+Uniform Information Density applies this uniformity of entropy rate in sentences to all levels of speech [@levy2007], and so our method obtained from @yu2016, which examines text at the word level, should find a monotonically increasing affine function at the word level.
 
@@
 We began by replicating Yu et al.'s [-@yu2016] analysis, computing entropy curves on the British National Corpus--a collection of predominantly written English [@clear1993].
-We then applied the same method to the Switchboard corpus--a collection of spoken language [@godfrey1992].
-This allows us to ask whether the characteristic function identified by @yu2016 is a general feature of English, or instead a function of written language.
+We then applied the same method to the Switchboard corpus--a collection of spoken English [@godfrey1992].
+This allowed us to ask whether the characteristic function identified by @yu2016 is a general feature of English, or instead a function of written language.
 
@@
 
-The British National Corpus consists of predominantly (90%) written language documents collected in the early 1980s and 1990s across a variety of genres (scientific articles, newspapers, fiction, etc).
+The British National Corpus consists of predominantly (90%) written documents collected in the early 1980s and 1990s across a variety of genres (scientific articles, newspapers, fiction, etc).
 It also contains a small collection of corpora from spoken language.
-All together, it contains $\sim$ 100 million tokens.
+All together, the British National Corpus contains $\sim$ 100 million tokens.
 The Switchboard corpus is a collection of $\sim$ 2,400 telephone conversations between unacquainted adults prompted with subjects of conversation.
@@
 Confirming Yu et al.'s [-@yu2016] findings, we find that
-the positional entropy at the beginning of sentences is low, then rises and plateaus for sentence-medial word positions before dropping slightly in the second-to-last position and rising again.
+the positional entropy at the beginning of sentences is low, then rises and plateaus for sentence-medial word positions before dropping slightly in the second-to-last position and rising again for the sentence-final word.
 
@@
 
-The shape of positional entropy curves we find in these two corpora is notable for two reasons: (1) it does not follow our predictions from Uniform Information Density, and (2) the distribution is robust across written and spoken Engilsh.
+The shape of positional entropy curves we find in these two corpora is notable for two reasons: (1) it does not follow our predictions from Uniform Information Density, and (2) the distribution is robust across written and spoken English.
 This suggests that the entropy curve is characteristic of the English language.
@@
 To understand how robust this entropy curve is, we turned to conversational speech from parent-child interactions.
-If we find the same shape even in children's productions, we have even stronger evidence that the 3-step curve is a charcateristic feature of English.
-We thus turned to the Child Language Data Exchange System (CHILDES), a collection of transcripts of parent-child interactions [@macwhinney2014].
+If we find the same shape even in children's productions, we have even stronger evidence that the three-step entropy curve is a characteristic feature of English.
+We thus used the Child Language Data Exchange System (CHILDES), a collection of transcripts of parent-child interactions [@macwhinney2014].
 
@@
 
-We analyzed three seperate corpora in CHILDES: The Providence Corpus [@demuth2006], The Shiro corpus [@shiro2000], and the Zhou Dinner Corpus [@li2015].
+We analyzed three separate corpora in CHILDES: The Providence Corpus [@demuth2006], The Shiro corpus [@shiro2000], and the Zhou Dinner Corpus [@li2015].
 The Providence corpus consists of conversations between six 1--3-year-old American English-speaking children and their parents recorded in their homes.
-The Shiro Corpus consists of prompted Spanish-language narratives individually collected from over a hundred Venezualan schoolchildren, half from high SES backgrounds and half from low SES backgrounds.
+The Shiro Corpus consists of prompted Spanish-language narratives individually collected from over a hundred Venezuelan schoolchildren.
 The Zhou Dinner Corpus contains dinner conversations between 5--6-year-old Mandarin-speaking children and their parents collected in Shanghai.
-Spanish is an Indo-European language like English, possessing similar grammar, word order and numerous cognate words.
-In contrast, Mandarin Chinese is typologically unrelated to English.
+Spanish is an Indo-European language like English, possessing similar grammar, word order and numerous cognate words to English.
+In contrast, Mandarin Chinese is typologically unrelated to English, with many structural differences between the two languages.
 
-We accessed the transcripts from each corpus using \texttt{childesr}, an R-interface to a database formatted version of CHILDES [@sanchez2019].
-We divided all utterances from each transcript into those produced by target child, and those produced by all other speakers.
-We then applied the same entropy curve method described above.
+We accessed the transcripts from each corpus using \texttt{childesr}, an R-interface to a database-formatted version of CHILDES [@sanchez2019].
+We divided all utterances from each transcript into those produced by the target child, and those produced by all other speakers.
+We then applied the same entropy curve method described earlier.
 For Mandarin, we used pinyin transliterations of the utterances in the corpus with demarcated word boundaries.
-The Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Chinese writing including spaces between word boundaries is unnatural and can have a negative effect on reading times [@bai2008].
+The Chinese characters used for writing Mandarin do not normally demarcate word boundaries by spacing words apart, and for normal Chinese writing spaces between word boundaries are unnatural and such spaces can be difficult to determine [@bai2008].
 
@@
 Across corpora, we found similar entropy curve shapes for adults and children, but distinct shapes for each language.
-Figure \ref{fig:plot_childes} shows representative curves across corpora, but shapes were robust across the full range utterances we analyzed.
-We found a distinct three-step distribution for English and Spanish CHILDES corpora, with a slight dip in the penultimate position of each sentence.
-The Mandarin corpus entropy curve, by comparison, has a noticeably lower positional entropy values in utterance-final positions than in utterance-penultimate positions.
+Figure \ref{fig:plot_childes} shows representative curves across corpora, but the entropy curve shapes were robust across the full range of utterances we analyzed.
+We found a distinct three-step distribution for both the English and Spanish CHILDES corpora, with a slight dip in the penultimate position of each sentence.
+The Mandarin corpus entropy curve, by comparison, displayed a noticeably lower positional entropy values in utterance-final positions than in utterance-penultimate positions.
 
 These results present two important pieces of information.
-First, the Providence corpus entropy curve broadly replicates the curves found in both the British National Corpus and in Switchboard.
+First, the English Providence corpus entropy curve broadly replicates the curves found in both the British National Corpus and in Switchboard, our two previous English corpora.
 Thus, the entropy curve of English appears not just in adult-adult conversation, but even in speech produced by parents to their children, and speech produced by very young children to their parents.
-This suggests that it is a highly robust feature of English language, as it structures the productions of even pre-schooled aged children.
+This suggests that it is a highly robust feature of the English language, as it reflects the structure of even pre-schooled aged children's speech.
 
@@
 This suggests that the entropy curve can vary from language to language.
-One possibility is that this variation arises from typological features of languages, such as characteristic word order.
+One possibility is that this variation arises from typological features of language, such as characteristic word order.
 We noticed, for instance, a relatively high density of determiners in the penultimate position of English and Spanish utterances, which could account their characteristic penultimate dip.
@@
 
-## Method
-
-### Data and Analysis
+## Data and Analysis
 
@@
 To understand how variability in entropy curves is related to the structure of language, we used features from the World Atlas of Language Structures [WALS, @dryer2013].
-WALS is The World Atlas of Language Structures is a curated collection of typological features across the world's languages.
-However, as the collection includes contributions from many authors studying a heterogeneous set of topic and languages, most features are missing for most languages.
+WALS is a curated collection of typological features which analyze and classify the diversity across the world's languages.
+However, as the collection includes contributions from many authors studying a heterogeneous set of topics and languages, most features are missing for most languages.
 For our sample, we sought to balance diversity of languages with coverage in WALS.
-We selected $45$ languages for which we $8$ WALS features were coded (Figure \ref{fig:clust_tree}).
-Unfortunately, these features all
-reflected the order of words in a sentence in each langauge (e.g.
+We selected $45$ languages for which eight specific WALS features were coded (Figure \ref{fig:clust_tree}).
+Unfortunately, these features all reflected the order of words in a sentence in each language (e.g.
 whether objects come before verbs, whether adjectives come before nouns).
@@
 We then averaged these slopes across utterance lengths within a language to estimate a language-typical signature.
-We performed an additional analysis dividing each utterance into 6ths evenly and computing the slopes between these 6ths.
-Results from this analysis were qualitatively similar.
+We performed an additional analysis dividing each utterance into fifths evenly and computing the slopes between these fifths.
+Results from this second analysis were qualitatively similar.
 
-## Results and Analysis
+## Results
 
 We treated these 5 slopes as a vector, allowing us to compute the pairwise similarity between languages with cosine--a standard measure of vector similarity [@landauer1997].
-Intuitively, vectors with with a smaller angle between them are more similar.
+Intuitively, vectors with a smaller angle between them are more similar.
 Figure \ref{fig:clust_tree} shows a hierarchical clustering of these 5 dimensional vectors, demonstrating that entropy curve similarity appears to capture typological relatedly, e.g.
@@
 ```{r load_features}
-features <- read_csv(here("data/wals_features.csv"))
+features <- read_csv(here("data/WALS_features.csv"))
 
@@
 
-We then used asked whether these pairwise similarities are related to overlap in linguistic features.
-For each of the 8 features we measured, we fit a logistic regression predicting whether two languages would have the same value for it from their pairwise entropy curve similarity.
+We then asked whether these pairwise similarities were related to overlap in linguistic features.
+For each of the 8 WALS features we measured, we fit a logistic regression predicting whether two languages would have the same value for that feature from the similarity of their pairwise entropy curves.
 Overlap on all 8 features was reliably predicted by entropy curve similarity (largest $p = .001$), with the amount of variance in feature overlap predicted varying across predictors \ref{tab:tab_features}.
@@
   select(-estimate,-meaning)  %>%
-  xtable(caption = "Capptoin here",
+  xtable(caption = "Table showing the eight WALS features ranked by significance of effect on the entropy slope cosine distances",
          label = "tab:tab_features")
@@
 First, languages have characteristic entropy curves that can be found in their productions whether written, spoken, or even spoken by children.
-These curves diverge in reliabile ways from the monotonically increasing functions predicting by a naive Uniform Information Density hypothesis.
-Second, these characteristic curves vary across languages--some languages like English have increasing curves, but others like Mandarin are characterized by decreasing entropy of the course of an utterance.
+These curves diverge in reliable ways from the affine functions predicted by a naive Uniform Information Density hypothesis.
+Second, these characteristic curves vary across languages--some languages like English have increasing curves, but others like Mandarin are characterized by decreasing entropy over the course of an utterance.
 Finally, these characteristic curves are related to structural properties of those languages, for instance their word orders.
 This work presents an important rejoinder to models of linguistic productions as optimizing communicative success.
-While speakers may do their best to structure the information in their productions to minimize spikes in entropy, they operate under coding models they inherit from their native language, and these coding models constrain the choices they make.
+While speakers may structure the information in their productions to minimize spikes in entropy, speakers operate under coding models they inherit from their language, and these coding models constrain the choices they make.
+
+Our work complements the approach of studies such as @aylett2004, where a language's information distribution is characterized by a single number representing the average rate of semantic information transfer per syllable.
+Each language possesses a characteristic entropy curve derived from the information distribution within utterances as well as a measure of the average information expressed by a syllable in that language.
+
+We expect the characteristic entropy curve for a language to have important downstream effects for how people process that language.
+While conversational turn-taking is universal, there exists cross-linguistic variation in the average length of time between conversational partners speaking [@stivers2009universals].
+We predict that one factor in this variability relates to the shape of a language's entropy curve.
+A jump at the end of a language's entropy curve, such as in English, indicates that there is on average high levels of information at the end of utterances in that language.
+Therefore we expect turn-taking times to be longer in such languages than in languages where there is a drop in the final segment of the entropy curve, such as Mandarin.
+We also expect that readers of a language will take longer on average to read parts of sentences where the entropy curve is relatively high, due to the presence of higher information content.
+This may explain the so-called wrap-up effect in sentence processing, in which the end of the sentence in English and other Germanic languages is processed more slowly during reading [@stowe2018;@kuperman2010].
 

### 2019-02-02 cogsci/position-info.tex
cogsci revisions, rewrote final discussion

@@
 the transmission of information. Consequently, aspects of speakers' word
-choice can be understood through the models of optimal communication.
-One prediction of these models is that speakers should keep the density
-of information constant over the duration of an utterance (Levy \&
-Jaeger, 2007). However, different languages have different coding models
-that constrain the space of a speaker's choices (e.g.~canonical word
-order). We build on a method developed by Yu, Cong, Liang, \& Liu (2016)
-to analyze the word-level entropy curves of natural language productions
-across a diverse set of languages and in diverse written and spoken
-contexts. We show languages impose characteristic constraints on
-speaker's choices that predict deviations from Uniform Information
-Density, and that cross-linguistic variability in these deviations is
-predictable in part from syntactic properties of those languages.
+choice can be understood through models of optimal communication. One
+prediction of these models is that speakers should keep the density of
+information constant over the duration of an utterance (R. P. Levy \&
+Jaeger, 2007). However, different languages have different structural
+properties that constrain the space of a speaker's choices
+(e.g.~canonical word order). We build on a method developed by Yu, Cong,
+Liang, \& Liu (2016) to analyze the word-level entropy curves of natural
+language productions across a diverse set of languages and in diverse
+written and spoken contexts. We show that languages impose
+characteristic constraints on speaker's choices that predict deviations
+from Uniform Information Density, and that cross-linguistic variability
+in these deviations is predictable in part from syntactic properties of
+those languages.
 
@@
 mathematical framework developed by Shannon (1948) to describe the
-transmission and decoding of signals, has been a unifying language for
-the recent development of such theories in human and machine language
-processing (Jelinek, 1976; Levy \& Jaeger, 2007).
+transmission and decoding of signals, has been a unifying approach for
+the recent development of theories of communication in human and machine
+language processing (Jelinek, 1976; R. P. Levy \& Jaeger, 2007).
 
@@
 receiver must then decode the producer's intended meaning from the
-signal they receive on their end of the channel. The problem is that the
-channel is noisy, and sometimes the signal can get corrupted (e.g.~the
-producer can misspeak, or the receiver can mishear). In order to
-maximize the probability that the correct meaning is transmitted, these
-theories predict that producers should choose linguistic messages that
-keep the rate of information constant across words. The intuition is
-that if the receiver misperceives a word, and that word contains most of
-the information in the sentence, then the communication will have
+signal they receive on their end of the channel. The problem is that
+this channel is noisy, and sometimes the signal can be corrupted
+(e.g.~the producer can misspeak, or the receiver can mishear). In order
+to maximize the probability that the correct meaning is transmitted,
+these theories predict that producers should choose linguistic messages
+that keep the rate of information constant across words. The intuition
+is that if the receiver misperceives a word, and that word contains most
+of the information in the sentence, then the communication will have
 failed. Because producers cannot predict which word a speaker will
@@
 of the words in a sentence, i.e.~maintain \emph{Uniform Information
-Density} (Genzel \& Charniak, 2002; Levy \& Jaeger, 2007).
+Density} (Genzel \& Charniak, 2002; R. P. Levy \& Jaeger, 2007).
 
@@
 itself has been applied broadly over the past decade. The original
-evidence in favor of Uniform Information Density in Levy \& Jaeger
+evidence in favor of Uniform Information Density in R. P. Levy \& Jaeger
 (2007) finds that the insertion of complementizers (e.g. ``that'') in
 relative clauses in English corresponds to where neighboring words have
-high information content. Similarly, Frank \& Jaeger (2008) argues that
-contractions in English such as ``you're'' do not occur when neighboring
-words are highly informative. Applications of Uniform Information
-Density include determining whether linguistic alignment takes place
-(Jaeger \& Snider, 2013), Zipfian word length distributions (Piantadosi,
-Tily, \& Gibson, 2011), communication efficiency (Mahowald, Fedorenko,
-Piantadosi, \& Gibson, 2013), dialogue and turn-taking (Xu \& Reitter,
-2018) and the significance of ambiguity in language (Piantadosi, Tily,
-\& Gibson, 2012), among other research.
+high information content. Similarly, A. F. Frank \& Jaeger (2008) argues
+that contractions in English such as ``you're'' do not occur when
+neighboring words are highly informative. Applications of Uniform
+Information Density include determining whether linguistic alignment
+takes place (Jaeger \& Snider, 2013), Zipfian word length distributions
+(Piantadosi, Tily, \& Gibson, 2011), communication efficiency (Mahowald,
+Fedorenko, Piantadosi, \& Gibson, 2013), dialogue and turn-taking (Xu \&
+Reitter, 2018) and the significance of ambiguity in language
+(Piantadosi, Tily, \& Gibson, 2012), among other research.
 
 However, other recent work has contradicted the Uniform Information
-Density hypothesis. Similar to the original Levy \& Jaeger (2007), Zhan
-\& Levy (2018) focuses on information distribution at particular points
-in sentences. Zhan \& Levy (2018) finds that more information-rich
-classifiers in Mandarin Chinese are produced when production of the
-neighboring noun is difficult, not when the information content is high.
-Jain, Singh, Ranjan, Rajkumar, \& Agarwal (2018) examine word order
-across spoken sentences in Hindi, a freer word order language than
-English, and find that information density has no significant effect on
-determining a Hindi's speaker's word order.
+Density hypothesis. Similar to the original R. P. Levy \& Jaeger (2007),
+Zhan \& Levy (2018) focuses on information distribution at particular
+points in sentences. Zhan \& Levy (2018) finds that more
+information-rich classifiers in Mandarin Chinese are produced when
+production of the neighboring noun is difficult, not when the
+information content is high. Jain, Singh, Ranjan, Rajkumar, \& Agarwal
+(2018) examine word order across spoken sentences in Hindi, a freer word
+order language than English, and find that information density has no
+significant effect on determining a Hindi speaker's word order.
 
@@
 sentence (Ferrer-i-Cancho, Debowski, \& Prado Martin, 2013). Analyzing a
-large corpus of written English, they find a different pattern: Entropy
-increases over the first few words of an utterance and then remains
-constant until the final word where it again jumps up (see top of Figure
-\ref{fig:read_and_plot_exp1}). Yu et al. (2016) conclude that the
-Uniform Information Density hypothesis must not hold for medial words in
-a sentence.
+large corpus of written English, Yu et al. (2016) find a different
+pattern: Unigram entropy increases over the first few words of an
+utterance and then remains constant until the final word where it again
+jumps up (see top of Figure \ref{fig:read_and_plot_exp1}). Yu et al.
+(2016) conclude that the Uniform Information Density hypothesis must not
+hold for medial words in a sentence.
 
 We extend and generalize Yu et al. (2016) in three ways: We confirm that
-this same pattern is found in spoken English--in written language, and
-in conversational speech between adults and between parents and their
+this same pattern is found in English--in written language, and in
+conversational speech between adults and between parents and their
 children. Thus, this entropy curve is a robust feature of English
-productions. We then examine entropy curves cross-linguistically, and
-show that characteristic curves vary across the world's languages.
-Finally, we show that this variation is predictable in part from the
-structure of individual languages (i.e.~word order). Taken together, our
-results suggest a refinement of the Uniform Information Density
-hypothesis: speakers may structure their utterances to optimize
-information density, but they must do so under the predictable
-constraints of their language.
+linguistic productions. We then examine entropy curves
+cross-linguistically, and show that characteristic curves vary across
+the world's languages. Finally, we show that this variation is
+predictable in part from the structure of individual languages
+(i.e.~word order). Taken together, our results suggest a refinement of
+the Uniform Information Density hypothesis: speakers may structure their
+utterances to optimize information density, but they must do so under
+the predictable constraints of their language.
 
-\hypertarget{calculating-entropy-curves}{%
 \subsection{Calculating entropy
-curves}\label{calculating-entropy-curves}}
+curves}\label{calculating-entropy-curves}
 
@@
 
-With this measure, we compute the unigram entropy at each word position
-of sentences of each length within the corpus. The result of this method
-can be plotted for each length as an \emph{entropy curve}, which can be
-visually compared across sentence lengths to observe the how the unigram
-entropy changes across absolute positions in each of the sentences.
-Genzel \& Charniak (2002) similarly examine a unigram entropy measure on
-sentences, and found that entropy at the sentence level increases
-linearly with sentence index within a corpus. Uniform Information
-Density applies this uniformity of entropy rate in sentences to all
-levels of speech, and so our method obtained from Yu et al. (2016),
-which examines text at the word level, should find a monotonically
-increasing affine function at the word level.
+With this measure, we can compute the unigram entropy at each word
+position of sentences of each length within a corpus. The result of this
+method can be plotted for each sentence length as an \emph{entropy
+curve}, which can be visually compared across sentence lengths to
+observe how the unigram entropy changes across absolute positions in
+each of the sentences. Genzel \& Charniak (2002) similarly examine a
+unigram entropy measure on sentences, and found that entropy at the
+sentence level increases linearly with sentence index within a corpus.
+Uniform Information Density applies this uniformity of entropy rate in
+sentences to all levels of speech (R. P. Levy \& Jaeger, 2007), and so
+our method obtained from Yu et al. (2016), which examines text at the
+word level, should find a monotonically increasing affine function at
+the word level.
 
@@
 
-\hypertarget{study-1}{%
-\section{Study 1}\label{study-1}}
+\section{Study 1}\label{study-1}
 
@@
 written English (Clear, 1993). We then applied the same method to the
-Switchboard corpus--a collection of spoken language (Godfrey, Holliman,
-\& McDaniel, 1992). This allows us to ask whether the characteristic
+Switchboard corpus--a collection of spoken English (Godfrey, Holliman,
+\& McDaniel, 1992). This allowed us to ask whether the characteristic
 function identified by Yu et al. (2016) is a general feature of English,
@@
 
-\hypertarget{data-and-analysis}{%
-\section{Data and Analysis}\label{data-and-analysis}}
+\section{Data and Analysis}\label{data-and-analysis}
 
 The British National Corpus consists of predominantly (90\%) written
-language documents collected in the early 1980s and 1990s across a
-variety of genres (scientific articles, newspapers, fiction, etc). It
-also contains a small collection of corpora from spoken language. All
-together, it contains \(\sim\) 100 million tokens. The Switchboard
-corpus is a collection of \(\sim\) 2,400 telephone conversations between
-unacquainted adults prompted with subjects of conversation. Switchboard
-is the corpus used in Levy and Jaeger's (2007) original demonstration of
-the Uniform Information Density Hypothesis.
+documents collected in the early 1980s and 1990s across a variety of
+genres (scientific articles, newspapers, fiction, etc). It also contains
+a small collection of corpora from spoken language. All together, the
+British National Corpus contains \(\sim\) 100 million tokens. The
+Switchboard corpus is a collection of \(\sim\) 2,400 telephone
+conversations between unacquainted adults prompted with subjects of
+conversation. Switchboard is the corpus used in Levy and Jaeger's (2007)
+original demonstration of the Uniform Information Density Hypothesis.
 
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\section{Results}\label{results}
 
@@
 low, then rises and plateaus for sentence-medial word positions before
-dropping slightly in the second-to-last position and rising again.
+dropping slightly in the second-to-last position and rising again for
+the sentence-final word.
 
@@
 Uniform Information Density, and (2) the distribution is robust across
-written and spoken Engilsh. This suggests that the entropy curve is
+written and spoken English. This suggests that the entropy curve is
 characteristic of the English language. We next asked whether this shape
@@
 
-\hypertarget{study-2}{%
-\section{Study 2}\label{study-2}}
+\section{Study 2}\label{study-2}
 
@@
 same shape even in children's productions, we have even stronger
-evidence that the 3-step curve is a charcateristic feature of English.
-We thus turned to the Child Language Data Exchange System (CHILDES), a
-collection of transcripts of parent-child interactions (MacWhinney,
-2014).
+evidence that the three-step entropy curve is a characteristic feature
+of English. We thus used the Child Language Data Exchange System
+(CHILDES), a collection of transcripts of parent-child interactions
+(MacWhinney, 2014).
 
-\hypertarget{data-and-analysis-1}{%
-\subsection{Data and Analysis}\label{data-and-analysis-1}}
+\subsection{Data and Analysis}\label{data-and-analysis-1}
 
-We analyzed three seperate corpora in CHILDES: The Providence Corpus
+We analyzed three separate corpora in CHILDES: The Providence Corpus
 (Demuth, Culbertson, \& Alter, 2006), The Shiro corpus (Shiro, 2000),
@@
 Shiro Corpus consists of prompted Spanish-language narratives
-individually collected from over a hundred Venezualan schoolchildren,
-half from high SES backgrounds and half from low SES backgrounds. The
-Zhou Dinner Corpus contains dinner conversations between 5--6-year-old
-Mandarin-speaking children and their parents collected in Shanghai.
-Spanish is an Indo-European language like English, possessing similar
-grammar, word order and numerous cognate words. In contrast, Mandarin
-Chinese is typologically unrelated to English.
+individually collected from over a hundred Venezuelan schoolchildren.
+The Zhou Dinner Corpus contains dinner conversations between
+5--6-year-old Mandarin-speaking children and their parents collected in
+Shanghai. Spanish is an Indo-European language like English, possessing
+similar grammar, word order and numerous cognate words to English. In
+contrast, Mandarin Chinese is typologically unrelated to English, with
+many structural differences between the two languages.
 
 We accessed the transcripts from each corpus using \texttt{childesr}, an
-R-interface to a database formatted version of CHILDES (Sanchez et al.,
+R-interface to a database-formatted version of CHILDES (Sanchez et al.,
 in press). We divided all utterances from each transcript into those
-produced by target child, and those produced by all other speakers. We
-then applied the same entropy curve method described above. For
+produced by the target child, and those produced by all other speakers.
+We then applied the same entropy curve method described earlier. For
 Mandarin, we used pinyin transliterations of the utterances in the
@@
 writing Mandarin do not normally demarcate word boundaries by spacing
-words apart, and for normal Chinese writing including spaces between
-word boundaries is unnatural and can have a negative effect on reading
-times (Bai, Yan, Liversedge, Zang, \& Rayner, 2008).
+words apart, and for normal Chinese writing spaces between word
+boundaries are unnatural and such spaces can be difficult to determine
+(Bai, Yan, Liversedge, Zang, \& Rayner, 2008).
 
-\hypertarget{results-1}{%
-\subsection{Results}\label{results-1}}
+\subsection{Results}\label{results-1}
 
@@
 \ref{fig:plot_childes} shows representative curves across corpora, but
-shapes were robust across the full range utterances we analyzed. We
-found a distinct three-step distribution for English and Spanish CHILDES
-corpora, with a slight dip in the penultimate position of each sentence.
-The Mandarin corpus entropy curve, by comparison, has a noticeably lower
-positional entropy values in utterance-final positions than in
-utterance-penultimate positions.
+the entropy curve shapes were robust across the full range of utterances
+we analyzed. We found a distinct three-step distribution for both the
+English and Spanish CHILDES corpora, with a slight dip in the
+penultimate position of each sentence. The Mandarin corpus entropy
+curve, by comparison, displayed a noticeably lower positional entropy
+values in utterance-final positions than in utterance-penultimate
+positions.
 
 These results present two important pieces of information. First, the
-Providence corpus entropy curve broadly replicates the curves found in
-both the British National Corpus and in Switchboard. Thus, the entropy
-curve of English appears not just in adult-adult conversation, but even
-in speech produced by parents to their children, and speech produced by
-very young children to their parents. This suggests that it is a highly
-robust feature of English language, as it structures the productions of
-even pre-schooled aged children.
+English Providence corpus entropy curve broadly replicates the curves
+found in both the British National Corpus and in Switchboard, our two
+previous English corpora. Thus, the entropy curve of English appears not
+just in adult-adult conversation, but even in speech produced by parents
+to their children, and speech produced by very young children to their
+parents. This suggests that it is a highly robust feature of the English
+language, as it reflects the structure of even pre-schooled aged
+children's speech.
 
@@
 from language to language. One possibility is that this variation arises
-from typological features of languages, such as characteristic word
+from typological features of language, such as characteristic word
 order. We noticed, for instance, a relatively high density of
@@
 
-\hypertarget{study-3}{%
-\section{Study 3}\label{study-3}}
+\section{Study 3}\label{study-3}
 
@@
 
-\hypertarget{method}{%
-\subsection{Method}\label{method}}
-
-\hypertarget{data-and-analysis-2}{%
-\subsubsection{Data and Analysis}\label{data-and-analysis-2}}
+\subsection{Data and Analysis}\label{data-and-analysis-2}
 
@@
 of language, we used features from the World Atlas of Language
-Structures (WALS, Dryer \& Haspelmath, 2013). WALS is The World Atlas of
-Language Structures is a curated collection of typological features
-across the world's languages. However, as the collection includes
-contributions from many authors studying a heterogeneous set of topic
-and languages, most features are missing for most languages. For our
-sample, we sought to balance diversity of languages with coverage in
-WALS. We selected \(45\) languages for which we \(8\) WALS features were
-coded (Figure \ref{fig:clust_tree}). Unfortunately, these features all
-reflected the order of words in a sentence in each langauge
+Structures (WALS, Dryer \& Haspelmath, 2013). WALS is a curated
+collection of typological features which analyze and classify the
+diversity across the world's languages. However, as the collection
+includes contributions from many authors studying a heterogeneous set of
+topics and languages, most features are missing for most languages. For
+our sample, we sought to balance diversity of languages with coverage in
+WALS. We selected \(45\) languages for which eight specific WALS
+features were coded (Figure \ref{fig:clust_tree}). Unfortunately, these
+features all reflected the order of words in a sentence in each language
 (e.g.~whether objects come before verbs, whether adjectives come before
@@
 lengths within a language to estimate a language-typical signature. We
-performed an additional analysis dividing each utterance into 6ths
-evenly and computing the slopes between these 6ths. Results from this
-analysis were qualitatively similar.
+performed an additional analysis dividing each utterance into fifths
+evenly and computing the slopes between these fifths. Results from this
+second analysis were qualitatively similar.
 
-\hypertarget{results-and-analysis}{%
-\subsection{Results and Analysis}\label{results-and-analysis}}
+\subsection{Results}\label{results-2}
 
@@
 vector similarity (Landauer \& Dumais, 1997). Intuitively, vectors with
-with a smaller angle between them are more similar. Figure
+a smaller angle between them are more similar. Figure
 \ref{fig:clust_tree} shows a hierarchical clustering of these 5
@@
 
-We then used asked whether these pairwise similarities are related to
-overlap in linguistic features. For each of the 8 features we measured,
-we fit a logistic regression predicting whether two languages would have
-the same value for it from their pairwise entropy curve similarity.
-Overlap on all 8 features was reliably predicted by entropy curve
-similarity (largest \(p = .001\)), with the amount of variance in
-feature overlap predicted varying across predictors
-\ref{tab:tab_features}. Order of object and verb (\(83A\)) was the
-strongest predictor, explaining 16.22\% of the variance in pairwise
-entropy curve similarity.
+We then asked whether these pairwise similarities were related to
+overlap in linguistic features. For each of the 8 WALS features we
+measured, we fit a logistic regression predicting whether two languages
+would have the same value for that feature from the similarity of their
+pairwise entropy curves. Overlap on all 8 features was reliably
+predicted by entropy curve similarity (largest \(p = .001\)), with the
+amount of variance in feature overlap predicted varying across
+predictors \ref{tab:tab_features}. Order of object and verb (\(83A\))
+was the strongest predictor, explaining 16.22\% of the variance in
+pairwise entropy curve similarity.
 
@@
 \end{tabular}
-\caption{Capptoin here} 
+\caption{Table showing the eight WALS features ranked by significance of effect on the entropy slope cosine distances} 
 \label{tab:tab_features}
@@
 
-\hypertarget{general-discussion}{%
-\section{General Discussion}\label{general-discussion}}
+\section{General Discussion}\label{general-discussion}
 
@@
 productions whether written, spoken, or even spoken by children. These
-curves diverge in reliabile ways from the monotonically increasing
-functions predicting by a naive Uniform Information Density hypothesis.
-Second, these characteristic curves vary across languages--some
-languages like English have increasing curves, but others like Mandarin
-are characterized by decreasing entropy of the course of an utterance.
-Finally, these characteristic curves are related to structural
-properties of those languages, for instance their word orders. This work
-presents an important rejoinder to models of linguistic productions as
-optimizing communicative success. While speakers may do their best to
-structure the information in their productions to minimize spikes in
-entropy, they operate under coding models they inherit from their native
-language, and these coding models constrain the choices they make.
+curves diverge in reliable ways from the affine functions predicted by a
+naive Uniform Information Density hypothesis. Second, these
+characteristic curves vary across languages--some languages like English
+have increasing curves, but others like Mandarin are characterized by
+decreasing entropy over the course of an utterance. Finally, these
+characteristic curves are related to structural properties of those
+languages, for instance their word orders. This work presents an
+important rejoinder to models of linguistic productions as optimizing
+communicative success. While speakers may structure the information in
+their productions to minimize spikes in entropy, speakers operate under
+coding models they inherit from their language, and these coding models
+constrain the choices they make.
 
-\hypertarget{references}{%
-\section{References}\label{references}}
+Our work complements the approach of studies such as Aylett \& Turk
+(2004), where a language's information distribution is characterized by
+a single number representing the average rate of semantic information
+transfer per syllable. Each language possesses a characteristic entropy
+curve derived from the information distribution within utterances as
+well as a measure of the average information expressed by a syllable in
+that language.
 
-\setlength{\parindent}{-0.1in} 
-\setlength{\leftskip}{0.125in}
+We expect the characteristic entropy curve for a language to have
+important downstream effects for how people process that language. While
+conversational turn-taking is universal, there exists cross-linguistic
+variation in the average length of time between conversational partners
+speaking (Stivers et al., 2009). We predict that one factor in this
+variability relates to the shape of a language's entropy curve. A jump
+at the end of a language's entropy curve, such as in English, indicates
+that there is on average high levels of information at the end of
+utterances in that language. Therefore we expect turn-taking times to be
+longer in such languages than in languages where there is a drop in the
+final segment of the entropy curve, such as Mandarin. We also expect
+that readers of a language will take longer on average to read parts of
+sentences where the entropy curve is relatively high, due to the
+presence of higher information content. This may explain the so-called
+wrap-up effect in sentence processing, in which the end of the sentence
+in English and other Germanic languages is processed more slowly during
+reading (Kuperman, Dambacher, Nuthmann, \& Kliegl, 2010; Stowe, Kaan,
+Sabourin, \& Taylor, 2018).
 
-\noindent
+\section{References}\label{references}
+
+\setlength{\parindent}{-0.1in} \setlength{\leftskip}{0.125in} \noindent
 
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-anderson1991}{}%
+\hypertarget{ref-anderson1991}{}
 Anderson, J. R. (1991). The adaptive nature of human categorization.
@@
 
-\leavevmode\hypertarget{ref-bai2008}{}%
+\hypertarget{ref-aylett2004}{}
+Aylett, M., \& Turk, A. (2004). The smooth signal redundancy hypothesis:
+A functional explanation for relationships between redundancy, prosodic
+prominence, and duration in spontaneous speech. \emph{Language and
+Speech}, \emph{47}(1), 31--56.
+
+\hypertarget{ref-bai2008}{}
 Bai, X., Yan, G., Liversedge, S. P., Zang, C., \& Rayner, K. (2008).
@@
 
-\leavevmode\hypertarget{ref-clear1993}{}%
+\hypertarget{ref-clear1993}{}
 Clear, J. H. (1993). The british national corpus. \emph{The Digital
@@
 
-\leavevmode\hypertarget{ref-demuth2006}{}%
+\hypertarget{ref-demuth2006}{}
 Demuth, K., Culbertson, J., \& Alter, J. (2006). Word-minimality,
@@
 
-\leavevmode\hypertarget{ref-dryer2013}{}%
+\hypertarget{ref-dryer2013}{}
 Dryer, M. S., \& Haspelmath, M. (2013). \emph{The world atlas of
@@
 
-\leavevmode\hypertarget{ref-ferrer-i-cancho2013}{}%
+\hypertarget{ref-ferrer-i-cancho2013}{}
 Ferrer-i-Cancho, R., Debowski, L., \& Prado Martin, F. M. del. (2013).
@@
 
-\leavevmode\hypertarget{ref-frank2008}{}%
+\hypertarget{ref-frank2008}{}
 Frank, A. F., \& Jaeger, T. F. (2008). Speaking rationally: Uniform
@@
 
-\leavevmode\hypertarget{ref-genzel2002}{}%
+\hypertarget{ref-genzel2002}{}
 Genzel, D., \& Charniak, E. (2002). Entropy rate constancy in text. In
@@
 
-\leavevmode\hypertarget{ref-godfrey1992}{}%
+\hypertarget{ref-godfrey1992}{}
 Godfrey, J. J., Holliman, E. C., \& McDaniel, J. (1992). SWITCHBOARD:
 Telephone speech corpus for research and development. In
-\emph{Acoustics, speech, and signal processing, 1992. ICASSP-92., 1992
+\emph{Acoustics, speech, and signal processing, 1992. icassp-92., 1992
 ieee international conference on} (Vol. 1, pp. 517--520). IEEE.
 
-\leavevmode\hypertarget{ref-jaeger2013}{}%
+\hypertarget{ref-jaeger2013}{}
 Jaeger, T. F., \& Snider, N. E. (2013). Alignment as a consequence of
@@
 
-\leavevmode\hypertarget{ref-jelinek1976}{}%
+\hypertarget{ref-jelinek1976}{}
 Jelinek, F. (1976). Continuous speech recognition by statistical
@@
 
-\leavevmode\hypertarget{ref-landauer1997}{}%
+\hypertarget{ref-kuperman2010}{}
+Kuperman, V., Dambacher, M., Nuthmann, A., \& Kliegl, R. (2010). The
+effect of word position on eye-movements in sentence and paragraph
+reading. \emph{The Quarterly Journal of Experimental Psychology},
+\emph{63}(9), 1838--1857.
+
+\hypertarget{ref-landauer1997}{}
 Landauer, T. K., \& Dumais, S. T. (1997). A solution to plato's problem:
@@
 
-\leavevmode\hypertarget{ref-levy2007}{}%
+\hypertarget{ref-levy2007}{}
 Levy, R. P., \& Jaeger, T. F. (2007). Speakers optimize information
@@
 
-\leavevmode\hypertarget{ref-li2015}{}%
+\hypertarget{ref-li2015}{}
 Li, H., \& Zhou, J. (2015). \emph{Study on dinner table talk of
@@
 
-\leavevmode\hypertarget{ref-macwhinney2014}{}%
+\hypertarget{ref-macwhinney2014}{}
 MacWhinney, B. (2014). \emph{The childes project: Tools for analyzing
@@
 
-\leavevmode\hypertarget{ref-sanchez2019}{}%
+\hypertarget{ref-sanchez2019}{}
 Sanchez, A., Meylan, S. C., Braginsky, M., MacDonald, K. E., Yurovsky,
@@
 interface to the child language data exchange system. \emph{Behavior
-Research Methods}.
+Research Methods}. in press.
 
-\leavevmode\hypertarget{ref-shannon1948}{}%
+\hypertarget{ref-shannon1948}{}
 Shannon, C. E. (1948). A mathematical theory of communication.
@@
 
-\leavevmode\hypertarget{ref-shiro2000}{}%
-Shiro, M. (2000). Diferencias sociales en la construcción del" yo" y
-del" otro": Expresiones evaluativas en la narrativa de niños caraqueños
-en edad escolar. In \emph{Lengua, discurso, texto: I simposio
+\hypertarget{ref-shiro2000}{}
+Shiro, M. (2000). Diferencias sociales en la construcción del`` yo'' y
+del`` otro'': Expresiones evaluativas en la narrativa de niños
+caraqueños en edad escolar. In \emph{Lengua, discurso, texto: I simposio
 internacional de análisis del discurso} (pp. 1303--1318). Visor.
 
-\leavevmode\hypertarget{ref-simons2018}{}%
+\hypertarget{ref-simons2018}{}
 Simons, G. F., \& Fenning, C. D. (Eds.). (2018). \emph{Ethnologue:
@@
 
-\leavevmode\hypertarget{ref-xu2018}{}%
+\hypertarget{ref-stivers2009universals}{}
+Stivers, T., Enfield, N. J., Brown, P., Englert, C., Hayashi, M.,
+Heinemann, T., \ldots{} others. (2009). Universals and cultural
+variation in turn-taking in conversation. \emph{Proceedings of the
+National Academy of Sciences}, pnas--0903616106.
+
+\hypertarget{ref-stowe2018}{}
+Stowe, L. A., Kaan, E., Sabourin, L., \& Taylor, R. C. (2018). The
+sentence wrap-up dogma. \emph{Cognition}, \emph{176}, 232--247.
+
+\hypertarget{ref-xu2018}{}
 Xu, Y., \& Reitter, D. (2018). Information density converges in

### 2019-02-02 cogsci/position-info.Rmd
final cogsci push

@@
 Over 7,000 languages are spoken around the modern world [@simons2018].
-These language vary along many dimensions, but all share a core goal: communicating information.
+These languages vary along many dimensions, but all share a core goal: communicating information.
 If speakers and writers of these languages act near-optimally to achieve their communicative goals, regularities of use across these diverse languages can be explained by a rational theory of communication [@anderson1991].
@@
 
-The evidence in favor of Uniform Information Density largely been situation-specific and English-language driven, while the hypothesis itself has been applied broadly over the past decade.
+The evidence in favor of Uniform Information Density has largely focused on specific  points within sentences, while the hypothesis itself has been applied broadly over the past decade.
 The original evidence in favor of Uniform Information Density in @levy2007 finds that the insertion of complementizers (e.g.
@@
 However, other recent work has contradicted the Uniform Information Density hypothesis.
-Similar to the original @levy2007, @zhan2018 focuses on information distribution at particular points in sentences.
+Similar to the original work in @levy2007, @zhan2018 focuses on information distribution at particular points in sentences.
 @zhan2018 finds that more information-rich classifiers in Mandarin Chinese are produced when production of the neighboring noun is difficult, not when the information content is high.
-@jain2018 examine word order across spoken sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on determining a Hindi speaker's word order.
+Examining the information density at the sentence level, @jain2018 study word order across written sentences in Hindi, a freer word order language than English, and find that information density has no significant effect on determining a Hindi writer's word order.
 
@@
 
-We extend and generalize @yu2016 in three ways: We confirm that this same pattern is found in English--in written language, and in conversational speech between adults and between parents and their children.
+We extend and generalize @yu2016 in three ways: We confirm that this same pattern is found throughout English--in written language, and in conversational speech between adults and between parents and their children.
 Thus, this entropy curve is a robust feature of English linguistic productions.
-We then examine entropy curves cross-linguistically, and show that characteristic curves vary across the world's languages.
+We then examine entropy curves cross-linguistically, and show that the characteristic entropy curves vary across the world's languages.
 Finally, we show that this variation is predictable in part from the structure of individual languages (i.e.
@@
 
-With this measure, we can compute the unigram entropy at each word position of sentences of each length within a corpus.
-The result of this method can be plotted for each sentence length as an *entropy curve*, which can be visually compared across sentence lengths to observe how the unigram entropy changes across absolute positions in each of the sentences.
+With this measure, we can compute the unigram positional entropy at each word position of sentences of each length within a corpus.
+The result of this method across word positions can be plotted for each sentence length as an *entropy curve*, which can be visually compared across sentence lengths to observe how the unigram entropy changes across absolute positions in each of the sentences.
 @genzel2002 similarly examine a unigram entropy measure on sentences, and found that entropy at the sentence level increases linearly with sentence index within a corpus.
@@
 We then applied the same method to the Switchboard corpus--a collection of spoken English [@godfrey1992].
-This allowed us to ask whether the characteristic function identified by @yu2016 is a general feature of English, or instead a function of written language.
+This allowed us to ask whether the characteristic function identified by @yu2016 is a general feature of English, or instead a function of their written language corpus.
 
-# Data and Analysis
+## Data and Analysis
 
@@
 The Switchboard corpus is a collection of $\sim$ 2,400 telephone conversations between unacquainted adults prompted with subjects of conversation.
-Switchboard is the corpus used in Levy and Jaeger's [-@levy2007] original demonstration of the Uniform Information Density Hypothesis.
+Switchboard is the corpus used in Levy and Jaeger's [-@levy2007] original demonstration of the Uniform Information Density hypothesis.
 
@@
 
-# Results
+## Results
 
 The entropy curves computed in both the British National Corpus and Switchboard were remarkably consistent both across corpora and across the range of sentence lengths we analyzed (Figure \ref{fig:read_and_plot_exp1}).
-Confirming Yu et al.'s [-@yu2016] findings, we find that
+Confirming Yu et al.'s [-@yu2016] findings, we found that
 the positional entropy at the beginning of sentences is low, then rises and plateaus for sentence-medial word positions before dropping slightly in the second-to-last position and rising again for the sentence-final word.
@@
 ```{r read_and_plot_exp1, fig.env = "figure*", fig.width=6, fig.height=4, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom).
-Points average entropies, error bars show 95\\% confidence intervals computed by non-parametric bootstrap."}
+Points represent average entropies, error bars show 95\\% confidence intervals computed by non-parametric bootstrap."}
 
@@
   facet_wrap(~ length, scales = "free_x") + 
-  xlab("British National Corpus utterance position") +
+  xlab("British National Corpus word position") +
   ylab("Positional entropy") +
@@
   facet_wrap(~ length, scales = "free_x") + 
-  xlab("Switchboard utterance position") +
+  xlab("Switchboard word position") +
   ylab("Positional entropy") +
@@
   facet_grid(language ~ length, scales = "free") + 
-  xlab("Utterance position") +
+  xlab("Word position") +
   ylab("Positional entropy") +
@@
 To understand how variability in entropy curves is related to the structure of language, we used features from the World Atlas of Language Structures [WALS, @dryer2013].
-WALS is a curated collection of typological features which analyze and classify the diversity across the world's languages.
+WALS is a curated collection of typological features which help to classify the diversity across the world's languages.
 However, as the collection includes contributions from many authors studying a heterogeneous set of topics and languages, most features are missing for most languages.
 For our sample, we sought to balance diversity of languages with coverage in WALS.
-We selected $45$ languages for which eight specific WALS features were coded (Figure \ref{fig:clust_tree}).
+We selected eight specific WALS features for which the same $45$ languages were coded (Figure \ref{fig:clust_tree}).
 Unfortunately, these features all reflected the order of words in a sentence in each language (e.g.
@@
 In order to understand how these curves varied across languages, we developed a method for compressing the curves into a few representative values.
-We noticed that curves tended to vary in predictable utterance positions: The majority were flat in the middle of utterances, but had either rising or falling slopes at the beginnings and ends.
-For each language, for each utterance length $k$ we computed 5 slopes: The change between the 1st and 2nd position, the change between the 2nd and 3rd position, the slope between the 3rd position and the 3rd-to-last position ($k-1$), the slope between the 3rd-to-last and the 2nd-to-last position ($k-1$), and the slope between the 2nd-to-last and final positions.
-We then averaged these slopes across utterance lengths within a language to estimate a language-typical signature.
-We performed an additional analysis dividing each utterance into fifths evenly and computing the slopes between these fifths.
+We noticed that curves tended to vary at predictable positions within sentences: The majority were flat in the middle, but had either rising or falling slopes at the beginnings and ends.
+For each language, for each sentence length $k$ we computed 5 slopes: The change between the 1st and 2nd position, the change between the 2nd and 3rd position, the slope between the 3rd position and the 3rd-to-last position ($k-1$), the slope between the 3rd-to-last and the 2nd-to-last position ($k-1$), and the slope between the 2nd-to-last and final positions.
+We then averaged these slopes across sentence lengths within a language to estimate a language-typical signature.
+We performed an additional analysis dividing each sentence into fifths evenly and computing the slopes between these fifths.
 Results from this second analysis were qualitatively similar.
@@
 
-We treated these 5 slopes as a vector, allowing us to compute the pairwise similarity between languages with cosine--a standard measure of vector similarity [@landauer1997].
+We treated these five slopes as a vector, allowing us to compute the pairwise similarity between languages with cosine--a standard measure of vector similarity [@landauer1997].
 Intuitively, vectors with a smaller angle between them are more similar.
-Figure \ref{fig:clust_tree} shows a hierarchical clustering of these 5 dimensional vectors, demonstrating that entropy curve similarity appears to capture typological relatedly, e.g.
-Slavic languages are similar to each-other.
+Figure \ref{fig:clust_tree} shows a hierarchical clustering of these 5 dimensional vectors, demonstrating that entropy curve similarity appears to capture typological relatedly (e.g.
+Slavic languages are similar to each other).
 
@@
 We then asked whether these pairwise similarities were related to overlap in linguistic features.
-For each of the 8 WALS features we measured, we fit a logistic regression predicting whether two languages would have the same value for that feature from the similarity of their pairwise entropy curves.
-Overlap on all 8 features was reliably predicted by entropy curve similarity (largest $p = .001$), with the amount of variance in feature overlap predicted varying across predictors \ref{tab:tab_features}.
+For each of the eight WALS features we measured, we fit a logistic regression predicting whether two languages would have the same value for that feature from the pairwise similarity of their entropy curves.
+Overlap on all eight features was reliably predicted by entropy curve similarity (largest $p = .001$), with the amount of variance in feature overlap predicted varying across predictors \ref{tab:tab_features}.
 Order of object and verb ($83A$) was the strongest predictor, explaining `r max_r2`% of the variance in pairwise entropy curve similarity.
@@
 We also expect that readers of a language will take longer on average to read parts of sentences where the entropy curve is relatively high, due to the presence of higher information content.
-This may explain the so-called wrap-up effect in sentence processing, in which the end of the sentence in English and other Germanic languages is processed more slowly during reading [@stowe2018;@kuperman2010].
+This may explain the so-called wrap-up effect in sentence processing, in which the ends of sentences are processed more slowly during reading [@stowe2018;@kuperman2010].
+The entropy curve analysis can be carried out on any language across a wide variety of text corpora, making it a powerful tool for analyzing the information structure of both spoken and written utterances in a language.
 

### 2019-02-02 cogsci/position-info.tex
final cogsci push

@@
 Over 7,000 languages are spoken around the modern world (Simons \&
-Fenning, 2018). These language vary along many dimensions, but all share
-a core goal: communicating information. If speakers and writers of these
-languages act near-optimally to achieve their communicative goals,
+Fenning, 2018). These languages vary along many dimensions, but all
+share a core goal: communicating information. If speakers and writers of
+these languages act near-optimally to achieve their communicative goals,
 regularities of use across these diverse languages can be explained by a
@@
 
-The evidence in favor of Uniform Information Density largely been
-situation-specific and English-language driven, while the hypothesis
-itself has been applied broadly over the past decade. The original
-evidence in favor of Uniform Information Density in R. P. Levy \& Jaeger
-(2007) finds that the insertion of complementizers (e.g. ``that'') in
-relative clauses in English corresponds to where neighboring words have
-high information content. Similarly, A. F. Frank \& Jaeger (2008) argues
-that contractions in English such as ``you're'' do not occur when
-neighboring words are highly informative. Applications of Uniform
-Information Density include determining whether linguistic alignment
-takes place (Jaeger \& Snider, 2013), Zipfian word length distributions
-(Piantadosi, Tily, \& Gibson, 2011), communication efficiency (Mahowald,
-Fedorenko, Piantadosi, \& Gibson, 2013), dialogue and turn-taking (Xu \&
-Reitter, 2018) and the significance of ambiguity in language
-(Piantadosi, Tily, \& Gibson, 2012), among other research.
+The evidence in favor of Uniform Information Density has largely focused
+on specific points within sentences, while the hypothesis itself has
+been applied broadly over the past decade. The original evidence in
+favor of Uniform Information Density in R. P. Levy \& Jaeger (2007)
+finds that the insertion of complementizers (e.g. ``that'') in relative
+clauses in English corresponds to where neighboring words have high
+information content. Similarly, A. F. Frank \& Jaeger (2008) argues that
+contractions in English such as ``you're'' do not occur when neighboring
+words are highly informative. Applications of Uniform Information
+Density include determining whether linguistic alignment takes place
+(Jaeger \& Snider, 2013), Zipfian word length distributions (Piantadosi,
+Tily, \& Gibson, 2011), communication efficiency (Mahowald, Fedorenko,
+Piantadosi, \& Gibson, 2013), dialogue and turn-taking (Xu \& Reitter,
+2018) and the significance of ambiguity in language (Piantadosi, Tily,
+\& Gibson, 2012), among other research.
 
 However, other recent work has contradicted the Uniform Information
-Density hypothesis. Similar to the original R. P. Levy \& Jaeger (2007),
-Zhan \& Levy (2018) focuses on information distribution at particular
-points in sentences. Zhan \& Levy (2018) finds that more
+Density hypothesis. Similar to the original work in R. P. Levy \& Jaeger
+(2007), Zhan \& Levy (2018) focuses on information distribution at
+particular points in sentences. Zhan \& Levy (2018) finds that more
 information-rich classifiers in Mandarin Chinese are produced when
 production of the neighboring noun is difficult, not when the
-information content is high. Jain, Singh, Ranjan, Rajkumar, \& Agarwal
-(2018) examine word order across spoken sentences in Hindi, a freer word
-order language than English, and find that information density has no
-significant effect on determining a Hindi speaker's word order.
+information content is high. Examining the information density at the
+sentence level, Jain, Singh, Ranjan, Rajkumar, \& Agarwal (2018) study
+word order across written sentences in Hindi, a freer word order
+language than English, and find that information density has no
+significant effect on determining a Hindi writer's word order.
 
@@
 We extend and generalize Yu et al. (2016) in three ways: We confirm that
-this same pattern is found in English--in written language, and in
-conversational speech between adults and between parents and their
+this same pattern is found throughout English--in written language, and
+in conversational speech between adults and between parents and their
 children. Thus, this entropy curve is a robust feature of English
 linguistic productions. We then examine entropy curves
-cross-linguistically, and show that characteristic curves vary across
-the world's languages. Finally, we show that this variation is
-predictable in part from the structure of individual languages
+cross-linguistically, and show that the characteristic entropy curves
+vary across the world's languages. Finally, we show that this variation
+is predictable in part from the structure of individual languages
 (i.e.~word order). Taken together, our results suggest a refinement of
@@
 
-With this measure, we can compute the unigram entropy at each word
-position of sentences of each length within a corpus. The result of this
-method can be plotted for each sentence length as an \emph{entropy
-curve}, which can be visually compared across sentence lengths to
-observe how the unigram entropy changes across absolute positions in
-each of the sentences. Genzel \& Charniak (2002) similarly examine a
-unigram entropy measure on sentences, and found that entropy at the
-sentence level increases linearly with sentence index within a corpus.
-Uniform Information Density applies this uniformity of entropy rate in
-sentences to all levels of speech (R. P. Levy \& Jaeger, 2007), and so
-our method obtained from Yu et al. (2016), which examines text at the
-word level, should find a monotonically increasing affine function at
-the word level.
+With this measure, we can compute the unigram positional entropy at each
+word position of sentences of each length within a corpus. The result of
+this method across word positions can be plotted for each sentence
+length as an \emph{entropy curve}, which can be visually compared across
+sentence lengths to observe how the unigram entropy changes across
+absolute positions in each of the sentences. Genzel \& Charniak (2002)
+similarly examine a unigram entropy measure on sentences, and found that
+entropy at the sentence level increases linearly with sentence index
+within a corpus. Uniform Information Density applies this uniformity of
+entropy rate in sentences to all levels of speech (R. P. Levy \& Jaeger,
+2007), and so our method obtained from Yu et al. (2016), which examines
+text at the word level, should find a monotonically increasing affine
+function at the word level.
 
@@
 function identified by Yu et al. (2016) is a general feature of English,
-or instead a function of written language.
+or instead a function of their written language corpus.
 
-\section{Data and Analysis}\label{data-and-analysis}
+\subsection{Data and Analysis}\label{data-and-analysis}
 
@@
 conversation. Switchboard is the corpus used in Levy and Jaeger's (2007)
-original demonstration of the Uniform Information Density Hypothesis.
+original demonstration of the Uniform Information Density hypothesis.
 
@@
 
-\section{Results}\label{results}
+\subsection{Results}\label{results}
 
@@
 \ref{fig:read_and_plot_exp1}). Confirming Yu et al.'s (2016) findings,
-we find that the positional entropy at the beginning of sentences is
+we found that the positional entropy at the beginning of sentences is
 low, then rises and plateaus for sentence-medial word positions before
@@
 \caption[Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom)]{Representative Entropy Curves for the British National Corpus (top) and Switchboard (bottom).
-Points average entropies, error bars show 95\% confidence intervals computed by non-parametric bootstrap.}\label{fig:read_and_plot_exp1}
+Points represent average entropies, error bars show 95\% confidence intervals computed by non-parametric bootstrap.}\label{fig:read_and_plot_exp1}
 \end{figure*}
@@
 Structures (WALS, Dryer \& Haspelmath, 2013). WALS is a curated
-collection of typological features which analyze and classify the
-diversity across the world's languages. However, as the collection
-includes contributions from many authors studying a heterogeneous set of
-topics and languages, most features are missing for most languages. For
-our sample, we sought to balance diversity of languages with coverage in
-WALS. We selected \(45\) languages for which eight specific WALS
-features were coded (Figure \ref{fig:clust_tree}). Unfortunately, these
+collection of typological features which help to classify the diversity
+across the world's languages. However, as the collection includes
+contributions from many authors studying a heterogeneous set of topics
+and languages, most features are missing for most languages. For our
+sample, we sought to balance diversity of languages with coverage in
+WALS. We selected eight specific WALS features for which the same \(45\)
+languages were coded (Figure \ref{fig:clust_tree}). Unfortunately, these
 features all reflected the order of words in a sentence in each language
@@
 developed a method for compressing the curves into a few representative
-values. We noticed that curves tended to vary in predictable utterance
-positions: The majority were flat in the middle of utterances, but had
-either rising or falling slopes at the beginnings and ends. For each
-language, for each utterance length \(k\) we computed 5 slopes: The
-change between the 1st and 2nd position, the change between the 2nd and
-3rd position, the slope between the 3rd position and the 3rd-to-last
-position (\(k-1\)), the slope between the 3rd-to-last and the
-2nd-to-last position (\(k-1\)), and the slope between the 2nd-to-last
-and final positions. We then averaged these slopes across utterance
-lengths within a language to estimate a language-typical signature. We
-performed an additional analysis dividing each utterance into fifths
-evenly and computing the slopes between these fifths. Results from this
-second analysis were qualitatively similar.
+values. We noticed that curves tended to vary at predictable positions
+within sentences: The majority were flat in the middle, but had either
+rising or falling slopes at the beginnings and ends. For each language,
+for each sentence length \(k\) we computed 5 slopes: The change between
+the 1st and 2nd position, the change between the 2nd and 3rd position,
+the slope between the 3rd position and the 3rd-to-last position
+(\(k-1\)), the slope between the 3rd-to-last and the 2nd-to-last
+position (\(k-1\)), and the slope between the 2nd-to-last and final
+positions. We then averaged these slopes across sentence lengths within
+a language to estimate a language-typical signature. We performed an
+additional analysis dividing each sentence into fifths evenly and
+computing the slopes between these fifths. Results from this second
+analysis were qualitatively similar.
 
@@
 
-We treated these 5 slopes as a vector, allowing us to compute the
+We treated these five slopes as a vector, allowing us to compute the
 pairwise similarity between languages with cosine--a standard measure of
@@
 dimensional vectors, demonstrating that entropy curve similarity appears
-to capture typological relatedly, e.g.~Slavic languages are similar to
-each-other.
+to capture typological relatedly (e.g.~Slavic languages are similar to
+each other).
 
 We then asked whether these pairwise similarities were related to
-overlap in linguistic features. For each of the 8 WALS features we
+overlap in linguistic features. For each of the eight WALS features we
 measured, we fit a logistic regression predicting whether two languages
-would have the same value for that feature from the similarity of their
-pairwise entropy curves. Overlap on all 8 features was reliably
+would have the same value for that feature from the pairwise similarity
+of their entropy curves. Overlap on all eight features was reliably
 predicted by entropy curve similarity (largest \(p = .001\)), with the
@@
 presence of higher information content. This may explain the so-called
-wrap-up effect in sentence processing, in which the end of the sentence
-in English and other Germanic languages is processed more slowly during
-reading (Kuperman, Dambacher, Nuthmann, \& Kliegl, 2010; Stowe, Kaan,
-Sabourin, \& Taylor, 2018).
+wrap-up effect in sentence processing, in which the ends of sentences
+are processed more slowly during reading (Kuperman, Dambacher, Nuthmann,
+\& Kliegl, 2010; Stowe, Kaan, Sabourin, \& Taylor, 2018). The entropy
+curve analysis can be carried out on any language across a wide variety
+of text corpora, making it a powerful tool for analyzing the information
+structure of both spoken and written utterances in a language.
 

### 2019-02-08 ngrams/prelim_ngrams.Rmd
prelim bigrams and trigrams on prov and switchboard

---
title: "Eye-tracking"
author: "Joe Klafka"
date: "2/8/2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(tidyboot)
library(entropy)
library(magrittr)
library(feather)
library(here)

library(childesr)

knitr::opts_chunk$set(echo = TRUE)
```

```{r corpora}
prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss

switchboard <- read_feather(here("../switchboard/switchboard.feather")) %>%
  rename(words = value)
```

```{r ngram entropy}
# takes in a tbl of utterances the column with the actual utterances is 
# named words, along with an integer target_length and a ngram length N.
ngram_entropy <- function(utterances, target_length, N) {

 tokens <- utterances %>%
  mutate(length = str_count(words, pattern = "[ +]+") + 1) %>%
  filter(length == target_length) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(bigrams, words, token = "ngrams", n = N) %>%
  group_by(utterance_id) %>%
  mutate(gram_order = 1:n()) %>%
  group_by(gram_order, bigrams) %>%
  filter(gram_order <= target_length - (N-1))

 tokens %>%
  summarise(n = n()) %>%
  tidyboot(summary_function = function(x) x %>% 
        summarise(entropy = entropy(n, unit = "log2")),
       statistics_functions = function(x) x %>%
       summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
  mutate(length = target_length)
}
```

```{r providence}
prov_utterances %<>% 
  filter(speaker_role == "Target_Child") %>%
  mutate(words = stem)

# bigrams
prov_bigrams <- map(3:10, ~ngram_entropy(prov_utterances, .x, 2)) %>% 
  bind_rows()

prov_bigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

# trigrams
prov_trigrams <- map(4:10, ~ngram_entropy(prov_utterances, .x, 3)) %>% 
  bind_rows()

prov_trigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

```

```{r switchboard}
# unigrams to check against previous results
switchboard_unigrams <- map(2:10, ~ngram_entropy(switchboard, .x, 1)) %>% 
  bind_rows()

switchboard_unigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

# bigrams
switchboard_bigrams <- map(3:10, ~ngram_entropy(switchboard, .x, 2)) %>% 
  bind_rows()

switchboard_bigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

#trigrams
switchboard_trigrams <- map(4:10, ~ngram_entropy(switchboard, .x, 3)) %>% 
  bind_rows()

switchboard_trigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)
```

### 2019-02-11 ngrams/prelim_ngrams.Rmd
more childes bigrams and trigrams

@@
 ```{r corpora}
-prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
+prov_utterances <- get_utterances(corpus = "Providence") %>% 
+  filter(speaker_role == "Target_Child") %>%
+  mutate(words = stem)
+
+okayama_utterances <- get_utterances(corpus = "Okayama")  %>% 
+  filter(speaker_role == "Target_Child") %>%
+  rename(words = gloss)
+
+zhou_utterances <- get_utterances(corpus = "ZhouDinner") %>%
+  filter(speaker_role == "Target_Child") %>%
+  rename(words = stem)
 
@@
 ```{r ngram entropy}
+# unnest_tokens(word, value, token = stringr::str_split, pattern = " +")
+
 # takes in a tbl of utterances the column with the actual utterances is 
@@
   mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(bigrams, words, token = "ngrams", n = N) %>%
+  unnest_tokens(ngrams, words, token = "ngrams", n = N) %>%
   group_by(utterance_id) %>%
   mutate(gram_order = 1:n()) %>%
-  group_by(gram_order, bigrams) %>%
+  group_by(gram_order, ngrams) %>%
   filter(gram_order <= target_length - (N-1))
@@
 
-```{r providence}
-prov_utterances %<>% 
-  filter(speaker_role == "Target_Child") %>%
-  mutate(words = stem)
+```{r childes}
+## Providence
+# unigrams
+prov_unigrams <- map(2:10, ~ngram_entropy(prov_utterances, .x, 1)) %>% 
+  bind_rows()
 
+prov_unigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F)
+
+# bigrams
+prov_bigrams <- map(3:10, ~ngram_entropy(prov_utterances, .x, 2)) %>% 
+  bind_rows()
+
+prov_bigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F)
+
+# trigrams
+prov_trigrams <- map(4:10, ~ngram_entropy(prov_utterances, .x, 3)) %>% 
+  bind_rows()
+
+prov_trigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F)
+
+## Zhou
+# unigrams
+zhou_unigrams <- map(2:10, ~ngram_entropy(zhou_utterances, .x, 1)) %>% 
+  bind_rows()
+
+zhou_unigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F)
+
+# bigrams
+zhou_bigrams <- map(3:10, ~ngram_entropy(zhou_utterances, .x, 2)) %>% 
+  bind_rows()
+
+zhou_bigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F)
+
+# trigrams
+zhou_trigrams <- map(4:10, ~ngram_entropy(zhou_utterances, .x, 3)) %>% 
+  bind_rows()
+
+zhou_trigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F)
+
+## Okayama
+# bigrams 
+okayama_bigrams <- map(3:10, ~ngram_entropy(okayama_utterances, .x, 2)) %>% 
+  bind_rows()
+
+okayama_bigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F)
+
+# trigrams
+okayama_trigrams <- map(4:10, ~ngram_entropy(okayama_utterances, .x, 3)) %>% 
+  bind_rows()
+
+okayama_trigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F)
+```
+
+```{r zhou and okayama}
 # bigrams

### 2019-02-11 ngrams/prelim_ngrams.Rmd
ngram computation

@@
 prov_utterances <- get_utterances(corpus = "Providence") %>% 
-  filter(speaker_role == "Target_Child") %>%
+  filter(speaker_role != "Target_Child") %>%
   mutate(words = stem)
@@
 zhou_utterances <- get_utterances(corpus = "ZhouDinner") %>%
-  filter(speaker_role == "Target_Child") %>%
+  filter(speaker_role != "Target_Child") %>%
   rename(words = stem)
@@
     mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(ngrams, words, token = "ngrams", n = N) %>%
+    unnest_tokens(word, words, token = stringr::str_split, pattern = " +") %>%
     group_by(utterance_id) %>%
-  mutate(gram_order = 1:n()) %>%
-  group_by(gram_order, ngrams) %>%
-  filter(gram_order <= target_length - (N-1))
+    mutate(word_order = 1:n())
 
- tokens %>%
+  lags <- expand.grid(lag_n = 1:(N-1), 
+                      utterance_id = unique(tokens$utterance_id))
+
+  grams <- tokens %>%
+    left_join(lags, by = "utterance_id") %>%
+    group_by(utterance_id, lag_n) %>%
+    arrange(utterance_id, lag_n, word_order) %>%
+    mutate(lag = lag(word, n = first(lag_n))) %>%
+    group_by(utterance_id, word_order, word) %>%
+    summarise(gram = paste(lag, collapse = " ")) %>%
+    summarise(gram = paste(gram, word, collapse = " ")) %>%
+    filter(word_order >= N) %>%
+    rename(gram_order = word_order) %>%
+    group_by(gram_order, gram)
+
+ grams %>%
   summarise(n = n()) %>%

### 2019-02-15 ngrams/prelim_ngrams.Rmd
feathers for childes corpora ngrams

@@
   rename(words = stem)
+
+shiro_utterances <- get_utterances(corpus = "Shiro") #use gloss
+
+wagner_utterances <- get_utterances(corpus = "Wagner") #use gloss
 
@@
 
+write_feather(feather(prov_unigrams), here("prov/adult_unigrams.feather"))
 ## Zhou

### 2019-02-15 ngrams/prelim_ngrams.Rmd
cleaned up ngrams rmd

--- 2019-02-11 ngrams/prelim_ngrams.Rmd
@@
 prov_utterances <- get_utterances(corpus = "Providence") %>% 
-  filter(speaker_role != "Target_Child") %>%
+  filter(speaker_role == "Target_Child") %>%
   mutate(words = stem)
@@
 okayama_utterances <- get_utterances(corpus = "Okayama")  %>% 
-  filter(speaker_role == "Target_Child") %>%
+  filter(speaker_role != "Target_Child") %>%
   rename(words = gloss)
@@
 
+shiro_utterances <- get_utterances(corpus = "Shiro") %>% 
+  filter(speaker_role == "Target_Child") %>%
+  rename(words = gloss)
+
+wagner_utterances <- get_utterances(corpus = "Wagner") %>% 
+  filter(speaker_role == "Target_Child") %>%
+  rename(words = gloss)
+
 switchboard <- read_feather(here("../switchboard/switchboard.feather")) %>%
@@
 ```{r ngram entropy}
-# unnest_tokens(word, value, token = stringr::str_split, pattern = " +")
-
 # takes in a tbl of utterances the column with the actual utterances is 
@@
 ## Providence
-# unigrams
+# unigrams to check against previous results
 prov_unigrams <- map(2:10, ~ngram_entropy(prov_utterances, .x, 1)) %>% 
@@
 
-## Zhou
-# unigrams
-zhou_unigrams <- map(2:10, ~ngram_entropy(zhou_utterances, .x, 1)) %>% 
-  bind_rows()
-
-zhou_unigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  geom_pointrange() +
-  geom_smooth(se = F)
-
-# bigrams
-zhou_bigrams <- map(3:10, ~ngram_entropy(zhou_utterances, .x, 2)) %>% 
-  bind_rows()
-
-zhou_bigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  geom_pointrange() +
-  geom_smooth(se = F)
-
-# trigrams
-zhou_trigrams <- map(4:10, ~ngram_entropy(zhou_utterances, .x, 3)) %>% 
-  bind_rows()
-
-zhou_trigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  geom_pointrange() +
-  geom_smooth(se = F)
-
-## Okayama
-# bigrams 
-okayama_bigrams <- map(3:10, ~ngram_entropy(okayama_utterances, .x, 2)) %>% 
-  bind_rows()
-
-okayama_bigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  geom_pointrange() +
-  geom_smooth(se = F)
-
-# trigrams
-okayama_trigrams <- map(4:10, ~ngram_entropy(okayama_utterances, .x, 3)) %>% 
-  bind_rows()
-
-okayama_trigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  geom_pointrange() +
-  geom_smooth(se = F)
-```
-
-```{r zhou and okayama}
-# bigrams
-prov_bigrams <- map(3:10, ~ngram_entropy(prov_utterances, .x, 2)) %>% 
-  bind_rows()
-
-prov_bigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  geom_pointrange() +
-  geom_smooth(se = F)
-
-# trigrams
-prov_trigrams <- map(4:10, ~ngram_entropy(prov_utterances, .x, 3)) %>% 
-  bind_rows()
-
-prov_trigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  geom_pointrange() +
-  geom_smooth(se = F)
-
+write_feather(prov_unigrams, "prov/child_unigrams.feather")
 ```

### 2019-02-15 wiki_entropy/wiki_clustering.Rmd
imputation work and organizing

@@
 
+```{r word order}
+wiki_gs <- gs_title("Wikipedia_relative5")
+wkpd <- gs_read(wiki_gs) 
+LANGUAGES <- wkpd$Language
+
+FEATURES <- c(82:97)
+for (i in c(1:length(FEATURES))) {
+  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
+} #get all word order WALS features in correct syntax
+
+wf <- wals.feature(FEATURES)
+
+wo_df <- wf %>% 
+  filter(language %in% LANGUAGES) %>%
+  select(-wals.code, -latitude, -longitude, -glottocode, -`84A`)
+
+wo_df <- wo_df[rowSums(is.na(new_wf)) < 6,] %>%
+  as_data_frame() %>%
+  group_by(language) %>%
+  slice(1) %>%
+  gather(feature, value, `82A`:`97A`) %>%
+  group_by(feature) %>%
+  mutate(value = factor(value))
+
+imputed_wo <- wo_df %>%
+  select(language, feature, value) %>%
+  ungroup() %>% 
+  spread(feature, value) %>% 
+  column_to_rownames("language") %>%
+  sapply(as.factor) %>%
+  as_data_frame() %>%
+  MIMCA(ncp = 3, maxiter = 999)
+
+imputed_wo <- as.data.frame(imputed_result[[1]][[99]])
+
+## convert this to work with imputed_wo
+wf_many_langs <- wf_df %>%
+  group_by(feature) %>%
+  summarise(n = sum(!is.na(value))) %>%
+  filter(n > 60)
+
+wf_sub_langs <- wf_df %>%
+  filter(feature %in% wf_many_langs$feature) %>%
+  group_by(language) %>%
+  summarise(keep = sum(is.na(value)) == 0) %>%
+  filter(keep)
+
+wf_sub <- wf_df %>%
+  filter(feature %in% wf_many_langs$feature,
+         language %in% wf_sub_langs$language) %>%
+  select(language, feature, value) %>%
+  mutate(value = as.numeric(as.factor(value)))
+
+wf_pairwise <- expand.grid(language1 = wf_sub$language,
+                           language2 = wf_sub$language) %>%
+  left_join(wf_sub, by = c("language1" = "language")) %>%
+  rename(value1 = value) %>%
+  left_join(wf_sub, by = c("language2" = "language", "feature")) %>%
+  rename(value2 = value) %>%
+  mutate(same = value1 == value2) %>%
+  select(-value1, -value2) %>%
+  left_join(cosines, by = c("language1", "language2"))
+
+models <- wf_pairwise %>%
+  group_by(feature) %>%
+  nest() %>%
+  mutate(model = map(data, ~glm(same ~ cosine, 
+                                family = "binomial", data = .)))
+
+models %>%
+  mutate(coeffs = map(model, tidy),
+         rs = map(model, )) %>%
+  select(-data, -model) %>%
+  unnest() %>%
+  filter(term == "cosine") %>%
+  arrange(desc(statistic))
+```
+
 ```{r}
@@
 FEATURES <- c(1:144)
-for (i in c(1:144)) {
+for (i in c(1:length(FEATURES))) {
   FEATURES[i] <- paste(FEATURES[i], "A", sep = "")

### 2019-02-19 wiki_entropy/wiki_clustering.Rmd
word order glm

--- 2019-02-02 wiki_entropy/wiki_clustering.Rmd
@@
 library(broom)
+library(here)
+library(feather)
 
@@
 
+```{r word order}
+wiki_gs <- gs_title("Wikipedia_relative5")
+wkpd <- gs_read(wiki_gs) 
+LANGUAGES <- wkpd$Language
+
+FEATURES <- c(82:97)
+for (i in c(1:length(FEATURES))) {
+  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
+} #get all word order WALS features in correct syntax
+
+wf <- wals.feature(FEATURES)
+
+wo_wf <- wf %>% 
+  filter(language %in% LANGUAGES) %>%
+  select(-wals.code, -latitude, -longitude, -glottocode, -`84A`)
+
+wo_df <- wo_wf[rowSums(is.na(wo_wf)) < 6,] %>%
+  as_data_frame() %>%
+  group_by(language) %>%
+  slice(1) %>%
+  gather(feature, value, `82A`:`97A`) %>%
+  group_by(feature) %>%
+  mutate(value = factor(value))
+
+imputed_wo <- wo_df %>%
+  select(language, feature, value) %>%
+  ungroup() %>% 
+  spread(feature, value) %>% 
+  column_to_rownames("language") %>%
+  sapply(as.factor) %>%
+  MIMCA(ncp = 1)
+
+imputed_wo <- as.data.frame(imputed_wo[[1]][[99]]) 
+imputed_wo$language <- unique(wo_df$language)
+wo_sub <- imputed_wo %>% group_by(language) %>%
+  slice(1) %>%
+  gather(feature, value, `82A`:`97A`) %>%
+  group_by(feature) %>%
+  mutate(value = as.numeric(as.factor(value)))
+
+wf_pairwise <- expand.grid(language1 = wo_sub$language,
+                           language2 = wo_sub$language) %>%
+  left_join(wo_sub, by = c("language1" = "language")) %>%
+  rename(value1 = value) %>%
+  left_join(wo_sub, by = c("language2" = "language", "feature")) %>%
+  rename(value2 = value) %>%
+  mutate(same = value1 == value2) %>%
+  select(-value1, -value2) %>%
+  left_join(cosines, by = c("language1", "language2"))
+
+models <- wf_pairwise %>%
+  group_by(feature) %>%
+  nest() %>%
+  mutate(model = map(data, ~glm(same ~ cosine, 
+                                family = "binomial", data = .)))
+
+models %>%
+  mutate(coeffs = map(model, tidy)) %>%
+  select(-data, -model) %>%
+  unnest() %>%
+  filter(term == "cosine") %>%
+  arrange(desc(statistic)) %>%
+  write_feather(here("word_order_glm.feather"))
+```
+
 ```{r}
@@
 
-```{r}
+```{r cogsci glm subset}
 
@@
 ```
-
-```{r imputation for glm}
-dendro_langs <- wf_sub_langs$language
-
-wk_clust_tree <- wkpd %>% 
-              filter(Language %in% dendro_langs) %>%
-              column_to_rownames(var = "Language") %>%
-              dist() %>%
-              hclust("ward.D2")
-
-wk_clust_tree %>% plot(hang = -1)
-
-imputed_result <- wf_df %>%
-  select(language, feature, value) %>%
-  ungroup() %>% 
-  spread(feature, value) %>% 
-  column_to_rownames("language") %>%
-  sapply(as.factor) %>%
-  as_data_frame() %>%
-  MIMCA(ncp = 7, maxiter = 999)
-
-imputed_wf <- as.data.frame(imputed_result[[1]][[99]])
-```

### 2019-02-22 wiki_entropy/wiki_clustering.Rmd
wikipedia word order absolute feather

@@
 ```{r word order}
-wiki_gs <- gs_title("Wikipedia_relative5")
+wiki_gs <- gs_title("Wikipedia_absolute")
 wkpd <- gs_read(wiki_gs) 
@@
   arrange(desc(statistic)) %>%
-  write_feather(here("word_order_glm.feather"))
+  write_feather(here("word_order_glm_absolute.feather"))
 ```

### 2019-03-01 wiki_entropy/wiki_clustering.Rmd
starting on WALS imputation & MCA

@@
 library(feather)
+library(FactoMineR)
 
@@
 theme_set(theme_classic(base_size = 16))
+```
+
+```{r imputation and MCA}
+wiki_gs <- gs_title("Wikipedia_absolute")
+wkpd <- gs_read(wiki_gs) 
+LANGUAGES <- wkpd$Language
+
+FEATURES <- c(1:144)
+for (i in c(1:length(FEATURES))) {
+  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
+} #get all WALS features in correct syntax
+
+wf <- wals.feature(FEATURES)
+
+# Phonology `1A`:`19A`
+# Morphology `20A`:`29A`
+# Nominal Categories `30A`:`57A`
+# Nominal Syntax `58A`:64A`
+# Verbal syntax `65A`:`80A`
+# Word order `81A`:`97A`, `143A`, `144A`
+# Simple clauses `98A`:`120A`
+# Complex sentences `121A`:`128A`
+# Lexicon `129A`:`138A`
+# Sign languages `139A`:`140A`
+
+phon_wf <- wf %>% 
+  filter(language %in% LANGUAGES) %>%
+  select(-wals.code, -latitude, -longitude, -glottocode) %>%
+  select(`1A`:`19A`, language)
+
+phon_wf <- phon_wf[rowSums(is.na(phon_wf)) < 15,]
 ```

### 2019-03-04 wiki_entropy/wiki_clustering.Rmd
working on imputation

@@
 library(FactoMineR)
+library(feather)
+library(janitor)
 
@@
 wf <- wals.feature(FEATURES) 
+fair_wf <- wf[rowSums(!is.na(wf)) > 50,] %>%
+  filter(language %in% LANGUAGES) %>%
+  select(-wals.code, -latitude, -longitude, -glottocode) %>%
+  as_data_frame()
 
@@
 
-phon_wf <- wf %>% 
+wa <- fair_wf %>% 
   filter(language %in% LANGUAGES) %>%
-  select(-wals.code, -latitude, -longitude, -glottocode) %>%
-  select(`1A`:`19A`, language)
+  select(`1A`:`19A`, language) %>%
+  group_by(language) %>%
+  slice(1) %>%
+  gather(feature, value, `1A`:`19A`) %>%
+  group_by(feature) %>%
+  mutate(value = factor(value)) %>%
+  ungroup() %>% 
+  spread(feature, value) %>% 
+  column_to_rownames("language") %>%
+  as.data.frame(stringsAsFactors = TRUE) %>%
+  clean_names() %>% 
+  lapply(factor) %>%
+  lapply(as.numeric) %>%
+  lapply(factor) %>%
+  as.data.frame() %>% 
+  MIMCA(ncp = 4, maxiter = 100000, threshold = 1e-1)
 
-phon_wf <- phon_wf[rowSums(is.na(phon_wf)) < 15,]
+wa <- as.data.frame(wa[[1]][[99]]) 
+wa$language <- unique(phon_wf$language)
 ```

### 2019-03-04 wiki_entropy/wiki_clustering.Rmd
issues with apply and dim(X) in MIMCA for non-word order WALS imputation

@@
   column_to_rownames("language") %>%
-  as.data.frame(stringsAsFactors = TRUE) %>%
-  clean_names() %>% 
-  lapply(factor) %>%
-  lapply(as.numeric) %>%
-  lapply(factor) %>%
-  as.data.frame() %>% 
-  MIMCA(ncp = 4, maxiter = 100000, threshold = 1e-1)
+  sapply(as.factor) %>%
+  as_data_frame() %>%
+  MIMCA(ncp = 1, maxiter = 100000, threshold = 1e-1)
 
@@
 wa$language <- unique(phon_wf$language)
-```
-
-```{r hierarchical clustering}
-NUM_CLUSTERS <- 10
-
-##read in googlesheet
-wiki_gs <- gs_title("Wikipedia_absolute")
-wkpd <- gs_read(wiki_gs) 
-LANGUAGES <- wkpd$Language
-
-wkpd <- wkpd %>% 
-  as.data.frame() %>%
-  column_to_rownames(var = "Language")
-
-##do clustering
-wk_clust_tree <- wkpd %>% 
-              dist() %>%
-              hclust("ward.D2")
-
-wk_clust_tree %>% plot(hang = -1)
-wk_clusters <- wk_clust_tree %>%
-              cutree(k = NUM_CLUSTERS)
-
-wk_clusters %>% 
-  t() %>%
-  as_data_frame() %>% 
-  t() %>%
-  as_data_frame(rownames = "language") %>%
-  rename(cluster = V1) %>%
-  arrange(cluster) %>% 
-  View()
 ```
@@
   filter(language %in% LANGUAGES) %>%
-  select(-wals.code, -latitude, -longitude, -glottocode, -`84A`)
+  select(language, `82A`:`97A`) %>%
+  select(-`84A`)
 
@@
   write_feather(here("word_order_glm_absolute.feather"))
+```
+
+```{r hierarchical clustering}
+NUM_CLUSTERS <- 10
+
+##read in googlesheet
+wiki_gs <- gs_title("Wikipedia_absolute")
+wkpd <- gs_read(wiki_gs) 
+LANGUAGES <- wkpd$Language
+
+wkpd <- wkpd %>% 
+  as.data.frame() %>%
+  column_to_rownames(var = "Language")
+
+##do clustering
+wk_clust_tree <- wkpd %>% 
+              dist() %>%
+              hclust("ward.D2")
+
+wk_clust_tree %>% plot(hang = -1)
+wk_clusters <- wk_clust_tree %>%
+              cutree(k = NUM_CLUSTERS)
+
+wk_clusters %>% 
+  t() %>%
+  as_data_frame() %>% 
+  t() %>%
+  as_data_frame(rownames = "language") %>%
+  rename(cluster = V1) %>%
+  arrange(cluster) %>% 
+  View()
 ```

### 2019-03-04 wiki_entropy/wiki_clustering.Rmd
clustering issues

@@
 wf <- wals.feature(FEATURES)
+
 fair_wf <- wf[rowSums(!is.na(wf)) > 50,] %>%
@@
   select(-wals.code, -latitude, -longitude, -glottocode) %>%
-  as_data_frame()
+  as_tibble() %>%
+  clean_names()
 
@@
   filter(language %in% LANGUAGES) %>%
-  select(`1A`:`19A`, language) %>%
+  select(x1a:x19a, language) %>%
   group_by(language) %>%
-  slice(1) %>%
-  gather(feature, value, `1A`:`19A`) %>%
-  group_by(feature) %>%
-  mutate(value = factor(value)) %>%
-  ungroup() %>% 
-  spread(feature, value) %>% 
-  column_to_rownames("language") %>%
-  sapply(as.factor) %>%
-  as_data_frame() %>%
-  MIMCA(ncp = 1, maxiter = 100000, threshold = 1e-1)
+  slice(1)
 
-wa <- as.data.frame(wa[[1]][[99]]) 
-wa$language <- unique(phon_wf$language)
+wa_langs <- wa %>%
+  gather(measure, value, -language) %>%
+  group_by(language) %>%
+  summarise(value = mean(!is.na(value))) %>%
+  arrange(desc(value)) %>%
+  filter(value > .6) %>% 
+  pull(language)
+
+wa_sub <- wa %>%
+  filter(language %in% wa_langs) %>%
+  column_to_rownames("language")
+ # mutate_each(~as.numeric(as.factor(.x)))
+
+wa_out <- wa_sub %>%
+  select(-x14a:-x17a, -x9a,-x10a) %>%
+  slice(1:5) %>%
+  MIMCA(ncp = 1)
+
+wa_imputed <- wa_out[["res.MI"]] %>%
+  map(~mutate(.x, language = wa_langs)) %>%
+  bind_rows(.id = "boot") %>%
+  group_by(language, boot) %>%
+  gather(measure, value, -language, -boot) %>%
+  group_by(language, measure) %>%
+  summarise(value = mean(value, na.rm = T)) %>%
+  spread(measure, value)
 ```

### 2019-03-14 wiki_entropy/wiki_clustering.Rmd
mean imputation for testing

@@
 
+mean_wf <- fair_wf
+
+mean_impute <- function(col) {
+  counts <- col %>%
+    remove_missing() %>%
+    unlist() %>%
+    table() %>%
+    sort(decreasing = T) 
+  col[is.na(col),] <- labels(counts[1])
+  return(col)
+}
+
+for (i in 1:ncol(mean_wf)) {
+  mean_wf[, i] <- mean_impute(mean_wf[, i])
+}
+
 # Phonology `1A`:`19A`
@@
 wa <- fair_wf %>% 
-  filter(language %in% LANGUAGES) %>%
-  select(x1a:x19a, language) %>%
+  select(x1a:x10a, language) %>%
   group_by(language) %>%
@@
   filter(language %in% wa_langs) %>%
-  column_to_rownames("language")
+  column_to_rownames("language") %>%
+  ungroup()
  # mutate_each(~as.numeric(as.factor(.x)))
@@
 wa_out <- wa_sub %>%
-  select(-x14a:-x17a, -x9a,-x10a) %>%
-  slice(1:5) %>%
   MIMCA(ncp = 1)

### 2019-03-21 ngrams/wiki_ngrams.Rmd
starting wikipedia ngrams process

---
title: "Wikipedia ngrams Entropy Analysis"
author: "Josef Klafka and Dan Yurovsky"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(googlesheets)
library(here)
library(feather)

knitr::opts_chunk$set(echo = TRUE)
```

```{r getting gsheets}
wiki_gs <- gs_title("Wikipedia_relative5")
w1 <- gs_read(wiki_gs) %>% arrange(Language)
wiki_gs <- gs_title("Wikipedia_relative_2grams")
w2 <- gs_read(wiki_gs) %>% arrange(Language)

w1 %>% 
  filter(Language %in% w2$Language)

```

### 2019-03-22 ngrams/wiki_relative_ngrams.R
relative ngrams

@@ -1,0 +1,109 @@
+library(tidyverse)
+library(directlabels)
+library(dplyr)
+library(tidytext)
+library(entropy)
+library(tidyboot)
+library(tokenizers)
+library(googlesheets)
+library(stringr)
+### Begin Functions
+
+extract_wiki_tokens <- function(min_length, max_length, utterances, N) {
+
+tokens <- utterances %>%
+  mutate(length = str_count(words, pattern = "[ +]+") + 1) %>%
+  filter(length >= min_length) %>%
+  filter(length <= max_length) %>%
+  mutate(utterance_id = 1:n()) %>%
+  unnest_tokens(word, words, token = stringr::str_split, pattern = " +") %>%
+  group_by(utterance_id) %>%
+  mutate(word_order = 1:n())
+
+
+lags <- expand.grid(lag_n = 1:(N-1),
+                    utterance_id = unique(tokens$utterance_id))
+
+grams <- tokens %>%
+  left_join(lags, by = "utterance_id") %>%
+  arrange(utterance_id, lag_n, word_order) %>%
+  mutate(lag = lag(word, n = first(lag_n))) %>%
+  drop_na() %>%
+  mutate(gram = str_c(lag, word, sep = " ")) %>%
+  rename(gram_order = word_order) %>%
+  group_by(length, gram_order, gram) %>%
+  summarise(n = n())
+
+}
+
+get_quantiles <- function(df, num_sections) {
+
+  quantile(df$gram_order, probs = seq(0, 1, 1/num_sections)) %>%
+    round() %>%
+    as_data_frame() %>%
+    rename(gram_order = value)
+}
+
+relative_slopes <- function(df, pos_list) {
+
+  get_slope <- function(x) {
+
+    selected_pos <- pos_list %>%
+      group_by(length) %>%
+      slice(x:(x+1)) %>%
+      group_by(length) %>%
+      summarise(min = min(gram_order),
+                max = max(gram_order)) %>%
+      split(.$length) %>%
+      map(~seq(.x$min, .x$max, 1) %>% as_data_frame) %>%
+      bind_rows(.id = "length") %>%
+      rename(gram_order = value) %>%
+      mutate(length = as.numeric(length))
+
+    df %>%
+      inner_join(selected_pos, by = c("length", "gram_order")) %>%
[…]

### 2019-03-22 ngrams/wiki_relative_ngrams.R
fixing issues with rel relative ngrams

@@ -10,30 +10,30 @@ library(stringr)
 ### Begin Functions
 
 extract_wiki_tokens <- function(min_length, max_length, utterances, N) {
+  
+  tokens <- utterances %>%
+    mutate(length = str_count(words, pattern = "[ +]+") + 1) %>%
+    filter(length >= min_length) %>%
+    filter(length <= max_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, words, token = stringr::str_split, pattern = "[ +]+") %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n())
+  
+  
+  lags <- expand.grid(lag_n = 1:(N-1),
+                      utterance_id = unique(tokens$utterance_id))
 
-tokens <- utterances %>%
-  mutate(length = str_count(words, pattern = "[ +]+") + 1) %>%
-  filter(length >= min_length) %>%
-  filter(length <= max_length) %>%
-  mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, words, token = stringr::str_split, pattern = " +") %>%
-  group_by(utterance_id) %>%
-  mutate(word_order = 1:n())
-
-
-lags <- expand.grid(lag_n = 1:(N-1),
-                    utterance_id = unique(tokens$utterance_id))
-
-grams <- tokens %>%
-  left_join(lags, by = "utterance_id") %>%
-  arrange(utterance_id, lag_n, word_order) %>%
-  mutate(lag = lag(word, n = first(lag_n))) %>%
-  drop_na() %>%
-  mutate(gram = str_c(lag, word, sep = " ")) %>%
-  rename(gram_order = word_order) %>%
-  group_by(length, gram_order, gram) %>%
-  summarise(n = n())
-
+  tokens %>%
+    left_join(lags, by = "utterance_id") %>%
+    arrange(utterance_id, lag_n, word_order) %>%
+    mutate(lag = lag(word, n = first(lag_n))) %>%
+    drop_na() %>%
+    mutate(gram = str_c(lag, word, sep = " ")) %>%
+    rename(gram_order = word_order) %>%
+    group_by(length, gram_order, gram) %>%
+    summarise(n = n())
+  
 }
 
 get_quantiles <- function(df, num_sections) {
[…]

### 2019-04-04 ngrams/wiki_ngrams.Rmd
initial 1-4grams for wikipedia relative slopes

@@
 library(feather)
+library(magrittr)
+library(reshape2)
 
@@
 ```{r getting gsheets}
-wiki_gs <- gs_title("Wikipedia_relative5")
-w1 <- gs_read(wiki_gs) %>% arrange(Language)
-wiki_gs <- gs_title("Wikipedia_relative_2grams")
-w2 <- gs_read(wiki_gs) %>% arrange(Language)
+# wiki_gs <- gs_title("Wikipedia_relative5")
+# w1 <- gs_read(wiki_gs) %>% arrange(Language)
+# wiki_gs <- gs_title("Wikipedia_relative_2grams")
+# w2 <- gs_read(wiki_gs) %>% arrange(Language)
+# wiki_gs <- gs_title("Wikipedia_relative_3grams")
+# w3 <- gs_read(wiki_gs) %>% arrange(Language)
+# wiki_gs <- gs_title("Wikipedia_relative_4grams")
+# w4 <- gs_read(wiki_gs) %>% arrange(Language)
+
+w1 <- read_feather(here("relative_unigrams.feather"))
+w2 <- read_feather(here("relative_bigrams.feather"))
+w3 <- read_feather(here("relative_trigrams.feather"))
+w4 <- read_feather(here("relative_quadgrams.feather"))
+
+get_agg <- function(ws) {
+ ws %>%
+  mutate(agg = abs(Slope1) + abs(Slope2) + abs(Slope3) + abs(Slope4) + abs(Slope5)) %>%
+  select(Language, agg)
+}
+
+w1 %<>% 
+  get_agg() %>%
+  rename(unigram = agg)
+w2 %<>%
+  get_agg() %>%
+  rename(bigram = agg)
+w3 %<>%
+  get_agg() %>%
+  rename(trigram = agg)
+w4 %<>%
+  get_agg() %>%
+  rename(quadgram = agg)
 
 w1 %>% 
-  filter(Language %in% w2$Language)
+  left_join(w2) %>%
+  left_join(w3) %>%
+  left_join(w4) %>%
+  remove_missing() %>%
+  slice(1:5) %>%
+  melt(id.vars = "Language") %>%
+  ggplot(aes(variable, value, color=Language)) + 
+  geom_point() +
+  geom_line()
 

### 2019-04-15 ngrams/wiki_ngrams.Rmd
first attempt wikipedia ngrams lm

--- 2019-03-21 ngrams/wiki_ngrams.Rmd
@@
 library(feather)
+library(magrittr)
+library(reshape2)
+library(lme4)
 
@@
 ```{r getting gsheets}
-wiki_gs <- gs_title("Wikipedia_relative5")
-w1 <- gs_read(wiki_gs) %>% arrange(Language)
-wiki_gs <- gs_title("Wikipedia_relative_2grams")
-w2 <- gs_read(wiki_gs) %>% arrange(Language)
+w1 <- read_feather(here("relative_unigrams.feather"))
+w2 <- read_feather(here("relative_bigrams.feather"))
+w3 <- read_feather(here("relative_trigrams.feather"))
+w4 <- read_feather(here("relative_quadgrams.feather"))
+
+get_agg <- function(ws) {
+ ws %>%
+  mutate(agg = abs(Slope1) + abs(Slope2) + abs(Slope3) + abs(Slope4) + abs(Slope5)) %>%
+  select(Language, agg)
+}
+
+w1 %<>% 
+  get_agg() %>%
+  rename(unigram = agg)
+w2 %<>%
+  get_agg() %>%
+  rename(bigram = agg)
+w3 %<>%
+  get_agg() %>%
+  rename(trigram = agg)
+w4 %<>%
+  get_agg() %>%
+  rename(quadgram = agg)
+
+wn <- w1 %>% 
+  left_join(w2) %>%
+  left_join(w3) %>%
+  left_join(w4) %>%
+  remove_missing() %>%
+  melt(id.vars = "Language") %>% 
+  rename(gram = variable, s_i = value)
 
 w1 %>% 
-  filter(Language %in% w2$Language)
+  left_join(w2) %>%
+  left_join(w3) %>%
+  left_join(w4) %>%
+  remove_missing() %>%
+  slice(1:5) %>%
+  melt(id.vars = "Language") %>%
+  ggplot(aes(variable, value, color=Language)) + 
+  geom_point() +
+  geom_line()
+```
+
+```{r Joe's regression}
+# log(sum(abs(s_i))) = language * type * gram + (source | type)
+
+# let's just try log(sum(abs(s_i))) = language + gram since type/source are all the same atm
+## why would there be an interaction between language and gram?
+
+simple_glm <- glm(data = wn, formula = s_i ~ gram + Language)
 

### 2019-04-15 ngrams/wiki_absolute_ngrams.R

@@ -1,0 +1,182 @@
+library(tidyverse)
+library(directlabels)
+library(tidytext)
+library(entropy)
+library(tidyboot)
+library(tokenizers)
+library(googlesheets)
+
+### BEGIN FUNCTIONS
+
+extract_wiki_tokens <- function(min_length, max_length, utterances, N) {
+
+  tokens <- utterances %>%
+    mutate(length = str_count(words, pattern = "[ +]+") + 1) %>%
+    filter(length >= min_length) %>%
+    filter(length <= max_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(gram, words, token = stringr::str_split, pattern = "[ +]+") %>%
+    group_by(utterance_id) %>%
+    mutate(gram_order = 1:n())
+
+
+  lags <- expand.grid(lag_n = 1:(N-1),
+                      utterance_id = unique(tokens$utterance_id))
+
+  tokens %>%
+    left_join(lags, by = "utterance_id") %>%
+    arrange(utterance_id, lag_n, gram_order) %>%
+    mutate(lag = lag(gram, n = first(lag_n))) %>%
+    drop_na() %>%
+    mutate(gram = str_c(lag, gram, sep = " ")) %>%
+    rename(gram_order = gram_order) %>%
+    group_by(length, gram_order, gram) %>%
+    summarise(n = n())
+
+}
+
+start_slope <- function(df) {
+
+ df %>%
+    filter(gram_order %in% c(2,3)) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, gram_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(gram_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `3` - `2`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+second_slope <- function(df) {
+
+ df %>%
+    ungroup() %>%
+    mutate(gram_order %in% c(3, 4)) %>%
+    filter(gram_order > 0) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, gram_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(gram_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `4` - `3`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
[…]

### 2019-04-15 wiki_entropy/get_wiki_df.py

@@ -1,12 +1,11 @@
+import regex as re
 import numpy as np
-import subprocess
-import regex as re 
-import pandas as pd 
+import pandas as pd
 import csv
 import sys
 import json
+import subprocess
 import urllib.request
-import os
 
 
 with open("language_dict.json", 'r') as f:
@@ -17,8 +16,8 @@ LANG_DICT = json.loads(file) #maps languages to their wikipedia url prefixes
 def extract_texts(upper_bound):
 	'''
 	Given an upper_bound on file numbers, extracts the raw text from the files
-	in the text/AA folder. That folder is given as output by the 
-	cirrus_extract program. 
+	in the text/AA folder. That folder is given as output by the
+	cirrus_extract program.
 	'''
 	rs = ""
 	for i in range(upper_bound):
@@ -34,11 +33,11 @@ def extract_texts(upper_bound):
 
 def get_sen_dict(text):
 	'''
-	Given the output of extract_texts as a string, constructs a 
+	Given the output of extract_texts as a string, constructs a
 	dictionary mapping each sentence in the string to its length in words.
 	'''
-	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:&]|\d+|\(|\)|\[|\]", \
-		"", text) 
+	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:–'&*#/—“»]|\d+|\(|\)|\[|\]", \
+		"", text)
 	text = re.sub('"', '', text)
 	#just the words from the articles
 	text = text.replace('\n', " ")
@@ -54,33 +53,32 @@ def get_sen_dict(text):
 def get_sen_df(text):
 	'''
 	Given the output of extract_texts as a string, returns a Pandas dataframe
-	of the output of get_sen_dict. 
+	of the output of get_sen_dict.
 	'''
 	sd = get_sen_dict(text)
 	s = pd.Series(sd, name = "length")
 	del sd
-	s.index.name = "gloss"
-	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence 
+	s.index.name = "words"
+	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence
 					 	  #and the sentence length respectively
 	return df
-	
+
 
 def main(lang_name):
 	'''
-	Given the English name of a language, constructs a dataframe with the 
-	sentences of a random subset of articles in that language from Wikipedia, 
[…]

### 2019-04-15 wiki_entropy/wiki_absolute.R

@@ -1,0 +1,170 @@
+library(knitr)
+library(tidyverse)
+library(directlabels)
+library(dplyr)
+library(tidytext)
+library(entropy)
+library(tidyboot)
+library(tokenizers)
+library(googlesheets)
+
+### BEGIN FUNCTIONS
+
+extract_wiki_tokens <- function(min_length, max_length, utterances) {
+
+  tokens <- utterances %>%
+    mutate(gloss = str_trim(gloss)) %>%
+    mutate(length = str_count(gloss, pattern = " +") + 1) %>%
+    filter(length >= min_length) %>%
+    filter(length <= max_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(length, word_order, word) %>%
+    summarise(n = n())
+}
+
+start_slope <- function(df) {
+
+ df %>%
+    filter(word_order %in% c(1,2)) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `2` - `1`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+second_slope <- function(df) {
+
+ df %>%
+    ungroup() %>%
+    mutate(word_order %in% c(2, 3)) %>%
+    filter(word_order > 0) %>%
+    sample_frac(1, replace = T) %>%
+    group_by(length, word_order) %>%
+    summarise(entropy = entropy(n, unit = "log2")) %>%
+    spread(word_order, entropy) %>%
+    ungroup() %>%
+    mutate(diff = `3` - `2`) %>%
+    summarise(diff = mean(diff)) %>%
+    pull(diff)
+
+}
+
+#note this measurement only makes sense if utterance length is at least 6
+mid_slope <- function(df) {
+
+ df %>%
+    ungroup() %>%
+    mutate(word_order = if_else(word_order == length-2, 2,
+                                if_else(word_order == 3, 1,
+                                        0))) %>%
+    filter(word_order > 0) %>%
+    group_by(length, word_order) %>%
+    sample_frac(1, replace = T) %>%
[…]

### 2019-04-15 wiki_entropy/wiki_relative.R

@@ -1,46 +1,35 @@
+library(knitr)
 library(tidyverse)
 library(directlabels)
+library(dplyr)
 library(tidytext)
 library(entropy)
 library(tidyboot)
 library(tokenizers)
 library(googlesheets)
-library(stringr)
+
 ### Begin Functions
 
-extract_wiki_tokens <- function(min_length, max_length, utterances, N) {
+extract_wiki_tokens <- function(min_length, max_length, utterances) {
 
   tokens <- utterances %>%
-    mutate(length = str_count(words, pattern = "[ +]+") + 1) %>%
+    mutate(gloss = str_trim(gloss)) %>%
+    mutate(length = str_count(gloss, pattern = " +") + 1) %>%
     filter(length >= min_length) %>%
     filter(length <= max_length) %>%
     mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, words, token = stringr::str_split, pattern = "[ +]+") %>%
+    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
     group_by(utterance_id) %>%
-    mutate(word_order = 1:n())
-
-
-  lags <- expand.grid(lag_n = 1:(N-1),
-                      utterance_id = unique(tokens$utterance_id))
-
-  tokens %>%
-    left_join(lags, by = "utterance_id") %>%
-    arrange(utterance_id, lag_n, word_order) %>%
-    mutate(lag = lag(word, n = first(lag_n))) %>%
-    drop_na() %>%
-    mutate(gram = str_c(lag, word, sep = " ")) %>%
-    rename(gram_order = word_order) %>%
-    group_by(length, gram_order, gram) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(length, word_order, word) %>%
     summarise(n = n())
-
 }
 
 get_quantiles <- function(df, num_sections) {
-
-  quantile(df$gram_order, probs = seq(0, 1, 1/num_sections)) %>%
+  quantile(df$word_order, probs = seq(0, 1, 1/num_sections)) %>%
     round() %>%
     as_data_frame() %>%
-    rename(gram_order = value)
+    rename(word_order = value)
 }
 
 relative_slopes <- function(df, pos_list) {
@@ -51,21 +40,21 @@ relative_slopes <- function(df, pos_list) {
       group_by(length) %>%
[…]

### 2019-04-15 wiki_entropy/wiki_script.sh

@@ -1,0 +1,42 @@
+#!/bin/bash
+
+declare -a arr=("Basque" "Dutch" "Alemannic" "Wolof" "Bosnian" "Hindi" "Danish"
+	"Russian" "Urdu" "German" "Swedish" "Afrikaans" "Anglo-Saxon" "Icelandic"
+	"Belarusian" "Bulgarian" "Croatian" "Macedonian" "Serbo-Croatian" "Ukrainian"
+	"Vietnamese" "Pangasinan" "Kapampangan" "Khmer" "Japanese" "Javanese"
+	"Indonesian" "Polish" "Malay" "Sundanese" "Hakka" "Turkish" "Kazakh"
+	"Kirghiz" "Turkmen" "Hebrew" "Persian" "Arabic" "Tajik" "Latin" "French"
+	"Galician" "Romanian" "Norman" "Corsican" "Maori" "Nepali" "Mongolian"
+	"Estonian" "Finnish" "Hungarian" "Bengali" "Waray" "Welsh" "Armenian"
+	"Korean" "Greek" "Catalan" "Walloon" "Yiddish" "Maltese" "Romansh"
+	"Banyumasan" "Võro" "Samogitian" "Chavacano")
+
+declare -a newarr=("Basque" "Dutch" "Alemannic" "Wolof" "Bosnian" "Hindi"
+	"Danish" "Russian" "Urdu" "German" "Swedish" "Afrikaans" "Anglo-Saxon"
+	"Icelandic" "Belarusian" "Bulgarian" "Croatian" "Macedonian" "Serbo-Croatian"
+	"Ukrainian" "Vietnamese" "Pangasinan" "Kapampangan" "Khmer" "Japanese"
+	"Javanese" "Indonesian" "Polish" "Malay" "Sundanese" "Hakka" "Turkish"
+	"Kazakh" "Kirghiz" "Turkmen" "Hebrew" "Persian" "Arabic" "Tajik" "Latin"
+	"French" "Galician" "Romanian" "Norman" "Corsican" "Maori" "Nepali" "Mongolian"
+	"Estonian" "Finnish" "Hungarian" "Bengali" "Waray" "Welsh" "Armenian" "Korean"
+	"Greek" "Catalan" "Walloon" "Yiddish" "Maltese" "Romansh" "Bavarian" "Faroese"
[…]

### 2019-04-16 ngrams/ngrams.Rmd
started on CHILDES ngrams

---
title: "Eye-tracking"
author: "Joe Klafka"
date: "2/8/2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(tidyboot)
library(entropy)
library(magrittr)
library(feather)
library(here)

library(childesr)

knitr::opts_chunk$set(echo = TRUE)
```

```{r corpora}
prov_utterances <- get_utterances(corpus = "Providence") %>% 
  filter(speaker_role == "Target_Child") %>%
  mutate(words = stem)

okayama_utterances <- get_utterances(corpus = "Okayama")  %>% 
  filter(speaker_role != "Target_Child") %>%
  rename(words = gloss)

zhou_utterances <- get_utterances(corpus = "ZhouDinner") %>%
  filter(speaker_role != "Target_Child") %>%
  rename(words = stem)

shiro_utterances <- get_utterances(corpus = "Shiro") %>% 
  filter(speaker_role == "Target_Child") %>%
  rename(words = gloss)

wagner_utterances <- get_utterances(corpus = "Wagner") %>% 
  filter(speaker_role == "Target_Child") %>%
  rename(words = gloss)

switchboard <- read_feather(here("../switchboard/switchboard.feather")) %>%
  rename(words = value)
```

```{r relative ngram entropy}
extract_wiki_tokens <- function(min_length, max_length, utterances, N) {

  tokens <- utterances %>%
    mutate(length = str_count(words, pattern = "[ +]+") + 1) %>%
    filter(length >= min_length) %>%
    filter(length <= max_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, words, token = stringr::str_split, pattern = "[ +]+") %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n())

  lags <- expand.grid(lag_n = 1:(N-1),
                      utterance_id = unique(tokens$utterance_id))

  tokens %>%
    left_join(lags, by = "utterance_id") %>%
    arrange(utterance_id, lag_n, word_order) %>%
    mutate(lag = lag(word, n = first(lag_n))) %>%
    drop_na() %>%
    mutate(gram = str_c(lag, word, sep = " ")) %>%
    rename(gram_order = word_order) %>%
    group_by(length, gram_order, gram) %>%
    summarise(n = n())

}

get_quantiles <- function(df, num_sections) {

  quantile(df$gram_order, probs = seq(0, 1, 1/num_sections)) %>%
    round() %>%
    as_data_frame() %>%
    rename(gram_order = value)
}

relative_slopes <- function(df, pos_list) {

  get_slope <- function(x) {

    selected_pos <- pos_list %>%
      group_by(length) %>%
      slice(x:(x+1)) %>%
      group_by(length) %>%
      summarise(min = min(gram_order),
                max = max(gram_order)) %>%
      split(.$length) %>%
      map(~seq(.x$min, .x$max, 1) %>% as_data_frame) %>%
      bind_rows(.id = "length") %>%
      rename(gram_order = value) %>%
      mutate(length = as.numeric(length))

    df_tokens %>%
      inner_join(selected_pos, by = c("length", "gram_order")) %>%
      group_by(length, gram_order) %>%
      summarise(entropy = entropy(n, unit = "log2")) %>%
      lm(entropy ~ length + I(length^2) + gram_order, data  = .) %>%
      tidy() %>%
      filter(term == "gram_order")
  }

  divides <- pos_list %>%
    group_by(length) %>%
    summarise(n = n()) %>%
    summarise(n = mean(n)) %>%
    pull()

  map(1:((divides)-1), get_slope) %>%
    bind_rows(.id = "cut")

}

### End Functions

NUM_SECTIONS = 5

LANGUAGE <- "Danish"
NGRAMS <- 2

df_tokens <- get_utterances(corpus = "Plunkett") %>% 
  filter(speaker_role != "Target_Child") %>%
  rename(words = gloss) %>%
  extract_wiki_tokens(min_length = NGRAMS+5, max_length = 10, N = NGRAMS)

pos_list <- df_tokens %>%
  group_by(length) %>%
  distinct(gram_order) %>%
  split(.$length) %>%
  map(~get_quantiles(.x, NUM_SECTIONS)) %>%
  bind_rows(.id = "length") %>%
  mutate(length = as.numeric(length))

slopes <- relative_slopes(df_tokens, pos_list)

```

```{r absolute ngram entropy}
extract_wiki_tokens <- function(min_length, max_length, utterances, N) {

  tokens <- utterances %>%
    mutate(length = str_count(words, pattern = "[ +]+") + 1) %>%
    filter(length >= min_length) %>%
    filter(length <= max_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(gram, words, token = stringr::str_split, pattern = "[ +]+") %>%
    group_by(utterance_id) %>%
    mutate(gram_order = 1:n())

  lags <- expand.grid(lag_n = 1:(N-1),
                      utterance_id = unique(tokens$utterance_id))

  tokens %>%
    left_join(lags, by = "utterance_id") %>%
    arrange(utterance_id, lag_n, gram_order) %>%
    mutate(lag = lag(gram, n = first(lag_n))) %>%
    drop_na() %>%
    mutate(gram = str_c(lag, gram, sep = " ")) %>%
    rename(gram_order = gram_order) %>%
    group_by(length, gram_order, gram) %>%
    summarise(n = n())

}

start_slope <- function(df) {

 df %>%
    filter(gram_order %in% c(2,3)) %>%
    sample_frac(1, replace = T) %>%
    group_by(length, gram_order) %>%
    summarise(entropy = entropy(n, unit = "log2")) %>%
    spread(gram_order, entropy) %>%
    ungroup() %>%
    mutate(diff = `3` - `2`) %>%
    summarise(diff = mean(diff)) %>%
    pull(diff)

}

second_slope <- function(df) {

 df %>%
    ungroup() %>%
    mutate(gram_order %in% c(3, 4)) %>%
    filter(gram_order > 0) %>%
    sample_frac(1, replace = T) %>%
    group_by(length, gram_order) %>%
    summarise(entropy = entropy(n, unit = "log2")) %>%
    spread(gram_order, entropy) %>%
    ungroup() %>%
    mutate(diff = `4` - `3`) %>%
    summarise(diff = mean(diff)) %>%
    pull(diff)

}

#note this measurement only makes sense if utterance length is at least 6
mid_slope <- function(df) {

 df %>%
    ungroup() %>%
    mutate(gram_order = if_else(gram_order == length-2, 3,
                                if_else(gram_order == 3, 2,
                                        0))) %>%
    filter(gram_order > 0) %>%
    group_by(length, gram_order) %>%
    sample_frac(1, replace = T) %>%
    group_by(length, gram_order) %>%
    summarise(entropy = entropy(n, unit = "log2")) %>%
    spread(gram_order, entropy) %>%
    ungroup() %>%
    mutate(diff = `3` - `2`) %>%
    summarise(diff = mean(diff)) %>%
    pull(diff)

}

penult_slope <- function(df) {

 df %>%
    ungroup() %>%
    mutate(gram_order = if_else(gram_order == length-1, 3,
                                if_else(gram_order == length -2, 2,
                                        0))) %>%
    filter(gram_order > 0) %>%
    group_by(length, gram_order) %>%
    sample_frac(1, replace = T) %>%
    group_by(length, gram_order) %>%
    summarise(entropy = entropy(n, unit = "log2")) %>%
    spread(gram_order, entropy) %>%
    ungroup() %>%
    mutate(diff = `3` - `2`) %>%
    summarise(diff = mean(diff)) %>%
    pull(diff)

}

end_slope <- function(df) {

 df %>%
    ungroup() %>%
    mutate(gram_order = if_else(gram_order == length, 3,
                                if_else(gram_order == length -1 , 2,
                                        0))) %>%
    filter(gram_order > 0) %>%
    group_by(length, gram_order) %>%
    sample_frac(1, replace = T) %>%
    group_by(length, gram_order) %>%
    summarise(entropy = entropy(n, unit = "log2")) %>%
    spread(gram_order, entropy) %>%
    ungroup() %>%
    mutate(diff = `3` - `2`) %>%
    summarise(diff = mean(diff)) %>%
    pull(diff)

}

get_wiki_slopes <- function(tokens, REPS) {

  start_slopes <- replicate(REPS, start_slope(tokens)) %>%
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975),
              ci_lower = quantile(value, .025))

  second_slopes <- replicate(REPS, second_slope(tokens)) %>%
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975),
              ci_lower = quantile(value, .025))

  mid_slopes <- replicate(REPS, mid_slope(tokens)) %>%
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975),
              ci_lower = quantile(value, .025))

  penult_slopes <- replicate(REPS, penult_slope(tokens)) %>%
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975),
              ci_lower = quantile(value, .025))

  end_slopes <- replicate(REPS, end_slope(tokens)) %>%
    as_data_frame() %>%
    summarise(mean = mean(value), ci_upper = quantile(value, .975),
              ci_lower = quantile(value, .025))

  rbind(start_slopes, second_slopes, mid_slopes, penult_slopes, end_slopes)
}

### END FUNCTIONS

LANGUAGE <- 
NGRAMS <- 
df <-

REPS <- 100

tokens <- extract_wiki_tokens(NGRAMS + 5, 50, df, NGRAMS)
slopes <- get_wiki_slopes(tokens, REPS)
```

```{r basic ngram entropy}
# takes in a tbl of utterances the column with the actual utterances is 
# named words, along with an integer target_length and a ngram length N.
ngram_entropy <- function(utterances, target_length, N) {

  tokens <- utterances %>%
    mutate(length = str_count(words, pattern = "[ +]+") + 1) %>%
    filter(length == target_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, words, token = stringr::str_split, pattern = " +") %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n())

  lags <- expand.grid(lag_n = 1:(N-1), 
                      utterance_id = unique(tokens$utterance_id))

  grams <- tokens %>%
    left_join(lags, by = "utterance_id") %>%
    group_by(utterance_id, lag_n) %>%
    arrange(utterance_id, lag_n, word_order) %>%
    mutate(lag = lag(word, n = first(lag_n))) %>%
    group_by(utterance_id, word_order, word) %>%
    summarise(gram = paste(lag, collapse = " ")) %>%
    summarise(gram = paste(gram, word, collapse = " ")) %>%
    filter(word_order >= N) %>%
    rename(gram_order = word_order) %>%
    group_by(gram_order, gram)

 grams %>%
  summarise(n = n()) %>%
  tidyboot(summary_function = function(x) x %>% 
        summarise(entropy = entropy(n, unit = "log2")),
       statistics_functions = function(x) x %>%
       summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
  mutate(length = target_length)
}
```

```{r childes}
utterances <- get_utterances(corpus = "Plunkett") %>% 
  filter(speaker_role != "Target_Child") %>%
  rename(words = gloss)

# unigrams to check against previous results
unigrams <- map(2:10, ~ngram_entropy(utterances, .x, 1)) %>% 
  bind_rows()

# bigrams
bigrams <- map(3:10, ~ngram_entropy(utterances, .x, 2)) %>% 
  bind_rows()

# trigrams
trigrams <- map(4:10, ~ngram_entropy(utterances, .x, 3)) %>% 
  bind_rows()

# quadgrams
quadgrams <- map(5:10, ~ngram_entropy(utterances, .x, 4)) %>%
  bind_rows()

unigrams %>%
  bind_rows(bigrams, trigrams, quadgrams) %>%
  mutate(Language = "Danish") %>%
  write_feather("CHILDES/danish_adult.feather")
```

```{r switchboard}
# unigrams to check against previous results
switchboard_unigrams <- map(2:10, ~ngram_entropy(switchboard, .x, 1)) %>% 
  bind_rows()

switchboard_unigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

# bigrams
switchboard_bigrams <- map(3:10, ~ngram_entropy(switchboard, .x, 2)) %>% 
  bind_rows()

switchboard_bigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)

#trigrams
switchboard_trigrams <- map(4:10, ~ngram_entropy(switchboard, .x, 3)) %>% 
  bind_rows()

switchboard_trigrams %>% ggplot(aes(x = gram_order, y = empirical_entropy,
                      ymin = ci_lower, ymax = ci_upper)) +
  facet_wrap(~ length) + 
  geom_pointrange() +
  geom_smooth(se = F)
```

### 2019-04-18 ngrams/wiki_ngrams.Rmd
reorganized wikipedia ngrams data

@@
 library(tidytext)
-library(googlesheets)
 library(here)
@@
 
-```{r getting gsheets}
-w1 <- read_feather(here("relative_unigrams.feather"))
-w2 <- read_feather(here("relative_bigrams.feather"))
-w3 <- read_feather(here("relative_trigrams.feather"))
-w4 <- read_feather(here("relative_quadgrams.feather"))
+```{r set up data}
+w1 %>% 
+  left_join(w2) %>%
+  left_join(w3) %>%
+  left_join(w4) %>%
+  remove_missing() %>%
+  slice(1:5) %>%
+  melt(id.vars = "Language") %>%
+  ggplot(aes(variable, value, color=Language)) + 
+  geom_point() +
+  geom_line()
+```
+
+```{r Joe's regression}
+# log(sum(abs(s_i))) = language * type * gram + (source | type)
+
+# let's just try log(sum(abs(s_i))) = language + gram since type/source are all the same atm
+## why would there be an interaction between language and gram?
+wn <- read_feather(here("Data/wiki_ngrams.feather"))
+
+simple_glm <- glm(data = wn, formula = s_i ~ gram + Language)
+
+```
+
+```{r get data}
+w1 <- read_feather(here("Data/relative_unigrams.feather"))
+w2 <- read_feather(here("Data/relative_bigrams.feather"))
+w3 <- read_feather(here("Data/relative_trigrams.feather"))
+w4 <- read_feather(here("Data/relative_quadgrams.feather"))
 
@@
 
-w1 %>% 
-  left_join(w2) %>%
-  left_join(w3) %>%
-  left_join(w4) %>%
-  remove_missing() %>%
-  slice(1:5) %>%
-  melt(id.vars = "Language") %>%
-  ggplot(aes(variable, value, color=Language)) + 
-  geom_point() +
-  geom_line()
+wn %>% write_feather(here("Data/wiki_ngrams.feather"))
 ```
-
-```{r Joe's regression}
-# log(sum(abs(s_i))) = language * type * gram + (source | type)
-
-# let's just try log(sum(abs(s_i))) = language + gram since type/source are all the same atm
-## why would there be an interaction between language and gram?
-
-simple_glm <- glm(data = wn, formula = s_i ~ gram + Language)
-
-```

### 2019-04-18 ngrams/ngrams.Rmd
making graphs for Psycholing working group talk

@@
 library(here)
-
+library(ggthemes)
 library(childesr)
@@
 knitr::opts_chunk$set(echo = TRUE)
+```
+
+```{r plotting for Monday}
+df <- read_feather(here("Data/childes.feather"))
+
+label_data <- data_frame(length = 8, language = "Spanish", gram_order = 6,
+                         speaker_role = c("Adult", "Child"), empirical_entropy = c(9,10.25))
+
+df %>% 
+  filter(language %in% c("English", "Mandarin", "Spanish"), gram == "bigram", length %in% c(4, 6, 8)) %>%
+  ggplot(aes(x = gram_order, y = empirical_entropy, color = speaker_role, label = speaker_role)) + 
+  facet_grid(language ~ length, scales = "free") + 
+  xlab("Gram position") + 
+  ylab("Positional entropy") + 
+  geom_pointrange(size = .25, aes(ymin = ci_lower, ymax = ci_upper)) + 
+  geom_line(size = .25) + 
+  theme_classic(base_size = 10) + 
+  scale_color_ptol() + 
+  theme(legend.position = "nones") + 
+  geom_text(data = label_data)
+
 ```
@@
 
-```{r childes}
-utterances <- get_utterances(corpus = "Plunkett") %>% 
-  filter(speaker_role != "Target_Child") %>%
-  rename(words = gloss)
-
-# unigrams to check against previous results
-unigrams <- map(2:10, ~ngram_entropy(utterances, .x, 1)) %>% 
-  bind_rows()
-
-# bigrams
-bigrams <- map(3:10, ~ngram_entropy(utterances, .x, 2)) %>% 
-  bind_rows()
-
-# trigrams
-trigrams <- map(4:10, ~ngram_entropy(utterances, .x, 3)) %>% 
-  bind_rows()
-
-# quadgrams
-quadgrams <- map(5:10, ~ngram_entropy(utterances, .x, 4)) %>%
-  bind_rows()
-
-unigrams %>%
-  bind_rows(bigrams, trigrams, quadgrams) %>%
-  mutate(Language = "Danish") %>%
-  write_feather("CHILDES/danish_adult.feather")
-```
-
 ```{r switchboard}

### 2019-04-18 ngrams/wiki_ngrams.Rmd
making graphs for Psycholing working group talk

@@
 
-```{r set up data}
-w1 %>% 
-  left_join(w2) %>%
-  left_join(w3) %>%
-  left_join(w4) %>%
-  remove_missing() %>%
-  slice(1:5) %>%
-  melt(id.vars = "Language") %>%
-  ggplot(aes(variable, value, color=Language)) + 
-  geom_point() +
-  geom_line()
-```
-
 ```{r Joe's regression}
@@
 ## why would there be an interaction between language and gram?
+## shouldn't the final model be more like
+# log(sum(abs(s_i))) = language * type + gram * type + (source | type)
+
 wn <- read_feather(here("Data/wiki_ngrams.feather"))
 
-simple_glm <- glm(data = wn, formula = s_i ~ gram + Language)
-
+simple_glm <- glm(data = wn, formula = log(s_i) ~ gram + Language)
+summary(simple_glm)
 ```
@@
 
-wn %>% write_feather(here("Data/wiki_ngrams.feather"))
+w1 %>% 
+  left_join(w2) %>%
+  left_join(w3) %>%
+  left_join(w4) %>%
+  remove_missing() %>%
+  slice(1:5) %>%
+  melt(id.vars = "Language") %>%
+  ggplot(aes(variable, value, color=Language)) + 
+  geom_point() +
+  geom_line()
+
 ```

### 2019-04-19 childes/childes_entropy.Rmd
more plots for Monday

@@
 
+```{r make plots for joint lab meeting}
+df <- prov_utterances %>%
+  extract_stem_tokens(role = "adult", min_length = 5, max_length = 9)
+
+prov_child_tokens <- extract_stem_tokens("child", 5, 9, prov_utterances) 
+prov_adult_tokens <- extract_stem_tokens("not child",5, 9, prov_utterances)
+
+prov_child_entropies <-  map_df(c(5,7,9), ~unigram_entropy(.x, prov_child_tokens)) %>%
+  mutate(language = "English", person = "Child")
+prov_adult_entropies <-  map_df(c(5,7,9), ~unigram_entropy(.x, prov_adult_tokens)) %>%
+  mutate(language = "English", person = "Adult")
+
+prov_entropies <- prov_child_entropies %>%
+  bind_rows(prov_adult_entropies)
+
+prov_entropies %>% ggplot(aes(x = word_order, y = empirical_entropy, color = person, label = person)) + 
+  facet_wrap(~ length) + 
+  xlab("Word position") + 
+  ylab("Positional entropy") + 
+  geom_pointrange(size = .25, aes(ymin = ci_lower, ymax = ci_upper)) + 
+  geom_line(size = .25) + 
+  theme_classic(base_size = 10) + 
+  scale_color_ptol() + 
+  theme(legend.position = "nones") + 
+  geom_dl(aes(label = person), method = "last.points") 
+```
 bootstrap entropy slopes

### 2019-04-19 ngrams/ngrams.Rmd
more plots for Monday

@@
 
-label_data <- data_frame(length = 8, language = "Spanish", gram_order = 6,
-                         speaker_role = c("Adult", "Child"), empirical_entropy = c(9,10.25))
+label_data <- data_frame(length = 9, language = "Spanish", gram_order = 6,
+                         speaker_role = c("Adult", "Child"), empirical_entropy = c(9.25,10.25))
 
 df %>% 
-  filter(language %in% c("English", "Mandarin", "Spanish"), gram == "bigram", length %in% c(4, 6, 8)) %>%
+  filter(language %in% c("English", "Mandarin", "Spanish"), gram == "trigram", length %in% c(5, 7, 9)) %>%
   ggplot(aes(x = gram_order, y = empirical_entropy, color = speaker_role, label = speaker_role)) + 

### 2019-04-23 surprisal/childes_surprisal.Rmd
surprisal computation

--- 2019-04-19 childes/childes_entropy.Rmd
@@
 library(gtools)
+library(here)
 
@@
 prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
+```
 
-brown_utterances <- get_utterances(corpus = "Brown") #stem or gloss
+```{r}
+adult_utterances <- prov_utterances %>%
+  filter(speaker_role == "Target_Child") %>% 
+  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
+  mutate(utterance_id = 1:n()) %>%
+  unnest_tokens(word, gloss) %>%
+  group_by(utterance_id) %>%
+  mutate(word_order = 1:n())
+```
 
-shiro_utterances <- get_utterances(corpus = "Shiro") #use gloss
+```{r unigrams}
+adult_unigrams <- adult_utterances %>%
+  group_by(word) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(p = n / sum(n))
 
-wagner_utterances <- get_utterances(corpus = "Wagner") #use gloss
+adult_surprisals <- adult_utterances %>%
+  left_join(adult_unigrams) %>%
+  mutate(s = -log(p)) %>%
+  group_by(length, word_order) %>%
+  summarise(s = mean(s))
 
-okayama_utterances <- get_utterances(corpus = "Okayama") #use gloss
+adult_surprisals %>%
+  filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
+  ggplot(aes(x = word_order, y = s)) + 
+  facet_wrap(~ length) + 
+  geom_point() + 
+  geom_smooth(se = F)
+```
 
-zhou_utterances <- get_utterances(corpus = "ZhouDinner") #use stem, NOT gloss
+```{r bigrams}
+adult_bigrams <- adult_utterances %>%
+  group_by(utterance_id) %>%
+  mutate(lag_word = lag(word)) %>%
+  group_by(lag_word, word) %>%
+  count() %>%
+  filter(!is.na(lag_word)) %>%
+  ungroup() %>%
+  mutate(joint_p = n / sum(n)) %>%
+  select(-n) %>%
+  left_join(adult_unigrams, by = c("lag_word" = "word")) %>%
+  mutate(cond_p = joint_p / p) %>%
+  select(-n, -p)
+
+bigram_surprisals_prep <- adult_utterances %>%
+  group_by(utterance_id) %>%
+  mutate(lag_word = lag(word)) %>%
+  left_join(adult_bigrams)
+
+bigram_surprisals <- bigram_surprisals_prep %>%
+  filter(!is.na(lag_word)) %>%
+  mutate(s = -log(cond_p)) %>%
+  group_by(length, word_order) %>%
+  summarise(s = mean(s))
+
+bigram_surprisals_nonnull <- bigram_surprisals_prep %>%
+  filter(!is.na(lag_word)) %>%
+  mutate(s = -log(cond_p)) %>%
+  group_by(length, word_order) %>%
+  summarise(s = mean(s))
+
+bigram_surprisals %>%
+  filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
+  ggplot(aes(x = word_order, y = s)) + 
+  facet_wrap(~ length) + 
+  geom_point() + 
+  geom_smooth(se = F)
+
 ```

### 2019-04-24 surprisal/childes_surprisal.Rmd
bigrams with first position for surprisal

@@
   mutate(lag_word = lag(word)) %>%
-  left_join(adult_bigrams)
+  left_join(adult_bigrams) %>%
+  left_join(adult_unigrams)
 
@@
   group_by(length, word_order) %>%
-  summarise(s = mean(s))
+  summarise(s = mean(s)) %>%
+  filter(word_order <= length)
 
 bigram_surprisals_nonnull <- bigram_surprisals_prep %>%
+  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
+  group_by(length, word_order) %>%
+  summarise(s = mean(s)) %>% 
+  filter(word_order <= length)
+
+bigram_surprisals_nonnull %>%
+  filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
+  ggplot(aes(x = word_order, y = s)) + 
+  facet_wrap(~ length) + 
+  geom_point() + 
+  geom_smooth(se = F)
+
+```
+
+```{r trigrams}
+adult_trigrams <- adult_utterances %>%
+  group_by(utterance_id) %>%
+  mutate(lag_word1 = lag(word)) %>%
+  mutate(lag_word2 = lag(lag_word1)) %>%
+  group_by(lag_word2, lag_word1, word) %>%
+  count() %>%
+  filter(!is.na(lag_word1), !is.na(lag_word2)) %>%
+  ungroup() %>%
+  mutate(tri_joint_p = n / sum(n)) %>%
+  select(-n) %>%
+  left_join(adult_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
+  mutate(tri_cond_p = tri_joint_p / joint_p) %>%
+  select(-cond_p, -joint_p)
+
+trigram_surprisals_prep <- adult_utterances %>%
+  group_by(utterance_id) %>%
+  mutate(lag_word1 = lag(word)) %>%
+  mutate(lag_word2 = lag(lag_word1)) %>%
+  left_join(adult_trigrams) %>%
+  left_join(adult_bigrams) %>%
+  left_join(adult_unigrams)
+
+trigram_surprisals <- trigram_surprisals_prep %>%
   filter(!is.na(lag_word)) %>%
@@
   group_by(length, word_order) %>%
-  summarise(s = mean(s))
+  summarise(s = mean(s)) %>%
+  filter(word_order <= length)
 
-bigram_surprisals %>%
+trigram_surprisals_nonnull <- trigram_surprisals_prep %>%
+  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
+  group_by(length, word_order) %>%
+  summarise(s = mean(s)) %>% 
+  filter(word_order <= length)
+
+bigram_surprisals_nonnull %>%
   filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%

### 2019-04-24 surprisal/childes_surprisal.Rmd
trigram providence surprisals, second position not recording

@@
   mutate(tri_cond_p = tri_joint_p / joint_p) %>%
-  select(-cond_p, -joint_p)
+  select(-joint_p, -tri_joint_p)
 
@@
   mutate(lag_word2 = lag(lag_word1)) %>%
-  left_join(adult_trigrams) %>%
-  left_join(adult_bigrams) %>%
+  left_join(adult_trigrams) %>% #add bigram and trigram conditionals
   left_join(adult_unigrams)
 
-trigram_surprisals <- trigram_surprisals_prep %>%
-  filter(!is.na(lag_word)) %>%
-  mutate(s = -log(cond_p)) %>%
+trigram_surprisals_nonnull <- trigram_surprisals_prep %>%
+  mutate(s = ifelse(is.na(lag_word1), # check if it's not the first word
+                    ifelse(is.na(lag_word2), # check if not the second word
+                           -log(p), # trigram if third or beyond
+                           -log(cond_p)), # bigram if second word
+                    -log(tri_cond_p))) %>% # unigram if first word
   group_by(length, word_order) %>%
@@
 
-trigram_surprisals_nonnull <- trigram_surprisals_prep %>%
-  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
-  group_by(length, word_order) %>%
-  summarise(s = mean(s)) %>% 
-  filter(word_order <= length)
-
-bigram_surprisals_nonnull %>%
+trigram_surprisals_nonnull %>%
   filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%

### 2019-04-25 surprisal/childes_surprisal.Rmd
fixed childes surprisal trigrams

@@
 ---
-title: Positional entropy, conditional entropy and mutual information in English CHILDES corpora
+title: Positional surprisal (1-3grams) in CHILDES corpora
 author: Josef Klafka and Dan Yurovsky 
@@
     code_folding: show 
+editor_options: 
+  chunk_output_type: console
 ---
@@
 prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
+zhou_utterances <- get_utterances(corpus = "ZhouDinner")
 ```
@@
 ```{r}
-adult_utterances <- prov_utterances %>%
+adult_utterances <- zhou_utterances %>%
   filter(speaker_role == "Target_Child") %>% 
@@
   mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, gloss) %>%
+  unnest_tokens(word, stem) %>%
   group_by(utterance_id) %>%
@@
   geom_smooth(se = F)
-
 ```
@@
   mutate(tri_cond_p = tri_joint_p / joint_p) %>%
-  select(-joint_p, -tri_joint_p)
+  select(-joint_p, -tri_joint_p, -cond_p) # %>% 
+  # left_join(adult_bigrams, by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
+  # select(-joint_p)
 
@@
   mutate(lag_word2 = lag(lag_word1)) %>%
-  left_join(adult_trigrams) %>% #add bigram and trigram conditionals
+  left_join(adult_trigrams) %>% 
+  left_join(select(adult_bigrams, -joint_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
   left_join(adult_unigrams)
@@
 trigram_surprisals_nonnull <- trigram_surprisals_prep %>%
-  mutate(s = ifelse(is.na(lag_word1), # check if it's not the first word
-                    ifelse(is.na(lag_word2), # check if not the second word
+  mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
+                    ifelse(is.na(lag_word1), # check if not the second word
                            -log(p), # trigram if third or beyond
@@
 ```
-
-```{r make plots for joint lab meeting}
-df <- prov_utterances %>%
-  extract_stem_tokens(role = "adult", min_length = 5, max_length = 9)
-
-prov_child_tokens <- extract_stem_tokens("child", 5, 9, prov_utterances) 
-prov_adult_tokens <- extract_stem_tokens("not child",5, 9, prov_utterances)
-
-prov_child_entropies <-  map_df(c(5,7,9), ~unigram_entropy(.x, prov_child_tokens)) %>%
-  mutate(language = "English", person = "Child")
-prov_adult_entropies <-  map_df(c(5,7,9), ~unigram_entropy(.x, prov_adult_tokens)) %>%
-  mutate(language = "English", person = "Adult")
-
-prov_entropies <- prov_child_entropies %>%
-  bind_rows(prov_adult_entropies)
-
-prov_entropies %>% ggplot(aes(x = word_order, y = empirical_entropy, color = person, label = person)) + 
-  facet_wrap(~ length) + 
-  xlab("Word position") + 
-  ylab("Positional entropy") + 
-  geom_pointrange(size = .25, aes(ymin = ci_lower, ymax = ci_upper)) + 
-  geom_line(size = .25) + 
-  theme_classic(base_size = 10) + 
-  scale_color_ptol() + 
-  theme(legend.position = "nones") + 
-  geom_dl(aes(label = person), method = "last.points") 
-```
-bootstrap entropy slopes
-```{r slope bootstrapping}
-
-#extract tokens to be used for slope bootstrapping
-extract_stem_tokens <- function(role, target_length, utterances) {
-
-  if (role == "child") {
-    sub_utterances <- utterances %>%
-      filter(speaker_role == "Target_Child") %>%
-      mutate(length = str_count(stem, pattern = " +") + 1)
-
-} else {
-    sub_utterances <- utterances %>%
-      filter(speaker_role != "Target_Child") %>%
-      mutate(length = str_count(stem, pattern = " +") + 1)
-}
-
-  tokens <- sub_utterances %>%
-    filter(length == target_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, stem) %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    group_by(word_order, word) %>%
-    filter(word_order <= length)
-
-  tokens %>%
-    summarise(n = n()) %>%
-    tidyboot(summary_function = function(x) x %>% 
-               summarise(entropy = entropy(n, unit = "log2")),
-             statistics_functions = function(x) x %>%
-             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
-    mutate(role = role, length = target_length)
-}
-
-extract_gloss_tokens <- function(role, target_length, utterances) {
-
-  if (role == "child") {
-    sub_utterances <- utterances %>%
-      filter(speaker_role == "Target_Child") %>%
-      mutate(length = str_count(gloss, pattern = " +") + 1)
-
-} else {
-    sub_utterances <- utterances %>%
-      filter(speaker_role != "Target_Child") %>%
-      mutate(length = str_count(gloss, pattern = " +") + 1)
-}
-
-  tokens <- sub_utterances %>%
-    filter(length == target_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss) %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    group_by(word_order, word) %>%
-    filter(word_order <= length)
-
-  tokens %>%
-    summarise(n = n()) %>%
-    tidyboot(summary_function = function(x) x %>% 
-               summarise(entropy = entropy(n, unit = "log2")),
-             statistics_functions = function(x) x %>%
-             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
-    mutate(role = role, length = target_length)
-}
-
-#compute the unigram entropy measure from the counts of each word in each position
-unigram_entropy <- function(tokens) {
-
-  tokens %>%
-    group_by(word_order, word) %>%
-    summarise(n = n()) %>%
-    tidyboot(summary_function = function(x) x %>% 
-               summarise(entropy = entropy(n, unit = "log2")),
-             statistics_functions = function(x) x %>%
-             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
-    mutate(role = role, length = sen_length)
-}
-
-start_slope <- function(df) {
-
- df %>%
-    filter(word_order %in% c(1,2)) %>%
-    sample_frac(1, replace = T) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    spread(word_order, entropy) %>%
-    ungroup() %>%
-    mutate(diff = `2` - `1`) %>%
-    summarise(diff = mean(diff)) %>%
-    pull(diff)
-
-}
-
-second_slope <- function(df) {
-
- df %>%
-    ungroup() %>%
-    mutate(word_order %in% c(2, 3)) %>%
-    filter(word_order > 0) %>%
-    group_by(length, word_order) %>%
-    sample_frac(1, replace = T) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    spread(word_order, entropy) %>%
-    ungroup() %>%
-    mutate(diff = `3` - `2`) %>%
-    summarise(diff = mean(diff)) %>%
-    pull(diff)
-
-}
-
-#note this measurement only makes sense if utterance length is at least 6
-mid_slope <- function(df) {
-
- df %>%
-    ungroup() %>%
-    mutate(word_order = if_else(word_order == length-2, 2,
-                                if_else(word_order == 3, 1,
-                                        0))) %>%
-    filter(word_order > 0) %>%
-    group_by(length, word_order) %>%
-    sample_frac(1, replace = T) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    spread(word_order, entropy) %>%
-    ungroup() %>%
-    mutate(diff = `2` - `1`) %>%
-    summarise(diff = mean(diff)) %>%
-    pull(diff)
-
-}
-
-penult_slope <- function(df) {
-
- df %>%
-    ungroup() %>%
-    mutate(word_order = if_else(word_order == length-1, 2,
-                                if_else(word_order == length -2, 1,
-                                        0))) %>%
-    filter(word_order > 0) %>%
-    group_by(length, word_order) %>%
-    sample_frac(1, replace = T) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    spread(word_order, entropy) %>%
-    ungroup() %>%
-    mutate(diff = `2` - `1`) %>%
-    summarise(diff = mean(diff)) %>%
-    pull(diff)
-
-}
-
-end_slope <- function(df) {
-
- df %>%
-    ungroup() %>%
-    mutate(word_order = if_else(word_order == length, 2,
-                                if_else(word_order == length -1 , 1,
-                                        0))) %>%
-    filter(word_order > 0) %>%
-    group_by(length, word_order) %>%
-    sample_frac(1, replace = T) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    spread(word_order, entropy) %>%
-    ungroup() %>%
-    mutate(diff = `2` - `1`) %>%
-    summarise(diff = mean(diff)) %>%
-    pull(diff)
-
-}
-
-get_slopes <- function(role, target_length, utterances, stemgloss) {
-
-  if (stemgloss == "stem") {
-    tokens <- extract_stem_tokens(role, target_length, utterances)  
-  } else {
-    tokens <- extract_tokens(role, target_length, target_length, utterances)
-  }
-
-  start_slopes <- replicate(1000, start_slope(tokens)) %>% 
-    as_data_frame() %>%
-    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
-              ci_lower = quantile(value, .025))
-
-  second_slopes <- replicate(1000, second_slope(tokens)) %>% 
-    as_data_frame() %>%
-    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
-              ci_lower = quantile(value, .025))
-
-  mid_slopes <- replicate(1000, mid_slope(tokens)) %>% 
-    as_data_frame() %>%
-    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
-              ci_lower = quantile(value, .025))
-
-  penult_slopes <- replicate(1000, penult_slope(tokens)) %>% 
-    as_data_frame() %>%
-    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
-              ci_lower = quantile(value, .025))
-
-  end_slopes <- replicate(1000, end_slope(tokens)) %>% 
-    as_data_frame() %>%
-    summarise(mean = mean(value), ci_upper = quantile(value, .975), 
-              ci_lower = quantile(value, .025))
-
-  rbind(start_slopes, second_slopes, mid_slopes, penult_slopes, end_slopes)
-}
-```
-
-Compute Providence Corpus Estimates
-```{r calc_pos_entropy}
-tokens <- extract_stem_tokens("not child", 6, 30, prov_utterances)
-
-get_counts <- function(role, sen_length, utterances) {
-
-  if(role == "Target_Child") {
-    sub_utterances <- utterances %>%
-      filter(speaker_role == "Target_Child") %>%
-      mutate(length = str_count(gloss, " ") + 1) # uses stems of words, not original words
-  } else {
-     sub_utterances <- utterances %>%
-      filter(speaker_role != "Target_Child") %>%
-      mutate(length = str_count(gloss, " ") + 1)
-  }
-
-  sub_utterances %>%
-    filter(length == sen_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss) %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    filter(word_order <= length)
-}
-
-get_mean_lengths <- function(role, sen_length, utterances) {
-
-  counts <- get_counts(role, sen_length, utterances)
-
-  lens <- counts %>% 
-    group_by(word_order, word) %>% 
-    mutate(word_len = str_length(word))
-
-  lens$word_len %>% 
-    aggregate(list(lens$word_order), mean)
-}
-
-get_entropies <- function(role, sen_length, utterances) {
-
-  if(role == "Target_Child") {
-    sub_utterances <- utterances %>%
-      filter(speaker_role == "Target_Child") %>%
-      mutate(length = str_count(gloss, " ") + 1) # uses stems of words, not original words
-  } else {
-     sub_utterances <- utterances %>%
-      filter(speaker_role != "Target_Child") %>%
-      mutate(length = str_count(gloss, " ") + 1)
-  }
-
-  tokens <- sub_utterances %>%
-    filter(length == sen_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss) %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    group_by(word_order, word) %>%
-    filter(word_order <= length)
-
-  tokens %>%
-    summarise(n = n()) %>%
-    tidyboot(summary_function = function(x) x %>% 
-               summarise(entropy = entropy(n, unit = "log2")),
-             statistics_functions = function(x) x %>%
-             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
-    mutate(role = role, length = sen_length)
-}
-
-tokens %>%
-    filter(word_order %in% c(1,2)) %>%
-    dplyr::group_by(length, word_order) %>%
-    dplyr::summarise(entropy = entropy(n, unit = "log2"), n = n()) %>%
-    spread(word_order, entropy) %>%
-    group_by(length) %>%
-    mutate(diff = `2` - `1`) %>%
-  ggplot(aes(x = length, y = diff)) +
-  geom_point(aes(size = n)) +
-  geom_smooth(se = F) +
-  scale_x_log10() #log length stats
-
-entropies <- map(2:10, ~get_entropies("not child", .x, brown_utterances)) %>%
-  bind_rows()
-
-ggplot(entropies, aes(x = word_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  geom_pointrange() +
-  geom_smooth(se = F)
-
-entropies_child <- map(2:10, ~ get_entropies("Target_Child", .x, prov_utterances)) %>%
-  bind_rows()
-
-ggplot(entropies_child, aes(x = word_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper)) +
-  facet_wrap(~ length) + 
-  geom_pointrange() +
-  geom_smooth(se = F)
-
-entropies_all <- bind_rows(entropies, entropies_child)
-
-ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
-                      ymin = ci_lower, ymax = ci_upper, color = role)) +
-  facet_wrap(~ length) + 
-  geom_pointrange(position = position_dodge(.25)) +
-  geom_smooth(se = F) +
-  ggtitle("Child and Adult Positional Entropy vs. Word Position (Providence corpus)")
-
-```
-
-English Mutual Information
-```{r mutual_information}
-
-#compute the pairwise Mutual Information measure between all pairs of word positions
-all_pairwise_mi <- function(df) {
-  max_len <- max(df$word_order)
-
-  combs <- combn(1:max_len, 2, simplify = F)
-
-  map(combs, ~position_mi(df, .x[1], .x[2])) %>%
-    bind_rows()
-
-}
-
-#perform the actual Mutual Information computation
-position_mi <- function(df, pos1, pos2) {
-
-  mi <- df %>%
-    filter(word_order %in% c(pos1, pos2)) %>%
-    mutate(word_order = factor(word_order, labels = c("first", "second"))) %>%
-    spread(word_order, word) %>%
-    group_by(first, second) %>%
-    summarise(n = n()) %>%
-    spread(second, n, fill = 0) %>% 
-    ungroup() %>%
-    select(-first) %>%
-    mi.empirical(., unit = "log2")
-
-  data_frame(length = max(df$word_order), pos1 = pos1, pos2 = pos2, mi = mi)
-}
-
-counts <- map(2:6, ~get_counts("not child", .x, prov_utterances))
-
-mi <- map(counts, all_pairwise_mi) %>% bind_rows()
-
-ggplot(mi, aes(x = pos2, y = mi, color = as.factor(pos1))) + 
-  facet_wrap(~length) + 
-  geom_point() + 
-  geom_line()
-
-ggplot(mi, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
-  facet_wrap(~length) + 
-  geom_point() + 
-  geom_line()
-
-```
-
-English Conditional Entropy
-```{r conditional_entropy}
-#compute all Conditional Entropy from all one-way pairs of word positions
-all_pairwise_ce <- function(df) {
-  max_len <- max(df$word_order)
-
-  perms <- permutations(max_len, 2, 1:max_len, repeats.allowed = F) 
-  combs <-  setNames(split(perms, seq(nrow(perms))), rownames(perms))
-
-  map(combs, ~conditional_entropy(df, .x[1], .x[2])) %>%
-    bind_rows()
-
-}
-
-#perform the actual Conditional Entropy calculation
-get_ce <- function(df, X, Y) { #Y is conditioning variable
-
-  d <- df %>%
-    filter(word_order %in% c(X, Y)) %>%
-    mutate(word_order = factor(word_order, labels = c("first", "second"))) %>%
-    spread(word_order, word)
-
-  ent <- d$first %>% table() %>% as.vector() %>% entropy.empirical(unit="log2")
-
-  mi <- d %>% 
-    group_by(first, second) %>%
-    summarise(n = n()) %>%
-    spread(second, n, fill = 0) %>% 
-    ungroup() %>%
-    select(-first) %>%
-    mi.empirical(., unit = "log2")
-
-  data_frame(length = max(df$word_order), X = X, Y = Y, mi = mi, ent = ent, ce = ent - mi)
-}
-
-counts <- map(2:8, ~get_counts("not child", .x, prov_utterances))
-
-ce <- map(counts, all_conditional_entropy) %>% bind_rows()
-
-ggplot(ce, aes(x = X, y = ce, color = as.factor(X))) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_line()
-
-ggplot(ce, aes(x = X, y = ce, color = as.factor(Y))) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_line() + 
-  ggtitle("Conditional entropy vs. conditioned random variable")
-
-```

### 2019-05-07 process_bnc.Rmd
bnc & switchboard surprisals

@@
 library(here)
+library(tidytext)
 
@@
 
+You want to download the file from (http://ota.ox.ac.uk/desc/2554). Move it to the same directory as this .Rmd file
+
 ```{r parse_bnc, eval = F}
-files <- list.files(here("BNC/Texts"), "*.xml", full.names = T, recursive = T)
+
+files <- list.files(here("bnc/Texts"), "*.xml", full.names = T, recursive = T)
 
@@
 
-  system(paste0("xsltproc BNC/XML/Scripts/justTheWords.xsl ", file, " > ", newname))
+  system(paste0("xsltproc bnc/XML/Scripts/justTheWords.xsl ", file, " > ", newname))
 }
@@
 map(files, parse_BNC)
-
-write_feather(txt, "switchboard/switchboard.feather")
 ```
@@
 ```{r post_process_bnc}
-bnc_tokens <- read_feather(here("../bnc/bnc.feather")) %>%
+bnc_tokens <- read_feather(here("bnc/bnc.feather")) %>%
   mutate(length = str_count(value, pattern = " +") + 1) %>% 
   mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, value, token = stringr::str_split, pattern = " +") %>%
+  unnest_tokens(word, value, token = stringr::str_split, pattern = "[ +]+") %>%
   group_by(utterance_id) %>%

### 2019-05-07 surprisal/bnc&switchboard.Rmd
bnc & switchboard surprisals

---
title: "Switchboard and BNC Suprisal"
author: "Josef Klafka & Dan Yurovsky"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(directlabels)
library(tidytext)
library(tidyboot)
library(dplyr)
library(tokenizers)
library(gtools)
library(here)
library(feather)

knitr::opts_chunk$set(echo = TRUE)
```

```{r get switchboard}
switchboard <- read_feather(here("../switchboard/switchboard.feather")) %>%
  rename(text = value) %>%
  mutate(length = str_count(text, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, text, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n())
```

```{r switchboard unigrams}
switch_unigrams <- switchboard %>%
  group_by(word) %>%
  count() %>%
  ungroup() %>%
  mutate(p = n / sum(n))

switch_surprisals <- switchboard %>%
  left_join(switch_unigrams) %>%
  mutate(s = -log(p)) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s))

switch_surprisals %>%
  filter(length %in% c(3, 5, 7)) %>%
  ggplot(aes(x = word_order, y = s)) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_smooth(se = F)
```

```{r switchboard bigrams}
switch_bigrams <- switchboard %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  mutate(joint_p = n / sum(n)) %>%
  select(-n) %>%
  left_join(switch_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = joint_p / p) %>%
  select(-n, -p)

bigram_surprisals_prep <- switchboard %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  left_join(switch_bigrams) %>%
  left_join(switch_unigrams)

bigram_surprisals <- bigram_surprisals_prep %>%
  filter(!is.na(lag_word)) %>%
  mutate(s = -log(cond_p)) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>%
  filter(word_order <= length)

bigram_surprisals_nonnull <- bigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)

bigram_surprisals_nonnull %>%
  filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
  ggplot(aes(x = word_order, y = s)) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_smooth(se = F)
```

```{r switchboard trigrams}
switch_trigrams <- switchboard %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  group_by(lag_word2, lag_word1, word) %>%
  count() %>%
  filter(!is.na(lag_word1), !is.na(lag_word2)) %>%
  ungroup() %>%
  mutate(tri_joint_p = n / sum(n)) %>%
  select(-n) %>%
  left_join(switch_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
  mutate(tri_cond_p = tri_joint_p / joint_p) %>%
  select(-joint_p, -tri_joint_p, -cond_p) # %>% 
  # left_join(adult_bigrams, by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
  # select(-joint_p)

trigram_surprisals_prep <- switchboard %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  left_join(switch_trigrams) %>% 
  left_join(select(switch_bigrams, -joint_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
  left_join(switch_unigrams)

trigram_surprisals_nonnull <- trigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
                    ifelse(is.na(lag_word1), # check if not the second word
                           -log(p), # trigram if third or beyond
                           -log(cond_p)), # bigram if second word
                    -log(tri_cond_p))) %>% # unigram if first word
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)

trigram_surprisals_nonnull %>%
  filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
  ggplot(aes(x = word_order, y = s)) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_smooth(se = F)
```

A pre-requisite for the next section is that you run the "process_bnc.Rmd" pipeline for preprocessing the British National Corpus (bnc). The files are too large to store in the GitHub repository.

```{r bnc}
bnc <- read_feather(here("../bnc/bnc10.feather")) %>%
  bind_rows(read_feather(here("../bnc/bnc20.feather"))) %>%
  bind_rows(read_feather(here("../bnc/bnc30.feather")))
```

```{r bnc unigrams}
bnc_unigrams <- bnc %>%
  group_by(word) %>%
  count() %>%
  ungroup() %>%
  mutate(p = n / sum(n))

bnc_surprisals <- bnc %>%
  left_join(bnc_unigrams) %>%
  mutate(s = -log(p)) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s))

bnc_surprisals %>%
  ggplot(aes(x = word_order, y = s)) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_smooth(se = F)
```

```{r bnc bigrams}
bnc_bigrams <- bnc %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  mutate(joint_p = n / sum(n)) %>%
  select(-n) %>%
  left_join(bnc_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = joint_p / p) %>%
  select(-n, -p)

bigram_surprisals_prep <- bnc %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  left_join(bnc_bigrams) %>%
  left_join(bnc_unigrams)

bigram_surprisals <- bigram_surprisals_prep %>%
  filter(!is.na(lag_word)) %>%
  mutate(s = -log(cond_p)) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>%
  filter(word_order <= length)

bigram_surprisals_nonnull <- bigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)

bigram_surprisals_nonnull %>%
  ggplot(aes(x = word_order, y = s)) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_smooth(se = F)
```

```{r bnc trigrams}
bnc_trigrams <- bnc %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  group_by(lag_word2, lag_word1, word) %>%
  count() %>%
  filter(!is.na(lag_word1), !is.na(lag_word2)) %>%
  ungroup() %>%
  mutate(tri_joint_p = n / sum(n)) %>%
  select(-n) %>%
  left_join(bnc_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
  mutate(tri_cond_p = tri_joint_p / joint_p) %>%
  select(-joint_p, -tri_joint_p, -cond_p) # %>%

trigram_surprisals_prep <- bnc %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  left_join(bnc_trigrams) %>% 
  left_join(select(bnc_bigrams, -joint_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
  left_join(bnc_unigrams)

trigram_surprisals_nonnull <- trigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
                    ifelse(is.na(lag_word1), # check if not the second word
                           -log(p), # trigram if third or beyond
                           -log(cond_p)), # bigram if second word
                    -log(tri_cond_p))) %>% # unigram if first word
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)

trigram_surprisals_nonnull %>%
  ggplot(aes(x = word_order, y = s)) + 
  facet_wrap(~ length) + 
  geom_point() + 
  geom_smooth(se = F)
```

### 2019-05-07 surprisal/childes_surprisal.Rmd
bnc & switchboard surprisals

@@
 prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
-zhou_utterances <- get_utterances(corpus = "ZhouDinner")
+zhou_utterances <- get_utterances(corpus = "ZhouDinner") #stem for pinyin
 ```
@@
 ```{r}
-adult_utterances <- zhou_utterances %>%
+adult_utterances <- prov_utterances %>%
   filter(speaker_role == "Target_Child") %>% 
-  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
+  mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
   mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, stem) %>%
+  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
   group_by(utterance_id) %>%

### 2019-05-08 surprisal/bnc&switchboard.Rmd
bootstrap_mean

@@
 library(feather)
+library(gridExtra)
+library(tidyboot)
 
@@
   summarise(s = mean(s))
+  # tidyboot_mean(s)
 
@@
   geom_smooth(se = F)
+```
+
+```{r bootstrapping}
+
+# extract_stem_tokens <- function(role, target_length, utterances) {
+#   
+#   if (role == "child") {
+#     sub_utterances <- utterances %>%
+#       filter(speaker_role == "Target_Child") %>%
+#       mutate(length = str_count(stem, pattern = " +") + 1)
+# 
+# } else {
+#     sub_utterances <- utterances %>%
+#       filter(speaker_role != "Target_Child") %>%
+#       mutate(length = str_count(stem, pattern = " +") + 1)
+# }
+#   
+#   tokens <- sub_utterances %>%
+#     filter(length == target_length) %>%
+#     mutate(utterance_id = 1:n()) %>%
+#     unnest_tokens(word, stem) %>%
+#     group_by(utterance_id) %>%
+#     mutate(word_order = 1:n()) %>%
+#     group_by(word_order, word) %>%
+#     filter(word_order <= length)
+#    
+  # tokens %>%
+  #   summarise(n = n()) %>%
+  #   tidyboot(summary_function = function(x) x %>%
+  #              summarise(entropy = entropy(n, unit = "log2")),
+  #            statistics_functions = function(x) x %>%
+  #            summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
+  #   mutate(role = role, length = target_length)
+# }
 ```
@@
 
-bigram_surprisals <- bigram_surprisals_prep %>%
-  filter(!is.na(lag_word)) %>%
-  mutate(s = -log(cond_p)) %>%
-  group_by(length, word_order) %>%
-  summarise(s = mean(s)) %>%
-  filter(word_order <= length)
-
-bigram_surprisals_nonnull <- bigram_surprisals_prep %>%
+switch_bigram_surprisals <- bigram_surprisals_prep %>%
   mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
@@
 
-bigram_surprisals_nonnull %>%
+switch_bigram_surprisals %>%
   filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
@@
   select(-joint_p, -tri_joint_p, -cond_p) # %>%
-  # left_join(adult_bigrams, by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
-  # select(-joint_p)
 
@@
 
-trigram_surprisals_nonnull <- trigram_surprisals_prep %>%
+switch_trigram_surprisals <- trigram_surprisals_prep %>%
   mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
@@
 
-trigram_surprisals_nonnull %>%
+switch_trigram_surprisals %>%
   filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
@@
 
-bigram_surprisals <- bigram_surprisals_prep %>%
-  filter(!is.na(lag_word)) %>%
-  mutate(s = -log(cond_p)) %>%
-  group_by(length, word_order) %>%
-  summarise(s = mean(s)) %>%
-  filter(word_order <= length)
-
-bigram_surprisals_nonnull <- bigram_surprisals_prep %>%
+bnc_bigram_surprisals <- bigram_surprisals_prep %>%
   mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
@@
 
-bigram_surprisals_nonnull %>%
+bnc_bigram_surprisals %>%
   ggplot(aes(x = word_order, y = s)) + 
@@
 
-trigram_surprisals_nonnull <- trigram_surprisals_prep %>%
+bnc_trigram_surprisals <- trigram_surprisals_prep %>%
   mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
@@
 
-trigram_surprisals_nonnull %>%
+bnc_trigram_surprisals %>%
   ggplot(aes(x = word_order, y = s)) + 
@@
 ```
+
+```{r plotting}
+bnc_plot <- ggplot(bnc_surprisals, aes(x = word_order, y = s)) +
+  facet_wrap(~ length, scales = "free_x") + 
+  xlab("British National Corpus word position") +
+  ylab("Surprisal") +
+  geom_smooth(se = F) +
+  theme_classic(base_size = 10)
+
+sw_plot <- switch_surprisals %>%
+  filter(length %in% c(5, 7, 9)) %>% 
+  ggplot(aes(x = word_order, y = s)) +
+  facet_wrap(~ length, scales = "free_x") + 
+  xlab("Switchboard word position") +
+  ylab("Surprisal") +
+  geom_smooth(se = F) +
+  theme_classic(base_size = 10)
+
+grid.arrange(bnc_plot, sw_plot, nrow = 2)
+```

### 2019-05-09 surprisal/bnc&switchboard.Rmd
worked out bootstrap for unigram surprisals

@@
 
-switch_surprisals <- switchboard %>%
+sw <- switchboard %>%
   left_join(switch_unigrams) %>%
   mutate(s = -log(p)) %>%
-  group_by(length, word_order) %>% 
-  summarise(s = mean(s))
+  ungroup() %>%
+  select(length, word_order, s) %>%
+  nest(s)
+
+swc <- sw %>% filter(length %in% c(10, 20, 30))
+switch_surprisals <- map_dfr(1:nrow(swc), 
+        ~bind_cols(select(slice(swc, .x), length, word_order), 
+                               tidyboot_mean(data = unnest(slice(swc, .x)), 
+                                             column = s, nboot = 100)))
+
+# switch_surprisals <- switchboard %>%
+#   left_join(switch_unigrams) %>%
+#   mutate(s = -log(p)) %>%
+#   ungroup() %>%
+#   select(length, word_order, s) %>%
+#   filter(length %in% c(10, 20, 30)) %>%
+#   group_by(length, word_order) %>% 
 #   tidyboot_mean(s)
@@
 switch_surprisals %>%
-  filter(length %in% c(3, 5, 7)) %>%
-  ggplot(aes(x = word_order, y = s)) + 
+  ggplot(aes(x = word_order, y = empirical_stat, #observed mean surprisal
+             ymin = ci_lower, ymax = ci_upper)) + 
   facet_wrap(~ length) + 
-  geom_point() + 
-  geom_smooth(se = F)
-```
-
-```{r bootstrapping}
-
-# extract_stem_tokens <- function(role, target_length, utterances) {
-#   
-#   if (role == "child") {
-#     sub_utterances <- utterances %>%
-#       filter(speaker_role == "Target_Child") %>%
-#       mutate(length = str_count(stem, pattern = " +") + 1)
-# 
-# } else {
-#     sub_utterances <- utterances %>%
-#       filter(speaker_role != "Target_Child") %>%
-#       mutate(length = str_count(stem, pattern = " +") + 1)
-# }
-#   
-#   tokens <- sub_utterances %>%
-#     filter(length == target_length) %>%
-#     mutate(utterance_id = 1:n()) %>%
-#     unnest_tokens(word, stem) %>%
-#     group_by(utterance_id) %>%
-#     mutate(word_order = 1:n()) %>%
-#     group_by(word_order, word) %>%
-#     filter(word_order <= length)
-#    
-  # tokens %>%
-  #   summarise(n = n()) %>%
-  #   tidyboot(summary_function = function(x) x %>%
-  #              summarise(entropy = entropy(n, unit = "log2")),
-  #            statistics_functions = function(x) x %>%
-  #            summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
-  #   mutate(role = role, length = target_length)
-# }
+  ylab("Surprisal") + 
+  xlab("Word Position in Switchboard") + 
+  geom_pointrange(size = .50) + 
+  geom_line(size = .50)
 ```
@@
 
-bnc_surprisals <- bnc %>%
+bnc_surprisals_prep <- bnc %>% 
   left_join(bnc_unigrams) %>%
   mutate(s = -log(p)) %>%
-  group_by(length, word_order) %>%
-  summarise(s = mean(s))
+  ungroup() %>%
+  select(length, word_order, s) %>%
+  nest(s)
+
+bnc_surprisals <- map_dfr(1:nrow(bnc_surprisals_prep), 
+        ~bind_cols(select(slice(bnc_surprisals_prep, .x), length, word_order), 
+                               tidyboot_mean(data = unnest(slice(bnc_surprisals_prep, .x)), 
+                                             column = s, nboot = 100)))
+
+# bnc_surprisals <- bnc %>%
+#   left_join(bnc_unigrams) %>%
+#   mutate(s = -log(p)) %>%
+#   select(length, word_order, s) %>%
+#   filter(length == 10) %>%
+#   group_by(length, word_order) %>%
+#   # summarise(s = mean(s)) 
+#   tidyboot_mean(s)
 
 bnc_surprisals %>%
-  ggplot(aes(x = word_order, y = s)) + 
+  ggplot(aes(x = word_order, y = empirical_stat, #observed mean surprisal
+             ymin = ci_lower, ymax = ci_upper)) + 
   facet_wrap(~ length) + 
-  geom_point() + 
-  geom_smooth(se = F)
+  ylab("Surprisal") + 
+  xlab("Word Position in BNC") + 
+  geom_pointrange(size = .50) + 
+  geom_line(size = .50)
+
+# bnc_surprisals %>%
+#   ggplot(aes(x = word_order, y = s)) + 
+#   facet_wrap(~ length) + 
+#   geom_point() + 
+#   geom_smooth(se = F)
 ```

### 2019-05-09 previous/cogsci/scripts/plot_slopes.R

@@ -1,0 +1,32 @@
+library(tidyverse)
+library(here)
+
+data <- read_csv(here("../wiki_entropy/Wikipedia_relative5.csv")) %>%
+  gather(position, slope, Slope1:Slope5) %>%
+  mutate(position = gsub("[^0-9]","", position),
+         position = as.numeric(position))
+
+languages <- distinct(data, Language)
+
+sub_data <- data %>%
+  filter(Language %in% (slice(languages, 1:10) %>% pull(Language))) %>%
+  spread(position, slope) %>%
+  mutate(value0 = 0,
+         value1 = value0 + `1`,
+         value2 = value1 + `2`,
+         value3 = value2 + `3`, 
+         value4 = value3 + `4`,
+         value5 = value4 + `5`) %>%
+  select(-`1`:-`5`) %>%
+  gather(position, value, value0:value5) %>%
+  mutate(position = gsub("[^0-9]","", position),
+         position = as.numeric(position)) %>%
+  group_by(Language) %>%
+  mutate(mid_value = value[3]) %>%
+  mutate(centered_value = value - mid_value)
+
+
+ggplot(sub_data,
+       aes(x = position, y = centered_value, color = Language)) + 
+  geom_line() + 
+  theme_classic()

### 2019-05-09 previous/cogsci/scripts/process_bnc_switchboard.R

@@ -1,0 +1,44 @@
+library(here)
+library(feather)
+library(tidyverse)
+library(tidyboot)
+library(entropy)
+library(stringr)
+library(tidytext)
+
+sb_tokens <- read_feather(here("../switchboard/switchboard.feather")) %>%
+  mutate(length = str_count(value, pattern = " +") + 1) %>% 
+  mutate(utterance_id = 1:n()) %>%
+  unnest_tokens(word, value, token = stringr::str_split, pattern = " +") %>%
+  group_by(utterance_id) %>%
+  mutate(word_order = 1:n()) 
+
+bnc10_tokens <- read_feather(here("../bnc/bnc10.feather")) %>%
+  mutate(word = as.numeric(as.factor(word)))
+bnc20_tokens <- read_feather(here("../bnc/bnc20.feather")) %>%
+  mutate(word = as.numeric(as.factor(word)))
+bnc30_tokens <- read_feather(here("../bnc/bnc30.feather")) %>%
+  mutate(word = as.numeric(as.factor(word)))
+
+unigram_entropy <- function(sen_length, tokens) {
+  tokens %>%
+    filter(length == sen_length) %>%
+    group_by(word_order, word) %>%
+    summarise(n = n()) %>%
+    tidyboot(summary_function = function(x) x %>% 
+               summarise(entropy = entropy(n, unit = "log2")),
+             statistics_functions = function(x) x %>%
+               summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
+    mutate(length = sen_length)
+}
+
+switchboard_entropies <- map_df(2:10, ~unigram_entropy(.x, sb_tokens)) 
+
+bnc10_entropies <- unigram_entropy(10, bnc10_tokens)
+bnc20_entropies <- unigram_entropy(20, bnc20_tokens)
+bnc30_entropies <- unigram_entropy(30, bnc30_tokens)
+
+bnc_entropies <- bind_rows(bnc10_entropies, bnc20_entropies, bnc30_entropies)
+
+write_csv(bnc_entropies, here("data/bnc_entropies.csv"))
+write_csv(switchboard_entropies, here("data/switchboard_entropies.csv"))

### 2019-05-13 surprisal/bnc&switchboard.Rmd
set up wiki_relative_surprisals

@@
                                tidyboot_mean(data = unnest(slice(swc, .x)), 
-                                             column = s, nboot = 100)))
+                                             column = s, nboot = 1000)))
 

### 2019-05-13 surprisal/childes_surprisal.Rmd
set up wiki_relative_surprisals

--- 2019-04-25 surprisal/childes_surprisal.Rmd
@@
 library(here)
+library(ggthemes)
 
@@
 prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
-zhou_utterances <- get_utterances(corpus = "ZhouDinner")
+zhou_utterances <- get_utterances(corpus = "ZhouDinner") #stem for pinyin
+shiro_utterances <- get_utterances(corpus = "Shiro")
 ```
 
-```{r}
-adult_utterances <- zhou_utterances %>%
-  filter(speaker_role == "Target_Child") %>% 
-  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
+```{r old unigrams}
+adult_utterances <- prov_utterances %>%
+  filter(speaker_role != "Target_Child") %>% 
+  mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
   mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, stem) %>%
+  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
   group_by(utterance_id) %>%
   mutate(word_order = 1:n())
-```
 
-```{r unigrams}
 adult_unigrams <- adult_utterances %>%
@@
 adult_surprisals %>%
-  filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
+  filter(length %in% c(5, 7, 9)) %>%
   ggplot(aes(x = word_order, y = s)) + 
@@
   geom_smooth(se = F)
+```
+
+```{r unigrams}
+get_unigrams <- function(utterances, speaker_role, stem_or_gloss) {
+
+  if (tolower(speaker_role) == "child") {
+    utts <- utterances %>%
+      filter(speaker_role == "Target_Child") 
+  } else {
+    utts <- utterances %>%
+      filter(speaker_role != "Target_Child")   
+  }
+
+  if (tolower(stem_or_gloss) == "stem") {
+    utts <- utts %>%
+      mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
+      mutate(utterance_id = 1:n()) %>%
+      unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
+      group_by(utterance_id) %>%
+      mutate(word_order = 1:n())
+
+  } else {
+    utts <- utts %>%
+      mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
+      mutate(utterance_id = 1:n()) %>%
+      unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
+      group_by(utterance_id) %>%
+      mutate(word_order = 1:n())
+  }
+
+  unigrams <- utts %>%
+    group_by(word) %>%
+    count() %>%
+    ungroup() %>%
+    mutate(p = n / sum(n)) %>%
+    mutate(s = -log(p)) %>%
+    select(word, s)
+
+  surprisals <- utts %>%
+    left_join(unigrams) %>%
+    group_by(length, word_order) %>%
+    summarise(s = mean(s)) %>%
+    ungroup()
+}
+
+prov_sups <- get_unigrams(prov_utterances, "not child", "stem") %>% 
+  mutate(speaker_role = "adult") %>%
+  bind_rows(mutate(get_unigrams(prov_utterances, "child", "stem"), speaker_role = "child")) %>%
+  mutate(language = "English")
+
+shiro_sups <- get_unigrams(shiro_utterances, "not child", "gloss") %>% 
+  mutate(speaker_role = "adult") %>%
+  bind_rows(mutate(get_unigrams(shiro_utterances, "child", "gloss"), speaker_role = "child")) %>%
+  mutate(language = "Spanish")
+
+zhou_sups <- get_unigrams(zhou_utterances, "not child", "stem") %>% 
+  mutate(speaker_role = "adult") %>%
+  bind_rows(mutate(get_unigrams(zhou_utterances, "child", "stem"), speaker_role = "child")) %>%
+  mutate(language = "Mandarin")
+
+all_sups <- prov_sups %>%
+  bind_rows(shiro_sups, zhou_sups)
+
+label_data <- data_frame(length = 9, language = "Mandarin", word_order = 6,
+                         speaker_role = c("adult", "child"), s = c(7,9.5))
+
+all_sups %>% 
+  filter(length %in% c(5, 7, 9)) %>%
+  ggplot(aes(x = word_order, y = s,
+                        color = speaker_role,
+                        label = speaker_role)) +
+    facet_grid(language ~ length, scales = "free") + 
+    xlab("Word position") +
+    ylab("Mean surprisal") +
+    geom_line(size = .5) +
+    theme_classic(base_size = 10) + 
+    scale_color_ptol() +
+    theme(legend.position = "nones") + 
+    geom_text(data = label_data)
+
 ```

### 2019-05-13 surprisal/get_wiki_df.py

@@ -1,0 +1,96 @@
+import regex as re
+import numpy as np
+import pandas as pd
+import csv
+import sys
+import json
+import subprocess
+import urllib.request
+
+
+with open("language_dict.json", 'r') as f:
+	file = f.read()
+LANG_DICT = json.loads(file) #maps languages to their wikipedia url prefixes
+
+
+def extract_texts(upper_bound):
+	'''
+	Given an upper_bound on file numbers, extracts the raw text from the files
+	in the text/AA folder. That folder is given as output by the
+	cirrus_extract program.
+	'''
+	rs = ""
+	for i in range(upper_bound):
+		if i < 10:
+			number = "0" + str(i)
+		else:
+			number = str(i)
+		f = open("text/AA/wiki_" + number, 'r')
+		text = f.read()
+		rs += text
+	return rs
+
+
+def get_sen_dict(text):
+	'''
+	Given the output of extract_texts as a string, constructs a
+	dictionary mapping each sentence in the string to its length in words.
+	'''
+	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:–'&*#/—“»]|\d+|\(|\)|\[|\]", \
+		"", text)
+	text = re.sub('"', '', text)
+	#just the words from the articles
+	text = text.replace('\n', " ")
+	sens = re.split("[.!?]", text.lower())
+	sen_dict = {}
+	for sen in sens:
+		length = len(re.findall("\w+", sen))
+		if length > 0:
+			sen_dict[sen] = length
+	return sen_dict
+
+
+def get_sen_df(text):
+	'''
+	Given the output of extract_texts as a string, returns a Pandas dataframe
+	of the output of get_sen_dict.
+	'''
+	sd = get_sen_dict(text)
+	s = pd.Series(sd, name = "length")
+	del sd
+	s.index.name = "gloss"
+	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence
+					 	  #and the sentence length respectively
+	return df
+
+
+def main(lang_name):
+	'''
+	Given the English name of a language, constructs a dataframe with the
+	sentences of a random subset of articles in that language from Wikipedia,
+	and stores the dataframe in a csv in the working directory.
+	'''
+	lang_prefix = LANG_DICT[lang_name]
+	url = "https://dumps.wikimedia.org/other/cirrussearch/20190318/" \
+		+ lang_prefix + "wiki-20190318-cirrussearch-content.json.gz"
[…]

### 2019-05-13 surprisal/test_wiki.R

@@ -1,0 +1,100 @@
+library(tidyverse)
+library(tidytext)
+library(directlabels)
+library(googlesheets)
+
+extract_wiki_tokens <- function(min_length, max_length, utterances) {
+  
+  tokens <- utterances %>%
+    mutate(gloss = str_trim(gloss)) %>%
+    mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
+    filter(length >= min_length) %>%
+    filter(length <= max_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(length, word_order, word) %>%
+    summarise(n = n()) %>%
+    ungroup()
+  
+  sups <- tokens %>% 
+    mutate(p = n/sum(n)) %>%
+    mutate(s = -log(p)) %>%
+    select(word, s)
+  
+  tokens %>%
+    left_join(sups)
+}
+
+get_quantiles <- function(df, num_sections) {
+  quantile(df$word_order, probs = seq(0, 1, 1/num_sections)) %>%
+    round() %>%
+    as_data_frame() %>%
+    rename(word_order = value)
+}
+
+relative_slopes <- function(df_tokens, pos_list) {
+  
+  get_slope <- function(x) {
+    
+    selected_pos <- pos_list %>%
+      group_by(length) %>%
+      slice(x:(x+1)) %>%
+      group_by(length) %>%
+      summarise(min = min(word_order),
+                max = max(word_order)) %>%
+      split(.$length) %>%
+      map(~seq(.x$min, .x$max, 1) %>% as_data_frame) %>%
+      bind_rows(.id = "length") %>%
+      rename(word_order = value) %>%
+      mutate(length = as.numeric(length))
+    
+    df_tokens %>%
+      inner_join(selected_pos, by = c("length", "word_order")) %>%
+      group_by(length, word_order) %>%
+      summarise(s = mean(s)) %>%
+      lm(s ~ length + I(length^2) + word_order, data  = .) %>%
+      tidy() %>%
+      filter(term == "word_order")
+  }
+  
+  divides <- pos_list %>%
[…]

### 2019-05-13 surprisal/wiki_extract/cirrus_extract.py

@@ -1,0 +1,246 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+#
+# =============================================================================
+#  Version: 1.00 (December 15, 2015)
+#  Author: Giuseppe Attardi (attardi@di.unipi.it), University of Pisa
+#
+# =============================================================================
+#  Copyright (c) 2015. Giuseppe Attardi (attardi@di.unipi.it).
+# =============================================================================
+#  This file is part of Tanl.
+#
+#  Tanl is free software; you can redistribute it and/or modify it
+#  under the terms of the GNU General Public License, version 3,
+#  as published by the Free Software Foundation.
+#
+#  Tanl is distributed in the hope that it will be useful,
+#  but WITHOUT ANY WARRANTY; without even the implied warranty of
+#  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+#  GNU General Public License for more details.
+#
+#  You should have received a copy of the GNU General Public License
+#  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+# =============================================================================
+
+"""Wikipedia Cirrus Extractor:
+Extracts and cleans text from a Wikipedia Cirrus dump and stores output in a
+number of files of similar size in a given directory.
+Each file will contain several documents in the format:
+
+    <doc id="" url="" title="">
+        ...
+        </doc>
+
+"""
+
+import sys, os.path, time
+import re
+import json
+import argparse
+import bz2
+import gzip
+import logging
+
+# Program version
+version = '1.00'
+
+urlbase = 'http://it.wikipedia.org/'
+
+# ----------------------------------------------------------------------
+
+class NextFile(object):
+    """
+    Synchronous generation of next available file name.
+    """
+
+    filesPerDir = 100
+
+    def __init__(self, path_name):
+        self.path_name = path_name
+        self.dir_index = -1
+        self.file_index = -1
+
+    def next(self):
+        self.file_index = (self.file_index + 1) % NextFile.filesPerDir
+        if self.file_index == 0:
+            self.dir_index += 1
+        dirname = self._dirname()
+        if not os.path.isdir(dirname):
+            os.makedirs(dirname)
+        return self._filepath()
+
+    def _dirname(self):
+        char1 = self.dir_index % 26
+        char2 = self.dir_index / 26 % 26
+        return os.path.join(self.path_name, '%c%c' % (ord('A') + int(char2), ord('A') + int(char1)))
+
+    def _filepath(self):
[…]

### 2019-05-13 surprisal/wiki_script.sh

@@ -1,0 +1,42 @@
+#!/bin/bash
+
+declare -a arr=("Basque" "Dutch" "Alemannic" "Wolof" "Bosnian" "Hindi" "Danish"
+	"Russian" "Urdu" "German" "Swedish" "Afrikaans" "Anglo-Saxon" "Icelandic"
+	"Belarusian" "Bulgarian" "Croatian" "Macedonian" "Serbo-Croatian" "Ukrainian"
+	"Vietnamese" "Pangasinan" "Kapampangan" "Khmer" "Japanese" "Javanese"
+	"Indonesian" "Polish" "Malay" "Sundanese" "Hakka" "Turkish" "Kazakh"
+	"Kirghiz" "Turkmen" "Hebrew" "Persian" "Arabic" "Tajik" "Latin" "French"
+	"Galician" "Romanian" "Norman" "Corsican" "Maori" "Nepali" "Mongolian"
+	"Estonian" "Finnish" "Hungarian" "Bengali" "Waray" "Welsh" "Armenian"
+	"Korean" "Greek" "Catalan" "Walloon" "Yiddish" "Maltese" "Romansh"
+	"Banyumasan" "Võro" "Samogitian" "Chavacano")
+
+declare -a newarr=("Basque" "Dutch" "Alemannic" "Wolof" "Bosnian" "Hindi"
+	"Danish" "Russian" "Urdu" "German" "Swedish" "Afrikaans" "Anglo-Saxon"
+	"Icelandic" "Belarusian" "Bulgarian" "Croatian" "Macedonian" "Serbo-Croatian"
+	"Ukrainian" "Vietnamese" "Pangasinan" "Kapampangan" "Khmer" "Japanese"
+	"Javanese" "Indonesian" "Polish" "Malay" "Sundanese" "Hakka" "Turkish"
+	"Kazakh" "Kirghiz" "Turkmen" "Hebrew" "Persian" "Arabic" "Tajik" "Latin"
+	"French" "Galician" "Romanian" "Norman" "Corsican" "Maori" "Nepali" "Mongolian"
+	"Estonian" "Finnish" "Hungarian" "Bengali" "Waray" "Welsh" "Armenian" "Korean"
+	"Greek" "Catalan" "Walloon" "Yiddish" "Maltese" "Romansh" "Bavarian" "Faroese"
[…]

### 2019-05-14 surprisal/bnc&switchboard.Rmd
more changes

@@
   mutate(p = n / sum(n))
+
+switch_surprisals <- switchboard %>%
+  left_join(switch_unigrams) %>%
+  mutate(s = -log(p)) %>%
+  group_by(length, word_order) %>%
+  summarise(s = mean(s))
 
@@
 
+bnc_surprisals <- bnc %>%
+  left_join(bnc_unigrams) %>%
+  mutate(s = -log(p)) %>%
+  group_by(length, word_order) %>%
+  summarise(s = mean(s))
+
 bnc_surprisals_prep <- bnc %>% 
@@
   xlab("British National Corpus word position") +
-  ylab("Surprisal") +
-  geom_smooth(se = F) +
+  ylab("Mean surprisal") +
+  geom_line() + 
+  geom_point() + 
   theme_classic(base_size = 10)
@@
 sw_plot <- switch_surprisals %>%
-  filter(length %in% c(5, 7, 9)) %>% 
+  filter(length %in% c(10, 20, 30)) %>% 
   ggplot(aes(x = word_order, y = s)) +
@@
   xlab("Switchboard word position") +
-  ylab("Surprisal") +
-  geom_smooth(se = F) +
+  ylab("Mean surprisal") +
+  geom_line() +
+  geom_point() +
   theme_classic(base_size = 10)

### 2019-05-14 surprisal/childes_surprisal.Rmd
more changes

@@
 ```{r old unigrams}
-adult_utterances <- prov_utterances %>%
-  filter(speaker_role != "Target_Child") %>% 
-  mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
+adult_utterances <- shiro_utterances %>%
+  filter(speaker_role == "Target_Child") %>% 
+  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
   mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
+  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
   group_by(utterance_id) %>%
@@
 
-label_data <- data_frame(length = 9, language = "Mandarin", word_order = 6,
-                         speaker_role = c("adult", "child"), s = c(7,9.5))
+label_data <- data_frame(length = 9, Language = "Spanish", word_order = 3,
+                         speaker_role = c("Adult", "Child"), s = c(1.25,3.5))
 
-all_sups %>% 
+all_trigrams %>% 
   filter(length %in% c(5, 7, 9)) %>%
@@
                         label = speaker_role)) +
-    facet_grid(language ~ length, scales = "free") + 
+    facet_grid(Language ~ length, scales = "free") + 
     xlab("Word position") +
     ylab("Mean surprisal") +
-    geom_line(size = .5) +
+    geom_line(size = 1) +
+    geom_point() + 
     theme_classic(base_size = 10) + 
@@
 
-bigram_surprisals <- bigram_surprisals_prep %>%
-  filter(!is.na(lag_word)) %>%
-  mutate(s = -log(cond_p)) %>%
-  group_by(length, word_order) %>%
-  summarise(s = mean(s)) %>%
-  filter(word_order <= length)
-
 bigram_surprisals_nonnull <- bigram_surprisals_prep %>%
@@
   filter(word_order <= length)
+
+all_bigrams <- all_bigrams %>% 
+  bind_rows(mutate(bigram_surprisals_nonnull, speaker_role = "Child", Language = "Spanish"))
 
@@
 
+all_trigrams <- all_trigrams %>% 
+  bind_rows(mutate(trigram_surprisals_nonnull, speaker_role = "Child", Language = "Spanish"))
+
 trigram_surprisals_nonnull %>%

### 2019-05-14 surprisal/test_wiki.R

@@ -1,7 +1,7 @@
 library(tidyverse)
 library(tidytext)
 library(directlabels)
-library(googlesheets)
+# library(googlesheets)
 
 extract_wiki_tokens <- function(min_length, max_length, utterances) {
   
@@ -18,7 +18,7 @@ extract_wiki_tokens <- function(min_length, max_length, utterances) {
     summarise(n = n()) %>%
     ungroup()
   
-  sups <- tokens %>% 
+  sups <- tokens %>%
     mutate(p = n/sum(n)) %>%
     mutate(s = -log(p)) %>%
     select(word, s)
@@ -73,15 +73,15 @@ relative_slopes <- function(df_tokens, pos_list) {
 
 NUM_SECTIONS = 5
 
-LANGUAGE <- commandArgs(trailingOnly=TRUE)[1]
+LANGUAGE <- "Alemannic" #commandArgs(trailingOnly=TRUE)[1]
 
 system(paste("python3 get_wiki_df.py", LANGUAGE, sep = " "))
 
-for_gs <- gs_title("wiki_surprisal_relative")
+# for_gs <- gs_title("wiki_surprisal_relative")
 
 df <- read_csv("wiki_df.csv")
 
-df_tokens <- extract_wiki_tokens(6, 50, df) 
+df_tokens <- extract_wiki_tokens(6, 50, df)
 
 pos_list <- df_tokens %>%
   group_by(length) %>%
@@ -92,9 +92,20 @@ pos_list <- df_tokens %>%
   mutate(length = as.numeric(length))
 
 slopes <- relative_slopes(df_tokens, pos_list)
-wkpd <- gs_read(for_gs)
-nr <- nrow(wkpd)
 
-lang_slopes <- c(LANGUAGE, slopes$estimate)
-gs_edit_cells(for_gs, ws = "Sheet1", anchor = paste("A", nr + 2, sep=""), input = lang_slopes, byrow = TRUE)
-#system("rm wiki_df.csv")
+lang_slopes <- c(LANGUAGE, slopes$estimate) %>% 
+  t() %>%
+  data.frame()
+
+all_data <- read.csv(file = "relative_unigrams.csv") %>%
+  select(-X)
+names(lang_slopes) <- names(all_data)
+all_data <- all_data %>%
+  rbind(lang_slopes)
+write.csv(all_data, file = "relative_unigrams.csv")
+# wkpd <- gs_read(for_gs)
[…]

### 2019-05-14 entropy/childes/non-english_entropies.Rmd
reorganization; working on wiki relative unigram surprisal

--- 2018-09-14 non-english_entropies.Rmd
@@
     group_by(utterance_id) %>%
-    mutate(word_order = 1:n())
+    mutate(word_order = 1:n()) %>%
+    filter(word_order <= sen_length)
 

### 2019-05-14 surprisal/wiki_surprisal.Rmd
reorganization; working on wiki relative unigram surprisal

--- 2019-03-04 wiki_entropy/wiki_clustering.Rmd
@@
 ---
-title: Wikipedia Text Mining and Entropy Analysis
+title: Wikipedia Text Mining and Surprisal Analysis
 author: Josef Klafka and Dan Yurovsky
@@
 theme_set(theme_classic(base_size = 16))
-```
-
-```{r imputation and MCA}
-wiki_gs <- gs_title("Wikipedia_absolute")
-wkpd <- gs_read(wiki_gs)
-LANGUAGES <- wkpd$Language
-
-FEATURES <- c(1:144)
-for (i in c(1:length(FEATURES))) {
-  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
-} #get all WALS features in correct syntax
-
-wf <- wals.feature(FEATURES) 
-fair_wf <- wf[rowSums(!is.na(wf)) > 50,] %>%
-  filter(language %in% LANGUAGES) %>%
-  select(-wals.code, -latitude, -longitude, -glottocode) %>%
-  as_data_frame()
-
-# Phonology `1A`:`19A`
-# Morphology `20A`:`29A`
-# Nominal Categories `30A`:`57A`
-# Nominal Syntax `58A`:64A`
-# Verbal syntax `65A`:`80A`
-# Word order `81A`:`97A`, `143A`, `144A`
-# Simple clauses `98A`:`120A`
-# Complex sentences `121A`:`128A`
-# Lexicon `129A`:`138A`
-# Sign languages `139A`:`140A`
-
-wa <- fair_wf %>% 
-  filter(language %in% LANGUAGES) %>%
-  select(`1A`:`19A`, language) %>%
-  group_by(language) %>%
-  slice(1) %>%
-  gather(feature, value, `1A`:`19A`) %>%
-  group_by(feature) %>%
-  mutate(value = factor(value)) %>%
-  ungroup() %>% 
-  spread(feature, value) %>% 
-  column_to_rownames("language") %>%
-  sapply(as.factor) %>%
-  as_data_frame() %>%
-  MIMCA(ncp = 1, maxiter = 100000, threshold = 1e-1)
-
-wa <- as.data.frame(wa[[1]][[99]]) 
-wa$language <- unique(phon_wf$language)
 ```
@@
 
+cosines <- wkpd %>%
+  column_to_rownames("Language") %>%
+  t() %>%
+  cosine() %>%
+  as_data_frame(rownames = "language1") %>%
+  gather(language2, cosine, -language1)
+
 models <- wf_pairwise %>%
@@
 ```
-
-```{r hierarchical clustering}
-NUM_CLUSTERS <- 10
-
-##read in googlesheet
-wiki_gs <- gs_title("Wikipedia_absolute")
-wkpd <- gs_read(wiki_gs) 
-LANGUAGES <- wkpd$Language
-
-wkpd <- wkpd %>% 
-  as.data.frame() %>%
-  column_to_rownames(var = "Language")
-
-##do clustering
-wk_clust_tree <- wkpd %>% 
-              dist() %>%
-              hclust("ward.D2")
-
-wk_clust_tree %>% plot(hang = -1)
-wk_clusters <- wk_clust_tree %>%
-              cutree(k = NUM_CLUSTERS)
-
-wk_clusters %>% 
-  t() %>%
-  as_data_frame() %>% 
-  t() %>%
-  as_data_frame(rownames = "language") %>%
-  rename(cluster = V1) %>%
-  arrange(cluster) %>% 
-  View()
-```
-
-```{r}
-wiki_gs <- gs_title("Wikipedia_relative5")
-wkpd <- gs_read(wiki_gs) 
-LANGUAGES <- wkpd$Language
-
-FEATURES <- c(1:144)
-for (i in c(1:length(FEATURES))) {
-  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
-} #get all WALS features in correct syntax
-
-wf <- wals.feature(FEATURES)
-
-wf_df <- wf %>%
-  as_data_frame() %>%
-  filter(language %in% LANGUAGES) %>%
-  group_by(language) %>%
-  slice(1) %>%
-  gather(feature, value, `1A`:`144A`) %>%
-  group_by(feature) %>%
-  mutate(value = factor(value))
-
-cosines <- wkpd %>%
-  column_to_rownames("Language") %>%
-  t() %>%
-  cosine() %>%
-  as_data_frame(rownames = "language1") %>%
-  gather(language2, cosine, -language1)
-
-```
-
-```{r cogsci glm subset}
-
-wf_many_langs <- wf_df %>%
-  group_by(feature) %>%
-  summarise(n = sum(!is.na(value))) %>%
-  filter(n > 60)
-
-wf_sub_langs <- wf_df %>%
-  filter(feature %in% wf_many_langs$feature) %>%
-  group_by(language) %>%
-  summarise(keep = sum(is.na(value)) == 0) %>%
-  filter(keep)
-
-wf_sub <- wf_df %>%
-  filter(feature %in% wf_many_langs$feature,
-         language %in% wf_sub_langs$language) %>%
-  select(language, feature, value) %>%
-  mutate(value = as.numeric(as.factor(value)))
-
-wf_pairwise <- expand.grid(language1 = wf_sub$language,
-                           language2 = wf_sub$language) %>%
-  left_join(wf_sub, by = c("language1" = "language")) %>%
-  rename(value1 = value) %>%
-  left_join(wf_sub, by = c("language2" = "language", "feature")) %>%
-  rename(value2 = value) %>%
-  mutate(same = value1 == value2) %>%
-  select(-value1, -value2) %>%
-  left_join(cosines, by = c("language1", "language2"))
-
-models <- wf_pairwise %>%
-  group_by(feature) %>%
-  nest() %>%
-  mutate(model = map(data, ~glm(same ~ cosine, 
-                                family = "binomial", data = .)))
-
-models %>%
-  mutate(coeffs = map(model, tidy),
-         rs = map(model, )) %>%
-  select(-data, -model) %>%
-  unnest() %>%
-  filter(term == "cosine") %>%
-  arrange(desc(statistic))
-```

### 2019-05-14 entropy/ngrams/get_wiki_df.py

@@ -1,0 +1,96 @@
+import regex as re
+import numpy as np
+import pandas as pd
+import csv
+import sys
+import json
+import subprocess
+import urllib.request
+
+
+with open("language_dict.json", 'r') as f:
+	file = f.read()
+LANG_DICT = json.loads(file) #maps languages to their wikipedia url prefixes
+
+
+def extract_texts(upper_bound):
+	'''
+	Given an upper_bound on file numbers, extracts the raw text from the files
+	in the text/AA folder. That folder is given as output by the
+	cirrus_extract program.
+	'''
+	rs = ""
+	for i in range(upper_bound):
+		if i < 10:
+			number = "0" + str(i)
+		else:
+			number = str(i)
+		f = open("text/AA/wiki_" + number, 'r')
+		text = f.read()
+		rs += text
+	return rs
+
+
+def get_sen_dict(text):
+	'''
+	Given the output of extract_texts as a string, constructs a
+	dictionary mapping each sentence in the string to its length in words.
+	'''
+	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:–'&*#/—“»]|\d+|\(|\)|\[|\]", \
+		"", text)
+	text = re.sub('"', '', text)
+	#just the words from the articles
+	text = text.replace('\n', " ")
+	sens = re.split("[.!?]", text.lower())
+	sen_dict = {}
+	for sen in sens:
+		length = len(re.findall("\w+", sen))
+		if length > 0:
+			sen_dict[sen] = length
+	return sen_dict
+
+
+def get_sen_df(text):
+	'''
+	Given the output of extract_texts as a string, returns a Pandas dataframe
+	of the output of get_sen_dict.
+	'''
+	sd = get_sen_dict(text)
+	s = pd.Series(sd, name = "length")
+	del sd
+	s.index.name = "words"
+	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence
+					 	  #and the sentence length respectively
+	return df
+
+
+def main(lang_name):
+	'''
+	Given the English name of a language, constructs a dataframe with the
+	sentences of a random subset of articles in that language from Wikipedia,
+	and stores the dataframe in a csv in the working directory.
+	'''
+	lang_prefix = LANG_DICT[lang_name]
+	url = "https://dumps.wikimedia.org/other/cirrussearch/20190318/" \
+		+ lang_prefix + "wiki-20190318-cirrussearch-content.json.gz"
[…]

### 2019-05-14 entropy/ngrams/wiki_extract/cirrus_extract.py

@@ -1,0 +1,245 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+#
+# =============================================================================
+#  Version: 1.00 (December 15, 2015)
+#  Author: Giuseppe Attardi (attardi@di.unipi.it), University of Pisa
+#
+# =============================================================================
+#  Copyright (c) 2015. Giuseppe Attardi (attardi@di.unipi.it).
+# =============================================================================
+#  This file is part of Tanl.
+#
+#  Tanl is free software; you can redistribute it and/or modify it
+#  under the terms of the GNU General Public License, version 3,
+#  as published by the Free Software Foundation.
+#
+#  Tanl is distributed in the hope that it will be useful,
+#  but WITHOUT ANY WARRANTY; without even the implied warranty of
+#  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+#  GNU General Public License for more details.
+#
+#  You should have received a copy of the GNU General Public License
+#  along with this program.  If not, see <http://www.gnu.org/licenses/>.
+# =============================================================================
+
+"""Wikipedia Cirrus Extractor:
+Extracts and cleans text from a Wikipedia Cirrus dump and stores output in a
+number of files of similar size in a given directory.
+Each file will contain several documents in the format:
+
+    <doc id="" url="" title="">
+        ...
+        </doc>
+
+"""
+
+import sys, os.path, time
+import re
+import json
+import argparse
+import bz2
+import gzip
+import logging
+
+# Program version
+version = '1.00'
+
+urlbase = 'http://it.wikipedia.org/'
+
+# ----------------------------------------------------------------------
+
+class NextFile(object):
+    """
+    Synchronous generation of next available file name.
+    """
+
+    filesPerDir = 100
+
+    def __init__(self, path_name):
+        self.path_name = path_name
+        self.dir_index = -1
+        self.file_index = -1
+
+    def next(self):
+        self.file_index = (self.file_index + 1) % NextFile.filesPerDir
+        if self.file_index == 0:
+            self.dir_index += 1
+        dirname = self._dirname()
+        if not os.path.isdir(dirname):
+            os.makedirs(dirname)
+        return self._filepath()
+
+    def _dirname(self):
+        char1 = self.dir_index % 26
+        char2 = self.dir_index / 26 % 26
+        return os.path.join(self.path_name, '%c%c' % (ord('A') + int(char2), ord('A') + int(char1)))
+
+    def _filepath(self):
[…]

### 2019-05-14 entropy/ngrams/wiki_relative_ngrams.R

@@ -1,0 +1,113 @@
+library(tidyverse)
+library(directlabels)
+library(tidytext)
+library(entropy)
+library(tidyboot)
+library(tokenizers)
+library(googlesheets)
+library(stringr)
+### Begin Functions
+
+extract_wiki_tokens <- function(min_length, max_length, utterances, N) {
+
+  tokens <- utterances %>%
+    mutate(length = str_count(words, pattern = "[ +]+") + 1) %>%
+    filter(length >= min_length) %>%
+    filter(length <= max_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, words, token = stringr::str_split, pattern = "[ +]+") %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n())
+
+
+  lags <- expand.grid(lag_n = 1:(N-1),
+                      utterance_id = unique(tokens$utterance_id))
+
+  tokens %>%
+    left_join(lags, by = "utterance_id") %>%
+    arrange(utterance_id, lag_n, word_order) %>%
+    mutate(lag = lag(word, n = first(lag_n))) %>%
+    drop_na() %>%
+    mutate(gram = str_c(lag, word, sep = " ")) %>%
+    rename(gram_order = word_order) %>%
+    group_by(length, gram_order, gram) %>%
+    summarise(n = n())
+
+}
+
+get_quantiles <- function(df, num_sections) {
+
+  quantile(df$gram_order, probs = seq(0, 1, 1/num_sections)) %>%
+    round() %>%
+    as_data_frame() %>%
+    rename(gram_order = value)
+}
+
+relative_slopes <- function(df, pos_list) {
+
+  get_slope <- function(x) {
+
+    selected_pos <- pos_list %>%
+      group_by(length) %>%
+      slice(x:(x+1)) %>%
+      group_by(length) %>%
+      summarise(min = min(gram_order),
+                max = max(gram_order)) %>%
+      split(.$length) %>%
+      map(~seq(.x$min, .x$max, 1) %>% as_data_frame) %>%
+      bind_rows(.id = "length") %>%
+      rename(gram_order = value) %>%
+      mutate(length = as.numeric(length))
+
+    df_tokens %>%
[…]

### 2019-05-14 entropy/wiki_entropy/wiki_relative.R

@@ -1,0 +1,98 @@
+library(knitr)
+library(tidyverse)
+library(directlabels)
+library(dplyr)
+library(tidytext)
+library(entropy)
+library(tidyboot)
+library(tokenizers)
+library(googlesheets)
+
+### Begin Functions
+
+extract_wiki_tokens <- function(min_length, max_length, utterances) {
+
+  tokens <- utterances %>%
+    mutate(gloss = str_trim(gloss)) %>%
+    mutate(length = str_count(gloss, pattern = " +") + 1) %>%
+    filter(length >= min_length) %>%
+    filter(length <= max_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(length, word_order, word) %>%
+    summarise(n = n())
+}
+
+get_quantiles <- function(df, num_sections) {
+  quantile(df$word_order, probs = seq(0, 1, 1/num_sections)) %>%
+    round() %>%
+    as_data_frame() %>%
+    rename(word_order = value)
+}
+
+relative_slopes <- function(df, pos_list) {
+
+  get_slope <- function(x) {
+
+    selected_pos <- pos_list %>%
+      group_by(length) %>%
+      slice(x:(x+1)) %>%
+      group_by(length) %>%
+      summarise(min = min(word_order),
+                max = max(word_order)) %>%
+      split(.$length) %>%
+      map(~seq(.x$min, .x$max, 1) %>% as_data_frame) %>%
+      bind_rows(.id = "length") %>%
+      rename(word_order = value) %>%
+      mutate(length = as.numeric(length))
+
+    df %>%
+      inner_join(selected_pos, by = c("length", "word_order")) %>%
+      group_by(length, word_order) %>%
+      summarise(entropy = entropy(n, unit = "log2")) %>%
+      lm(entropy ~ length + I(length^2) + word_order, data  = .) %>%
+      tidy() %>%
+      filter(term == "word_order")
+  }
+
+  divides <- pos_list %>%
+    group_by(length) %>%
+    summarise(n = n()) %>%
+    summarise(n = mean(n)) %>%
+    pull()
+
[…]

### 2019-05-14 surprisal/wiki_relative_surprisal.R

@@ -1,0 +1,112 @@
+library(tidyverse)
+library(tidytext)
+library(directlabels)
+# library(googlesheets)
+
+extract_wiki_tokens <- function(min_length, max_length, utterances) {
+
+  tokens <- utterances %>%
+    mutate(gloss = str_trim(gloss)) %>%
+    mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
+    filter(length >= min_length) %>%
+    filter(length <= max_length) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n()) %>%
+    group_by(length, word_order, word) %>%
+    summarise(n = n()) %>%
+    ungroup()
+
+  sups <- tokens %>%
+    mutate(p = n/sum(n)) %>%
+    mutate(s = -log(p)) %>%
+    select(word, s)
+
+  tokens %>%
+    left_join(sups)
+}
+
+get_quantiles <- function(df, num_sections) {
+  quantile(df$word_order, probs = seq(0, 1, 1/num_sections)) %>%
+    round() %>%
+    as_data_frame() %>%
+    rename(word_order = value)
+}
+
+relative_slopes <- function(df_tokens, pos_list) {
+
+  get_slope <- function(x) {
+
+    selected_pos <- pos_list %>%
+      group_by(length) %>%
+      slice(x:(x+1)) %>%
+      group_by(length) %>%
+      summarise(min = min(word_order),
+                max = max(word_order)) %>%
+      split(.$length) %>%
+      map(~seq(.x$min, .x$max, 1) %>% as_data_frame) %>%
+      bind_rows(.id = "length") %>%
+      rename(word_order = value) %>%
+      mutate(length = as.numeric(length))
+
+    df_tokens %>%
+      inner_join(selected_pos, by = c("length", "word_order")) %>%
+      group_by(length, word_order) %>%
+      summarise(s = mean(s)) %>%
+      lm(s ~ length + I(length^2) + word_order, data  = .) %>%
+      tidy() %>%
+      filter(term == "word_order")
+  }
+
+  divides <- pos_list %>%
[…]

### 2019-05-14 surprisal/wiki_script.sh

@@ -1,17 +1,14 @@
 #!/bin/bash
 
-declare -a arr=("Basque" "Dutch" "Alemannic" "Wolof" "Bosnian" "Hindi" "Danish"
-	"Russian" "Urdu" "German" "Swedish" "Afrikaans" "Anglo-Saxon" "Icelandic"
-	"Belarusian" "Bulgarian" "Croatian" "Macedonian" "Serbo-Croatian" "Ukrainian"
-	"Vietnamese" "Pangasinan" "Kapampangan" "Khmer" "Japanese" "Javanese"
-	"Indonesian" "Polish" "Malay" "Sundanese" "Hakka" "Turkish" "Kazakh"
-	"Kirghiz" "Turkmen" "Hebrew" "Persian" "Arabic" "Tajik" "Latin" "French"
-	"Galician" "Romanian" "Norman" "Corsican" "Maori" "Nepali" "Mongolian"
-	"Estonian" "Finnish" "Hungarian" "Bengali" "Waray" "Welsh" "Armenian"
-	"Korean" "Greek" "Catalan" "Walloon" "Yiddish" "Maltese" "Romansh"
-	"Banyumasan" "Võro" "Samogitian" "Chavacano")
+declare -a arr=("Amharic" "Armenian" "Bashkir" "Bhojpuri" "Breton" "Burmese"
+	"Catalan" "Chuvash" "Czech" "Danish" "Estonian" "Finnish" "French" "Georgian"
+	"German" "Gujarati" "Hausa" "Hungarian" "Indonesian" "Irish" "Japanese" "Kabyle"
+	"Kannada" "Karachay-Balkar" "Korean" "Lao" "Latvian" "Lezgian" "Lithuanian"
+	"Maithili" "Malayalam" "Maori" "Marathi" "Ossetian" "Polish" "Romanian"
+	"Romansh" "Sakha" "Samoan" "Somali" "Sundanese" "Swahili" "Tajik" "Tamil"
+	"Tatar" "Telugu" "Tetum" "Thai" "Tulu" "Turkish" "Turkmen" "Welsh" "Yoruba")
 
-declare -a newarr=("Basque" "Dutch" "Alemannic" "Wolof" "Bosnian" "Hindi"
+declare -a fullarr=("Basque" "Dutch" "Alemannic" "Wolof" "Bosnian" "Hindi"
[…]

### 2019-05-14 switchboard/process_switchboard.R

@@ -1,0 +1,15 @@
+library(tidyverse)
+library(stringr)
+library(feather)
+
+files <- list.files("switchboard", "*.txt", full.names = T)
+
+
+
+txt <- map(files, read_lines) %>%
+  unlist() %>%
+  as_tibble() %>%
+  mutate(value = str_trim(value)) %>%
+  filter(value != "")
+
+write_feather(txt, "switchboard/switchboard.feather")

### 2019-05-17 surprisal/childes_surprisal.Rmd
tiny corpus for midwest cogsci

@@
 ```{r old unigrams}
-adult_utterances <- shiro_utterances %>%
-  filter(speaker_role == "Target_Child") %>% 
+adult_utterances <- prov_utterances %>%
+  filter(speaker_role != "Target_Child") %>% 
   mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
@@
   geom_smooth(se = F)
+```
+
+```{r redoing bigrams}
+adult_bigrams_prep <- adult_utterances %>% 
+  group_by(utterance_id) %>%
+  mutate(lag_word = lag(word)) %>%
+  group_by(lag_word, word) %>%
+  count() %>%
+  filter(!is.na(lag_word)) %>%
+  ungroup() %>%
+  left_join(adult_unigrams, by = c("lag_word" = "word")) %>%
+  mutate(cond_p = n.x / n.y) %>%
+  select(lag_word, word, cond_p)
+
+adult_bigrams <- adult_utterances %>%
+  group_by(utterance_id) %>%
+  mutate(lag_word = lag(word)) %>%
+  ungroup() %>%
+  left_join(adult_unigrams) %>%
+  left_join(adult_bigrams_prep) %>%
+  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
+  group_by(length, word_order) %>%
+  summarise(mean_s = mean(s))
+
+adult_bigrams %>% 
+  filter(length %in% c(5, 7, 9)) %>% 
+  ggplot(aes(x = word_order, y = mean_s)) + 
+  facet_wrap(~length) + 
+  geom_point() + 
+  geom_line()
+
 ```

### 2019-05-17 surprisal/wiki_surprisal.Rmd
tiny corpus for midwest cogsci

@@
 ```{r word order}
-wiki_gs <- gs_title("Wikipedia_absolute")
-wkpd <- gs_read(wiki_gs) 
+wkpd <- read.csv(here("relative_unigrams.csv"))
 LANGUAGES <- wkpd$Language
@@
 
+cosines <- wkpd %>%
+  column_to_rownames("Language") %>%
+  t() %>%
+  cosine() %>%
+  as_data_frame(rownames = "language1") %>%
+  gather(language2, cosine, -language1)
+
 wf_pairwise <- expand.grid(language1 = wo_sub$language,
@@
 
-cosines <- wkpd %>%
-  column_to_rownames("Language") %>%
-  t() %>%
-  cosine() %>%
-  as_data_frame(rownames = "language1") %>%
-  gather(language2, cosine, -language1)
-
 models <- wf_pairwise %>%
@@
 
-models %>%
+feature_names <- read_feather(here("feature_names.feather"))
+
+model_df <- models %>%
   mutate(coeffs = map(model, tidy)) %>%
@@
   arrange(desc(statistic)) %>%
-  write_feather(here("word_order_glm_absolute.feather"))
+  left_join(feature_names)
 ```

### 2019-05-17 surprisal/tiny_corpus.R

@@ -1,0 +1,79 @@
+library(tidyverse)
+library(tidytext)
+
+# a1 <- c("a", "cup", "broke")
+# a2 <- c("the", "cup", "broke")
+# a3 <- c("a", "cup", "smashed")
+# a4 <- c("the", "glass", "broke")
+# a5 <- c("a", "glass", "broke")
+# tiny_corpus <- as.data.frame(rbind(a1, a2, a3, a4, a5))
+# row.names(tiny_corpus) <- 1:5
+# names(tiny_corpus) <- c("first", "second", "third")
+
+a1 <- c("a cup broke")
+a2 <- c("the cup broke")
+a3 <- c("a cup smashed")
+a4 <- c("the glass broke")
+a5 <- c("a glass broke")
+tiny_corpus <- as.data.frame(rbind(a1, a2, a3, a4, a5))
+names(tiny_corpus) <- c("text")
+tiny_corpus$text <- as.character(tiny_corpus$text)
+row.names(tiny_corpus) <- 1:5
+tiny_corpus <- tiny_corpus %>%
+  unnest_tokens(word, text, drop=F)
+
+tiny_surprisals <- tiny_corpus %>% 
+  group_by(word) %>% 
+  count() %>% 
+  ungroup() %>% 
+  mutate(p = n / sum(n)) %>% 
+  mutate(s = -log(p)) 
+
+tiny_means <- tiny_corpus %>% 
+  group_by(text) %>%
+  mutate(word_order = 1:n()) %>%
+  ungroup() %>%
+  left_join(tiny_surprisals) %>% 
+  group_by(word_order) %>%
+  summarise(mean_s = mean(s)) 
+
+tiny_means %>% 
+  ggplot(aes(x = word_order, y = mean_s)) + 
+    geom_point() + 
+    geom_line() +
+    xlab("Word order") + 
+    ylab("Mean surprisal") + 
+    ylim(0, 5)
+
+
+
+tiny_bigrams_prep <- tiny_corpus %>% 
+  group_by(text) %>%
+  mutate(lag_word = lag(word)) %>%
+  group_by(lag_word, word) %>%
+  count() %>%
+  filter(!is.na(lag_word)) %>%
+  ungroup() %>%
+  mutate(p = n/sum(n)) %>%
+  left_join(tiny_surprisals, by = c("lag_word" = "word")) %>%
+  mutate(cond_p = n.x / n.y) %>%
[…]

### 2019-05-22 surprisal/childes_surprisal.Rmd
childes unigrams

@@
 ```{r old unigrams}
+child_utterances <- prov_utterances %>%
+  filter(speaker_role == "Target_Child") %>% 
+  mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
+  mutate(utterance_id = 1:n()) %>%
+  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
+  group_by(utterance_id) %>%
+  mutate(word_order = 1:n())
+
+child_unigrams <- child_utterances %>%
+  group_by(word) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(p = n / sum(n))
+
+child_surprisals <- child_utterances %>%
+  left_join(child_unigrams) %>%
+  mutate(s = -log(p)) %>%
+  group_by(length, word_order) %>%
+  summarise(s = mean(s))
+
 adult_utterances <- prov_utterances %>%
   filter(speaker_role != "Target_Child") %>% 
-  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
+  mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
   mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
+  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
   group_by(utterance_id) %>%
@@
 
-adult_surprisals %>%
-  filter(length %in% c(5, 7, 9)) %>%
-  ggplot(aes(x = word_order, y = s)) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_smooth(se = F)
+surprise <- bind_rows(child_surprisals %>%
+  mutate(speaker_role = "Child") %>%
+  bind_rows(mutate(adult_surprisals, speaker_role = "Adult")) %>%
+  mutate(Language = "English"))
+
+surprisals <- surprisals %>% 
+  bind_rows(surprise)
 ```
@@
   filter(length %in% c(5, 7, 9)) %>% 
-  ggplot(aes(x = word_order, y = mean_s)) + 
-  facet_wrap(~length) + 
+  ggplot(aes(x = word_order, y = mean_s, color = speaker, label = speaker)) + 
+  xlab("Word position") +
+  ylab("Mean surprisal") +
   geom_point() + 
-  geom_line()
+  geom_line() + 
+  theme_classic(base_size = 10) + 
+  scale_color_ptol() +
+  theme(legend.position = "nones")
 
+label_data <- data_frame(length = 9, Language = "Spanish", word_order = 4,
+                         speaker_role = c("Adult", "Child"), s = c(5,6.25))
+
+surprisals %>%  
+  filter(length == 9) %>%
+  ggplot(aes(x = word_order, y = s,
+                        color = speaker_role,
+                        label = speaker_role)) +
+    facet_grid(Language ~ length, scales = "free") + 
+    xlab("Word position") +
+    ylab("Mean surprisal") +
+    scale_x_continuous(breaks = 1:9) + 
+    geom_line(size = 1) +
+    geom_point() + 
+    theme_classic(base_size = 10) + 
+    scale_color_ptol() +
+    theme(legend.position = "nones") + 
+    geom_text(data = label_data)
 ```
@@
   ungroup() %>%
-  mutate(joint_p = n / sum(n)) %>%
-  select(-n) %>%
   left_join(adult_unigrams, by = c("lag_word" = "word")) %>%
-  mutate(cond_p = joint_p / p) %>%
+  mutate(cond_p = n.x/n.y) %>%
   select(-n, -p)
@@
   ungroup() %>%
-  mutate(tri_joint_p = n / sum(n)) %>%
-  select(-n) %>%
   left_join(adult_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%

### 2019-05-28 surprisal/MacWhinney poster/MacWhinney poster.Rmd
posterdown not working yet

---
# PLEASE SEE THE README for in depth description github.com/brentthorne/posterdown
poster_height: "32in"
poster_width: "46in"
font_family: 'Rasa'
#ESSENTIALS
title: A Better Reproducible Poster Title
author:
  - name: W. Brent Thorne
    affil: 1,*
    main: true
    orcid: '0000-0002-1099-3857'
    twitter: brentthorne18 
    email: bthorne2@brocku.ca
  - name: Another G. Contributor
    affil: 2
    email: ag.con@posterdown.net
  - name: Person Three
    affil: 3
affiliation:
  - num: 1
    address: Department of Earth Science, Brock University
  - num: 2
    address: Department of Graphics and Layouts, University of Posters; Canada
  - num: 3
    address: Another Institute of a place where work gets done, Earth, Milky Way
#STYLE & FORMATTING
title_textsize: "125pt"
author_textsize: "1.17em"
authorextra_textsize: "35px"
affiliation_textsize: "25px"
affiliation_textcol: '#00000060'
caption_fontsize: "20pt"
#Middle of the poster
middle_fontfamily: "Special Elite"
middle_textcol: "#FFFFFF90"
middle_fontsize: "170px"
main_findings:
  - "Make **better posters** with <br> RMarkdown + **posterdown**."
  - "Transition from **poster** to **manuscript** with ease!"
logoleft_name: "Figures/qr-code-black.png"
logoright_name: "Figures/posterdownlogo.png"
#---POSTER BODY OPTIONS---#
primary_colour: '#0b4545'
secondary_colour: '#008080' 
accent_colour: "#cc0000"
body_bgcol: "#FFFFFF"
body_textsize: "45px"
body_textcol: "#000000"
#--Standard Options--#
output: 
  posterdown::posterdown_betterland:
    self_contained: false
    pandoc_args: --mathjax
    highlight: espresso
    number_sections: false
link-citations: true
bibliography: MyBib.bib
---

```{r, include=FALSE}
knitr::opts_chunk$set(results = 'asis', echo = FALSE, comment = FALSE,  warning = FALSE,tidy = FALSE, message = FALSE, fig.align = 'center')
options(knitr.table.format = "html") 
```

# Introduction

This is the `posterdown_betterland` template for the {posterdown} package! I was inspired by the twitter thread of [Mike Morrison](https://mobile.twitter.com/mikemorrison/status/1110191245035479041) and wanted to apply the `#betterposter` concept to the reproducible (yet simple to use) functionality of the {posterdown} package [@thorneposterdown2019]. If you're not an R user don't sweat as you do **NOT** need to use it at all! Feel free to use only the Markdown functionality of this package :)

# Methods

1. Install R and `posterdown`, see [github](github.com/brentthorne/posterdown)
2. Open the `posterdown_betterland` template
3. Make all your poster dreams come true!

# Results

```{r, out.width='90%', fig.height=5.5, fig.cap='Using {ggplot} and {patchwork} to generate a layout of multiple plots in one figure [@patchwork2017].'}

library(tidyverse)
library(patchwork)
library(ggthemes)

theme_set(theme_tufte() + theme(legend.position = "none"))

base <- ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, colour = Species))

p_point <- base + geom_point()
p_line  <- base + geom_line() 
p_area  <- base + geom_area()
p_box1  <- ggplot(iris) + geom_boxplot(aes(x = Species, y = Sepal.Length, fill = Species))
p_box2  <- ggplot(iris) + geom_boxplot(aes(x = Species, y = Sepal.Width, fill = Species))

p_line +
  {p_point + p_box1} +
  plot_layout(ncol = 1) + plot_annotation(tag_levels = "a", tag_prefix = "(",tag_suffix = ")")

```

## More Figures and Tables

```{r, mytable, fig.align='center', out.width='90%'}

library(knitr)
library(kableExtra)

dt <- iris[1:5, 1:5]

kable(dt, caption = 'Here is a caption for the table made with the {kableExtra} package [@kableExtra2019].',align = 'c', col.names = c("Sepal W","Sepal L","Pedal W","Pedal L","Species")) %>% 
  kable_styling(full_width = T) %>% 
  row_spec(0, color = "#fff", background = "#0b454580")
```

```{r, echo=TRUE}
library(ggplot2)
library(ggspatial)

load_longlake_data()

map <- ggplot() + 
  theme_pander() +
  layer_spatial(longlake_depthdf,
                aes(colour = DEPTH)) +
  theme(legend.position = "bottom")
```

```{r, mymapfig, out.width='70%', cache=FALSE, fig.cap='Map of Long Lake eample from the [ggspatial](https://github.com/paleolimbot/ggspatial) package [@ggspatial].',strip.white= TRUE, comment=FALSE}

ggsave(map,filename = "Figures/map.svg", width = 3.5, height = 3.5, device = 'svg', bg = 'transparent')

knitr::include_graphics("Figures/map.svg")
```

A **BIG** thank you to [Romain Leasur](https://twitter.com/RLesur) and [Yihui Xie](https://twitter.com/xieyihui) for their wonderful work on [{pagedown}](https://github.com/rstudio/pagedown) which had made this poster possible [@pagedown]!

# References

### 2019-05-28 surprisal/MacPoster/MacPoster.Rmd
changes

---
# PLEASE SEE THE README for in depth description github.com/brentthorne/posterdown
poster_height: "48in"
poster_width: "36in"
font_family: 'Rasa'
#ESSENTIALS
title: '**Parents and children structure their utterances for efficient communication**'
author:
  - name: '**Josef Klafka**'
    affil: 1,*
    main: true
    orcid: '0000-0002-1099-3857'
    twitter: josefklafka
    email: jklafka@uchicago.edu
  - name: Daniel Yurovsky
    affil: 1
    main: false
    email: 'dyurovsky@uchicago.edu'
affiliation:
  - num: 1
    address: Department of Psychology, University of Chicago
#STYLE & FORMATTING
title_textsize: "125pt"
author_textsize: "1.17em"
authorextra_textsize: "35px"
affiliation_textsize: "25px"
affiliation_textcol: '#00000080'
caption_fontsize: "20pt"
#Middle of the poster
middle_fontfamily: "Special Elite"
middle_textcol: "#FFFFFF"
middle:textalpha: 90
middle_fontsize: "150px"
main_findings:
  - "Information language children parents incremental-processing"
logoleft_name: "Figures/callab_logo.png"
logocenter_name: "Figures/qr-code-black.png"
logoright_name: "Figures/logo.rgb.maroon.jpg"
#---POSTER BODY OPTIONS---#
primary_colour: '#0b4545'
secondary_colour: '#008080'
accent_colour: "#cc0000"
body_bgcol: "#ffffff"
body_textsize: "45px"
body_textcol: "#000000"
reference_textsize: "23px"
column_numbers: 1
columnline_col: "black"
columnline_style: "solid"
column_margins: "10px"
#--Standard Options--#
output: 
  posterdown::posterdown_betterport:
    self_contained: false
    pandoc_args: --mathjax
    highlight: espresso
    number_sections: true
bibliography: MyBib.bib
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  results = 'asis',
  echo = FALSE,
  comment = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = 'center'
)
```

# Introduction

Welcome to `posterdown` ! This is my attempt to provide a semi-smooth workflow for those who wish to take their R Markdown skills to the conference world. Most features from R Markdown are available in this package such as Markdown section notation, figure captioning, and even citations like this one [@rmarkdown]. The rest of this example poster will show how you can insert typical conference poster features into your own document.

## Objectives

1. Easy to use reproducible poster design. 
2. Integration with R Markdown.
3. Easy transition from `posterdown` to `thesisdown` or `rticles` [@rticles; @thesisdown].

# Methods

This package uses the same workflow approach as the R Markdown you know and love. Basically it goes from RMarkdown > Knitr > Markdown > Pandoc > HTML/CSS > PDF. You can even use the bibliography the same way [@turnerControlsWaterBalance2014].

# Results

Usually you want to have a nice table displaying some important results that you have calculated. In `posterdown` this is as easy as using the `kable` table formatting you are probably use to as per typical R Markdown formatting.

I suggesting checking out the `kableExtra` package and its in depth documentation on customizing these tables found [here](https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf) [@kableExtra2019]. You can reference tables like so: Table \@ref(tab:mytable). Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam placerat augue at velit tincidunt semper. Donec elementum porta posuere. Nullam interdum, odio at tincidunt feugiat, turpis nisi blandit eros, eu posuere risus felis non quam. Nam eget lorem odio. Duis et aliquet orci. Phasellus nec viverra est.

```{r standard-plot, out.width='80%', fig.align='center', fig.cap='Great figure!', fig.height=10}
plot(mtcars[1:2])
```

# Next Steps: More random text

Aliquam sed faucibus risus, quis efficitur erat. Vestibulum semper mauris quis tempus eleifend. Aliquam sagittis dictum ipsum, quis viverra ligula eleifend ut. Curabitur sagittis vitae arcu eget faucibus. In non elementum felis. Duis et aliquam nunc. Nunc pulvinar sapien nunc, vel pretium nisi efficitur in. Fusce fringilla maximus leo et maximus. Fusce at ligula laoreet, iaculis mi at, auctor odio. Praesent sed elementum justo. Aenean consectetur risus rhoncus tincidunt efficitur. Praesent dictum mauris at diam maximus maximus [@thorneposterdown2019].

# Conclusion

Try `posterdown` out! Hopefully you like it!

# References

### 2019-05-28 surprisal/MacWhinney poster/MacWhinney_Poster.Rmd
changes

---
# PLEASE SEE THE README for in depth description github.com/brentthorne/posterdown
#
#---POSTER SIZE & DEFAULT FONT---#
poster_height: "32in" # height in inches of poster
poster_width: "40in" # width in inches of poster
font_family: "palatino" # choose from typical latex fonts (example: "palatino")
font_size: "30pt" #please see github.com/brentthorne/posterdown for compatible options.

#---TITLE BOX OPTIONS---#
#ESSENTIALS
title: '\fontfamily{phv}\selectfont Parents and children in diverse languages structure their utterances to support efficient communication'
author: "Josef Klafka Daniel Yurovsky"
affiliation: "^1^Department of Psychology, University of Chicago"
#STYLE & FORMATTING
titlebox_bgcol: "008080"  #Colour of the Title Box background
titlebox_bordercol: "0b4545" #Colour of the title Box border.
titlebox_shape: "all"
titlebox_borderwidth: "1cm"
title_textcol: "ffffff" #colour of title text
author_textcol: "0b4545" # Colour of author text
affiliation_textcol: "FFFFFF" # Colour of affiliation text
title_textsize: "Huge"         # Poster title fontsize
author_textsize: "Large"       # Author list font size
affiliation_textsize: "large"  # Affiliation font size
#ADDING LOGOS
logoleft_name: 'Figures/posterdownlogo'
logoleft_width: '3in'
logoleft_xshift: '1in'
logoleft_yshift: '1in'
logoright_name: 'Figures/posterdownlogo'
logoright_width: '3in'
logoright_xshift: '-1in'
logoright_yshift: '1in'

#---POSTER BODY OPTIONS---#
body_bgcol: "ffffff" #colour of the poster main background
body_textsize: "normalsize"    # Size of the main poster body text
body_textcol: "000000" # Colour of main text in the body of poster
column_numbers: 2 # Number of columns that the poster has
column_margins: "0.5in" # Margin spacing for columns
columnline_col: "008080" #colour 
columnline_width: "0pt" #width of line between each column
#SECTION TITLE STYLING
sectitle_textcol: "ffffff" # Colour of the poster section titles
sectitle_bgcol: "0b4545" # Colour of the section title box
sectitle_bordercol: "0b4545" # Colour of the border around the section title box.
sectitle_borderwidth: "2mm" # Thicknes of the section title box border
sectitle_boxshape: "uphill" # Changes the shape of the section title box.

#---BIBLIOGRAPHY OPTIONS---#
bibliography: MyLibrary # name of the .bib file used for referencing
bibliography_spacing: 0.8 # sets the multiplier for line spacing of bibliography spacing (between 0 and 1)
bibliography_textsize: "small"  # size of the bibliography text size (handy for one too many references!)

#---OTHER---#
cite_col: "CC0000" #colour of ciation elements
url_col: "008080" #colour of url links
link_col: "008080" #colour of other links within the poster
footnote_textcol: "ffffff" # Colour of footnote text if used
output: posterdown::posterdown_latex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  results = 'asis',
  echo = FALSE,
  comment = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = 'center'
)
```

# Introduction

Welcome to `posterdown` ! This is my attempt to provide a semi-smooth workflow for those who wish to take their R Markdown skills to the conference world. Most features from R Markdown are available in this package such as Markdown section notation, figure captioning, and even citations like this one [@rmarkdown]. The rest of this example poster will show how you can insert typical conference poster features into your own document.

## Objectives

1. Easy to use reproducible poster design. 
2. Integration with R Markdown.
3. Easy transition from `posterdown` to `thesisdown` or `rticles` [@rticles; @thesisdown].

# Methods

This package uses the same workflow approach as the R Markdown you know and love. Basically it goes from RMarkdown > Knitr > Markdown > Pandoc > HTML/CSS > PDF. You can even use the bibliography the same way [@turnerControlsWaterBalance2014].

## Random text

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam placerat augue at velit tincidunt semper. Donec elementum porta posuere. Nullam interdum, odio at tincidunt feugiat, turpis nisi blandit eros, eu posuere risus felis non quam. Nam eget lorem odio. Duis et aliquet orci. Phasellus nec viverra est. Praesent odio orci, mattis vel mauris nec, consectetur fermentum mauris. Etiam eu hendrerit tortor. Donec mi tellus, efficitur et porttitor eu, auctor eu tellus. Quisque faucibus vestibulum sapien vel lacinia. Ut auctor lorem non interdum blandit.

# Results

Usually you want to have a nice table displaying some important results that you have calculated. In `posterdown` this is as easy as using the `kable` table formatting you are probably use to as per typical R Markdown formatting. I suggesting checking out the `kableExtra` package and its in depth documentation on customizing these tables found [here](https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf) [@kableExtra2019]. You can reference tables like so: Table \@ref(tab:mytable).

```{r mytable, out.width='80%'}
knitr::kable(iris[1:5, 1:4], caption = 'Table caption.',align = 'c',"html")
```

Or with figures: Figure \@ref(fig:standard-plot).

```{r standard-plot, out.width='80%', fig.align='center', echo=TRUE, fig.cap='Great figure!'}
plot(mtcars[1:2])
```

# Next Steps: More random text

Aliquam sed faucibus risus, quis efficitur erat. Vestibulum semper mauris quis tempus eleifend. Aliquam sagittis dictum ipsum, quis viverra ligula eleifend ut. Curabitur sagittis vitae arcu eget faucibus. In non elementum felis. Duis et aliquam nunc. Nunc pulvinar sapien nunc, vel pretium nisi efficitur in. Fusce fringilla maximus leo et maximus. Fusce at ligula laoreet, iaculis mi at, auctor odio. Praesent sed elementum justo. Aenean consectetur risus rhoncus tincidunt efficitur. Praesent dictum mauris at diam maximus maximus [@thorneposterdown2019].

# Conclusion

Try `posterdown` out! Hopefully you like it!

# References

\printbibliography

### 2019-05-28 surprisal/childes_surprisal.Rmd
changes

--- 2019-05-17 surprisal/childes_surprisal.Rmd
@@
 library(ggthemes)
+library(feather)
 
@@
 zhou_utterances <- get_utterances(corpus = "ZhouDinner") #stem for pinyin
-shiro_utterances <- get_utterances(corpus = "Shiro")
+prov_utterances <- get_utterances(corpus = "Okayama")
 ```
@@
 ```{r old unigrams}
+child_utterances <- prov_utterances %>%
+  filter(speaker_role == "Target_Child") %>% 
+  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
+  mutate(utterance_id = 1:n()) %>%
+  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
+  group_by(utterance_id) %>%
+  mutate(word_order = 1:n())
+
+child_unigrams <- child_utterances %>%
+  group_by(word) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(p = n / sum(n))
+
+child_surprisals <- child_utterances %>%
+  left_join(child_unigrams) %>%
+  mutate(s = -log(p)) %>%
+  group_by(length, word_order) %>%
+  summarise(s = mean(s))
+
 adult_utterances <- prov_utterances %>%
@@
 
-adult_surprisals %>%
-  filter(length %in% c(5, 7, 9)) %>%
-  ggplot(aes(x = word_order, y = s)) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_smooth(se = F)
+surprise <- bind_rows(child_surprisals %>%
+  mutate(speaker_role = "Child") %>%
+  bind_rows(mutate(adult_surprisals, speaker_role = "Adult")) %>%
+  mutate(Language = "Japanese"))
+
 ```
@@
 
-adult_bigrams %>% 
-  filter(length %in% c(5, 7, 9)) %>% 
-  ggplot(aes(x = word_order, y = mean_s)) + 
-  facet_wrap(~length) + 
+label_data <- data_frame(length = 7, Language = "Japanese", word_order = 2,
+                         speaker_role = c("Adult", "Child"), s = c(2,4))
+
+surprises %>%  
+  filter(length == 7, Language == "Japanese") %>%
+  ggplot(aes(x = word_order, y = s,
+                        color = speaker_role,
+                        label = speaker_role)) +
+    # facet_grid(Language ~ length, scales = "free") + 
+    xlab("Word position") +
+    ylab("Mean surprisal") +
+    scale_x_continuous(breaks = 1:7) + 
+    geom_line(size = 1) +
     geom_point() + 
-  geom_line()
-
+    theme_classic(base_size = 10) + 
+    scale_color_ptol() +
+    theme(legend.position = "nones") + 
+    geom_text(data = label_data)
 ```
@@
   ungroup() %>%
-  mutate(joint_p = n / sum(n)) %>%
-  select(-n) %>%
   left_join(adult_unigrams, by = c("lag_word" = "word")) %>%
-  mutate(cond_p = joint_p / p) %>%
-  select(-n, -p)
+  mutate(cond_p = n.x/n.y)
 
@@
   ungroup() %>%
-  mutate(tri_joint_p = n / sum(n)) %>%
-  select(-n) %>%
   left_join(adult_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
-  mutate(tri_cond_p = tri_joint_p / joint_p) %>%
-  select(-joint_p, -tri_joint_p, -cond_p) # %>%
+  mutate(tri_cond_p = n / n.x) %>%
+  select(-n, -n.x, -n.y, -p, -cond_p)
 
@@
   left_join(adult_trigrams) %>% 
-  left_join(select(adult_bigrams, -joint_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
+  left_join(select(adult_bigrams, lag_word, word, cond_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
   left_join(adult_unigrams)
@@
 
-all_trigrams <- all_trigrams %>% 
-  bind_rows(mutate(trigram_surprisals_nonnull, speaker_role = "Child", Language = "Spanish"))
-
 trigram_surprisals_nonnull %>%

### 2019-05-30 surprisal/bnc&switchboard.Rmd
cleaned up surprisals folder, added childes and switchboard surprisal feathers

@@
 
-sw <- switchboard %>%
-  left_join(switch_unigrams) %>%
-  mutate(s = -log(p)) %>%
-  ungroup() %>%
-  select(length, word_order, s) %>%
-  nest(s)
-
-swc <- sw %>% filter(length %in% c(10, 20, 30))
-switch_surprisals <- map_dfr(1:nrow(swc), 
-        ~bind_cols(select(slice(swc, .x), length, word_order), 
-                               tidyboot_mean(data = unnest(slice(swc, .x)), 
-                                             column = s, nboot = 1000)))
-
-# switch_surprisals <- switchboard %>%
+# 
+# sw <- switchboard %>%
 #   left_join(switch_unigrams) %>%
@@
 #   select(length, word_order, s) %>%
-#   filter(length %in% c(10, 20, 30)) %>%
-#   group_by(length, word_order) %>% 
-#   tidyboot_mean(s)
+#   nest(s)
+# 
+# swc <- sw %>% filter(length %in% c(10, 20, 30))
+# switch_surprisals <- map_dfr(1:nrow(swc), 
+#         ~bind_cols(select(slice(swc, .x), length, word_order), 
+#                                tidyboot_mean(data = unnest(slice(swc, .x)), 
+#                                              column = s, nboot = 1000)))
 
-switch_surprisals %>%
-  ggplot(aes(x = word_order, y = empirical_stat, #observed mean surprisal
-             ymin = ci_lower, ymax = ci_upper)) + 
-  facet_wrap(~ length) + 
-  ylab("Surprisal") + 
-  xlab("Word Position in Switchboard") + 
-  geom_pointrange(size = .50) + 
-  geom_line(size = .50)
 ```
@@
   ungroup() %>%
-  mutate(joint_p = n / sum(n)) %>%
-  select(-n) %>%
   left_join(switch_unigrams, by = c("lag_word" = "word")) %>%
-  mutate(cond_p = joint_p / p) %>%
-  select(-n, -p)
+  mutate(cond_p = n.x / n.y) %>%
+  select(-n.y, -p) %>%
+  rename(n = n.x)
 
@@
 
-switch_bigram_surprisals %>%
-  filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
-  ggplot(aes(x = word_order, y = s)) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_smooth(se = F)
 ```
@@
   ungroup() %>%
-  mutate(tri_joint_p = n / sum(n)) %>%
-  select(-n) %>%
   left_join(switch_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
-  mutate(tri_cond_p = tri_joint_p / joint_p) %>%
-  select(-joint_p, -tri_joint_p, -cond_p) # %>%
+  mutate(tri_cond_p = n.x / n.y) %>%
+  select(-n.x, -n.y, -cond_p)
 
@@
   left_join(switch_trigrams) %>% 
-  left_join(select(switch_bigrams, -joint_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
-  left_join(switch_unigrams)
+  left_join(switch_bigrams, by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
+  left_join(switch_unigrams, by = c("word" = "word"))
 
@@
 
-switch_trigram_surprisals %>%
-  filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
-  ggplot(aes(x = word_order, y = s)) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_smooth(se = F)
 ```
@@
 
-bnc_surprisals %>%
-  ggplot(aes(x = word_order, y = empirical_stat, #observed mean surprisal
-             ymin = ci_lower, ymax = ci_upper)) + 
-  facet_wrap(~ length) + 
-  ylab("Surprisal") + 
-  xlab("Word Position in BNC") + 
-  geom_pointrange(size = .50) + 
-  geom_line(size = .50)
-
-# bnc_surprisals %>%
-#   ggplot(aes(x = word_order, y = s)) + 
-#   facet_wrap(~ length) + 
-#   geom_point() + 
-#   geom_smooth(se = F)
 ```
@@
   filter(word_order <= length)
-
-bnc_bigram_surprisals %>%
-  ggplot(aes(x = word_order, y = s)) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_smooth(se = F)
 ```

### 2019-05-30 surprisal/childes_surprisal.Rmd
cleaned up surprisals folder, added childes and switchboard surprisal feathers

@@
 zhou_utterances <- get_utterances(corpus = "ZhouDinner") #stem for pinyin
-prov_utterances <- get_utterances(corpus = "Okayama")
+okayama_utterances <- get_utterances(corpus = "Okayama")
 ```
 
-```{r old unigrams}
+```{r unigrams}
+# column name must be a string, NOT just the name itself
+get_unigrams <- function(utterances, speaker_role, name) {
+
+  if (tolower(speaker_role) == "child") {
+    utts <- utterances %>%
+      filter(speaker_role == "Target_Child") 
+  } else {
+    utts <- utterances %>%
+      filter(speaker_role != "Target_Child")   
+  }
+
+  utts <- utts %>%
+    rename(text = name) %>%
+    mutate(length = str_count(text, pattern = "[ +]+") + 1) %>%
+    mutate(utterance_id = 1:n()) %>%
+    unnest_tokens(word, text, token = stringr::str_split, pattern = "[ +]+") %>%
+    group_by(utterance_id) %>%
+    mutate(word_order = 1:n())
+
+  unigrams <- utts %>%
+    group_by(word) %>%
+    count() %>%
+    ungroup() %>%
+    mutate(p = n / sum(n)) %>%
+    mutate(s = -log(p)) %>%
+    select(word, s)
+
+  surprisals <- utts %>%
+    left_join(unigrams) %>%
+    group_by(length, word_order) %>%
+    summarise(s = mean(s)) %>%
+    ungroup()
+}
+
+prov_sups <- get_unigrams(prov_utterances, "not child", "stem") %>% 
+  mutate(speaker_role = "adult") %>%
+  bind_rows(mutate(get_unigrams(prov_utterances, "child", "stem"), speaker_role = "child")) %>%
+  mutate(language = "English")
+
+```
+
+Sample code for plotting your surprisal results.
+
+```{r plotting}
+label_data <- data_frame(length = 7, Language = "Japanese", word_order = 2,
+                         speaker_role = c("Adult", "Child"), s = c(2,4))
+
+surprises %>%  
+  filter(length == 7, Language == "Japanese") %>%
+  ggplot(aes(x = word_order, y = s,
+                        color = speaker_role,
+                        label = speaker_role)) +
+    # facet_grid(Language ~ length, scales = "free") + 
+    xlab("Word position") +
+    ylab("Mean surprisal") +
+    scale_x_continuous(breaks = 1:7) + 
+    geom_line(size = 1) +
+    geom_point() + 
+    theme_classic(base_size = 10) + 
+    scale_color_ptol() +
+    theme(legend.position = "nones") + 
+    geom_text(data = label_data)
+```
+
+Down here is the old code. You can run it, but you have to change the corpus name and whether you're using stem or gloss by hand.
+
+```{r hard-coded language unigrams}
 child_utterances <- prov_utterances %>%
@@
 
-surprise <- bind_rows(child_surprisals %>%
-  mutate(speaker_role = "Child") %>%
-  bind_rows(mutate(adult_surprisals, speaker_role = "Adult")) %>%
-  mutate(Language = "Japanese"))
-
 ```
 
-```{r redoing bigrams}
+```{r new bigrams}
 adult_bigrams_prep <- adult_utterances %>% 
@@
   summarise(mean_s = mean(s)) 
-
-label_data <- data_frame(length = 7, Language = "Japanese", word_order = 2,
-                         speaker_role = c("Adult", "Child"), s = c(2,4))
-
-surprises %>%  
-  filter(length == 7, Language == "Japanese") %>%
-  ggplot(aes(x = word_order, y = s,
-                        color = speaker_role,
-                        label = speaker_role)) +
-    # facet_grid(Language ~ length, scales = "free") + 
-    xlab("Word position") +
-    ylab("Mean surprisal") +
-    scale_x_continuous(breaks = 1:7) + 
-    geom_line(size = 1) +
-    geom_point() + 
-    theme_classic(base_size = 10) + 
-    scale_color_ptol() +
-    theme(legend.position = "nones") + 
-    geom_text(data = label_data)
-```
-
-```{r unigrams}
-get_unigrams <- function(utterances, speaker_role, stem_or_gloss) {
-
-  if (tolower(speaker_role) == "child") {
-    utts <- utterances %>%
-      filter(speaker_role == "Target_Child") 
-  } else {
-    utts <- utterances %>%
-      filter(speaker_role != "Target_Child")   
-  }
-
-  if (tolower(stem_or_gloss) == "stem") {
-    utts <- utts %>%
-      mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
-      mutate(utterance_id = 1:n()) %>%
-      unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
-      group_by(utterance_id) %>%
-      mutate(word_order = 1:n())
-
-  } else {
-    utts <- utts %>%
-      mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
-      mutate(utterance_id = 1:n()) %>%
-      unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
-      group_by(utterance_id) %>%
-      mutate(word_order = 1:n())
-  }
-
-  unigrams <- utts %>%
-    group_by(word) %>%
-    count() %>%
-    ungroup() %>%
-    mutate(p = n / sum(n)) %>%
-    mutate(s = -log(p)) %>%
-    select(word, s)
-
-  surprisals <- utts %>%
-    left_join(unigrams) %>%
-    group_by(length, word_order) %>%
-    summarise(s = mean(s)) %>%
-    ungroup()
-}
-
-prov_sups <- get_unigrams(prov_utterances, "not child", "stem") %>% 
-  mutate(speaker_role = "adult") %>%
-  bind_rows(mutate(get_unigrams(prov_utterances, "child", "stem"), speaker_role = "child")) %>%
-  mutate(language = "English")
-
-shiro_sups <- get_unigrams(shiro_utterances, "not child", "gloss") %>% 
-  mutate(speaker_role = "adult") %>%
-  bind_rows(mutate(get_unigrams(shiro_utterances, "child", "gloss"), speaker_role = "child")) %>%
-  mutate(language = "Spanish")
-
-zhou_sups <- get_unigrams(zhou_utterances, "not child", "stem") %>% 
-  mutate(speaker_role = "adult") %>%
-  bind_rows(mutate(get_unigrams(zhou_utterances, "child", "stem"), speaker_role = "child")) %>%
-  mutate(language = "Mandarin")
-
-all_sups <- prov_sups %>%
-  bind_rows(shiro_sups, zhou_sups)
-
-label_data <- data_frame(length = 9, Language = "Spanish", word_order = 3,
-                         speaker_role = c("Adult", "Child"), s = c(1.25,3.5))
-
-all_trigrams %>% 
-  filter(length %in% c(5, 7, 9)) %>%
-  ggplot(aes(x = word_order, y = s,
-                        color = speaker_role,
-                        label = speaker_role)) +
-    facet_grid(Language ~ length, scales = "free") + 
-    xlab("Word position") +
-    ylab("Mean surprisal") +
-    geom_line(size = 1) +
-    geom_point() + 
-    theme_classic(base_size = 10) + 
-    scale_color_ptol() +
-    theme(legend.position = "nones") + 
-    geom_text(data = label_data)
-
 ```
@@
 
-bigram_surprisals_nonnull <- bigram_surprisals_prep %>%
+bigram_surprisals <- bigram_surprisals_prep %>%
   mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
@@
 
-all_bigrams <- all_bigrams %>% 
-  bind_rows(mutate(bigram_surprisals_nonnull, speaker_role = "Child", Language = "Spanish"))
-
-bigram_surprisals_nonnull %>%
-  filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
-  ggplot(aes(x = word_order, y = s)) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_smooth(se = F)
 ```
@@
 
-trigram_surprisals_nonnull <- trigram_surprisals_prep %>%
+trigram_surprisals <- trigram_surprisals_prep %>%
   mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
@@
   filter(word_order <= length)
-
-trigram_surprisals_nonnull %>%
-  filter(length %in% c(5, 7, 9, 11, 13, 15)) %>%
-  ggplot(aes(x = word_order, y = s)) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_smooth(se = F)
-
 ```

### 2019-05-30 entropy/Wikipedia/wiki_clustering.Rmd
reorganized surprisal folder more

--- 2019-03-14 wiki_entropy/wiki_clustering.Rmd
@@
   filter(term == "cosine") %>%
-  arrange(desc(statistic)) %>%
-  write_feather(here("word_order_glm_absolute.feather"))
+  arrange(desc(statistic)) 
 ```

### 2019-05-30 extensions/CogSci 2019 paper/position-info.tex
reorganized surprisal folder more

--- 2019-02-02 cogsci/position-info.tex
@@
 prediction of these models is that speakers should keep the density of
-information constant over the duration of an utterance (R. P. Levy \&
-Jaeger, 2007). However, different languages have different structural
-properties that constrain the space of a speaker's choices
-(e.g.~canonical word order). We build on a method developed by Yu, Cong,
-Liang, \& Liu (2016) to analyze the word-level entropy curves of natural
-language productions across a diverse set of languages and in diverse
-written and spoken contexts. We show that languages impose
-characteristic constraints on speaker's choices that predict deviations
-from Uniform Information Density, and that cross-linguistic variability
-in these deviations is predictable in part from syntactic properties of
-those languages.
+information constant over the duration of an utterance (Levy \& Jaeger,
+2007). However, different languages have different structural properties
+that constrain the space of a speaker's choices (e.g.~canonical word
+order). We build on a method developed by Yu, Cong, Liang, \& Liu (2016)
+to analyze the word-level entropy curves of natural language productions
+across a diverse set of languages and in diverse written and spoken
+contexts. We show that languages impose characteristic constraints on
+speaker's choices that predict deviations from Uniform Information
+Density, and that cross-linguistic variability in these deviations is
+predictable in part from syntactic properties of those languages.
 
@@
 the recent development of theories of communication in human and machine
-language processing (Jelinek, 1976; R. P. Levy \& Jaeger, 2007).
+language processing (Jelinek, 1976; Levy \& Jaeger, 2007).
 
@@
 of the words in a sentence, i.e.~maintain \emph{Uniform Information
-Density} (Genzel \& Charniak, 2002; R. P. Levy \& Jaeger, 2007).
+Density} (Genzel \& Charniak, 2002; Levy \& Jaeger, 2007).
 
@@
 been applied broadly over the past decade. The original evidence in
-favor of Uniform Information Density in R. P. Levy \& Jaeger (2007)
-finds that the insertion of complementizers (e.g. ``that'') in relative
-clauses in English corresponds to where neighboring words have high
-information content. Similarly, A. F. Frank \& Jaeger (2008) argues that
-contractions in English such as ``you're'' do not occur when neighboring
-words are highly informative. Applications of Uniform Information
-Density include determining whether linguistic alignment takes place
-(Jaeger \& Snider, 2013), Zipfian word length distributions (Piantadosi,
-Tily, \& Gibson, 2011), communication efficiency (Mahowald, Fedorenko,
-Piantadosi, \& Gibson, 2013), dialogue and turn-taking (Xu \& Reitter,
-2018) and the significance of ambiguity in language (Piantadosi, Tily,
-\& Gibson, 2012), among other research.
+favor of Uniform Information Density in Levy \& Jaeger (2007) finds that
+the insertion of complementizers (e.g. ``that'') in relative clauses in
+English corresponds to where neighboring words have high information
+content. Similarly, Frank \& Jaeger (2008) argues that contractions in
+English such as ``you're'' do not occur when neighboring words are
+highly informative. Applications of Uniform Information Density include
+determining whether linguistic alignment takes place (Jaeger \& Snider,
+2013), Zipfian word length distributions (Piantadosi, Tily, \& Gibson,
+2011), communication efficiency (Mahowald, Fedorenko, Piantadosi, \&
+Gibson, 2013), dialogue and turn-taking (Xu \& Reitter, 2018) and the
+significance of ambiguity in language (Piantadosi, Tily, \& Gibson,
+2012), among other research.
 
 However, other recent work has contradicted the Uniform Information
-Density hypothesis. Similar to the original work in R. P. Levy \& Jaeger
+Density hypothesis. Similar to the original work in Levy \& Jaeger
 (2007), Zhan \& Levy (2018) focuses on information distribution at
@@
 
+\hypertarget{calculating-entropy-curves}{%
 \subsection{Calculating entropy
-curves}\label{calculating-entropy-curves}
+curves}\label{calculating-entropy-curves}}
 
@@
 within a corpus. Uniform Information Density applies this uniformity of
-entropy rate in sentences to all levels of speech (R. P. Levy \& Jaeger,
+entropy rate in sentences to all levels of speech (Levy \& Jaeger,
 2007), and so our method obtained from Yu et al. (2016), which examines
@@
 
-\section{Study 1}\label{study-1}
+\hypertarget{study-1}{%
+\section{Study 1}\label{study-1}}
 
@@
 
-\subsection{Data and Analysis}\label{data-and-analysis}
+\hypertarget{data-and-analysis}{%
+\subsection{Data and Analysis}\label{data-and-analysis}}
 
@@
 
-\subsection{Results}\label{results}
+\hypertarget{results}{%
+\subsection{Results}\label{results}}
 
@@
 
-\section{Study 2}\label{study-2}
+\hypertarget{study-2}{%
+\section{Study 2}\label{study-2}}
 
@@
 
-\subsection{Data and Analysis}\label{data-and-analysis-1}
+\hypertarget{data-and-analysis-1}{%
+\subsection{Data and Analysis}\label{data-and-analysis-1}}
 
@@
 
-\subsection{Results}\label{results-1}
+\hypertarget{results-1}{%
+\subsection{Results}\label{results-1}}
 
@@
 
-\section{Study 3}\label{study-3}
+\hypertarget{study-3}{%
+\section{Study 3}\label{study-3}}
 
@@
 
-\subsection{Data and Analysis}\label{data-and-analysis-2}
+\hypertarget{data-and-analysis-2}{%
+\subsection{Data and Analysis}\label{data-and-analysis-2}}
 
@@
 
-\subsection{Results}\label{results-2}
+\hypertarget{results-2}{%
+\subsection{Results}\label{results-2}}
 
@@
 
-\section{General Discussion}\label{general-discussion}
+\hypertarget{general-discussion}{%
+\section{General Discussion}\label{general-discussion}}
 
@@
 
-\hypertarget{ref-clear1993}{}
+\leavevmode\hypertarget{ref-clear1993}{}%
 Clear, J. H. (1993). The british national corpus. \emph{The Digital
@@
 
-\hypertarget{ref-demuth2006}{}
+\leavevmode\hypertarget{ref-demuth2006}{}%
 Demuth, K., Culbertson, J., \& Alter, J. (2006). Word-minimality,
@@
 
-\hypertarget{ref-dryer2013}{}
+\leavevmode\hypertarget{ref-dryer2013}{}%
 Dryer, M. S., \& Haspelmath, M. (2013). \emph{The world atlas of
@@
 
-\hypertarget{ref-ferrer-i-cancho2013}{}
+\leavevmode\hypertarget{ref-ferrer-i-cancho2013}{}%
 Ferrer-i-Cancho, R., Debowski, L., \& Prado Martin, F. M. del. (2013).
@@
 
-\hypertarget{ref-frank2008}{}
+\leavevmode\hypertarget{ref-frank2008}{}%
 Frank, A. F., \& Jaeger, T. F. (2008). Speaking rationally: Uniform
@@
 
-\hypertarget{ref-godfrey1992}{}
+\leavevmode\hypertarget{ref-godfrey1992}{}%
 Godfrey, J. J., Holliman, E. C., \& McDaniel, J. (1992). SWITCHBOARD:
 Telephone speech corpus for research and development. In
-\emph{Acoustics, speech, and signal processing, 1992. icassp-92., 1992
+\emph{Acoustics, speech, and signal processing, 1992. ICASSP-92., 1992
 ieee international conference on} (Vol. 1, pp. 517--520). IEEE.
 
-\hypertarget{ref-jaeger2013}{}
+\leavevmode\hypertarget{ref-jaeger2013}{}%
 Jaeger, T. F., \& Snider, N. E. (2013). Alignment as a consequence of
@@
 
-\hypertarget{ref-landauer1997}{}
+\leavevmode\hypertarget{ref-landauer1997}{}%
 Landauer, T. K., \& Dumais, S. T. (1997). A solution to plato's problem:
@@
 
-\hypertarget{ref-li2015}{}
+\leavevmode\hypertarget{ref-li2015}{}%
 Li, H., \& Zhou, J. (2015). \emph{Study on dinner table talk of
@@
 
-\hypertarget{ref-macwhinney2014}{}
+\leavevmode\hypertarget{ref-macwhinney2014}{}%
 MacWhinney, B. (2014). \emph{The childes project: Tools for analyzing
@@
 
-\hypertarget{ref-sanchez2019}{}
+\leavevmode\hypertarget{ref-sanchez2019}{}%
 Sanchez, A., Meylan, S. C., Braginsky, M., MacDonald, K. E., Yurovsky,
@@
 interface to the child language data exchange system. \emph{Behavior
-Research Methods}. in press.
+Research Methods}.
 
-\hypertarget{ref-shannon1948}{}
+\leavevmode\hypertarget{ref-shannon1948}{}%
 Shannon, C. E. (1948). A mathematical theory of communication.
@@
 
-\hypertarget{ref-stivers2009universals}{}
+\leavevmode\hypertarget{ref-stivers2009universals}{}%
 Stivers, T., Enfield, N. J., Brown, P., Englert, C., Hayashi, M.,

### 2019-05-30 extensions/Language Dev. Proj./LDP_entropy.Rmd
reorganized surprisal folder more

--- 2018-11-18 ldp/LDP_entropy.Rmd
@@
 get_ses_utterances <- function(status, utterances, home_visits) {
-  income <- home_visits %>% filter(income_category %in% status)
-
-  ids <- unique(income$id) %>% unlist()
+  ids <- home_visits %>% 
+    filter(income_category %in% status)  %$% 
+    id %>% 
+    unique() %>% 
+    unlist()
 
@@
 }
-
-ses_utts <- get_ses_utterances(c(6), utterances, home_visits)
 ```
 
-```{r ldp_analysis}
+```{r entropy}
 get_unigram_entropies <- function(role, sen_length, utterances) {
@@
       mutate(length = str_count(c_utts_orig, " |\\+|_") + 1)
-
       colnames(sub_utterances)[colnames(sub_utterances)=="c_utts_orig"] <- "gloss"
@@
       mutate(length = str_count(p_utts_orig, " |\\+|_") + 1) 
-
       colnames(sub_utterances)[colnames(sub_utterances)=="p_utts_orig"] <- "gloss"
@@
 }
+```
 
-entropies <- map(2:10, ~get_unigram_entropies("child", .x, ses_utts)) %>%
+```{r ldp_analysis}
+STATUS <- c(1)
+
+ses_utts <- get_ses_utterances(STATUS, utterances, home_visits)
+
+entropies_child <- map(2:10, ~get_unigram_entropies("child", .x, ses_utts)) %>%
   bind_rows()
 
-ggplot(entropies, aes(x = word_order, y = empirical_entropy,
+ggplot(entropies_child, aes(x = word_order, y = empirical_entropy,
                       ymin = ci_lower, ymax = ci_upper)) +
@@
   geom_smooth(se = F)
+
+entropies_adult <- map(2:10, ~get_unigram_entropies("not child", .x, ses_utts)) %>%
+  bind_rows()
+
+ggplot(entropies_adult, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper)) +
+  facet_wrap(~ length) + 
+  geom_pointrange() +
+  geom_smooth(se = F)
+
+entropies_all <- bind_rows(entropies_adult, entropies_child)
+
+ggplot(entropies_all, aes(x = word_order, y = empirical_entropy,
+                      ymin = ci_lower, ymax = ci_upper, color = role)) +
+  facet_wrap(~ length) + 
+  geom_pointrange(position = position_dodge(.25)) +
+  geom_smooth(se = F) +
+  ggtitle("Child and Adult Entropy Distributions (LDP SES Bracket 1)")
 ```

### 2019-05-30 surprisal/wiki_surprisal.Rmd
childes unigrams
reorganized surprisal folder more
tiny change to filepath in wiki_surprisal

@@
 
+```{r make dendrogram}
+wiki <- read.csv(here("Data/relative_unigrams.csv")) %>%
+  select(-X) %>% 
+  column_to_rownames("Language")
+
+wiki %>% 
+  dist() %>% 
+  hclust() %>% 
+  plot(hang = -1, 
+       main = "Dendrogram of slope embeddings")
+
+```
+
 ```{r word order}
-wkpd <- read.csv(here("relative_unigrams.csv"))
+wkpd <- read.csv(here("Data/relative_unigrams.csv"))
 LANGUAGES <- wkpd$Language
@@
 
-feature_names <- read_feather(here("feature_names.feather"))
+feature_names <- read_feather(here("Wikipedia/feature_names.feather"))
 

### 2019-05-30 surprisal/MacWhinney poster/poster_plotting.Rmd
poster plot

--- 2019-05-30 surprisal/childes_surprisal.Rmd
@@
 library(directlabels)
-library(childesr) #data
 library(tidytext)
-library(entropy)
 library(tidyboot)
-library(dplyr)
-library(tokenizers)
-library(gtools)
 library(here)
@@
 library(feather)
+library(janitor)
 
@@
 
-theme_set(theme_classic(base_size = 16))
+theme_set(theme_classic(base_size = 24))
 ```
@@
 ```{r get_utterances}
-prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
-zhou_utterances <- get_utterances(corpus = "ZhouDinner") #stem for pinyin
-okayama_utterances <- get_utterances(corpus = "Okayama")
-```
+childes <- read_feather(here("Data/childes_surprisals.feather")) %>%
+  clean_names() %>%
+  mutate(speaker_role = if_else(speaker_role == "Adult", "Parent", "Child"))
 
-```{r unigrams}
-# column name must be a string, NOT just the name itself
-get_unigrams <- function(utterances, speaker_role, name) {
-
-  if (tolower(speaker_role) == "child") {
-    utts <- utterances %>%
-      filter(speaker_role == "Target_Child") 
-  } else {
-    utts <- utterances %>%
-      filter(speaker_role != "Target_Child")   
-  }
-
-  utts <- utts %>%
-    rename(text = name) %>%
-    mutate(length = str_count(text, pattern = "[ +]+") + 1) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, text, token = stringr::str_split, pattern = "[ +]+") %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n())
-
-  unigrams <- utts %>%
-    group_by(word) %>%
-    count() %>%
-    ungroup() %>%
-    mutate(p = n / sum(n)) %>%
-    mutate(s = -log(p)) %>%
-    select(word, s)
-
-  surprisals <- utts %>%
-    left_join(unigrams) %>%
-    group_by(length, word_order) %>%
-    summarise(s = mean(s)) %>%
-    ungroup()
-}
-
-prov_sups <- get_unigrams(prov_utterances, "not child", "stem") %>% 
-  mutate(speaker_role = "adult") %>%
-  bind_rows(mutate(get_unigrams(prov_utterances, "child", "stem"), speaker_role = "child")) %>%
-  mutate(language = "English")
-
+switchboard <- read_feather(here("Data/switchboard_surprisals.feather")) %>%
+  clean_names() %>%
+  mutate(language = "English", speaker_role = "Adult") %>%
+  select(-corpus)
 ```
@@
 ```{r plotting}
-label_data <- data_frame(length = 7, Language = "Japanese", word_order = 2,
-                         speaker_role = c("Adult", "Child"), s = c(2,4))
+data <- bind_rows(childes, switchboard) %>%
+  filter(context %in% c(0, 2)) %>%
+  mutate(language = factor(language, 
+                           levels = c("English", "Spanish", 
+                                      "Mandarin", "Japanese")),
+         context = factor(context, levels = c(0, 2), 
+                          labels = c("Unigram", "Trigram")),
+         speaker_role = factor(speaker_role, 
+                               levels = c("Parent", "Child", "Adult")))
 
-surprises %>%  
-  filter(length == 7, Language == "Japanese") %>%
+label_data <- tibble(language = "English",
+                     speaker_role = unique(data$speaker_role),
+                     word_order = c(2, 3, 4),
+                     s = c(6, 5, 4),
+                     context = "Unigram") %>%
+  mutate(context = factor(context, levels = c("Unigram", "Trigram")),
+         language = factor(language, 
+                           levels = c("English", "Spanish", 
+                                      "Mandarin", "Japanese")))
+
+pdf("surprisals.pdf", width = 10, height = 5.5)
+data %>%
+  filter(length == 8) %>%
    ggplot(aes(x = word_order, y = s,
@@
                         label = speaker_role)) +
-    # facet_grid(Language ~ length, scales = "free") + 
-    xlab("Word position") +
-    ylab("Mean surprisal") +
-    scale_x_continuous(breaks = 1:7) + 
-    geom_line(size = 1) +
-    geom_point() + 
-    theme_classic(base_size = 10) + 
-    scale_color_ptol() +
-    theme(legend.position = "nones") + 
-    geom_text(data = label_data)
+  facet_grid(context ~ language, scales = "free") +
+  geom_point(size = 1.5) +
+  geom_line(size = 1.5) +
+  scale_color_brewer(palette = "Set1") + 
+  theme(legend.position = "none") +
+  geom_text(data = label_data, size = 6) + 
+  labs(x = "Utterance Position", y = "Surprisal")
+dev.off()
 ```

### 2019-06-04 surprisal/childes_surprisal.Rmd
slope estimates

@@
      group_by(length, word_order) %>%
-    summarise(s = mean(s)) %>%
+    summarise(mean = mean(log(s)),
+              median = median(log(s)),
+              se = sd(s)/sqrt(length(s))) %>%
     ungroup()
@@
 
+prov_sups %>%
+  filter(speaker_role == "child", length == 6, word_order == 3) %>%
+  ggplot(aes(x = log(s))) + 
+  geom_histogram()
 ```
@@
 
-surprises %>%  
-  filter(length == 7, Language == "Japanese") %>%
-  ggplot(aes(x = word_order, y = s,
-                        color = speaker_role,
-                        label = speaker_role)) +
+child_surprisals %>%  
+  filter(length == 12) %>%
+  ggplot(aes(x = word_order, y = median)) +
     # facet_grid(Language ~ length, scales = "free") + 
@@
     geom_line(size = 1) +
-    geom_point() + 
+    geom_pointrange(aes(ymin = mean - 1.96 * se, ymax = mean + 1.96 * se)) + 
     theme_classic(base_size = 10) + 
     scale_color_ptol() +
-    theme(legend.position = "nones") + 
-    geom_text(data = label_data)
+    theme(legend.position = "nones") #+ 
+   # geom_text(data = label_data)
 ```
@@
   group_by(length, word_order) %>%
-  summarise(s = mean(s))
+  summarise(mean = mean(s),
+            median = median(s),
+            se = sd(s)/sqrt(n()))
 
@@
   group_by(length, word_order) %>%
-  summarise(s = mean(s))
+  summarise(mean = mean(s),
+            se = sd(s)/sqrt(n()))
 

### 2019-06-04 ngrams/get_wiki_df.py

@@ -1,3 +1,4 @@
+# -*- coding: utf-8 -*-
 import regex as re
 import numpy as np
 import pandas as pd

### 2019-06-04 surprisal/Wikipedia/wiki_relative_surprisal.R

@@ -81,7 +81,7 @@ system(paste("python3 get_wiki_df.py", LANGUAGE, sep = " "))
 
 df <- read_csv("wiki_df.csv")
 
-df_tokens <- extract_wiki_tokens(6, 50, df)
+df_tokens <- extract_wiki_tokens(6, 10, df)
 
 pos_list <- df_tokens %>%
   group_by(length) %>%
@@ -110,3 +110,10 @@ write.csv(all_data, file = "relative_unigrams.csv")
 # lang_slopes <- c(LANGUAGE, slopes$estimate)
 # gs_edit_cells(for_gs, ws = "Sheet1", anchor = paste("A", nr + 2, sep=""), input = lang_slopes, byrow = TRUE)
 system("rm wiki_df.csv")
+
+slopes %>%
+  mutate(lower = estimate - 1.96 * std.error,
+         upper = estimate + 1.96 * std.error) %>%
+  ggplot(aes(x = cut, y = estimate, group = 1)) + 
+  geom_pointrange(aes(ymin = lower, ymax = upper)) + 
+  geom_line()

### 2019-06-09 surprisal/MacWhinney poster/poster_plotting.Rmd
unigram letter surprisals

@@
 ---
-title: Positional surprisal (1-3grams) in CHILDES corpora
+title: Plotting surprisal in CHILDES corpora and Switchboard English corpus
 author: Josef Klafka and Dan Yurovsky 
@@
 
-label_data <- tibble(language = "English",
+label_data <- tibble(language = c("Spanish", "Spanish", "English"),
                      speaker_role = unique(data$speaker_role),
-                     word_order = c(2, 3, 4),
-                     s = c(6, 5, 4),
+                     word_order = c(2, 4, 2),
+                     s = c(6.5, 5, 6.5),
                      context = "Unigram") %>%
@@
   geom_text(data = label_data, size = 6) + 
-  labs(x = "Utterance Position", y = "Surprisal")
+  labs(x = "Word Position", y = "Average Information")
 dev.off()
 ```
-
-Down here is the old code. You can run it, but you have to change the corpus name and whether you're using stem or gloss by hand.
-
-```{r hard-coded language unigrams}
-child_utterances <- prov_utterances %>%
-  filter(speaker_role == "Target_Child") %>% 
-  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
-  mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
-  group_by(utterance_id) %>%
-  mutate(word_order = 1:n())
-
-child_unigrams <- child_utterances %>%
-  group_by(word) %>%
-  count() %>%
-  ungroup() %>%
-  mutate(p = n / sum(n))
-
-child_surprisals <- child_utterances %>%
-  left_join(child_unigrams) %>%
-  mutate(s = -log(p)) %>%
-  group_by(length, word_order) %>%
-  summarise(s = mean(s))
-
-adult_utterances <- prov_utterances %>%
-  filter(speaker_role != "Target_Child") %>% 
-  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
-  mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
-  group_by(utterance_id) %>%
-  mutate(word_order = 1:n())
-
-adult_unigrams <- adult_utterances %>%
-  group_by(word) %>%
-  count() %>%
-  ungroup() %>%
-  mutate(p = n / sum(n))
-
-adult_surprisals <- adult_utterances %>%
-  left_join(adult_unigrams) %>%
-  mutate(s = -log(p)) %>%
-  group_by(length, word_order) %>%
-  summarise(s = mean(s))
-
-```
-
-```{r new bigrams}
-adult_bigrams_prep <- adult_utterances %>% 
-  group_by(utterance_id) %>%
-  mutate(lag_word = lag(word)) %>%
-  group_by(lag_word, word) %>%
-  count() %>%
-  filter(!is.na(lag_word)) %>%
-  ungroup() %>%
-  left_join(adult_unigrams, by = c("lag_word" = "word")) %>%
-  mutate(cond_p = n.x / n.y) %>%
-  select(lag_word, word, cond_p)
-
-adult_bigrams <- adult_utterances %>%
-  group_by(utterance_id) %>%
-  mutate(lag_word = lag(word)) %>%
-  ungroup() %>%
-  left_join(adult_unigrams) %>%
-  left_join(adult_bigrams_prep) %>%
-  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
-  group_by(length, word_order) %>%
-  summarise(mean_s = mean(s)) 
-```
-
-```{r bigrams}
-adult_bigrams <- adult_utterances %>%
-  group_by(utterance_id) %>%
-  mutate(lag_word = lag(word)) %>%
-  group_by(lag_word, word) %>%
-  count() %>%
-  filter(!is.na(lag_word)) %>%
-  ungroup() %>%
-  left_join(adult_unigrams, by = c("lag_word" = "word")) %>%
-  mutate(cond_p = n.x/n.y)
-
-bigram_surprisals_prep <- adult_utterances %>%
-  group_by(utterance_id) %>%
-  mutate(lag_word = lag(word)) %>%
-  left_join(adult_bigrams) %>%
-  left_join(adult_unigrams)
-
-bigram_surprisals <- bigram_surprisals_prep %>%
-  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
-  group_by(length, word_order) %>%
-  summarise(s = mean(s)) %>% 
-  filter(word_order <= length)
-
-```
-
-```{r trigrams}
-adult_trigrams <- adult_utterances %>%
-  group_by(utterance_id) %>%
-  mutate(lag_word1 = lag(word)) %>%
-  mutate(lag_word2 = lag(lag_word1)) %>%
-  group_by(lag_word2, lag_word1, word) %>%
-  count() %>%
-  filter(!is.na(lag_word1), !is.na(lag_word2)) %>%
-  ungroup() %>%
-  left_join(adult_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
-  mutate(tri_cond_p = n / n.x) %>%
-  select(-n, -n.x, -n.y, -p, -cond_p)
-
-trigram_surprisals_prep <- adult_utterances %>%
-  group_by(utterance_id) %>%
-  mutate(lag_word1 = lag(word)) %>%
-  mutate(lag_word2 = lag(lag_word1)) %>%
-  left_join(adult_trigrams) %>% 
-  left_join(select(adult_bigrams, lag_word, word, cond_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
-  left_join(adult_unigrams)
-
-trigram_surprisals <- trigram_surprisals_prep %>%
-  mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
-                    ifelse(is.na(lag_word1), # check if not the second word
-                           -log(p), # trigram if third or beyond
-                           -log(cond_p)), # bigram if second word
-                    -log(tri_cond_p))) %>% # unigram if first word
-  group_by(length, word_order) %>%
-  summarise(s = mean(s)) %>% 
-  filter(word_order <= length)
-```

### 2019-06-09 surprisal/letter_surprisal.Rmd
unigram letter surprisals

---
title: Wikipedia Text Mining and Surprisal Analysis
author: Josef Klafka and Dan Yurovsky
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(childesr)

knitr::opts_chunk$set(echo = TRUE)
```

```{r prov}
prov_utterances <- get_utterances(corpus = "Providence")

adult_utterances <- prov_utterances %>%
  filter(speaker_role != "Target_Child") %>% 
  mutate(utterance_id = 1:n()) %>% 
  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
  unnest_tokens(letter, word, token = stringr::str_split, pattern = "", drop = F) %>%
  group_by(utterance_id, word) %>%
  mutate(letter_order = 1:n()) %>%
  mutate(word_length = max(letter_order)) %>%
  ungroup()

letter_probs <- adult_utterances %>% 
  count(letter) %>%
  ungroup() %>%
  mutate(p = n/sum(n))

letter_unigrams <- adult_utterances %>% 
  left_join(letter_probs) %>%
  mutate(s = -log(p)) %>%
  group_by(word_length, letter_order) %>%
  summarise(mean_s = mean(s))

letter_unigrams %>%
  filter(word_length %in% 3:11) %>%
  ggplot(aes(x = letter_order, y = mean_s)) + 
    facet_wrap(~word_length) + 
    geom_point() +
    geom_line() + 
    xlab("Letter Position") + 
    ylab("Mean surprisal")

```

### 2019-06-09 surprisal/wiki_surprisal.Rmd
unigram letter surprisals

@@
 library(janitor)
+library(ggdendro)
 
@@
 ```{r make dendrogram}
-wiki <- read.csv(here("Data/relative_unigrams.csv")) %>%
+wiki <- read.csv(here("Data/wikipedia_0context.csv")) %>%
   select(-X) %>% 
@@
   hclust() %>% 
+  ggdendrogram(rotate = FALSE, size = .1) + 
+  ylim(0, .10)
+
   plot(hang = -1, 

### 2019-06-09 ngrams/wiki_relative_ngrams.R

@@ -1,113 +1,0 @@
-library(tidyverse)
-library(directlabels)
-library(tidytext)
-library(entropy)
-library(tidyboot)
-library(tokenizers)
-library(googlesheets)
-library(stringr)
-### Begin Functions
-
-extract_wiki_tokens <- function(min_length, max_length, utterances, N) {
-
-  tokens <- utterances %>%
-    mutate(length = str_count(words, pattern = "[ +]+") + 1) %>%
-    filter(length >= min_length) %>%
-    filter(length <= max_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, words, token = stringr::str_split, pattern = "[ +]+") %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n())
-
-
-  lags <- expand.grid(lag_n = 1:(N-1),
-                      utterance_id = unique(tokens$utterance_id))
-
-  tokens %>%
-    left_join(lags, by = "utterance_id") %>%
-    arrange(utterance_id, lag_n, word_order) %>%
-    mutate(lag = lag(word, n = first(lag_n))) %>%
-    drop_na() %>%
-    mutate(gram = str_c(lag, word, sep = " ")) %>%
-    rename(gram_order = word_order) %>%
-    group_by(length, gram_order, gram) %>%
-    summarise(n = n())
-
-}
-
-get_quantiles <- function(df, num_sections) {
-
-  quantile(df$gram_order, probs = seq(0, 1, 1/num_sections)) %>%
-    round() %>%
-    as_data_frame() %>%
-    rename(gram_order = value)
-}
-
-relative_slopes <- function(df, pos_list) {
-
-  get_slope <- function(x) {
-
-    selected_pos <- pos_list %>%
-      group_by(length) %>%
-      slice(x:(x+1)) %>%
-      group_by(length) %>%
-      summarise(min = min(gram_order),
-                max = max(gram_order)) %>%
-      split(.$length) %>%
-      map(~seq(.x$min, .x$max, 1) %>% as_data_frame) %>%
-      bind_rows(.id = "length") %>%
-      rename(gram_order = value) %>%
-      mutate(length = as.numeric(length))
-
-    df_tokens %>%
[…]

### 2019-06-09 surprisal/Wikipedia/get_wiki_df.py

@@ -71,20 +71,25 @@ def main(lang_name):
 	and stores the dataframe in a csv in the working directory.
 	'''
 	lang_prefix = LANG_DICT[lang_name]
-	url = "https://dumps.wikimedia.org/other/cirrussearch/20190318/" \
-		+ lang_prefix + "wiki-20190318-cirrussearch-content.json.gz"
+	url = "https://dumps.wikimedia.org/other/cirrussearch/20190520/" + \
+		lang_prefix + "wiki-20190520-cirrussearch-content.json.gz"
 	urllib.request.urlretrieve(url, "datafile")
 	subprocess.call(["wiki_extract/cirrus_extract.py", "datafile"])
 
+	all_dfs = pd.DataFrame(columns = ["gloss", "length"])
 	# find the number of article files we want to combine
-	directory = subprocess.check_output(["ls", "text/AA"]).decode("utf-8")
-	highest_filenum = max(int(max(re.findall("\d\d", directory))), 1)
-	upper_bound = min(highest_filenum, 50)
+	directory_names = subprocess.check_output(["ls", "text"]).decode("utf-8").split('\n')[:-1]
+	for name in directory_names:
+		directory = subprocess.check_output(["ls", "text/" + name]).decode("utf-8")
+		highest_filenum = max(int(max(re.findall("\d\d", directory))), 1)
+		upper_bound = min(highest_filenum, 99)
 
-	# construct the df
-	text = extract_texts(upper_bound)
-	df = get_sen_df(text)
-	df.to_csv("wiki_df.csv", index = False)
+		# construct the df
+		text = extract_texts(upper_bound)
+		df = get_sen_df(text)
+		all_dfs = pd.concat([all_dfs, df])
+
+	all_dfs.to_csv("wiki_df.csv", index = False)
 
 	# delete the text files and datafile
 	# may be several gigabytes large

### 2019-06-09 surprisal/Wikipedia/wiki_extract/cirrus_extract.py

@@ -103,7 +103,7 @@ class OutputSplitter(object):
     def write(self, data):
         #print(data)
         self.reserve(len(data))
-        self.file.write(data.encode('utf-8'))
+        self.file.write(data)
 
     def close(self):
         self.file.close()

### 2019-06-09 surprisal/Wikipedia/wiki_relative_surprisal.R

@@ -16,9 +16,11 @@ extract_wiki_tokens <- function(min_length, max_length, utterances) {
     mutate(word_order = 1:n()) %>%
     group_by(length, word_order, word) %>%
     summarise(n = n()) %>%
-    ungroup()
+    ungroup() %>%
+    mutate(word = as.numeric(as.factor(word)))
 
   sups <- tokens %>%
+    count(word) %>% 
     mutate(p = n/sum(n)) %>%
     mutate(s = -log(p)) %>%
     select(word, s)
@@ -73,13 +75,20 @@ relative_slopes <- function(df_tokens, pos_list) {
 
 NUM_SECTIONS = 5
 
+NUM_SENTENCES <- 500000
+
 LANGUAGE <- commandArgs(trailingOnly=TRUE)[1]
 
 system(paste("python3 get_wiki_df.py", LANGUAGE, sep = " "))
 
 # for_gs <- gs_title("wiki_surprisal_relative")
 
-df <- read_csv("wiki_df.csv")
+df <- read_csv("wiki_df.csv") 
+
+if (nrow(df) > NUM_SENTENCES) {
+  df <- df %>% 
+    sample_n(NUM_SENTENCES)
+}
 
 df_tokens <- extract_wiki_tokens(6, 10, df)
 
@@ -111,9 +120,9 @@ write.csv(all_data, file = "relative_unigrams.csv")
 # gs_edit_cells(for_gs, ws = "Sheet1", anchor = paste("A", nr + 2, sep=""), input = lang_slopes, byrow = TRUE)
 system("rm wiki_df.csv")
 
-slopes %>%
-  mutate(lower = estimate - 1.96 * std.error,
-         upper = estimate + 1.96 * std.error) %>%
-  ggplot(aes(x = cut, y = estimate, group = 1)) + 
-  geom_pointrange(aes(ymin = lower, ymax = upper)) + 
-  geom_line()
+# slopes %>%
+#   mutate(lower = estimate - 1.96 * std.error,
+#          upper = estimate + 1.96 * std.error) %>%
+#   ggplot(aes(x = cut, y = estimate, group = 1)) + 
+#   geom_pointrange(aes(ymin = lower, ymax = upper)) + 
+#   geom_line()

### 2019-06-18 surprisal/phoneme_surprisal.Rmd
phoneme

---
title: "Phoneme surprisal"
author: "Josef Klafka & Dan Yurovsky"
date: "6/13/2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidytext)
library(tidyverse)
library(childesr)
library(glue)

knitr::opts_chunk$set(echo = TRUE)
```

```{r espeak pipeline}
# all thanks to Mika Braginsky and Dan Yurovsky
lang_codes <- list(
  "Croatian" = "hr",
  "Danish" = "da",
  "English (American)" = "en-us",
  "French (Quebec)" = "fr",
  "Italian" = "it",
  "Norwegian" = "no",
  "Russian" = "ru",
  "Spanish (Mexican)" = "es",
  "Swedish" = "sv",
  "Turkish" = "tr", 
  "Mandarin" = "cmn"
)

get_ipa <- function(word) {
  system2("espeak", args = c("--ipa=3", "-v", "en", "-q", glue("'{word}'")),
          stdout = TRUE) %>%
    gsub("^ ", "", .) %>%
    gsub("[ˈˌ]", "", .)
}
```

```{r phonemes}
prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss

adult_utterances <- prov_utterances %>%
  filter(speaker_role != "Target_Child") %>% 
  mutate(utterance_id = 1:n()) %>% 
  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+")

words <- adult_utterances$word %>% 
  unique()

ipa <- words %>%
  lapply(get_ipa)

adult_utterances %<>%
  left_join(ipa) %>%
  unnest_tokens(phon, word, token = stringr::str_split, pattern = "", drop = F) %>%
  group_by(utterance_id, word) %>%
  mutate(phon_order = 1:n()) %>%
  mutate(word_length = max(phon_order)) %>%
  ungroup()

phon_probs <- adult_utterances %>% 
  count(phon) %>%
  ungroup() %>%
  mutate(p = n/sum(n))

letter_unigrams <- adult_utterances %>% 
  left_join(letter_probs) %>%
  mutate(s = -log(p)) %>%
  group_by(word_length, letter_order) %>%
  summarise(mean_s = mean(s))

letter_unigrams %>%
  filter(word_length %in% 3:11) %>%
  ggplot(aes(x = letter_order, y = mean_s)) + 
    facet_wrap(~word_length) + 
    geom_point() +
    geom_line() + 
    xlab("Letter Position") + 
    ylab("Mean surprisal")
```

```{r providence}
prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss

child_utterances <- prov_utterances %>%
  filter(speaker_role == "Target_Child") %>% 
  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n()) %>%
  mutate(ipa = get_ipa(word, "en-us"))

```

### 2019-06-18 surprisal/phoneme_surprisal.Rmd
ipas from text

@@
 ```{r espeak pipeline}
-# all thanks to Mika Braginsky and Dan Yurovsky
-lang_codes <- list(
-  "Croatian" = "hr",
-  "Danish" = "da",
-  "English (American)" = "en-us",
-  "French (Quebec)" = "fr",
-  "Italian" = "it",
-  "Norwegian" = "no",
-  "Russian" = "ru",
-  "Spanish (Mexican)" = "es",
-  "Swedish" = "sv",
-  "Turkish" = "tr", 
-  "Mandarin" = "cmn"
-)
-
 get_ipa <- function(word) {
-  system2("espeak", args = c("--ipa=3", "-v", "en", "-q", glue("'{word}'")),
+  system2("espeak", args = c("--ipa=3", "-v", "en-us", "-q", glue('"{word}"')),
           stdout = TRUE) %>%
@@
 
-words <- adult_utterances$word %>% 
-  unique()
+adult_utterances %>%
+  distinct(word) %>%
+  # rowwise() %>%
+  # mutate(ipa = get_ipa(word)) 
+  pull(word) %>%
+  write_lines("words.txt")
 
-ipa <- words %>%
-  lapply(get_ipa)
+phones <- read_lines("phones.txt") %>%
+  str_split(" ", simplify = F) %>%
+  unlist() %>%
+  unlist() %>%
+  enframe(name = NULL) %>%
+  filter(nchar(value) > 0)
 
-adult_utterances %<>%
+adult_phonemes <- adult_utterances %>%
   left_join(ipa) %>%
-  unnest_tokens(phon, word, token = stringr::str_split, pattern = "", drop = F) %>%
+  unnest_tokens(ipa, word, token = stringr::str_split, pattern = "_", drop = F) %>%
   group_by(utterance_id, word) %>%

### 2019-06-19 surprisal/phoneme_surprisal.Rmd
prelim POS tag surprisal results on Prov, progress on phoneme surprisal

@@
 title: "Phoneme surprisal"
-author: "Josef Klafka & Dan Yurovsky"
-date: "6/13/2019"
-output: pdf_document
+author: Josef Klafka and Dan Yurovsky 
+date: "`r Sys.Date()`"
+output: 
+  html_document:
+    toc: false
+    number_sections: false
+    theme: lumen
+    toc_float: false
+    code_folding: show 
+editor_options: 
+  chunk_output_type: console
 ---
@@
   mutate(utterance_id = 1:n()) %>% 
-  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+")
+  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]|[_+]+")
 
+system("rm words.txt phones.txt")
 adult_utterances %>%
   distinct(word) %>%
+  slice(1:10) %>%
+  # filter(word != "", word != "tʃutʃutʃu", word != "ɛː") %>%
   # rowwise() %>%
@@
 
-phones <- read_lines("phones.txt") %>%
+system('espeak -f words.txt --ipa=3 -v en-us -q --phonout="phones.txt"')
+system("paste -d , words.txt phones.txt > joined.txt")
+
+phones <- read_lines("joined.txt") %>%
   str_split(" ", simplify = F) %>%
@@
   filter(nchar(value) > 0)
+phones
 
@@
   group_by(utterance_id, word) %>%
-  mutate(phon_order = 1:n()) %>%
-  mutate(word_length = max(phon_order)) %>%
+  mutate(ipa_order = 1:n()) %>%
+  mutate(word_length = max(ipa_order)) %>%
   ungroup()
 
-phon_probs <- adult_utterances %>% 
-  count(phon) %>%
+ipa_probs <- adult_phonemes %>% 
+  count(ipa) %>%
   ungroup() %>%
@@
 
-letter_unigrams <- adult_utterances %>% 
-  left_join(letter_probs) %>%
+ipa_unigrams <- adult_utterances %>% 
+  left_join(ipa_probs) %>%
   mutate(s = -log(p)) %>%
-  group_by(word_length, letter_order) %>%
+  group_by(word_length, ipa_order) %>%
   summarise(mean_s = mean(s))
 
-letter_unigrams %>%
+ipa_unigrams %>%
   filter(word_length %in% 3:11) %>%
-  ggplot(aes(x = letter_order, y = mean_s)) + 
+  ggplot(aes(x = ipa_order, y = mean_s)) + 
     facet_wrap(~word_length) + 

### 2019-06-19 surprisal/pos_surprisal.Rmd
prelim POS tag surprisal results on Prov, progress on phoneme surprisal

---
title: Positional surprisal (1-3grams) in CHILDES corpora
author: Josef Klafka and Dan Yurovsky 
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(childesr)
library(ggthemes)
library(feather)

knitr::opts_chunk$set(echo = TRUE)
```

```{r pos unigrams}
prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
shiro_utterances <- get_utterances(corpus = "Shiro")

child_utterances <- prov_utterances %>%
  filter(speaker_role == "Target_Child") %>% 
  mutate(length = str_count(part_of_speech, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, part_of_speech, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n())

child_unigrams <- child_utterances %>%
  group_by(word) %>%
  count() %>%
  ungroup() %>%
  mutate(p = n / sum(n))

child_surprisals <- child_utterances %>%
  left_join(child_unigrams) %>%
  mutate(s = -log(p)) %>%
  group_by(length, word_order) %>%
  summarise(mean = mean(s),
            median = median(s),
            se = sd(s)/sqrt(n()))

child_surprisals %>%  
  filter(length %in% 4:9) %>%
  ggplot(aes(x = word_order, y = mean)) +
    facet_wrap(~length, scales = "free") + 
    ggtitle("Unigram POS surprisals") + 
    xlab("Word position") +
    ylab("Mean surprisal") +
    # scale_x_continuous(breaks = 1:7) + 
    geom_line(size = 1) +
    geom_pointrange(aes(ymin = mean - 1.96 * se, ymax = mean + 1.96 * se)) + 
    theme_classic(base_size = 10) + 
    scale_color_ptol() +
    theme(legend.position = "nones")
```

```{r pos bigrams}
child_bigrams_prep <- child_utterances %>% 
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  left_join(child_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = n.x / n.y) %>%
  select(lag_word, word, cond_p, n.x) %>%
  rename(n = n.x)

child_bigrams <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  ungroup() %>%
  left_join(child_unigrams) %>%
  left_join(child_bigrams_prep) %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  group_by(length, word_order) %>%
  summarise(mean_s = mean(s))

child_bigrams %>%  
  filter(length %in% 4:9) %>%
  ggplot(aes(x = word_order, y = mean_s)) +
    facet_wrap(~length, scales = "free") + 
    ggtitle("Bigram POS surprisals") + 
    xlab("Word position") +
    ylab("Mean surprisal") +
    # scale_x_continuous(breaks = 1:7) + 
    geom_line(size = 1) +
    geom_point() + 
    # geom_pointrange(aes(ymin = mean - 1.96 * se, ymax = mean + 1.96 * se)) + 
    theme_classic(base_size = 10) + 
    scale_color_ptol() +
    theme(legend.position = "nones")
```

```{r pos trigrams}
child_trigrams <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  group_by(lag_word2, lag_word1, word) %>%
  count() %>%
  filter(!is.na(lag_word1), !is.na(lag_word2)) %>%
  ungroup() %>%
  left_join(child_bigrams_prep, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
  mutate(tri_cond_p = n.x / n.y) %>%
  select(lag_word2, lag_word1, word, tri_cond_p)

trigram_surprisals_prep <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  left_join(child_trigrams) %>% 
  left_join(select(child_bigrams_prep, lag_word, word, cond_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
  left_join(child_unigrams)

trigram_surprisals <- trigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
                    ifelse(is.na(lag_word1), # check if not the second word
                           -log(p), # trigram if third or beyond
                           -log(cond_p)), # bigram if second word
                    -log(tri_cond_p))) %>% # unigram if first word
  group_by(length, word_order) %>%
  summarise(mean_s = mean(s)) %>% 
  filter(word_order <= length)

trigram_surprisals %>%  
  filter(length %in% 4:9) %>%
  ggplot(aes(x = word_order, y = mean_s)) +
    facet_wrap(~length, scales = "free") +
    ggtitle("Trigram POS surprisals") + 
    xlab("Word position") +
    ylab("Mean surprisal") +
    # scale_x_continuous(breaks = 1:7) + 
    geom_line(size = 1) +
    geom_point() + 
    # geom_pointrange(aes(ymin = mean - 1.96 * se, ymax = mean + 1.96 * se)) + 
    theme_classic(base_size = 10) + 
    scale_color_ptol() +
    theme(legend.position = "nones")
```

### 2019-06-19 surprisal/phoneme_surprisal.Rmd
progress on phoneme surprisal setup

@@
 
-system("rm words.txt phones.txt")
+system("rm words.txt phones.txt splitphones.txt joined.csv")
 adult_utterances %>%
   distinct(word) %>%
-  slice(1:10) %>%
-  # filter(word != "", word != "tʃutʃutʃu", word != "ɛː") %>%
+  filter(word != "", word != "tʃutʃutʃu", word != "ɛː", word != "ˈɛ_s") %>%
   # rowwise() %>%
@@
 
-system('espeak -f words.txt --ipa=3 -v en-us -q --phonout="phones.txt"')
-system("paste -d , words.txt phones.txt > joined.txt")
+system("espeak -f words.txt --ipa=3 -v en-us -q --phonout='phones.txt'") #get the ipa of each word
+system("cat phones.txt | sed -E 's/ /|/g' | tr '|' '\n' | tail -n +2 > splitphones.txt") #put each ipa word onto its own line in a file called splitphones.txt
+system("rm phones.txt")
+system("sed '/^$/d' splitphones.txt > phones.txt") #remove empty lines in splitphones
+system("paste -d , words.txt phones.txt > joined.csv") #make csv with words and their ipa paired
 
-phones <- read_lines("joined.txt") %>%
-  str_split(" ", simplify = F) %>%
-  unlist() %>%
-  unlist() %>%
-  enframe(name = NULL) %>%
-  filter(nchar(value) > 0)
-phones
+phones <- read.csv("joined.csv", 
+                   header = F, 
+                   col.names = c("words", "ipa"))
+# phones <- read_lines("joined.txt") %>%
+#   str_split(" ", simplify = F) %>%
+#   unlist() %>%
+#   unlist() %>%
+#   enframe(name = NULL) %>%
+#   filter(nchar(value) > 0)
 
 adult_phonemes <- adult_utterances %>%
-  left_join(ipa) %>%
+  left_join(phones) %>%
   unnest_tokens(ipa, word, token = stringr::str_split, pattern = "_", drop = F) %>%

### 2019-06-20 surprisal/phoneme_surprisal.Rmd
providence unigram phonemes pipeline done

@@
 library(glue)
+library(feather)
+library(here)
 
@@
 
-```{r phonemes}
+```{r prov phonemes}
 prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
@@
   distinct(word) %>%
-  filter(word != "", word != "tʃutʃutʃu", word != "ɛː", word != "ˈɛ_s") %>%
+  filter(word != "", word != "tʃutʃutʃu", word != "ɛː", word != "lunchroom", word != "dryad's", word != "vi") %>%
   # rowwise() %>%
@@
                    header = F, 
-                   col.names = c("words", "ipa"))
+                   col.names = c("word", "ipa"))
 # phones <- read_lines("joined.txt") %>%
@@
 adult_phonemes <- adult_utterances %>%
+  left_join(phones) %>%
+  filter(!is.na(ipa)) %>%
+  unnest_tokens(phone, ipa, token = stringr::str_split, pattern = "_", drop = F) %>%
+  group_by(utterance_id, word) %>%
+  mutate(phone_order = 1:n()) %>%
+  mutate(word_length = max(phone_order)) %>%
+  ungroup() %>%
+  select(utterance_id, word, ipa, phone, phone_order, word_length)
+
+phone_probs <- adult_phonemes %>% 
+  count(phone) %>%
+  ungroup() %>%
+  mutate(p = n/sum(n))
+
+ipa_unigrams <- adult_phonemes %>% 
+  left_join(phone_probs) %>%
+  mutate(s = -log(p)) %>%
+  group_by(word_length, phone_order) %>%
+  summarise(mean_s = mean(s))
+
+ipa_unigrams %>%
+  filter(word_length %in% 4:9) %>%
+  ggplot(aes(x = phone_order, y = mean_s)) + 
+    facet_wrap(~word_length) + 
+    geom_point() +
+    geom_line() + 
+    xlab("Letter Position") + 
+    ylab("Mean surprisal") 
+```
+
+```{r bnc}
+bnc <- read_feather(here("../bnc/bnc10.feather"))
+
+system("rm words.txt phones.txt splitphones.txt joined.csv")
+bnc %>%
+  distinct(word) %>%
+  filter(word != "") %>%
+  pull(word) %>%
+  grep(pattern = "[0-9]+", value = T, invert = T) %>% 
+  write_lines("words.txt")
+
+system("espeak -f words.txt --ipa=3 -v en-us -q --phonout='phones.txt'") #get the ipa of each word
+system("cat phones.txt | sed -E 's/ /|/g' | tr '|' '\n' | tail -n +2 > splitphones.txt") #put each ipa word onto its own line in a file called splitphones.txt
+system("rm phones.txt")
+system("sed '/^$/d' splitphones.txt > phones.txt") #remove empty lines in splitphones
+system("paste -d , words.txt phones.txt > joined.csv") #make csv with words and their ipa paired
+
+phones <- read.csv("joined.csv", 
+                   header = F, 
+                   col.names = c("words", "ipa"))
+phones
+
+bnc_phonemes <- bnc %>%
   left_join(phones) %>%

### 2019-06-20 surprisal/pos_surprisal.Rmd
providence unigram phonemes pipeline done

@@
 ---
-title: Positional surprisal (1-3grams) in CHILDES corpora
+title: Positional surprisal (1-3grams) in adultES corpora
 author: Josef Klafka and Dan Yurovsky 
@@
 prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
-shiro_utterances <- get_utterances(corpus = "Shiro")
+zhou_utterances <- get_utterances(corpus = "ZhouDinner")
+okayama_utterances <- get_utterances(corpus = "Okayama")
 
-child_utterances <- prov_utterances %>%
-  filter(speaker_role == "Target_Child") %>% 
+adult_utterances <- prov_utterances %>%
+  filter(speaker_role != "Target_adult") %>% 
   mutate(length = str_count(part_of_speech, pattern = "[ +]+") + 1) %>%
@@
 
-child_unigrams <- child_utterances %>%
+adult_unigrams <- adult_utterances %>%
   group_by(word) %>%
@@
 
-child_surprisals <- child_utterances %>%
-  left_join(child_unigrams) %>%
+adult_surprisals <- adult_utterances %>%
+  left_join(adult_unigrams) %>%
   mutate(s = -log(p)) %>%
@@
 
-child_surprisals %>%  
+adult_surprisals %>%  
   filter(length %in% 4:9) %>%
@@
     facet_wrap(~length, scales = "free") + 
-    ggtitle("Unigram POS surprisals") + 
+    ggtitle("Adult unigram POS surprisals") + 
     xlab("Word position") +
@@
 ```{r pos bigrams}
-child_bigrams_prep <- child_utterances %>% 
+adult_bigrams_prep <- adult_utterances %>% 
   group_by(utterance_id) %>%
@@
   ungroup() %>%
-  left_join(child_unigrams, by = c("lag_word" = "word")) %>%
+  left_join(adult_unigrams, by = c("lag_word" = "word")) %>%
   mutate(cond_p = n.x / n.y) %>%
@@
 
-child_bigrams <- child_utterances %>%
+adult_bigrams <- adult_utterances %>%
   group_by(utterance_id) %>%
@@
   ungroup() %>%
-  left_join(child_unigrams) %>%
-  left_join(child_bigrams_prep) %>%
+  left_join(adult_unigrams) %>%
+  left_join(select(adult_bigrams_prep, -n), by = c("lag_word" = "lag_word", "word" = "word")) %>%
   mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
@@
 
-child_bigrams %>%  
+adult_bigrams %>%  
   filter(length %in% 4:9) %>%
@@
     facet_wrap(~length, scales = "free") + 
-    ggtitle("Bigram POS surprisals") + 
+    ggtitle("Adult bigram POS surprisals") + 
     xlab("Word position") +
@@
 ```{r pos trigrams}
-child_trigrams <- child_utterances %>%
+adult_trigrams <- adult_utterances %>%
   group_by(utterance_id) %>%
@@
   ungroup() %>%
-  left_join(child_bigrams_prep, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
+  left_join(adult_bigrams_prep, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
   mutate(tri_cond_p = n.x / n.y) %>%
@@
 
-trigram_surprisals_prep <- child_utterances %>%
+trigram_surprisals_prep <- adult_utterances %>%
   group_by(utterance_id) %>%
@@
   mutate(lag_word2 = lag(lag_word1)) %>%
-  left_join(child_trigrams) %>% 
-  left_join(select(child_bigrams_prep, lag_word, word, cond_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
-  left_join(child_unigrams)
+  left_join(adult_trigrams) %>% 
+  left_join(select(adult_bigrams_prep, lag_word, word, cond_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
+  left_join(adult_unigrams)
 
@@
     facet_wrap(~length, scales = "free") +
-    ggtitle("Trigram POS surprisals") + 
+    ggtitle("Adult trigram POS surprisals") + 
     xlab("Word position") +

### 2019-06-21 surprisal/phoneme_surprisal.Rmd
interim information analysis

@@
 
-```{r prov phonemes}
+```{r prov phonemes, eval = F}
 prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
@@
   mutate(utterance_id = 1:n()) %>% 
-  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]|[_+]+")
+  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]|[_+]+") %>%
+  group_by(utterance_id) %>%
+  mutate(word_id = 1:n())
 
@@
 system("paste -d , words.txt phones.txt > joined.csv") #make csv with words and their ipa paired
-
-phones <- read.csv("joined.csv", 
-                   header = F, 
-                   col.names = c("word", "ipa"))
 # phones <- read_lines("joined.txt") %>%
@@
 #   filter(nchar(value) > 0)
+```
+
+```{r read_phonemes}
+phones <- read_csv("joined.csv",
+                   col_names = c("word", "ipa"))
 
@@
   filter(!is.na(ipa)) %>%
-  unnest_tokens(phone, ipa, token = stringr::str_split, pattern = "_", drop = F) %>%
-  group_by(utterance_id, word) %>%
+  unnest_tokens(phone, ipa, token = stringr::str_split, 
+                pattern = "_+", drop = F) %>%
+  group_by(utterance_id, word_id) %>%
   mutate(phone_order = 1:n()) %>%
@@
   ungroup() %>%
-  select(utterance_id, word, ipa, phone, phone_order, word_length)
+  select(utterance_id, word_id, word, ipa, phone, phone_order, word_length)
+
+adult_phonemes %>%
+  filter(phone_order == 1) %>%
+  ggplot(aes(x = word_length)) + 
+  geom_histogram(fill = "white", color = "black") + 
+  theme_classic()
 
@@
 ipa_unigrams %>%
-  filter(word_length %in% 4:9) %>%
+  filter(word_length %in% 1:5) %>%
   ggplot(aes(x = phone_order, y = mean_s)) + 
@@
     ylab("Mean surprisal") 
+```
+
+```{r}
+four_lengths <- adult_phonemes %>%
+  filter(word_length == 3) %>%
+  group_by(utterance_id, word_id, phone_order)
+
+four_words <- four_lengths %>%
+  ungroup() %>%
+  distinct(word, phone_order, phone) %>%
+  group_by(word) %>%
+  nest()
+
+word1 <- four_words %>%
+  slice(1) %>%
+  unnest() %>%
+  slice(c(2,3)) %>%
+  select(-word)
+
+word1 %>%
+  spread(phone_order, phone) %>%
+  left_join(four_lengths %>%
+  spread(phone_order, phone)) %>%
+  distinct(utterance_id, word_id, word) %>%
+  group_by(word) %>%
+  summarise(n = n())
+
 ```

### 2019-06-21 surprisal/phoneme_surprisal.Rmd
hold out entropy

@@
 library(here)
+library(entropy)
+library(tidyboot)
+library(janitor)
+library(arrangements)
 
@@
                 pattern = "_+", drop = F) %>%
+  filter(phone != "") %>%
   group_by(utterance_id, word_id) %>%
@@
   ggplot(aes(x = word_length)) + 
-  geom_histogram(fill = "white", color = "black") + 
+  geom_histogram(fill = "white", color = "black", binwidth = 1) + 
   theme_classic()
@@
 
-```{r}
-four_lengths <- adult_phonemes %>%
-  filter(word_length == 3) %>%
-  group_by(utterance_id, word_id, phone_order)
+```{r leave-out-surprisal}
+hold_out_entropy <- function(df) {
+  spread_df <- df %>%
+    spread(phone_order, phone) %>%
+    clean_names()
 
-four_words <- four_lengths %>%
+  phone_length <- max(df$phone_order)
+
+  lens <- paste0("x", 1:phone_length)
+
+  boot_entropy <- function(group) {
+    spread_df %>%
+      group_by_at(group) %>%
+      summarise(n = n()) %>%
+      summarise(entropy = entropy(n)) %>%
       ungroup() %>%
-  distinct(word, phone_order, phone) %>%
-  group_by(word) %>%
-  nest()
+      tidyboot_mean(entropy)
+  }
 
-word1 <- four_words %>%
-  slice(1) %>%
+  groupings <- combinations(lens, phone_length -1, 
+                            layout = "list")
+
+  groupings <- map(seq(length(lens),1,-1),
+                   function(x) c(groupings[[x]], lens[x]))
+
+  map_df(groupings, boot_entropy) %>%
+    mutate(phone_order = seq(phone_length, 1, -1))
+
+}
+
+split_lengths <- adult_phonemes %>%
+  group_by(word_length) %>%
+  nest() %>%
+  filter(word_length > 1) %>%
+  filter(word_length %in% 1:6 ) %>%
+  mutate(entropy = map(data, hold_out_entropy))
+
+split_lengths %>%
+  select(-data) %>%
   unnest() %>%
-  slice(c(2,3)) %>%
-  select(-word)
-
-word1 %>%
-  spread(phone_order, phone) %>%
-  left_join(four_lengths %>%
-  spread(phone_order, phone)) %>%
-  distinct(utterance_id, word_id, word) %>%
-  group_by(word) %>%
-  summarise(n = n())
-
+  ggplot(aes(x = phone_order, y = empirical_stat)) + 
+  facet_wrap(~ word_length, scales = "free") +
+  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) + 
+  theme_classic() + 
+  geom_line()
 ```

### 2019-06-24 surprisal/new_phonemes.Rmd
phoneme write-up stuff

---
title: "Phoneme surprisal"
author: Josef Klafka and Dan Yurovsky 
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
editor_options: 
  chunk_output_type: console
---
Briefly what we're doing is splitting each word in the corpus (Providence from CHILDES and the Switchboard corpus of adult telephone conversations) up into its constituent phones. We then calculate how common each phone is, and compute the surprisal of each phone--a measure of how much information is in the phone.

We split up the words in the corpus by their length in number of phones. For each phone index or position in words of each length, we calculate the average information (suprisal) in the phones in that position. For example, we take all the words with four phones in the Providence corpus, and then calculate the average information in the first phone of those words, the average information in the second phone, and so on.

```{r setup, include=FALSE}
library(tidytext)
library(tidyverse)
library(childesr)
library(glue)
library(feather)
library(here)

knitr::opts_chunk$set(echo = TRUE)
```

```{r prov phonemes}
prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss

adult_utterances <- prov_utterances %>%
  filter(speaker_role != "Target_Child") %>% 
  mutate(utterance_id = 1:n()) %>% 
  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]|[_+]+")

system("rm words.txt phones.txt splitphones.txt joined.csv")
adult_utterances %>%
  distinct(word) %>%
  filter(word != "", word != "tʃutʃutʃu", word != "ɛː", word != "lunchroom", word != "dryad's", word != "vi") %>%
  pull(word) %>%
  write_lines("words.txt")

system("espeak -f words.txt --ipa=3 -v en-us -q --phonout='phones.txt'") #get the ipa of each word
system("cat phones.txt | sed -E 's/ /|/g' | tr '|' '\n' | tail -n +2 > splitphones.txt") #put each ipa word onto its own line in a file called splitphones.txt
system("rm phones.txt")
system("sed '/^$/d' splitphones.txt > phones.txt") #remove empty lines in splitphones
system("paste -d , words.txt phones.txt > joined.csv") #make csv with words and their ipa paired

phones <- read.csv("joined.csv", 
                   header = F, 
                   col.names = c("word", "ipa"))

adult_phonemes <- adult_utterances %>%
  left_join(phones) %>%
  filter(!is.na(ipa)) %>%
  unnest_tokens(phone, ipa, token = stringr::str_split, pattern = "_", drop = F) %>%
  group_by(utterance_id, word) %>%
  mutate(phone_order = 1:n()) %>%
  mutate(word_length = max(phone_order)) %>%
  ungroup() %>%
  select(utterance_id, word, ipa, phone, phone_order, word_length)

phone_probs <- adult_phonemes %>% 
  count(phone) %>%
  ungroup() %>%
  mutate(p = n/sum(n))

ipa_unigrams <- adult_phonemes %>% 
  left_join(phone_probs) %>%
  mutate(s = -log(p)) %>%
  group_by(word_length, phone_order) %>%
  summarise(mean_s = mean(s))

ipa_unigrams %>%
  filter(word_length %in% 4:9) %>%
  ggplot(aes(x = phone_order, y = mean_s)) + 
    facet_wrap(~word_length) + 
    geom_point() +
    geom_line() + 
    xlab("Letter Position") + 
    ylab("Mean surprisal") 
```

```{r switchboard}
switchboard <- read_feather(here("../switchboard/switchboard.feather")) %>%
  mutate(utterance_id = 1:n()) %>% 
  unnest_tokens(word, value, token = stringr::str_split, pattern = "[ +]|[-+]+")

system("rm words.txt phones.txt splitphones.txt joined.csv")
switchboard %>%
  distinct(word) %>%
  filter(word != "", word != "ii") %>%
  pull(word) %>%
  grep(pattern = "[0-9]+", value = T, invert = T) %>%
  write_lines("words.txt")

system("espeak -f words.txt --ipa=3 -v en-us -q --phonout='phones.txt'") #get the ipa of each word 
system("cat phones.txt | sed -E 's/ /|/g' | tr '|' '\n' | tail -n +2 > splitphones.txt") #put each ipa word onto its own line in a file called splitphones.txt
system("rm phones.txt")
system("sed '/^$/d' splitphones.txt > phones.txt") #remove empty lines in splitphones
system("paste -d , words.txt phones.txt > joined.csv") #make csv with words and their ipa paired

switch_phones <- read.csv("joined.csv", 
                   header = F, 
                   col.names = c("word", "ipa"))
tail(switch_phones)

adult_phonemes <- adult_utterances %>%
  left_join(phones) %>%
  filter(!is.na(ipa)) %>%
  unnest_tokens(phone, ipa, token = stringr::str_split, pattern = "_", drop = F) %>%
  group_by(utterance_id, word) %>%
  mutate(phone_order = 1:n()) %>%
  mutate(word_length = max(phone_order)) %>%
  ungroup() %>%
  select(utterance_id, word, ipa, phone, phone_order, word_length)

phone_probs <- adult_phonemes %>% 
  count(phone) %>%
  ungroup() %>%
  mutate(p = n/sum(n))

ipa_unigrams <- adult_phonemes %>% 
  left_join(phone_probs) %>%
  mutate(s = -log(p)) %>%
  group_by(word_length, phone_order) %>%
  summarise(mean_s = mean(s))

ipa_unigrams %>%
  filter(word_length %in% 4:9) %>%
  ggplot(aes(x = phone_order, y = mean_s)) + 
    facet_wrap(~word_length) + 
    geom_point() +
    geom_line() + 
    xlab("Letter Position") + 
    ylab("Mean surprisal")
```

```{r bnc}
bnc <- read_feather(here("../bnc/bnc10.feather"))

system("rm words.txt phones.txt splitphones.txt joined.csv")
bnc %>%
  distinct(word) %>%
  filter(word != "") %>%
  pull(word) %>%
  grep(pattern = "[0-9]+", value = T, invert = T) %>% 
  write_lines("words.txt")

system("espeak -f words.txt --ipa=3 -v en-us -q --phonout='phones.txt'") #get the ipa of each word
system("cat phones.txt | sed -E 's/ /|/g' | tr '|' '\n' | tail -n +2 > splitphones.txt") #put each ipa word onto its own line in a file called splitphones.txt
system("rm phones.txt")
system("sed '/^$/d' splitphones.txt > phones.txt") #remove empty lines in splitphones
system("paste -d , words.txt phones.txt > joined.csv") #make csv with words and their ipa paired

phones <- read.csv("joined.csv", 
                   header = F, 
                   col.names = c("words", "ipa"))
phones

bnc_phonemes <- bnc %>%
  left_join(phones) %>%
  unnest_tokens(ipa, word, token = stringr::str_split, pattern = "_", drop = F) %>%
  group_by(utterance_id, word) %>%
  mutate(ipa_order = 1:n()) %>%
  mutate(word_length = max(ipa_order)) %>%
  ungroup()

ipa_probs <- adult_phonemes %>% 
  count(ipa) %>%
  ungroup() %>%
  mutate(p = n/sum(n))

ipa_unigrams <- adult_utterances %>% 
  left_join(ipa_probs) %>%
  mutate(s = -log(p)) %>%
  group_by(word_length, ipa_order) %>%
  summarise(mean_s = mean(s))

ipa_unigrams %>%
  filter(word_length %in% 3:11) %>%
  ggplot(aes(x = ipa_order, y = mean_s)) + 
    facet_wrap(~word_length) + 
    geom_point() +
    geom_line() + 
    xlab("Letter Position") + 
    ylab("Mean surprisal")
```

### 2019-06-24 surprisal/phoneme_surprisal.Rmd
phoneme write-up stuff

@@
 
-```{r espeak pipeline}
-get_ipa <- function(word) {
-  system2("espeak", args = c("--ipa=3", "-v", "en-us", "-q", glue('"{word}"')),
-          stdout = TRUE) %>%
-    gsub("^ ", "", .) %>%
-    gsub("[ˈˌ]", "", .)
-}
+```{r prov phonemes, eval = F, include = F}
+system("rm words.txt phones.txt splitphones.txt joined.csv")
+adult_utterances %>%
+  distinct(word) %>%
+  filter(word != "", word != "tʃutʃutʃu", word != "ɛː", word != "lunchroom", word != "dryad's", word != "vi") %>%
+  pull(word) %>%
+  write_lines("words.txt")
+
+system("espeak -f words.txt --ipa=3 -v en-us -q --phonout='phones.txt'") #get the ipa of each word
+system("cat phones.txt | sed -E 's/ /|/g' | tr '|' '\n' | tail -n +2 > splitphones.txt") #put each ipa word onto its own line in a file called splitphones.txt
+system("rm phones.txt")
+system("sed '/^$/d' splitphones.txt > phones.txt") #remove empty lines in splitphones
+system("paste -d , words.txt phones.txt > joined.csv") #make csv with words and their ipa paired
+
 ```
 
-```{r prov phonemes, eval = F}
+```{r read_phonemes, include = F}
 prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
@@
 
-system("rm words.txt phones.txt splitphones.txt joined.csv")
-adult_utterances %>%
-  distinct(word) %>%
-  filter(word != "", word != "tʃutʃutʃu", word != "ɛː", word != "lunchroom", word != "dryad's", word != "vi") %>%
-  # rowwise() %>%
-  # mutate(ipa = get_ipa(word)) 
-  pull(word) %>%
-  write_lines("words.txt")
-
-system("espeak -f words.txt --ipa=3 -v en-us -q --phonout='phones.txt'") #get the ipa of each word
-system("cat phones.txt | sed -E 's/ /|/g' | tr '|' '\n' | tail -n +2 > splitphones.txt") #put each ipa word onto its own line in a file called splitphones.txt
-system("rm phones.txt")
-system("sed '/^$/d' splitphones.txt > phones.txt") #remove empty lines in splitphones
-system("paste -d , words.txt phones.txt > joined.csv") #make csv with words and their ipa paired
-# phones <- read_lines("joined.txt") %>%
-#   str_split(" ", simplify = F) %>%
-#   unlist() %>%
-#   unlist() %>%
-#   enframe(name = NULL) %>%
-#   filter(nchar(value) > 0)
-```
-
-```{r read_phonemes}
 phones <- read_csv("joined.csv",
@@
   summarise(mean_s = mean(s))
+```
 
+Briefly, what we're doing is splitting each word in the Providence corpus from CHILDES up into its constituent phones, i.e.
+the sounds that make up the word.
+We then calculate how common each phone is across the entire corpus.
+From this frequency measure, we compute the surprisal of each phone, which is inversely proportional to the frequency of the phone.
+The less common a phone is, the more difficult it is to process the phone.
+
+We split up the words in the corpus by their length in number of phones--for example, the word "food" has three phones: one for the "f", one for the long "oo", and one for the "d".
+We would take all words like this that have three phones.
+For each phone position  within the words of that length, we calculate the average surprisal/information in that position.
+For example, looking at all of the words with three phones, we take all of the first phones and calculate how much information is in that position on average.
+In other words, how difficult is it to process the first sound in a word, compared to the other sounds in a word?
+
+```{r plot unigrams}
 ipa_unigrams %>%
-  filter(word_length %in% 1:5) %>%
+  filter(word_length %in% 2:6) %>%
   ggplot(aes(x = phone_order, y = mean_s)) +
@@
     geom_line() +
-    xlab("Letter Position") + 
+    xlab("Phone Position") +
     ylab("Mean surprisal")
@@
 
-```{r leave-out-surprisal}
+In the plot above
+
+```{r leave-out-surprisal, include = F}
 hold_out_entropy <- function(df) {
@@
 }
+```
 
+For this next analysis, we took all words with 2, 3, 4, 5 or 6 phones.
+For each word length, we held out the phones in one phone position.
+For example, in all words with 4 phones, we held out the first phone.
+We then computed the entropy, or amount of variation, in the remaining words.
+The higher the entropy when holding out phones in a given position, the more information is contained in that phone on average.
+
+```{r show leave-out-surprisal}
 split_lengths <- adult_phonemes %>%
@@
   theme_classic() + 
-  geom_line()
+  geom_line() + 
+  xlab("Phone position") + 
+  ylab("Entropy")
 ```
 
-```{r bnc}
-bnc <- read_feather(here("../bnc/bnc10.feather"))
-
-system("rm words.txt phones.txt splitphones.txt joined.csv")
-bnc %>%
-  distinct(word) %>%
-  filter(word != "") %>%
-  pull(word) %>%
-  grep(pattern = "[0-9]+", value = T, invert = T) %>% 
-  write_lines("words.txt")
-
-system("espeak -f words.txt --ipa=3 -v en-us -q --phonout='phones.txt'") #get the ipa of each word
-system("cat phones.txt | sed -E 's/ /|/g' | tr '|' '\n' | tail -n +2 > splitphones.txt") #put each ipa word onto its own line in a file called splitphones.txt
-system("rm phones.txt")
-system("sed '/^$/d' splitphones.txt > phones.txt") #remove empty lines in splitphones
-system("paste -d , words.txt phones.txt > joined.csv") #make csv with words and their ipa paired
-
-phones <- read.csv("joined.csv", 
-                   header = F, 
-                   col.names = c("words", "ipa"))
-phones
-
-bnc_phonemes <- bnc %>%
-  left_join(phones) %>%
-  unnest_tokens(ipa, word, token = stringr::str_split, pattern = "_", drop = F) %>%
-  group_by(utterance_id, word) %>%
-  mutate(ipa_order = 1:n()) %>%
-  mutate(word_length = max(ipa_order)) %>%
-  ungroup()
-
-ipa_probs <- adult_phonemes %>% 
-  count(ipa) %>%
-  ungroup() %>%
-  mutate(p = n/sum(n))
-
-ipa_unigrams <- adult_utterances %>% 
-  left_join(ipa_probs) %>%
-  mutate(s = -log(p)) %>%
-  group_by(word_length, ipa_order) %>%
-  summarise(mean_s = mean(s))
-
-ipa_unigrams %>%
-  filter(word_length %in% 3:11) %>%
-  ggplot(aes(x = ipa_order, y = mean_s)) + 
-    facet_wrap(~word_length) + 
-    geom_point() +
-    geom_line() + 
-    xlab("Letter Position") + 
-    ylab("Mean surprisal")
-```
-
-```{r providence}
-prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
-
-child_utterances <- prov_utterances %>%
-  filter(speaker_role == "Target_Child") %>% 
-  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
-  mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
-  group_by(utterance_id) %>%
-  mutate(word_order = 1:n()) %>%
-  mutate(ipa = get_ipa(word, "en-us"))
-
-```
+We see that word-initial phones tend to have more information in them than word-medial or word-final phones, so less is lost by dropping the final sounds in a word than in dropping the initial sounds.

### 2019-06-25 surprisal/letter_surprisal.Rmd
prov letter surprisals

@@
 ---
-title: Wikipedia Text Mining and Surprisal Analysis
+title: Letter-level (orthographic) surprisal analysis
 author: Josef Klafka and Dan Yurovsky
@@
 
-```{r prov}
+```{r set up corpus}
 prov_utterances <- get_utterances(corpus = "Providence")
@@
   unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
+  group_by(utterance_id) %>%
+  mutate(word_id = 1:n()) %>%
+  ungroup() %>%
   unnest_tokens(letter, word, token = stringr::str_split, pattern = "", drop = F) %>%
-  group_by(utterance_id, word) %>%
+  group_by(utterance_id, word_id) %>%
   mutate(letter_order = 1:n()) %>%
@@
   ungroup()
+```
 
+```{r unigrams}
 letter_probs <- adult_utterances %>% 
@@
   summarise(mean_s = mean(s))
+```
 
+```{r childes bigrams}
+bigram_probs <- adult_utterances %>% 
+  group_by(utterance_id) %>%
+  mutate(lag_letter = lag(letter)) %>%
+  group_by(lag_letter, letter) %>%
+  count() %>%
+  filter(!is.na(lag_letter)) %>%
+  ungroup() %>%
+  left_join(letter_probs, by = c("lag_letter" = "letter")) %>%
+  mutate(cond_p = n.x / n.y) %>%
+  select(lag_letter, letter, cond_p, n.x) %>%
+  rename(n = n.x)
+
+letter_bigrams <- adult_utterances %>%
+  group_by(utterance_id, word_id) %>%
+  mutate(lag_letter = lag(letter)) %>%
+  ungroup() %>%
+  left_join(bigram_probs, by = c("lag_letter" = "lag_letter", "letter" = "letter")) %>%
+  left_join(letter_probs, by = c("letter" = "letter")) %>%
+  mutate(s = ifelse(is.na(lag_letter), -log(p), -log(cond_p))) %>%
+  group_by(word_length, letter_order) %>%
+  summarise(mean_s = mean(s)) 
+```
+
+```{r childes trigrams}
+trigram_probs <- adult_utterances %>%
+  group_by(utterance_id, word_id) %>%
+  mutate(lag_letter1 = lag(letter)) %>%
+  mutate(lag_letter2 = lag(lag_letter1)) %>%
+  group_by(lag_letter2, lag_letter1, letter) %>%
+  count() %>%
+  filter(!is.na(lag_letter1), !is.na(lag_letter2)) %>%
+  ungroup() %>%
+  left_join(bigram_probs, by = c("lag_letter2" = "lag_letter", "lag_letter1" = "letter")) %>%
+  mutate(tri_cond_p = n.x / n.y) %>%
+  select(-cond_p, -n.y) %>%
+  rename(n = n.x)
+
+letter_trigrams <- adult_utterances %>%
+  group_by(utterance_id, word_id) %>%
+  mutate(lag_letter1 = lag(letter)) %>%
+  mutate(lag_letter2 = lag(lag_letter1)) %>%
+  left_join(trigram_probs, by = c("lag_letter2" = "lag_letter2", "lag_letter1" = "lag_letter1", "letter" = "letter")) %>% 
+  left_join(bigram_probs, by = c("lag_letter1" = "lag_letter", "letter" = "letter")) %>% 
+  left_join(letter_probs, by = c("letter" = "letter")) %>%
+  mutate(s = ifelse(is.na(lag_letter2), # check if it's not the first word
+                    ifelse(is.na(lag_letter1), # check if not the second word
+                           -log(p), # trigram if third or beyond
+                           -log(cond_p)), # bigram if second word
+                    -log(tri_cond_p))) %>% # unigram if first word
+  group_by(word_length, letter_order) %>%
+  summarise(mean_s = mean(s)) %>% 
+  filter(letter_order <= word_length)
+```
+
+```{r plot results}
 letter_unigrams %>%
@@
     ylab("Mean surprisal")
-
 ```

### 2019-06-25 surprisal/new_phonemes.Rmd
prov letter surprisals

@@
 ```
-
-```{r bnc}
-bnc <- read_feather(here("../bnc/bnc10.feather"))
-
-system("rm words.txt phones.txt splitphones.txt joined.csv")
-bnc %>%
-  distinct(word) %>%
-  filter(word != "") %>%
-  pull(word) %>%
-  grep(pattern = "[0-9]+", value = T, invert = T) %>% 
-  write_lines("words.txt")
-
-system("espeak -f words.txt --ipa=3 -v en-us -q --phonout='phones.txt'") #get the ipa of each word
-system("cat phones.txt | sed -E 's/ /|/g' | tr '|' '\n' | tail -n +2 > splitphones.txt") #put each ipa word onto its own line in a file called splitphones.txt
-system("rm phones.txt")
-system("sed '/^$/d' splitphones.txt > phones.txt") #remove empty lines in splitphones
-system("paste -d , words.txt phones.txt > joined.csv") #make csv with words and their ipa paired
-
-phones <- read.csv("joined.csv", 
-                   header = F, 
-                   col.names = c("words", "ipa"))
-phones
-
-bnc_phonemes <- bnc %>%
-  left_join(phones) %>%
-  unnest_tokens(ipa, word, token = stringr::str_split, pattern = "_", drop = F) %>%
-  group_by(utterance_id, word) %>%
-  mutate(ipa_order = 1:n()) %>%
-  mutate(word_length = max(ipa_order)) %>%
-  ungroup()
-
-ipa_probs <- adult_phonemes %>% 
-  count(ipa) %>%
-  ungroup() %>%
-  mutate(p = n/sum(n))
-
-ipa_unigrams <- adult_utterances %>% 
-  left_join(ipa_probs) %>%
-  mutate(s = -log(p)) %>%
-  group_by(word_length, ipa_order) %>%
-  summarise(mean_s = mean(s))
-
-ipa_unigrams %>%
-  filter(word_length %in% 3:11) %>%
-  ggplot(aes(x = ipa_order, y = mean_s)) + 
-    facet_wrap(~word_length) + 
-    geom_point() +
-    geom_line() + 
-    xlab("Letter Position") + 
-    ylab("Mean surprisal")
-```

### 2019-06-25 surprisal/wiki_surprisal.Rmd
prov letter surprisals

@@
 ```{r word order}
-wkpd <- read.csv(here("Data/relative_unigrams.csv"))
+wkpd <- read.csv(here("Data/wikipedia_0context.csv"))
 LANGUAGES <- wkpd$Language
 
-FEATURES <- c(82:97)
+FEATURES <- c(1:144)
 for (i in c(1:length(FEATURES))) {
@@
 
+# word order
+# wo_wf <- wf %>% 
+#   filter(language %in% LANGUAGES) %>%
+#   select(language, `82A`:`97A`) %>%
+#   select(-`84A`)
+
+# nominal categories
 wo_wf <- wf %>%
   filter(language %in% LANGUAGES) %>%
-  select(language, `82A`:`97A`) %>%
-  select(-`84A`)
+  select(language, `30A`:`57A`) #%>%
+wo_wf$non <- rowSums(is.na(wo_wf))
+wo_wf %<>%
+  filter(non < 12) %>%
+  select(-non)
 
-wo_df <- wo_wf[rowSums(is.na(wo_wf)) < 6,] %>%
+wo_df <- wo_wf %>%
   as_data_frame() %>%
@@
   slice(1) %>%
-  gather(feature, value, `82A`:`97A`) %>%
+  gather(feature, value, `30A`:`57A`) %>%
   group_by(feature) %>%
@@
 imputed_wo <- as.data.frame(imputed_wo[[1]][[99]])
-imputed_wo$language <- unique(wo_df$language)
+
 wo_sub <- imputed_wo %>% group_by(language) %>%
   slice(1) %>%
-  gather(feature, value, `82A`:`97A`) %>%
+  gather(feature, value, `30A`:`57A`) %>%
   group_by(feature) %>%

### 2019-06-26 surprisal/letter_surprisal.Rmd
providence unigrams, palasis and wagner letter surprisal

@@
 library(childesr)
+library(here)
+library(feather)
 
@@
 ```{r set up corpus}
-prov_utterances <- get_utterances(corpus = "Providence")
+prov_utterances <- get_utterances(corpus = "Palasis")
 
@@
 ```{r plot results}
+letter_unigrams <- letter_unigrams %>% mutate(gram = "Unigram") 
+letter_bigrams <- letter_bigrams %>% mutate(gram = "Bigram") 
+letter_trigrams <- letter_trigrams %>% mutate(gram = "Trigram") 
+letter_unigrams %>% bind_rows(letter_bigrams, letter_trigrams) %>% write_feather(here("Data/palasis_letters.feather"))
+
 letter_unigrams %>%

### 2019-06-26 surprisal/phoneme_surprisal.Rmd
providence unigrams, palasis and wagner letter surprisal

@@
 
+```{r childes unigrams}
+phone_probs <- adult_phonemes %>% 
+  count(phone) %>%
+  ungroup() %>%
+  mutate(p = n/sum(n))
+
+phone_unigrams <- adult_phonemes %>% 
+  left_join(phone_probs) %>%
+  mutate(s = -log(p)) %>%
+  group_by(word_length, phone_order) %>%
+  summarise(mean_s = mean(s))
+```
+
+```{r childes bigrams}
+bigram_probs <- adult_phonemes %>% 
+  group_by(utterance_id) %>%
+  mutate(lag_phone = lag(phone)) %>%
+  group_by(lag_phone, phone) %>%
+  count() %>%
+  filter(!is.na(lag_phone)) %>%
+  ungroup() %>%
+  left_join(phone_probs, by = c("lag_phone" = "phone")) %>%
+  mutate(cond_p = n.x / n.y) %>%
+  select(lag_phone, phone, cond_p, n.x) %>%
+  rename(n = n.x)
+
+phone_bigrams <- adult_phonemes %>%
+  group_by(utterance_id, word_id) %>%
+  mutate(lag_phone = lag(phone)) %>%
+  ungroup() %>%
+  left_join(bigram_probs, by = c("lag_phone" = "lag_phone", "phone" = "phone")) %>%
+  left_join(phone_probs, by = c("phone" = "phone")) %>%
+  mutate(s = ifelse(is.na(lag_phone), -log(p), -log(cond_p))) %>%
+  group_by(word_length, phone_order) %>%
+  summarise(mean_s = mean(s)) 
+```
+
+```{r childes trigrams}
+trigram_probs <- adult_phonemes %>%
+  group_by(utterance_id, word_id) %>%
+  mutate(lag_phone1 = lag(phone)) %>%
+  mutate(lag_phone2 = lag(lag_phone1)) %>%
+  group_by(lag_phone2, lag_phone1, phone) %>%
+  count() %>%
+  filter(!is.na(lag_phone1), !is.na(lag_phone2)) %>%
+  ungroup() %>%
+  left_join(bigram_probs, by = c("lag_phone2" = "lag_phone", "lag_phone1" = "phone")) %>%
+  mutate(tri_cond_p = n.x / n.y) %>%
+  select(-cond_p, -n.y) %>%
+  rename(n = n.x)
+
+phone_trigrams <- adult_phonemes %>%
+  group_by(utterance_id, word_id) %>%
+  mutate(lag_phone1 = lag(phone)) %>%
+  mutate(lag_phone2 = lag(lag_phone1)) %>%
+  left_join(trigram_probs, by = c("lag_phone2" = "lag_phone2", "lag_phone1" = "lag_phone1", "phone" = "phone")) %>% 
+  left_join(bigram_probs, by = c("lag_phone1" = "lag_phone", "phone" = "phone")) %>% 
+  left_join(phone_probs, by = c("phone" = "phone")) %>%
+  mutate(s = ifelse(is.na(lag_phone2), # check if it's not the first word
+                    ifelse(is.na(lag_phone1), # check if not the second word
+                           -log(p), # trigram if third or beyond
+                           -log(cond_p)), # bigram if second word
+                    -log(tri_cond_p))) %>% # unigram if first word
+  group_by(word_length, phone_order) %>%
+  summarise(mean_s = mean(s)) %>% 
+  filter(phone_order <= word_length)
+```
+
 Briefly, what we're doing is splitting each word in the Providence corpus from CHILDES up into its constituent phones, i.e.
@@
 
-```{r show leave-out-surprisal}
+```{r show leave-out-entropy}
 split_lengths <- adult_phonemes %>%

### 2019-06-27 surprisal/phoneme_surprisal.Rmd
phoneme write-up draft

@@
 
-system("espeak -f words.txt --ipa=3 -v en-us -q --phonout='phones.txt'") #get the ipa of each word
+system("espeak -f words.txt --ipa=3 -v fr -q --phonout='phones.txt'") #get the ipa of each word
 system("cat phones.txt | sed -E 's/ /|/g' | tr '|' '\n' | tail -n +2 > splitphones.txt") #put each ipa word onto its own line in a file called splitphones.txt
@@
 
+system("rm words.txt phones.txt splitphones.txt joined.csv")
+adult_utterances %>%
+  filter(word != "", word != "tʃutʃutʃu", word != "ɛː", word != "lunchroom", word != "dryad's", word != "vi") %>%
+  pull(word) %>%
+  unique() %>%
+  write_lines("words.txt")
+
+system("espeak -f words.txt --ipa=3 -v en-us -q --phonout='phones.txt'") #get the ipa of each word
+system("cat phones.txt | sed -E 's/ /|/g' | tr '|' '\n' | tail -n +2 > splitphones.txt") #put each ipa word onto its own line in a file called splitphones.txt
+system("rm phones.txt")
+system("sed '/^$/d' splitphones.txt > phones.txt") #remove empty lines in splitphones
+system("paste -d , words.txt phones.txt > joined.csv") #make csv with words and their ipa paired
+
 phones <- read_csv("joined.csv",
@@
   select(utterance_id, word_id, word, ipa, phone, phone_order, word_length)
+
+adult_phonemes %>% write_feather(here("Data/adult_phonemes.feather"))
 
@@
   filter(phone_order <= word_length)
+
+phone_unigrams <- phone_unigrams %>% mutate(gram = "Unigram") 
+phone_bigrams <- phone_bigrams %>% mutate(gram = "Bigram") 
+phone_trigrams <- phone_trigrams %>% mutate(gram = "Trigram") 
+phone_unigrams %>% bind_rows(phone_bigrams, phone_trigrams) %>% write_feather(here("Data/palasis_phones.feather"))
 ```
-
-Briefly, what we're doing is splitting each word in the Providence corpus from CHILDES up into its constituent phones, i.e.
-the sounds that make up the word.
-We then calculate how common each phone is across the entire corpus.
-From this frequency measure, we compute the surprisal of each phone, which is inversely proportional to the frequency of the phone.
-The less common a phone is, the more difficult it is to process the phone.
-
-We split up the words in the corpus by their length in number of phones--for example, the word "food" has three phones: one for the "f", one for the long "oo", and one for the "d".
-We would take all words like this that have three phones.
-For each phone position  within the words of that length, we calculate the average surprisal/information in that position.
-For example, looking at all of the words with three phones, we take all of the first phones and calculate how much information is in that position on average.
-In other words, how difficult is it to process the first sound in a word, compared to the other sounds in a word?
 
@@
 ```
-
-In the plot above
 
@@
 
-For this next analysis, we took all words with 2, 3, 4, 5 or 6 phones.
-For each word length, we held out the phones in one phone position.
-For example, in all words with 4 phones, we held out the first phone.
-We then computed the entropy, or amount of variation, in the remaining words.
-The higher the entropy when holding out phones in a given position, the more information is contained in that phone on average.
-
 ```{r show leave-out-entropy}
@@
 ```
-
-We see that word-initial phones tend to have more information in them than word-medial or word-final phones, so less is lost by dropping the final sounds in a word than in dropping the initial sounds.

### 2019-06-27 surprisal/phoneme_writeup.Rmd
phoneme write-up draft

---
title: "Phoneme surprisal"
author: Josef Klafka and Dan Yurovsky 
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: hide 
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidytext)
library(tidyverse)
library(childesr)
library(glue)
library(feather)
library(here)
library(entropy)
library(tidyboot)
library(janitor)
library(arrangements)

knitr::opts_chunk$set(echo = TRUE)
```

We're going to discuss the amount of information across letters in English, then information in English phones (sounds), then the hold-out entropy in English and finally the information in letters in French and German. We are interested in French because when many French words are pronounced, the final letters are not pronounced in any way (e.g. "ils sauraient"="they would know", the final "ent" in "sauraient" is not pronounced at all). We are interested in German because German has a much richer case system than English or French, and case markings in German are suffixes (at the ends of words). So in German, the ends of words may be much more important than in French or English.

```{r read data}
prov_letters <- read_feather(here("Data/providence_letters.feather"))
prov_phones <- read_feather(here("Data/providence_phonemes.feather"))
adult_phonemes <- read_feather(here("Data/adult_phonemes.feather"))
french_letters <- read_feather(here("Data/palasis_letters.feather"))
german_letters <- read_feather(here("Data/wagner_letters.feather"))
```

# English Letters

How much information is in each letter in English words? We use the Providence corpus from CHILDES. In the plots below, each plot represents the information in each letter in words of different lengths by number of letters--we show the length above each plot.

We see a distinctive shape for information at the level of individual letters in English. The first letter has relatively high information, the second letter has less information and the third letter has high infomration, and the final letter (if the word has more than 3 letters) has less information.

```{r plot prov letter unigrams}
prov_letters %>%
  filter(gram == "Unigram") %>%
  filter(word_length %in% 2:6) %>%
  ggplot(aes(x = letter_order, y = mean_s)) +
    facet_wrap(~word_length) +
    geom_point() +
    geom_line() +
    xlab("Phone Position") +
    ylab("Average Information") + 
    ggtitle("Information in English letters by letter position - unigrams")
```

Now we ask: if you know the last letter, how much information is in the next letter in English? There is relatively less information in each letter knowing the previous letter, but we see the same general information trajectory across letters.

```{r plot prov letter bigrams}
prov_letters %>%
  filter(gram == "Bigram") %>%
  filter(word_length %in% 2:6) %>%
  ggplot(aes(x = letter_order, y = mean_s)) +
    facet_wrap(~word_length) +
    geom_point() +
    geom_line() +
    xlab("Phone Position") +
    ylab("Average Information") + 
    ggtitle("Information in English letters by letter position - bigrams")
```

What about if you know the last two letters and are trying to predict the next letter? We see decreasing information within words as you move left-to-right.

```{r plot prov letter trigrams}
prov_letters %>%
  filter(gram == "Trigram") %>%
  filter(word_length %in% 2:6) %>%
  ggplot(aes(x = letter_order, y = mean_s)) +
    facet_wrap(~word_length) +
    geom_point() +
    geom_line() +
    xlab("Phone Position") +
    ylab("Average Information") + 
    ggtitle("Information in English letters by letter position - trigrams")
```

# English Phonemes

## Surprisal

Briefly, what we're doing is splitting each word in the Providence corpus from CHILDES up into its constituent phones, i.e. the sounds that make up the word. We then calculate how common each phone is across the entire corpus. From this frequency measure, we compute the surprisal of each phone, which is inversely proportional to the frequency of the phone. The less common a phone is, the more difficult it is to process the phone.

We split up the words in the corpus by their length in number of phones--for example, the word "food" has three phones: one for the "f", one for the long "oo", and one for the "d". We would take all words like this that have three phones. For each phone position  within the words of that length, we calculate the average surprisal/information in that position. For example, looking at all of the words with three phones, we take all of the first phones and calculate how much information is in that position on average. In other words, how difficult is it to process the first sound in a word, compared to the other sounds in a word?

Unlike with letters, there isn't as much of a consistent shape to the information distribution for phones considered without context. For the most part, the first phone has relatively little information while the second phone has a lot of information.

```{r plot prov phone unigrams}
prov_phones %>%
  filter(gram == "Unigram") %>%
  filter(word_length %in% 2:6) %>%
  ggplot(aes(x = phone_order, y = mean_s)) +
    facet_wrap(~word_length) +
    geom_point() +
    geom_line() +
    xlab("Phone Position") +
    ylab("Average Information") + 
    ggtitle("Information in English phones by phone position - unigrams")
```

With bigrams and trigrams for English phones, the next phone is very predictable if you know the phone or two that came before it. Especially if you know the preceding two phones, then each phone past the first two is predictable and carries relatively little information in English.

```{r plot prov phone bigrams}
prov_phones %>%
  filter(gram == "Bigram") %>%
  filter(word_length %in% 2:6) %>%
  ggplot(aes(x = phone_order, y = mean_s)) +
    facet_wrap(~word_length) +
    geom_point() +
    geom_line() +
    xlab("Phone Position") +
    ylab("Average Information") + 
    ggtitle("Information in English phones by phone position - bigrams")
```

```{r plot prov phone trigrams}
prov_phones %>%
  filter(gram == "Trigram") %>%
  filter(word_length %in% 2:6) %>%
  ggplot(aes(x = phone_order, y = mean_s)) +
    facet_wrap(~word_length) +
    geom_point() +
    geom_line() +
    xlab("Phone Position") +
    ylab("Average Information") + 
    ggtitle("Information in English phones by phone position - trigrams")
```

## Hold-out entropy

For this next analysis, we took all words with 2, 3, 4, 5 or 6 phones. For each word length, we held out the phones in one phone position. For example, in all words with 4 phones, we held out the first phone. We then computed the entropy, or amount of variation, in the remaining words. The higher the entropy when holding out phones in a given position, the more information is contained in that phone on average.

```{r leave-out-entropy, include = F}
hold_out_entropy <- function(df) {
  spread_df <- df %>%
    spread(phone_order, phone) %>%
    clean_names()

  phone_length <- max(df$phone_order)

  lens <- paste0("x", 1:phone_length)

  boot_entropy <- function(group) {
    spread_df %>%
      group_by_at(group) %>%
      summarise(n = n()) %>%
      summarise(entropy = entropy(n)) %>%
      ungroup() %>%
      tidyboot_mean(entropy)
  }

  groupings <- combinations(lens, phone_length -1, 
                            layout = "list")

  groupings <- map(seq(length(lens),1,-1),
                   function(x) c(groupings[[x]], lens[x]))

  map_df(groupings, boot_entropy) %>%
    mutate(phone_order = seq(phone_length, 1, -1))

}
```

```{r show leave-out-entropy}
split_lengths <- adult_phonemes %>%
  group_by(word_length) %>%
  nest() %>%
  filter(word_length > 1) %>%
  filter(word_length %in% 1:6 ) %>%
  mutate(entropy = map(data, hold_out_entropy))

split_lengths %>%
  select(-data) %>%
  unnest() %>%
  ggplot(aes(x = phone_order, y = empirical_stat)) + 
  facet_wrap(~ word_length, scales = "free") +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) + 
  theme_classic() + 
  geom_line() + 
  xlab("Phone position") + 
  ylab("Entropy")
```

We see that word-initial phones tend to have more information in them than word-medial or word-final phones, so less is lost by dropping the final sounds in a word than in dropping the initial sounds.

# French letters

In French, the final letters in a lot of words are dropped when those words are spoken aloud. We expect that the final letters in French words will on average contain little information, and that is what we see. In particular, if you know the two letters that came before, then each letter beyond the first in French has little information. Even if you don't know which letters came before, then the final letters of French words have relatively little information in them.

```{r plot french letter unigrams}
french_letters %>%
  filter(gram == "Unigram") %>%
  filter(word_length %in% 2:6) %>%
  ggplot(aes(x = letter_order, y = mean_s)) +
    facet_wrap(~word_length) +
    geom_point() +
    geom_line() +
    xlab("Phone Position") +
    ylab("Average Information") + 
    ggtitle("Information in French letters by letter position - unigrams")
```

```{r plot french letter bigrams}
french_letters %>%
  filter(gram == "Bigram") %>%
  filter(word_length %in% 2:6) %>%
  ggplot(aes(x = letter_order, y = mean_s)) +
    facet_wrap(~word_length) +
    geom_point() +
    geom_line() +
    xlab("Phone Position") +
    ylab("Average Information") + 
    ggtitle("Information in French letters by letter position - bigrams")
```

```{r plot french letter trigrams}
french_letters %>%
  filter(gram == "Trigram") %>%
  filter(word_length %in% 2:6) %>%
  ggplot(aes(x = letter_order, y = mean_s)) +
    facet_wrap(~word_length) +
    geom_point() +
    geom_line() +
    xlab("Phone Position") +
    ylab("Average Information") + 
    ggtitle("Information in French letters by letter position - trigrams")
```

# German letters

We expect that German, as a language with a case system which gets shown through affixation on nouns, will have relatively higher information in the final letters of words compared to English or French. We don't observe this, even in unigrams. The final letters of German words hold little information compared to the first letter of German words, which is even more noticeable when you have two letters of predictive context for each word.

```{r plot german letter unigrams}
german_letters %>%
  filter(gram == "Unigram") %>%
  filter(word_length %in% 2:6) %>%
  ggplot(aes(x = letter_order, y = mean_s)) +
    facet_wrap(~word_length) +
    geom_point() +
    geom_line() +
    xlab("Phone Position") +
    ylab("Average Information") + 
    ggtitle("Information in German letters by letter position - unigrams")
```

```{r plot german letter bigrams}
german_letters %>%
  filter(gram == "Bigram") %>%
  filter(word_length %in% 2:6) %>%
  ggplot(aes(x = letter_order, y = mean_s)) +
    facet_wrap(~word_length) +
    geom_point() +
    geom_line() +
    xlab("Phone Position") +
    ylab("Average Information") + 
    ggtitle("Information in German letters by letter position - bigrams")
```

```{r plot german letter trigrams}
german_letters %>%
  filter(gram == "Trigram") %>%
  filter(word_length %in% 2:6) %>%
  ggplot(aes(x = letter_order, y = mean_s)) +
    facet_wrap(~word_length) +
    geom_point() +
    geom_line() +
    xlab("Phone Position") +
    ylab("Average Information") + 
    ggtitle("Information in German letters by letter position - trigrams")
```

### 2019-07-01 surprisal/phoneme_writeup.Rmd
tidying phoneme writeup

@@
 
-```{r setup, include=FALSE}
+```{r load-libraries, library, message=F, results='hide', warning=FALSE}
 library(tidytext)
@@
 library(arrangements)
+library(stringr)
 
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, cache = T, warning = F, 
+                      fig.width = 8, fig.height = 4)
+
+theme_set(theme_classic(base_size = 16))
 ```
 
-We're going to discuss the amount of information across letters in English, then information in English phones (sounds), then the hold-out entropy in English and finally the information in letters in French and German.
-We are interested in French because when many French words are pronounced, the final letters are not pronounced in any way (e.g.
-"ils sauraient"="they would know", the final "ent" in "sauraient" is not pronounced at all).
-We are interested in German because German has a much richer case system than English or French, and case markings in German are suffixes (at the ends of words).
-So in German, the ends of words may be much more important than in French or English.
+Where is the information in English words?
+A salient feature of African American Enlgish Vernacular is the reduction of final consonants and consonant clusters.
+Why must this happen?
+One possibility is that the information important for identifying a word is not uniformly distributed across it's characters or sounds but instead concentrated at the beginnings or middles of words.
+In these cases, reducing final sounds might reduce the articulatory complexity of words without much cost for the listener in processing these words and inferring their intended meaning.
 
-```{r read data}
-prov_letters <- read_feather(here("Data/providence_letters.feather"))
-prov_phones <- read_feather(here("Data/providence_phonemes.feather"))
+In a series of initial analyses we tackle three ways:
+1.
+Computing the surprisal of characters in English words either from baserates of character production (unigram) or conditioning on the prior charaters in each word (bigram and trigram).
+This analyses roughly captures the difficulty that a listener might have in predicting each character.
+2. Repeating this analysis at the level of phonemes rather than characters, arguably a better proxy for the speaker's and listener's representations
+3.
+Computing hold-out entropy for individual phonemes in English words.
+This is roughly equivalent to asking "if you knew every other sound in a word, how much trouble would you have predicting what sounds appears in position X"
+4. Comparing English to French and German which have different expected information profiles (more below).
+
+SOME KIND OF PREVIEW OF RESULTS HERE?
+
+```{r read-data}
+prov_letters <- read_feather(here("Data/providence_letters.feather")) %>%
+  mutate(rep = "letter", language = "English") %>%
+  rename(order = letter_order)
+
+prov_phones <- read_feather(here("Data/providence_phonemes.feather")) %>%
+  mutate(rep = "phone", language = "English") %>%
+  rename(order = phone_order)
+
 adult_phonemes <- read_feather(here("Data/adult_phonemes.feather"))
-french_letters <- read_feather(here("Data/palasis_letters.feather"))
-german_letters <- read_feather(here("Data/wagner_letters.feather"))
+
+french_letters <- read_feather(here("Data/palasis_letters.feather")) %>%
+  mutate(rep = "letter", language = "French") %>%
+  rename(order = letter_order)
+
+german_letters <- read_feather(here("Data/wagner_letters.feather")) %>%
+  mutate(rep = "letter", language = "German") %>%
+  rename(order = letter_order)
+
+all_data <- bind_rows(prov_letters, prov_phones, french_letters, german_letters)
 ```
@@
 
-How much information is in each letter in English words?
-We use the Providence corpus from CHILDES.
+We use for our English corpus the Providence corpus from CHILDES.
 In the plots below, each plot represents the information in each letter in words of different lengths by number of letters--we show the length above each plot.
@@
 
-```{r plot prov letter unigrams}
-prov_letters %>%
-  filter(gram == "Unigram") %>%
+```{r plot-helper}
+plot_curves <- function(select_language = "English", select_rep = "letter", 
+                        select_gram = "Unigram") {
+  all_data %>%
+    filter(language == select_language, rep == select_rep,
+           gram == select_gram) %>%
     filter(word_length %in% 2:6) %>%
-  ggplot(aes(x = letter_order, y = mean_s)) +
-    facet_wrap(~word_length) +
+    ggplot(aes(x = order, y = mean_s)) +
+    facet_grid(. ~ word_length, 
+               labeller = as_labeller(function(x) paste0("Length ", x))) +
     geom_point() +
     geom_line() +
-    xlab("Phone Position") +
-    ylab("Average Information") + 
-    ggtitle("Information in English letters by letter position - unigrams")
+    xlab(paste0(str_to_sentence(select_rep), " Position")) +
+    ylab("Surprisal") + 
+    ggtitle(paste0("Information in ", select_language, " ", 
+                   select_rep, "s by ",select_rep, " position - ", select_gram))
+}
 ```
 
-Now we ask: if you know the last letter, how much information is in the next letter in English?
+```{r prov-letter-unigrams}
+plot_curves()
+```
+
+Now we ask: if you know the previous letter, how much information is in the next letter in English?
 There is relatively less information in each letter knowing the previous letter, but we see the same general information trajectory across letters.
 
-```{r plot prov letter bigrams}
-prov_letters %>%
-  filter(gram == "Bigram") %>%
-  filter(word_length %in% 2:6) %>%
-  ggplot(aes(x = letter_order, y = mean_s)) +
-    facet_wrap(~word_length) +
-    geom_point() +
-    geom_line() +
-    xlab("Phone Position") +
-    ylab("Average Information") + 
-    ggtitle("Information in English letters by letter position - bigrams")
+```{r prov-letter-bigrams}
+plot_curves(select_gram = "Bigram")
 ```
@@
 
-```{r plot prov letter trigrams}
-prov_letters %>%
-  filter(gram == "Trigram") %>%
-  filter(word_length %in% 2:6) %>%
-  ggplot(aes(x = letter_order, y = mean_s)) +
-    facet_wrap(~word_length) +
-    geom_point() +
-    geom_line() +
-    xlab("Phone Position") +
-    ylab("Average Information") + 
-    ggtitle("Information in English letters by letter position - trigrams")
+```{r prov-letter-trigrams}
+plot_curves(select_gram = "Trigram")
 ```
@@
 
-```{r plot prov phone unigrams}
-prov_phones %>%
-  filter(gram == "Unigram") %>%
-  filter(word_length %in% 2:6) %>%
-  ggplot(aes(x = phone_order, y = mean_s)) +
-    facet_wrap(~word_length) +
-    geom_point() +
-    geom_line() +
-    xlab("Phone Position") +
-    ylab("Average Information") + 
-    ggtitle("Information in English phones by phone position - unigrams")
+```{r prov-phone-unigrams}
+plot_curves(select_rep = "phone")
 ```
@@
 
-```{r plot prov phone bigrams}
-prov_phones %>%
-  filter(gram == "Bigram") %>%
-  filter(word_length %in% 2:6) %>%
-  ggplot(aes(x = phone_order, y = mean_s)) +
-    facet_wrap(~word_length) +
-    geom_point() +
-    geom_line() +
-    xlab("Phone Position") +
-    ylab("Average Information") + 
-    ggtitle("Information in English phones by phone position - bigrams")
+```{r prov-phone-bigrams}
+plot_curves(select_rep = "phone", select_gram = "Bigram")
 ```
 
-```{r plot prov phone trigrams}
-prov_phones %>%
-  filter(gram == "Trigram") %>%
-  filter(word_length %in% 2:6) %>%
-  ggplot(aes(x = phone_order, y = mean_s)) +
-    facet_wrap(~word_length) +
-    geom_point() +
-    geom_line() +
-    xlab("Phone Position") +
-    ylab("Average Information") + 
-    ggtitle("Information in English phones by phone position - trigrams")
+```{r prov-phone-trigrams}
+plot_curves(select_rep = "phone", select_gram = "Trigram")
 ```
@@
 
-```{r show leave-out-entropy}
+```{r show-leave-out-entropy}
 split_lengths <- adult_phonemes %>%
@@
   ggplot(aes(x = phone_order, y = empirical_stat)) + 
-  facet_wrap(~ word_length, scales = "free") +
+  facet_wrap(~ word_length, 
+               labeller = as_labeller(function(x) paste0("Length ", x)),
+               scales = "free") +
   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) + 
@@
 
+We are interested in French because when many French words are pronounced, the final letters are not pronounced in any way (e.g.
+"ils sauraient"="they would know", the final "ent" in "sauraient" is not pronounced at all).
+We are interested in German because German has a much richer case system than English or French, and case markings in German are suffixes (at the ends of words).
+So in German, the ends of words may be much more important than in French or English.
+
 # French letters
@@
 ```{r plot french letter unigrams}
-french_letters %>%
-  filter(gram == "Unigram") %>%
-  filter(word_length %in% 2:6) %>%
-  ggplot(aes(x = letter_order, y = mean_s)) +
-    facet_wrap(~word_length) +
-    geom_point() +
-    geom_line() +
-    xlab("Phone Position") +
-    ylab("Average Information") + 
-    ggtitle("Information in French letters by letter position - unigrams")
+plot_curves(select_language = "French")
 ```
@@
 ```{r plot french letter bigrams}
-french_letters %>%
-  filter(gram == "Bigram") %>%
-  filter(word_length %in% 2:6) %>%
-  ggplot(aes(x = letter_order, y = mean_s)) +
-    facet_wrap(~word_length) +
-    geom_point() +
-    geom_line() +
-    xlab("Phone Position") +
-    ylab("Average Information") + 
-    ggtitle("Information in French letters by letter position - bigrams")
+plot_curves(select_language = "French", select_gram = "Bigram")
 ```
 
-```{r plot french letter trigrams}
-french_letters %>%
-  filter(gram == "Trigram") %>%
-  filter(word_length %in% 2:6) %>%
-  ggplot(aes(x = letter_order, y = mean_s)) +
-    facet_wrap(~word_length) +
-    geom_point() +
-    geom_line() +
-    xlab("Phone Position") +
-    ylab("Average Information") + 
-    ggtitle("Information in French letters by letter position - trigrams")
+```{r french-letter-trigrams}
+plot_curves(select_language = "French", select_gram = "Trigram")
 ```
@@
 ```{r plot german letter unigrams}
-german_letters %>%
-  filter(gram == "Unigram") %>%
-  filter(word_length %in% 2:6) %>%
-  ggplot(aes(x = letter_order, y = mean_s)) +
-    facet_wrap(~word_length) +
-    geom_point() +
-    geom_line() +
-    xlab("Phone Position") +
-    ylab("Average Information") + 
-    ggtitle("Information in German letters by letter position - unigrams")
+plot_curves(select_language = "German")
 ```
@@
 ```{r plot german letter bigrams}
-german_letters %>%
-  filter(gram == "Bigram") %>%
-  filter(word_length %in% 2:6) %>%
-  ggplot(aes(x = letter_order, y = mean_s)) +
-    facet_wrap(~word_length) +
-    geom_point() +
-    geom_line() +
-    xlab("Phone Position") +
-    ylab("Average Information") + 
-    ggtitle("Information in German letters by letter position - bigrams")
+plot_curves(select_language = "German", select_gram = "Bigram")
 ```
@@
 ```{r plot german letter trigrams}
-german_letters %>%
-  filter(gram == "Trigram") %>%
-  filter(word_length %in% 2:6) %>%
-  ggplot(aes(x = letter_order, y = mean_s)) +
-    facet_wrap(~word_length) +
-    geom_point() +
-    geom_line() +
-    xlab("Phone Position") +
-    ylab("Average Information") + 
-    ggtitle("Information in German letters by letter position - trigrams")
+plot_curves(select_language = "German", select_gram = "Trigram")
 ```

### 2019-07-01 surprisal/letter_surprisal.Rmd
phoneme writeup counts and edits

@@
 
-adult_utterances <- prov_utterances %>%
+french_utterances <- prov_utterances %>%
   filter(speaker_role != "Target_Child") %>% 
@@
   ungroup()
+
+phone_counts <- adult_phonemes %>% 
+  group_by(utterance_id, word_id) %>% 
+  summarise(length = max(word_length)) %>% 
+  ungroup() %>% 
+  count(length)
 ```

### 2019-07-01 surprisal/phoneme_writeup.Rmd
phoneme writeup counts and edits

@@
 ---
-title: "Phoneme surprisal"
+title: "Information within words in English, French and German speech"
 author: Josef Klafka and Dan Yurovsky 
@@
     toc: false
-    number_sections: false
+    number_sections: true
     theme: lumen
@@
 
-Where is the information in English words?
-A salient feature of African American Enlgish Vernacular is the reduction of final consonants and consonant clusters.
-Why must this happen?
-One possibility is that the information important for identifying a word is not uniformly distributed across it's characters or sounds but instead concentrated at the beginnings or middles of words.
+```{r read-data}
+all_data <- read_feather(here("Data/all_phonemes.feather"))
+all_counts <- read_feather(here("Data/all_counts.feather"))
+```
+
+How is information spread among the sounds in English words?
+A salient feature of African American English Vernacular is the reduction of final consonants and consonant clusters.
+Why hasn't this created communication problems, namely identifying and disambiguating words?
+One possibility is that the information important for identifying a word is not uniformly distributed across all of its characters or sounds, but instead concentrated at the beginnings or middles of words.
 In these cases, reducing final sounds might reduce the articulatory complexity of words without much cost for the listener in processing these words and inferring their intended meaning.
+By the time the listener hears only the first sounds of a word, they may already have identified which word they are hearing.
 
-In a series of initial analyses we tackle three ways:
+Through a series of exploratory analyses we tackle this problem in three ways:
 1.
-Computing the surprisal of characters in English words either from baserates of character production (unigram) or conditioning on the prior charaters in each word (bigram and trigram).
+Computing the surprisal of characters in English words from baserates of character production (unigram) and then conditioning on the prior charaters in each word (bigram and trigram).
 This analyses roughly captures the difficulty that a listener might have in predicting each character.
-2. Repeating this analysis at the level of phonemes rather than characters, arguably a better proxy for the speaker's and listener's representations
+2. Repeating this analysis at the level of phonemes rather than characters, arguably a better proxy for the speaker's and listener's representations of spoken English. 
 3.
 Computing hold-out entropy for individual phonemes in English words.
-This is roughly equivalent to asking "if you knew every other sound in a word, how much trouble would you have predicting what sounds appears in position X"
-4. Comparing English to French and German which have different expected information profiles (more below).
+This is roughly equivalent to asking "if you knew every other sound in a word, how much trouble would you have predicting what sounds appears in position $X$".
+4.
+Comparing our English by-letter analysis to results for French and German which have different expected information profiles.
+We expect that French words will have less information at the end of words compared to English, while German words will have more compared to English.
 
-SOME KIND OF PREVIEW OF RESULTS HERE?
-
-```{r read-data}
-prov_letters <- read_feather(here("Data/providence_letters.feather")) %>%
-  mutate(rep = "letter", language = "English") %>%
-  rename(order = letter_order)
-
-prov_phones <- read_feather(here("Data/providence_phonemes.feather")) %>%
-  mutate(rep = "phone", language = "English") %>%
-  rename(order = phone_order)
-
-adult_phonemes <- read_feather(here("Data/adult_phonemes.feather"))
-
-french_letters <- read_feather(here("Data/palasis_letters.feather")) %>%
-  mutate(rep = "letter", language = "French") %>%
-  rename(order = letter_order)
-
-german_letters <- read_feather(here("Data/wagner_letters.feather")) %>%
-  mutate(rep = "letter", language = "German") %>%
-  rename(order = letter_order)
-
-all_data <- bind_rows(prov_letters, prov_phones, french_letters, german_letters)
-```
+Across all of our analyses, we find that the letters and sounds at the end of words in English have less information than the first letters and sounds in a word.
+Letters at the end of words in French and German also have little information compared to letters at the beginning of words in those languages, like in English.
 
@@
 
-We use for our English corpus the Providence corpus from CHILDES.
-In the plots below, each plot represents the information in each letter in words of different lengths by number of letters--we show the length above each plot.
+We use the Providence corpus from CHILDES as data for English.
+In the plots below, each plot represents the information in each letter, divided by words of different lengths.
+Word lengths are given by number of letters--we show the length above each plot.
 
 We see a distinctive shape for information at the level of individual letters in English.
-The first letter has relatively high information, the second letter has less information and the third letter has high infomration, and the final letter (if the word has more than 3 letters) has less information.
+The first letter has relatively high information, the second letter has less information and the third letter has high infomration, and the final letter has less information (if the word has more than $3$ letters).
 
@@
     facet_grid(. ~ word_length, 
-               labeller = as_labeller(function(x) paste0("Length ", x))) +
+               labeller = as_labeller(function(x) paste0("Length ", x, " \n n = ", 
+                                                         all_counts %>% 
+                                                           filter(language == select_language,
+                                                                  rep == select_rep, 
+                                                                  length == x) %>%
+                                                           pull(length)))) +
     geom_point() +
@@
     xlab(paste0(str_to_sentence(select_rep), " Position")) +
-    ylab("Surprisal") + 
+    ylab("Average information") + 
     ggtitle(paste0("Information in ", select_language, " ", 
@@
 
-## Surprisal
-
-Briefly, what we're doing is splitting each word in the Providence corpus from CHILDES up into its constituent phones, i.e.
-the sounds that make up the word.
-We then calculate how common each phone is across the entire corpus.
-From this frequency measure, we compute the surprisal of each phone, which is inversely proportional to the frequency of the phone.
-The less common a phone is, the more difficult it is to process the phone.
-
+Briefly, what we're doing is splitting each word in the Providence corpus from CHILDES up into its constituent phones--the sounds that make up the word.
 We split up the words in the corpus by their length in number of phones--for example, the word "food" has three phones: one for the "f", one for the long "oo", and one for the "d".
 We would take all words like this that have three phones.
-For each phone position  within the words of that length, we calculate the average surprisal/information in that position.
-For example, looking at all of the words with three phones, we take all of the first phones and calculate how much information is in that position on average.
-In other words, how difficult is it to process the first sound in a word, compared to the other sounds in a word?
+For each phone position  within the words of that length, we calculate the average information in that position.
 
-Unlike with letters, there isn't as much of a consistent shape to the information distribution for phones considered without context.
-For the most part, the first phone has relatively little information while the second phone has a lot of information.
+Unlike with letters, there isn't as much of a consistent shape to the information distribution for phones in the unigram condition.
+If anything, the first phone has relatively little information while the second phone has a lot of information.
 
@@
 
-## Hold-out entropy
+# Hold-out entropy
 
-For this next analysis, we took all words with 2, 3, 4, 5 or 6 phones.
+For this next analysis, we took all words with $2$, $3$, $4$, $5$ or $6$ phones.
 For each word length, we held out the phones in one phone position.
@@
 
-We are interested in French because when many French words are pronounced, the final letters are not pronounced in any way (e.g.
-"ils sauraient"="they would know", the final "ent" in "sauraient" is not pronounced at all).
-We are interested in German because German has a much richer case system than English or French, and case markings in German are suffixes (at the ends of words).
-So in German, the ends of words may be much more important than in French or English.
+# French and German
 
-# French letters
+For our final analysis, we're going to look at how the information within words in French and German compare to English.
+We use the Palasis French corpus and Wagner German corpus from CHILDES for our data.
+We chose French because for many French words the final letters are not pronounced regardless of how the letters in the word appear.
+For example, "ils sauraient" meaning "they would know" is pronounced without the final "ent" in "sauraient".
+We are interested in German because German has a much richer case system than English or French, where case markings in German are suffixes at the ends of words.
+We think that in German, the ends of words may be more important than in French or English.
+
+## French letters
 
@@
 
-# German letters
+## German letters
 

### 2019-07-01 surprisal/phoneme_writeup.Rmd
combined phoneme data and counts

@@
 all_data <- read_feather(here("Data/all_phonemes.feather"))
-all_counts <- read_feather(here("Data/all_counts.feather"))
 ```
@@
                labeller = as_labeller(function(x) paste0("Length ", x, " \n n = ", 
-                                                         all_counts %>% 
+                                                         all_data %>% 
                                                            filter(language == select_language,
@@
                                                                   length == x) %>%
-                                                           pull(length)))) +
+                                                           pull(length) %>%
+                                                           unique()))) +
     geom_point() +

### 2019-07-02 surprisal/wiki_surprisal.Rmd
MIMCA and models on nominal syntax, verbal categories and clauses in wiki_surprisal

@@
 
+The section below reads in the Wikipedia unigram surprisal data and creates a dendrogram of the slope embeddings.
+The dendrogram qualitatively resembles a tree of language families with some degree of accuracy.
+In the rest of this part of the project, we investigate what determines the placement of a language in this dendrogram.
+
 ```{r make dendrogram}
-wiki <- read.csv(here("Data/wikipedia_0context.csv")) %>%
+wiki <- read.csv(here("Data/wikipedia_unigrams.csv")) %>%
   select(-X) %>% 
@@
 
-```{r word order}
-wkpd <- read.csv(here("Data/wikipedia_0context.csv"))
+Always run this section. It takes time, but this gets all of the WALS features for every language in WALS along with the unigram slope data for all of the Wikipedia languages we've run.
+
+```{r get features}
+wkpd <- read.csv(here("Data/wikipedia_unigrams.csv"))
 LANGUAGES <- wkpd$Language
@@
 wf <- wals.feature(FEATURES) 
+```
 
-# word order
-# wo_wf <- wf %>% 
-#   filter(language %in% LANGUAGES) %>%
-#   select(language, `82A`:`97A`) %>%
-#   select(-`84A`)
+Run only _one_ of the following five sections, depending on which typological features you are interested in. Look here (https://wals.info/feature) for more information about the features.
 
-# nominal categories
+```{r nominal categories}
 wo_wf <- wf %>%
@@
   select(-non)
+```
+```{r nominal syntax}
+wo_wf <- wf %>%
+  filter(language %in% LANGUAGES) %>%
+  select(language, `58A`:`64A`) #%>%
+wo_wf$non <- rowSums(is.na(wo_wf))
+wo_wf %<>%
+  filter(non < 5) %>%
+  select(-non)
+```
+```{r verbal categories}
+wo_wf <- wf %>%
+  filter(language %in% LANGUAGES) %>%
+  select(language, `65A`:`80A`) #%>%
+wo_wf$non <- rowSums(is.na(wo_wf))
+wo_wf %<>%
+  filter(non < 8) %>%
+  select(-non)
+```
+```{r word order}
+wo_wf <- wf %>%
+  filter(language %in% LANGUAGES) %>%
+  select(language, `81A`:`97A`) %>%
+  select(-`84A`)
+```
+```{r clauses}
+wo_wf <- wf %>%
+  filter(language %in% LANGUAGES) %>%
+  select(language, `98A`:`121A`) #%>%
+wo_wf$non <- rowSums(is.na(wo_wf))
+wo_wf %<>%
+  filter(non < 12) %>%
+  select(-non)
+```
 
-wo_df <- wo_wf %>%
+Run this section for imputation and model fitting.
+The models will be written into a tiny dataframe: `model_df`.
+Each row of this dataframe has the id of the feature along with a short, informative name about the feature.
+Again, see (https://wals.info/feature) for in-depth information about each feature.
+The estimates, standard errors, statistics and p-values of the models are also contained in `model_df`.
+
+You can adjust the number of components and/or the number of iterations of the imputation algorithm in the `MIMCA` command at the bottom of the first code chunk in this section.
+What we leave in the code we know to converge for each feature category.
+
+```{r imputation and models}
+imputed_wo <- wo_wf %>%
   as_data_frame() %>%
@@
   slice(1) %>%
-  gather(feature, value, `30A`:`57A`) %>%
+  gather(feature, value, -language) %>%
   group_by(feature) %>%
-  mutate(value = factor(value))
-
-imputed_wo <- wo_df %>%
+  mutate(value = factor(value)) %>%
   select(language, feature, value) %>%
@@
   spread(feature, value) %>% 
+  sapply(as.factor) %>%
+  as.data.frame() %>%
   column_to_rownames("language") %>%
-  sapply(as.factor) %>%
   MIMCA(ncp = 1)
@@
 
-wo_sub <- imputed_wo %>% group_by(language) %>%
+wo_sub <- imputed_wo %>% 
+  rownames_to_column("language") %>%
+  group_by(language) %>%
   slice(1) %>%
-  gather(feature, value, `30A`:`57A`) %>%
+  gather(feature, value, -language) %>%
   group_by(feature) %>%

### 2019-07-04 surprisal/phoneme_writeup.Rmd
all data has counts
histograms

@@
 ```{r read-data}
-all_data <- read_feather(here("Data/all_phonemes.feather"))
+all_data <- read_feather(here("Data/all_data.feather"))
 ```
@@
 Through a series of exploratory analyses we tackle this problem in three ways:
+
 1.
@@
 
+```{r plot-hist-helper}
+plot_hists <- function(select_language = "English", select_rep = "letter") {
+
+    all_data %>%
+    filter(language == select_language, rep == select_rep) %>%
+    distinct(word_length, n, rep) %>%
+    ggplot(aes(x = word_length, y = n)) +
+    geom_col(fill = "white", color = "black") +
+    xlab(paste0(str_to_sentence(select_rep), " Length")) +
+    ylab("Frequency")
+}
+```
+
 We use the Providence corpus from CHILDES as data for English.
 In the plots below, each plot represents the information in each letter, divided by words of different lengths.
-Word lengths are given by number of letters--we show the length above each plot.
+Word lengths are given by number of letters.
+
+```{r plot-english-letter-hist}
+plot_hists()
+```
 
@@
 
-```{r plot-helper}
+```{r plot-curve-helper}
 plot_curves <- function(select_language = "English", select_rep = "letter", 
@@
     facet_grid(. ~ word_length, 
-               labeller = as_labeller(function(x) paste0("Length ", x, " \n n = ", 
-                                                         all_data %>% 
-                                                           filter(language == select_language,
-                                                                  rep == select_rep, 
-                                                                  length == x) %>%
-                                                           pull(length) %>%
-                                                           unique()))) +
+               labeller = as_labeller(function(x) paste0("Length ", x))) +
     geom_point() +
@@
 For each phone position  within the words of that length, we calculate the average information in that position.
+
+```{r plot-english-phone-hist}
+plot_hists(select_rep = "phone")
+```
 
@@
 
+```{r plot-french-hist}
+plot_hists(select_language = "French")
+```
+
 ```{r plot french letter unigrams}
@@
 
+```{r plot-german-hist}
+plot_hists(select_language = "German")
+```
+
 ```{r plot german letter unigrams}

### 2019-07-05 surprisal/phoneme_writeup.Rmd
description of histograms and brief general discussion

@@
 Why hasn't this created communication problems, namely identifying and disambiguating words?
-One possibility is that the information important for identifying a word is not uniformly distributed across all of its characters or sounds, but instead concentrated at the beginnings or middles of words.
+One possibility is that the information important for identifying a word is not uniformly distributed across all of its characters or sounds, but instead concentrated at the beginning or middle of the word.
 In these cases, reducing final sounds might reduce the articulatory complexity of words without much cost for the listener in processing these words and inferring their intended meaning.
@@
     geom_col(fill = "white", color = "black") +
-    xlab(paste0(str_to_sentence(select_rep), " Length")) +
+    xlab("Word length") + 
     ylab("Frequency")
 }
+    # xlab(paste0(str_to_sentence(select_rep), " Length")) +
+
 ```
@@
 Word lengths are given by number of letters.
+See below a histogram of the distribution of word lengths in the Providence corpus.
+We observe that most words ($84%$) have $5$ letters or fewer, while $99%$ of words in the corpus have fewer than $10$ letters.
+There are a tiny number of outliers, mainly onomatopoeia, which have $10$ or more letters.
+
+```{r get cumsums for letter and phone frequencies, include = F, eval = F}
+get_freq_sums <- function(select_language = "English", select_rep = "letter") {
+
+  all_data %>% 
+    filter(language == select_language, rep == select_rep) %>% 
+    distinct(word_length, n, rep) %>% 
+    mutate(share = n/sum(n)) %>% 
+    mutate(cumsum = cumsum(share)) %>% 
+    mutate(language = select_language)
+}
+```
 
@@
 We see a distinctive shape for information at the level of individual letters in English.
-The first letter has relatively high information, the second letter has less information and the third letter has high infomration, and the final letter has less information (if the word has more than $3$ letters).
+The first letter has relatively high information, the second letter has less information and the third letter has high information, and the final letter has less information (if the word has more than $3$ letters).
 
@@
 For each phone position  within the words of that length, we calculate the average information in that position.
+
+As seen in the histogram below, almost all words in the Providence corpus ($92%$) have four or fewer phones. $99.5%$ have seven or fewer phones.
 
@@
 For our final analysis, we're going to look at how the information within words in French and German compare to English.
-We use the Palasis French corpus and Wagner German corpus from CHILDES for our data.
+We use the French Palasis corpus and German Wagner corpus from CHILDES for our data.
 We chose French because for many French words the final letters are not pronounced regardless of how the letters in the word appear.
@@
 
+First, we examine the word length frequencies by number of letters in the French Palasis corpus.
+In the Palasis corpus, $99%$ of words have $10$ letters or fewer.
+This French corpus has a higher share of words with six to ten letters than the English Providence corpus: $24%$ versus $16%$.
+However, almost all words in both corpora have $10$ or fewer letters.
+While there are relatively more medium length words in the French corpus compared to the English corpus, there are very few long words in either corpus.
+
+```{r plot-french-hist}
+plot_hists(select_language = "French")
+```
+
 In French, the final letters in a lot of words are dropped when those words are spoken aloud.
@@
 Even if you don't know which letters came before, then the final letters of French words have relatively little information in them.
-
-```{r plot-french-hist}
-plot_hists(select_language = "French")
-```
 
@@
 
-We expect that German, as a language with a case system which gets shown through affixation on nouns, will have relatively higher information in the final letters of words compared to English or French.
-We don't observe this, even in unigrams.
-The final letters of German words hold little information compared to the first letter of German words, which is even more noticeable when you have two letters of predictive context for each word.
+In the German Wagner corpus, the distribution of word lengths in number of letters has a thicker tail: $99%$ of words have $13$ or fewer letters, compared to the Palasis and Providence corpora where over $99%$ of words had $10$ or fewer letters.
+However, $97.4%$ of German words have $10$ or fewer letters, so the corpus is similarly comprised of short and medium-length words.
 
@@
 ```
+
+We expect that German, as a language with a case system which gets shown through affixation on nouns, will have relatively higher information in the final letters of words compared to English or French.
+We don't observe this, even in unigrams.
+The final letters of German words hold little information compared to the first letter of German words, which is even more noticeable when you have two letters of predictive context for each word.
 
@@
 ```
+
+# General Discussion
+
+Across all three languages we examined and both letters and phonemes in English, we found that the first two or three letters and sounds in a word contain relatively more information than any other letters or sounds in the word.
+In particular, when we knew the letters or sounds that came before, the middles and ends of words became predictable and held little information.
+Even for a language like German which marks case at the end of nouns, listeners can identify and disambiguate the words they hear by the time they have heard the first few sounds.

### 2019-07-08 surprisal/phoneme_writeup.Rmd
fixing percents

@@
 See below a histogram of the distribution of word lengths in the Providence corpus.
-We observe that most words ($84%$) have $5$ letters or fewer, while $99%$ of words in the corpus have fewer than $10$ letters.
+We observe that most words ($84\%$) have $5$ letters or fewer, while $99\%$ of words in the corpus have fewer than $10$ letters.
 There are a tiny number of outliers, mainly onomatopoeia, which have $10$ or more letters.
@@
 
-As seen in the histogram below, almost all words in the Providence corpus ($92%$) have four or fewer phones. $99.5%$ have seven or fewer phones.
+As seen in the histogram below, almost all words in the Providence corpus ($92\%$) have four or fewer phones. $99.5\%$ have seven or fewer phones.
 
@@
 First, we examine the word length frequencies by number of letters in the French Palasis corpus.
-In the Palasis corpus, $99%$ of words have $10$ letters or fewer.
-This French corpus has a higher share of words with six to ten letters than the English Providence corpus: $24%$ versus $16%$.
+In the Palasis corpus, $99\%$ of words have $10$ letters or fewer.
+This French corpus has a higher share of words with six to ten letters than the English Providence corpus: $24\%$ versus $16\%$.
 However, almost all words in both corpora have $10$ or fewer letters.
@@
 
-In the German Wagner corpus, the distribution of word lengths in number of letters has a thicker tail: $99%$ of words have $13$ or fewer letters, compared to the Palasis and Providence corpora where over $99%$ of words had $10$ or fewer letters.
-However, $97.4%$ of German words have $10$ or fewer letters, so the corpus is similarly comprised of short and medium-length words.
+In the German Wagner corpus, the distribution of word lengths in number of letters has a thicker tail: $99\%$ of words have $13$ or fewer letters, compared to the Palasis and Providence corpora where over $99\%$ of words had $10$ or fewer letters.
+However, $97.4\%$ of German words have $10$ or fewer letters, so the corpus is similarly comprised of short and medium-length words.
 

### 2019-07-09 surprisal/wiki_surprisal.Rmd
caching wals

@@
 
-  plot(hang = -1, 
-       main = "Dendrogram of slope embeddings")
+  # plot(hang = -1, 
+  #      main = "Dendrogram of slope embeddings")
 
@@
 
-```{r get features}
-wkpd <- read.csv(here("Data/wikipedia_unigrams.csv"))
+```{r get features, eval = F}
+wkpd <- read_csv(here("Data/wikipedia_unigrams.csv"))
 LANGUAGES <- wkpd$Language
@@
 wf <- wals.feature(FEATURES)
+
+write_csv(wf, here("Data/wals.csv"))
+```
+
+```{r load-wals, eval = T}
+wf <- read_csv(here("Data/wals.csv"))
 ```
@@
 ```
+
 ```{r nominal syntax}

### 2019-07-09 surprisal/WALS_features.Rmd
started feature and cosine similarity comparison analysis

---
title: "Wikipedia WALS Feature Analysis"
author: "Josef Klafka and Dan Yurovsky"
date: "7/9/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lingtypology)
library(feather)
library(here)
knitr::opts_chunk$set(echo = TRUE)
```

```{r }
wkpd <- read.csv(here("Data/wikipedia_unigrams.csv"))
LANGUAGES <- wkpd$language

FEATURES <- c(1:144)
for (i in c(1:length(FEATURES))) {
  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
} #get all word order WALS features in correct syntax

wf <- wals.feature(FEATURES)
```

```{r see which WALS features are the most complete for our languages}
df <- wf %>% 
  filter(language %in% LANGUAGES) %>% 
  filter(rowSums(is.na(.)) < 130) %>% 
  select(`30A`:`121A`) %>% 
  is.na() %>% 
  colSums() %>% 
  sort(decreasing = T)
```

```{r get common languages between the imputations and models}
nom_categories <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, `30A`:`57A`) %>%
  filter(rowSums(is.na(.)) < 12)

nom_syntax <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, `58A`:`64A`) %>%
  filter(rowSums(is.na(.)) < 5)

verb_categories <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, `65A`:`80A`) %>%
  filter(rowSums(is.na(.)) < 8)

word_order <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, `81A`:`97A`) %>%
  select(-`84A`)

clauses <- wf %>%
  filter(language %in% LANGUAGES) %>%
  select(language, `98A`:`121A`) %>%
  filter(rowSums(is.na(.)) < 12)

common_df <- nom_categories %>% 
  semi_join(nom_syntax) %>%
  semi_join(verb_categories) %>%
  semi_join(word_order) %>%
  semi_join(clauses)

common_languages <- common_df %>%
  pull(language)
```

```{r comparison: features to cosines}
imputed_df <- common_df %>% 
  gather(feature, value, -language) %>%
  group_by(feature) %>%
  mutate(value = factor(value)) %>%
  select(language, feature, value) %>%
  ungroup() %>% 
  spread(feature, value) %>% 
  sapply(as.factor) %>%
  as.data.frame() %>%
  column_to_rownames("language") %>%
  MIMCA(ncp = 1)

imputed_df <- as.data.frame(imputed_df[[1]][[99]])

wo_sub <- imputed_df %>% 
  rownames_to_column("language") %>%
  group_by(language) %>%
  slice(1) %>%
  gather(feature, value, -language) %>%
  group_by(feature) %>%
  mutate(value = as.numeric(as.factor(value)))

cosines <- wkpd %>%
  filter(language %in% common_languages) %>%
  column_to_rownames("language") %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, cosine, -language1)

wf_pairwise <- expand.grid(language1 = wo_sub$language,
                           language2 = wo_sub$language) %>%
  left_join(wo_sub, by = c("language1" = "language")) %>%
  rename(value1 = value) %>%
  left_join(wo_sub, by = c("language2" = "language", "feature")) %>%
  rename(value2 = value) %>%
  mutate(same = value1 == value2) %>%
  select(-value1, -value2) %>%
  left_join(cosines, by = c("language1", "language2"))

models <- wf_pairwise %>%
  group_by(feature) %>%
  nest() %>%
  mutate(model = map(data, ~glm(same ~ cosine, 
                                family = "binomial", data = .)))

feature_names <- read_feather(here("Wikipedia/feature_names.feather"))

model_df <- models %>%
  mutate(coeffs = map(model, tidy)) %>%
  select(-data, -model) %>%
  unnest() %>%
  filter(term == "cosine") %>%
  arrange(desc(statistic)) %>%
  left_join(feature_names)
```

### 2019-07-09 surprisal/wiki_surprisal.Rmd
started feature and cosine similarity comparison analysis

--- 2019-07-02 surprisal/wiki_surprisal.Rmd
@@
 wkpd <- read.csv(here("Data/wikipedia_unigrams.csv"))
-LANGUAGES <- wkpd$Language
+LANGUAGES <- wkpd$language
 
@@
   filter(language %in% LANGUAGES) %>%
-  select(language, `30A`:`57A`) #%>%
-wo_wf$non <- rowSums(is.na(wo_wf))
-wo_wf %<>%
-  filter(non < 12) %>%
-  select(-non)
+  select(language, `30A`:`57A`) %>%
+  filter(rowSums(is.na(.)) < 12)
+
 ```
@@
   filter(language %in% LANGUAGES) %>%
-  select(language, `58A`:`64A`) #%>%
-wo_wf$non <- rowSums(is.na(wo_wf))
-wo_wf %<>%
-  filter(non < 5) %>%
-  select(-non)
+  select(language, `58A`:`64A`) %>%
+  filter(rowSums(is.na(.)) < 5)
+
 ```
@@
   filter(language %in% LANGUAGES) %>%
-  select(language, `65A`:`80A`) #%>%
-wo_wf$non <- rowSums(is.na(wo_wf))
-wo_wf %<>%
-  filter(non < 8) %>%
-  select(-non)
+  select(language, `65A`:`80A`) %>%
+  filter(rowSums(is.na(.)) < 8)
+
 ```
@@
   filter(language %in% LANGUAGES) %>%
-  select(language, `98A`:`121A`) #%>%
-wo_wf$non <- rowSums(is.na(wo_wf))
-wo_wf %<>%
-  filter(non < 12) %>%
-  select(-non)
+  select(language, `98A`:`121A`) %>%
+  filter(rowSums(is.na(.)) < 12)
+
 ```
@@
 cosines <- wkpd %>%
-  column_to_rownames("Language") %>%
+  column_to_rownames("language") %>%
   t() %>%

### 2019-07-09 surprisal/wiki_surprisal.Rmd

@@
 
-```{r get features}
+```{r get features, eval=F}
 wkpd <- read.csv(here("Data/wikipedia_unigrams.csv"))

### 2019-07-10 surprisal/WALS_features.Rmd
using existing features data

@@
 library(here)
+library(missMDA)
 knitr::opts_chunk$set(echo = TRUE)
@@
 
-FEATURES <- c(1:144)
-for (i in c(1:length(FEATURES))) {
-  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
-} #get all word order WALS features in correct syntax
-
-wf <- wals.feature(FEATURES)
+wals <- read.csv(here("Data/wals.csv"))
+feature_names <- read_feather(here("Wikipedia/feature_names.feather"))
 ```
@@
 ```{r see which WALS features are the most complete for our languages}
-df <- wf %>% 
+all_languages <- wals %>% 
   filter(language %in% LANGUAGES) %>% 
   filter(rowSums(is.na(.)) < 130) %>%
-  select(`30A`:`121A`) %>% 
-  is.na() %>% 
-  colSums() %>% 
-  sort(decreasing = T)
+  group_by(language) %>%
+  slice(1) %>%
+  ungroup() 
 ```
@@
 ```{r get common languages between the imputations and models}
-nom_categories <- wf %>%
-  filter(language %in% LANGUAGES) %>%
-  select(language, `30A`:`57A`) %>%
+nom_categories <- all_languages %>%
+  select(language, X30A:X57A) %>%
   filter(rowSums(is.na(.)) < 12)
 
-nom_syntax <- wf %>%
-  filter(language %in% LANGUAGES) %>%
-  select(language, `58A`:`64A`) %>%
+nom_syntax <- all_languages %>%
+  select(language, X58A:X64A) %>%
   filter(rowSums(is.na(.)) < 5)
 
-verb_categories <- wf %>%
-  filter(language %in% LANGUAGES) %>%
-  select(language, `65A`:`80A`) %>%
+verb_categories <- all_languages %>%
+  select(language, X65A:X80A) %>%
   filter(rowSums(is.na(.)) < 8)
 
-word_order <- wf %>%
-  filter(language %in% LANGUAGES) %>%
-  select(language, `81A`:`97A`) %>%
-  select(-`84A`)
+word_order <- all_languages %>%
+  select(language, X81A:X97A) %>%
+  select(-X84A)
 
-clauses <- wf %>%
-  filter(language %in% LANGUAGES) %>%
-  select(language, `98A`:`121A`) %>%
+clauses <- all_languages %>%
+  select(language, X98A:X121A) %>%
   filter(rowSums(is.na(.)) < 12)
@@
 common_df <- nom_categories %>% 
-  semi_join(nom_syntax) %>%
-  semi_join(verb_categories) %>%
-  semi_join(word_order) %>%
-  semi_join(clauses)
+  inner_join(nom_syntax) %>%
+  inner_join(verb_categories) %>%
+  inner_join(word_order) %>%
+  inner_join(clauses)
 
@@
 ```{r comparison: features to cosines}
-imputed_df <- common_df %>% 
+#imputed_df <- 
+  common_df %>% 
   gather(feature, value, -language) %>%
@@
   as.data.frame() %>%
-  column_to_rownames("language") %>%
+  column_to_rownames("language") %>% pull(X98A) %>% length()
   MIMCA(ncp = 1)

### 2019-07-10 surprisal/wiki_surprisal.Rmd
using existing features data

@@
 
-Always run this section. It takes time, but this gets all of the WALS features for every language in WALS along with the unigram slope data for all of the Wikipedia languages we've run.
+Always run this section. This gets all of the WALS features, for every language, along with the unigram slope data for all of the Wikipedia languages we've run.
 
-```{r get features, eval=F}
+```{r get features, eval=T}
 wkpd <- read.csv(here("Data/wikipedia_unigrams.csv"))
@@
 
-FEATURES <- c(1:144)
-for (i in c(1:length(FEATURES))) {
-  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
-} #get all word order WALS features in correct syntax
+wf <- read_csv(here("Data/wals.csv"))
 
-wf <- wals.feature(FEATURES)
-
-write_csv(wf, here("Data/wals.csv"))
-```
-
-```{r load-wals, eval = T}
-wf <- read_csv(here("Data/wals.csv"))
 ```
@@
 
-```{r nominal categories}
+```{r nominal categories, eval=F}
 wo_wf <- wf %>%
   filter(language %in% LANGUAGES) %>%
-  select(language, `30A`:`57A`) %>%
+  select(language, X30A:X57A) %>%
   filter(rowSums(is.na(.)) < 12)
@@
 ```
-```{r nominal syntax}
+```{r nominal syntax, eval=F}
 wo_wf <- wf %>%
   filter(language %in% LANGUAGES) %>%
-  select(language, `58A`:`64A`) %>%
+  select(language, X58A:X64A) %>%
   filter(rowSums(is.na(.)) < 5)
@@
 ```
-```{r verbal categories}
+```{r verbal categories, eval=F}
 wo_wf <- wf %>%
   filter(language %in% LANGUAGES) %>%
-  select(language, `65A`:`80A`) %>%
+  select(language, X65A:X80A) %>%
   filter(rowSums(is.na(.)) < 8)
@@
 ```
-```{r word order}
+```{r word order, eval=F}
 wo_wf <- wf %>%
   filter(language %in% LANGUAGES) %>%
-  select(language, `81A`:`97A`) %>%
+  select(language, X81A:X97A) %>%
   select(-`84A`)
 ```
-```{r clauses}
+```{r clauses, eval=F}
 wo_wf <- wf %>%
   filter(language %in% LANGUAGES) %>%
-  select(language, `98A`:`121A`) %>%
+  select(language, X98A:X121A) %>%
   filter(rowSums(is.na(.)) < 12)

### 2019-07-10 surprisal/WALS_features.Rmd
working on imputation of common languages

@@
 
-```{r }
+```{r read in data}
 wkpd <- read.csv(here("Data/wikipedia_unigrams.csv"))
@@
 ```{r comparison: features to cosines}
-#imputed_df <- 
-  common_df %>% 
+imputed_df <- common_df %>% 
   gather(feature, value, -language) %>%
@@
   as.data.frame() %>%
-  column_to_rownames("language") %>% pull(X98A) %>% length()
+  column_to_rownames("language") %>% dim()
   MIMCA(ncp = 1)
@@
 
-feature_names <- read_feather(here("Wikipedia/feature_names.feather"))
-
 model_df <- models %>%

### 2019-07-11 surprisal/WALS_features.Rmd
MIMCA linear dependence understood

@@
 nom_categories <- all_languages %>%
-  select(language, X30A:X57A) %>%
+  select(language, x30a:x57a) %>%
   filter(rowSums(is.na(.)) < 12)
@@
 nom_syntax <- all_languages %>%
-  select(language, X58A:X64A) %>%
+  select(language, x58a:x64a) %>%
   filter(rowSums(is.na(.)) < 5)
@@
 verb_categories <- all_languages %>%
-  select(language, X65A:X80A) %>%
+  select(language, x65a:x80a) %>%
   filter(rowSums(is.na(.)) < 8)
@@
 word_order <- all_languages %>%
-  select(language, X81A:X97A) %>%
-  select(-X84A)
+  select(language, x81a:x97a)
 
 clauses <- all_languages %>%
-  select(language, X98A:X121A) %>%
+  select(language, x98a:x121a) %>%
   filter(rowSums(is.na(.)) < 12)
@@
   as.data.frame() %>%
-  column_to_rownames("language") %>% dim()
+  column_to_rownames("language") %>% 
   MIMCA(ncp = 1)

### 2019-07-11 surprisal/wiki_surprisal.Rmd
MIMCA linear dependence understood

@@
 wf <- read_csv(here("Data/wals.csv")) 
-
+feature_names <- read_feather(here("Wikipedia/feature_names.feather"))
 ```
@@
   filter(language %in% LANGUAGES) %>%
-  select(language, X30A:X57A) %>%
+  select(language, x30A:x57a) %>%
   filter(rowSums(is.na(.)) < 12)
@@
   filter(language %in% LANGUAGES) %>%
-  select(language, X58A:X64A) %>%
+  select(language, x58a:x64a) %>%
   filter(rowSums(is.na(.)) < 5)
@@
   filter(language %in% LANGUAGES) %>%
-  select(language, X65A:X80A) %>%
+  select(language, x65a:x80a) %>%
   filter(rowSums(is.na(.)) < 8)
@@
   filter(language %in% LANGUAGES) %>%
-  select(language, X81A:X97A) %>%
-  select(-`84A`)
+  select(language, x81Aa:x97a) 
 ```
@@
   filter(language %in% LANGUAGES) %>%
-  select(language, X98A:X121A) %>%
+  select(language, x98a:x121a) %>%
   filter(rowSums(is.na(.)) < 12)

### 2019-07-16 surprisal/wiki_surprisal.Rmd
surprisal hclust

@@
 ```{r make dendrogram}
-wiki <- read.csv(here("Data/wikipedia_unigrams.csv")) %>%
-  select(-X) %>% 
-  column_to_rownames("Language")
+wiki_unigrams <- read.csv(here("Data/relative_ngrams.csv")) %>%
+  filter(gram == "Unigram") %>%
+  sample_n(100) %>%
+  column_to_rownames("language")
 
-wiki %>% 
+wiki_unigrams %>% 
   dist() %>% 

### 2019-07-16 surprisal/wiki_surprisal.Rmd
trying to join wals and wiki

--- 2019-07-10 surprisal/wiki_surprisal.Rmd
@@
 library(ggdendro)
+library(jsonlite)
 
@@
 
-Always run this section. This gets all of the WALS features, for every language, along with the unigram slope data for all of the Wikipedia languages we've run.
+Always run this section. It takes time, but this gets all of the WALS features for every language in WALS along with the unigram slope data for all of the Wikipedia languages we've run.
 
-```{r get features, eval=T}
-wkpd <- read.csv(here("Data/wikipedia_unigrams.csv"))
-LANGUAGES <- wkpd$language
+```{r get features, eval = F}
+wkpd <- read_csv(here("Data/wikipedia_unigrams.csv"))
+LANGUAGES <- wkpd$Language
 
+FEATURES <- c(1:144)
+for (i in c(1:length(FEATURES))) {
+  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
+} #get all word order WALS features in correct syntax
+
+wals_isos <- wals %>%
+  mutate(iso = iso.gltc(glottocode)) %>%
+  select(-glottocode)
+
+wiki_isos <- read_csv(here("Wikipedia/wiki_isos.csv")) %>%
+  clean_names() %>%
+  select(x639_1, x639_3)
+
+wiki_langs <- fromJSON(here("/Wikipedia/language_dict.json")) %>%
+  as_tibble() %>%
+  gather(language, x639_1) %>%
+  left_join(wiki_isos)
+
+%>%
+  left_join(wals_isos) %>%
+  filter(is.na(wals.code)) %>%
+  View()
+
+wf <- wals.feature(FEATURES)
+
+write_csv(wf, here("Data/wals.csv"))
+```
+
+```{r load-wals, eval = T}
 wf <- read_csv(here("Data/wals.csv"))
-
 ```
@@
 
-```{r nominal categories, eval=F}
+```{r nominal categories}
 wo_wf <- wf %>%
   filter(language %in% LANGUAGES) %>%
-  select(language, X30A:X57A) %>%
-  filter(rowSums(is.na(.)) < 12)
+  select(language, `30A`:`57A`) #%>%
+wo_wf$non <- rowSums(is.na(wo_wf))
+wo_wf %<>%
+  filter(non < 12) %>%
+  select(-non)
+```
 
-```
-```{r nominal syntax, eval=F}
+```{r nominal syntax}
 wo_wf <- wf %>%
   filter(language %in% LANGUAGES) %>%
-  select(language, X58A:X64A) %>%
-  filter(rowSums(is.na(.)) < 5)
-
+  select(language, `58A`:`64A`) #%>%
+wo_wf$non <- rowSums(is.na(wo_wf))
+wo_wf %<>%
+  filter(non < 5) %>%
+  select(-non)
 ```
-```{r verbal categories, eval=F}
+```{r verbal categories}
 wo_wf <- wf %>%
   filter(language %in% LANGUAGES) %>%
-  select(language, X65A:X80A) %>%
-  filter(rowSums(is.na(.)) < 8)
-
+  select(language, `65A`:`80A`) #%>%
+wo_wf$non <- rowSums(is.na(wo_wf))
+wo_wf %<>%
+  filter(non < 8) %>%
+  select(-non)
 ```
-```{r word order, eval=F}
+```{r word order}
 wo_wf <- wf %>%
   filter(language %in% LANGUAGES) %>%
-  select(language, X81A:X97A) %>%
+  select(language, `81A`:`97A`) %>%
   select(-`84A`)
 ```
-```{r clauses, eval=F}
+```{r clauses}
 wo_wf <- wf %>%
   filter(language %in% LANGUAGES) %>%
-  select(language, X98A:X121A) %>%
-  filter(rowSums(is.na(.)) < 12)
-
+  select(language, `98A`:`121A`) #%>%
+wo_wf$non <- rowSums(is.na(wo_wf))
+wo_wf %<>%
+  filter(non < 12) %>%
+  select(-non)
 ```
@@
 cosines <- wkpd %>%
-  column_to_rownames("language") %>%
+  column_to_rownames("Language") %>%
   t() %>%

### 2019-07-16 surprisal/WALS_features.Rmd

@@
 ```{r read in data}
-wkpd <- read.csv(here("Data/wikipedia_unigrams.csv"))
-LANGUAGES <- wkpd$language
+wals <- read_csv(here("Data/wals.csv"))
+feature_names <- read_feather(here("Wikipedia/feature_names.feather"))
 
-wals <- read.csv(here("Data/wals.csv"))
-feature_names <- read_feather(here("Wikipedia/feature_names.feather"))
+wiki <- read_csv(here("Data/relative_ngrams.csv"))
+LANGUAGES <- wiki %>% pull(language) %>% unique()
 ```
@@
 ```
+
+## Errata
+
+```{r cleaning}
+wiki <- read_feather(here("Data/relative_ngrams.feather"))
+wiki %>% 
+  group_by(language) %>% 
+  mutate(order = 1:n()) %>% 
+  ungroup() %>% 
+  filter(order <= 15) %>% 
+  select(-order) %>%
+  select(language, gram, estimate, cut) %>% 
+  spread(cut, estimate) %>% 
+  rename(Slope1 =`1`, Slope2 =`2`, Slope3 =`3`, Slope4 =`4`, Slope5 =`5`) %>%
+  write_csv(here("Data/relative_ngrams.csv"))
+```

### 2019-07-16 surprisal/wiki_surprisal.Rmd

--- 2019-07-16 surprisal/wiki_surprisal.Rmd
@@
 
+FEATURES <- c(1:144)
+for (i in c(1:length(FEATURES))) {
+  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
+} #get all word order WALS features in correct syntax
+
+wals_isos <- wals %>%
+  mutate(iso = iso.gltc(glottocode)) %>%
+  select(-glottocode)
+
+wiki_isos <- read_csv(here("Wikipedia/wiki_isos.csv")) %>%
+  clean_names() %>%
+  select(x639_1, x639_3)
+
+wiki_langs <- fromJSON(here("/Wikipedia/language_dict.json")) %>%
+  as_tibble() %>%
+  gather(language, x639_1) %>%
+  left_join(wiki_isos)
+
+%>%
+  left_join(wals_isos) %>%
+  filter(is.na(wals.code)) %>%
+  View()
+
+wf <- wals.feature(FEATURES)
+
+write_csv(wf, here("Data/wals.csv"))
+```
+
+```{r load-wals, eval = T}
 wf <- read_csv(here("Data/wals.csv"))
-feature_names <- read_feather(here("Wikipedia/feature_names.feather"))
+
 ```

### 2019-07-16 surprisal/wiki_surprisal.Rmd
iso codes matched up

@@
 ```{r get features, eval=T}
-wkpd <- read.csv(here("Data/wikipedia_unigrams.csv"))
-LANGUAGES <- wkpd$language
+wiki <- read_csv(here("Data/relative_ngrams.csv"))
+LANGUAGES <- wiki %>% pull(language) %>% unique()
 
-FEATURES <- c(1:144)
-for (i in c(1:length(FEATURES))) {
-  FEATURES[i] <- paste(FEATURES[i], "A", sep = "")
-} #get all word order WALS features in correct syntax
+codes <- read_csv(here("Wikipedia/isos.csv"))
+
+wals <- read_csv(here("Data/wals.csv"))
 
@@
 
-wiki_isos <- read_csv(here("Wikipedia/wiki_isos.csv")) %>%
-  clean_names() %>%
-  select(x639_1, x639_3)
-
-wiki_langs <- fromJSON(here("/Wikipedia/language_dict.json")) %>%
-  as_tibble() %>%
-  gather(language, x639_1) %>%
-  left_join(wiki_isos)
-
-%>%
-  left_join(wals_isos) %>%
-  filter(is.na(wals.code)) %>%
-  View()
-
-wf <- wals.feature(FEATURES)
-
-write_csv(wf, here("Data/wals.csv"))
+wals_isos %>% 
+  inner_join(codes, by = "iso") %>% 
+  select(-language.x) %>%
+  rename(language = language.y) %>% 
+  inner_join(wiki %>% inner_join(codes, by = "language"), by = "iso")
 ```
@@
 ```
+
+```{r setup}
+wiki_isos <- read_csv(here("Wikipedia/wiki_isos.csv")) %>%
+  clean_names() %>%
+  select(iso_language_name, x639_1, x639_3) %>% 
+  mutate(x639_3 = str_extract(x639_3, "\\w+")) %>% 
+  rename(language = iso_language_name) %>% 
+  filter(language %in% LANGUAGES)
+
+wiki_langs <- fromJSON(here("Wikipedia/language_dict.json")) %>%
+  as_tibble() %>%
+  gather(language, x639_1) %>%
+  filter(language %in% LANGUAGES) %>%
+  left_join(wiki_isos)
+
+%>%
+  left_join(wals_isos) %>%
+  filter(is.na(wals.code))
+```

### 2019-07-16 surprisal/WALS_features.Rmd
not enough memory to run models for all of the features

@@
 library(missMDA)
+library(lsa)
 knitr::opts_chunk$set(echo = TRUE)
@@
 ```{r read in data}
-wals <- read_csv(here("Data/wals.csv"))
+wals <- read_csv(here("Data/wals_isos.csv"))
 feature_names <- read_feather(here("Wikipedia/feature_names.feather"))
@@
 all_languages <- wals %>% 
-  filter(language %in% LANGUAGES) %>% 
   filter(rowSums(is.na(.)) < 130) %>%
@@
 
-```{r comparison: features to cosines}
-imputed_df <- common_df %>% 
-  gather(feature, value, -language) %>%
-  group_by(feature) %>%
-  mutate(value = factor(value)) %>%
-  select(language, feature, value) %>%
-  ungroup() %>% 
-  spread(feature, value) %>% 
-  sapply(as.factor) %>%
-  as.data.frame() %>%
-  column_to_rownames("language") %>% 
-  MIMCA(ncp = 1)
+```{r models}
+imputed_df <- read_csv(here("Data/imputed_wals.csv"))
 
-imputed_df <- as.data.frame(imputed_df[[1]][[99]])
-
-wo_sub <- imputed_df %>% 
-  rownames_to_column("language") %>%
+wf_sub <- imputed_df %>% 
   group_by(language) %>%
@@
 
-cosines <- wkpd %>%
-  filter(language %in% common_languages) %>%
+cosines <- wiki %>%
+  filter(language %in% common_languages, gram == "Unigram") %>%
+  select(-gram) %>% 
   column_to_rownames("language") %>%
@@
 
-wf_pairwise <- expand.grid(language1 = wo_sub$language,
-                           language2 = wo_sub$language) %>%
-  left_join(wo_sub, by = c("language1" = "language")) %>%
+wf_pairwise <- expand.grid(language1 = wf_sub$language,
+                           language2 = wf_sub$language) %>%
+  left_join(wf_sub, by = c("language1" = "language")) %>%
   rename(value1 = value) %>%
-  left_join(wo_sub, by = c("language2" = "language", "feature")) %>%
+  left_join(wf_sub, by = c("language2" = "language", "feature")) %>%
   rename(value2 = value) %>%
@@
 
+```{r imputation}
+imputed_df <- common_df %>% 
+  gather(feature, value, -language) %>%
+  group_by(feature) %>%
+  mutate(value = factor(value)) %>%
+  select(language, feature, value) %>%
+  ungroup() %>% 
+  spread(feature, value) %>% 
+  sapply(as.factor) %>%
+  as.data.frame() %>%
+  column_to_rownames("language") %>% 
+  MIMCA(ncp = 1)
+
+imputed_df <- as.data.frame(imputed_df[[1]][[99]]) 
+```
+
 ## Errata

### 2019-07-16 surprisal/wiki_surprisal.Rmd
not enough memory to run models for all of the features

@@
 
-wals_isos <- wals %>%
+wals %>%
   mutate(iso = iso.gltc(glottocode)) %>%
-  select(-glottocode)
-
-wals_isos %>% 
+  select(-glottocode) %>% 
   inner_join(codes, by = "iso") %>% 
@@
   rename(language = language.y) %>%
-  inner_join(wiki %>% inner_join(codes, by = "language"), by = "iso")
+  write_csv(here("Data/wals_isos.csv"))
+
+#  inner_join(wiki %>% inner_join(codes, by = "language"), by = "iso")
 ```

### 2019-07-17 surprisal/MacWhinney poster/poster_plotting.Rmd
feature space and cosine distances joined

@@
 ```{r get_utterances}
-childes <- read_feather(here("Data/childes_surprisals.feather")) %>%
+childes <- read_feather(here("Data/childes_words.feather")) %>%
   clean_names() %>%
@@
 
-switchboard <- read_feather(here("Data/switchboard_surprisals.feather")) %>%
+switchboard <- read_feather(here("Data/switchboard_words.feather")) %>%
   clean_names() %>%
@@
 
-pdf("surprisals.pdf", width = 10, height = 5.5)
+# pdf("surprisals.pdf", width = 10, height = 5.5)
 data %>%
-  filter(length == 8) %>%
+  filter(length == 8, language != "Mandarin") %>%
    ggplot(aes(x = word_order, y = s,
@@
                         label = speaker_role)) +
-  facet_grid(context ~ language, scales = "free") +
+  facet_grid(language~context, scales = "free") +
   geom_point(size = 1.5) +
@@
   labs(x = "Word Position", y = "Average Information")
-dev.off()
+# dev.off()
 ```

### 2019-07-17 surprisal/WALS_features.Rmd
feature space and cosine distances joined

@@
 library(lsa)
+library(reshape2)
 knitr::opts_chunk$set(echo = TRUE)
@@
   pull(language)
+```
+
+```{r feature space}
+imputed_df <- read_csv(here("Data/imputed_wals.csv"))
+
+feature_space <- imputed_df %>% 
+  gather(feature, value, -language) %>% 
+  group_by(feature) %>% 
+  mutate(value = as.numeric(as.factor(value))) %>% 
+  spread(feature, value) %>%
+  column_to_rownames("language") %>%
+  dist(method = "manhattan") %>%
+  as.matrix() %>%
+  melt(value.name = "distance") %>%
+  as_tibble() %>%
+  rename(language1 = Var1, language2 = Var2)
+
+feature_space %>%
+  count(distance) %>%
+  mutate(n = n/2) %>% 
+  ggplot(aes(x = distance, y = n)) + 
+    geom_col() + 
+    scale_y_continuous(breaks = 1:12)
+```
+
+```{r cosines}
+cosines <- wiki %>%
+  filter(language %in% common_languages, gram == "Unigram") %>%
+  select(-gram) %>% 
+  column_to_rownames("language") %>%
+  t() %>%
+  cosine() %>%
+  as_data_frame(rownames = "language1") %>%
+  gather(language2, cosine, -language1) %>%
+  mutate(cosine = 1 - abs(cosine))
+
+all_distances <- cosines %>%
+  left_join(feature_space)
 ```

### 2019-07-17 surprisal/WALS_features.Rmd
plotting feature space vs cosine space distance

--- 2019-07-16 surprisal/WALS_features.Rmd
@@
   pull(language)
+```
+
+```{r feature space}
+imputed_df <- read_csv(here("Data/imputed_wals.csv"))
+
+feature_space <- imputed_df %>% 
+  gather(feature, value, -language) %>% 
+  group_by(feature) %>% 
+  mutate(value = as.numeric(as.factor(value))) %>% 
+  spread(feature, value) %>%
+  column_to_rownames("language") %>%
+  dist(method = "manhattan") %>%
+  as.matrix() %>%
+  melt(value.name = "feature") %>%
+  as_tibble() %>%
+  rename(language1 = Var1, language2 = Var2) 
+```
+
+```{r cosines}
+unigram_cosines <- wiki %>%
+  filter(language %in% common_languages, gram == "Unigram") %>%
+  select(-gram) %>% 
+  column_to_rownames("language") %>%
+  t() %>%
+  cosine() %>%
+  as_data_frame(rownames = "language1") %>%
+  gather(language2, unigram_cosine, -language1) %>%
+  mutate(unigram_cosine = 1 - abs(unigram_cosine))
+
+bigram_cosines <- wiki %>%
+  filter(language %in% common_languages, gram == "Bigram") %>%
+  select(-gram) %>% 
+  column_to_rownames("language") %>%
+  t() %>%
+  cosine() %>%
+  as_data_frame(rownames = "language1") %>%
+  gather(language2, bigram_cosine, -language1) %>%
+  mutate(bigram_cosine = 1 - abs(bigram_cosine))
+
+trigram_cosines <- wiki %>%
+  filter(language %in% common_languages, gram == "Trigram") %>%
+  select(-gram) %>% 
+  column_to_rownames("language") %>%
+  t() %>%
+  cosine() %>%
+  as_data_frame(rownames = "language1") %>%
+  gather(language2, trigram_cosine, -language1) %>%
+  mutate(trigram_cosine = 1 - abs(trigram_cosine))
+
+all_distances <- feature_space %>%
+  left_join(unigram_cosines) %>%
+  left_join(bigram_cosines) %>%
+  left_join(trigram_cosines)
+
+all_distances %>% 
+  ggplot(aes(x = feature)) +
+    geom_smooth(aes(y = unigram_cosine, color = "#0033CC"), 
+                method = 'lm', se = F) + 
+    geom_smooth(aes(y = bigram_cosine, color = "#6699FF"), 
+                method = 'lm', se = F) +
+    geom_smooth(aes(y = trigram_cosine, color = "#33CCFF"), 
+                method = 'lm', se = F) +
+    ylab("Cosine Space Distance") + 
+    xlab("Feature Space Distance") + 
+    scale_color_manual(scale_name = "gram", 
+                       values = c("#0033CC", "#6699FF", "#33CCFF"), 
+                       labels = c("Unigrams", "Bigrams", "Trigrams")) + 
+    ggtitle("Cosine space distance regressed on feature spaced distance")
 ```

### 2019-07-17 surprisal/WALS_features.Rmd
changes

@@
                 method = 'lm', se = F) +
+    geom_point(aes(y = unigram_cosine, color = "#0033CC")) + 
+    geom_point(aes(y = bigram_cosine, color = "#6699FF")) + 
+    geom_point(aes(y = trigram_cosine, color = "#33CCFF")) + 
     ylab("Cosine Space Distance") + 
     xlab("Feature Space Distance") + 
-    scale_color_manual(scale_name = "gram", 
+    scale_color_manual(name = "Gram", 
                        values = c("#0033CC", "#6699FF", "#33CCFF"), 

### 2019-07-17 surprisal/WALS_features.Rmd
models for 70 odd features

@@
 library(reshape2)
+library(broom)
 knitr::opts_chunk$set(echo = TRUE)
@@
 wals <- read_csv(here("Data/wals_isos.csv"))
+
 feature_names <- read_feather(here("Wikipedia/feature_names.feather"))
+clean_names <- feature_names %>% mutate(feature = make_clean_names(feature))
 
@@
   arrange(desc(statistic)) %>%
-  left_join(feature_names)
+  left_join(clean_names, by = "feature")
 ```

### 2019-07-18 surprisal/MacWhinney poster/poster_plotting.Rmd
matched up wals features and wiki slopes

@@
                         label = speaker_role)) +
-  facet_grid(language~context, scales = "free") +
+  facet_grid(context~language, scales = "free") +
   geom_point(size = 1.5) +
@@
 ```
+
+```{r plotting features}
+models <- read_feather(here("Wikipedia/models.feather"))
+
+sf <- models %>% 
+  filter(p.value == 0, estimate > 0) %>% 
+  select(name) %>% 
+  rename("Some Important Features" = name) %>% 
+  slice(c(1, 10, 12, 15, 23, 29))
+
+sf$Examples <- c("Noun-relative clause; relative clause-noun", 
+                 "No cases; 6 cases", 
+                 "SOV; VSO", 
+                 "Only singular; singular and plural", 
+                 "Affix; distinct word", 
+                 "Prefixes; suffixes")
+
+sf %>% grid.table()
+
+```

### 2019-07-18 surprisal/WALS_features.Rmd
matched up wals features and wiki slopes

@@
 library(broom)
+library(janitor)
 knitr::opts_chunk$set(echo = TRUE)
@@
 ```{r read in data}
-wals <- read_csv(here("Data/wals_isos.csv"))
+# wals <- read_csv(here("Data/wals_isos.csv"))
+wals <- read_csv(here("Data/from_website.csv"))
 
@@
 
+codes <- read_csv(here("Wikipedia/isos.csv"))
 wiki <- read_csv(here("Data/relative_ngrams.csv"))
@@
 
-```{r see which WALS features are the most complete for our languages}
+```{r get only languages with enough features}
 all_languages <- wals %>% 
-  filter(rowSums(is.na(.)) < 130) %>%
+  mutate(missing = rowSums(is.na(.))) %>% 
   group_by(language) %>% 
-  slice(1) %>%
+  mutate(minimum = min(missing)) %>%
+  filter(missing == minimum) %>% 
+  select(-missing, minimum) %>%
   ungroup() 
@@
 
-common_languages <- common_df %>%
-  pull(language)
+# common_languages <- common_df %>%
+#   pull(language)
 ```
@@
 ```{r cosines}
+common_languages <- read_csv(here("Data/imputed_wals.csv")) %>% 
+  pull(language)
+
 unigram_cosines <- wiki %>%
@@
 all_distances %>% 
+  select(-bigram_cosine) %>% 
   ggplot(aes(x = feature)) +
@@
                 method = 'lm', se = F) + 
-    geom_smooth(aes(y = bigram_cosine, color = "#6699FF"), 
-                method = 'lm', se = F) +
+    # geom_smooth(aes(y = bigram_cosine, color = "#6699FF"), 
+    #             method = 'lm', se = F) +
     geom_smooth(aes(y = trigram_cosine, color = "#33CCFF"), 
@@
     geom_point(aes(y = unigram_cosine, color = "#0033CC")) + 
-    geom_point(aes(y = bigram_cosine, color = "#6699FF")) + 
+    # geom_point(aes(y = bigram_cosine, color = "#6699FF")) + 
     geom_point(aes(y = trigram_cosine, color = "#33CCFF")) + 
-    ylab("Cosine Space Distance") + 
-    xlab("Feature Space Distance") + 
+    ylab("Information Curve Difference") + 
+    xlab("Difference in WALS Features") + 
     scale_color_manual(name = "Gram", 
-                       values = c("#0033CC", "#6699FF", "#33CCFF"), 
-                       labels = c("Unigrams", "Bigrams", "Trigrams")) + 
-    ggtitle("Cosine space distance regressed on feature spaced distance")
+                       values = c("#0033CC", "#33CCFF"), 
+                       labels = c("Unigrams", "Trigrams")) 
 ```
@@
 
-wf_pairwise <- expand.grid(language1 = wf_sub$language,
-                           language2 = wf_sub$language) %>%
-  left_join(wf_sub, by = c("language1" = "language")) %>%
+sub_langs <- wf_sub %>%
+  pull(language) %>%
+  unique()
+
+wf_pairwise <- expand.grid(language1 = sub_langs,
+                           language2 = sub_langs) %>%
+  filter(language1 != language2)
+
+compute_sim <- function(language1, language2) {
+
+  wf_sub %>%
+    filter(language == language1) %>%
     rename(value1 = value) %>% 
-  left_join(wf_sub, by = c("language2" = "language", "feature")) %>%
-  rename(value2 = value) %>%
-  mutate(same = value1 == value2) %>%
-  select(-value1, -value2) %>%
-  left_join(cosines, by = c("language1", "language2"))
+    left_join(wf_sub %>% filter(language == language2) %>% select(feature, value), 
+              by = "feature") %>%
+    mutate(same = value == value1) %>%
+    ungroup() %>%
+    group_by(language) %>%
+    summarise(feature = sum(same)) %>%
+    mutate(language2 = language2) %>%
+    rename(language1 = language)
+}
+
+feature_space <- map_dfr(1:nrow(wf_pairwise), ~compute_sim(wf_pairwise[.x,"language1"],
+                                          wf_pairwise[.x,"language2"]))
 
@@
 
-```{r cleaning}
+```{r get values from website}
+languages <- read_csv(here("wals_dataset/languages.csv"))
+
+values <- read_csv(here("wals_dataset/values.csv"))
+
+values %>% 
+  left_join(languages, by = c("Language_ID" = "ID")) %>% 
+  select(Language_ID, Parameter_ID, Value, Name, Glottocode, ISO639P3code) %>% 
+  rename(wals_code = Language_ID, feature = Parameter_ID, value = Value, 
+         language = Name, glottocode = Glottocode, iso = ISO639P3code) %>% 
+  spread(feature, value)
+
+wf %>% write_csv(here("Data/from_website.csv"))
+```
+
+```{r pairing up languages with wikipedia features}
+wals <- read_csv(here("Data/from_website.csv"))
+
+all_data <- wiki %>% 
+  left_join(codes, by = "language") %>% 
+  left_join(wals %>% select(-language), by = "iso") %>% 
+  mutate(n = rowSums(is.na(.))) %>% 
+  group_by(language) %>% 
+  filter(n == min(n)) %>% 
+  group_by(language, gram) %>%
+  slice(1) %>% 
+  ungroup()
+
+all_data %>% 
+  write_csv(here("Data/wiki_wals.csv"))
+```
+
+```{r cleaning, eval = F}
 wiki <- read_feather(here("Data/relative_ngrams.feather"))

### 2019-07-18 surprisal/wiki_surprisal.Rmd
matched up wals features and wiki slopes

@@
   filter(gram == "Unigram") %>%
-  sample_n(100) %>%
   column_to_rownames("language")
@@
   select(-glottocode) %>% 
+  select(language, iso) %>% View()
+
   inner_join(codes, by = "iso") %>% 
@@
   select(iso_language_name, x639_1, x639_3) %>% 
-  mutate(x639_3 = str_extract(x639_3, "\\w+")) %>% 
+  # mutate(x639_3 = str_extract(x639_3, "\\w+")) %>% 
   rename(language = iso_language_name) %>% 
@@
   filter(language %in% LANGUAGES) %>%
+  write_csv(here("wiki_to_iso.csv"))
+
+%>%
   left_join(wiki_isos)
@@
 ```
+
+```{r plotting grams across languages, eval = F}
+# how much do the slopes differ from 0? 
+wiki %>% 
+  group_by(language, gram) %>% 
+  summarise(diff = sum(abs(Slope1:Slope5))) %>%
+  ggplot(aes(x = gram, y = diff)) + 
+  geom_point()
+```

### 2019-07-19 surprisal/WALS_features.Rmd
most features imputed for all languages

@@
 library(janitor)
+library(hexbin)
 knitr::opts_chunk$set(echo = TRUE)
@@
 
-```{r get common languages between the imputations and models}
-nom_categories <- all_languages %>%
-  select(language, x30a:x57a) %>%
-  filter(rowSums(is.na(.)) < 12)
+```{r imputation}
+df <- read_csv(here("Data/wiki_wals.csv"))
 
-nom_syntax <- all_languages %>%
-  select(language, x58a:x64a) %>%
-  filter(rowSums(is.na(.)) < 5)
+wals <- df %>% 
+  select(language, ends_with('A')) %>%
+  group_by(language) %>%
+  slice(1) %>% 
+  ungroup()
 
-verb_categories <- all_languages %>%
-  select(language, x65a:x80a) %>%
-  filter(rowSums(is.na(.)) < 8)
-
-word_order <- all_languages %>%
-  select(language, x81a:x97a)
-
-clauses <- all_languages %>%
-  select(language, x98a:x121a) %>%
-  filter(rowSums(is.na(.)) < 12)
-
-common_df <- nom_categories %>%
-  inner_join(nom_syntax) %>%
-  inner_join(verb_categories) %>%
-  inner_join(word_order) %>%
-  inner_join(clauses)
-
-# common_languages <- common_df %>%
-#   pull(language)
-```
-
-```{r feature space}
-imputed_df <- read_csv(here("Data/imputed_wals.csv"))
-
-feature_space <- imputed_df %>% 
+imputed_df <- wals %>% 
+  select(-`139A`, -`141A`, -`140A`) %>% # these features have no values
+  select(language, `121A`:`120A`) %>%
   gather(feature, value, -language) %>%
   group_by(feature) %>%
-  mutate(value = as.numeric(as.factor(value))) %>% 
+  mutate(value = factor(value)) %>%
+  select(language, feature, value) %>%
+  ungroup() %>% 
   spread(feature, value) %>% 
+  sapply(as.factor) %>%
+  as.data.frame() %>%
   column_to_rownames("language") %>% 
-  dist(method = "manhattan") %>%
-  as.matrix() %>%
-  melt(value.name = "feature") %>%
-  as_tibble() %>%
-  rename(language1 = Var1, language2 = Var2) 
-```
+  MIMCA(ncp = 2, threshold = 1e-02)
 
-```{r cosines}
-common_languages <- read_csv(here("Data/imputed_wals.csv")) %>% 
-  pull(language)
-
-unigram_cosines <- wiki %>%
-  filter(language %in% common_languages, gram == "Unigram") %>%
-  select(-gram) %>% 
-  column_to_rownames("language") %>%
-  t() %>%
-  cosine() %>%
-  as_data_frame(rownames = "language1") %>%
-  gather(language2, unigram_cosine, -language1) %>%
-  mutate(unigram_cosine = 1 - abs(unigram_cosine))
-
-bigram_cosines <- wiki %>%
-  filter(language %in% common_languages, gram == "Bigram") %>%
-  select(-gram) %>% 
-  column_to_rownames("language") %>%
-  t() %>%
-  cosine() %>%
-  as_data_frame(rownames = "language1") %>%
-  gather(language2, bigram_cosine, -language1) %>%
-  mutate(bigram_cosine = 1 - abs(bigram_cosine))
-
-trigram_cosines <- wiki %>%
-  filter(language %in% common_languages, gram == "Trigram") %>%
-  select(-gram) %>% 
-  column_to_rownames("language") %>%
-  t() %>%
-  cosine() %>%
-  as_data_frame(rownames = "language1") %>%
-  gather(language2, trigram_cosine, -language1) %>%
-  mutate(trigram_cosine = 1 - abs(trigram_cosine))
-
-all_distances <- feature_space %>%
-  left_join(unigram_cosines) %>%
-  left_join(bigram_cosines) %>%
-  left_join(trigram_cosines)
-
-all_distances %>% 
-  select(-bigram_cosine) %>% 
-  ggplot(aes(x = feature)) +
-    geom_smooth(aes(y = unigram_cosine, color = "#0033CC"), 
-                method = 'lm', se = F) + 
-    # geom_smooth(aes(y = bigram_cosine, color = "#6699FF"), 
-    #             method = 'lm', se = F) +
-    geom_smooth(aes(y = trigram_cosine, color = "#33CCFF"), 
-                method = 'lm', se = F) +
-    geom_point(aes(y = unigram_cosine, color = "#0033CC")) + 
-    # geom_point(aes(y = bigram_cosine, color = "#6699FF")) + 
-    geom_point(aes(y = trigram_cosine, color = "#33CCFF")) + 
-    ylab("Information Curve Difference") + 
-    xlab("Difference in WALS Features") + 
-    scale_color_manual(name = "Gram", 
-                       values = c("#0033CC", "#33CCFF"), 
-                       labels = c("Unigrams", "Trigrams")) 
+imputed_df110 <- as.data.frame(imputed_df[[1]][[99]]) 
 ```
@@
 ```{r models}
-imputed_df <- read_csv(here("Data/imputed_wals.csv"))
+imputed_df <- read_csv(here("Data/imputed_all.csv"))
 
-wf_sub <- imputed_df %>% 
+wf_sub <- all_imputed %>% 
   group_by(language) %>%
@@
 
-```{r imputation}
-imputed_df <- common_df %>% 
-  gather(feature, value, -language) %>%
-  group_by(feature) %>%
-  mutate(value = factor(value)) %>%
-  select(language, feature, value) %>%
-  ungroup() %>% 
-  spread(feature, value) %>% 
-  sapply(as.factor) %>%
-  as.data.frame() %>%
+```{r cosines}
+unigram_cosines <- wiki %>%
+  filter(language %in% sub_langs, gram == "Unigram") %>%
+  select(-gram) %>% 
   column_to_rownames("language") %>%
-  MIMCA(ncp = 1)
+  t() %>%
+  cosine() %>%
+  as_data_frame(rownames = "language1") %>%
+  gather(language2, unigram_cosine, -language1) %>%
+  mutate(unigram_cosine = 1 - abs(unigram_cosine))
 
-imputed_df <- as.data.frame(imputed_df[[1]][[99]]) 
+# bigram_cosines <- wiki %>%
+#   filter(language %in% sub_langs, gram == "Bigram") %>%
+#   select(-gram) %>% 
+#   column_to_rownames("language") %>%
+#   t() %>%
+#   cosine() %>%
+#   as_data_frame(rownames = "language1") %>%
+#   gather(language2, bigram_cosine, -language1) %>%
+#   mutate(bigram_cosine = 1 - abs(bigram_cosine))
+
+trigram_cosines <- wiki %>%
+  filter(language %in% sub_langs, gram == "Trigram") %>%
+  select(-gram) %>% 
+  column_to_rownames("language") %>%
+  t() %>%
+  cosine() %>%
+  as_data_frame(rownames = "language1") %>%
+  gather(language2, trigram_cosine, -language1) %>%
+  mutate(trigram_cosine = 1 - abs(trigram_cosine))
+
+all_distances <- feature_space %>%
+  left_join(unigram_cosines) %>%
+  left_join(trigram_cosines)
+
+all_distances %>% 
+  ggplot(aes(x = feature)) +
+    geom_smooth(aes(y = unigram_cosine, color = "#0033CC"), 
+                method = 'lm', se = F) + 
+    # geom_smooth(aes(y = bigram_cosine, color = "#6699FF"), 
+    #             method = 'lm', se = F) +
+    geom_smooth(aes(y = trigram_cosine, color = "#33CCFF"), 
+                method = 'lm', se = F) +
+    geom_hex(aes(y = unigram_cosine, color = "#0033CC", alpha=.2)) + 
+    # geom_point(aes(y = bigram_cosine, color = "#6699FF")) + 
+    geom_hex(aes(y = trigram_cosine, color = "#33CCFF", alpha=.2)) + 
+    ylab("Information Curve Difference") + 
+    xlab("Difference in WALS Features") + 
+    scale_color_manual(name = "Gram", 
+                       values = c("#0033CC", "#33CCFF"), 
+                       labels = c("Unigrams", "Trigrams")) 
 ```
@@
 ```
+
+```{r feature space}
+imputed_df <- read_csv(here("Data/imputed_all.csv"))
+
+feature_space <- all_imputed %>% 
+  gather(feature, value, -language) %>% 
+  group_by(feature) %>% 
+  mutate(value = as.numeric(as.factor(value))) %>% 
+  spread(feature, value) %>%
+  column_to_rownames("language") %>%
+  dist(method = "manhattan") %>%
+  as.matrix() %>%
+  melt(value.name = "feature") %>%
+  as_tibble() %>%
+  rename(language1 = Var1, language2 = Var2) 
+```
+
+```{r get common languages between the imputations and models}
+nom_categories <- all_languages %>%
+  select(language, x30a:x57a) %>%
+  filter(rowSums(is.na(.)) < 12)
+
+nom_syntax <- all_languages %>%
+  select(language, x58a:x64a) %>%
+  filter(rowSums(is.na(.)) < 5)
+
+verb_categories <- all_languages %>%
+  select(language, x65a:x80a) %>%
+  filter(rowSums(is.na(.)) < 8)
+
+word_order <- all_languages %>%
+  select(language, x81a:x97a)
+
+clauses <- all_languages %>%
+  select(language, x98a:x121a) %>%
+  filter(rowSums(is.na(.)) < 12)
+
+common_df <- nom_categories %>%
+  inner_join(nom_syntax) %>%
+  inner_join(verb_categories) %>%
+  inner_join(word_order) %>%
+  inner_join(clauses)
+
+# common_languages <- common_df %>%
+#   pull(language)
+```

### 2019-07-19 surprisal/WALS_features.Rmd
correlation plot

@@
 library(janitor)
-library(hexbin)
+library(directlabels)
 knitr::opts_chunk$set(echo = TRUE)
+
+theme_set(theme_classic(base_size = 18))
 ```
@@
 
-```{r cosines}
+```{r cosines, eval = F}
 unigram_cosines <- wiki %>%
@@
   as_data_frame(rownames = "language1") %>%
-  gather(language2, unigram_cosine, -language1) %>%
-  mutate(unigram_cosine = 1 - abs(unigram_cosine))
+  gather(language2, unigram_cosine, -language1)
 
@@
   as_data_frame(rownames = "language1") %>%
-  gather(language2, trigram_cosine, -language1) %>%
-  mutate(trigram_cosine = 1 - abs(trigram_cosine))
+  gather(language2, trigram_cosine, -language1)
 
@@
                        labels = c("Unigrams", "Trigrams")) 
+```
+
+```{r correlations}
+correlation_data <- read_feather(here("Data/feature_cosine_spaces.feather"))
+
+tidy_corr_data <- correlation_data %>%
+  gather(gram, angle, unigram_cosine, trigram_cosine) %>%
+  mutate(gram = if_else(gram == "unigram_cosine", "unigram", "trigram"))
+
+ggplot(tidy_corr_data, aes(x = feature, y = angle, fill = gram,
+                           color = gram, group = gram,
+                           label = gram)) + 
+  geom_smooth(method = "lm") +
+  geom_jitter(alpha = .05) +
+  geom_dl(method = "smart.grid") + 
+  theme(legend.position = "none")
+
+lmer(feature ~ distance * gram + (1|language1) + (1|language2),
+     data = tidy_corr_data) %>%
+  tidy()
+
 ```

### 2019-07-19 surprisal/MacWhinney poster/poster_plotting.Rmd
finishing touches

@@
 library(janitor)
+library(gridExtra)
 
@@
 
-sf$Examples <- c("Noun-relative clause; relative clause-noun", 
+sf$Examples <- c("Noun-RC; RC-noun", 
                  "No cases; 6 cases", 
                  "SOV; VSO", 
-                 "Only singular; singular and plural", 
+                 "Only singular; sing. and plural", 
                  "Affix; distinct word", 
@@
 ```
+
+```{r plotting trigram smoothing}
+wiki %>% 
+  mutate(Slope0 = runif(1, 0.9, 1.1)) %>% 
+  mutate(Slope1 = Slope1 + Slope0) %>% 
+  mutate(Slope2 = Slope2 + Slope1) %>% 
+  mutate(Slope3 = Slope3 + Slope2) %>% 
+  mutate(Slope4 = Slope4 + Slope3) %>% 
+  mutate(Slope5 = Slope5 + Slope4) %>% 
+  # rename(1 = Slope0, 2 = Slope1, 3 = Slope2, 4 = Slope3, 5 = Slope4, 6 = Slope5) %>% 
+  filter(gram == "Trigram") %>% 
+  select(-gram) %>% 
+  sample_n(6) %>% 
+  gather(position, information, -language) %>%
+  ggplot(aes(x = position, y = information, group = language, color = language))  
+    geom_point() + 
+    geom_line() + 
+    xlab("Position") + 
+    ylab("Average Information") + 
+    theme(axis.text.x = element_blank())
+```

### 2019-07-19 surprisal/WALS_features.Rmd
finishing touches

@@
 ```{r correlations}
-correlation_data <- read_feather(here("Data/feature_cosine_spaces.feather"))
+correlation_data <- read_feather(here("Data/feature_cosine_spaces.feather")) %>% 
+  mutate(unigram_cosine = unigram_cosine * 180) %>% 
+  mutate(trigram_cosine = trigram_cosine * 180)
 
@@
   geom_dl(method = "smart.grid") + 
-  theme(legend.position = "none")
+  theme(legend.position = "none") + 
+  xlab("Number of WALS features in common") + 
+  ylab("Information curve distance")
 
-lmer(feature ~ distance * gram + (1|language1) + (1|language2),
+library(lme4)
+lmer(feature ~ angle * gram + (1|language1) + (1|language2),
      data = tidy_corr_data) %>%

### 2019-07-29 surprisal/WALS_features.Rmd
geographic distance between languages

@@
 
+```{r geographic area comparison}
+languages <- read_csv(here("Data/languages.csv")) %>%
+  select(Latitude, Longitude, ISO639P3code)
+
+wiki <- read_csv(here("Data/relative_ngrams.csv"))
+
+labs <- read_csv(here("Data/wiki_wals.csv")) %>% 
+  group_by(language) %>% 
+  slice(1) %>% 
+  ungroup() %>% 
+  select(language, iso)
+
+latlong <- labs %>% 
+  left_join(imputed_df) %>% 
+  left_join(languages, by = c("iso" = "ISO639P3code")) %>%
+  group_by(language) %>%
+  slice(1) %>% 
+  ungroup() %>%
+  select(language, Latitude, Longitude) %>%
+  filter(complete.cases(.))
+
+sub_langs <- latlong %>% 
+  pull(language)
+
+unigram_cosines <- wiki %>%
+  filter(language %in% sub_langs, gram == "Unigram") %>%
+  select(-gram) %>% 
+  column_to_rownames("language") %>%
+  t() %>%
+  cosine() %>%
+  as_data_frame(rownames = "language1") %>%
+  gather(language2, unigram_cosine, -language1)
+
+wf_pairwise <- expand.grid(language1 = sub_langs,
+                           language2 = sub_langs) %>%
+  filter(language1 != language2)
+
+compute_sim <- function(language1, language2) {
+
+  wf_sub %>%
+    filter(language == language1) %>%
+    rename(value1 = value) %>% 
+    left_join(wf_sub %>% filter(language == language2) %>% select(feature, value), 
+              by = "feature") %>%
+    mutate(same = value == value1) %>%
+    ungroup() %>%
+    group_by(language) %>%
+    summarise(feature = sum(same)) %>%
+    mutate(language2 = language2) %>%
+    rename(language1 = language)
+}
+
+wf_sub <- imputed_df %>% 
+  filter(language %in% sub_langs) %>%
+  group_by(language) %>%
+  slice(1) %>%
+  gather(feature, value, -language) %>%
+  group_by(feature) %>%
+  mutate(value = as.numeric(as.factor(value)))
+
+feature_space <- map_dfr(1:nrow(wf_pairwise), ~compute_sim(wf_pairwise[.x,"language1"],
+                                          wf_pairwise[.x,"language2"]))
+
+compute_geo_sim <- function(language1, language2) {
+
+  latlong %>%
+    filter(language == language1) %>%
+    bind_cols(latlong %>% filter(language == language2)) %>%
+    mutate(geo = sqrt((Latitude1 - Latitude)^2 + (Latitude1 - Latitude)^2)) %>%
+    ungroup() %>%
+    rename(language2 = language1, language1 = language) %>% 
+    select(language1, language2, geo)
+}
+
+geo_space <- map_dfr(1:nrow(wf_pairwise), ~compute_geo_sim(wf_pairwise[.x,"language1"],
+                                          wf_pairwise[.x,"language2"]))
+
+```
+
 ```{r imputation}
@@
   MIMCA(ncp = 2, threshold = 1e-02)
-
-imputed_df110 <- as.data.frame(imputed_df[[1]][[99]]) 
 ```
@@
 
-wf_sub <- all_imputed %>% 
+wf_sub <- imputed_df %>% 
   group_by(language) %>%

### 2019-08-01 surprisal/Phonemes/phoneme_writeup.Rmd
tidied phones write-up

---
title: "Information within words in English, French and German speech"
author: Josef Klafka and Dan Yurovsky 
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: false
    number_sections: true
    theme: lumen
    toc_float: false
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r load-libraries, library, message=F, results='hide', warning=FALSE, include = F}
library(tidytext)
library(tidyverse)
library(childesr)
library(glue)
library(feather)
library(here)
library(entropy)
library(tidyboot)
library(janitor)
library(arrangements)
library(stringr)

knitr::opts_chunk$set(echo = TRUE, cache = T, warning = F, 
                      fig.width = 8, fig.height = 4)

theme_set(theme_classic(base_size = 16))
```

```{r read-data}
all_data <- read_feather(here("Data/all_data.feather"))
adult_phonemes <- read_feather(here("Data/adult_phonemes.feather"))
```

# Introduction

How do speakers tend to spread information among the sounds in English words? A salient feature of African American Vernacular English is the reduction of final consonants and consonant clusters. Why doesn't this reduction create communication problems, namely identifying and disambiguating words? One possibility is that the information important for identifying a word is not always uniformly distributed across all of its characters or sounds, but instead usually concentrated at the beginning or middle of the word. For these words, reducing final sounds might reduce the articulatory complexity of words without much cost for the listener in processing these words and inferring their intended meaning. By the time a listener hears only the first sounds of a word, they may already have identified which word they hear.

Through a series of exploratory analyses we tackle this problem using four approaches:

1. We compute the information of characters in English words from baserates of character production (unigram) and then conditioning on the prior charaters in each word (bigram and trigram). This analysis roughly captures the difficulty that a listener might have in predicting each character.
2. We repeat this analysis at the level of phonemes rather than characters, arguably a better proxy for speakers' and listeners' representations of spoken English. 
3. We compute hold-out entropy for individual phonemes in English words. This is roughly equivalent to asking "if you knew every other sound in a word, how much trouble would you have predicting what sounds appears in position $X$", where position $X$ is held out. 
4. We compare our English by-letter analysis to results for French and German, for which we  expect different information profiles. We expect that French words will have less information on average at the end of words compared to English, as the final letters in many French words are unpronounced. We expect German words will have more information on average in their final letters compared to English, as German nouns are case-marked with affixes.

Across all of our analyses, we find that the letters and sounds at the end of words in English have less information than the first letters and sounds in a word, and that the middle and final letters in English words are easily predictable from their preceding letters. We find that letters at the end of words in French and German also have little information compared to letters at the beginning of words in those languages, like in English. This is unexpected for German, and difficult to differentiate from English for French.

We use three corpora from the CHILDES-Talkbank system (MacWhinney, 2000) to investigate these questions for child language and child-directed speech. See the Appendix for more technical information about methods.

# English Letters

```{r plot-hist-helper}
plot_hists <- function(select_language = "English", select_rep = "letter") {

    all_data %>%
      filter(language == select_language, rep == select_rep) %>%
      distinct(word_length, n, rep) %>%
      ggplot(aes(x = word_length, y = n)) +
      geom_col(fill = "white", color = "black") +
      xlab("Word length") + 
      ylab("Frequency")
}
    # xlab(paste0(str_to_sentence(select_rep), " Length")) +

```

We used the Providence corpus from CHILDES as data for English (Evans & Demuth, 2012). In the plots below, each plot represents the information within letters at different positions in words, divided by words of different lengths. Word lengths are given by number of letters. Below is a histogram of the distribution of word lengths in the Providence corpus. We observed that most words ($84\%$) have $5$ letters or fewer, while $99\%$ of words in the corpus have fewer than $10$ letters. There are a tiny number of outliers which have $10$ or more letters--by manual inspection we see that these are mainly onomatopoeia.

```{r get cumsums for letter and phone frequencies, include = F, eval = F}
get_freq_sums <- function(select_language = "English", select_rep = "letter") {

  all_data %>% 
    filter(language == select_language, rep == select_rep) %>% 
    distinct(word_length, n, rep) %>% 
    mutate(share = n/sum(n)) %>% 
    mutate(cumsum = cumsum(share)) %>% 
    mutate(language = select_language)
}
```

```{r plot-english-letter-hist}
plot_hists()
```

We observed a distinctive shape for information at the level of individual letters in English. The first letter has relatively high information, the second letter has less information, the third letter has high information, and the final letter has less information (if the word has more than three letters).

```{r plot-curve-helper}
plot_curves <- function(select_language = "English", select_rep = "letter", 
                        select_gram = "Unigram") {
  all_data %>%
    filter(language == select_language, rep == select_rep,
           gram == select_gram) %>%
    filter(word_length %in% 2:6) %>%
    ggplot(aes(x = order, y = mean_s)) +
    facet_grid(. ~ word_length, 
               labeller = as_labeller(function(x) paste0("Length ", x))) +
    geom_point() +
    geom_line() +
    xlab(paste0(str_to_sentence(select_rep), " Position")) +
    ylab("Average information") + 
    ggtitle(paste0("Information in ", select_language, " ", 
                   select_rep, "s by ",select_rep, " position - ", select_gram))
}
```

```{r prov-letter-unigrams}
plot_curves()
```

Now we ask: if you know the previous letter, how much information is in the next letter in English? There is relatively less information in each letter knowing the previous letter, but we see the same general information trajectory across letters as in the individual letter trajectory.

```{r prov-letter-bigrams}
plot_curves(select_gram = "Bigram")
```

What about if you know the last two letters and are trying to predict the next letter? We see decreasing information for each letter within words as you move left-to-right. All letters in English words, aside from the first letter, are easily predicable from only knowing the preceding two letters for each. This is especially noticeable for shorter words with only three or four letters, but still holds true for words of length five or six letters.

```{r prov-letter-trigrams}
plot_curves(select_gram = "Trigram")
```

# English Phones

In order to analyze the individual phones in English, we split each word in the Providence corpus into its constituent phones, the sounds that make up the word. We then divide up the words in the corpus by their length in number of phones, similar to how we divided words by number of letters. For example, the word "food" has three phones: one for the "f", one for the long "oo", and one for the "d". We take all words like this that have three phones, and for each phone position we calculate the average information in that position.

As seen in the histogram below, almost all words in the Providence corpus ($92\%$) have four or fewer phones. $99.5\%$ of words have seven or fewer phones.

```{r plot-english-phone-hist}
plot_hists(select_rep = "phone")
```

Unlike in our letter analysis, there isn't as much of a consistent shape to the information distribution for phones in the individual (unigram) condition. There is some evidence that the first phone has relatively little information while the second phone has a lot of information, which is the opposite of what we observed consistently with the information in English letters.

```{r prov-phone-unigrams}
plot_curves(select_rep = "phone")
```

With bigrams and trigrams for English phones, the next phone is very predictable if you know the phone or two that came before it. Especially if you know the preceding two phones, then each phone past the first two is predictable and carries relatively little information in English. This is the same as our conclusion from the letter analysis in English. The middles and especially the ends of words in English contain relatively little information, because those part of words in speech or reading are easily predictable from what came before them.

```{r prov-phone-bigrams}
plot_curves(select_rep = "phone", select_gram = "Bigram")
```

```{r prov-phone-trigrams}
plot_curves(select_rep = "phone", select_gram = "Trigram")
```

# Hold-out entropy

For this next analysis, we took all words in the Providence corpus with between two and six phones. For each word length, we held out the phones in one phone position. For example, in all words with four phones, we held out the first phone. For each phone we held out, we computed the entropy (a measure of variation) in the observed words that had the remaining phones. Holding out the /f/ in "food" gives the continuations "food" and "mood" for example.

The higher the entropy when holding out phones in a given position, the more information is contained in that phone on average. We observed that word-initial phones tend to have more information in them than word-medial or word-final phones, so less is lost by dropping the final sounds in a word than in dropping the initial sounds.

```{r leave-out-entropy, include = F}
hold_out_entropy <- function(df) {
  spread_df <- df %>%
    spread(phone_order, phone) %>%
    clean_names()

  phone_length <- max(df$phone_order)

  lens <- paste0("x", 1:phone_length)

  boot_entropy <- function(group) {
    spread_df %>%
      group_by_at(group) %>%
      summarise(n = n()) %>%
      summarise(entropy = entropy(n)) %>%
      ungroup() %>%
      tidyboot_mean(entropy)
  }

  groupings <- combinations(lens, phone_length -1, 
                            layout = "list")

  groupings <- map(seq(length(lens),1,-1),
                   function(x) c(groupings[[x]], lens[x]))

  map_df(groupings, boot_entropy) %>%
    mutate(phone_order = seq(phone_length, 1, -1))

}
```

```{r show-leave-out-entropy}
split_lengths <- adult_phonemes %>%
  group_by(word_length) %>%
  nest() %>%
  filter(word_length > 1) %>%
  filter(word_length %in% 1:6 ) %>%
  mutate(entropy = map(data, hold_out_entropy))

split_lengths %>%
  select(-data) %>%
  unnest() %>%
  ggplot(aes(x = phone_order, y = empirical_stat)) + 
  facet_wrap(~ word_length, 
               labeller = as_labeller(function(x) paste0("Length ", x)),
               scales = "free") +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) + 
  theme_classic() + 
  geom_line() + 
  xlab("Phone position") + 
  ylab("Entropy")
```

# French and German

For our final analysis, we examined how the information within words in French and German compare to English. We used the French Palasis corpus (Palasis, $2009$) and German Wagner corpus (Wagner, $1985$) from CHILDES for our data. For many French words, the final letters are not pronounced although those letters are still written in the word. For example, "ils sauraient" meaning "they would know" is pronounced without the final "ent" in "sauraient". We predicted that the ends of French words will be less important than in English. German has a much richer nominal case system than English or French, and case markings in German are suffixes at the ends of nouns. We predicted that in German, the ends of words may be more important than in French or English.

## French letters

First, we examined the word length frequencies by number of letters in the French Palasis corpus. In the Palasis corpus, $99\%$ of words have $10$ letters or fewer. This French corpus has a higher share of words with six to ten letters than the English Providence corpus: $24\%$ versus $16\%$. However, almost all words in both corpora have $10$ or fewer letters. While there are relatively more medium length words in the French corpus compared to the English corpus, there are few long words in either corpus.

```{r plot-french-hist}
plot_hists(select_language = "French")
```

We observed that the final letters in French words, even considered without context, have little information. The gap between the information-rich first letter and final letters becomes more apparent with context, and we observe that the middle letters in French have relatively little information when considered with context, similar to English letters.

```{r plot french letter unigrams}
plot_curves(select_language = "French")
```

```{r plot french letter bigrams}
plot_curves(select_language = "French", select_gram = "Bigram")
```

```{r french-letter-trigrams}
plot_curves(select_language = "French", select_gram = "Trigram")
```

## German letters

In the German Wagner corpus, the distribution of word lengths has a thicker tail than in English. $99\%$ of words have $13$ or fewer letters, compared to the Palasis and Providence corpora where over $99\%$ of words had $10$ or fewer letters. However, $97.4\%$ of German words have $10$ or fewer letters, so the corpus is similarly comprised of short and medium-length words.

```{r plot-german-hist}
plot_hists(select_language = "German")
```

We did not observe that German had more information in final letters than English or French, even when examining the information in independent letters. The final letters of German words hold little information compared to the first letter of German words, which is even more noticeable with two letters of predictive context for each letter.

```{r plot german letter unigrams}
plot_curves(select_language = "German")
```

```{r plot german letter bigrams}
plot_curves(select_language = "German", select_gram = "Bigram")
```

```{r plot german letter trigrams}
plot_curves(select_language = "German", select_gram = "Trigram")
```

# General Discussion

Across all three languages we examined and both letters and phonemes in English, we found that the first two or three letters and sounds in a word contain relatively more information than any other letters or sounds in the word. The final letters and sounds tended to contain little information compared to the initial letters or sounds. The middle and final letters in words in all three languages were easily predictable from the two letters that came directly before them. Even for a language like German which marks case at the end of nouns, young listeners can identify and disambiguate the words using only the first few letters of those words.

# Additional material
## Appendix A: Methods
The metric we use for information is based on the lexical surprisal metric from Levy (2008), with formula given below. The surprisal $s$ of the $i$th word $w_i$ in an utterance is given by

$$s(w_i) = -\log P(w_i|w_{i-1}w_{i-2}...)$$

where $w_{i-1}$ denotes the word one previous to the target word. For our work, the surprisal $s$ of the $i$th letter or sound $\ell_i$ in a word is given by

$$s(\ell_i) = -\log P(\ell_i|\ell_{i-1}\ell_{i-2}...)$$

where $\ell_{i-1}$ denotes the letter or sound one previous to the target letter or sound. For our unigram work, we did not use context, and so the conditional probability is instead a simple probability.

# References
Evans, K. E., & Demuth, K. (2012). Individual differences in pronoun reversal: Evidence from two longitudinal case studies. *Journal of child language*, 39(1), 162-191.

Levy, R. (2008). Expectation-based syntactic comprehension. *Cognition*, 106(3), 1126-1177.

MacWhinney, B. (2000). *The CHILDES project: The database* (Vol. 2). Psychology Press.

Palasis, K. (2009). *Syntaxe générative et acquisition: le sujet dans le développement du système linguistique du jeune enfant* (Doctoral dissertation, Nice).

Wagner, K. R. (1985). How much do children say in a day? *Journal of Child Language*, 12, 475–487.

### 2019-08-01 surprisal/Phonemes/phoneme_writeup.Rmd
minor phoneme edits

@@
 This analysis roughly captures the difficulty that a listener might have in predicting each character.
-2. We repeat this analysis at the level of phonemes rather than characters, arguably a better proxy for speakers' and listeners' representations of spoken English. 
+2. We repeat this analysis at the level of phones rather than characters, arguably a better proxy for speakers' and listeners' representations of spoken English. 
 3.
-We compute hold-out entropy for individual phonemes in English words.
+We compute hold-out entropy for individual phones in English words.
 This is roughly equivalent to asking "if you knew every other sound in a word, how much trouble would you have predicting what sounds appears in position $X$", where position $X$ is held out.
@@
 
-Across all three languages we examined and both letters and phonemes in English, we found that the first two or three letters and sounds in a word contain relatively more information than any other letters or sounds in the word.
+Across all three languages we examined and both letters and phones in English, we found that the first two or three letters and sounds in a word contain relatively more information than any other letters or sounds in the word.
 The final letters and sounds tended to contain little information compared to the initial letters or sounds.

### 2019-08-09 surprisal/linguistic_distance.Rmd
mean levenshtein distance correlated with information curve distance

---
title: "Linguistic distance"
author: "Josef Klafka and Dan Yurovsky"
date: "8/9/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(stringdist)
library(lsa)
library(feather)

knitr::opts_chunk$set(echo = TRUE)
```

```{r read in data}
wals <- read_csv(here("Data/imputed_all.csv"))
wiki <- read_csv(here("Data/relative_ngrams.csv"))
iso_to_language <- read_csv(here("Data/wiki_wals.csv")) %>%
  select(language, iso) %>%
  group_by(language) %>%
  slice(1) %>% 
  ungroup()

asjp_forms <- read_csv(here("asjp_dataset/forms.csv"))
asjp_langs <- read_csv(here("asjp_dataset/languages.csv"))
lang_forms <- asjp_forms %>% 
  select(Language_ID, Parameter_ID, Form) %>% 
  left_join(asjp_langs %>% select(ID, ISO639P3code), by = c("Language_ID" = "ID")) %>%
  left_join(iso_to_language, by = c("ISO639P3code" = "iso")) %>%
  filter(complete.cases(.)) %>% 
  select(language, Parameter_ID, Form) %>%
  group_by(language, Parameter_ID) %>% ## taking off one at random for each language - fix
  slice(1) %>%
  ungroup()

langs <- wals %>% pull(language)
```

```{r make swadesh space}
langs_pairwise <- expand.grid(language1 = langs,
                           language2 = langs) %>%
  filter(language1 != language2)

asjp_distance <- function(language1, language2) {
  lang_forms %>% 
    filter(language == language1) %>%
    left_join(lang_forms %>% filter(language == language2), by = "Parameter_ID") %>% 
    filter(complete.cases(.)) %>% 
    mutate(ldn = stringdist(Form.x, Form.y) / max(str_length(c(Form.x, Form.y)))) %>%
    group_by(language.x, language.y) %>% 
    summarise(ldn = mean(ldn)) %>% 
    rename(language1 = language.x, language2 = language.y) %>% 
    ungroup()
}

swadesh_space <- map_dfr(1:nrow(langs_pairwise), ~asjp_distance(langs_pairwise[.x,"language1"],
                                          langs_pairwise[.x,"language2"])) 
```

```{r compare swadesh and information curve spaces}
all_distances <- read_feather(here("Data/all_spaces.feather"))

all_distances %>% 
  ggplot(aes(x = ldn)) +
    geom_smooth(aes(y = unigram_cosine, color = "#0033CC"), 
                method = 'lm', se = F) + 
    # geom_smooth(aes(y = bigram_cosine, color = "#6699FF"), 
    #             method = 'lm', se = F) +
    geom_smooth(aes(y = trigram_cosine, color = "#33CCFF"), 
                method = 'lm', se = F) +
    geom_hex(aes(y = unigram_cosine, color = "#0033CC", alpha=.2)) + 
    # geom_point(aes(y = bigram_cosine, color = "#6699FF")) + 
    geom_hex(aes(y = trigram_cosine, color = "#33CCFF", alpha=.2)) + 
    ylab("Information Curve Cosine") + 
    xlab("Average Swadesh Distance") + 
    scale_color_manual(name = "Gram", 
                       values = c("#0033CC", "#33CCFF"), 
                       labels = c("Unigrams", "Trigrams")) 
```

## Making the data

```{r pairing up spaces}
unigram_cosines <- wiki %>%
  filter(language %in% langs, gram == "Unigram") %>%
  select(-gram) %>% 
  column_to_rownames("language") %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, unigram_cosine, -language1)

trigram_cosines <- wiki %>%
  filter(language %in% langs, gram == "Trigram") %>%
  select(-gram) %>% 
  column_to_rownames("language") %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, trigram_cosine, -language1)

feature_space <- read_feather(here("Data/feature_cosine_spaces.feather")) %>% 
  select(language1, language2, feature)

all_distances <- swadesh_space %>%
  left_join(unigram_cosines) %>%
  left_join(trigram_cosines) %>% 
  left_join(feature_space)

all_distances %>% write_feather(here("Data/all_spaces.feather"))
```

### 2019-08-09 surprisal/Phonemes/Phones.Rmd
phones writeup updated

--- 2019-08-01 surprisal/Phonemes/phoneme_writeup.Rmd
@@
 ---
-title: "Information within words in English, French and German speech"
+title: "Information in sounds and letters in English, French and German speech"
 author: Josef Klafka and Dan Yurovsky 
@@
 
-Through a series of exploratory analyses we tackle this problem using four approaches:
+We tackle this problem using four exploratory analyses:
 
 1.
-We compute the information of characters in English words from baserates of character production (unigram) and then conditioning on the prior charaters in each word (bigram and trigram).
+We compute the information of characters in English words from baserates of character production (unigram) and then conditioning on the prior characters in each word (bigram and trigram).
 This analysis roughly captures the difficulty that a listener might have in predicting each character.
@@
 We compare our English by-letter analysis to results for French and German, for which we  expect different information profiles.
-We expect that French words will have less information on average at the end of words compared to English, as the final letters in many French words are unpronounced.
+We expect that French words will have less information on average at the end of words compared to English, as French speakers do not pronounce the finals sounds in many words in their language.
 We expect German words will have more information on average in their final letters compared to English, as German nouns are case-marked with affixes.
@@
 We find that letters at the end of words in French and German also have little information compared to letters at the beginning of words in those languages, like in English.
-This is unexpected for German, and difficult to differentiate from English for French.
+This is unexpected for German, though we do find that the ends of words in French contain less information than the ends of words in English, even looking solely at character baserates.
 
-We use three corpora from the CHILDES-Talkbank system (MacWhinney, 2000) to investigate these questions for child language and child-directed speech.
-See the Appendix for more technical information about methods.
+We use three corpora from the CHILDES-Talkbank system (MacWhinney, 2000). See the Appendix for more technical information about methods.
 
@@
 The middle and final letters in words in all three languages were easily predictable from the two letters that came directly before them.
-Even for a language like German which marks case at the end of nouns, young listeners can identify and disambiguate the words using only the first few letters of those words.
+Even for a language like German which marks case at the end of nouns, listeners can identify and disambiguate the words using only the first few letters of those words.
 
-# Additional material
-## Appendix A: Methods
+# Appendices
+## A: Methods
 The metric we use for information is based on the lexical surprisal metric from Levy (2008), with formula given below. The surprisal $s$ of the $i$th word $w_i$ in an utterance is given by

### 2019-08-30 surprisal/Phonemes/Phones.Rmd
reverse and shuffled prov surprisals

@@
 date: "`r Sys.Date()`"
-output: 
-  html_document:
-    toc: false
-    number_sections: true
-    theme: lumen
-    toc_float: false
-    code_folding: hide
+output: pdf_document
+
 editor_options: 
@@
 
-```{r read-data}
+```{r read-data, include=F}
 all_data <- read_feather(here("Data/all_data.feather"))
@@
 
-```{r plot-hist-helper}
+```{r plot-hist-helper, include=F}
 plot_hists <- function(select_language = "English", select_rep = "letter") {
@@
 
-```{r plot-curve-helper}
+```{r plot-curve-helper, include=F}
 plot_curves <- function(select_language = "English", select_rep = "letter", 

### 2019-08-30 surprisal/reverse_shuffled_surprisals.Rmd
reverse and shuffled prov surprisals

---
title: "Reverse and shuffled lexical surprisal"
author: "Josef Klafka and Dan Yurovsky"
date: "8/30/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(here)
library(feather)
library(childesr)

knitr::opts_chunk$set(echo = TRUE)
```

```{r load corpora}
prov_utterances <- get_utterances(corpus = "Providence")
```

```{r get reverse}
adult_utterances <- prov_utterances %>%
  filter(speaker_role != "Target_Child") %>% 
  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n()) %>% 
  arrange(desc(word_order)) %>% 
  mutate(word_order = 1:n()) %>%
  ungroup()
```

```{r get shuffled}
adult_utterances <- prov_utterances %>%
  filter(speaker_role != "Target_Child") %>% 
  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  sample_frac(size = 1, replace = F) %>% 
  mutate(word_order = 1:n()) %>%
  ungroup() 
```

```{r}
adult_unigrams <- adult_utterances %>%
  group_by(word) %>%
  count() %>%
  ungroup() %>%
  mutate(p = n / sum(n))

unigram_surprisals <- adult_utterances %>%
  left_join(adult_unigrams) %>%
  mutate(s = -log(p)) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s))

adult_bigrams <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  left_join(adult_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = n.x/n.y)

bigram_surprisals_prep <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  left_join(select(adult_bigrams, -p)) %>%
  left_join(adult_unigrams)

bigram_surprisals <- bigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)

adult_trigrams <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  group_by(lag_word2, lag_word1, word) %>%
  count() %>%
  filter(!is.na(lag_word1), !is.na(lag_word2)) %>%
  ungroup() %>%
  left_join(adult_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
  mutate(tri_cond_p = n / n.x) %>%
  select(-n, -n.x, -n.y, -p, -cond_p)

trigram_surprisals_prep <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  left_join(adult_trigrams) %>% 
  left_join(select(adult_bigrams, lag_word, word, cond_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
  left_join(adult_unigrams)

trigram_surprisals <- trigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
                    ifelse(is.na(lag_word1), # check if not the second word
                           -log(p), # trigram if third or beyond
                           -log(cond_p)), # bigram if second word
                    -log(tri_cond_p))) %>% # unigram if first word
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)
```

```{r plotting}
trigram_surprisals %>%  
  filter(length %in% c(5, 7, 9)) %>%
  ggplot(aes(x = word_order, y = s)) +
    facet_wrap(~ length) + 
    xlab("Word position") +
    ylab("Mean surprisal") +
    ggtitle("Shuffled trigram surprisals") + 
    scale_x_continuous(breaks = 1:9) + 
    geom_line(size = 1) +
    theme_classic(base_size = 10) + 
    theme(legend.position = "nones") 
```

### 2019-09-05 Paper/paper.Rmd
paper rmd created

---
title: "Speakers of diverse languages structure their utterances for efficient communication"
author: "Josef Klafka and Daniel Yurovsky"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    number_sections: true
header-includes:
    - \usepackage{setspace}\doublespacing
indent: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggdendro)

knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Language can be thought of as a code: speakers encode their thoughts into a sequence of linguistic signals, namely an utterance, which listeners then must decode in order to recover the intended message. How do speakers structure their utterances so that communication is successful? Using tools from information theory and sentence processing, we find that speakers must structure the information in their utterances according to the grammatical, morphological and lexical constraints of their language, which vary greatly from language to language. This information structure characterizes language from children’s first utterances to large-scale knowledge base entries. Once listeners have access to linguistic context within an utterance, then they can decode the rest of that utterance at a constant and optimal rate.

# Introduction

We use language to communicate information. Although we often need to say things we have never said before, we are easily and efficiently able to produce novel utterances with minimal prior planning. Our listeners incrementally process our utterances, and predict which words will come next within fractions of a second of hearing the beginning of the word (for evidence see e.g. Kutas & Federmeier, 2010). Encoding new and unpredictable information into our utterances surprises our listeners, and presents problems for communication. Listeners struggle to decode unpredictable words and may have difficulty decoding the following words as well. Efficient and effective communication relies on interlocutors being able to process novel information in without being surprised by new information in unpredictable words. On the other hand, to make best use of the capacity of the communication channel, speakers should encode as much information as possible into each word they say.

Due to their differences in distributional frequency, words vary in how much information-theoretic content they contain (Shannon, 1948). This information content plays an important role in determining the shape of words and speech. Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be as speakers minimize effort in the amount of information they encode in speech (Zipf, 1936). Over and above frequency, contextual information content predicts word length (Piantadosi, Tily & Gibson, 2011). Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side (Levy & Jaeger, 2007). Across utterances, context allows speakers to produce more informative and surprising sentences later in communication. The later a sentence occurs in a paragraph, the the informative it is and the more unusual and surprising words it contains (Genzel & Charniak, 2002). From the level of individual words to sentences in paragraphs, information content shapes and is shaped by the context it appears in.

However, a gap exists in the literature for examining how frequency and context determine the information content in utterances. We do not speak only in phrases or paragraphs, but conversationally we communicate in utterances, either on their own or banded together. Writers use sentences to communicate thoughts, often on their own instead in paragraphs with other sentences. Cross-linguistically, the average rate of information speakers transfer in a syllable differs (Aylett & Turk, 2004). Languages possess very different grammatical and morphological structures--the same thought can be encoded in at least as many ways as there are different languages. How do the different structures seen in different languages, with different rates of information encoding, affect how speakers structure information in production?

Yu et al. (2016) proposed a novel method for quantifying the amount of information typically found at each part of English utterances. The Yu et al work examined written sentences from the British National Corpus. Treating each word position by itself, Yu et al. calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty (Shannon, 1948). Yu et al. found a distinctive three-step shape to the distribution of information in the British National Corpus.

In this paper, we expand on this body of prior work in a number of novel ways. We replicate Yu et al’s results with a different metric, one intimately tied to incremental word processing. We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech. We expand the study of information density to the largest set of languages yet considered, and incorporate contextual and typological information. We find that each language has a characteristic, robust distribution of information predictable from its morphology, phonology and syntax, and tied to the language’s genealogy. Speakers of a language will tend to distribute information according to that language’s distribution of information in their utterances. As soon as a child can form multi-word utterances in a given language, they will produce utterances obeying the characteristic distribution of information in a given language. We also apply a contextual version of the same metric, and find that in communication, listeners are able to decode information from the utterances they hear at a constant rate, regardless of language, past the first couple of words.

# Methods

Shannon (1948) defined information as “the reduction in uncertainty about one variable given that the value of another variable is known”. We use a metric proposed by Shannon and applied by Levy (2008): lexical surprisal, which defines the information in word based on the ratio of possible continuations of the sentence after to before the word is seen. Equivalently, we can compute surprisal with the predictability of the word, as in the formula below. The surprisal of a word is inversely proportional to the predictability of a word, such that less common and less predictable words carry more information.

$$surprisal(word) = -log P(word)$$

The surprisal of a word is also correlated with the processing cost of a word, with evidence from eye-tracking (Smith & Levy, 2013) and ERP (Frank et al, 2014) studies, among other sources. Considered without context, the surprisal of an individual word is inversely proportional to the frequency of that word, so that the less often a person has seen a word, the more information that word holds. For example, “flower” has less information than “azalea” because “flower” is much more common than “azalea”. Though the two words have the same length in number of letters, it’s more difficult to process “azalea” when reading it here than when reading “flower”. Frequency is intimately tied information content in words, with much of the differences between words frequencies being explained by information content cross-linguistically (Piantadosi et al, 2011).

However, when reading or listening, people don’t just consider each word as an isolated linguistic signal. Instead, listeners use the words they have already heard to predict and decode the word they are currently processing. Following this incremental processing paradigm, we can also condition the surprisal of a word in its context. In the formula below, widenotes the word currently being read or heard, while wi-1denotes the first word before the current word, wi-2denotes the second word before the current word, and so on.

$$surprisal(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...) = -log P(w_i,w_{i-1}w_{i-2},...)P(w_{i-1}w_{i-2}...)$$

When we use a word or two of context in our surprisal calculations, then the set of reasonable final items in our ngrams is greatly restricted. “Flower” may contain less information than “azalea” when we consider the words independently of their context, but with context this can be reversed. Flower appears in a variety of contexts, and so the information content of a word like “flower” in a particular context may be higher than “azalea”. If you only have azaleas in your garden, then hearing someone say “in that garden, look at the flowers” may be higher surprisal for you: you expect them to say “azalea”. This prediction does not need many words before to work out, for example in “I take my coffee with cream and sugar”. When you hear “cream and”, you automatically predict “sugar” or maybe “honey”, but there are few possible continuations with even those two words. Hearing “I” restricts the next word to a verb, or possibly an adverb, and since you have just heard the speaker refer to themselves in the first person singular, your set of possible completions is significantly restricted.

To find the distribution of information within a corpus, we calculate the surprisal or contextual surprisal for each unique word or sequence of words within the corpus. We take all utterances of a given length, and for each word position in utterances of that length, we compute the average of the surprisals for all of the non-unique words that occur in that position, conditioned or not conditioned on context. By computing these averages for each word position in an utterance, we compute a low-dimensional approximation to the average distribution of information in the corpus. With the surprisal metric, we base the information contained in each word on how often the word is encountered in its context in the corpus. As long as the corpus is representative of the language or population we study, then the distribution of information is approximated for that language or population as a whole.

The frequency-based surprisal metric gives us an idea of when in their utterances speakers say frequent i.e. independently information-rich words. The context-based surprisal metric show us how speakers tend to distribute the information in utterances relative to real-time processing in communication. We expect a priori that our frequency-based surprisal curve will be flat. No one part of the sentence will on average have words that are more frequent than another across utterance lengths. Similarly, we expect that there will be a small smoothing effect for our contextual surprisal metric.

# Written and Spoken English have the same information distribution

We first turn to working with written English in the British National Corpus (BNC). The BNC is a collection of spoken and written records (90% written) from the turn of the century, intended to be a representative sample of British English. Using their word entropy metric without context, Yu et al. found a distinctive three-step distribution for information in written English sentences in the corpus. The first word tended to contain little information. While the middle words of sentences each had more information than the first word, information was flat and non-increasing across the middle of sentences. The final word, they found, contained the most information out of any in the sentence, with a noticeable spike in information. They found the same distribution across sentence lengths, from sentences with 15 words to sentences with 45 words.

We first replicate the Yu et al result using the surprisal metric in place of the entropy metric. We use the frequency-based or “contextless” surprisal metric, which determines the average distribution of information based on word frequencies in a corpus. A priori we expect that the frequency-based metric will produce a flat distribution of information across word positions in the BNC.

See in the figure below a qualitative comparison of the information trajectory they found and the information trajectory we found. It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as Yu et al did. The first words in written English sentences tend to contain little information, while the final words tend to contain a great deal of information.

Comparison between the unigram entropy and surprisal BNC results plot

However, this only characterizes written English sentences. If this information distribution is characteristic more generally of English, then we will find the same distribution for spoken English. We use the Switchboard corpus of spoken English telephone conversations between adults to compute the distribution of information across spoken utterances in English. The sentences in the BNC, compared to the conversations in Switchboard, feature much longer sentences with larger vocabularies and more complex sentence structures. We split the conversations in Switchboard not by sentences, as in the BNC, but by turns.

For Switchboard we find the same information trajectory as we did for the BNC, and we find the same distribution of information regardless of utterance length. The longer the utterance, the more pronounced the information distribution shape becomes. A qualitative comparison of the two trajectories is below.

We have found a unique average distribution of information that appears to characterize the English language as a whole, regardless of medium (spoken or written) or utterance length. This distribution indicates that in English, the words we speak or write at the beginnings of utterances has little information, while the words we speak or write at the ends of utterances has a lot of information. The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from Genzel & Charniak (2002).

What about context? So far we’ve only discussed the unigram (“contextless”) surprisal metric, considering words on their own without any explicit incorporation of prior context. As previously discussed, listeners decode information and process what they hear incrementally, using prior heard words to ease the comprehension process. We now include one word of context (bigrams) and then two words of context (trigrams). We observe a flattening effect of context across both modalities and all speaker populations. After the first word or two, where the listener does not have access to prior context, then they decode information at a flat and more or less uniform rate.

BNC and Switchboard plot with context

Speakers produce information at a more or less constant rate, avoiding peaks or troughs in their information distribution, except at the beginnings of utterances, where listeners may not have any context to predict what the speaker is going to say. Speakers of English tend towards a characteristic and uneven distribution of word information based on frequencies within their utterances. Their interlocutors, however, once they have a word or two of context, decode information at a more or less constant and optimal rate.

# English CDS and child speech

We found that English speakers and writers used the same robust and distinctive distribution of information within each utterance, regardless of the number of words in their utterances. To determine if this distribution truly characterizes all speakers of the English language as a whole, we wanted to examine speech from English-speaking children who are producing their very first multi-word utterances. We hypothesize the three-step distribution of information we found for English will characterize child speech and child-directed speech.

We use the Providence corpus from CHILDES (MacWhinney, 2000; Demuth, 2012), which consists of about 650000 utterances from parent-child conversations for six monolingual English-speaking children talking with the parents. The child age range was between 11 months and 4 years. Sessions recorded in the home. Parents said a majority of the utterances, but children take up an increasing share of the conversation as they grow older and are actually able to speak full utterances. The utterances in the Providence corpus are on average significantly shorter than those in the BNC and Switchboard; over 90% of the utterances in the Providence corpus are at most 10 words long, with the median at around 6 words. Similar to our treatment of the Switchboard corpus, we split the Providence corpus up by conversational turns and separate utterances.

We observe the same distinctive distribution of information for parents and children in the Providence corpus as we did for adults in the BNC and Switchboard. The distribution of information we found at the level of individual words in English, therefore, characterizes the English language as a whole and not only adult utterances, not only written utterances.

Plot of Providence unigram curves

What about context in the Providence corpus? When incorporating one or two words of predictive context, we observe the same trend as in the Switchboard and BNC corpora. Beyond the first couple of words, once your interlocutor has enough context to predict with some accuracy what you will say next, then you decode information from their speech stream at a constant and optimal rate. This applies to parents and children speaking to one another, as well as adults speaking and writing to one another.

Plot of Providence bigram and trigram curves

# Child speech and CDS across languages

So far, we have only looked at the distribution of information in words in English, both with and without context. We have examined child speech, child-directed speech and adult-directed speech, as well as writing samples selected to be representative of British English as a whole. But this only captures the picture for English.

We now turn to a small number of typologically diverse languages, and conduct the same analysis, using a monolingual adult-child speech corpora from CHILDES (MacWhinney, 2000) to compare the results from these languages directly to our results from the English Providence corpus. We use corpora for Spanish, German, French, Mandarin and Japanese. Similarly to the English Providence corpus, all of the following corpora consist mainly of shorter utterances: most utterances in the corpora are under 10 words long. For Spanish, we use the corpus of conversations between primary school-aged children between 6 and 9 years old and the researchers, in which the children were asked to recite stories in their native Spanish (Shiro, 1997). The students came from working class public schools and upper class private schools. For German, we use the Wagner corpus (Wagner, 1985). This corpus consists of a collection of mini-corpora collected from socio-economically diverse families, with children aged from 1;0 to 14;0, mainly on the younger end of that range. The corpus mainly consists of parent-child conversations between the children and their primary caregivers. For French, we use the Palasis corpus (Palasis, 2009), a longitudinal study of children between the ages of 2;5 and 4;0 from the same kindergarten class. The conversations in this corpus take place between the children and other children in the same class, and between the children and the investigator. For Mandarin Chinese, we use the Zhou corpus of dinner conversations from the Shanghai area (Li & Zhou, 2015). The children in this corpus are 5 or 6 years old, and half come from working class families (the other half from upper class families). For Japanese, we use the Okayama corpus (Okayama, 1973). This corpus consists of conversations between children between the ages of 2;2 and 4;2 and their mothers.

We observe a distinct and characteristic frequency-based information trajectory emerge for each language, robust across each utterance length within each language. We see the same distribution of information for both parents and children in each language, with the child’s curve normally below the parent’s curve, likely due to the parent speaking a larger share of the utterances in the corpus and using a larger vocabulary than the child. We see the opposite amplitude trend in the Shiro corpus, where the children speak more than the adult investigators as the children are the ones telling the stories. We include the frequency-based information curve from the Providence corpus for comparison.

Plots of CHILDES frequency-based information curves

English, Spanish, French and German feature similar information curves shapes, with slight variations. The German information curve features lower information for longer towards the beginnings of utterances, possibly due to the grammatical restriction that the second word in German utterances must be a verb (V2). Spanish features a larger spike in the amount of information in the final words of utterances. For Japanese and Mandarin, we observe completely different unigram information curve trajectories. The Japanese unigram information curve trajectory begins high and finishes low, the mirror image of the European language information curves. The Mandarin curve begins low and finishes low, but features high information in the middle of utterances. We hypothesize this may be due to both Japanese and Mandarin speakers typically ending their utterances with particles, which contain very little information on their own. The penultimate noun with the speakers pair the particles are where the information lies.

For the bigram and trigram information curves, we see the same contextual smoothing effect as for all the corpora we worked with in English. While the unigram information curves may depend based on the language, the contextual information curves show the same trajectory cross-linguistically. Using ngrams with n higher than 3 is difficult for parent-child speech corpora because the utterances are so short on average (less than 10 words). We hypothesize that the unigram information curves may vary based on the genealogy and typology of the languages in question, but this does not extend to the trigram information curves in particular. We shall show evidence that this is true in the rest of the paper.

Plots of CHILDES trigram (and bigram) information curves

# Language structure and large-scale data analysis: methods and preprocessing

To make a claim about language as a whole and languages on a larger scale, we needed to use larger corpora and a much larger number of languages. We pulled corpora for 159 diverse languages from Wikipedia [Citation], each of which had at least 10000 articles. We split each article up into sentences; the variance in sentence lengths for Wikipedia was significantly larger than for the CHILDES corpora we used in the previous section. Most sentences in Wikipedia contained between 10 and 30 words, unlike the CHILDES corpora which mainly contained utterances with under 10 words. We excluded the small fraction of utterances with more than 50 words as uncharacteristic of English speech.

How to analyze more than 40 different surprisal curves for each language, adding up to several thousand surprisal curves total? We used two different strategies, which yielded identical results upon analysis. Each strategy gave us a five-dimensional vector space embedding for each language in a Wikipedia “slope space”. For the first strategy, we split each sentence length by number of words into fifths, and computed surprisal values for the closest word position to each quintile. We then computed the slopes between the surprisal values at neighboring quintiles, yielding five slope values for each curve. For the second strategy, we split each sentence length by number of words into sections based on those areas of the surprisal curves that had seemed most important before: between the first and second word; between the second and third word; between the third word and third-to-last word; between the third-to-last word and the second to last word; and between the second-to-last word and the last word. We then similarly computed surprisal values at each of these positions, and similarly computed slopes between the surprisal values at each position, giving us another five slope values for each language summarizing the surprisal curves. Illustrations of these two strategies are below.

Illustration of the regions we target with each of the two slope treatments below.

We computed unigram, bigram and trigram values for each language. All of this data, along with the scripts used for the analysis, are available on GitHub. We ran this code on the RCC at the University of Chicago; it took X amount of time and used up Y amount of carbon to run.

To more rigorously described the differences between languages, we used data from the World Atlas of Language Structures (WALS; Dryer & Haspelmath eds., 2013). WALS has data for 144 typological features in 2569 languages around the world. These features describe aspects of morphology, syntax, phonology, etymology and semantics--in short the features describe the structures in each language. As WALS is a compiled database from dozens of papers from different authors, most of the features and languages are fairly sparse. Even limiting ourselves to the 159 language corpora we pulled from Wikipedia and 122 features from WALS, there are nearly 20000 individual possible feature values, fewer than half of which were already inputted for those languages in the WALS database.

We used the Multiple Imputation Multiple Correspondence Analysis algorithm (MIMCA; Citation) to fill in the missing data using statistical imputation. MIMCA essentially uses mean imputation to begin with the missing values, then performs repeatedly performs principle components analysis on and reconstructs the contingency table formed from observations in categorical variables. By the end of this we had unigram, bigram and trigram surprisal curves for the 159 language corpora pulled from Wikipedia, along with 122 typological features for each language.

However, the WALS features describe specific structural differences between languages, while our surprisal metric is lexical. To target lexical differences between languages, we computed the average normalized Levenshtein distance (LDN; Holman et al, 2008) over the 40 item Swadesh list (1955), retrieved from the ASJP database (Wichmann, Holman & Brown, 2018). The Swadesh list is designed to include near-universal words that target basic cognitive concepts, and are useful in determining the genealogical similarities and differences between languages. The results of classifying languages using the Swadesh list and LDN are correlated with those using WALS features, but the Swadesh list and LDN do not suffer from the same sparsity problem as WALS (Holman et al, 2008).

# Language structures and large-scale data analysis: results

We ran a hierarchical clustering algorithm on the unigram information curves using the hclust package in R (hclust). We used the complete linkage algorithm for hierarchical clustering with distances between information curves in language computed using cosine distance between their embeddings in the slope space. The complete linkage algorithm at every step pairs each language or cluster of languages with its closest neighboring language or cluster. A sample from the dendrogram is shown below, and the full tree can be constructed using the code in our GitHub repository. The language family-like structure can be seen at first glance, although the dendrogram does not exactly replicate language genealogy for all 159 languages. This suggests using a first-pass quantitative method that the information curves do correspond in some measure to language families, but language families do not explain the entire story.

Dendro goes here

A sample of the trigram information curves are plotted below, and all trigram information curves for the languages we used follow the same pattern. The first few words in utterances for each language are surprising, but with even a couple words of predictive context for each word, the amount of information in each word drops off and speakers produce information at a constant, optimal rate.

Plot of some trigram surprisal curves goes here

First we examined the effects of individual typological features on the shapes of the unigram information curves. We ran logistic regressions using the lme4 package in R (lme4), checking whether the cosine distance between two languages’ embeddings in the slope space played a role in determining if they had the same value for a given WALS feature. Individual WALS features do not necessarily have ordinal values. Some, such as the “Number of Cases” feature, are easy to quantify and order. Others are more difficult. For example, how does one order “relative clauses appear after the nouns they modify”, “relative clauses appear before the nouns they modify” and “free order of relative clauses and nouns”? We chose the identify relation to avoid deciding on the basis of individual features. We found that 100 out of the 120 features from WALS we examined was statistically significant (p < .001) in determining whether two languages had the same shape to their unigram information distributions. The results for some important features are below, and the rest of the results are in the Appendices.

Some lr results go here.

We next compared how the cosine distance between two languages related to how many WALS features they had in common. There is a small correlation with a very small R^2 value between WALS features similarity and cosine similarity between two languages, but a statistically significant slope (p < .001). Essentially, for every feature two languages have in common, they have one degree more cosine similarity in the slope space. As feature values tend to be highly correlated within categories (e.g. the word order features are highly correlated), then this translates into language having much more similar information curves if they have more WALS features in common, an intuitive result.

Messy plot of WALS features vs. cosine similarity goes here.

So the more similar two languages are in terms of typological features, the more similar their unigram information curves tend to be. For lexical features, we see a stronger correlation between the similarity of two languages in terms of their average LDN and the cosine distance between their information curves.

Messy plot of average LDN vs. cosine similarity goes here

The shape of a language’s frequency-based information curve covaries with its typological and lexical similarity to other languages.

# General Discussion

The language we speak impacts how we encode information into our utterances, and how listeners decode our speech signals. However, despite the variety in communicative codes, all languages are characterized by predictability and a lack of suprisal in regular speech. In this paper, we developed a novel computational modeling method for examining the average information at a fine-grained level in utterances of different lengths in a variety of corpora and across a large number of languages. With only a couple of words of context for each word, cross-linguistic differences disappear and all languages are characterizes by low surprisal (low processing cost and information).

The Uniform Information Density Hypothesis (Levy & Jaeger, 2007) provides an optimal strategy for distributing information among the words in an utterance. The hypothesis suggests that speakers tend to distribute information evenly in their utterances, close to the capacity of the communication channel. This strategy avoids peaks in information which lead to issues for the decoder, while making the best use of the communication channel possible and aiming to produce as much information as the listener can decode at every moment.

### 2019-09-06 Paper/paper.Rmd
paper citations

@@
 output:
-  pdf_document:
+  bookdown::pdf_document2:
+    keep_tex: yes
     number_sections: true
+    toc: no
+bibliography: ref.bib
+fontsize: 11pt
+link-citations: no
+csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
 header-includes:
@@
 indent: true
+
 ---
@@
 library(ggdendro)
+library(here)
+library(feather)
 
 knitr::opts_chunk$set(echo = TRUE)
+```
+
+```{r load data, include = F}
+childes_results <- read_feather(here("../surprisal/Data/childes_words.feather"))
+wiki_results <- read_feather(here("../surprisal/Data/all_spaces.feather"))
+switchboard <- read_feather(here("../surprisal/Data/switchboard_words.feather"))
 ```
@@
 
-Due to their differences in distributional frequency, words vary in how much information-theoretic content they contain (Shannon, 1948).
+Due to their differences in distributional frequency, words vary in how much information-theoretic content they contain [@shannon1948mathematical].
 This information content plays an important role in determining the shape of words and speech.
-Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be as speakers minimize effort in the amount of information they encode in speech (Zipf, 1936).
+Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be as speakers minimize effort in the amount of information they encode in speech [@zipf1949human].
 Over and above frequency, contextual information content predicts word length (Piantadosi, Tily & Gibson, 2011).
-Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side (Levy & Jaeger, 2007).
+Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side [@jaeger2007speakers].
 Across utterances, context allows speakers to produce more informative and surprising sentences later in communication.
-The later a sentence occurs in a paragraph, the the informative it is and the more unusual and surprising words it contains (Genzel & Charniak, 2002).
+The later a sentence occurs in a paragraph, the the informative it is and the more unusual and surprising words it contains [@genzel2002entropy].
 From the level of individual words to sentences in paragraphs, information content shapes and is shaped by the context it appears in.
@@
 Writers use sentences to communicate thoughts, often on their own instead in paragraphs with other sentences.
-Cross-linguistically, the average rate of information speakers transfer in a syllable differs (Aylett & Turk, 2004).
+Cross-linguistically, the average rate of information speakers transfer in a syllable differs [@aylett2004smooth].
 Languages possess very different grammatical and morphological structures--the same thought can be encoded in at least as many ways as there are different languages.
@@
 
-Yu et al.
-(2016) proposed a novel method for quantifying the amount of information typically found at each part of English utterances.
+@yu2016distribution proposed a novel method for quantifying the amount of information typically found at each part of English utterances.
 The Yu et al work examined written sentences from the British National Corpus.
 Treating each word position by itself, Yu et al.
-calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty (Shannon, 1948).
+calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty [@shannon1948mathematical].
 Yu et al.
@@
 
-Shannon (1948) defined information as “the reduction in uncertainty about one variable given that the value of another variable is known”.
-We use a metric proposed by Shannon and applied by Levy (2008): lexical surprisal, which defines the information in word based on the ratio of possible continuations of the sentence after to before the word is seen.
+@shannon1948mathematical defined information as “the reduction in uncertainty about one variable given that the value of another variable is known”.
+We use a metric proposed by Shannon and applied by @levy2008expectation: lexical surprisal, which defines the information in word based on the ratio of possible continuations of the sentence after to before the word is seen.
 Equivalently, we can compute surprisal with the predictability of the word, as in the formula below.
@@
 
-The surprisal of a word is also correlated with the processing cost of a word, with evidence from eye-tracking (Smith & Levy, 2013) and ERP (Frank et al, 2014) studies, among other sources.
+The surprisal of a word is also correlated with the processing cost of a word, with evidence from eye-tracking [@smith2013effect] and ERP [@frank2015erp] studies, among other sources.
 Considered without context, the surprisal of an individual word is inversely proportional to the frequency of that word, so that the less often a person has seen a word, the more information that word holds.
@@
 Though the two words have the same length in number of letters, it’s more difficult to process “azalea” when reading it here than when reading “flower”.
-Frequency is intimately tied information content in words, with much of the differences between words frequencies being explained by information content cross-linguistically (Piantadosi et al, 2011).
+Frequency is intimately tied information content in words, with much of the differences between words frequencies being explained by information content cross-linguistically [@piantadosi2011word].
 
@@
 
-We first replicate the Yu et al result using the surprisal metric in place of the entropy metric.
+We first replicate the @yu2016distribution result using the surprisal metric in place of the entropy metric.
 We use the frequency-based or “contextless” surprisal metric, which determines the average distribution of information based on word frequencies in a corpus.
@@
 This distribution indicates that in English, the words we speak or write at the beginnings of utterances has little information, while the words we speak or write at the ends of utterances has a lot of information.
-The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from Genzel & Charniak (2002).
+The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from [@genzel2002entropy].
 
@@
 
-BNC and Switchboard plot with context
+```{r switchboard & bnc plot}
+
+```
 
@@
 
-We use the Providence corpus from CHILDES (MacWhinney, 2000; Demuth, 2012), which consists of about 650000 utterances from parent-child conversations for six monolingual English-speaking children talking with the parents.
+We use the Providence corpus from CHILDES [@macwhinney2000childes; @evans2012individual], which consists of about 650000 utterances from parent-child conversations for six monolingual English-speaking children talking with the parents.
 The child age range was between 11 months and 4 years.
@@
 
-We now turn to a small number of typologically diverse languages, and conduct the same analysis, using a monolingual adult-child speech corpora from CHILDES (MacWhinney, 2000) to compare the results from these languages directly to our results from the English Providence corpus.
+We now turn to a small number of typologically diverse languages, and conduct the same analysis, using a monolingual adult-child speech corpora from CHILDES [@macwhinney2000childes] to compare the results from these languages directly to our results from the English Providence corpus.
 We use corpora for Spanish, German, French, Mandarin and Japanese.
@@
 
-Messy plot of WALS features vs. cosine similarity goes here.
+```{r wals features vs cosine similarity}
+wiki_results %>% 
+  ggplot(aes(x = feature, y=unigram_cosine)) + 
+    geom_smooth(aes(weight = .01), se=F, method = "lm") + 
+    geom_hex()
+```
 
@@
 
-Messy plot of average LDN vs. cosine similarity goes here
+```{r ldn features vs cosine similarity}
+wiki_results %>% 
+  ggplot(aes(x = ldn, y=unigram_cosine)) + 
+    geom_smooth(aes(weight = .01), se=F, method = "lm") + 
+    geom_hex()
+```
 
@@
 This strategy avoids peaks in information which lead to issues for the decoder, while making the best use of the communication channel possible and aiming to produce as much information as the listener can decode at every moment.
+
+# References

### 2019-09-06 Paper/paper.tex
paper citations

--- 2019-09-05 Paper/paper.Rmd
@@
----
-title: "Speakers of diverse languages structure their utterances for efficient communication"
-author: "Josef Klafka and Daniel Yurovsky"
-date: "`r Sys.Date()`"
-output: 
-  pdf_document:
-    number_sections: true
-header-includes:
-    - \usepackage{setspace}\doublespacing
-indent: true
----
+\documentclass[11pt,]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={Speakers of diverse languages structure their utterances for efficient communication},
+            pdfauthor={Josef Klafka and Daniel Yurovsky},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\usepackage{longtable,booktabs}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-```{r setup, include=FALSE}
-library(tidyverse)
-library(ggdendro)
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
 
-knitr::opts_chunk$set(echo = TRUE)
-```
+%%% Change title format to be more compact
+\usepackage{titling}
 
-# Abstract
+% Create subtitle command for use in maketitle
+\providecommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+
+  \title{Speakers of diverse languages structure their utterances for efficient communication}
+    \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+    \author{Josef Klafka and Daniel Yurovsky}
+    \preauthor{\centering\large\emph}
+  \postauthor{\par}
+      \predate{\centering\large\emph}
+  \postdate{\par}
+    \date{2019-09-06}
+
+\usepackage{setspace}
+
+\doublespacing
+
+\begin{document}
+\maketitle
+
+\hypertarget{abstract}{%
+\section{Abstract}\label{abstract}}
 
@@
 Using tools from information theory and sentence processing, we find that speakers must structure the information in their utterances according to the grammatical, morphological and lexical constraints of their language, which vary greatly from language to language.
-This information structure characterizes language from children’s first utterances to large-scale knowledge base entries.
+This information structure characterizes language from children's first utterances to large-scale knowledge base entries.
 Once listeners have access to linguistic context within an utterance, then they can decode the rest of that utterance at a constant and optimal rate.
 
-# Introduction
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 Although we often need to say things we have never said before, we are easily and efficiently able to produce novel utterances with minimal prior planning.
-Our listeners incrementally process our utterances, and predict which words will come next within fractions of a second of hearing the beginning of the word (for evidence see e.g.
-Kutas & Federmeier, 2010).
+Our listeners incrementally process our utterances, and predict which words will come next within fractions of a second of hearing the beginning of the word (for evidence see e.g.~Kutas \& Federmeier, 2010).
 Encoding new and unpredictable information into our utterances surprises our listeners, and presents problems for communication.
@@
 This information content plays an important role in determining the shape of words and speech.
-Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be as speakers minimize effort in the amount of information they encode in speech (Zipf, 1936).
-Over and above frequency, contextual information content predicts word length (Piantadosi, Tily & Gibson, 2011).
-Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side (Levy & Jaeger, 2007).
+Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be as speakers minimize effort in the amount of information they encode in speech (Zipf, 1949).
+Over and above frequency, contextual information content predicts word length (Piantadosi, Tily \& Gibson, 2011).
+Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side (Jaeger \& Levy, 2007).
 Across utterances, context allows speakers to produce more informative and surprising sentences later in communication.
-The later a sentence occurs in a paragraph, the the informative it is and the more unusual and surprising words it contains (Genzel & Charniak, 2002).
+The later a sentence occurs in a paragraph, the the informative it is and the more unusual and surprising words it contains (Genzel \& Charniak, 2002).
 From the level of individual words to sentences in paragraphs, information content shapes and is shaped by the context it appears in.
@@
 Writers use sentences to communicate thoughts, often on their own instead in paragraphs with other sentences.
-Cross-linguistically, the average rate of information speakers transfer in a syllable differs (Aylett & Turk, 2004).
+Cross-linguistically, the average rate of information speakers transfer in a syllable differs (Aylett \& Turk, 2004).
 Languages possess very different grammatical and morphological structures--the same thought can be encoded in at least as many ways as there are different languages.
@@
 
-Yu et al.
-(2016) proposed a novel method for quantifying the amount of information typically found at each part of English utterances.
+Yu, Cong, Liang, \& Liu (2016) proposed a novel method for quantifying the amount of information typically found at each part of English utterances.
 The Yu et al work examined written sentences from the British National Corpus.
-Treating each word position by itself, Yu et al.
-calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty (Shannon, 1948).
-Yu et al.
-found a distinctive three-step shape to the distribution of information in the British National Corpus.
+Treating each word position by itself, Yu et al.~calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty (Shannon, 1948).
+Yu et al.~found a distinctive three-step shape to the distribution of information in the British National Corpus.
 
 In this paper, we expand on this body of prior work in a number of novel ways.
-We replicate Yu et al’s results with a different metric, one intimately tied to incremental word processing.
+We replicate Yu et al's results with a different metric, one intimately tied to incremental word processing.
 We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech.
 We expand the study of information density to the largest set of languages yet considered, and incorporate contextual and typological information.
-We find that each language has a characteristic, robust distribution of information predictable from its morphology, phonology and syntax, and tied to the language’s genealogy.
-Speakers of a language will tend to distribute information according to that language’s distribution of information in their utterances.
+We find that each language has a characteristic, robust distribution of information predictable from its morphology, phonology and syntax, and tied to the language's genealogy.
+Speakers of a language will tend to distribute information according to that language's distribution of information in their utterances.
 As soon as a child can form multi-word utterances in a given language, they will produce utterances obeying the characteristic distribution of information in a given language.
@@
 
-# Methods
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
 
-Shannon (1948) defined information as “the reduction in uncertainty about one variable given that the value of another variable is known”.
+Shannon (1948) defined information as ``the reduction in uncertainty about one variable given that the value of another variable is known''.
 We use a metric proposed by Shannon and applied by Levy (2008): lexical surprisal, which defines the information in word based on the ratio of possible continuations of the sentence after to before the word is seen.
@@
 
-$$surprisal(word) = -log P(word)$$
+\[surprisal(word) = -log P(word)\]
 
-The surprisal of a word is also correlated with the processing cost of a word, with evidence from eye-tracking (Smith & Levy, 2013) and ERP (Frank et al, 2014) studies, among other sources.
+The surprisal of a word is also correlated with the processing cost of a word, with evidence from eye-tracking (Smith \& Levy, 2013) and ERP (Frank, Otten, Galli, \& Vigliocco, 2015) studies, among other sources.
 Considered without context, the surprisal of an individual word is inversely proportional to the frequency of that word, so that the less often a person has seen a word, the more information that word holds.
-For example, “flower” has less information than “azalea” because “flower” is much more common than “azalea”.
-Though the two words have the same length in number of letters, it’s more difficult to process “azalea” when reading it here than when reading “flower”.
-Frequency is intimately tied information content in words, with much of the differences between words frequencies being explained by information content cross-linguistically (Piantadosi et al, 2011).
+For example, ``flower'' has less information than ``azalea'' because ``flower'' is much more common than ``azalea''.
+Though the two words have the same length in number of letters, it's more difficult to process ``azalea'' when reading it here than when reading ``flower''.
+Frequency is intimately tied information content in words, with much of the differences between words frequencies being explained by information content cross-linguistically (Piantadosi, Tily, \& Gibson, 2011).
 
-However, when reading or listening, people don’t just consider each word as an isolated linguistic signal.
+However, when reading or listening, people don't just consider each word as an isolated linguistic signal.
 Instead, listeners use the words they have already heard to predict and decode the word they are currently processing.
@@
 
-$$surprisal(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...) = -log P(w_i,w_{i-1}w_{i-2},...)P(w_{i-1}w_{i-2}...)$$
+\[surprisal(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...) = -log P(w_i,w_{i-1}w_{i-2},...)P(w_{i-1}w_{i-2}...)\]
 
 When we use a word or two of context in our surprisal calculations, then the set of reasonable final items in our ngrams is greatly restricted.
-“Flower” may contain less information than “azalea” when we consider the words independently of their context, but with context this can be reversed.
-Flower appears in a variety of contexts, and so the information content of a word like “flower” in a particular context may be higher than “azalea”.
-If you only have azaleas in your garden, then hearing someone say “in that garden, look at the flowers” may be higher surprisal for you: you expect them to say “azalea”.
-This prediction does not need many words before to work out, for example in “I take my coffee with cream and sugar”.
-When you hear “cream and”, you automatically predict “sugar” or maybe “honey”, but there are few possible continuations with even those two words.
-Hearing “I” restricts the next word to a verb, or possibly an adverb, and since you have just heard the speaker refer to themselves in the first person singular, your set of possible completions is significantly restricted.
+``Flower'' may contain less information than ``azalea'' when we consider the words independently of their context, but with context this can be reversed.
+Flower appears in a variety of contexts, and so the information content of a word like ``flower'' in a particular context may be higher than ``azalea''.
+If you only have azaleas in your garden, then hearing someone say ``in that garden, look at the flowers'' may be higher surprisal for you: you expect them to say ``azalea''.
+This prediction does not need many words before to work out, for example in ``I take my coffee with cream and sugar''.
+When you hear ``cream and'', you automatically predict ``sugar'' or maybe ``honey'', but there are few possible continuations with even those two words.
+Hearing ``I'' restricts the next word to a verb, or possibly an adverb, and since you have just heard the speaker refer to themselves in the first person singular, your set of possible completions is significantly restricted.
 
@@
 
-The frequency-based surprisal metric gives us an idea of when in their utterances speakers say frequent i.e.
-independently information-rich words.
+The frequency-based surprisal metric gives us an idea of when in their utterances speakers say frequent i.e.~independently information-rich words.
 The context-based surprisal metric show us how speakers tend to distribute the information in utterances relative to real-time processing in communication.
@@
 
-# Written and Spoken English have the same information distribution
+\hypertarget{written-and-spoken-english-have-the-same-information-distribution}{%
+\section{Written and Spoken English have the same information distribution}\label{written-and-spoken-english-have-the-same-information-distribution}}
 
 We first turn to working with written English in the British National Corpus (BNC).
-The BNC is a collection of spoken and written records (90% written) from the turn of the century, intended to be a representative sample of British English.
-Using their word entropy metric without context, Yu et al.
-found a distinctive three-step distribution for information in written English sentences in the corpus.
+The BNC is a collection of spoken and written records (90\% written) from the turn of the century, intended to be a representative sample of British English.
+Using their word entropy metric without context, Yu et al.~found a distinctive three-step distribution for information in written English sentences in the corpus.
 The first word tended to contain little information.
@@
 
-We first replicate the Yu et al result using the surprisal metric in place of the entropy metric.
-We use the frequency-based or “contextless” surprisal metric, which determines the average distribution of information based on word frequencies in a corpus.
+We first replicate the Yu et al.
+(2016) result using the surprisal metric in place of the entropy metric.
+We use the frequency-based or ``contextless'' surprisal metric, which determines the average distribution of information based on word frequencies in a corpus.
 A priori we expect that the frequency-based metric will produce a flat distribution of information across word positions in the BNC.
@@
 This distribution indicates that in English, the words we speak or write at the beginnings of utterances has little information, while the words we speak or write at the ends of utterances has a lot of information.
-The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from Genzel & Charniak (2002).
+The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from (Genzel \& Charniak, 2002).
 
 What about context?
-So far we’ve only discussed the unigram (“contextless”) surprisal metric, considering words on their own without any explicit incorporation of prior context.
+So far we've only discussed the unigram (``contextless'') surprisal metric, considering words on their own without any explicit incorporation of prior context.
 As previously discussed, listeners decode information and process what they hear incrementally, using prior heard words to ease the comprehension process.
@@
 
-BNC and Switchboard plot with context
-
 Speakers produce information at a more or less constant rate, avoiding peaks or troughs in their information distribution, except at the beginnings of utterances, where listeners may not have any context to predict what the speaker is going to say.
@@
 
-# English CDS and child speech
+\hypertarget{english-cds-and-child-speech}{%
+\section{English CDS and child speech}\label{english-cds-and-child-speech}}
 
@@
 
-We use the Providence corpus from CHILDES (MacWhinney, 2000; Demuth, 2012), which consists of about 650000 utterances from parent-child conversations for six monolingual English-speaking children talking with the parents.
+We use the Providence corpus from CHILDES (Evans \& Demuth, 2012; MacWhinney, 2000), which consists of about 650000 utterances from parent-child conversations for six monolingual English-speaking children talking with the parents.
 The child age range was between 11 months and 4 years.
@@
 Parents said a majority of the utterances, but children take up an increasing share of the conversation as they grow older and are actually able to speak full utterances.
-The utterances in the Providence corpus are on average significantly shorter than those in the BNC and Switchboard; over 90% of the utterances in the Providence corpus are at most 10 words long, with the median at around 6 words.
+The utterances in the Providence corpus are on average significantly shorter than those in the BNC and Switchboard; over 90\% of the utterances in the Providence corpus are at most 10 words long, with the median at around 6 words.
 Similar to our treatment of the Switchboard corpus, we split the Providence corpus up by conversational turns and separate utterances.
@@
 
-# Child speech and CDS across languages
+\hypertarget{child-speech-and-cds-across-languages}{%
+\section{Child speech and CDS across languages}\label{child-speech-and-cds-across-languages}}
 
@@
 The conversations in this corpus take place between the children and other children in the same class, and between the children and the investigator.
-For Mandarin Chinese, we use the Zhou corpus of dinner conversations from the Shanghai area (Li & Zhou, 2015).
+For Mandarin Chinese, we use the Zhou corpus of dinner conversations from the Shanghai area (Li \& Zhou, 2015).
 The children in this corpus are 5 or 6 years old, and half come from working class families (the other half from upper class families).
@@
 We observe a distinct and characteristic frequency-based information trajectory emerge for each language, robust across each utterance length within each language.
-We see the same distribution of information for both parents and children in each language, with the child’s curve normally below the parent’s curve, likely due to the parent speaking a larger share of the utterances in the corpus and using a larger vocabulary than the child.
+We see the same distribution of information for both parents and children in each language, with the child's curve normally below the parent's curve, likely due to the parent speaking a larger share of the utterances in the corpus and using a larger vocabulary than the child.
 We see the opposite amplitude trend in the Shiro corpus, where the children speak more than the adult investigators as the children are the ones telling the stories.
@@
 
-# Language structure and large-scale data analysis: methods and preprocessing
+\hypertarget{language-structure-and-large-scale-data-analysis-methods-and-preprocessing}{%
+\section{Language structure and large-scale data analysis: methods and preprocessing}\label{language-structure-and-large-scale-data-analysis-methods-and-preprocessing}}
 
 To make a claim about language as a whole and languages on a larger scale, we needed to use larger corpora and a much larger number of languages.
-We pulled corpora for 159 diverse languages from Wikipedia [Citation], each of which had at least 10000 articles.
+We pulled corpora for 159 diverse languages from Wikipedia {[}Citation{]}, each of which had at least 10000 articles.
 We split each article up into sentences; the variance in sentence lengths for Wikipedia was significantly larger than for the CHILDES corpora we used in the previous section.
@@
 We used two different strategies, which yielded identical results upon analysis.
-Each strategy gave us a five-dimensional vector space embedding for each language in a Wikipedia “slope space”.
+Each strategy gave us a five-dimensional vector space embedding for each language in a Wikipedia ``slope space''.
 For the first strategy, we split each sentence length by number of words into fifths, and computed surprisal values for the closest word position to each quintile.
@@
 
-To more rigorously described the differences between languages, we used data from the World Atlas of Language Structures (WALS; Dryer & Haspelmath eds., 2013).
+To more rigorously described the differences between languages, we used data from the World Atlas of Language Structures (WALS; Dryer \& Haspelmath eds., 2013).
 WALS has data for 144 typological features in 2569 languages around the world.
@@
 However, the WALS features describe specific structural differences between languages, while our surprisal metric is lexical.
-To target lexical differences between languages, we computed the average normalized Levenshtein distance (LDN; Holman et al, 2008) over the 40 item Swadesh list (1955), retrieved from the ASJP database (Wichmann, Holman & Brown, 2018).
+To target lexical differences between languages, we computed the average normalized Levenshtein distance (LDN; Holman et al, 2008) over the 40 item Swadesh list (1955), retrieved from the ASJP database (Wichmann, Holman \& Brown, 2018).
 The Swadesh list is designed to include near-universal words that target basic cognitive concepts, and are useful in determining the genealogical similarities and differences between languages.
@@
 
-# Language structures and large-scale data analysis: results
+\hypertarget{language-structures-and-large-scale-data-analysis-results}{%
+\section{Language structures and large-scale data analysis: results}\label{language-structures-and-large-scale-data-analysis-results}}
 
@@
 First we examined the effects of individual typological features on the shapes of the unigram information curves.
-We ran logistic regressions using the lme4 package in R (lme4), checking whether the cosine distance between two languages’ embeddings in the slope space played a role in determining if they had the same value for a given WALS feature.
+We ran logistic regressions using the lme4 package in R (lme4), checking whether the cosine distance between two languages' embeddings in the slope space played a role in determining if they had the same value for a given WALS feature.
 Individual WALS features do not necessarily have ordinal values.
-Some, such as the “Number of Cases” feature, are easy to quantify and order.
+Some, such as the ``Number of Cases'' feature, are easy to quantify and order.
 Others are more difficult.
-For example, how does one order “relative clauses appear after the nouns they modify”, “relative clauses appear before the nouns they modify” and “free order of relative clauses and nouns”?
+For example, how does one order ``relative clauses appear after the nouns they modify'', ``relative clauses appear before the nouns they modify'' and ``free order of relative clauses and nouns''?
 We chose the identify relation to avoid deciding on the basis of individual features.
-We found that 100 out of the 120 features from WALS we examined was statistically significant (p < .001) in determining whether two languages had the same shape to their unigram information distributions.
+We found that 100 out of the 120 features from WALS we examined was statistically significant (p \textless{} .001) in determining whether two languages had the same shape to their unigram information distributions.
 The results for some important features are below, and the rest of the results are in the Appendices.
@@
 We next compared how the cosine distance between two languages related to how many WALS features they had in common.
-There is a small correlation with a very small R^2 value between WALS features similarity and cosine similarity between two languages, but a statistically significant slope (p < .001).
+There is a small correlation with a very small R\^{}2 value between WALS features similarity and cosine similarity between two languages, but a statistically significant slope (p \textless{} .001).
 Essentially, for every feature two languages have in common, they have one degree more cosine similarity in the slope space.
-As feature values tend to be highly correlated within categories (e.g.
-the word order features are highly correlated), then this translates into language having much more similar information curves if they have more WALS features in common, an intuitive result.
+As feature values tend to be highly correlated within categories (e.g.~the word order features are highly correlated), then this translates into language having much more similar information curves if they have more WALS features in common, an intuitive result.
 
-Messy plot of WALS features vs. cosine similarity goes here.
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{wiki_results }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ feature, }\DataTypeTok{y=}\NormalTok{unigram_cosine)) }\OperatorTok{+}\StringTok{ }
+\StringTok{    }\KeywordTok{geom_smooth}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{weight =} \FloatTok{.01}\NormalTok{), }\DataTypeTok{se=}\NormalTok{F, }\DataTypeTok{method =} \StringTok{"lm"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
+\StringTok{    }\KeywordTok{geom_hex}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{paper_files/figure-latex/wals features vs cosine similarity-1.pdf}
 
@@
 
-Messy plot of average LDN vs. cosine similarity goes here
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{wiki_results }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ ldn, }\DataTypeTok{y=}\NormalTok{unigram_cosine)) }\OperatorTok{+}\StringTok{ }
+\StringTok{    }\KeywordTok{geom_smooth}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{weight =} \FloatTok{.01}\NormalTok{), }\DataTypeTok{se=}\NormalTok{F, }\DataTypeTok{method =} \StringTok{"lm"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
+\StringTok{    }\KeywordTok{geom_hex}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
 
-The shape of a language’s frequency-based information curve covaries with its typological and lexical similarity to other languages.
+\includegraphics{paper_files/figure-latex/ldn features vs cosine similarity-1.pdf}
 
-# General Discussion
+The shape of a language's frequency-based information curve covaries with its typological and lexical similarity to other languages.
+
+\hypertarget{general-discussion}{%
+\section{General Discussion}\label{general-discussion}}
 
@@
 
-The Uniform Information Density Hypothesis (Levy & Jaeger, 2007) provides an optimal strategy for distributing information among the words in an utterance.
+The Uniform Information Density Hypothesis (Levy \& Jaeger, 2007) provides an optimal strategy for distributing information among the words in an utterance.
 The hypothesis suggests that speakers tend to distribute information evenly in their utterances, close to the capacity of the communication channel.
 This strategy avoids peaks in information which lead to issues for the decoder, while making the best use of the communication channel possible and aiming to produce as much information as the listener can decode at every moment.
+
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-aylett2004smooth}{}%
+Aylett, M., \& Turk, A.
+(2004).
+The smooth signal redundancy hypothesis: A functional explanation for relationships between redundancy, prosodic prominence, and duration in spontaneous speech.
+\emph{Language and Speech}, \emph{47}(1), 31--56.
+
+\leavevmode\hypertarget{ref-evans2012individual}{}%
+Evans, K. E., \& Demuth, K. (2012). Individual differences in pronoun reversal: Evidence from two longitudinal case studies. \emph{Journal of Child Language}, \emph{39}(1), 162--191.
+
+\leavevmode\hypertarget{ref-frank2015erp}{}%
+Frank, S. L., Otten, L. J., Galli, G., \& Vigliocco, G. (2015). The erp response to the amount of information conveyed by words in sentences. \emph{Brain and Language}, \emph{140}, 1--11.
+
+\leavevmode\hypertarget{ref-genzel2002entropy}{}%
+Genzel, D., \& Charniak, E. (2002). Entropy rate constancy in text. \emph{Proceedings of the 40th annual meeting of the association for computational linguistics}, 199--206.
+
+\leavevmode\hypertarget{ref-jaeger2007speakers}{}%
+Jaeger, T. F., \& Levy, R. P. (2007). Speakers optimize information density through syntactic reduction. \emph{Advances in neural information processing systems}, 849--856.
+
+\leavevmode\hypertarget{ref-levy2008expectation}{}%
+Levy, R. (2008). Expectation-based syntactic comprehension. \emph{Cognition}, \emph{106}(3), 1126--1177.
+
+\leavevmode\hypertarget{ref-macwhinney2000childes}{}%
+MacWhinney, B. (2000). \emph{The childes project: The database} (Vol. 2). Psychology Press.
+
+\leavevmode\hypertarget{ref-piantadosi2011word}{}%
+Piantadosi, S. T., Tily, H., \& Gibson, E. (2011). Word lengths are optimized for efficient communication. \emph{Proceedings of the National Academy of Sciences}, \emph{108}(9), 3526--3529.
+
+\leavevmode\hypertarget{ref-shannon1948mathematical}{}%
+Shannon, C. E. (1948). A mathematical theory of communication. \emph{Bell System Technical Journal}, \emph{27}(3), 379--423.
+
+\leavevmode\hypertarget{ref-smith2013effect}{}%
+Smith, N. J., \& Levy, R. (2013). The effect of word predictability on reading time is logarithmic. \emph{Cognition}, \emph{128}(3), 302--319.
+
+\leavevmode\hypertarget{ref-yu2016distribution}{}%
+Yu, S., Cong, J., Liang, J., \& Liu, H. (2016). The distribution of information content in english sentences. \emph{arXiv Preprint arXiv:1609.07681}.
+
+\leavevmode\hypertarget{ref-zipf1949human}{}%
+Zipf, G. K. (1949). \emph{Human behavior and the principle of least effort.}
+
+\end{document}

### 2019-09-06 Paper/paper.Rmd
added prov plots to paper

@@
 
-$$surprisal(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...) = -log P(w_i,w_{i-1}w_{i-2},...)P(w_{i-1}w_{i-2}...)$$
+$$surprisal(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...)$$
+$$= -log \frac{P(w_i,w_{i-1}w_{i-2},...)}{P(w_{i-1}w_{i-2}...)}$$
 
@@
 
-Plot of Providence unigram curves
+```{r, echo=F, fig.height =2, fig.cap="Providence frequency-based information curves"}
+childes_results %>% 
+  filter(Language == "English", context == 0, length %in% c(5, 7, 9)) %>%
+  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
+    geom_point() + 
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = 1:9) + 
+    geom_line() + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
 
@@
 
-Plot of Providence bigram and trigram curves
+```{r, echo=F, fig.height =2, fig.cap="Providence context-based information curves"}
+childes_results %>% 
+  filter(Language == "English", context %in% c(1, 2), length %in% c(5, 7, 9)) %>%
+  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
+    geom_point() + 
+    facet_wrap(context~length) + 
+    scale_x_continuous(breaks = 1:9) + 
+    geom_line() + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
 

### 2019-09-06 Paper/paper.tex
added prov plots to paper

@@
 
-\[surprisal(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...) = -log P(w_i,w_{i-1}w_{i-2},...)P(w_{i-1}w_{i-2}...)\]
+\[surprisal(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...)\]
+\[= -log \frac{P(w_i,w_{i-1}w_{i-2},...)}{P(w_{i-1}w_{i-2}...)}\]
 
@@
 
-Plot of Providence unigram curves
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
+\caption{\label{fig:unnamed-chunk-1}Providence frequency-based information curves}
+\end{figure}
 
@@
 
-Plot of Providence bigram and trigram curves
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
+\caption{\label{fig:unnamed-chunk-2}Providence context-based information curves}
+\end{figure}
 

### 2019-09-09 Paper/paper.Rmd
more plots in paper

@@
 
-```{r, echo=F, fig.height =2, fig.cap="Providence context-based information curves"}
+```{r, echo=F, fig.height =2, fig.cap="Providence bigram context-based information curves"}
 childes_results %>% 
-  filter(Language == "English", context %in% c(1, 2), length %in% c(5, 7, 9)) %>%
+  filter(Language == "English", context == 1, length %in% c(5, 7, 9)) %>%
   ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
     geom_point() + 
-    facet_wrap(context~length) + 
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = 1:9) + 
+    geom_line() + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
+
+```{r, echo=F, fig.height =2, fig.cap="Providence trigram context-based information curves"}
+childes_results %>% 
+  filter(Language == "English", context == 2, length %in% c(5, 7, 9)) %>%
+  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
+    geom_point() + 
+    facet_wrap(~length) + 
     scale_x_continuous(breaks = 1:9) + 
@@
 
-Plots of CHILDES frequency-based information curves
+```{r, echo=F, fig.height =3, fig.cap="CHILDES frequency-based information curves"}
+childes_results %>% 
+  filter(context == 0, length == 7) %>%
+  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
+    geom_point() + 
+    facet_wrap(~Language) + 
+    scale_x_continuous(breaks = 1:9) + 
+    geom_line() + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
 
@@
 
-Plots of CHILDES trigram (and bigram) information curves
+```{r, echo=F, fig.height =3, fig.cap="CHILDES bigram context-based information curves"}
+childes_results %>% 
+  filter(context == 1, length == 7) %>%
+  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
+    geom_point() + 
+    facet_wrap(~Language) + 
+    scale_x_continuous(breaks = 1:9) + 
+    geom_line() + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
+
+```{r, echo=F, fig.height =3, fig.cap="CHILDES trigram context-based information curves"}
+childes_results %>% 
+  filter(context == 2, length == 7) %>%
+  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
+    geom_point() + 
+    facet_wrap(~Language) + 
+    scale_x_continuous(breaks = 1:9) + 
+    geom_line() + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
 
@@
 
-```{r wals features vs cosine similarity}
+```{r, echo = F, fig.cap="wals features vs cosine similarity"}
 wiki_results %>% 
@@
 
-```{r ldn features vs cosine similarity}
+```{r, echo = F, fig.cap="ldn features vs cosine similarity"}
 wiki_results %>% 

### 2019-09-09 Paper/paper.tex
more plots in paper

@@
 \urlstyle{same}  % don't use monospace font for urls
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{longtable,booktabs}
@@
   \postdate{\par}
-    \date{2019-09-06}
+    \date{2019-09-09}
 
@@
 \includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
-\caption{\label{fig:unnamed-chunk-2}Providence context-based information curves}
+\caption{\label{fig:unnamed-chunk-2}Providence bigram context-based information curves}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
+\caption{\label{fig:unnamed-chunk-3}Providence trigram context-based information curves}
 \end{figure}
@@
 
-Plots of CHILDES frequency-based information curves
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/unnamed-chunk-4-1.pdf}
+\caption{\label{fig:unnamed-chunk-4}CHILDES frequency-based information curves}
+\end{figure}
 
@@
 
-Plots of CHILDES trigram (and bigram) information curves
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1.pdf}
+\caption{\label{fig:unnamed-chunk-5}CHILDES bigram context-based information curves}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}
+\caption{\label{fig:unnamed-chunk-6}CHILDES trigram context-based information curves}
+\end{figure}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{wiki_results }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ feature, }\DataTypeTok{y=}\NormalTok{unigram_cosine)) }\OperatorTok{+}\StringTok{ }
-\StringTok{    }\KeywordTok{geom_smooth}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{weight =} \FloatTok{.01}\NormalTok{), }\DataTypeTok{se=}\NormalTok{F, }\DataTypeTok{method =} \StringTok{"lm"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
-\StringTok{    }\KeywordTok{geom_hex}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{paper_files/figure-latex/wals features vs cosine similarity-1.pdf}
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}
+\caption{\label{fig:unnamed-chunk-7}wals features vs cosine similarity}
+\end{figure}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{wiki_results }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ ldn, }\DataTypeTok{y=}\NormalTok{unigram_cosine)) }\OperatorTok{+}\StringTok{ }
-\StringTok{    }\KeywordTok{geom_smooth}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{weight =} \FloatTok{.01}\NormalTok{), }\DataTypeTok{se=}\NormalTok{F, }\DataTypeTok{method =} \StringTok{"lm"}\NormalTok{) }\OperatorTok{+}\StringTok{ }
-\StringTok{    }\KeywordTok{geom_hex}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{paper_files/figure-latex/ldn features vs cosine similarity-1.pdf}
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
+\caption{\label{fig:unnamed-chunk-8}ldn features vs cosine similarity}
+\end{figure}
 

### 2019-09-11 Paper/paper.Rmd
all references in bibtex and paper

@@
 library(feather)
+library(ggdendro)
 
@@
 switchboard <- read_feather(here("../surprisal/Data/switchboard_words.feather"))
+wiki_curves <- read_csv(here("../surprisal/Data/relative_ngrams.csv"))
 ```
@@
 Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be as speakers minimize effort in the amount of information they encode in speech [@zipf1949human].
-Over and above frequency, contextual information content predicts word length (Piantadosi, Tily & Gibson, 2011).
+Over and above frequency, contextual information content predicts word length [@piantadosi2011word].
 Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side [@jaeger2007speakers].
@@
 In this paper, we expand on this body of prior work in a number of novel ways.
-We replicate Yu et al’s results with a different metric, one intimately tied to incremental word processing.
+We replicate the results from @yu2016distribution with a different metric, one intimately tied to incremental word processing.
 We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech.
@@
 
-We first turn to working with written English in the British National Corpus (BNC).
+We first turn to working with written English in the British National Corpus [BNC; @leech1992100].
 The BNC is a collection of spoken and written records (90% written) from the turn of the century, intended to be a representative sample of British English.
@@
 See in the figure below a qualitative comparison of the information trajectory they found and the information trajectory we found.
-It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as Yu et al did.
+It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as @yu2016distribution did.
 The first words in written English sentences tend to contain little information, while the final words tend to contain a great deal of information.
@@
 We use the Providence corpus from CHILDES [@macwhinney2000childes; @evans2012individual], which consists of about 650000 utterances from parent-child conversations for six monolingual English-speaking children talking with the parents.
+We obtained the Providence corpus using the childesr frontend to the childes-db database [@sanchez2019childes].
 The child age range was between 11 months and 4 years.
@@
 Similarly to the English Providence corpus, all of the following corpora consist mainly of shorter utterances: most utterances in the corpora are under 10 words long.
-For Spanish, we use the corpus of conversations between primary school-aged children between 6 and 9 years old and the researchers, in which the children were asked to recite stories in their native Spanish (Shiro, 1997).
+For Spanish, we use the corpus of conversations between primary school-aged children between 6 and 9 years old and the researchers, in which the children were asked to recite stories in their native Spanish [@de1996estudio].
 The students came from working class public schools and upper class private schools.
-For German, we use the Wagner corpus (Wagner, 1985).
+For German, we use the Wagner corpus [@wagner1985much].
 This corpus consists of a collection of mini-corpora collected from socio-economically diverse families, with children aged from 1;0 to 14;0, mainly on the younger end of that range.
 The corpus mainly consists of parent-child conversations between the children and their primary caregivers.
-For French, we use the Palasis corpus (Palasis, 2009), a longitudinal study of children between the ages of 2;5 and 4;0 from the same kindergarten class.
+For French, we use the Palasis corpus [@palasis2009syntaxe], a longitudinal study of children between the ages of 2;5 and 4;0 from the same kindergarten class.
 The conversations in this corpus take place between the children and other children in the same class, and between the children and the investigator.
-For Mandarin Chinese, we use the Zhou corpus of dinner conversations from the Shanghai area (Li & Zhou, 2015).
+For Mandarin Chinese, we use the Zhou corpus of dinner conversations from the Shanghai area [@li2015dinner].
 The children in this corpus are 5 or 6 years old, and half come from working class families (the other half from upper class families).
-For Japanese, we use the Okayama corpus (Okayama, 1973).
+For Japanese, we use the Okayama corpus [@shirai2001acquisition].
 This corpus consists of conversations between children between the ages of 2;2 and 4;2 and their mothers.
@@
 To make a claim about language as a whole and languages on a larger scale, we needed to use larger corpora and a much larger number of languages.
-We pulled corpora for 159 diverse languages from Wikipedia [Citation], each of which had at least 10000 articles.
+We pulled corpora for 159 diverse languages from Wikipedia, each of which had at least 10000 articles.
 We split each article up into sentences; the variance in sentence lengths for Wikipedia was significantly larger than for the CHILDES corpora we used in the previous section.
@@
 
-To more rigorously described the differences between languages, we used data from the World Atlas of Language Structures (WALS; Dryer & Haspelmath eds., 2013).
+To more rigorously described the differences between languages, we used data from the World Atlas of Language Structures [WALS; @wals].
 WALS has data for 144 typological features in 2569 languages around the world.
@@
 
-We used the Multiple Imputation Multiple Correspondence Analysis algorithm (MIMCA; Citation) to fill in the missing data using statistical imputation.
+We used the Multiple Imputation Multiple Correspondence Analysis algorithm [MIMCA; @audigier2017mimca] to fill in the missing data using statistical imputation.
 MIMCA essentially uses mean imputation to begin with the missing values, then performs repeatedly performs principle components analysis on and reconstructs the contingency table formed from observations in categorical variables.
@@
 However, the WALS features describe specific structural differences between languages, while our surprisal metric is lexical.
-To target lexical differences between languages, we computed the average normalized Levenshtein distance (LDN; Holman et al, 2008) over the 40 item Swadesh list (1955), retrieved from the ASJP database (Wichmann, Holman & Brown, 2018).
+To target lexical differences between languages, we computed the average normalized Levenshtein distance [LDN; @holman2008advances] over the 40 item Swadesh list [@swadesh1955towards], retrieved from the ASJP database [@wichmann2016asjp].
 The Swadesh list is designed to include near-universal words that target basic cognitive concepts, and are useful in determining the genealogical similarities and differences between languages.
-The results of classifying languages using the Swadesh list and LDN are correlated with those using WALS features, but the Swadesh list and LDN do not suffer from the same sparsity problem as WALS (Holman et al, 2008).
+The results of classifying languages using the Swadesh list and LDN are correlated with those using WALS features, but the Swadesh list and LDN do not suffer from the same sparsity problem as WALS [@holman2008advances].
 
@@
 
-We ran a hierarchical clustering algorithm on the unigram information curves using the hclust package in R (hclust).
+We ran a hierarchical clustering algorithm on the unigram information curves using the hclust package from the R stats core library [@team2013r].
 We used the complete linkage algorithm for hierarchical clustering with distances between information curves in language computed using cosine distance between their embeddings in the slope space.
@@
 
-Dendro goes here
+```{r, echo=F}
+wiki_curves %>%
+  filter(gram == "Unigram") %>%
+  select(-gram) %>% 
+  sample_n(40) %>%
+  column_to_rownames("language") %>%
+  dist() %>% 
+  hclust() %>%
+  ggdendrogram(rotate=T, leaf_labels = F)
+```
 
@@
 
-Plot of some trigram surprisal curves goes here
+```{r, echo=F, fig.cap="Some trigram information curves from the Wikipedia data"}
+wiki_curves %>%
+  filter(gram == "Trigram") %>% 
+  mutate(Slope0 = runif(1, 0.9, 1.1)) %>% 
+  mutate(Slope1 = Slope1 + Slope0) %>% 
+  mutate(Slope2 = Slope2 + Slope1) %>% 
+  mutate(Slope3 = Slope3 + Slope2) %>% 
+  mutate(Slope4 = Slope4 + Slope3) %>% 
+  mutate(Slope5 = Slope5 + Slope4) %>% 
+  select(-gram) %>% 
+  sample_n(6) %>% 
+  gather(position, information, -language) %>%
+  ggplot(aes(x = position, y = information, group = language, color = language)) + 
+    geom_point() + 
+    geom_line() + 
+    xlab("Word position") + 
+    ylab("Average Information") + 
+    theme(axis.text.x = element_blank())
+```
 
 First we examined the effects of individual typological features on the shapes of the unigram information curves.
-We ran logistic regressions using the lme4 package in R (lme4), checking whether the cosine distance between two languages’ embeddings in the slope space played a role in determining if they had the same value for a given WALS feature.
+We ran logistic regressions using the lme4 package in R [@bates2014fitting], checking whether the cosine distance between two languages’ embeddings in the slope space played a role in determining if they had the same value for a given WALS feature.
 Individual WALS features do not necessarily have ordinal values.
@@
 
-The Uniform Information Density Hypothesis (Levy & Jaeger, 2007) provides an optimal strategy for distributing information among the words in an utterance.
+The Uniform Information Density Hypothesis [@jaeger2007speakers] provides an optimal strategy for distributing information among the words in an utterance.
 The hypothesis suggests that speakers tend to distribute information evenly in their utterances, close to the capacity of the communication channel.

### 2019-09-11 Paper/paper.tex
all references in bibtex and paper

@@
   \postdate{\par}
-    \date{2019-09-09}
+    \date{2019-09-11}
 
@@
 Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be as speakers minimize effort in the amount of information they encode in speech (Zipf, 1949).
-Over and above frequency, contextual information content predicts word length (Piantadosi, Tily \& Gibson, 2011).
+Over and above frequency, contextual information content predicts word length (Piantadosi, Tily, \& Gibson, 2011).
 Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side (Jaeger \& Levy, 2007).
@@
 In this paper, we expand on this body of prior work in a number of novel ways.
-We replicate Yu et al's results with a different metric, one intimately tied to incremental word processing.
+We replicate the results from Yu et al.
+(2016) with a different metric, one intimately tied to incremental word processing.
 We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech.
@@
 Though the two words have the same length in number of letters, it's more difficult to process ``azalea'' when reading it here than when reading ``flower''.
-Frequency is intimately tied information content in words, with much of the differences between words frequencies being explained by information content cross-linguistically (Piantadosi, Tily, \& Gibson, 2011).
+Frequency is intimately tied information content in words, with much of the differences between words frequencies being explained by information content cross-linguistically (Piantadosi et al., 2011).
 
@@
 
-We first turn to working with written English in the British National Corpus (BNC).
+We first turn to working with written English in the British National Corpus (BNC; Leech, 1992).
 The BNC is a collection of spoken and written records (90\% written) from the turn of the century, intended to be a representative sample of British English.
@@
 See in the figure below a qualitative comparison of the information trajectory they found and the information trajectory we found.
-It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as Yu et al did.
+It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as Yu et al.
+(2016) did.
 The first words in written English sentences tend to contain little information, while the final words tend to contain a great deal of information.
@@
 We use the Providence corpus from CHILDES (Evans \& Demuth, 2012; MacWhinney, 2000), which consists of about 650000 utterances from parent-child conversations for six monolingual English-speaking children talking with the parents.
+We obtained the Providence corpus using the childesr frontend to the childes-db database (Sanchez et al., 2019).
 The child age range was between 11 months and 4 years.
@@
 Similarly to the English Providence corpus, all of the following corpora consist mainly of shorter utterances: most utterances in the corpora are under 10 words long.
-For Spanish, we use the corpus of conversations between primary school-aged children between 6 and 9 years old and the researchers, in which the children were asked to recite stories in their native Spanish (Shiro, 1997).
+For Spanish, we use the corpus of conversations between primary school-aged children between 6 and 9 years old and the researchers, in which the children were asked to recite stories in their native Spanish (Shiro, 1996).
 The students came from working class public schools and upper class private schools.
@@
 The children in this corpus are 5 or 6 years old, and half come from working class families (the other half from upper class families).
-For Japanese, we use the Okayama corpus (Okayama, 1973).
+For Japanese, we use the Okayama corpus (Shirai, 2001).
 This corpus consists of conversations between children between the ages of 2;2 and 4;2 and their mothers.
@@
 To make a claim about language as a whole and languages on a larger scale, we needed to use larger corpora and a much larger number of languages.
-We pulled corpora for 159 diverse languages from Wikipedia {[}Citation{]}, each of which had at least 10000 articles.
+We pulled corpora for 159 diverse languages from Wikipedia, each of which had at least 10000 articles.
 We split each article up into sentences; the variance in sentence lengths for Wikipedia was significantly larger than for the CHILDES corpora we used in the previous section.
@@
 
-To more rigorously described the differences between languages, we used data from the World Atlas of Language Structures (WALS; Dryer \& Haspelmath eds., 2013).
+To more rigorously described the differences between languages, we used data from the World Atlas of Language Structures (WALS; Dryer \& Haspelmath, 2013).
 WALS has data for 144 typological features in 2569 languages around the world.
@@
 
-We used the Multiple Imputation Multiple Correspondence Analysis algorithm (MIMCA; Citation) to fill in the missing data using statistical imputation.
+We used the Multiple Imputation Multiple Correspondence Analysis algorithm (MIMCA; Audigier, Husson, \& Josse, 2017) to fill in the missing data using statistical imputation.
 MIMCA essentially uses mean imputation to begin with the missing values, then performs repeatedly performs principle components analysis on and reconstructs the contingency table formed from observations in categorical variables.
@@
 However, the WALS features describe specific structural differences between languages, while our surprisal metric is lexical.
-To target lexical differences between languages, we computed the average normalized Levenshtein distance (LDN; Holman et al, 2008) over the 40 item Swadesh list (1955), retrieved from the ASJP database (Wichmann, Holman \& Brown, 2018).
+To target lexical differences between languages, we computed the average normalized Levenshtein distance (LDN; Holman et al., 2008) over the 40 item Swadesh list (Swadesh, 1955), retrieved from the ASJP database (Wichmann et al., 2016).
 The Swadesh list is designed to include near-universal words that target basic cognitive concepts, and are useful in determining the genealogical similarities and differences between languages.
-The results of classifying languages using the Swadesh list and LDN are correlated with those using WALS features, but the Swadesh list and LDN do not suffer from the same sparsity problem as WALS (Holman et al, 2008).
+The results of classifying languages using the Swadesh list and LDN are correlated with those using WALS features, but the Swadesh list and LDN do not suffer from the same sparsity problem as WALS (Holman et al., 2008).
 
@@
 
-We ran a hierarchical clustering algorithm on the unigram information curves using the hclust package in R (hclust).
+We ran a hierarchical clustering algorithm on the unigram information curves using the hclust package from the R stats core library (Team \& others, 2013).
 We used the complete linkage algorithm for hierarchical clustering with distances between information curves in language computed using cosine distance between their embeddings in the slope space.
@@
 
-Dendro goes here
+\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}
 
@@
 
-Plot of some trigram surprisal curves goes here
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
+\caption{\label{fig:unnamed-chunk-8}Some trigram information curves from the Wikipedia data}
+\end{figure}
 
 First we examined the effects of individual typological features on the shapes of the unigram information curves.
-We ran logistic regressions using the lme4 package in R (lme4), checking whether the cosine distance between two languages' embeddings in the slope space played a role in determining if they had the same value for a given WALS feature.
+We ran logistic regressions using the lme4 package in R (Bates, Mächler, Bolker, \& Walker, 2014), checking whether the cosine distance between two languages' embeddings in the slope space played a role in determining if they had the same value for a given WALS feature.
 Individual WALS features do not necessarily have ordinal values.
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}
-\caption{\label{fig:unnamed-chunk-7}wals features vs cosine similarity}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1.pdf}
+\caption{\label{fig:unnamed-chunk-9}wals features vs cosine similarity}
 \end{figure}
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
-\caption{\label{fig:unnamed-chunk-8}ldn features vs cosine similarity}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1.pdf}
+\caption{\label{fig:unnamed-chunk-10}ldn features vs cosine similarity}
 \end{figure}
@@
 
-The Uniform Information Density Hypothesis (Levy \& Jaeger, 2007) provides an optimal strategy for distributing information among the words in an utterance.
+The Uniform Information Density Hypothesis (Jaeger \& Levy, 2007) provides an optimal strategy for distributing information among the words in an utterance.
 The hypothesis suggests that speakers tend to distribute information evenly in their utterances, close to the capacity of the communication channel.
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-audigier2017mimca}{}%
+Audigier, V., Husson, F., \& Josse, J. (2017). MIMCA: Multiple imputation for categorical variables with multiple correspondence analysis. \emph{Statistics and Computing}, \emph{27}(2), 501--518.
+
 \leavevmode\hypertarget{ref-aylett2004smooth}{}%
@@
 \emph{Language and Speech}, \emph{47}(1), 31--56.
+
+\leavevmode\hypertarget{ref-bates2014fitting}{}%
+Bates, D., Mächler, M., Bolker, B., \& Walker, S. (2014). Fitting linear mixed-effects models using lme4. \emph{arXiv Preprint arXiv:1406.5823}.
+
+\leavevmode\hypertarget{ref-wals}{}%
+Dryer, M. S., \& Haspelmath, M. (Eds.). (2013). \emph{WALS online}. Retrieved from \url{https://wals.info/}
 
@@
 
+\leavevmode\hypertarget{ref-holman2008advances}{}%
+Holman, E.
+W., Wichmann, S., Brown, C.
+H., Velupillai, V., Müller, A., Bakker, D., \& others.
+(2008).
+Advances in automated language classification.
+\emph{Quantitative Investigations in Theoretical Linguistics}, 40--43.
+
 \leavevmode\hypertarget{ref-jaeger2007speakers}{}%
 Jaeger, T. F., \& Levy, R. P. (2007). Speakers optimize information density through syntactic reduction. \emph{Advances in neural information processing systems}, 849--856.
+
+\leavevmode\hypertarget{ref-leech1992100}{}%
+Leech, G. N. (1992). \emph{100 million words of english: The british national corpus (bnc)}.
 
@@
 
+\leavevmode\hypertarget{ref-li2015dinner}{}%
+Li, H., \& Zhou, J. (2015). \emph{Study on dinner table talk of preschool children family in shanghai} (Master's thesis). East China Normal University, Shanghai, China.
+
 \leavevmode\hypertarget{ref-macwhinney2000childes}{}%
 MacWhinney, B. (2000). \emph{The childes project: The database} (Vol. 2). Psychology Press.
+
+\leavevmode\hypertarget{ref-palasis2009syntaxe}{}%
+Palasis, K. (2009). \emph{Syntaxe générative et acquisition: Le sujet dans le développement du système linguistique du jeune enfant} (PhD thesis). Nice.
 
@@
 
+\leavevmode\hypertarget{ref-sanchez2019childes}{}%
+Sanchez, A., Meylan, S.
+C., Braginsky, M., MacDonald, K.
+E., Yurovsky, D., \& Frank, M.
+C.
+(2019).
+Childes-db: A flexible and reproducible interface to the child language data exchange system.
+\emph{Behavior Research Methods}, \emph{51}(4), 1928--1941.
+
 \leavevmode\hypertarget{ref-shannon1948mathematical}{}%
@@
 
+\leavevmode\hypertarget{ref-shirai2001acquisition}{}%
+Shirai, Y. (2001). The acquisition of causative morphology in japanese: A prototype account. \emph{East Asian Language Acquisition}.
+
+\leavevmode\hypertarget{ref-de1996estudio}{}%
+Shiro, M. L. de. (1996). Un estudio de las expresiones de modalidad en hablantes de dos culturas. \emph{Boletı'n de Lingüı'stica}, (10), 43--60.
+
 \leavevmode\hypertarget{ref-smith2013effect}{}%
 Smith, N. J., \& Levy, R. (2013). The effect of word predictability on reading time is logarithmic. \emph{Cognition}, \emph{128}(3), 302--319.
+
+\leavevmode\hypertarget{ref-swadesh1955towards}{}%
+Swadesh, M. (1955). Towards greater accuracy in lexicostatistic dating. \emph{International Journal of American Linguistics}, \emph{21}(2), 121--137.
+
+\leavevmode\hypertarget{ref-team2013r}{}%
+Team, R. C., \& others. (2013). \emph{R: A language and environment for statistical computing}.
+
+\leavevmode\hypertarget{ref-wagner1985much}{}%
+Wagner, K. R. (1985). How much do children say in a day? \emph{Journal of Child Language}, \emph{12}(2), 475--487.
+
+\leavevmode\hypertarget{ref-wichmann2016asjp}{}%
+Wichmann, S., Müller, A., Wett, A., Velupillai, V., Bischoffberger, J., Brown, C. H., \ldots{} others. (2016). The asjp database. \emph{Max Planck Institute for the Science of Human History, Jena}.
 

### 2019-09-12 Paper/paper.Rmd
bootstrap childes surprisal results

@@
 library(ggdendro)
+library(gridExtra)
 
@@
 ```{r load data, include = F}
-childes_results <- read_feather(here("../surprisal/Data/childes_words.feather"))
+childes_results <- read_feather(here("Data/all_childes_bootstrap.feather"))
 wiki_results <- read_feather(here("../surprisal/Data/all_spaces.feather"))
-switchboard <- read_feather(here("../surprisal/Data/switchboard_words.feather"))
+switch_results <- read_feather(here("Data/switch_bootstrap.feather"))
 wiki_curves <- read_csv(here("../surprisal/Data/relative_ngrams.csv"))
+wiki_lr <- read_feather(here("../surprisal/Wikipedia/models.feather"))
 ```
@@
 
+```{r, echo=F, fig.height=2, fig.cap="Switchboard frequency-based information curves"}
+switch_results %>% 
+  filter(context == 0) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = 1:11) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
+
 We have found a unique average distribution of information that appears to characterize the English language as a whole, regardless of medium (spoken or written) or utterance length.
-This distribution indicates that in English, the words we speak or write at the beginnings of utterances has little information, while the words we speak or write at the ends of utterances has a lot of information.
+This distribution indicates that in English, the words we speak or write at the beginnings of utterances have little information, while the words we speak or write at the ends of utterances have a lot of information.
 The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from [@genzel2002entropy].
@@
 
-```{r switchboard & bnc plot}
+```{r, echo=F,fig.height=2,fig.cap="Switchboard bigram context-based information curves"}
+switch_results %>% 
+  filter(context == 1) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = 1:11) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
 
+```{r, echo=F,fig.height=2,fig.cap="Switchboard trigram context-based information curves"}
+switch_results %>% 
+  filter(context == 2) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = 1:11) + 
+    xlab("Word position") + 
+    ylab("Mean information")
 ```
@@
 childes_results %>% 
-  filter(Language == "English", context == 0, length %in% c(5, 7, 9)) %>%
-  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
+  filter(language == "English", context == 0) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
+             color = speaker)) + 
     geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
     facet_wrap(~length) + 
     scale_x_continuous(breaks = 1:9) + 
-    geom_line() + 
     xlab("Word position") + 
@@
 childes_results %>% 
-  filter(Language == "English", context == 1, length %in% c(5, 7, 9)) %>%
-  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
+  filter(language == "English", context == 1) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
+             color = speaker)) + 
     geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
     facet_wrap(~length) + 
     scale_x_continuous(breaks = 1:9) + 
-    geom_line() + 
     xlab("Word position") + 
@@
 childes_results %>% 
-  filter(Language == "English", context == 2, length %in% c(5, 7, 9)) %>%
-  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
+  filter(language == "English", context == 2) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
+             color = speaker)) + 
     geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
     facet_wrap(~length) + 
     scale_x_continuous(breaks = 1:9) + 
-    geom_line() + 
     xlab("Word position") + 
@@
   filter(context == 0, length == 7) %>%
-  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
+             color = speaker)) + 
     geom_point() + 
-    facet_wrap(~Language) + 
-    scale_x_continuous(breaks = 1:9) + 
+    geom_pointrange() + 
     geom_line() +
+    facet_wrap(~language) + 
+    scale_x_continuous(breaks = 1:7) + 
     xlab("Word position") + 
@@
   filter(context == 1, length == 7) %>%
-  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
+             color = speaker)) + 
     geom_point() + 
-    facet_wrap(~Language) + 
-    scale_x_continuous(breaks = 1:9) + 
+    geom_pointrange() + 
     geom_line() +
+    facet_wrap(~language) + 
+    scale_x_continuous(breaks = 1:7) + 
     xlab("Word position") + 
@@
   filter(context == 2, length == 7) %>%
-  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
+             color = speaker)) + 
     geom_point() + 
-    facet_wrap(~Language) + 
-    scale_x_continuous(breaks = 1:9) + 
+    geom_pointrange() + 
     geom_line() +
+    facet_wrap(~language) + 
+    scale_x_continuous(breaks = 1:7) + 
     xlab("Word position") + 
@@
 
-Some lr results go here.
+```{r, echo=F, fig.cap="Some linear model results from Wikipedia and WALS features"}
+sf <- wiki_lr %>% 
+  filter(p.value == 0, estimate > 0) %>% 
+  select(name) %>% 
+  rename("Some Important Features" = name) %>% 
+  slice(c(1, 10, 12, 15, 23, 29))
+
+sf$Examples <- c("Noun-RC; RC-noun", 
+                 "No cases; 6 cases", 
+                 "SOV; VSO", 
+                 "Only singular; sing. and plural", 
+                 "Affix; distinct word", 
+                 "Prefixes; suffixes")
+
+sf %>% grid.table()
+```
 

### 2019-09-12 Paper/paper.tex
bootstrap childes surprisal results

@@
   \postdate{\par}
-    \date{2019-09-11}
+    \date{2019-09-12}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
+\caption{\label{fig:unnamed-chunk-1}Switchboard frequency-based information curves}
+\end{figure}
+
 We have found a unique average distribution of information that appears to characterize the English language as a whole, regardless of medium (spoken or written) or utterance length.
-This distribution indicates that in English, the words we speak or write at the beginnings of utterances has little information, while the words we speak or write at the ends of utterances has a lot of information.
+This distribution indicates that in English, the words we speak or write at the beginnings of utterances have little information, while the words we speak or write at the ends of utterances have a lot of information.
 The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from (Genzel \& Charniak, 2002).
@@
 After the first word or two, where the listener does not have access to prior context, then they decode information at a flat and more or less uniform rate.
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
+\caption{\label{fig:unnamed-chunk-2}Switchboard bigram context-based information curves}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
+\caption{\label{fig:unnamed-chunk-3}Switchboard trigram context-based information curves}
+\end{figure}
 
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
-\caption{\label{fig:unnamed-chunk-1}Providence frequency-based information curves}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-4-1.pdf}
+\caption{\label{fig:unnamed-chunk-4}Providence frequency-based information curves}
 \end{figure}
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
-\caption{\label{fig:unnamed-chunk-2}Providence bigram context-based information curves}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1.pdf}
+\caption{\label{fig:unnamed-chunk-5}Providence bigram context-based information curves}
 \end{figure}
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
-\caption{\label{fig:unnamed-chunk-3}Providence trigram context-based information curves}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}
+\caption{\label{fig:unnamed-chunk-6}Providence trigram context-based information curves}
 \end{figure}
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-4-1.pdf}
-\caption{\label{fig:unnamed-chunk-4}CHILDES frequency-based information curves}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}
+\caption{\label{fig:unnamed-chunk-7}CHILDES frequency-based information curves}
 \end{figure}
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1.pdf}
-\caption{\label{fig:unnamed-chunk-5}CHILDES bigram context-based information curves}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
+\caption{\label{fig:unnamed-chunk-8}CHILDES bigram context-based information curves}
 \end{figure}
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}
-\caption{\label{fig:unnamed-chunk-6}CHILDES trigram context-based information curves}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1.pdf}
+\caption{\label{fig:unnamed-chunk-9}CHILDES trigram context-based information curves}
 \end{figure}
@@
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1.pdf}
 
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
-\caption{\label{fig:unnamed-chunk-8}Some trigram information curves from the Wikipedia data}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-11-1.pdf}
+\caption{\label{fig:unnamed-chunk-11}Some trigram information curves from the Wikipedia data}
 \end{figure}
@@
 
-Some lr results go here.
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/unnamed-chunk-12-1.pdf}
+\caption{\label{fig:unnamed-chunk-12}Some linear model results from Wikipedia and WALS features}
+\end{figure}
 
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1.pdf}
-\caption{\label{fig:unnamed-chunk-9}wals features vs cosine similarity}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-13-1.pdf}
+\caption{\label{fig:unnamed-chunk-13}wals features vs cosine similarity}
 \end{figure}
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1.pdf}
-\caption{\label{fig:unnamed-chunk-10}ldn features vs cosine similarity}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-14-1.pdf}
+\caption{\label{fig:unnamed-chunk-14}ldn features vs cosine similarity}
 \end{figure}

### 2019-09-12 surprisal/bnc&switchboard.Rmd
paper citations
bootstrap childes surprisal results

@@
 
-# 
-# sw <- switchboard %>%
-#   left_join(switch_unigrams) %>%
-#   mutate(s = -log(p)) %>%
-#   ungroup() %>%
-#   select(length, word_order, s) %>%
-#   nest(s)
-# 
-# swc <- sw %>% filter(length %in% c(10, 20, 30))
-# switch_surprisals <- map_dfr(1:nrow(swc), 
-#         ~bind_cols(select(slice(swc, .x), length, word_order), 
-#                                tidyboot_mean(data = unnest(slice(swc, .x)), 
-#                                              column = s, nboot = 1000)))
+sw <- switchboard %>%
+  left_join(switch_unigrams) %>%
+  mutate(s = -log(p)) %>%
+  ungroup() %>%
+  select(length, word_order, s) %>%
+  nest(s) %>%
+  filter(length %in% c(7, 9, 11))
+
+switch_surprisals <- map_dfr(1:nrow(sw),
+        ~bind_cols(select(slice(sw, .x), length, word_order),
+                               tidyboot_mean(data = unnest(slice(sw, .x)),
+                                             column = s, nboot = 1000)))
 
@@
   mutate(lag_word = lag(word)) %>%
-  left_join(switch_bigrams) %>%
-  left_join(switch_unigrams)
+  ungroup() %>%
+  left_join(switch_bigrams, by = c("word", "lag_word")) %>%
+  left_join(switch_unigrams, by = "word")
 
-switch_bigram_surprisals <- bigram_surprisals_prep %>%
+sbs <- bigram_surprisals_prep %>%
   mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
-  group_by(length, word_order) %>%
-  summarise(s = mean(s)) %>% 
-  filter(word_order <= length)
+  ungroup() %>%
+  select(length, word_order, s) %>%
+  nest(s) %>%
+  filter(length %in% c(7, 9, 11))
+
+switch_bigram_surprisals <- map_dfr(1:nrow(sbs),
+        ~bind_cols(select(slice(sbs, .x), length, word_order),
+                               tidyboot_mean(data = unnest(slice(sbs, .x)),
+                                             column = s, nboot = 1000)))
+
+# switch_bigram_surprisals <- bigram_surprisals_prep %>%
+#   mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
+#   group_by(length, word_order) %>%
+#   summarise(s = mean(s)) %>%
+#   filter(word_order <= length)
 
@@
   left_join(switch_bigrams, by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
-  left_join(switch_unigrams, by = c("word" = "word"))
+  left_join(switch_unigrams, by = c("word"))
 
-switch_trigram_surprisals <- trigram_surprisals_prep %>%
+sts <- trigram_surprisals_prep %>%
   mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
@@
                     -log(tri_cond_p))) %>% # unigram if first word
-  group_by(length, word_order) %>%
-  summarise(s = mean(s)) %>% 
-  filter(word_order <= length)
+  ungroup() %>%
+  select(length, word_order, s) %>%
+  nest(s) %>%
+  filter(length %in% c(7, 9, 11))
 
+switch_trigram_surprisals <- map_dfr(1:nrow(sts),
+        ~bind_cols(select(slice(sts, .x), length, word_order),
+                               tidyboot_mean(data = unnest(slice(sts, .x)),
+                                             column = s, nboot = 1000)))
+# switch_trigram_surprisals <- trigram_surprisals_prep %>%
+#   mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
+#                     ifelse(is.na(lag_word1), # check if not the second word
+#                            -log(p), # trigram if third or beyond
+#                            -log(cond_p)), # bigram if second word
+#                     -log(tri_cond_p))) %>% # unigram if first word
+#   group_by(length, word_order) %>%
+#   summarise(s = mean(s)) %>% 
+#   filter(word_order <= length)
+
+```
+
+```{r saving switchboard surprisals}
+switch_surprisals %>% 
+  mutate(context = 0, corpus = "Switchboard") %>%
+  bind_rows(switch_bigram_surprisals %>% mutate(context = 1, corpus = "Switchboard")) %>% 
+  bind_rows(switch_trigram_surprisals %>% mutate(context = 2, corpus = "Switchboard")) %>% 
+  write_feather(here("Data/switch_bootstrap.feather"))
 ```

### 2019-09-12 surprisal/childes_surprisal.Rmd
bootstrap childes surprisal results

@@
 zhou_utterances <- get_utterances(corpus = "ZhouDinner") #stem for pinyin
-okayama_utterances <- get_utterances(corpus = "Okayama")
+okayama_utterances <- get_utterances(corpus = "Okayama") # gloss
+shiro_utterances <- get_utterances(corpus = "Shiro") #gloss
+palasis_utteranecs <- get_utterances(corpus = "Palasis") #stem
+wagner_utterances <- get_utterances(corpus = "Wagner") #gloss
 ```
@@
 
-Down here is the old code. You can run it, but you have to change the corpus name and whether you're using stem or gloss by hand.
-
-```{r hard-coded language unigrams}
-child_utterances <- prov_utterances %>%
-  filter(speaker_role == "Target_Child") %>% 
-  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
-  mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
-  group_by(utterance_id) %>%
-  mutate(word_order = 1:n())
-
-child_unigrams <- child_utterances %>%
-  group_by(word) %>%
-  count() %>%
-  ungroup() %>%
-  mutate(p = n / sum(n))
-
-child_surprisals <- child_utterances %>%
-  left_join(child_unigrams) %>%
-  mutate(s = -log(p)) %>%
-  group_by(length, word_order) %>%
-  summarise(mean = mean(s),
-            median = median(s),
-            se = sd(s)/sqrt(n()))
-
-adult_utterances <- prov_utterances %>%
-  filter(speaker_role != "Target_Child") %>% 
-  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
-  mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
-  group_by(utterance_id) %>%
-  mutate(word_order = 1:n())
-
-adult_unigrams <- adult_utterances %>%
-  group_by(word) %>%
-  count() %>%
-  ungroup() %>%
-  mutate(p = n / sum(n))
-
-adult_surprisals <- adult_utterances %>%
-  left_join(adult_unigrams) %>%
-  mutate(s = -log(p)) %>%
-  group_by(length, word_order) %>%
-  summarise(mean = mean(s),
-            se = sd(s)/sqrt(n()))
-
-```
-
 ```{r new bigrams}
@@
 
+Down here is the old code. You can run it, but you have to change the corpus name and whether you're using stem or gloss by hand.
+
+```{r hard-coded language unigrams}
+child_utterances <- palasis_utterances %>%
+  filter(speaker_role != "Target_Child") %>% 
+  mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
+  mutate(utterance_id = 1:n()) %>%
+  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
+  group_by(utterance_id) %>%
+  mutate(word_order = 1:n())
+
+child_unigrams <- child_utterances %>%
+  group_by(word) %>%
+  count() %>%
+  ungroup() %>%
+  mutate(p = n / sum(n))
+
+child <- child_utterances %>%
+  left_join(child_unigrams) %>%
+  mutate(s = -log(p)) %>%
+  ungroup() %>%
+  select(length, word_order, s) %>%
+  nest(s) %>%
+  filter(length %in% c(5, 7, 9))
+
+child_surprisals <- map_dfr(1:nrow(child),
+        ~bind_cols(select(slice(child, .x), length, word_order),
+                               tidyboot_mean(data = unnest(slice(child, .x)),
+                                             column = s, nboot = 1000)))
+
+```
+
 ```{r bigrams}
-adult_bigrams <- adult_utterances %>%
+child_bigrams <- child_utterances %>%
   group_by(utterance_id) %>%
@@
   ungroup() %>%
-  left_join(adult_unigrams, by = c("lag_word" = "word")) %>%
-  mutate(cond_p = n.x/n.y)
+  left_join(child_unigrams, by = c("lag_word" = "word")) %>%
+  mutate(cond_p = n.x/n.y) %>%
+  select(-n.y, -p)
 
-bigram_surprisals_prep <- adult_utterances %>%
+bigram_surprisals_prep <- child_utterances %>%
   group_by(utterance_id) %>%
   mutate(lag_word = lag(word)) %>%
-  left_join(adult_bigrams) %>%
-  left_join(adult_unigrams)
+  left_join(child_bigrams, by = c("lag_word", "word")) %>%
+  left_join(child_unigrams, by = "word")
 
-bigram_surprisals <- bigram_surprisals_prep %>%
+child <- bigram_surprisals_prep %>%
   mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
-  group_by(length, word_order) %>%
-  summarise(s = mean(s)) %>% 
-  filter(word_order <= length)
+  ungroup() %>%
+  select(length, word_order, s) %>%
+  nest(s) %>%
+  filter(length %in% c(5, 7, 9))
+
+child_bigram_surprisals <- map_dfr(1:nrow(child),
+        ~bind_cols(select(slice(child, .x), length, word_order),
+                               tidyboot_mean(data = unnest(slice(child, .x)),
+                                             column = s, nboot = 1000)))
 
@@
 ```{r trigrams}
-adult_trigrams <- adult_utterances %>%
+child_trigrams <- child_utterances %>%
   group_by(utterance_id) %>%
@@
   ungroup() %>%
-  left_join(adult_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
+  left_join(child_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
   mutate(tri_cond_p = n / n.x) %>%
-  select(-n, -n.x, -n.y, -p, -cond_p)
+  select(-n, -n.x, -cond_p)
 
-trigram_surprisals_prep <- adult_utterances %>%
+trigram_surprisals_prep <- child_utterances %>%
   group_by(utterance_id) %>%
@@
   mutate(lag_word2 = lag(lag_word1)) %>%
-  left_join(adult_trigrams) %>% 
-  left_join(select(adult_bigrams, lag_word, word, cond_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
-  left_join(adult_unigrams)
+  left_join(child_trigrams) %>% 
+  left_join(select(child_bigrams, lag_word, word, cond_p), by = c("lag_word1" = "lag_word", "word")) %>% 
+  left_join(child_unigrams, by = "word")
 
-trigram_surprisals <- trigram_surprisals_prep %>%
+child <- trigram_surprisals_prep %>%
   mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
@@
                     -log(tri_cond_p))) %>% # unigram if first word
-  group_by(length, word_order) %>%
-  summarise(s = mean(s)) %>% 
-  filter(word_order <= length)
+  ungroup() %>%
+  select(length, word_order, s) %>%
+  nest(s) %>%
+  filter(length %in% c(5, 7, 9))
+
+child_trigram_surprisals <- map_dfr(1:nrow(child),
+        ~bind_cols(select(slice(child, .x), length, word_order),
+                               tidyboot_mean(data = unnest(slice(child, .x)),
+                                             column = s, nboot = 1000)))
 ```
+
+```{r bootstrap data write out}
+surprisals <- surprisals %>% 
+  bind_rows(child_surprisals %>% mutate(context = 0) %>%
+  bind_rows(child_bigram_surprisals %>% mutate(context = 1)) %>% 
+  bind_rows(child_trigram_surprisals %>% mutate(context = 2)) %>% 
+  mutate(speaker = "Parent", language = "French"))
+surprisals %>% 
+  write_feather(here("Data/all_childes_bootstrap.feather"))
+```

### 2019-09-13 Paper/paper.Rmd
references and r^2s

@@
 library(tidyverse)
-library(ggdendro)
 library(here)
@@
 library(gridExtra)
+library(ggpmisc)
 
@@
 
-See in the figure below a qualitative comparison of the information trajectory they found and the information trajectory we found.
+See in the FIGURE below a qualitative comparison of the information trajectory they found and the information trajectory we found.
 It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as @yu2016distribution did.
@@
 The longer the utterance, the more pronounced the information distribution shape becomes.
-A qualitative comparison of the two trajectories is below.
+A qualitative comparison of the two trajectories is in Figure \@ref(fig:switchboard_unigrams).
 
-```{r, echo=F, fig.height=2, fig.cap="Switchboard frequency-based information curves"}
+```{r switchboard_unigrams, echo=F, fig.height=2, fig.cap="Switchboard frequency-based information curves"}
 switch_results %>% 
@@
 After the first word or two, where the listener does not have access to prior context, then they decode information at a flat and more or less uniform rate.
+The bigram and trigram information curves for Switchboard are in Figures \@ref(fig:switchboard_bigrams) and \@ref(fig:switchboard_trigrams).
 
-```{r, echo=F,fig.height=2,fig.cap="Switchboard bigram context-based information curves"}
+```{r switchboard_bigrams, echo=F,fig.height=2,fig.cap="Switchboard bigram context-based information curves"}
 switch_results %>% 
@@
 
-```{r, echo=F,fig.height=2,fig.cap="Switchboard trigram context-based information curves"}
+```{r switchboard_trigrams, echo=F,fig.height=2,fig.cap="Switchboard trigram context-based information curves"}
 switch_results %>% 
@@
 The distribution of information we found at the level of individual words in English, therefore, characterizes the English language as a whole and not only adult utterances, not only written utterances.
+See \@ref(fig:providence_unigrams).
 
-```{r, echo=F, fig.height =2, fig.cap="Providence frequency-based information curves"}
+```{r providence_unigrams, echo=F, fig.height =2, fig.cap="Providence frequency-based information curves"}
 childes_results %>% 
@@
 This applies to parents and children speaking to one another, as well as adults speaking and writing to one another.
+See \@ref(fig:providence_bigrams) and \@ref(fig:providence_trigrams).
 
-```{r, echo=F, fig.height =2, fig.cap="Providence bigram context-based information curves"}
+```{r providence_bigrams, echo=F, fig.height =2, fig.cap="Providence bigram context-based information curves"}
 childes_results %>% 
@@
 
-```{r, echo=F, fig.height =2, fig.cap="Providence trigram context-based information curves"}
+```{r providence_trigrams, echo=F, fig.height =2, fig.cap="Providence trigram context-based information curves"}
 childes_results %>% 
@@
 The complete linkage algorithm at every step pairs each language or cluster of languages with its closest neighboring language or cluster.
-A sample from the dendrogram is shown below, and the full tree can be constructed using the code in our GitHub repository.
+A sample from the dendrogram is shown in Figure \@ref(fig:dendro), and the full tree can be constructed using the code in our GitHub repository.
 The language family-like structure can be seen at first glance, although the dendrogram does not exactly replicate language genealogy for all 159 languages.
@@
 
-```{r, echo=F}
+```{r dendro, echo=F}
 wiki_curves %>%
@@
 
-A sample of the trigram information curves are plotted below, and all trigram information curves for the languages we used follow the same pattern.
+A sample of the trigram information curves are plotted in Figure \@ref(fig:wiki_trigrams), and all trigram information curves for the languages we used follow the same pattern.
 The first few words in utterances for each language are surprising, but with even a couple words of predictive context for each word, the amount of information in each word drops off and speakers produce information at a constant, optimal rate.
 
-```{r, echo=F, fig.cap="Some trigram information curves from the Wikipedia data"}
+```{r wiki_trigrams, echo=F, fig.cap="Some trigram information curves from the Wikipedia data"}
 wiki_curves %>%
@@
 We found that 100 out of the 120 features from WALS we examined was statistically significant (p < .001) in determining whether two languages had the same shape to their unigram information distributions.
-The results for some important features are below, and the rest of the results are in the Appendices.
+The results for some important features are in Figure \@ref(fig:linear_models), and the rest of the results are in the Appendices.
 
-```{r, echo=F, fig.cap="Some linear model results from Wikipedia and WALS features"}
+```{r linear_models, echo=F, fig.cap="Some linear model results from Wikipedia and WALS features"}
 sf <- wiki_lr %>% 
@@
 We next compared how the cosine distance between two languages related to how many WALS features they had in common.
-There is a small correlation with a very small R^2 value between WALS features similarity and cosine similarity between two languages, but a statistically significant slope (p < .001).
-Essentially, for every feature two languages have in common, they have one degree more cosine similarity in the slope space.
-As feature values tend to be highly correlated within categories (e.g.
-the word order features are highly correlated), then this translates into language having much more similar information curves if they have more WALS features in common, an intuitive result.
+$r^2$ value is $.005734$, which suggests that in aggregate there is not a correlation between how many WALS features languages have in common and the similarity of their frequency-based information curves.
+Figure \@ref(fig:cosine_wals) displays the results.
 
-```{r, echo = F, fig.cap="wals features vs cosine similarity"}
+```{r cosine_wals, echo = F, fig.cap="wals features vs cosine similarity"}
+r2 <- summary(lm(data = wiki_results, formula = unigram_cosine~feature))$r.squared %>%
+  round(digits = 3)
 wiki_results %>% 
   ggplot(aes(x = feature, y=unigram_cosine)) + 
-    geom_smooth(aes(weight = .01), se=F, method = "lm") + 
-    geom_hex()
+    geom_smooth(aes(weight = .01), se=F, method = "lm", color = "black") + 
+    geom_jitter(alpha = 0.04, color = "blue") + 
+    annotate("text", x = 70, y = -0.5, label = paste0("r^2 = ", r2)) + 
+    xlab("Number of identical WALS features") + 
+    ylab("Information curve similarity")
+    #stat_poly_eq(formula = y~x, position = "dodge")
 ```
@@
 For lexical features, we see a stronger correlation between the similarity of two languages in terms of their average LDN and the cosine distance between their information curves.
+Figure \@ref(fig:cosine_ldn).
 
-```{r, echo = F, fig.cap="ldn features vs cosine similarity"}
+```{r cosine_ldn, echo = F, fig.cap="ldn features vs cosine similarity"}
+r2 <- summary(lm(data = wiki_results, formula = unigram_cosine~ldn))$r.squared %>%
+  round(digits = 4)
 wiki_results %>% 
   ggplot(aes(x = ldn, y=unigram_cosine)) + 
-    geom_smooth(aes(weight = .01), se=F, method = "lm") + 
-    geom_hex()
+    geom_smooth(aes(weight = .01), se=F, method = "lm", color = "black") + 
+    geom_jitter(alpha = 0.04, color = "blue") + 
+    annotate("text", x = 0.2, y = -0.5, label = paste0("r^2 = ", r2)) + 
+    xlab("Normalized lexical distance") + 
+    ylab("Information curve similarity")
+    #stat_poly_eq(formula = y~x, position = "dodge")
 ```

### 2019-09-13 Paper/paper.tex
references and r^2s

--- 2019-09-11 Paper/paper.tex
@@
   \postdate{\par}
-    \date{2019-09-11}
+    \date{2019-09-13}
 
@@
 
-See in the figure below a qualitative comparison of the information trajectory they found and the information trajectory we found.
+See in the FIGURE below a qualitative comparison of the information trajectory they found and the information trajectory we found.
 It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as Yu et al.
@@
 The longer the utterance, the more pronounced the information distribution shape becomes.
-A qualitative comparison of the two trajectories is below.
+A qualitative comparison of the two trajectories is in Figure @ref(fig:switchboard\_unigrams).
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/switchboard_unigrams-1.pdf}
+\caption{(\#fig:switchboard\_unigrams)Switchboard frequency-based information curves}
+\end{figure}
 
 We have found a unique average distribution of information that appears to characterize the English language as a whole, regardless of medium (spoken or written) or utterance length.
-This distribution indicates that in English, the words we speak or write at the beginnings of utterances has little information, while the words we speak or write at the ends of utterances has a lot of information.
+This distribution indicates that in English, the words we speak or write at the beginnings of utterances have little information, while the words we speak or write at the ends of utterances have a lot of information.
 The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from (Genzel \& Charniak, 2002).
@@
 After the first word or two, where the listener does not have access to prior context, then they decode information at a flat and more or less uniform rate.
+The bigram and trigram information curves for Switchboard are in Figures @ref(fig:switchboard\_bigrams) and @ref(fig:switchboard\_trigrams).
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/switchboard_bigrams-1.pdf}
+\caption{(\#fig:switchboard\_bigrams)Switchboard bigram context-based information curves}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/switchboard_trigrams-1.pdf}
+\caption{(\#fig:switchboard\_trigrams)Switchboard trigram context-based information curves}
+\end{figure}
 
@@
 The distribution of information we found at the level of individual words in English, therefore, characterizes the English language as a whole and not only adult utterances, not only written utterances.
+See @ref(fig:providence\_unigrams).
 
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
-\caption{\label{fig:unnamed-chunk-1}Providence frequency-based information curves}
+\includegraphics{paper_files/figure-latex/providence_unigrams-1.pdf}
+\caption{(\#fig:providence\_unigrams)Providence frequency-based information curves}
 \end{figure}
@@
 This applies to parents and children speaking to one another, as well as adults speaking and writing to one another.
+See @ref(fig:providence\_bigrams) and @ref(fig:providence\_trigrams).
 
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
-\caption{\label{fig:unnamed-chunk-2}Providence bigram context-based information curves}
+\includegraphics{paper_files/figure-latex/providence_bigrams-1.pdf}
+\caption{(\#fig:providence\_bigrams)Providence bigram context-based information curves}
 \end{figure}
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
-\caption{\label{fig:unnamed-chunk-3}Providence trigram context-based information curves}
+\includegraphics{paper_files/figure-latex/providence_trigrams-1.pdf}
+\caption{(\#fig:providence\_trigrams)Providence trigram context-based information curves}
 \end{figure}
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-4-1.pdf}
-\caption{\label{fig:unnamed-chunk-4}CHILDES frequency-based information curves}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
+\caption{\label{fig:unnamed-chunk-1}CHILDES frequency-based information curves}
 \end{figure}
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1.pdf}
-\caption{\label{fig:unnamed-chunk-5}CHILDES bigram context-based information curves}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
+\caption{\label{fig:unnamed-chunk-2}CHILDES bigram context-based information curves}
 \end{figure}
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}
-\caption{\label{fig:unnamed-chunk-6}CHILDES trigram context-based information curves}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
+\caption{\label{fig:unnamed-chunk-3}CHILDES trigram context-based information curves}
 \end{figure}
@@
 The complete linkage algorithm at every step pairs each language or cluster of languages with its closest neighboring language or cluster.
-A sample from the dendrogram is shown below, and the full tree can be constructed using the code in our GitHub repository.
+A sample from the dendrogram is shown in Figure \ref{fig:dendro}, and the full tree can be constructed using the code in our GitHub repository.
 The language family-like structure can be seen at first glance, although the dendrogram does not exactly replicate language genealogy for all 159 languages.
@@
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}
+\includegraphics{paper_files/figure-latex/dendro-1.pdf}
 
-A sample of the trigram information curves are plotted below, and all trigram information curves for the languages we used follow the same pattern.
+A sample of the trigram information curves are plotted in Figure @ref(fig:wiki\_trigrams), and all trigram information curves for the languages we used follow the same pattern.
 The first few words in utterances for each language are surprising, but with even a couple words of predictive context for each word, the amount of information in each word drops off and speakers produce information at a constant, optimal rate.
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
-\caption{\label{fig:unnamed-chunk-8}Some trigram information curves from the Wikipedia data}
+\includegraphics{paper_files/figure-latex/wiki_trigrams-1.pdf}
+\caption{(\#fig:wiki\_trigrams)Some trigram information curves from the Wikipedia data}
 \end{figure}
@@
 We found that 100 out of the 120 features from WALS we examined was statistically significant (p \textless{} .001) in determining whether two languages had the same shape to their unigram information distributions.
-The results for some important features are below, and the rest of the results are in the Appendices.
-
-Some lr results go here.
-
-We next compared how the cosine distance between two languages related to how many WALS features they had in common.
-There is a small correlation with a very small R\^{}2 value between WALS features similarity and cosine similarity between two languages, but a statistically significant slope (p \textless{} .001).
-Essentially, for every feature two languages have in common, they have one degree more cosine similarity in the slope space.
-As feature values tend to be highly correlated within categories (e.g.~the word order features are highly correlated), then this translates into language having much more similar information curves if they have more WALS features in common, an intuitive result.
+The results for some important features are in Figure @ref(fig:linear\_models), and the rest of the results are in the Appendices.
 
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1.pdf}
-\caption{\label{fig:unnamed-chunk-9}wals features vs cosine similarity}
+\includegraphics{paper_files/figure-latex/linear_models-1.pdf}
+\caption{(\#fig:linear\_models)Some linear model results from Wikipedia and WALS features}
+\end{figure}
+
+We next compared how the cosine distance between two languages related to how many WALS features they had in common.
+\(r^2\) value is \(.005734\), which suggests that in aggregate there is not a correlation between how many WALS features languages have in common and the similarity of their frequency-based information curves.
+Figure @ref(fig:cosine\_wals) displays the results.
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/cosine_wals-1.pdf}
+\caption{(\#fig:cosine\_wals)wals features vs cosine similarity}
 \end{figure}
@@
 For lexical features, we see a stronger correlation between the similarity of two languages in terms of their average LDN and the cosine distance between their information curves.
+Figure @ref(fig:cosine\_ldn).
 
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1.pdf}
-\caption{\label{fig:unnamed-chunk-10}ldn features vs cosine similarity}
+\includegraphics{paper_files/figure-latex/cosine_ldn-1.pdf}
+\caption{(\#fig:cosine\_ldn)ldn features vs cosine similarity}
 \end{figure}

### 2019-09-16 Paper/paper.Rmd
bnc with nonpara bootstrap

@@
 We use language to communicate information.
-Although we often need to say things we have never said before, we are easily and efficiently able to produce novel utterances with minimal prior planning.
-Our listeners incrementally process our utterances, and predict which words will come next within fractions of a second of hearing the beginning of the word (for evidence see e.g.
+Although we often need to say things we have never said before, we are easily and efficiently able to produce novel utterances with next to no prior planning.
+Our listeners incrementally process each word we say, and predict which words will come next within fractions of a second of hearing the beginning of the word (for evidence see e.g.
 Kutas & Federmeier, 2010).
@@
 
-However, a gap exists in the literature for examining how frequency and context determine the information content in utterances.
-We do not speak only in phrases or paragraphs, but conversationally we communicate in utterances, either on their own or banded together.
-Writers use sentences to communicate thoughts, often on their own instead in paragraphs with other sentences.
-Cross-linguistically, the average rate of information speakers transfer in a syllable differs [@aylett2004smooth].
+Efficiency plays a big role in communication, from the way that languages evolve to the individual utterance choice timescale. @mahowald2013info. @gibson2017color. @piantadosi2011word.
+
+How do frequency and predictive context determine how we distribution information at the level of utterances?
+We do not speak only in phrases or paragraphs, but conversationally we speak in utterances and in writing we produce sentences as a coherent unit of a complete thought.
 Languages possess very different grammatical and morphological structures--the same thought can be encoded in at least as many ways as there are different languages.
+The average rate of information speakers transfer in syllables differs from language to language [@aylett2004smooth].
 How do the different structures seen in different languages, with different rates of information encoding, affect how speakers structure information in production?
@@
 @yu2016distribution proposed a novel method for quantifying the amount of information typically found at each part of English utterances.
-The Yu et al work examined written sentences from the British National Corpus.
-Treating each word position by itself, Yu et al.
-calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty [@shannon1948mathematical].
-Yu et al.
-found a distinctive three-step shape to the distribution of information in the British National Corpus.
+Treating each word position by itself, @yu2016distribution calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty [@shannon1948mathematical].
+They found a distinctive three-step shape to the distribution of information in English written sentences.
 
@@
 We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech.
+We also apply a contextual version of the same metric, and find that in communication, listeners are able to decode information from the utterances they hear at a constant rate past the initial words.
 We expand the study of information density to the largest set of languages yet considered, and incorporate contextual and typological information.
@@
 As soon as a child can form multi-word utterances in a given language, they will produce utterances obeying the characteristic distribution of information in a given language.
-We also apply a contextual version of the same metric, and find that in communication, listeners are able to decode information from the utterances they hear at a constant rate, regardless of language, past the first couple of words.
 
@@
 ```{r switchboard_unigrams, echo=F, fig.height=2, fig.cap="Switchboard frequency-based information curves"}
-switch_results %>% 
+sw_plot <- switch_results %>% 
   filter(context == 0) %>%
@@
     ylab("Mean information")
+
+bnc_plot <- bnc_results %>% 
+  filter(context == 0) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = 1:30) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+
+grid.arrange(bnc_plot, sw_plot, nrow = 2)
 ```
@@
 
-The language we speak impacts how we encode information into our utterances, and how listeners decode our speech signals.
-However, despite the variety in communicative codes, all languages are characterized by predictability and a lack of suprisal in regular speech.
-In this paper, we developed a novel computational modeling method for examining the average information at a fine-grained level in utterances of different lengths in a variety of corpora and across a large number of languages.
-With only a couple of words of context for each word, cross-linguistic differences disappear and all languages are characterizes by low surprisal (low processing cost and information).
-
-The Uniform Information Density Hypothesis [@jaeger2007speakers] provides an optimal strategy for distributing information among the words in an utterance.
-The hypothesis suggests that speakers tend to distribute information evenly in their utterances, close to the capacity of the communication channel.
-This strategy avoids peaks in information which lead to issues for the decoder, while making the best use of the communication channel possible and aiming to produce as much information as the listener can decode at every moment.
+By considering the distribution of information at the level of utterances and sentences, we join together the information-theoretic work focusing on sub-word units and words, and that focusing on paragraphs.
+In doing so, we show that frequency and context-based metrics complement one another in studying efficiency and information in language.
+We directly link linguistic efficiency in a language to the genealogy and properties of that language, while providing evidence for a novel, communicatively real linguistic universal.
+Also considering developmental angle, observing language-specific information distributions arising as soon as children begin speaking in multi-word utterances.
 

### 2019-09-17 Paper/paper.Rmd
full draft of paper ready for other eyes

@@
 wiki_results <- read_feather(here("../surprisal/Data/all_spaces.feather"))
+bnc_results <- read_feather(here("Data/bnc_bootstrap.feather"))
 switch_results <- read_feather(here("Data/switch_bootstrap.feather"))
@@
 
-Language can be thought of as a code: speakers encode their thoughts into a sequence of linguistic signals, namely an utterance, which listeners then must decode in order to recover the intended message.
-How do speakers structure their utterances so that communication is successful?
-Using tools from information theory and sentence processing, we find that speakers must structure the information in their utterances according to the grammatical, morphological and lexical constraints of their language, which vary greatly from language to language.
-This information structure characterizes language from children’s first utterances to large-scale knowledge base entries.
-Once listeners have access to linguistic context within an utterance, then they can decode the rest of that utterance at a constant and optimal rate.
+What role does communicative efficiency play in how we organize our utterances?
+In this paper, we present a novel method of examining how much information speakers in a given language communicate in each word in their utterances, surveying numerous diverse languages.
+We find that speakers produce frequent and informative words at regular parts of their utterances, depending on language they use, which is predictable in part from the features and genealogy of their language.
+This robust information distribution characterizes both spoken and written communication, and emerges in children's earliest utterances.
+However, in real-time communication, in-context word predictability allows listeners to process information at a constant, optimal rate, regardless of the information distribution in the language they understand.
 
@@
 We use language to communicate information.
-Although we often need to say things we have never said before, we are easily and efficiently able to produce novel utterances with next to no prior planning.
-Our listeners incrementally process each word we say, and predict which words will come next within fractions of a second of hearing the beginning of the word (for evidence see e.g.
-Kutas & Federmeier, 2010).
+Although we often need to say things we have never said before, we are easily and quickly able to produce novel utterances with next to no prior planning.
+Our listeners incrementally process each word we say in utterances they may have never heard before, and predict which words will come next within fractions of a second of hearing the beginning of the word (for evidence see e.g.
+@kutas2011thirty).
 Encoding new and unpredictable information into our utterances surprises our listeners, and presents problems for communication.
-Listeners struggle to decode unpredictable words and may have difficulty decoding the following words as well.
-Efficient and effective communication relies on interlocutors being able to process novel information in without being surprised by new information in unpredictable words.
+Listeners may struggle to decode unpredictable words, which can cause difficulty decoding the following words as well.
+Efficient and effective communication relies on interlocutors being able to process novel information without being surprised by new information in unpredictable words.
 On the other hand, to make best use of the capacity of the communication channel, speakers should encode as much information as possible into each word they say.
 
-Due to their differences in distributional frequency, words vary in how much information-theoretic content they contain [@shannon1948mathematical].
-This information content plays an important role in determining the shape of words and speech.
-Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be as speakers minimize effort in the amount of information they encode in speech [@zipf1949human].
-Over and above frequency, contextual information content predicts word length [@piantadosi2011word].
+Due to their differences in distributional frequency, words vary in how much information content they contain [@shannon1948mathematical].
+Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be [@zipf1949human].
+Over and above frequency, contextual information content predicts word length, due to speakers minimizing effort in the amount of information they encode in speech [@piantadosi2011word].
 Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side [@jaeger2007speakers].
@@
 
-Efficiency plays a big role in communication, from the way that languages evolve to the individual utterance choice timescale. @mahowald2013info. @gibson2017color. @piantadosi2011word.
+The pressure for efficient communication influences language at all timescales, from online speech production to evolution.
+@mahowald2013info found that speakers tend to use shorter, more efficient word forms in predictive contexts, minimizing communicative cost without sacrificing intelligibility.
+Predictability and ambiguity shape the distribution of color terms in languages [@gibson2017color] as well as word lengths [@piantadosi2011word] across languages, and speakers are able to deterministically decode words despite ambiguity in word forms [@piantadosi2012communicative].
 
@@
 We expand the study of information density to the largest set of languages yet considered, and incorporate contextual and typological information.
-We find that each language has a characteristic, robust distribution of information predictable from its morphology, phonology and syntax, and tied to the language’s genealogy.
+We find that each language has a characteristic, robust frequency-based information distribution tied to the language's morphology, phonology and syntax.
 Speakers of a language will tend to distribute information according to that language’s distribution of information in their utterances.
@@
 
+The flexibility of the surprisal metric we employ in this paper allows us to calculate the anticipated information for an individual utterance, as most work with the metric has done in the past.
+Averaging together the surprisal values for a word position within utterances is actually a step further than prior work, and indicates the tendencies speakers gravitate towards instead of examining individual stimuli in psycholinguistic experiments.
+
 The frequency-based surprisal metric gives us an idea of when in their utterances speakers say frequent i.e.
@@
 
-See in the FIGURE below a qualitative comparison of the information trajectory they found and the information trajectory we found.
+See in Figure \@ref(fig:bnc_unigrams) a qualitative comparison of the information trajectory they found and the information trajectory we found.
 It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as @yu2016distribution did.
@@
 
-Comparison between the unigram entropy and surprisal BNC results plot
+```{r bnc_unigrams, echo = F, fig.height=2, fig.cap="BNC frequency-based information curves"}
+bnc_results %>% 
+  filter(context == 0) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
 
@@
 
-```{r switchboard_unigrams, echo=F, fig.height=2, fig.cap="Switchboard frequency-based information curves"}
+```{r switchboard_unigrams, echo=F, fig.height=4, fig.cap="Switchboard frequency-based information curves"}
 sw_plot <- switch_results %>% 
@@
     facet_wrap(~length) + 
-    scale_x_continuous(breaks = 1:30) + 
+    scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
     xlab("Word position") + 
@@
 What about context?
-So far we’ve only discussed the unigram (“contextless”) surprisal metric, considering words on their own without any explicit incorporation of prior context.
+So far we’ve only discussed the frequency-based metric, considering words on their own without any explicit incorporation of prior context.
 As previously discussed, listeners decode information and process what they hear incrementally, using prior heard words to ease the comprehension process.
-We now include one word of context (bigrams) and then two words of context (trigrams).
+We now include one word of context (bigrams) and then two words of context (trigrams) for each word in our measurements.
 We observe a flattening effect of context across both modalities and all speaker populations.
 After the first word or two, where the listener does not have access to prior context, then they decode information at a flat and more or less uniform rate.
-The bigram and trigram information curves for Switchboard are in Figures \@ref(fig:switchboard_bigrams) and \@ref(fig:switchboard_trigrams).
+The contextual information curves for the BNC and Switchboard are in Figures \@ref(fig:switchboard_bigrams) and \@ref(fig:switchboard_trigrams).
 
-```{r switchboard_bigrams, echo=F,fig.height=2,fig.cap="Switchboard bigram context-based information curves"}
-switch_results %>% 
+```{r switchboard_bigrams, echo=F,fig.height=4,fig.cap="Switchboard bigram context-based information curves"}
+sw_plot <- switch_results %>% 
   filter(context == 1) %>%
@@
     ylab("Mean information")
+
+bnc_plot <- bnc_results %>% 
+  filter(context == 1) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+
+grid.arrange(bnc_plot, sw_plot, nrow = 2)
 ```
 
-```{r switchboard_trigrams, echo=F,fig.height=2,fig.cap="Switchboard trigram context-based information curves"}
-switch_results %>% 
+```{r switchboard_trigrams, echo=F,fig.height=4,fig.cap="Switchboard trigram context-based information curves"}
+sw_plot <- switch_results %>% 
   filter(context == 2) %>%
@@
     ylab("Mean information")
+
+bnc_plot <- bnc_results %>% 
+  filter(context == 2) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+
+grid.arrange(bnc_plot, sw_plot, nrow = 2)
 ```
@@
 
-```{r providence_unigrams, echo=F, fig.height =2, fig.cap="Providence frequency-based information curves"}
+```{r providence_unigrams, echo=F, fig.height =2, fig.cap="Providence frequency-based information curves.
+Lines around each point indicate 95% confidence intervals computed with non-parametric bootstrap"}
 childes_results %>% 
@@
 Spanish features a larger spike in the amount of information in the final words of utterances.
-For Japanese and Mandarin, we observe completely different unigram information curve trajectories.
-The Japanese unigram information curve trajectory begins high and finishes low, the mirror image of the European language information curves.
+For Japanese and Mandarin, we observe completely different frequency-based information curve trajectories.
+The Japanese frequency-based information curve trajectory begins high and finishes low, the mirror image of the European language information curves.
 The Mandarin curve begins low and finishes low, but features high information in the middle of utterances.
@@
 For the bigram and trigram information curves, we see the same contextual smoothing effect as for all the corpora we worked with in English.
-While the unigram information curves may depend based on the language, the contextual information curves show the same trajectory cross-linguistically.
-Using ngrams with n higher than 3 is difficult for parent-child speech corpora because the utterances are so short on average (less than 10 words).
-We hypothesize that the unigram information curves may vary based on the genealogy and typology of the languages in question, but this does not extend to the trigram information curves in particular.
-We shall show evidence that this is true in the rest of the paper.
+While the frequency-based information curves may depend based on the language, the contextual information curves show the same trajectory cross-linguistically.
+Using more than two words of context is difficult for parent-child speech corpora because the utterances are so short on average (less than $10$ words).
+We hypothesize that the frequency-based information curves may vary based on the genealogy and typology of the languages in question, but this does not extend to the information curves with two words of context in particular.
 
@@
 To make a claim about language as a whole and languages on a larger scale, we needed to use larger corpora and a much larger number of languages.
-We pulled corpora for 159 diverse languages from Wikipedia, each of which had at least 10000 articles.
+We pulled corpora for $159$ diverse languages from Wikipedia, each of which had at least $10000$ articles on the knowledge base.
 We split each article up into sentences; the variance in sentence lengths for Wikipedia was significantly larger than for the CHILDES corpora we used in the previous section.
-Most sentences in Wikipedia contained between 10 and 30 words, unlike the CHILDES corpora which mainly contained utterances with under 10 words.
-We excluded the small fraction of utterances with more than 50 words as uncharacteristic of English speech.
+Most sentences in Wikipedia contained between $10$ and $30$ words, unlike the CHILDES corpora which mainly contained utterances with under 10 words.
+We excluded the small fraction of utterances with more than $50$ words since they were small in number and, from manual inspection, uncharacteristic of typical written sentences.
 
-How to analyze more than 40 different surprisal curves for each language, adding up to several thousand surprisal curves total?
+How do we analyze more than $40$ different surprisal curves for each language, adding up to several thousand surprisal curves total?
 We used two different strategies, which yielded identical results upon analysis.
@@
 We then similarly computed surprisal values at each of these positions, and similarly computed slopes between the surprisal values at each position, giving us another five slope values for each language summarizing the surprisal curves.
-Illustrations of these two strategies are below.
+Illustrations of these two strategies are in Figure \@ref(fig:treatments)
 
-Illustration of the regions we target with each of the two slope treatments below.
+```{r treatments, echo = F, fig.height =3, fig.cap="Illustration of slope treatments for Wikipedia information curves: relative on top and absolute on bottom"}
+## relative treatment: every fifth in a different group
+relative <- bnc_results %>% 
+  filter(context == 0, length == 20) 
+relative$group <- rep(1:5, each = 4)
+relative <- relative %>%
+  mutate(group = as.factor(group))
 
-We computed unigram, bigram and trigram values for each language.
-All of this data, along with the scripts used for the analysis, are available on GitHub.
-We ran this code on the RCC at the University of Chicago; it took X amount of time and used up Y amount of carbon to run.
+# rel_plot <- relative %>%
+#   ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, color = group)) + 
+#     geom_point() + 
+#     geom_pointrange() + 
+#     geom_line() +
+#     scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
+#     xlab("Word position") + 
+#     ylab("Mean information")
+
+## absolute treatment: some qualitatively important sections
+absolute <- bnc_results %>% 
+  filter(context == 0, length == 20) 
+absolute$group <- c(1, 2, rep(3, 16), 4, 5) 
+absolute <- absolute %>%
+  mutate(group = as.factor(group))
+
+relative %>% 
+  mutate(treatment = "relative") %>% 
+  bind_rows(absolute %>% mutate(treatment = "absolute")) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, 
+             ymin = ci_lower, ymax = ci_upper, color = group)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~treatment) + 
+    scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+
+# grid.arrange(rel_plot, abs_plot, nrow = 2)
+
+```
+
+We computed unigram, bigram and trigram values for each language, and grouped the contextual and frequency-based values for each treatment.
 
 To more rigorously described the differences between languages, we used data from the World Atlas of Language Structures [WALS; @wals].
-WALS has data for 144 typological features in 2569 languages around the world.
+WALS has data for $144$ typological features in $2569$ languages around the world.
 These features describe aspects of morphology, syntax, phonology, etymology and semantics--in short the features describe the structures in each language.
 As WALS is a compiled database from dozens of papers from different authors, most of the features and languages are fairly sparse.
-Even limiting ourselves to the 159 language corpora we pulled from Wikipedia and 122 features from WALS, there are nearly 20000 individual possible feature values, fewer than half of which were already inputted for those languages in the WALS database.
+Even limiting ourselves to the $159$ language corpora we pulled from Wikipedia and $122$ features from WALS, there are nearly $20000$ individual possible feature values, fewer than half of which were already inputted for those languages in the WALS database.
 
@@
 MIMCA essentially uses mean imputation to begin with the missing values, then performs repeatedly performs principle components analysis on and reconstructs the contingency table formed from observations in categorical variables.
-By the end of this we had unigram, bigram and trigram surprisal curves for the 159 language corpora pulled from Wikipedia, along with 122 typological features for each language.
+By the end of this we had frequency-based and contextual information curves for the $159$ language corpora pulled from Wikipedia, along with $122$ typological features for each language.
 
 However, the WALS features describe specific structural differences between languages, while our surprisal metric is lexical.
-To target lexical differences between languages, we computed the average normalized Levenshtein distance [LDN; @holman2008advances] over the 40 item Swadesh list [@swadesh1955towards], retrieved from the ASJP database [@wichmann2016asjp].
+To target lexical differences between languages, we computed the average normalized Levenshtein distance [LDN; @holman2008advances] over the $40$ item Swadesh list [@swadesh1955towards], retrieved from the ASJP database [@wichmann2016asjp].
 The Swadesh list is designed to include near-universal words that target basic cognitive concepts, and are useful in determining the genealogical similarities and differences between languages.
@@
 A sample from the dendrogram is shown in Figure \@ref(fig:dendro), and the full tree can be constructed using the code in our GitHub repository.
-The language family-like structure can be seen at first glance, although the dendrogram does not exactly replicate language genealogy for all 159 languages.
+The language family-like structure can be seen at first glance, although the dendrogram does not exactly replicate language genealogy for all $159$ languages.
 This suggests using a first-pass quantitative method that the information curves do correspond in some measure to language families, but language families do not explain the entire story.
@@
 
-A sample of the trigram information curves are plotted in Figure \@ref(fig:wiki_trigrams), and all trigram information curves for the languages we used follow the same pattern.
+A sample of the contextual information curves (two words of context) are plotted in Figure \@ref(fig:wiki_trigrams), and all trigram information curves for the languages we used follow the same pattern.
 The first few words in utterances for each language are surprising, but with even a couple words of predictive context for each word, the amount of information in each word drops off and speakers produce information at a constant, optimal rate.
@@
 Figure \@ref(fig:cosine_wals) displays the results.
+This result is surprising based on the significance of many WALS features in predicting the shapes of the frequency-based information curves, and we return to this result in the general discussion.
 
@@
 
-So the more similar two languages are in terms of typological features, the more similar their unigram information curves tend to be.
 For lexical features, we see a stronger correlation between the similarity of two languages in terms of their average LDN and the cosine distance between their information curves.
 Figure \@ref(fig:cosine_ldn).
+We see a higher $r^2$ value here of $.026$, indicating that there is more correspondence between a language's lexical similarity to another language and their similarity in information curves.
 
@@
 
-The shape of a language’s frequency-based information curve covaries with its typological and lexical similarity to other languages.
+From these typological investigations, we conclude that the shape of a language’s frequency-based information curve covaries with its typological and lexical similarity to other languages.
 
-# General Discussion
+# Discussion
 
@@
 In doing so, we show that frequency and context-based metrics complement one another in studying efficiency and information in language.
-We directly link linguistic efficiency in a language to the genealogy and properties of that language, while providing evidence for a novel, communicatively real linguistic universal.
-Also considering developmental angle, observing language-specific information distributions arising as soon as children begin speaking in multi-word utterances.
+We directly link linguistic efficiency in a language to the genealogy and properties of that language, We provide evidence for a novel linguistic universal: low processing cost for listeners beyond the first words in utterances, driven by high average word predictability in conversation.
+Also considering the developmental angle, observing language-specific information distributions arising as soon as children begin speaking in multi-word utterances.
+
+Throughout this work we have averaged the surprisal values at each position.
+Averaging removes variation, which in turn may obscure trends in the data.
+As discussed in the methods section, the surprisal metric has historically been used for calculating the information and processing cost for individual utterances, and our use of the metric here is actually a step forward rather than a step back.
+Future work can investigate variation in how speakers distribute information in individual utterances.
+
+The WALS database we used to investigate typological variation in the information curves is overall sparse.
+We imputed well over $50\%$ of the WALS features for most of our $159$ languages, although all of the languages had at least $20$ features evaluated in WALS.
+A large part of this is due to WALS being a collection of a number of different studies, instead of a systematic effort to catalogue variation across the world's languages.
+Additionally, WALS features are meant to describe specific microvariations in languages, not to provide a comprehensive typological representation of each language compared to each other language.
+This may be why the Swadesh list provided a higher correlation for describing the differences in information curves: @swadesh1955towards intended the list to allow researchers to more comprehensively compared and constrast lexical differences between languages.
+For our Wikipedia analysis, we also reduce all of a language's variation down to a five-dimensional vector.
+These information curve representations show a surprising amount of variation despite the degree of compression.
 

### 2019-09-17 Paper/paper.tex
full draft of paper ready for other eyes

@@
   \postdate{\par}
-    \date{2019-09-13}
+    \date{2019-09-17}
 
@@
 
-Language can be thought of as a code: speakers encode their thoughts into a sequence of linguistic signals, namely an utterance, which listeners then must decode in order to recover the intended message.
-How do speakers structure their utterances so that communication is successful?
-Using tools from information theory and sentence processing, we find that speakers must structure the information in their utterances according to the grammatical, morphological and lexical constraints of their language, which vary greatly from language to language.
-This information structure characterizes language from children's first utterances to large-scale knowledge base entries.
-Once listeners have access to linguistic context within an utterance, then they can decode the rest of that utterance at a constant and optimal rate.
+What role does communicative efficiency play in how we organize our utterances?
+In this paper, we present a novel method of examining how much information speakers in a given language communicate in each word in their utterances, surveying numerous diverse languages.
+We find that speakers produce frequent and informative words at regular parts of their utterances, depending on language they use, which is predictable in part from the features and genealogy of their language.
+This robust information distribution characterizes both spoken and written communication, and emerges in children's earliest utterances.
+However, in real-time communication, in-context word predictability allows listeners to process information at a constant, optimal rate, regardless of the information distribution in the language they understand.
 
@@
 We use language to communicate information.
-Although we often need to say things we have never said before, we are easily and efficiently able to produce novel utterances with minimal prior planning.
-Our listeners incrementally process our utterances, and predict which words will come next within fractions of a second of hearing the beginning of the word (for evidence see e.g.~Kutas \& Federmeier, 2010).
+Although we often need to say things we have never said before, we are easily and quickly able to produce novel utterances with next to no prior planning.
+Our listeners incrementally process each word we say in utterances they may have never heard before, and predict which words will come next within fractions of a second of hearing the beginning of the word (for evidence see e.g.
+Kutas \& Federmeier (2011)).
 Encoding new and unpredictable information into our utterances surprises our listeners, and presents problems for communication.
-Listeners struggle to decode unpredictable words and may have difficulty decoding the following words as well.
-Efficient and effective communication relies on interlocutors being able to process novel information in without being surprised by new information in unpredictable words.
+Listeners may struggle to decode unpredictable words, which can cause difficulty decoding the following words as well.
+Efficient and effective communication relies on interlocutors being able to process novel information without being surprised by new information in unpredictable words.
 On the other hand, to make best use of the capacity of the communication channel, speakers should encode as much information as possible into each word they say.
 
-Due to their differences in distributional frequency, words vary in how much information-theoretic content they contain (Shannon, 1948).
-This information content plays an important role in determining the shape of words and speech.
-Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be as speakers minimize effort in the amount of information they encode in speech (Zipf, 1949).
-Over and above frequency, contextual information content predicts word length (Piantadosi, Tily, \& Gibson, 2011).
+Due to their differences in distributional frequency, words vary in how much information content they contain (Shannon, 1948).
+Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be (Zipf, 1949).
+Over and above frequency, contextual information content predicts word length, due to speakers minimizing effort in the amount of information they encode in speech (Piantadosi, Tily, \& Gibson, 2011).
 Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side (Jaeger \& Levy, 2007).
@@
 
-However, a gap exists in the literature for examining how frequency and context determine the information content in utterances.
-We do not speak only in phrases or paragraphs, but conversationally we communicate in utterances, either on their own or banded together.
-Writers use sentences to communicate thoughts, often on their own instead in paragraphs with other sentences.
-Cross-linguistically, the average rate of information speakers transfer in a syllable differs (Aylett \& Turk, 2004).
+The pressure for efficient communication influences language at all timescales, from online speech production to evolution.
+Mahowald, Fedorenko, Piantadosi, \& Gibson (2013) found that speakers tend to use shorter, more efficient word forms in predictive contexts, minimizing communicative cost without sacrificing intelligibility.
+Predictability and ambiguity shape the distribution of color terms in languages (Gibson et al., 2017) as well as word lengths (Piantadosi et al., 2011) across languages, and speakers are able to deterministically decode words despite ambiguity in word forms (Piantadosi, Tily, \& Gibson, 2012).
+
+How do frequency and predictive context determine how we distribution information at the level of utterances?
+We do not speak only in phrases or paragraphs, but conversationally we speak in utterances and in writing we produce sentences as a coherent unit of a complete thought.
 Languages possess very different grammatical and morphological structures--the same thought can be encoded in at least as many ways as there are different languages.
+The average rate of information speakers transfer in syllables differs from language to language (Aylett \& Turk, 2004).
 How do the different structures seen in different languages, with different rates of information encoding, affect how speakers structure information in production?
@@
 Yu, Cong, Liang, \& Liu (2016) proposed a novel method for quantifying the amount of information typically found at each part of English utterances.
-The Yu et al work examined written sentences from the British National Corpus.
-Treating each word position by itself, Yu et al.~calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty (Shannon, 1948).
-Yu et al.~found a distinctive three-step shape to the distribution of information in the British National Corpus.
+Treating each word position by itself, Yu et al.
+(2016) calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty (Shannon, 1948).
+They found a distinctive three-step shape to the distribution of information in English written sentences.
 
@@
 We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech.
+We also apply a contextual version of the same metric, and find that in communication, listeners are able to decode information from the utterances they hear at a constant rate past the initial words.
 We expand the study of information density to the largest set of languages yet considered, and incorporate contextual and typological information.
-We find that each language has a characteristic, robust distribution of information predictable from its morphology, phonology and syntax, and tied to the language's genealogy.
+We find that each language has a characteristic, robust frequency-based information distribution tied to the language's morphology, phonology and syntax.
 Speakers of a language will tend to distribute information according to that language's distribution of information in their utterances.
 As soon as a child can form multi-word utterances in a given language, they will produce utterances obeying the characteristic distribution of information in a given language.
-We also apply a contextual version of the same metric, and find that in communication, listeners are able to decode information from the utterances they hear at a constant rate, regardless of language, past the first couple of words.
 
@@
 
+The flexibility of the surprisal metric we employ in this paper allows us to calculate the anticipated information for an individual utterance, as most work with the metric has done in the past.
+Averaging together the surprisal values for a word position within utterances is actually a step further than prior work, and indicates the tendencies speakers gravitate towards instead of examining individual stimuli in psycholinguistic experiments.
+
 The frequency-based surprisal metric gives us an idea of when in their utterances speakers say frequent i.e.~independently information-rich words.
@@
 
-See in the FIGURE below a qualitative comparison of the information trajectory they found and the information trajectory we found.
+See in Figure @ref(fig:bnc\_unigrams) a qualitative comparison of the information trajectory they found and the information trajectory we found.
 It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as Yu et al.
@@
 
-Comparison between the unigram entropy and surprisal BNC results plot
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/bnc_unigrams-1.pdf}
+\caption{(\#fig:bnc\_unigrams)BNC frequency-based information curves}
+\end{figure}
 
@@
 What about context?
-So far we've only discussed the unigram (``contextless'') surprisal metric, considering words on their own without any explicit incorporation of prior context.
+So far we've only discussed the frequency-based metric, considering words on their own without any explicit incorporation of prior context.
 As previously discussed, listeners decode information and process what they hear incrementally, using prior heard words to ease the comprehension process.
-We now include one word of context (bigrams) and then two words of context (trigrams).
+We now include one word of context (bigrams) and then two words of context (trigrams) for each word in our measurements.
 We observe a flattening effect of context across both modalities and all speaker populations.
 After the first word or two, where the listener does not have access to prior context, then they decode information at a flat and more or less uniform rate.
-The bigram and trigram information curves for Switchboard are in Figures @ref(fig:switchboard\_bigrams) and @ref(fig:switchboard\_trigrams).
+The contextual information curves for the BNC and Switchboard are in Figures @ref(fig:switchboard\_bigrams) and @ref(fig:switchboard\_trigrams).
 
@@
 \includegraphics{paper_files/figure-latex/providence_unigrams-1.pdf}
-\caption{(\#fig:providence\_unigrams)Providence frequency-based information curves}
+\caption{(\#fig:providence\_unigrams)Providence frequency-based information curves. Lines around each point indicate 95\% confidence intervals computed with non-parametric bootstrap}
 \end{figure}
@@
 For the bigram and trigram information curves, we see the same contextual smoothing effect as for all the corpora we worked with in English.
-While the unigram information curves may depend based on the language, the contextual information curves show the same trajectory cross-linguistically.
-Using ngrams with n higher than 3 is difficult for parent-child speech corpora because the utterances are so short on average (less than 10 words).
-We hypothesize that the unigram information curves may vary based on the genealogy and typology of the languages in question, but this does not extend to the trigram information curves in particular.
-We shall show evidence that this is true in the rest of the paper.
+While the frequency-based information curves may depend based on the language, the contextual information curves show the same trajectory cross-linguistically.
+Using more than two words of context is difficult for parent-child speech corpora because the utterances are so short on average (less than \(10\) words).
+We hypothesize that the frequency-based information curves may vary based on the genealogy and typology of the languages in question, but this does not extend to the information curves with two words of context in particular.
 
@@
 To make a claim about language as a whole and languages on a larger scale, we needed to use larger corpora and a much larger number of languages.
-We pulled corpora for 159 diverse languages from Wikipedia, each of which had at least 10000 articles.
+We pulled corpora for \(159\) diverse languages from Wikipedia, each of which had at least \(10000\) articles on the knowledge base.
 We split each article up into sentences; the variance in sentence lengths for Wikipedia was significantly larger than for the CHILDES corpora we used in the previous section.
-Most sentences in Wikipedia contained between 10 and 30 words, unlike the CHILDES corpora which mainly contained utterances with under 10 words.
-We excluded the small fraction of utterances with more than 50 words as uncharacteristic of English speech.
+Most sentences in Wikipedia contained between \(10\) and \(30\) words, unlike the CHILDES corpora which mainly contained utterances with under 10 words.
+We excluded the small fraction of utterances with more than \(50\) words since they were small in number and, from manual inspection, uncharacteristic of typical written sentences.
 
-How to analyze more than 40 different surprisal curves for each language, adding up to several thousand surprisal curves total?
+How do we analyze more than \(40\) different surprisal curves for each language, adding up to several thousand surprisal curves total?
 We used two different strategies, which yielded identical results upon analysis.
@@
 We then similarly computed surprisal values at each of these positions, and similarly computed slopes between the surprisal values at each position, giving us another five slope values for each language summarizing the surprisal curves.
-Illustrations of these two strategies are below.
+Illustrations of these two strategies are in Figure \ref{fig:treatments}
 
-Illustration of the regions we target with each of the two slope treatments below.
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/treatments-1.pdf}
+\caption{\label{fig:treatments}Illustration of slope treatments for Wikipedia information curves: relative on top and absolute on bottom}
+\end{figure}
 
-We computed unigram, bigram and trigram values for each language.
-All of this data, along with the scripts used for the analysis, are available on GitHub.
-We ran this code on the RCC at the University of Chicago; it took X amount of time and used up Y amount of carbon to run.
+We computed unigram, bigram and trigram values for each language, and grouped the contextual and frequency-based values for each treatment.
 
 To more rigorously described the differences between languages, we used data from the World Atlas of Language Structures (WALS; Dryer \& Haspelmath, 2013).
-WALS has data for 144 typological features in 2569 languages around the world.
+WALS has data for \(144\) typological features in \(2569\) languages around the world.
 These features describe aspects of morphology, syntax, phonology, etymology and semantics--in short the features describe the structures in each language.
 As WALS is a compiled database from dozens of papers from different authors, most of the features and languages are fairly sparse.
-Even limiting ourselves to the 159 language corpora we pulled from Wikipedia and 122 features from WALS, there are nearly 20000 individual possible feature values, fewer than half of which were already inputted for those languages in the WALS database.
+Even limiting ourselves to the \(159\) language corpora we pulled from Wikipedia and \(122\) features from WALS, there are nearly \(20000\) individual possible feature values, fewer than half of which were already inputted for those languages in the WALS database.
 
@@
 MIMCA essentially uses mean imputation to begin with the missing values, then performs repeatedly performs principle components analysis on and reconstructs the contingency table formed from observations in categorical variables.
-By the end of this we had unigram, bigram and trigram surprisal curves for the 159 language corpora pulled from Wikipedia, along with 122 typological features for each language.
+By the end of this we had frequency-based and contextual information curves for the \(159\) language corpora pulled from Wikipedia, along with \(122\) typological features for each language.
 
 However, the WALS features describe specific structural differences between languages, while our surprisal metric is lexical.
-To target lexical differences between languages, we computed the average normalized Levenshtein distance (LDN; Holman et al., 2008) over the 40 item Swadesh list (Swadesh, 1955), retrieved from the ASJP database (Wichmann et al., 2016).
+To target lexical differences between languages, we computed the average normalized Levenshtein distance (LDN; Holman et al., 2008) over the \(40\) item Swadesh list (Swadesh, 1955), retrieved from the ASJP database (Wichmann et al., 2016).
 The Swadesh list is designed to include near-universal words that target basic cognitive concepts, and are useful in determining the genealogical similarities and differences between languages.
@@
 A sample from the dendrogram is shown in Figure \ref{fig:dendro}, and the full tree can be constructed using the code in our GitHub repository.
-The language family-like structure can be seen at first glance, although the dendrogram does not exactly replicate language genealogy for all 159 languages.
+The language family-like structure can be seen at first glance, although the dendrogram does not exactly replicate language genealogy for all \(159\) languages.
 This suggests using a first-pass quantitative method that the information curves do correspond in some measure to language families, but language families do not explain the entire story.
@@
 
-A sample of the trigram information curves are plotted in Figure @ref(fig:wiki\_trigrams), and all trigram information curves for the languages we used follow the same pattern.
+A sample of the contextual information curves (two words of context) are plotted in Figure @ref(fig:wiki\_trigrams), and all trigram information curves for the languages we used follow the same pattern.
 The first few words in utterances for each language are surprising, but with even a couple words of predictive context for each word, the amount of information in each word drops off and speakers produce information at a constant, optimal rate.
@@
 Figure @ref(fig:cosine\_wals) displays the results.
+This result is surprising based on the significance of many WALS features in predicting the shapes of the frequency-based information curves, and we return to this result in the general discussion.
 
@@
 
-So the more similar two languages are in terms of typological features, the more similar their unigram information curves tend to be.
 For lexical features, we see a stronger correlation between the similarity of two languages in terms of their average LDN and the cosine distance between their information curves.
 Figure @ref(fig:cosine\_ldn).
+We see a higher \(r^2\) value here of \(.026\), indicating that there is more correspondence between a language's lexical similarity to another language and their similarity in information curves.
 
@@
 
-The shape of a language's frequency-based information curve covaries with its typological and lexical similarity to other languages.
+From these typological investigations, we conclude that the shape of a language's frequency-based information curve covaries with its typological and lexical similarity to other languages.
 
-\hypertarget{general-discussion}{%
-\section{General Discussion}\label{general-discussion}}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
-The language we speak impacts how we encode information into our utterances, and how listeners decode our speech signals.
-However, despite the variety in communicative codes, all languages are characterized by predictability and a lack of suprisal in regular speech.
-In this paper, we developed a novel computational modeling method for examining the average information at a fine-grained level in utterances of different lengths in a variety of corpora and across a large number of languages.
-With only a couple of words of context for each word, cross-linguistic differences disappear and all languages are characterizes by low surprisal (low processing cost and information).
+By considering the distribution of information at the level of utterances and sentences, we join together the information-theoretic work focusing on sub-word units and words, and that focusing on paragraphs.
+In doing so, we show that frequency and context-based metrics complement one another in studying efficiency and information in language.
+We directly link linguistic efficiency in a language to the genealogy and properties of that language, We provide evidence for a novel linguistic universal: low processing cost for listeners beyond the first words in utterances, driven by high average word predictability in conversation.
+Also considering the developmental angle, observing language-specific information distributions arising as soon as children begin speaking in multi-word utterances.
 
-The Uniform Information Density Hypothesis (Jaeger \& Levy, 2007) provides an optimal strategy for distributing information among the words in an utterance.
-The hypothesis suggests that speakers tend to distribute information evenly in their utterances, close to the capacity of the communication channel.
-This strategy avoids peaks in information which lead to issues for the decoder, while making the best use of the communication channel possible and aiming to produce as much information as the listener can decode at every moment.
+Throughout this work we have averaged the surprisal values at each position.
+Averaging removes variation, which in turn may obscure trends in the data.
+As discussed in the methods section, the surprisal metric has historically been used for calculating the information and processing cost for individual utterances, and our use of the metric here is actually a step forward rather than a step back.
+Future work can investigate variation in how speakers distribute information in individual utterances.
+
+The WALS database we used to investigate typological variation in the information curves is overall sparse.
+We imputed well over \(50\%\) of the WALS features for most of our \(159\) languages, although all of the languages had at least \(20\) features evaluated in WALS.
+A large part of this is due to WALS being a collection of a number of different studies, instead of a systematic effort to catalogue variation across the world's languages.
+Additionally, WALS features are meant to describe specific microvariations in languages, not to provide a comprehensive typological representation of each language compared to each other language.
+This may be why the Swadesh list provided a higher correlation for describing the differences in information curves: Swadesh (1955) intended the list to allow researchers to more comprehensively compared and constrast lexical differences between languages.
+For our Wikipedia analysis, we also reduce all of a language's variation down to a five-dimensional vector.
+These information curve representations show a surprising amount of variation despite the degree of compression.
 
@@
 
+\leavevmode\hypertarget{ref-gibson2017color}{}%
+Gibson, E., Futrell, R., Jara-Ettinger, J., Mahowald, K., Bergen, L., Ratnasingam, S., \ldots{} Conway, B.
+R.
+(2017).
+Color naming across languages reflects color use.
+\emph{Proceedings of the National Academy of Sciences}, \emph{114}(40), 10785--10790.
+
 \leavevmode\hypertarget{ref-holman2008advances}{}%
@@
 Jaeger, T. F., \& Levy, R. P. (2007). Speakers optimize information density through syntactic reduction. \emph{Advances in neural information processing systems}, 849--856.
+
+\leavevmode\hypertarget{ref-kutas2011thirty}{}%
+Kutas, M., \& Federmeier, K.
+D.
+(2011).
+Thirty years and counting: Finding meaning in the n400 component of the event-related brain potential (erp).
+\emph{Annual Review of Psychology}, \emph{62}, 621--647.
 
@@
 
+\leavevmode\hypertarget{ref-mahowald2013info}{}%
+Mahowald, K., Fedorenko, E., Piantadosi, S. T., \& Gibson, E. (2013). Info/information theory: Speakers choose shorter words in predictive contexts. \emph{Cognition}, \emph{126}(2), 313--318.
+
 \leavevmode\hypertarget{ref-palasis2009syntaxe}{}%
@@
 Piantadosi, S. T., Tily, H., \& Gibson, E. (2011). Word lengths are optimized for efficient communication. \emph{Proceedings of the National Academy of Sciences}, \emph{108}(9), 3526--3529.
+
+\leavevmode\hypertarget{ref-piantadosi2012communicative}{}%
+Piantadosi, S. T., Tily, H., \& Gibson, E. (2012). The communicative function of ambiguity in language. \emph{Cognition}, \emph{122}(3), 280--291.
 

### 2019-09-17 surprisal/childes_surprisal.Rmd
laplace smoothing in childes surprisal

@@
 ```{r hard-coded language unigrams}
-child_utterances <- palasis_utterances %>%
+child_utterances <- prov_utterances %>%
   filter(speaker_role != "Target_Child") %>% 
@@
   ungroup() %>%
+  mutate(n = n + 1) %>%
   mutate(p = n / sum(n))
@@
 child <- child_utterances %>%
-  left_join(child_unigrams) %>%
+  left_join(child_unigrams, by = "word") %>%
   mutate(s = -log(p)) %>%
@@
   select(length, word_order, s) %>%
-  nest(s) %>%
-  filter(length %in% c(5, 7, 9))
+  filter(length %in% c(5, 7, 9)) %>%
+  group_by(length, word_order) %>%
+  nest()
+
+child %>%
+  mutate(surprisal = map(data, ~tidyboot_mean(.x, s, nboot = 10))) %>%
+  select(-data) %>%
+  unnest(cols = c(surprisal))
+
+  ggplot(aes(x = word_order, y = surprisal, color = as.factor(length))) + 
+  geom_point() +
+  geom_line()
 
@@
                                tidyboot_mean(data = unnest(slice(child, .x)),
-                                             column = s, nboot = 1000)))
+                                             column = s, nboot = 100)))
 
@@
 ```{r bigrams}
+vocab_size <- child_utterances %>%
+  ungroup() %>%
+  distinct( word) %>% 
+  nrow()
+
 child_bigrams <- child_utterances %>%
@@
   left_join(child_unigrams, by = c("lag_word" = "word")) %>%
-  mutate(cond_p = n.x/n.y) %>%
+  mutate(cond_p = (n.x + 0)  / (n.y + 0 * vocab_size)) %>%
   select(-n.y, -p)
@@
   mutate(lag_word = lag(word)) %>%
-  left_join(child_bigrams, by = c("lag_word", "word")) %>%
-  left_join(child_unigrams, by = "word")
+  left_join(child_bigrams, by = c("lag_word", "word")) #%>%
+  #left_join(child_unigrams, by = "word")
 
 child <- bigram_surprisals_prep %>%
-  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
+  filter(!is.na(lag_word)) %>%
+  mutate(s = -log(cond_p)) %>%
+ # mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
   ungroup() %>%
   select(length, word_order, s) %>%
-  nest(s) %>%
-  filter(length %in% c(5, 7, 9))
+  filter(length %in% c(5, 7, 9)) %>%
+  group_by(length, word_order) %>%
+  nest()
+
+child %>%
+  mutate(surprisal = map(data, ~tidyboot_mean(.x, s, nboot = 10))) %>%
+  select(-data) %>%
+  unnest(cols = surprisal) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, color = as.factor(length))) + 
+  geom_point() +
+  geom_line()
 

### 2019-09-17 surprisal/bnc&switchboard.Rmd
begin implementation of Kneser-Ney smoothing

@@
                                              column = s, nboot = 1000)))
-# switch_trigram_surprisals <- trigram_surprisals_prep %>%
-#   mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
-#                     ifelse(is.na(lag_word1), # check if not the second word
-#                            -log(p), # trigram if third or beyond
-#                            -log(cond_p)), # bigram if second word
-#                     -log(tri_cond_p))) %>% # unigram if first word
-#   group_by(length, word_order) %>%
-#   summarise(s = mean(s)) %>% 
-#   filter(word_order <= length)
-
 ```
@@
   bind_rows(read_feather(here("../bnc/bnc30.feather")))
+
+selected_bnc <- bnc %>% 
+  select(length, utterance_id) %>%
+  distinct() %>%
+  group_by(length) %>%
+  sample_n(8000) %>% # downsampling because otherwise there are at least 70000 utterances for each length
+  ungroup() %>%
+  select(utterance_id)
+
+bnc <- bnc %>%
+  semi_join(selected_bnc)
 ```
@@
   mutate(p = n / sum(n))
-
-bnc_surprisals <- bnc %>%
-  left_join(bnc_unigrams) %>%
-  mutate(s = -log(p)) %>%
-  group_by(length, word_order) %>%
-  summarise(s = mean(s))
 
@@
                                tidyboot_mean(data = unnest(slice(bnc_surprisals_prep, .x)), 
-                                             column = s, nboot = 100)))
-
-# bnc_surprisals <- bnc %>%
-#   left_join(bnc_unigrams) %>%
-#   mutate(s = -log(p)) %>%
-#   select(length, word_order, s) %>%
-#   filter(length == 10) %>%
-#   group_by(length, word_order) %>%
-#   # summarise(s = mean(s)) 
-#   tidyboot_mean(s)
-
+                                             column = s, nboot = 1000)))  
 ```
@@
   ungroup() %>%
-  mutate(joint_p = n / sum(n)) %>%
-  select(-n) %>%
   left_join(bnc_unigrams, by = c("lag_word" = "word")) %>%
-  mutate(cond_p = joint_p / p) %>%
-  select(-n, -p)
+  mutate(cond_p = n.x / n.y) %>%
+  select(-n.y, -p) %>%
+  rename(n = n.x)
 
@@
   mutate(lag_word = lag(word)) %>%
-  left_join(bnc_bigrams) %>%
-  left_join(bnc_unigrams)
+  ungroup() %>%
+  left_join(bnc_bigrams, by = c("word", "lag_word")) %>%
+  left_join(bnc_unigrams, by = "word")
 
-bnc_bigram_surprisals <- bigram_surprisals_prep %>%
+bbs <- bigram_surprisals_prep %>%
   mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
-  group_by(length, word_order) %>%
-  summarise(s = mean(s)) %>% 
-  filter(word_order <= length)
+  ungroup() %>%
+  select(length, word_order, s) %>%
+  nest(s)
+
+bnc_bigram_surprisals <- map_dfr(1:nrow(bbs),
+        ~bind_cols(select(slice(bbs, .x), length, word_order),
+                               tidyboot_mean(data = unnest(slice(bbs, .x)),
+                                             column = s, nboot = 1000)))
+
 ```
@@
   ungroup() %>%
-  mutate(tri_joint_p = n / sum(n)) %>%
-  select(-n) %>%
   left_join(bnc_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
-  mutate(tri_cond_p = tri_joint_p / joint_p) %>%
-  select(-joint_p, -tri_joint_p, -cond_p) # %>%
+  mutate(tri_cond_p = n.x / n.y) %>%
+  select(-n.x, -n.y, -cond_p)
 
@@
   left_join(bnc_trigrams) %>% 
-  left_join(select(bnc_bigrams, -joint_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
-  left_join(bnc_unigrams)
+  left_join(bnc_bigrams, by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
+  left_join(bnc_unigrams, by = c("word"))
 
-bnc_trigram_surprisals <- trigram_surprisals_prep %>%
+bts <- trigram_surprisals_prep %>%
   mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
@@
                     -log(tri_cond_p))) %>% # unigram if first word
-  group_by(length, word_order) %>%
-  summarise(s = mean(s)) %>% 
-  filter(word_order <= length)
+  ungroup() %>%
+  select(length, word_order, s) %>%
+  nest(s)
 
-bnc_trigram_surprisals %>%
-  ggplot(aes(x = word_order, y = s)) + 
-  facet_wrap(~ length) + 
-  geom_point() + 
-  geom_smooth(se = F)
+bnc_trigram_surprisals <- map_dfr(1:nrow(bts),
+        ~bind_cols(select(slice(bts, .x), length, word_order),
+                               tidyboot_mean(data = unnest(slice(bts, .x)),
+                                             column = s, nboot = 1000)))
+
+```
+
+```{r saving switchboard surprisals}
+bnc_surprisals %>% 
+  mutate(context = 0, corpus = "BNC") %>%
+  bind_rows(bnc_bigram_surprisals %>% mutate(context = 1, corpus = "BNC")) %>% 
+  bind_rows(bnc_trigram_surprisals %>% mutate(context = 2, corpus = "BNC")) %>% 
+  write_feather(here("Data/bnc_bootstrap.feather"))
 ```

### 2019-09-17 surprisal/smoothing.Rmd
begin implementation of Kneser-Ney smoothing

---
title: "Kneser-Ney smoothing"
author: "Josef Klafka and Dan Yurovsky"
date: "9/17/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(tidyboot)
library(childesr)

knitr::opts_chunk$set(echo = TRUE)
```

```{r set up}

child_utterances <- get_utterances(corpus = "Providence") %>%
  filter(speaker_role == "Target_Child") %>% 
  mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n()) %>%
  ungroup()

vocab_size <- child_utterances %>%
  ungroup() %>%
  distinct(word) %>% 
  nrow()

child_unigrams <- child_utterances %>%
  count(word) %>%
  mutate(p = (n + 1) / (sum(n) + vocab_size))

child <- child_utterances %>%
  left_join(child_unigrams, by = "word") %>%
  mutate(s = -log(p)) %>%
  ungroup() %>%
  select(length, word_order, s) %>%
  filter(length %in% c(5, 7, 9)) %>%
  group_by(length, word_order) %>%
  nest()

child %>%
  mutate(surprisal = map(data, ~tidyboot_mean(.x, s, nboot = 10))) %>%
  select(-data) %>%
  unnest(cols = c(surprisal)) %>%
  ggplot(aes(x = word_order, y = empirical_stat, color = as.factor(length))) + 
    geom_point() +
    geom_line()

# by-word: number of times w_i is unique continuation / number of unique continuations overall
continuations <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  ungroup() %>%
  select(lag_word, word) %>%
  filter(!is.na(lag_word)) %>%
  distinct() %>%
  count(word) %>%
  mutate(cont_p = n/sum(n)) %>%
  select(-p)

child_bigrams <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  left_join(child_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = (n.x + 1)  / (n.y + vocab_size))

```

### 2019-09-24 Paper/paper.Rmd
first couple of paras

--- 2019-09-16 Paper/paper.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
+theme_set(theme_classic(base_size = 14))
 ```
@@
 
-Language can be thought of as a code: speakers encode their thoughts into a sequence of linguistic signals, namely an utterance, which listeners then must decode in order to recover the intended message.
-How do speakers structure their utterances so that communication is successful?
-Using tools from information theory and sentence processing, we find that speakers must structure the information in their utterances according to the grammatical, morphological and lexical constraints of their language, which vary greatly from language to language.
-This information structure characterizes language from children’s first utterances to large-scale knowledge base entries.
-Once listeners have access to linguistic context within an utterance, then they can decode the rest of that utterance at a constant and optimal rate.
-
-# Introduction
-
-We use language to communicate information.
-Although we often need to say things we have never said before, we are easily and efficiently able to produce novel utterances with next to no prior planning.
-Our listeners incrementally process each word we say, and predict which words will come next within fractions of a second of hearing the beginning of the word (for evidence see e.g.
-Kutas & Federmeier, 2010).
-Encoding new and unpredictable information into our utterances surprises our listeners, and presents problems for communication.
-Listeners struggle to decode unpredictable words and may have difficulty decoding the following words as well.
-Efficient and effective communication relies on interlocutors being able to process novel information in without being surprised by new information in unpredictable words.
-On the other hand, to make best use of the capacity of the communication channel, speakers should encode as much information as possible into each word they say.
-
-Due to their differences in distributional frequency, words vary in how much information-theoretic content they contain [@shannon1948mathematical].
-This information content plays an important role in determining the shape of words and speech.
-Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be as speakers minimize effort in the amount of information they encode in speech [@zipf1949human].
-Over and above frequency, contextual information content predicts word length [@piantadosi2011word].
-Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side [@jaeger2007speakers].
-Across utterances, context allows speakers to produce more informative and surprising sentences later in communication.
-The later a sentence occurs in a paragraph, the the informative it is and the more unusual and surprising words it contains [@genzel2002entropy].
-From the level of individual words to sentences in paragraphs, information content shapes and is shaped by the context it appears in.
-
-Efficiency plays a big role in communication, from the way that languages evolve to the individual utterance choice timescale. @mahowald2013info. @gibson2017color. @piantadosi2011word.
-
-How do frequency and predictive context determine how we distribution information at the level of utterances?
-We do not speak only in phrases or paragraphs, but conversationally we speak in utterances and in writing we produce sentences as a coherent unit of a complete thought.
-Languages possess very different grammatical and morphological structures--the same thought can be encoded in at least as many ways as there are different languages.
-The average rate of information speakers transfer in syllables differs from language to language [@aylett2004smooth].
-How do the different structures seen in different languages, with different rates of information encoding, affect how speakers structure information in production?
-
-@yu2016distribution proposed a novel method for quantifying the amount of information typically found at each part of English utterances.
-Treating each word position by itself, @yu2016distribution calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty [@shannon1948mathematical].
-They found a distinctive three-step shape to the distribution of information in English written sentences.
-
-In this paper, we expand on this body of prior work in a number of novel ways.
-We replicate the results from @yu2016distribution with a different metric, one intimately tied to incremental word processing.
-We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech.
-We also apply a contextual version of the same metric, and find that in communication, listeners are able to decode information from the utterances they hear at a constant rate past the initial words.
-We expand the study of information density to the largest set of languages yet considered, and incorporate contextual and typological information.
-We find that each language has a characteristic, robust distribution of information predictable from its morphology, phonology and syntax, and tied to the language’s genealogy.
-Speakers of a language will tend to distribute information according to that language’s distribution of information in their utterances.
-As soon as a child can form multi-word utterances in a given language, they will produce utterances obeying the characteristic distribution of information in a given language.
-
-# Methods
-
-@shannon1948mathematical defined information as “the reduction in uncertainty about one variable given that the value of another variable is known”.
-We use a metric proposed by Shannon and applied by @levy2008expectation: lexical surprisal, which defines the information in word based on the ratio of possible continuations of the sentence after to before the word is seen.
-Equivalently, we can compute surprisal with the predictability of the word, as in the formula below.
-The surprisal of a word is inversely proportional to the predictability of a word, such that less common and less predictable words carry more information.
-
-$$surprisal(word) = -log P(word)$$
-
-The surprisal of a word is also correlated with the processing cost of a word, with evidence from eye-tracking [@smith2013effect] and ERP [@frank2015erp] studies, among other sources.
-Considered without context, the surprisal of an individual word is inversely proportional to the frequency of that word, so that the less often a person has seen a word, the more information that word holds.
-For example, “flower” has less information than “azalea” because “flower” is much more common than “azalea”.
-Though the two words have the same length in number of letters, it’s more difficult to process “azalea” when reading it here than when reading “flower”.
-Frequency is intimately tied information content in words, with much of the differences between words frequencies being explained by information content cross-linguistically [@piantadosi2011word].
-
-However, when reading or listening, people don’t just consider each word as an isolated linguistic signal.
-Instead, listeners use the words they have already heard to predict and decode the word they are currently processing.
-Following this incremental processing paradigm, we can also condition the surprisal of a word in its context.
-In the formula below, widenotes the word currently being read or heard, while wi-1denotes the first word before the current word, wi-2denotes the second word before the current word, and so on.
-
-$$surprisal(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...)$$
-$$= -log \frac{P(w_i,w_{i-1}w_{i-2},...)}{P(w_{i-1}w_{i-2}...)}$$
-
-When we use a word or two of context in our surprisal calculations, then the set of reasonable final items in our ngrams is greatly restricted.
-“Flower” may contain less information than “azalea” when we consider the words independently of their context, but with context this can be reversed.
-Flower appears in a variety of contexts, and so the information content of a word like “flower” in a particular context may be higher than “azalea”.
-If you only have azaleas in your garden, then hearing someone say “in that garden, look at the flowers” may be higher surprisal for you: you expect them to say “azalea”.
-This prediction does not need many words before to work out, for example in “I take my coffee with cream and sugar”.
-When you hear “cream and”, you automatically predict “sugar” or maybe “honey”, but there are few possible continuations with even those two words.
-Hearing “I” restricts the next word to a verb, or possibly an adverb, and since you have just heard the speaker refer to themselves in the first person singular, your set of possible completions is significantly restricted.
-
-To find the distribution of information within a corpus, we calculate the surprisal or contextual surprisal for each unique word or sequence of words within the corpus.
-We take all utterances of a given length, and for each word position in utterances of that length, we compute the average of the surprisals for all of the non-unique words that occur in that position, conditioned or not conditioned on context.
-By computing these averages for each word position in an utterance, we compute a low-dimensional approximation to the average distribution of information in the corpus.
-With the surprisal metric, we base the information contained in each word on how often the word is encountered in its context in the corpus.
-As long as the corpus is representative of the language or population we study, then the distribution of information is approximated for that language or population as a whole.
-
-The frequency-based surprisal metric gives us an idea of when in their utterances speakers say frequent i.e.
-independently information-rich words.
-The context-based surprisal metric show us how speakers tend to distribute the information in utterances relative to real-time processing in communication.
-We expect a priori that our frequency-based surprisal curve will be flat.
-No one part of the sentence will on average have words that are more frequent than another across utterance lengths.
-Similarly, we expect that there will be a small smoothing effect for our contextual surprisal metric.
-
-# Written and Spoken English have the same information distribution
-
-We first turn to working with written English in the British National Corpus [BNC; @leech1992100].
-The BNC is a collection of spoken and written records (90% written) from the turn of the century, intended to be a representative sample of British English.
-Using their word entropy metric without context, Yu et al.
-found a distinctive three-step distribution for information in written English sentences in the corpus.
-The first word tended to contain little information.
-While the middle words of sentences each had more information than the first word, information was flat and non-increasing across the middle of sentences.
-The final word, they found, contained the most information out of any in the sentence, with a noticeable spike in information.
-They found the same distribution across sentence lengths, from sentences with 15 words to sentences with 45 words.
-
-We first replicate the @yu2016distribution result using the surprisal metric in place of the entropy metric.
-We use the frequency-based or “contextless” surprisal metric, which determines the average distribution of information based on word frequencies in a corpus.
-A priori we expect that the frequency-based metric will produce a flat distribution of information across word positions in the BNC.
-
-See in the FIGURE below a qualitative comparison of the information trajectory they found and the information trajectory we found.
-It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as @yu2016distribution did.
-The first words in written English sentences tend to contain little information, while the final words tend to contain a great deal of information.
-
-Comparison between the unigram entropy and surprisal BNC results plot
-
-However, this only characterizes written English sentences.
-If this information distribution is characteristic more generally of English, then we will find the same distribution for spoken English.
-We use the Switchboard corpus of spoken English telephone conversations between adults to compute the distribution of information across spoken utterances in English.
-The sentences in the BNC, compared to the conversations in Switchboard, feature much longer sentences with larger vocabularies and more complex sentence structures.
-We split the conversations in Switchboard not by sentences, as in the BNC, but by turns.
-
-For Switchboard we find the same information trajectory as we did for the BNC, and we find the same distribution of information regardless of utterance length.
-The longer the utterance, the more pronounced the information distribution shape becomes.
-A qualitative comparison of the two trajectories is in Figure \@ref(fig:switchboard_unigrams).
-
-```{r switchboard_unigrams, echo=F, fig.height=2, fig.cap="Switchboard frequency-based information curves"}
-sw_plot <- switch_results %>% 
-  filter(context == 0) %>%
-  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
-    geom_point() + 
-    geom_pointrange() + 
-    geom_line() +
-    facet_wrap(~length) + 
-    scale_x_continuous(breaks = 1:11) + 
-    xlab("Word position") + 
-    ylab("Mean information")
-
-bnc_plot <- bnc_results %>% 
-  filter(context == 0) %>%
-  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
-    geom_point() + 
-    geom_pointrange() + 
-    geom_line() +
-    facet_wrap(~length) + 
-    scale_x_continuous(breaks = 1:30) + 
-    xlab("Word position") + 
-    ylab("Mean information")
-
-grid.arrange(bnc_plot, sw_plot, nrow = 2)
-```
-
-We have found a unique average distribution of information that appears to characterize the English language as a whole, regardless of medium (spoken or written) or utterance length.
-This distribution indicates that in English, the words we speak or write at the beginnings of utterances have little information, while the words we speak or write at the ends of utterances have a lot of information.
-The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from [@genzel2002entropy].
-
-What about context?
-So far we’ve only discussed the unigram (“contextless”) surprisal metric, considering words on their own without any explicit incorporation of prior context.
-As previously discussed, listeners decode information and process what they hear incrementally, using prior heard words to ease the comprehension process.
-We now include one word of context (bigrams) and then two words of context (trigrams).
-We observe a flattening effect of context across both modalities and all speaker populations.
-After the first word or two, where the listener does not have access to prior context, then they decode information at a flat and more or less uniform rate.
-The bigram and trigram information curves for Switchboard are in Figures \@ref(fig:switchboard_bigrams) and \@ref(fig:switchboard_trigrams).
-
-```{r switchboard_bigrams, echo=F,fig.height=2,fig.cap="Switchboard bigram context-based information curves"}
-switch_results %>% 
-  filter(context == 1) %>%
-  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
-    geom_point() + 
-    geom_pointrange() + 
-    geom_line() +
-    facet_wrap(~length) + 
-    scale_x_continuous(breaks = 1:11) + 
-    xlab("Word position") + 
-    ylab("Mean information")
-```
-
-```{r switchboard_trigrams, echo=F,fig.height=2,fig.cap="Switchboard trigram context-based information curves"}
-switch_results %>% 
-  filter(context == 2) %>%
-  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
-    geom_point() + 
-    geom_pointrange() + 
-    geom_line() +
-    facet_wrap(~length) + 
-    scale_x_continuous(breaks = 1:11) + 
-    xlab("Word position") + 
-    ylab("Mean information")
-```
-
-Speakers produce information at a more or less constant rate, avoiding peaks or troughs in their information distribution, except at the beginnings of utterances, where listeners may not have any context to predict what the speaker is going to say.
-Speakers of English tend towards a characteristic and uneven distribution of word information based on frequencies within their utterances.
-Their interlocutors, however, once they have a word or two of context, decode information at a more or less constant and optimal rate.
-
-# English CDS and child speech
-
-We found that English speakers and writers used the same robust and distinctive distribution of information within each utterance, regardless of the number of words in their utterances.
-To determine if this distribution truly characterizes all speakers of the English language as a whole, we wanted to examine speech from English-speaking children who are producing their very first multi-word utterances.
-We hypothesize the three-step distribution of information we found for English will characterize child speech and child-directed speech.
-
-We use the Providence corpus from CHILDES [@macwhinney2000childes; @evans2012individual], which consists of about 650000 utterances from parent-child conversations for six monolingual English-speaking children talking with the parents.
-We obtained the Providence corpus using the childesr frontend to the childes-db database [@sanchez2019childes].
-The child age range was between 11 months and 4 years.
-Sessions recorded in the home.
-Parents said a majority of the utterances, but children take up an increasing share of the conversation as they grow older and are actually able to speak full utterances.
-The utterances in the Providence corpus are on average significantly shorter than those in the BNC and Switchboard; over 90% of the utterances in the Providence corpus are at most 10 words long, with the median at around 6 words.
-Similar to our treatment of the Switchboard corpus, we split the Providence corpus up by conversational turns and separate utterances.
-
-We observe the same distinctive distribution of information for parents and children in the Providence corpus as we did for adults in the BNC and Switchboard.
-The distribution of information we found at the level of individual words in English, therefore, characterizes the English language as a whole and not only adult utterances, not only written utterances.
-See \@ref(fig:providence_unigrams).
-
-```{r providence_unigrams, echo=F, fig.height =2, fig.cap="Providence frequency-based information curves"}
-childes_results %>% 
-  filter(language == "English", context == 0) %>%
-  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
-             color = speaker)) + 
-    geom_point() + 
-    geom_pointrange() + 
-    geom_line() +
-    facet_wrap(~length) + 
-    scale_x_continuous(breaks = 1:9) + 
-    xlab("Word position") + 
-    ylab("Mean information")
-```
-
-What about context in the Providence corpus?
-When incorporating one or two words of predictive context, we observe the same trend as in the Switchboard and BNC corpora.
-Beyond the first couple of words, once your interlocutor has enough context to predict with some accuracy what you will say next, then you decode information from their speech stream at a constant and optimal rate.
-This applies to parents and children speaking to one another, as well as adults speaking and writing to one another.
-See \@ref(fig:providence_bigrams) and \@ref(fig:providence_trigrams).
-
-```{r providence_bigrams, echo=F, fig.height =2, fig.cap="Providence bigram context-based information curves"}
-childes_results %>% 
-  filter(language == "English", context == 1) %>%
-  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
-             color = speaker)) + 
-    geom_point() + 
-    geom_pointrange() + 
-    geom_line() +
-    facet_wrap(~length) + 
-    scale_x_continuous(breaks = 1:9) + 
-    xlab("Word position") + 
-    ylab("Mean information")
-```
-
-```{r providence_trigrams, echo=F, fig.height =2, fig.cap="Providence trigram context-based information curves"}
-childes_results %>% 
-  filter(language == "English", context == 2) %>%
-  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
-             color = speaker)) + 
-    geom_point() + 
-    geom_pointrange() + 
-    geom_line() +
-    facet_wrap(~length) + 
-    scale_x_continuous(breaks = 1:9) + 
-    xlab("Word position") + 
-    ylab("Mean information")
-```
-
-# Child speech and CDS across languages
-
-So far, we have only looked at the distribution of information in words in English, both with and without context.
-We have examined child speech, child-directed speech and adult-directed speech, as well as writing samples selected to be representative of British English as a whole.
-But this only captures the picture for English.
-
-We now turn to a small number of typologically diverse languages, and conduct the same analysis, using a monolingual adult-child speech corpora from CHILDES [@macwhinney2000childes] to compare the results from these languages directly to our results from the English Providence corpus.
-We use corpora for Spanish, German, French, Mandarin and Japanese.
-Similarly to the English Providence corpus, all of the following corpora consist mainly of shorter utterances: most utterances in the corpora are under 10 words long.
-For Spanish, we use the corpus of conversations between primary school-aged children between 6 and 9 years old and the researchers, in which the children were asked to recite stories in their native Spanish [@de1996estudio].
-The students came from working class public schools and upper class private schools.
-For German, we use the Wagner corpus [@wagner1985much].
-This corpus consists of a collection of mini-corpora collected from socio-economically diverse families, with children aged from 1;0 to 14;0, mainly on the younger end of that range.
-The corpus mainly consists of parent-child conversations between the children and their primary caregivers.
-For French, we use the Palasis corpus [@palasis2009syntaxe], a longitudinal study of children between the ages of 2;5 and 4;0 from the same kindergarten class.
-The conversations in this corpus take place between the children and other children in the same class, and between the children and the investigator.
-For Mandarin Chinese, we use the Zhou corpus of dinner conversations from the Shanghai area [@li2015dinner].
-The children in this corpus are 5 or 6 years old, and half come from working class families (the other half from upper class families).
-For Japanese, we use the Okayama corpus [@shirai2001acquisition].
-This corpus consists of conversations between children between the ages of 2;2 and 4;2 and their mothers.
-
-We observe a distinct and characteristic frequency-based information trajectory emerge for each language, robust across each utterance length within each language.
-We see the same distribution of information for both parents and children in each language, with the child’s curve normally below the parent’s curve, likely due to the parent speaking a larger share of the utterances in the corpus and using a larger vocabulary than the child.
-We see the opposite amplitude trend in the Shiro corpus, where the children speak more than the adult investigators as the children are the ones telling the stories.
-We include the frequency-based information curve from the Providence corpus for comparison.
-
-```{r, echo=F, fig.height =3, fig.cap="CHILDES frequency-based information curves"}
-childes_results %>% 
-  filter(context == 0, length == 7) %>%
-  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
-             color = speaker)) + 
-    geom_point() + 
-    geom_pointrange() + 
-    geom_line() +
-    facet_wrap(~language) + 
-    scale_x_continuous(breaks = 1:7) + 
-    xlab("Word position") + 
-    ylab("Mean information")
-```
-
-English, Spanish, French and German feature similar information curves shapes, with slight variations.
-The German information curve features lower information for longer towards the beginnings of utterances, possibly due to the grammatical restriction that the second word in German utterances must be a verb (V2).
-Spanish features a larger spike in the amount of information in the final words of utterances.
-For Japanese and Mandarin, we observe completely different unigram information curve trajectories.
-The Japanese unigram information curve trajectory begins high and finishes low, the mirror image of the European language information curves.
-The Mandarin curve begins low and finishes low, but features high information in the middle of utterances.
-We hypothesize this may be due to both Japanese and Mandarin speakers typically ending their utterances with particles, which contain very little information on their own.
-The penultimate noun with the speakers pair the particles are where the information lies.
-
-For the bigram and trigram information curves, we see the same contextual smoothing effect as for all the corpora we worked with in English.
-While the unigram information curves may depend based on the language, the contextual information curves show the same trajectory cross-linguistically.
-Using ngrams with n higher than 3 is difficult for parent-child speech corpora because the utterances are so short on average (less than 10 words).
-We hypothesize that the unigram information curves may vary based on the genealogy and typology of the languages in question, but this does not extend to the trigram information curves in particular.
-We shall show evidence that this is true in the rest of the paper.
-
-```{r, echo=F, fig.height =3, fig.cap="CHILDES bigram context-based information curves"}
-childes_results %>% 
-  filter(context == 1, length == 7) %>%
-  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
-             color = speaker)) + 
-    geom_point() + 
-    geom_pointrange() + 
-    geom_line() +
-    facet_wrap(~language) + 
-    scale_x_continuous(breaks = 1:7) + 
-    xlab("Word position") + 
-    ylab("Mean information")
-```
-
-```{r, echo=F, fig.height =3, fig.cap="CHILDES trigram context-based information curves"}
-childes_results %>% 
-  filter(context == 2, length == 7) %>%
-  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
-             color = speaker)) + 
-    geom_point() + 
-    geom_pointrange() + 
-    geom_line() +
-    facet_wrap(~language) + 
-    scale_x_continuous(breaks = 1:7) + 
-    xlab("Word position") + 
-    ylab("Mean information")
-```
-
-# Language structure and large-scale data analysis: methods and preprocessing
-
-To make a claim about language as a whole and languages on a larger scale, we needed to use larger corpora and a much larger number of languages.
-We pulled corpora for 159 diverse languages from Wikipedia, each of which had at least 10000 articles.
-We split each article up into sentences; the variance in sentence lengths for Wikipedia was significantly larger than for the CHILDES corpora we used in the previous section.
-Most sentences in Wikipedia contained between 10 and 30 words, unlike the CHILDES corpora which mainly contained utterances with under 10 words.
-We excluded the small fraction of utterances with more than 50 words as uncharacteristic of English speech.
-
-How to analyze more than 40 different surprisal curves for each language, adding up to several thousand surprisal curves total?
-We used two different strategies, which yielded identical results upon analysis.
-Each strategy gave us a five-dimensional vector space embedding for each language in a Wikipedia “slope space”.
-For the first strategy, we split each sentence length by number of words into fifths, and computed surprisal values for the closest word position to each quintile.
-We then computed the slopes between the surprisal values at neighboring quintiles, yielding five slope values for each curve.
-For the second strategy, we split each sentence length by number of words into sections based on those areas of the surprisal curves that had seemed most important before: between the first and second word; between the second and third word; between the third word and third-to-last word; between the third-to-last word and the second to last word; and between the second-to-last word and the last word.
-We then similarly computed surprisal values at each of these positions, and similarly computed slopes between the surprisal values at each position, giving us another five slope values for each language summarizing the surprisal curves.
-Illustrations of these two strategies are below.
-
-Illustration of the regions we target with each of the two slope treatments below.
-
-We computed unigram, bigram and trigram values for each language.
-All of this data, along with the scripts used for the analysis, are available on GitHub.
-We ran this code on the RCC at the University of Chicago; it took X amount of time and used up Y amount of carbon to run.
-
-To more rigorously described the differences between languages, we used data from the World Atlas of Language Structures [WALS; @wals].
-WALS has data for 144 typological features in 2569 languages around the world.
-These features describe aspects of morphology, syntax, phonology, etymology and semantics--in short the features describe the structures in each language.
-As WALS is a compiled database from dozens of papers from different authors, most of the features and languages are fairly sparse.
-Even limiting ourselves to the 159 language corpora we pulled from Wikipedia and 122 features from WALS, there are nearly 20000 individual possible feature values, fewer than half of which were already inputted for those languages in the WALS database.
-
-We used the Multiple Imputation Multiple Correspondence Analysis algorithm [MIMCA; @audigier2017mimca] to fill in the missing data using statistical imputation.
-MIMCA essentially uses mean imputation to begin with the missing values, then performs repeatedly performs principle components analysis on and reconstructs the contingency table formed from observations in categorical variables.
-By the end of this we had unigram, bigram and trigram surprisal curves for the 159 language corpora pulled from Wikipedia, along with 122 typological features for each language.
-
-However, the WALS features describe specific structural differences between languages, while our surprisal metric is lexical.
-To target lexical differences between languages, we computed the average normalized Levenshtein distance [LDN; @holman2008advances] over the 40 item Swadesh list [@swadesh1955towards], retrieved from the ASJP database [@wichmann2016asjp].
-The Swadesh list is designed to include near-universal words that target basic cognitive concepts, and are useful in determining the genealogical similarities and differences between languages.
-The results of classifying languages using the Swadesh list and LDN are correlated with those using WALS features, but the Swadesh list and LDN do not suffer from the same sparsity problem as WALS [@holman2008advances].
-
-# Language structures and large-scale data analysis: results
-
-We ran a hierarchical clustering algorithm on the unigram information curves using the hclust package from the R stats core library [@team2013r].
-We used the complete linkage algorithm for hierarchical clustering with distances between information curves in language computed using cosine distance between their embeddings in the slope space.
-The complete linkage algorithm at every step pairs each language or cluster of languages with its closest neighboring language or cluster.
-A sample from the dendrogram is shown in Figure \@ref(fig:dendro), and the full tree can be constructed using the code in our GitHub repository.
-The language family-like structure can be seen at first glance, although the dendrogram does not exactly replicate language genealogy for all 159 languages.
-This suggests using a first-pass quantitative method that the information curves do correspond in some measure to language families, but language families do not explain the entire story.
-
-```{r dendro, echo=F}
-wiki_curves %>%
-  filter(gram == "Unigram") %>%
-  select(-gram) %>% 
-  sample_n(40) %>%
-  column_to_rownames("language") %>%
-  dist() %>% 
-  hclust() %>%
-  ggdendrogram(rotate=T, leaf_labels = F)
-```
-
-A sample of the trigram information curves are plotted in Figure \@ref(fig:wiki_trigrams), and all trigram information curves for the languages we used follow the same pattern.
-The first few words in utterances for each language are surprising, but with even a couple words of predictive context for each word, the amount of information in each word drops off and speakers produce information at a constant, optimal rate.
-
-```{r wiki_trigrams, echo=F, fig.cap="Some trigram information curves from the Wikipedia data"}
-wiki_curves %>%
-  filter(gram == "Trigram") %>% 
-  mutate(Slope0 = runif(1, 0.9, 1.1)) %>% 
-  mutate(Slope1 = Slope1 + Slope0) %>% 
-  mutate(Slope2 = Slope2 + Slope1) %>% 
-  mutate(Slope3 = Slope3 + Slope2) %>% 
-  mutate(Slope4 = Slope4 + Slope3) %>% 
-  mutate(Slope5 = Slope5 + Slope4) %>% 
-  select(-gram) %>% 
-  sample_n(6) %>% 
-  gather(position, information, -language) %>%
-  ggplot(aes(x = position, y = information, group = language, color = language)) + 
-    geom_point() + 
-    geom_line() + 
-    xlab("Word position") + 
-    ylab("Average Information") + 
-    theme(axis.text.x = element_blank())
-```
-
-First we examined the effects of individual typological features on the shapes of the unigram information curves.
-We ran logistic regressions using the lme4 package in R [@bates2014fitting], checking whether the cosine distance between two languages’ embeddings in the slope space played a role in determining if they had the same value for a given WALS feature.
-Individual WALS features do not necessarily have ordinal values.
-Some, such as the “Number of Cases” feature, are easy to quantify and order.
-Others are more difficult.
-For example, how does one order “relative clauses appear after the nouns they modify”, “relative clauses appear before the nouns they modify” and “free order of relative clauses and nouns”?
-We chose the identify relation to avoid deciding on the basis of individual features.
-We found that 100 out of the 120 features from WALS we examined was statistically significant (p < .001) in determining whether two languages had the same shape to their unigram information distributions.
-The results for some important features are in Figure \@ref(fig:linear_models), and the rest of the results are in the Appendices.
-
-```{r linear_models, echo=F, fig.cap="Some linear model results from Wikipedia and WALS features"}
-sf <- wiki_lr %>% 
-  filter(p.value == 0, estimate > 0) %>% 
-  select(name) %>% 
-  rename("Some Important Features" = name) %>% 
-  slice(c(1, 10, 12, 15, 23, 29))
-
-sf$Examples <- c("Noun-RC; RC-noun", 
-                 "No cases; 6 cases", 
-                 "SOV; VSO", 
-                 "Only singular; sing. and plural", 
-                 "Affix; distinct word", 
-                 "Prefixes; suffixes")
-
-sf %>% grid.table()
-```
-
-We next compared how the cosine distance between two languages related to how many WALS features they had in common.
-$r^2$ value is $.005734$, which suggests that in aggregate there is not a correlation between how many WALS features languages have in common and the similarity of their frequency-based information curves.
-Figure \@ref(fig:cosine_wals) displays the results.
-
-```{r cosine_wals, echo = F, fig.cap="wals features vs cosine similarity"}
-r2 <- summary(lm(data = wiki_results, formula = unigram_cosine~feature))$r.squared %>%
-  round(digits = 3)
-wiki_results %>% 
-  ggplot(aes(x = feature, y=unigram_cosine)) + 
-    geom_smooth(aes(weight = .01), se=F, method = "lm", color = "black") + 
-    geom_jitter(alpha = 0.04, color = "blue") + 
-    annotate("text", x = 70, y = -0.5, label = paste0("r^2 = ", r2)) + 
-    xlab("Number of identical WALS features") + 
-    ylab("Information curve similarity")
-    #stat_poly_eq(formula = y~x, position = "dodge")
-```
-
-So the more similar two languages are in terms of typological features, the more similar their unigram information curves tend to be.
-For lexical features, we see a stronger correlation between the similarity of two languages in terms of their average LDN and the cosine distance between their information curves.
-Figure \@ref(fig:cosine_ldn).
-
-```{r cosine_ldn, echo = F, fig.cap="ldn features vs cosine similarity"}
-r2 <- summary(lm(data = wiki_results, formula = unigram_cosine~ldn))$r.squared %>%
-  round(digits = 4)
-wiki_results %>% 
-  ggplot(aes(x = ldn, y=unigram_cosine)) + 
-    geom_smooth(aes(weight = .01), se=F, method = "lm", color = "black") + 
-    geom_jitter(alpha = 0.04, color = "blue") + 
-    annotate("text", x = 0.2, y = -0.5, label = paste0("r^2 = ", r2)) + 
-    xlab("Normalized lexical distance") + 
-    ylab("Information curve similarity")
-    #stat_poly_eq(formula = y~x, position = "dodge")
-```
-
-The shape of a language’s frequency-based information curve covaries with its typological and lexical similarity to other languages.
-
-# General Discussion
-
-By considering the distribution of information at the level of utterances and sentences, we join together the information-theoretic work focusing on sub-word units and words, and that focusing on paragraphs.
-In doing so, we show that frequency and context-based metrics complement one another in studying efficiency and information in language.
-We directly link linguistic efficiency in a language to the genealogy and properties of that language, while providing evidence for a novel, communicatively real linguistic universal.
-Also considering developmental angle, observing language-specific information distributions arising as soon as children begin speaking in multi-word utterances.
+What role does communicative efficiency play in how we organize our utterances?
+In this paper, we present a novel method of examining how much information speakers in a given language communicate in each word in their utterances, surveying numerous diverse languages.
+We find that speakers produce frequent and informative words at regular parts of their utterances, depending on language they use, which is predictable in part from the features and genealogy of their language.
+This robust information distribution characterizes both spoken and written communication, and emerges in children's earliest utterances.
+However, in real-time communication, in-context word predictability allows listeners to process information at a constant, optimal rate, regardless of the information distribution in the language they understand.
 

### 2019-09-24 Paper/info-curves.Rmd
first couple of paras

--- 2019-09-17 Paper/paper.tex
@@
-\documentclass[11pt,]{article}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
-            pdftitle={Speakers of diverse languages structure their utterances for efficient communication},
-            pdfauthor={Josef Klafka and Daniel Yurovsky},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{longtable,booktabs}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
+---
+title             : "Speakers of diverse languages structure their utterances for efficient communication"
+shorttitle        : "Cross-linguistic information structure"
 
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
+author: 
+  - name          : "Josef Klafka"
+    affiliation   : "1"
+    corresponding : yes    # Define only one corresponding author
+    address       : "5000 Forbes Ave"
+    email         : "klafka@cmu.edu"
+  - name          : "Daniel Yurovksy"
+    affiliation   : "1,2"
 
-%%% Change title format to be more compact
-\usepackage{titling}
+affiliation:
+  - id            : "1"
+    institution   : "Carnegie Mellon University"
+  - id            : "2"
+    institution   : "University of Chicago"
 
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
+authornote: |
+  Thanks to folks. JSMF.
 
-\setlength{\droptitle}{-2em}
-
-  \title{Speakers of diverse languages structure their utterances for efficient communication}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-    \author{Josef Klafka and Daniel Yurovsky}
-    \preauthor{\centering\large\emph}
-  \postauthor{\par}
-      \predate{\centering\large\emph}
-  \postdate{\par}
-    \date{2019-09-17}
-
-\usepackage{setspace}
-
-\doublespacing
-
-\begin{document}
-\maketitle
-
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
-
+abstract: |
   What role does communicative efficiency play in how we organize our utterances?
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+keywords          : "keywords"
+wordcount         : "X"
+
+bibliography      : ["info-curves.bib"]
+
+floatsintext      : yes
+figurelist        : no
+tablelist         : no
+footnotelist      : no
+linenumbers       : yes
+mask              : no
+draft             : no
+
+documentclass     : "apa6"
+classoption       : "man"
+output            : papaja::apa6_pdf
+---
+
+```{r load_packages, include = FALSE}
+library(xtable)
+library(knitr)
+library(papaja)
+library(ggthemes)
+library(lmerTest)
+library(directlabels)
+library(ggrepel)
+library(feather)
+library(here)
+library(tidyboot)
+library(broom)
+library(broom.mixed)
+library(ggdendro)
+library(gridExtra)
+library(ggpmisc)
+
+library(tidyverse)
+
+theme_set(theme_classic(base_size = 14) + theme(legend.position = "none"))
+
+knitr::opts_chunk$set(fig.pos = 'tb', echo = FALSE, warning = FALSE, 
+                      cache = TRUE, warning = FALSE, message = FALSE, 
+                      sanitize = TRUE, fig.path='figs/', fig.width = 3,
+                      fig.height = 3)
+set.seed(42)
+options(digits=2)
+```
+
+```{r load data, include = F}
+childes_results <- read_feather(here("Data/all_childes_bootstrap.feather"))
+wiki_results <- read_feather(here("../surprisal/Data/all_spaces.feather"))
+bnc_results <- read_feather(here("Data/bnc_bootstrap.feather"))
+switch_results <- read_feather(here("Data/switch_bootstrap.feather"))
+wiki_curves <- read_csv(here("../surprisal/Data/relative_ngrams.csv"))
+wiki_lr <- read_feather(here("../surprisal/Wikipedia/models.feather"))
+```
+
+# Introduction
+
+One of the defining features of human language is its power to transmit information.
+We use language for a variety of different proximal tasks tasks: greeting friends, taking notes, signalling group identities.
+But all share a common unifying purpose: changing the mental state of the listener or reader of our communications [@austin1975].
+Language can thus naturally thought of as a code that allows speakers to turn their intended meanings into a form that can be transmitted through the air (or paper, or electrons) and subsequently converted by listeners back into an approximation of the intended meaning [@shannon1948].
+
+Beyond its utility as a metaphor, this coding perspective on language is powerful because it affords a framework for rational analysis: If language has evolved to be an optimal code for information transmission, how should this code be structured [@anderson1989]?
+The optimal code would have to optimize two competing pressures: (1) a pressure for listeners to easily and successfully decode messages sent by the speaker, and (2) a pressure for speakers to easily code their messages and transmit them with minimal effort and error.
+A fundamental constriaint on both of these processes is the serial nature of human language: sounds are produced one at a time and each is unavailable perceptually once it is no longer being produced.
+
+The solution that languages have arrived on is incremental processing: Listeners process speech continuously as it arrives, predicting upcoming words and building expectations about the likely meaning of utterances in real-time rather than at their conclusion [@kutas2011; @tanenhaus1995; @pickering2013].
+Because prediction errors can lead to severe processing costs and difficulty integrating new information, an optimal coding strategy is for speakers seeking to minimize their prodction costs is to produce utterances that are are just at the prediction capacity of listeners without exceeding this capacity [@aylett2004; @genzel2002].
+
+One salient constraint on this transmission processes is the serial nature of language:
+
+This coding perspective is powerful because it lends itself naturally to a rat
+
+Languages can thus be thought of as a code that can be used to transform speaker's intented meaning into a
+
+* Coding/decoding and information transfer
+* language as prediction
+* how should the code be structured?
+optimization questions like this have a had * a lot power across different parts of language: zipf, kemp & regier, etc etc.
+noisy channel stuff.
+(or does this go in discussion)
+* don't under or overutilize channel, stay near limit -- constant entropy rate
+* evidence for this at multiple levels from individiual lexical choice *(info/information theory paper), inclusion of non-obligatory aspects of language (filled pauses, that), across multiple utterances in a discourse
+* what about over the course of an individual utterance--perhaps the strongest constraints from language itself?
+* yu et al. found this puzzling effect. 
+* we replicate that in a different kind of analysis and show that it extends to trigrams etc???
+* BUT variation cross-linguistically. why? constraints from language.
 
@@
 Although we often need to say things we have never said before, we are easily and quickly able to produce novel utterances with next to no prior planning.
-Our listeners incrementally process each word we say in utterances they may have never heard before, and predict which words will come next within fractions of a second of hearing the beginning of the word (for evidence see e.g.
-Kutas \& Federmeier (2011)).
+Our listeners incrementally process each word we say in utterances they may have never heard before, and predict which words will come next within fractions of a second of hearing the beginning of the word [for evidence see e.g.
+@kutas2011].
 Encoding new and unpredictable information into our utterances surprises our listeners, and presents problems for communication.
@@
 
-Due to their differences in distributional frequency, words vary in how much information content they contain (Shannon, 1948).
-Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be (Zipf, 1949).
-Over and above frequency, contextual information content predicts word length, due to speakers minimizing effort in the amount of information they encode in speech (Piantadosi, Tily, \& Gibson, 2011).
-Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side (Jaeger \& Levy, 2007).
+Due to their differences in distributional frequency, words vary in how much information content they contain [@shannon1948].
+Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be [@zipf1949].
+Over and above frequency, contextual information content predicts word length, due to speakers minimizing effort in the amount of information they encode in speech [@piantadosi2011].
+Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side [@jaeger2007].
 Across utterances, context allows speakers to produce more informative and surprising sentences later in communication.
-The later a sentence occurs in a paragraph, the the informative it is and the more unusual and surprising words it contains (Genzel \& Charniak, 2002).
+The later a sentence occurs in a paragraph, the the informative it is and the more unusual and surprising words it contains [@genzel2002].
 From the level of individual words to sentences in paragraphs, information content shapes and is shaped by the context it appears in.
@@
 The pressure for efficient communication influences language at all timescales, from online speech production to evolution.
-Mahowald, Fedorenko, Piantadosi, \& Gibson (2013) found that speakers tend to use shorter, more efficient word forms in predictive contexts, minimizing communicative cost without sacrificing intelligibility.
-Predictability and ambiguity shape the distribution of color terms in languages (Gibson et al., 2017) as well as word lengths (Piantadosi et al., 2011) across languages, and speakers are able to deterministically decode words despite ambiguity in word forms (Piantadosi, Tily, \& Gibson, 2012).
+@mahowald2013 found that speakers tend to use shorter, more efficient word forms in predictive contexts, minimizing communicative cost without sacrificing intelligibility.
+Predictability and ambiguity shape the distribution of color terms in languages [@gibson2017] as well as word lengths [@piantadosi2011] across languages, and speakers are able to deterministically decode words despite ambiguity in word forms [@piantadosi2012].
 
@@
 Languages possess very different grammatical and morphological structures--the same thought can be encoded in at least as many ways as there are different languages.
-The average rate of information speakers transfer in syllables differs from language to language (Aylett \& Turk, 2004).
+The average rate of information speakers transfer in syllables differs from language to language [@aylett2004].
 How do the different structures seen in different languages, with different rates of information encoding, affect how speakers structure information in production?
 
-Yu, Cong, Liang, \& Liu (2016) proposed a novel method for quantifying the amount of information typically found at each part of English utterances.
-Treating each word position by itself, Yu et al.
-(2016) calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty (Shannon, 1948).
+@yu2016 proposed a novel method for quantifying the amount of information typically found at each part of English utterances.
+Treating each word position by itself, @yu2016 calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty [@shannon1948].
 They found a distinctive three-step shape to the distribution of information in English written sentences.
@@
 In this paper, we expand on this body of prior work in a number of novel ways.
-We replicate the results from Yu et al.
-(2016) with a different metric, one intimately tied to incremental word processing.
+We replicate the results from @yu2016 with a different metric, one intimately tied to incremental word processing.
 We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech.
@@
 We find that each language has a characteristic, robust frequency-based information distribution tied to the language's morphology, phonology and syntax.
-Speakers of a language will tend to distribute information according to that language's distribution of information in their utterances.
+Speakers of a language will tend to distribute information according to that language’s distribution of information in their utterances.
 As soon as a child can form multi-word utterances in a given language, they will produce utterances obeying the characteristic distribution of information in a given language.
 
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
+# Methods
 
-Shannon (1948) defined information as ``the reduction in uncertainty about one variable given that the value of another variable is known''.
-We use a metric proposed by Shannon and applied by Levy (2008): lexical surprisal, which defines the information in word based on the ratio of possible continuations of the sentence after to before the word is seen.
+@shannon1948 defined information as "the reduction in uncertainty about one variable given that the value of another variable is known".
+We use a metric proposed by Shannon and applied by @levy2008: lexical surprisal, which defines the information in word based on the ratio of possible continuations of the sentence after to before the word is seen.
 Equivalently, we can compute surprisal with the predictability of the word, as in the formula below.
@@
 
-\[surprisal(word) = -log P(word)\]
+$$surprisal(word) = -log P(word)$$
 
-The surprisal of a word is also correlated with the processing cost of a word, with evidence from eye-tracking (Smith \& Levy, 2013) and ERP (Frank, Otten, Galli, \& Vigliocco, 2015) studies, among other sources.
+The surprisal of a word is also correlated with the processing cost of a word, with evidence from eye-tracking [@smith2013] and ERP [@frank2015] studies, among other sources.
 Considered without context, the surprisal of an individual word is inversely proportional to the frequency of that word, so that the less often a person has seen a word, the more information that word holds.
-For example, ``flower'' has less information than ``azalea'' because ``flower'' is much more common than ``azalea''.
-Though the two words have the same length in number of letters, it's more difficult to process ``azalea'' when reading it here than when reading ``flower''.
-Frequency is intimately tied information content in words, with much of the differences between words frequencies being explained by information content cross-linguistically (Piantadosi et al., 2011).
+For example, “flower” has less information than “azalea” because “flower” is much more common than “azalea”.
+Though the two words have the same length in number of letters, it’s more difficult to process “azalea” when reading it here than when reading “flower”.
+Frequency is intimately tied information content in words, with much of the differences between words frequencies being explained by information content cross-linguistically [@piantadosi2011].
 
-However, when reading or listening, people don't just consider each word as an isolated linguistic signal.
+However, when reading or listening, people don’t just consider each word as an isolated linguistic signal.
 Instead, listeners use the words they have already heard to predict and decode the word they are currently processing.
@@
 
-\[surprisal(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...)\]
-\[= -log \frac{P(w_i,w_{i-1}w_{i-2},...)}{P(w_{i-1}w_{i-2}...)}\]
+$$surprisal(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...)$$
+$$= -log \frac{P(w_i,w_{i-1}w_{i-2},...)}{P(w_{i-1}w_{i-2}...)}$$
 
 When we use a word or two of context in our surprisal calculations, then the set of reasonable final items in our ngrams is greatly restricted.
-``Flower'' may contain less information than ``azalea'' when we consider the words independently of their context, but with context this can be reversed.
-Flower appears in a variety of contexts, and so the information content of a word like ``flower'' in a particular context may be higher than ``azalea''.
-If you only have azaleas in your garden, then hearing someone say ``in that garden, look at the flowers'' may be higher surprisal for you: you expect them to say ``azalea''.
-This prediction does not need many words before to work out, for example in ``I take my coffee with cream and sugar''.
-When you hear ``cream and'', you automatically predict ``sugar'' or maybe ``honey'', but there are few possible continuations with even those two words.
-Hearing ``I'' restricts the next word to a verb, or possibly an adverb, and since you have just heard the speaker refer to themselves in the first person singular, your set of possible completions is significantly restricted.
+“Flower” may contain less information than “azalea” when we consider the words independently of their context, but with context this can be reversed.
+Flower appears in a variety of contexts, and so the information content of a word like “flower” in a particular context may be higher than “azalea”.
+If you only have azaleas in your garden, then hearing someone say “in that garden, look at the flowers” may be higher surprisal for you: you expect them to say “azalea”.
+This prediction does not need many words before to work out, for example in “I take my coffee with cream and sugar”.
+When you hear “cream and”, you automatically predict “sugar” or maybe “honey”, but there are few possible continuations with even those two words.
+Hearing “I” restricts the next word to a verb, or possibly an adverb, and since you have just heard the speaker refer to themselves in the first person singular, your set of possible completions is significantly restricted.
 
@@
 
-The frequency-based surprisal metric gives us an idea of when in their utterances speakers say frequent i.e.~independently information-rich words.
+The frequency-based surprisal metric gives us an idea of when in their utterances speakers say frequent i.e.
+independently information-rich words.
 The context-based surprisal metric show us how speakers tend to distribute the information in utterances relative to real-time processing in communication.
@@
 
-\hypertarget{written-and-spoken-english-have-the-same-information-distribution}{%
-\section{Written and Spoken English have the same information distribution}\label{written-and-spoken-english-have-the-same-information-distribution}}
+# Written and Spoken English have the same information distribution
 
-We first turn to working with written English in the British National Corpus (BNC; Leech, 1992).
-The BNC is a collection of spoken and written records (90\% written) from the turn of the century, intended to be a representative sample of British English.
-Using their word entropy metric without context, Yu et al.~found a distinctive three-step distribution for information in written English sentences in the corpus.
+We first turn to working with written English in the British National Corpus [BNC; @leech1992].
+The BNC is a collection of spoken and written records (90% written) from the turn of the century, intended to be a representative sample of British English.
+Using their word entropy metric without context, Yu et al.
+found a distinctive three-step distribution for information in written English sentences in the corpus.
 The first word tended to contain little information.
@@
 
-We first replicate the Yu et al.
-(2016) result using the surprisal metric in place of the entropy metric.
-We use the frequency-based or ``contextless'' surprisal metric, which determines the average distribution of information based on word frequencies in a corpus.
+We first replicate the @yu2016 result using the surprisal metric in place of the entropy metric.
+We use the frequency-based or “contextless” surprisal metric, which determines the average distribution of information based on word frequencies in a corpus.
 A priori we expect that the frequency-based metric will produce a flat distribution of information across word positions in the BNC.
 
-See in Figure @ref(fig:bnc\_unigrams) a qualitative comparison of the information trajectory they found and the information trajectory we found.
-It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as Yu et al.
-(2016) did.
+See in Figure \@ref(fig:bnc_unigrams) a qualitative comparison of the information trajectory they found and the information trajectory we found.
+It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as @yu2016 did.
 The first words in written English sentences tend to contain little information, while the final words tend to contain a great deal of information.
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/bnc_unigrams-1.pdf}
-\caption{(\#fig:bnc\_unigrams)BNC frequency-based information curves}
-\end{figure}
+```{r bnc_unigrams, echo = F, fig.height=2, fig.cap="BNC frequency-based information curves"}
+bnc_results %>% 
+  filter(context == 0) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, 
+             ymin = ci_lower, ymax = ci_upper, color = as.factor(length))) + 
+    geom_pointrange() + 
+    geom_line() +
+    scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
 
@@
 The longer the utterance, the more pronounced the information distribution shape becomes.
-A qualitative comparison of the two trajectories is in Figure @ref(fig:switchboard\_unigrams).
+A qualitative comparison of the two trajectories is in Figure \@ref(fig:switchboard_unigrams).
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/switchboard_unigrams-1.pdf}
-\caption{(\#fig:switchboard\_unigrams)Switchboard frequency-based information curves}
-\end{figure}
+```{r switchboard_unigrams, echo=F, fig.height=4, fig.cap="Switchboard frequency-based information curves"}
+sw_plot <- switch_results %>% 
+  filter(context == 0) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = 1:11) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+
+bnc_plot <- bnc_results %>% 
+  filter(context == 0) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+
+grid.arrange(bnc_plot, sw_plot, nrow = 2)
+```
 
@@
 This distribution indicates that in English, the words we speak or write at the beginnings of utterances have little information, while the words we speak or write at the ends of utterances have a lot of information.
-The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from (Genzel \& Charniak, 2002).
+The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from [@genzel2002].
 
 What about context?
-So far we've only discussed the frequency-based metric, considering words on their own without any explicit incorporation of prior context.
+So far we’ve only discussed the frequency-based metric, considering words on their own without any explicit incorporation of prior context.
 As previously discussed, listeners decode information and process what they hear incrementally, using prior heard words to ease the comprehension process.
@@
 After the first word or two, where the listener does not have access to prior context, then they decode information at a flat and more or less uniform rate.
-The contextual information curves for the BNC and Switchboard are in Figures @ref(fig:switchboard\_bigrams) and @ref(fig:switchboard\_trigrams).
+The contextual information curves for the BNC and Switchboard are in Figures \@ref(fig:switchboard_bigrams) and \@ref(fig:switchboard_trigrams).
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/switchboard_bigrams-1.pdf}
-\caption{(\#fig:switchboard\_bigrams)Switchboard bigram context-based information curves}
-\end{figure}
+```{r switchboard_bigrams, echo=F,fig.height=4,fig.cap="Switchboard bigram context-based information curves"}
+sw_plot <- switch_results %>% 
+  filter(context == 1) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = 1:11) + 
+    xlab("Word position") + 
+    ylab("Mean information")
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/switchboard_trigrams-1.pdf}
-\caption{(\#fig:switchboard\_trigrams)Switchboard trigram context-based information curves}
-\end{figure}
+bnc_plot <- bnc_results %>% 
+  filter(context == 1) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+
+grid.arrange(bnc_plot, sw_plot, nrow = 2)
+```
+
+```{r switchboard_trigrams, echo=F,fig.height=4,fig.cap="Switchboard trigram context-based information curves"}
+sw_plot <- switch_results %>% 
+  filter(context == 2) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = 1:11) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+
+bnc_plot <- bnc_results %>% 
+  filter(context == 2) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+
+grid.arrange(bnc_plot, sw_plot, nrow = 2)
+```
 
@@
 
-\hypertarget{english-cds-and-child-speech}{%
-\section{English CDS and child speech}\label{english-cds-and-child-speech}}
+# English CDS and child speech
 
@@
 
-We use the Providence corpus from CHILDES (Evans \& Demuth, 2012; MacWhinney, 2000), which consists of about 650000 utterances from parent-child conversations for six monolingual English-speaking children talking with the parents.
-We obtained the Providence corpus using the childesr frontend to the childes-db database (Sanchez et al., 2019).
+We use the Providence corpus from CHILDES [@macwhinney2000; @evans2012], which consists of about 650000 utterances from parent-child conversations for six monolingual English-speaking children talking with the parents.
+We obtained the Providence corpus using the childesr frontend to the childes-db database [@sanchez2019].
 The child age range was between 11 months and 4 years.
@@
 Parents said a majority of the utterances, but children take up an increasing share of the conversation as they grow older and are actually able to speak full utterances.
-The utterances in the Providence corpus are on average significantly shorter than those in the BNC and Switchboard; over 90\% of the utterances in the Providence corpus are at most 10 words long, with the median at around 6 words.
+The utterances in the Providence corpus are on average significantly shorter than those in the BNC and Switchboard; over 90% of the utterances in the Providence corpus are at most 10 words long, with the median at around 6 words.
 Similar to our treatment of the Switchboard corpus, we split the Providence corpus up by conversational turns and separate utterances.
@@
 The distribution of information we found at the level of individual words in English, therefore, characterizes the English language as a whole and not only adult utterances, not only written utterances.
-See @ref(fig:providence\_unigrams).
+See \@ref(fig:providence_unigrams).
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/providence_unigrams-1.pdf}
-\caption{(\#fig:providence\_unigrams)Providence frequency-based information curves. Lines around each point indicate 95\% confidence intervals computed with non-parametric bootstrap}
-\end{figure}
+```{r providence_unigrams, echo=F, fig.height =2, fig.cap="Providence frequency-based information curves.
+Lines around each point indicate 95% confidence intervals computed with non-parametric bootstrap"}
+childes_results %>% 
+  filter(language == "English", context == 0) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, 
+             ymax = ci_upper, 
+             color = speaker)) + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = 1:9) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
 
@@
 This applies to parents and children speaking to one another, as well as adults speaking and writing to one another.
-See @ref(fig:providence\_bigrams) and @ref(fig:providence\_trigrams).
+See \@ref(fig:providence_bigrams) and \@ref(fig:providence_trigrams).
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/providence_bigrams-1.pdf}
-\caption{(\#fig:providence\_bigrams)Providence bigram context-based information curves}
-\end{figure}
+```{r providence_bigrams, echo=F, fig.height =2, fig.cap="Providence bigram context-based information curves"}
+childes_results %>% 
+  filter(language == "English", context == 1) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
+             color = speaker)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = 1:9) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/providence_trigrams-1.pdf}
-\caption{(\#fig:providence\_trigrams)Providence trigram context-based information curves}
-\end{figure}
+```{r providence_trigrams, echo=F, fig.height =2, fig.cap="Providence trigram context-based information curves"}
+childes_results %>% 
+  filter(language == "English", context == 2) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, 
+             ymin = ci_lower, ymax = ci_upper, 
+             color = speaker)) + 
+    geom_pointrange(position = position_dodge(.25)) + 
+    geom_line(position = position_dodge(.25)) +
+    facet_wrap(~length) + 
+    scale_x_continuous(breaks = 1:9) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
 
-\hypertarget{child-speech-and-cds-across-languages}{%
-\section{Child speech and CDS across languages}\label{child-speech-and-cds-across-languages}}
+# Child speech and CDS across languages
 
@@
 
-We now turn to a small number of typologically diverse languages, and conduct the same analysis, using a monolingual adult-child speech corpora from CHILDES (MacWhinney, 2000) to compare the results from these languages directly to our results from the English Providence corpus.
+We now turn to a small number of typologically diverse languages, and conduct the same analysis, using a monolingual adult-child speech corpora from CHILDES [@macwhinney2000] to compare the results from these languages directly to our results from the English Providence corpus.
 We use corpora for Spanish, German, French, Mandarin and Japanese.
 Similarly to the English Providence corpus, all of the following corpora consist mainly of shorter utterances: most utterances in the corpora are under 10 words long.
-For Spanish, we use the corpus of conversations between primary school-aged children between 6 and 9 years old and the researchers, in which the children were asked to recite stories in their native Spanish (Shiro, 1996).
+For Spanish, we use the corpus of conversations between primary school-aged children between 6 and 9 years old and the researchers, in which the children were asked to recite stories in their native Spanish [@shiro1996].
 The students came from working class public schools and upper class private schools.
-For German, we use the Wagner corpus (Wagner, 1985).
+For German, we use the Wagner corpus [@wagner1985].
 This corpus consists of a collection of mini-corpora collected from socio-economically diverse families, with children aged from 1;0 to 14;0, mainly on the younger end of that range.
 The corpus mainly consists of parent-child conversations between the children and their primary caregivers.
-For French, we use the Palasis corpus (Palasis, 2009), a longitudinal study of children between the ages of 2;5 and 4;0 from the same kindergarten class.
+For French, we use the Palasis corpus [@palasis2009], a longitudinal study of children between the ages of 2;5 and 4;0 from the same kindergarten class.
 The conversations in this corpus take place between the children and other children in the same class, and between the children and the investigator.
-For Mandarin Chinese, we use the Zhou corpus of dinner conversations from the Shanghai area (Li \& Zhou, 2015).
+For Mandarin Chinese, we use the Zhou corpus of dinner conversations from the Shanghai area [@li2015].
 The children in this corpus are 5 or 6 years old, and half come from working class families (the other half from upper class families).
-For Japanese, we use the Okayama corpus (Shirai, 2001).
+For Japanese, we use the Okayama corpus [@shirai2001].
 This corpus consists of conversations between children between the ages of 2;2 and 4;2 and their mothers.
@@
 We observe a distinct and characteristic frequency-based information trajectory emerge for each language, robust across each utterance length within each language.
-We see the same distribution of information for both parents and children in each language, with the child's curve normally below the parent's curve, likely due to the parent speaking a larger share of the utterances in the corpus and using a larger vocabulary than the child.
+We see the same distribution of information for both parents and children in each language, with the child’s curve normally below the parent’s curve, likely due to the parent speaking a larger share of the utterances in the corpus and using a larger vocabulary than the child.
 We see the opposite amplitude trend in the Shiro corpus, where the children speak more than the adult investigators as the children are the ones telling the stories.
@@
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
-\caption{\label{fig:unnamed-chunk-1}CHILDES frequency-based information curves}
-\end{figure}
+```{r, echo=F, fig.height =3, fig.cap="CHILDES frequency-based information curves"}
+childes_results %>% 
+  filter(context == 0, length == 7) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
+             color = speaker)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~language) + 
+    scale_x_continuous(breaks = 1:7) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
 
@@
 While the frequency-based information curves may depend based on the language, the contextual information curves show the same trajectory cross-linguistically.
-Using more than two words of context is difficult for parent-child speech corpora because the utterances are so short on average (less than \(10\) words).
+Using more than two words of context is difficult for parent-child speech corpora because the utterances are so short on average (less than $10$ words).
 We hypothesize that the frequency-based information curves may vary based on the genealogy and typology of the languages in question, but this does not extend to the information curves with two words of context in particular.
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
-\caption{\label{fig:unnamed-chunk-2}CHILDES bigram context-based information curves}
-\end{figure}
+```{r, echo=F, fig.height =3, fig.cap="CHILDES bigram context-based information curves"}
+childes_results %>% 
+  filter(context == 1, length == 7) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
+             color = speaker)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~language) + 
+    scale_x_continuous(breaks = 1:7) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
-\caption{\label{fig:unnamed-chunk-3}CHILDES trigram context-based information curves}
-\end{figure}
+```{r, echo=F, fig.height =3, fig.cap="CHILDES trigram context-based information curves"}
+childes_results %>% 
+  filter(context == 2, length == 7) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
+             color = speaker)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~language) + 
+    scale_x_continuous(breaks = 1:7) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+```
 
-\hypertarget{language-structure-and-large-scale-data-analysis-methods-and-preprocessing}{%
-\section{Language structure and large-scale data analysis: methods and preprocessing}\label{language-structure-and-large-scale-data-analysis-methods-and-preprocessing}}
+# Language structure and large-scale data analysis: methods and preprocessing
 
 To make a claim about language as a whole and languages on a larger scale, we needed to use larger corpora and a much larger number of languages.
-We pulled corpora for \(159\) diverse languages from Wikipedia, each of which had at least \(10000\) articles on the knowledge base.
+We pulled corpora for $159$ diverse languages from Wikipedia, each of which had at least $10000$ articles on the knowledge base.
 We split each article up into sentences; the variance in sentence lengths for Wikipedia was significantly larger than for the CHILDES corpora we used in the previous section.
-Most sentences in Wikipedia contained between \(10\) and \(30\) words, unlike the CHILDES corpora which mainly contained utterances with under 10 words.
-We excluded the small fraction of utterances with more than \(50\) words since they were small in number and, from manual inspection, uncharacteristic of typical written sentences.
+Most sentences in Wikipedia contained between $10$ and $30$ words, unlike the CHILDES corpora which mainly contained utterances with under 10 words.
+We excluded the small fraction of utterances with more than $50$ words since they were small in number and, from manual inspection, uncharacteristic of typical written sentences.
 
-How do we analyze more than \(40\) different surprisal curves for each language, adding up to several thousand surprisal curves total?
+How do we analyze more than $40$ different surprisal curves for each language, adding up to several thousand surprisal curves total?
 We used two different strategies, which yielded identical results upon analysis.
-Each strategy gave us a five-dimensional vector space embedding for each language in a Wikipedia ``slope space''.
+Each strategy gave us a five-dimensional vector space embedding for each language in a Wikipedia “slope space”.
 For the first strategy, we split each sentence length by number of words into fifths, and computed surprisal values for the closest word position to each quintile.
@@
 We then similarly computed surprisal values at each of these positions, and similarly computed slopes between the surprisal values at each position, giving us another five slope values for each language summarizing the surprisal curves.
-Illustrations of these two strategies are in Figure \ref{fig:treatments}
+Illustrations of these two strategies are in Figure \@ref(fig:treatments)
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/treatments-1.pdf}
-\caption{\label{fig:treatments}Illustration of slope treatments for Wikipedia information curves: relative on top and absolute on bottom}
-\end{figure}
+```{r treatments, echo = F, fig.height =3, fig.cap="Illustration of slope treatments for Wikipedia information curves: relative on top and absolute on bottom"}
+## relative treatment: every fifth in a different group
+relative <- bnc_results %>% 
+  filter(context == 0, length == 20) 
+relative$group <- rep(1:5, each = 4)
+relative <- relative %>%
+  mutate(group = as.factor(group))
+
+# rel_plot <- relative %>%
+#   ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, color = group)) + 
+#     geom_point() + 
+#     geom_pointrange() + 
+#     geom_line() +
+#     scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
+#     xlab("Word position") + 
+#     ylab("Mean information")
+
+## absolute treatment: some qualitatively important sections
+absolute <- bnc_results %>% 
+  filter(context == 0, length == 20) 
+absolute$group <- c(1, 2, rep(3, 16), 4, 5) 
+absolute <- absolute %>%
+  mutate(group = as.factor(group))
+
+relative %>% 
+  mutate(treatment = "relative") %>% 
+  bind_rows(absolute %>% mutate(treatment = "absolute")) %>%
+  ggplot(aes(x = word_order, y = empirical_stat, 
+             ymin = ci_lower, ymax = ci_upper, color = group)) + 
+    geom_point() + 
+    geom_pointrange() + 
+    geom_line() +
+    facet_wrap(~treatment) + 
+    scale_x_continuous(breaks = c(1, seq(5, 30, 5))) + 
+    xlab("Word position") + 
+    ylab("Mean information")
+
+# grid.arrange(rel_plot, abs_plot, nrow = 2)
+
+```
 
@@
 
-To more rigorously described the differences between languages, we used data from the World Atlas of Language Structures (WALS; Dryer \& Haspelmath, 2013).
-WALS has data for \(144\) typological features in \(2569\) languages around the world.
+To more rigorously described the differences between languages, we used data from the World Atlas of Language Structures [WALS; @wals].
+WALS has data for $144$ typological features in $2569$ languages around the world.
 These features describe aspects of morphology, syntax, phonology, etymology and semantics--in short the features describe the structures in each language.
 As WALS is a compiled database from dozens of papers from different authors, most of the features and languages are fairly sparse.
-Even limiting ourselves to the \(159\) language corpora we pulled from Wikipedia and \(122\) features from WALS, there are nearly \(20000\) individual possible feature values, fewer than half of which were already inputted for those languages in the WALS database.
+Even limiting ourselves to the $159$ language corpora we pulled from Wikipedia and $122$ features from WALS, there are nearly $20000$ individual possible feature values, fewer than half of which were already inputted for those languages in the WALS database.
 
-We used the Multiple Imputation Multiple Correspondence Analysis algorithm (MIMCA; Audigier, Husson, \& Josse, 2017) to fill in the missing data using statistical imputation.
+We used the Multiple Imputation Multiple Correspondence Analysis algorithm [MIMCA; @audigier2017] to fill in the missing data using statistical imputation.
 MIMCA essentially uses mean imputation to begin with the missing values, then performs repeatedly performs principle components analysis on and reconstructs the contingency table formed from observations in categorical variables.
-By the end of this we had frequency-based and contextual information curves for the \(159\) language corpora pulled from Wikipedia, along with \(122\) typological features for each language.
+By the end of this we had frequency-based and contextual information curves for the $159$ language corpora pulled from Wikipedia, along with $122$ typological features for each language.
 
 However, the WALS features describe specific structural differences between languages, while our surprisal metric is lexical.
-To target lexical differences between languages, we computed the average normalized Levenshtein distance (LDN; Holman et al., 2008) over the \(40\) item Swadesh list (Swadesh, 1955), retrieved from the ASJP database (Wichmann et al., 2016).
+To target lexical differences between languages, we computed the average normalized Levenshtein distance [LDN; @holman2008] over the $40$ item Swadesh list [@swadesh1955], retrieved from the ASJP database [@wichmann2016].
 The Swadesh list is designed to include near-universal words that target basic cognitive concepts, and are useful in determining the genealogical similarities and differences between languages.
-The results of classifying languages using the Swadesh list and LDN are correlated with those using WALS features, but the Swadesh list and LDN do not suffer from the same sparsity problem as WALS (Holman et al., 2008).
+The results of classifying languages using the Swadesh list and LDN are correlated with those using WALS features, but the Swadesh list and LDN do not suffer from the same sparsity problem as WALS [@holman2008].
 
-\hypertarget{language-structures-and-large-scale-data-analysis-results}{%
-\section{Language structures and large-scale data analysis: results}\label{language-structures-and-large-scale-data-analysis-results}}
+# Language structures and large-scale data analysis: results
 
-We ran a hierarchical clustering algorithm on the unigram information curves using the hclust package from the R stats core library (Team \& others, 2013).
+We ran a hierarchical clustering algorithm on the unigram information curves using the hclust package from the R stats core library [@team2013].
 We used the complete linkage algorithm for hierarchical clustering with distances between information curves in language computed using cosine distance between their embeddings in the slope space.
 The complete linkage algorithm at every step pairs each language or cluster of languages with its closest neighboring language or cluster.
-A sample from the dendrogram is shown in Figure \ref{fig:dendro}, and the full tree can be constructed using the code in our GitHub repository.
-The language family-like structure can be seen at first glance, although the dendrogram does not exactly replicate language genealogy for all \(159\) languages.
+A sample from the dendrogram is shown in Figure \@ref(fig:dendro), and the full tree can be constructed using the code in our GitHub repository.
+The language family-like structure can be seen at first glance, although the dendrogram does not exactly replicate language genealogy for all $159$ languages.
 This suggests using a first-pass quantitative method that the information curves do correspond in some measure to language families, but language families do not explain the entire story.
 
-\includegraphics{paper_files/figure-latex/dendro-1.pdf}
+```{r dendro, echo=F}
+wiki_curves %>%
+  filter(gram == "Unigram") %>%
+  select(-gram) %>% 
+  sample_n(40) %>%
+  column_to_rownames("language") %>%
+  dist() %>% 
+  hclust() %>%
+  ggdendrogram(rotate=T, leaf_labels = F)
+```
 
-A sample of the contextual information curves (two words of context) are plotted in Figure @ref(fig:wiki\_trigrams), and all trigram information curves for the languages we used follow the same pattern.
+A sample of the contextual information curves (two words of context) are plotted in Figure \@ref(fig:wiki_trigrams), and all trigram information curves for the languages we used follow the same pattern.
 The first few words in utterances for each language are surprising, but with even a couple words of predictive context for each word, the amount of information in each word drops off and speakers produce information at a constant, optimal rate.
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/wiki_trigrams-1.pdf}
-\caption{(\#fig:wiki\_trigrams)Some trigram information curves from the Wikipedia data}
-\end{figure}
+```{r wiki_trigrams, echo=F, fig.cap="Some trigram information curves from the Wikipedia data"}
+wiki_curves %>%
+  filter(gram == "Trigram") %>% 
+  mutate(Slope0 = runif(1, 0.9, 1.1)) %>% 
+  mutate(Slope1 = Slope1 + Slope0) %>% 
+  mutate(Slope2 = Slope2 + Slope1) %>% 
+  mutate(Slope3 = Slope3 + Slope2) %>% 
+  mutate(Slope4 = Slope4 + Slope3) %>% 
+  mutate(Slope5 = Slope5 + Slope4) %>% 
+  select(-gram) %>% 
+  sample_n(6) %>% 
+  gather(position, information, -language) %>%
+  ggplot(aes(x = position, y = information, group = language, color = language)) + 
+    geom_point() + 
+    geom_line() + 
+    xlab("Word position") + 
+    ylab("Average Information") + 
+    theme(axis.text.x = element_blank())
+```
 
 First we examined the effects of individual typological features on the shapes of the unigram information curves.
-We ran logistic regressions using the lme4 package in R (Bates, Mächler, Bolker, \& Walker, 2014), checking whether the cosine distance between two languages' embeddings in the slope space played a role in determining if they had the same value for a given WALS feature.
+We ran logistic regressions using the lme4 package in R [@bates2014], checking whether the cosine distance between two languages’ embeddings in the slope space played a role in determining if they had the same value for a given WALS feature.
 Individual WALS features do not necessarily have ordinal values.
-Some, such as the ``Number of Cases'' feature, are easy to quantify and order.
+Some, such as the “Number of Cases” feature, are easy to quantify and order.
 Others are more difficult.
-For example, how does one order ``relative clauses appear after the nouns they modify'', ``relative clauses appear before the nouns they modify'' and ``free order of relative clauses and nouns''?
+For example, how does one order “relative clauses appear after the nouns they modify”, “relative clauses appear before the nouns they modify” and “free order of relative clauses and nouns”?
 We chose the identify relation to avoid deciding on the basis of individual features.
-We found that 100 out of the 120 features from WALS we examined was statistically significant (p \textless{} .001) in determining whether two languages had the same shape to their unigram information distributions.
-The results for some important features are in Figure @ref(fig:linear\_models), and the rest of the results are in the Appendices.
+We found that 100 out of the 120 features from WALS we examined was statistically significant (p < .001) in determining whether two languages had the same shape to their unigram information distributions.
+The results for some important features are in Figure \@ref(fig:linear_models), and the rest of the results are in the Appendices.
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/linear_models-1.pdf}
-\caption{(\#fig:linear\_models)Some linear model results from Wikipedia and WALS features}
-\end{figure}
+```{r linear_models, echo=F, fig.cap="Some linear model results from Wikipedia and WALS features"}
+sf <- wiki_lr %>% 
+  filter(p.value == 0, estimate > 0) %>% 
+  select(name) %>% 
+  rename("Some Important Features" = name) %>% 
+  slice(c(1, 10, 12, 15, 23, 29))
+
+sf$Examples <- c("Noun-RC; RC-noun", 
+                 "No cases; 6 cases", 
+                 "SOV; VSO", 
+                 "Only singular; sing. and plural", 
+                 "Affix; distinct word", 
+                 "Prefixes; suffixes")
+
+sf %>% grid.table()
+```
 
 We next compared how the cosine distance between two languages related to how many WALS features they had in common.
-\(r^2\) value is \(.005734\), which suggests that in aggregate there is not a correlation between how many WALS features languages have in common and the similarity of their frequency-based information curves.
-Figure @ref(fig:cosine\_wals) displays the results.
+$r^2$ value is $.005734$, which suggests that in aggregate there is not a correlation between how many WALS features languages have in common and the similarity of their frequency-based information curves.
+Figure \@ref(fig:cosine_wals) displays the results.
 This result is surprising based on the significance of many WALS features in predicting the shapes of the frequency-based information curves, and we return to this result in the general discussion.
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/cosine_wals-1.pdf}
-\caption{(\#fig:cosine\_wals)wals features vs cosine similarity}
-\end{figure}
+```{r cosine_wals, echo = F, fig.cap="wals features vs cosine similarity"}
+r2 <- summary(lm(data = wiki_results, formula = unigram_cosine~feature))$r.squared %>%
+  round(digits = 3)
+wiki_results %>% 
+  ggplot(aes(x = feature, y=unigram_cosine)) + 
+    geom_smooth(aes(weight = .01), se=F, method = "lm", color = "black") + 
+    geom_jitter(alpha = 0.04, color = "blue") + 
+    annotate("text", x = 70, y = -0.5, label = paste0("r^2 = ", r2)) + 
+    xlab("Number of identical WALS features") + 
+    ylab("Information curve similarity")
+    #stat_poly_eq(formula = y~x, position = "dodge")
+```
 
 For lexical features, we see a stronger correlation between the similarity of two languages in terms of their average LDN and the cosine distance between their information curves.
-Figure @ref(fig:cosine\_ldn).
-We see a higher \(r^2\) value here of \(.026\), indicating that there is more correspondence between a language's lexical similarity to another language and their similarity in information curves.
+Figure \@ref(fig:cosine_ldn).
+We see a higher $r^2$ value here of $.026$, indicating that there is more correspondence between a language's lexical similarity to another language and their similarity in information curves.
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/cosine_ldn-1.pdf}
-\caption{(\#fig:cosine\_ldn)ldn features vs cosine similarity}
-\end{figure}
+```{r cosine_ldn, echo = F, fig.cap="ldn features vs cosine similarity"}
+r2 <- summary(lm(data = wiki_results, formula = unigram_cosine~ldn))$r.squared %>%
+  round(digits = 4)
+wiki_results %>% 
+  ggplot(aes(x = ldn, y=unigram_cosine)) + 
+    geom_smooth(aes(weight = .01), se=F, method = "lm", color = "black") + 
+    geom_jitter(alpha = 0.04, color = "blue") + 
+    annotate("text", x = 0.2, y = -0.5, label = paste0("r^2 = ", r2)) + 
+    xlab("Normalized lexical distance") + 
+    ylab("Information curve similarity")
+    #stat_poly_eq(formula = y~x, position = "dodge")
+```
 
-From these typological investigations, we conclude that the shape of a language's frequency-based information curve covaries with its typological and lexical similarity to other languages.
+From these typological investigations, we conclude that the shape of a language’s frequency-based information curve covaries with its typological and lexical similarity to other languages.
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+# Discussion
 
@@
 The WALS database we used to investigate typological variation in the information curves is overall sparse.
-We imputed well over \(50\%\) of the WALS features for most of our \(159\) languages, although all of the languages had at least \(20\) features evaluated in WALS.
+We imputed well over $50\%$ of the WALS features for most of our $159$ languages, although all of the languages had at least $20$ features evaluated in WALS.
 A large part of this is due to WALS being a collection of a number of different studies, instead of a systematic effort to catalogue variation across the world's languages.
 Additionally, WALS features are meant to describe specific microvariations in languages, not to provide a comprehensive typological representation of each language compared to each other language.
-This may be why the Swadesh list provided a higher correlation for describing the differences in information curves: Swadesh (1955) intended the list to allow researchers to more comprehensively compared and constrast lexical differences between languages.
+This may be why the Swadesh list provided a higher correlation for describing the differences in information curves: @swadesh1955 intended the list to allow researchers to more comprehensively compared and constrast lexical differences between languages.
 For our Wikipedia analysis, we also reduce all of a language's variation down to a five-dimensional vector.
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
+\newpage
 
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-audigier2017mimca}{}%
-Audigier, V., Husson, F., \& Josse, J. (2017). MIMCA: Multiple imputation for categorical variables with multiple correspondence analysis. \emph{Statistics and Computing}, \emph{27}(2), 501--518.
+# References
 
-\leavevmode\hypertarget{ref-aylett2004smooth}{}%
-Aylett, M., \& Turk, A.
-(2004).
-The smooth signal redundancy hypothesis: A functional explanation for relationships between redundancy, prosodic prominence, and duration in spontaneous speech.
-\emph{Language and Speech}, \emph{47}(1), 31--56.
+\begingroup
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
 
-\leavevmode\hypertarget{ref-bates2014fitting}{}%
-Bates, D., Mächler, M., Bolker, B., \& Walker, S. (2014). Fitting linear mixed-effects models using lme4. \emph{arXiv Preprint arXiv:1406.5823}.
-
-\leavevmode\hypertarget{ref-wals}{}%
-Dryer, M. S., \& Haspelmath, M. (Eds.). (2013). \emph{WALS online}. Retrieved from \url{https://wals.info/}
-
-\leavevmode\hypertarget{ref-evans2012individual}{}%
-Evans, K. E., \& Demuth, K. (2012). Individual differences in pronoun reversal: Evidence from two longitudinal case studies. \emph{Journal of Child Language}, \emph{39}(1), 162--191.
-
-\leavevmode\hypertarget{ref-frank2015erp}{}%
-Frank, S. L., Otten, L. J., Galli, G., \& Vigliocco, G. (2015). The erp response to the amount of information conveyed by words in sentences. \emph{Brain and Language}, \emph{140}, 1--11.
-
-\leavevmode\hypertarget{ref-genzel2002entropy}{}%
-Genzel, D., \& Charniak, E. (2002). Entropy rate constancy in text. \emph{Proceedings of the 40th annual meeting of the association for computational linguistics}, 199--206.
-
-\leavevmode\hypertarget{ref-gibson2017color}{}%
-Gibson, E., Futrell, R., Jara-Ettinger, J., Mahowald, K., Bergen, L., Ratnasingam, S., \ldots{} Conway, B.
-R.
-(2017).
-Color naming across languages reflects color use.
-\emph{Proceedings of the National Academy of Sciences}, \emph{114}(40), 10785--10790.
-
-\leavevmode\hypertarget{ref-holman2008advances}{}%
-Holman, E.
-W., Wichmann, S., Brown, C.
-H., Velupillai, V., Müller, A., Bakker, D., \& others.
-(2008).
-Advances in automated language classification.
-\emph{Quantitative Investigations in Theoretical Linguistics}, 40--43.
-
-\leavevmode\hypertarget{ref-jaeger2007speakers}{}%
-Jaeger, T. F., \& Levy, R. P. (2007). Speakers optimize information density through syntactic reduction. \emph{Advances in neural information processing systems}, 849--856.
-
-\leavevmode\hypertarget{ref-kutas2011thirty}{}%
-Kutas, M., \& Federmeier, K.
-D.
-(2011).
-Thirty years and counting: Finding meaning in the n400 component of the event-related brain potential (erp).
-\emph{Annual Review of Psychology}, \emph{62}, 621--647.
-
-\leavevmode\hypertarget{ref-leech1992100}{}%
-Leech, G. N. (1992). \emph{100 million words of english: The british national corpus (bnc)}.
-
-\leavevmode\hypertarget{ref-levy2008expectation}{}%
-Levy, R. (2008). Expectation-based syntactic comprehension. \emph{Cognition}, \emph{106}(3), 1126--1177.
-
-\leavevmode\hypertarget{ref-li2015dinner}{}%
-Li, H., \& Zhou, J. (2015). \emph{Study on dinner table talk of preschool children family in shanghai} (Master's thesis). East China Normal University, Shanghai, China.
-
-\leavevmode\hypertarget{ref-macwhinney2000childes}{}%
-MacWhinney, B. (2000). \emph{The childes project: The database} (Vol. 2). Psychology Press.
-
-\leavevmode\hypertarget{ref-mahowald2013info}{}%
-Mahowald, K., Fedorenko, E., Piantadosi, S. T., \& Gibson, E. (2013). Info/information theory: Speakers choose shorter words in predictive contexts. \emph{Cognition}, \emph{126}(2), 313--318.
-
-\leavevmode\hypertarget{ref-palasis2009syntaxe}{}%
-Palasis, K. (2009). \emph{Syntaxe générative et acquisition: Le sujet dans le développement du système linguistique du jeune enfant} (PhD thesis). Nice.
-
-\leavevmode\hypertarget{ref-piantadosi2011word}{}%
-Piantadosi, S. T., Tily, H., \& Gibson, E. (2011). Word lengths are optimized for efficient communication. \emph{Proceedings of the National Academy of Sciences}, \emph{108}(9), 3526--3529.
-
-\leavevmode\hypertarget{ref-piantadosi2012communicative}{}%
-Piantadosi, S. T., Tily, H., \& Gibson, E. (2012). The communicative function of ambiguity in language. \emph{Cognition}, \emph{122}(3), 280--291.
-
-\leavevmode\hypertarget{ref-sanchez2019childes}{}%
-Sanchez, A., Meylan, S.
-C., Braginsky, M., MacDonald, K.
-E., Yurovsky, D., \& Frank, M.
-C.
-(2019).
-Childes-db: A flexible and reproducible interface to the child language data exchange system.
-\emph{Behavior Research Methods}, \emph{51}(4), 1928--1941.
-
-\leavevmode\hypertarget{ref-shannon1948mathematical}{}%
-Shannon, C. E. (1948). A mathematical theory of communication. \emph{Bell System Technical Journal}, \emph{27}(3), 379--423.
-
-\leavevmode\hypertarget{ref-shirai2001acquisition}{}%
-Shirai, Y. (2001). The acquisition of causative morphology in japanese: A prototype account. \emph{East Asian Language Acquisition}.
-
-\leavevmode\hypertarget{ref-de1996estudio}{}%
-Shiro, M. L. de. (1996). Un estudio de las expresiones de modalidad en hablantes de dos culturas. \emph{Boletı'n de Lingüı'stica}, (10), 43--60.
-
-\leavevmode\hypertarget{ref-smith2013effect}{}%
-Smith, N. J., \& Levy, R. (2013). The effect of word predictability on reading time is logarithmic. \emph{Cognition}, \emph{128}(3), 302--319.
-
-\leavevmode\hypertarget{ref-swadesh1955towards}{}%
-Swadesh, M. (1955). Towards greater accuracy in lexicostatistic dating. \emph{International Journal of American Linguistics}, \emph{21}(2), 121--137.
-
-\leavevmode\hypertarget{ref-team2013r}{}%
-Team, R. C., \& others. (2013). \emph{R: A language and environment for statistical computing}.
-
-\leavevmode\hypertarget{ref-wagner1985much}{}%
-Wagner, K. R. (1985). How much do children say in a day? \emph{Journal of Child Language}, \emph{12}(2), 475--487.
-
-\leavevmode\hypertarget{ref-wichmann2016asjp}{}%
-Wichmann, S., Müller, A., Wett, A., Velupillai, V., Bischoffberger, J., Brown, C. H., \ldots{} others. (2016). The asjp database. \emph{Max Planck Institute for the Science of Human History, Jena}.
-
-\leavevmode\hypertarget{ref-yu2016distribution}{}%
-Yu, S., Cong, J., Liang, J., \& Liu, H. (2016). The distribution of information content in english sentences. \emph{arXiv Preprint arXiv:1609.07681}.
-
-\leavevmode\hypertarget{ref-zipf1949human}{}%
-Zipf, G. K. (1949). \emph{Human behavior and the principle of least effort.}
-
-\end{document}
+<div id = "refs"></div>
+\endgroup

### 2019-09-24 Paper/info-curves.Rmd
few more sentences

@@
 The solution that languages have arrived on is incremental processing: Listeners process speech continuously as it arrives, predicting upcoming words and building expectations about the likely meaning of utterances in real-time rather than at their conclusion [@kutas2011; @tanenhaus1995; @pickering2013].
-Because prediction errors can lead to severe processing costs and difficulty integrating new information, an optimal coding strategy is for speakers seeking to minimize their prodction costs is to produce utterances that are are just at the prediction capacity of listeners without exceeding this capacity [@aylett2004; @genzel2002].
+Because prediction errors can lead to severe processing costs and difficulty integrating new information, speakers should seek to minimize prediction errors.
+However, the cost of producing more predictable utterances is using more words.
+Thus, an optimal coding strategy is for speakers seeking to minimize their production costs is to produce utterances that are are just at the prediction capacity of listeners without exceeding this capacity [@aylett2004; @genzel2002].
 
-One salient constraint on this transmission processes is the serial nature of language:
+Using information theory--a mathematical framework for formalizing predictability--researchers have tested and confirmed this general prediction of optimal coding across levels and contexts of language production.
+For instance, @genzel2002 provided a clever indirect test of this hypothesis at the level of sentences.
+They showed that the predictability of successive sentences in a discourse--analyzed in isolation--decreases, as would be expected if readers use prior sentences to predict the content of future sentences and total predictability remains constant.
 
-This coding perspective is powerful because it lends itself naturally to a rat
+the predicted if the amount of information is constant readers use prior sentences in a discourse to predict the content of future setences as they read.
 
-Languages can thus be thought of as a code that can be used to transform speaker's intented meaning into a
+This prediction of optimal coding--that speakers should maintain a constant rate of predictability has been tested and confirmed in a variety of ways across levels of language.
 
@@
 * BUT variation cross-linguistically. why? constraints from language.
-
-We use language to communicate information.
-Although we often need to say things we have never said before, we are easily and quickly able to produce novel utterances with next to no prior planning.
-Our listeners incrementally process each word we say in utterances they may have never heard before, and predict which words will come next within fractions of a second of hearing the beginning of the word [for evidence see e.g.
-@kutas2011].
-Encoding new and unpredictable information into our utterances surprises our listeners, and presents problems for communication.
-Listeners may struggle to decode unpredictable words, which can cause difficulty decoding the following words as well.
-Efficient and effective communication relies on interlocutors being able to process novel information without being surprised by new information in unpredictable words.
-On the other hand, to make best use of the capacity of the communication channel, speakers should encode as much information as possible into each word they say.
 

### 2019-10-02 Paper/info-curves.Rmd
introduction edits and smoothing in methods added to paper

@@
     address       : "5000 Forbes Ave"
-    email         : "klafka@cmu.edu"
-  - name          : "Daniel Yurovksy"
+    email         : "jklafka@andrew.cmu.edu"
+  - name          : "Daniel Yurovsky"
     affiliation   : "1,2"
@@
 One of the defining features of human language is its power to transmit information.
-We use language for a variety of different proximal tasks tasks: greeting friends, taking notes, signalling group identities.
+We use language for a variety of different tasks: greeting friends, taking notes, signaling group identities.
 But all share a common unifying purpose: changing the mental state of the listener or reader of our communications [@austin1975].
-Language can thus naturally thought of as a code that allows speakers to turn their intended meanings into a form that can be transmitted through the air (or paper, or electrons) and subsequently converted by listeners back into an approximation of the intended meaning [@shannon1948].
+Language can thus naturally thought of as a code that allows speakers to turn their intended meanings into a message that can be transmitted through the air (or paper, or electrons) and subsequently converted by listeners back into an approximation of the intended meaning [@shannon1948].
 
-Beyond its utility as a metaphor, this coding perspective on language is powerful because it affords a framework for rational analysis: If language has evolved to be an optimal code for information transmission, how should this code be structured [@anderson1989]?
-The optimal code would have to optimize two competing pressures: (1) a pressure for listeners to easily and successfully decode messages sent by the speaker, and (2) a pressure for speakers to easily code their messages and transmit them with minimal effort and error.
-A fundamental constriaint on both of these processes is the serial nature of human language: sounds are produced one at a time and each is unavailable perceptually once it is no longer being produced.
+Beyond its utility as a metaphor, this coding perspective on language is powerful because it allows a framework for rational analysis: If language has evolved to be an optimal code for information transmission, how should this code be structured [@anderson1989]?
+The optimal code would have to work with two competing pressures: (1) a pressure for listeners to easily and successfully decode messages sent by the speaker, and (2) a pressure for speakers to easily code their messages and transmit them with minimal effort and error.
+A fundamental constriaint on both of these processes is the linear order of spoken language: sounds are produced one at a time and each is unavailable perceptually once it is no longer being produced.
 
-The solution that languages have arrived on is incremental processing: Listeners process speech continuously as it arrives, predicting upcoming words and building expectations about the likely meaning of utterances in real-time rather than at their conclusion [@kutas2011; @tanenhaus1995; @pickering2013].
+The strategic solution people employ is *incremental processing*.
+Listeners process speech continuously as it arrives, predicting upcoming words and building expectations about the likely meaning of utterances in real-time rather than at their conclusion [@kutas2011; @tanenhaus1995; @pickering2013].
 Because prediction errors can lead to severe processing costs and difficulty integrating new information, speakers should seek to minimize prediction errors.
@@
 
-Using information theory--a mathematical framework for formalizing predictability--researchers have tested and confirmed this general prediction of optimal coding across levels and contexts of language production.
-For instance, @genzel2002 provided a clever indirect test of this hypothesis at the level of sentences.
+Using information theory--a mathematical framework for formalizing predictability--researchers have tested and confirmed this general prediction of optimal coding across several levels and contexts of language production.
+For instance, @genzel2002 provided a clever indirect test of this hypothesis across sentences.
 They showed that the predictability of successive sentences in a discourse--analyzed in isolation--decreases, as would be expected if readers use prior sentences to predict the content of future sentences and total predictability remains constant.
 
-the predicted if the amount of information is constant readers use prior sentences in a discourse to predict the content of future setences as they read.
+This prediction of optimal coding--that speakers should maintain a constant rate of predictability has been tested and confirmed in a variety of ways across levels of language.
+At the level of individual words, @mahowald2013 showed that speakers use shorter words in more predictive contexts and prefer the shorter versions of words in more predictive contexts, maximizing the amount of information in each word while minimizing the time spent on those words.
+Over time, @piantadosi2011 showed that shorter words across languages tend to be more easily predictable in context and thus less informative, over and above frequency.
 
-This prediction of optimal coding--that speakers should maintain a constant rate of predictability has been tested and confirmed in a variety of ways across levels of language.
+Efficient encoding also impacts how speakers structure phrases.
+The inclusion of complementizers in relative clauses [@jaeger2007] and the use of contractions [@frank2008] are two situations in sentence formation in which speakers use structure to communicate efficiently, spreading more information across a larger amount of linguistic material.
+
+How these pressures shape language evolution??
+
+However, despite this literature using the predictive coding model of language, one level has not yet been studied in depth: how speakers structure individual utterances.
+This level may show the strongest effects of variation between languages, as specific languages have properties that constrain how speakers may form utterances in those languages, such as canonical word order.
+These properties vary widely from language to language.
+
+@yu2016 studied the utterance level in written English sentences using a contextless entropy model based on word frequency.
+They found a distinctive three-step distribution regardless of sentence length, with little information in the first words of sentences and the most information in the final words.
+This was surprising, as the distribution they found was robustly different from the linearly increasing trend in sentences from @genzel2002, and also did not resemble the uniform distribution of information that one might expect from a communicative efficiency account, in which each word has approximately equal information close to the channel capacity.
+
+In this paper, we expand on this body of prior work in a number of novel ways.
+We replicate the results from @yu2016 with a metric tied to incremental word processing.
+We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech.
+We also apply a contextual version of the same metric, and find that in communication, listeners are able to decode information from the utterances they hear at a constant rate past the initial words.
+We expand the study of information density to the largest set of languages yet considered, and incorporate contextual and typological information.
+We find that each language has a characteristic, robust frequency-based information distribution tied to the language's morphology, phonology and syntax.
+Speakers of a language will tend to distribute information according to that language’s distribution of information in their utterances.
+As soon as a child can form multi-word utterances in a given language, they will produce utterances obeying the characteristic distribution of information in a given language.
 
@@
 * BUT variation cross-linguistically. why? constraints from language.
-
-Due to their differences in distributional frequency, words vary in how much information content they contain [@shannon1948].
-Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be [@zipf1949].
-Over and above frequency, contextual information content predicts word length, due to speakers minimizing effort in the amount of information they encode in speech [@piantadosi2011].
-Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side [@jaeger2007].
-Across utterances, context allows speakers to produce more informative and surprising sentences later in communication.
-The later a sentence occurs in a paragraph, the the informative it is and the more unusual and surprising words it contains [@genzel2002].
-From the level of individual words to sentences in paragraphs, information content shapes and is shaped by the context it appears in.
-
-The pressure for efficient communication influences language at all timescales, from online speech production to evolution.
-@mahowald2013 found that speakers tend to use shorter, more efficient word forms in predictive contexts, minimizing communicative cost without sacrificing intelligibility.
-Predictability and ambiguity shape the distribution of color terms in languages [@gibson2017] as well as word lengths [@piantadosi2011] across languages, and speakers are able to deterministically decode words despite ambiguity in word forms [@piantadosi2012].
-
-How do frequency and predictive context determine how we distribution information at the level of utterances?
-We do not speak only in phrases or paragraphs, but conversationally we speak in utterances and in writing we produce sentences as a coherent unit of a complete thought.
-Languages possess very different grammatical and morphological structures--the same thought can be encoded in at least as many ways as there are different languages.
-The average rate of information speakers transfer in syllables differs from language to language [@aylett2004].
-How do the different structures seen in different languages, with different rates of information encoding, affect how speakers structure information in production?
-
-@yu2016 proposed a novel method for quantifying the amount of information typically found at each part of English utterances.
-Treating each word position by itself, @yu2016 calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty [@shannon1948].
-They found a distinctive three-step shape to the distribution of information in English written sentences.
-
-In this paper, we expand on this body of prior work in a number of novel ways.
-We replicate the results from @yu2016 with a different metric, one intimately tied to incremental word processing.
-We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech.
-We also apply a contextual version of the same metric, and find that in communication, listeners are able to decode information from the utterances they hear at a constant rate past the initial words.
-We expand the study of information density to the largest set of languages yet considered, and incorporate contextual and typological information.
-We find that each language has a characteristic, robust frequency-based information distribution tied to the language's morphology, phonology and syntax.
-Speakers of a language will tend to distribute information according to that language’s distribution of information in their utterances.
-As soon as a child can form multi-word utterances in a given language, they will produce utterances obeying the characteristic distribution of information in a given language.
 
@@
 Similarly, we expect that there will be a small smoothing effect for our contextual surprisal metric.
+
+In our contextual models, we face the problem of overfitting: we only train our model on those utterances which occur in the corpus and test our model on the same utterances.
+This ignores possible other utterances which the speakers could have produced, e.g.
+the words "I", "saw" and "bears" are in the corpus vocabulary, which the speaker may not have produced in the corpus but could have produced.
+To combat this issue, we use modified Kneser-Ney smoothing [@chen1999empirical].
+
+In this smoothing technique, we filter the ngram counts from the corpus to only those above a threshold, ignoring rare ngrams.
+We then apply *absolute discounting* to all of the ngram, which reduces the counts of all observed ngrams in the corpus by a fixed number, reducing the impact of uncommon ngrams on our surprisal calculations.
+We *interpolate* lower order ngram counts when computing bigrams and trigrams: this means computing a weighted average of the ngram counts along with lower order ngram counts, in case the higher order counts for a particular ngram are rare.
+The main insight of Kneser-Ney smoothing is to weight ngram counts by how often the final word occurs as a continuation of any previous words in the corpus, downweighting common unigrams which occur only in a few contexts.
+This version of modified Kneser-Ney smoothing is in the Kenlm language modeling toolkit [@heafield2013scalable].
 

### 2019-10-02 Paper/info-curves.tex
introduction edits and smoothing in methods added to paper

--- 2019-09-17 Paper/paper.tex
@@
-\documentclass[11pt,]{article}
+\documentclass[man,floatsintext]{apa6}
 \usepackage{lmodern}
@@
 }{}
-\usepackage[margin=1in]{geometry}
 \usepackage{hyperref}
@@
             pdftitle={Speakers of diverse languages structure their utterances for efficient communication},
-            pdfauthor={Josef Klafka and Daniel Yurovsky},
+            pdfauthor={Josef Klafka~\& Daniel Yurovksy},
+            pdfkeywords={keywords},
             pdfborder={0 0 0},
@@
 \urlstyle{same}  % don't use monospace font for urls
-\usepackage{longtable,booktabs}
 \usepackage{graphicx,grffile}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
 \setlength{\emergencystretch}{3em}  % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
+\setcounter{secnumdepth}{0}
 % Redefines (sub)paragraphs to behave more like sections
@@
 
-%%% Change title format to be more compact
-\usepackage{titling}
+  \title{Speakers of diverse languages structure their utterances for efficient communication}
+    \author{Josef Klafka\textsuperscript{1}~\& Daniel Yurovksy\textsuperscript{1,2}}
+    \date{}
 
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
+\shorttitle{Cross-linguistic information structure}
+\affiliation{
+\vspace{0.5cm}
+\textsuperscript{1} Carnegie Mellon University\\\textsuperscript{2} University of Chicago}
+\keywords{keywords\newline\indent Word count: X}
+\usepackage{csquotes}
+\usepackage{upgreek}
+\captionsetup{font=singlespacing,justification=justified}
 
-\setlength{\droptitle}{-2em}
+\usepackage{longtable}
+\usepackage{lscape}
+\usepackage{multirow}
+\usepackage{tabularx}
+\usepackage[flushleft]{threeparttable}
+\usepackage{threeparttablex}
 
-  \title{Speakers of diverse languages structure their utterances for efficient communication}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-    \author{Josef Klafka and Daniel Yurovsky}
-    \preauthor{\centering\large\emph}
-  \postauthor{\par}
-      \predate{\centering\large\emph}
-  \postdate{\par}
-    \date{2019-09-17}
+\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
 
-\usepackage{setspace}
+\makeatletter
+\newcommand\LastLTentrywidth{1em}
+\newlength\longtablewidth
+\setlength{\longtablewidth}{1in}
+\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
 
-\doublespacing
+\usepackage{lineno}
 
-\begin{document}
-\maketitle
+\linenumbers
 
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
+\authornote{Thanks to folks. JSMF.
 
+Correspondence concerning this article should be addressed to Josef Klafka, 5000 Forbes Ave. E-mail: \href{mailto:klafka@cmu.edu}{\nolinkurl{klafka@cmu.edu}}}
+
+\abstract{
 What role does communicative efficiency play in how we organize our utterances?
@@
 
+}
+
+\begin{document}
+\maketitle
+
 \hypertarget{introduction}{%
@@
 
-We use language to communicate information.
-Although we often need to say things we have never said before, we are easily and quickly able to produce novel utterances with next to no prior planning.
-Our listeners incrementally process each word we say in utterances they may have never heard before, and predict which words will come next within fractions of a second of hearing the beginning of the word (for evidence see e.g.
-Kutas \& Federmeier (2011)).
-Encoding new and unpredictable information into our utterances surprises our listeners, and presents problems for communication.
-Listeners may struggle to decode unpredictable words, which can cause difficulty decoding the following words as well.
-Efficient and effective communication relies on interlocutors being able to process novel information without being surprised by new information in unpredictable words.
-On the other hand, to make best use of the capacity of the communication channel, speakers should encode as much information as possible into each word they say.
+One of the defining features of human language is its power to transmit information.
+We use language for a variety of different proximal tasks tasks: greeting friends, taking notes, signalling group identities.
+But all share a common unifying purpose: changing the mental state of the listener or reader of our communications (Austin, 1975).
+Language can thus naturally thought of as a code that allows speakers to turn their intended meanings into a form that can be transmitted through the air (or paper, or electrons) and subsequently converted by listeners back into an approximation of the intended meaning (Shannon, 1948).
+
+Beyond its utility as a metaphor, this coding perspective on language is powerful because it affords a framework for rational analysis: If language has evolved to be an optimal code for information transmission, how should this code be structured (Anderson \& Milson, 1989)?
+The optimal code would have to optimize two competing pressures: (1) a pressure for listeners to easily and successfully decode messages sent by the speaker, and (2) a pressure for speakers to easily code their messages and transmit them with minimal effort and error.
+A fundamental constriaint on both of these processes is the serial nature of human language: sounds are produced one at a time and each is unavailable perceptually once it is no longer being produced.
+
+The solution that languages have arrived on is incremental processing: Listeners process speech continuously as it arrives, predicting upcoming words and building expectations about the likely meaning of utterances in real-time rather than at their conclusion (Kutas \& Federmeier, 2011; Pickering \& Garrod, 2013; Tanenhaus, Spivey-Knowlton, Eberhard, \& Sedivy, 1995).
+Because prediction errors can lead to severe processing costs and difficulty integrating new information, speakers should seek to minimize prediction errors.
+However, the cost of producing more predictable utterances is using more words.
+Thus, an optimal coding strategy is for speakers seeking to minimize their production costs is to produce utterances that are are just at the prediction capacity of listeners without exceeding this capacity (Aylett \& Turk, 2004; Genzel \& Charniak, 2002).
+
+Using information theory--a mathematical framework for formalizing predictability--researchers have tested and confirmed this general prediction of optimal coding across levels and contexts of language production.
+For instance, Genzel and Charniak (2002) provided a clever indirect test of this hypothesis at the level of sentences.
+They showed that the predictability of successive sentences in a discourse--analyzed in isolation--decreases, as would be expected if readers use prior sentences to predict the content of future sentences and total predictability remains constant.
+
+the predicted if the amount of information is constant readers use prior sentences in a discourse to predict the content of future setences as they read.
+
+This prediction of optimal coding--that speakers should maintain a constant rate of predictability has been tested and confirmed in a variety of ways across levels of language.
+
+\begin{itemize}
+\tightlist
+\item
+  Coding/decoding and information transfer
+\item
+  language as prediction
+\item
+  how should the code be structured?
+optimization questions like this have a had * a lot power across different parts of language: zipf, kemp \& regier, etc etc.
+noisy channel stuff.
+(or does this go in discussion)
+\item
+  don't under or overutilize channel, stay near limit -- constant entropy rate
+\item
+  evidence for this at multiple levels from individiual lexical choice *(info/information theory paper), inclusion of non-obligatory aspects of language (filled pauses, that), across multiple utterances in a discourse
+\item
+  what about over the course of an individual utterance--perhaps the strongest constraints from language itself?
+\item
+  yu et al.~found this puzzling effect.
+\item
+  we replicate that in a different kind of analysis and show that it extends to trigrams etc???
+\item
+  BUT variation cross-linguistically. why? constraints from language.
+\end{itemize}
 
@@
 The pressure for efficient communication influences language at all timescales, from online speech production to evolution.
-Mahowald, Fedorenko, Piantadosi, \& Gibson (2013) found that speakers tend to use shorter, more efficient word forms in predictive contexts, minimizing communicative cost without sacrificing intelligibility.
+Mahowald, Fedorenko, Piantadosi, and Gibson (2013) found that speakers tend to use shorter, more efficient word forms in predictive contexts, minimizing communicative cost without sacrificing intelligibility.
 Predictability and ambiguity shape the distribution of color terms in languages (Gibson et al., 2017) as well as word lengths (Piantadosi et al., 2011) across languages, and speakers are able to deterministically decode words despite ambiguity in word forms (Piantadosi, Tily, \& Gibson, 2012).
@@
 
-Yu, Cong, Liang, \& Liu (2016) proposed a novel method for quantifying the amount of information typically found at each part of English utterances.
+Yu, Cong, Liang, and Liu (2016) proposed a novel method for quantifying the amount of information typically found at each part of English utterances.
 Treating each word position by itself, Yu et al.
@@
 
-Shannon (1948) defined information as ``the reduction in uncertainty about one variable given that the value of another variable is known''.
+Shannon (1948) defined information as \enquote{the reduction in uncertainty about one variable given that the value of another variable is known}.
 We use a metric proposed by Shannon and applied by Levy (2008): lexical surprisal, which defines the information in word based on the ratio of possible continuations of the sentence after to before the word is seen.
@@
 Considered without context, the surprisal of an individual word is inversely proportional to the frequency of that word, so that the less often a person has seen a word, the more information that word holds.
-For example, ``flower'' has less information than ``azalea'' because ``flower'' is much more common than ``azalea''.
-Though the two words have the same length in number of letters, it's more difficult to process ``azalea'' when reading it here than when reading ``flower''.
+For example, \enquote{flower} has less information than \enquote{azalea} because \enquote{flower} is much more common than \enquote{azalea}.
+Though the two words have the same length in number of letters, it's more difficult to process \enquote{azalea} when reading it here than when reading \enquote{flower}.
 Frequency is intimately tied information content in words, with much of the differences between words frequencies being explained by information content cross-linguistically (Piantadosi et al., 2011).
@@
 When we use a word or two of context in our surprisal calculations, then the set of reasonable final items in our ngrams is greatly restricted.
-``Flower'' may contain less information than ``azalea'' when we consider the words independently of their context, but with context this can be reversed.
-Flower appears in a variety of contexts, and so the information content of a word like ``flower'' in a particular context may be higher than ``azalea''.
-If you only have azaleas in your garden, then hearing someone say ``in that garden, look at the flowers'' may be higher surprisal for you: you expect them to say ``azalea''.
-This prediction does not need many words before to work out, for example in ``I take my coffee with cream and sugar''.
-When you hear ``cream and'', you automatically predict ``sugar'' or maybe ``honey'', but there are few possible continuations with even those two words.
-Hearing ``I'' restricts the next word to a verb, or possibly an adverb, and since you have just heard the speaker refer to themselves in the first person singular, your set of possible completions is significantly restricted.
+\enquote{Flower} may contain less information than \enquote{azalea} when we consider the words independently of their context, but with context this can be reversed.
+Flower appears in a variety of contexts, and so the information content of a word like \enquote{flower} in a particular context may be higher than \enquote{azalea}.
+If you only have azaleas in your garden, then hearing someone say \enquote{in that garden, look at the flowers} may be higher surprisal for you: you expect them to say \enquote{azalea}.
+This prediction does not need many words before to work out, for example in \enquote{I take my coffee with cream and sugar}.
+When you hear \enquote{cream and}, you automatically predict \enquote{sugar} or maybe \enquote{honey}, but there are few possible continuations with even those two words.
+Hearing \enquote{I} restricts the next word to a verb, or possibly an adverb, and since you have just heard the speaker refer to themselves in the first person singular, your set of possible completions is significantly restricted.
 
@@
 (2016) result using the surprisal metric in place of the entropy metric.
-We use the frequency-based or ``contextless'' surprisal metric, which determines the average distribution of information based on word frequencies in a corpus.
+We use the frequency-based or \enquote{contextless} surprisal metric, which determines the average distribution of information based on word frequencies in a corpus.
 A priori we expect that the frequency-based metric will produce a flat distribution of information across word positions in the BNC.
@@
 \centering
-\includegraphics{paper_files/figure-latex/bnc_unigrams-1.pdf}
+\includegraphics{figs/bnc_unigrams-1.pdf}
 \caption{(\#fig:bnc\_unigrams)BNC frequency-based information curves}
@@
 \centering
-\includegraphics{paper_files/figure-latex/switchboard_unigrams-1.pdf}
+\includegraphics{figs/switchboard_unigrams-1.pdf}
 \caption{(\#fig:switchboard\_unigrams)Switchboard frequency-based information curves}
@@
 \centering
-\includegraphics{paper_files/figure-latex/switchboard_bigrams-1.pdf}
+\includegraphics{figs/switchboard_bigrams-1.pdf}
 \caption{(\#fig:switchboard\_bigrams)Switchboard bigram context-based information curves}
@@
 \centering
-\includegraphics{paper_files/figure-latex/switchboard_trigrams-1.pdf}
+\includegraphics{figs/switchboard_trigrams-1.pdf}
 \caption{(\#fig:switchboard\_trigrams)Switchboard trigram context-based information curves}
@@
 \centering
-\includegraphics{paper_files/figure-latex/providence_unigrams-1.pdf}
+\includegraphics{figs/providence_unigrams-1.pdf}
 \caption{(\#fig:providence\_unigrams)Providence frequency-based information curves. Lines around each point indicate 95\% confidence intervals computed with non-parametric bootstrap}
@@
 \centering
-\includegraphics{paper_files/figure-latex/providence_bigrams-1.pdf}
+\includegraphics{figs/providence_bigrams-1.pdf}
 \caption{(\#fig:providence\_bigrams)Providence bigram context-based information curves}
@@
 \centering
-\includegraphics{paper_files/figure-latex/providence_trigrams-1.pdf}
+\includegraphics{figs/providence_trigrams-1.pdf}
 \caption{(\#fig:providence\_trigrams)Providence trigram context-based information curves}
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
+\includegraphics{figs/unnamed-chunk-1-1.pdf}
 \caption{\label{fig:unnamed-chunk-1}CHILDES frequency-based information curves}
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
+\includegraphics{figs/unnamed-chunk-2-1.pdf}
 \caption{\label{fig:unnamed-chunk-2}CHILDES bigram context-based information curves}
@@
 \centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
+\includegraphics{figs/unnamed-chunk-3-1.pdf}
 \caption{\label{fig:unnamed-chunk-3}CHILDES trigram context-based information curves}
@@
 We used two different strategies, which yielded identical results upon analysis.
-Each strategy gave us a five-dimensional vector space embedding for each language in a Wikipedia ``slope space''.
+Each strategy gave us a five-dimensional vector space embedding for each language in a Wikipedia \enquote{slope space}.
 For the first strategy, we split each sentence length by number of words into fifths, and computed surprisal values for the closest word position to each quintile.
@@
 \centering
-\includegraphics{paper_files/figure-latex/treatments-1.pdf}
+\includegraphics{figs/treatments-1.pdf}
 \caption{\label{fig:treatments}Illustration of slope treatments for Wikipedia information curves: relative on top and absolute on bottom}
@@
 
-\includegraphics{paper_files/figure-latex/dendro-1.pdf}
+\includegraphics{figs/dendro-1.pdf}
 
@@
 \centering
-\includegraphics{paper_files/figure-latex/wiki_trigrams-1.pdf}
+\includegraphics{figs/wiki_trigrams-1.pdf}
 \caption{(\#fig:wiki\_trigrams)Some trigram information curves from the Wikipedia data}
@@
 Individual WALS features do not necessarily have ordinal values.
-Some, such as the ``Number of Cases'' feature, are easy to quantify and order.
+Some, such as the \enquote{Number of Cases} feature, are easy to quantify and order.
 Others are more difficult.
-For example, how does one order ``relative clauses appear after the nouns they modify'', ``relative clauses appear before the nouns they modify'' and ``free order of relative clauses and nouns''?
+For example, how does one order \enquote{relative clauses appear after the nouns they modify}, \enquote{relative clauses appear before the nouns they modify} and \enquote{free order of relative clauses and nouns}?
 We chose the identify relation to avoid deciding on the basis of individual features.
@@
 \centering
-\includegraphics{paper_files/figure-latex/linear_models-1.pdf}
+\includegraphics{figs/linear_models-1.pdf}
 \caption{(\#fig:linear\_models)Some linear model results from Wikipedia and WALS features}
@@
 \centering
-\includegraphics{paper_files/figure-latex/cosine_wals-1.pdf}
+\includegraphics{figs/cosine_wals-1.pdf}
 \caption{(\#fig:cosine\_wals)wals features vs cosine similarity}
@@
 \centering
-\includegraphics{paper_files/figure-latex/cosine_ldn-1.pdf}
+\includegraphics{figs/cosine_ldn-1.pdf}
 \caption{(\#fig:cosine\_ldn)ldn features vs cosine similarity}
@@
 
+\newpage
+
 \hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
+\section{References}\label{references}}
+
+\begingroup
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
 
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-audigier2017mimca}{}%
+\leavevmode\hypertarget{ref-anderson1989}{}%
+Anderson, J. R., \& Milson, R. (1989). Human memory: An adaptive perspective. \emph{Psychological Review}, \emph{96}(4), 703.
+
+\leavevmode\hypertarget{ref-audigier2017}{}%
 Audigier, V., Husson, F., \& Josse, J. (2017). MIMCA: Multiple imputation for categorical variables with multiple correspondence analysis. \emph{Statistics and Computing}, \emph{27}(2), 501--518.
 
-\leavevmode\hypertarget{ref-aylett2004smooth}{}%
+\leavevmode\hypertarget{ref-austin1975}{}%
+Austin, J. L. (1975). \emph{How to do things with words}. Oxford university press.
+
+\leavevmode\hypertarget{ref-aylett2004}{}%
 Aylett, M., \& Turk, A.
@@
 
-\leavevmode\hypertarget{ref-bates2014fitting}{}%
+\leavevmode\hypertarget{ref-bates2014}{}%
 Bates, D., Mächler, M., Bolker, B., \& Walker, S. (2014). Fitting linear mixed-effects models using lme4. \emph{arXiv Preprint arXiv:1406.5823}.
@@
 \leavevmode\hypertarget{ref-wals}{}%
-Dryer, M. S., \& Haspelmath, M. (Eds.). (2013). \emph{WALS online}. Retrieved from \url{https://wals.info/}
+Dryer, M. S., \& Haspelmath, M. (Eds.). (2013). \emph{WALS online}. Leipzig: Max Planck Institute for Evolutionary Anthropology. Retrieved from \url{https://wals.info/}
 
-\leavevmode\hypertarget{ref-evans2012individual}{}%
+\leavevmode\hypertarget{ref-evans2012}{}%
 Evans, K. E., \& Demuth, K. (2012). Individual differences in pronoun reversal: Evidence from two longitudinal case studies. \emph{Journal of Child Language}, \emph{39}(1), 162--191.
 
-\leavevmode\hypertarget{ref-frank2015erp}{}%
+\leavevmode\hypertarget{ref-frank2015}{}%
 Frank, S. L., Otten, L. J., Galli, G., \& Vigliocco, G. (2015). The erp response to the amount of information conveyed by words in sentences. \emph{Brain and Language}, \emph{140}, 1--11.
 
-\leavevmode\hypertarget{ref-genzel2002entropy}{}%
-Genzel, D., \& Charniak, E. (2002). Entropy rate constancy in text. \emph{Proceedings of the 40th annual meeting of the association for computational linguistics}, 199--206.
+\leavevmode\hypertarget{ref-genzel2002}{}%
+Genzel, D., \& Charniak, E. (2002). Entropy rate constancy in text. In \emph{Proceedings of the 40th annual meeting of the association for computational linguistics} (pp. 199--206).
 
-\leavevmode\hypertarget{ref-gibson2017color}{}%
+\leavevmode\hypertarget{ref-gibson2017}{}%
 Gibson, E., Futrell, R., Jara-Ettinger, J., Mahowald, K., Bergen, L., Ratnasingam, S., \ldots{} Conway, B.
@@
 
-\leavevmode\hypertarget{ref-holman2008advances}{}%
+\leavevmode\hypertarget{ref-holman2008}{}%
 Holman, E.
@@
 
-\leavevmode\hypertarget{ref-jaeger2007speakers}{}%
-Jaeger, T. F., \& Levy, R. P. (2007). Speakers optimize information density through syntactic reduction. \emph{Advances in neural information processing systems}, 849--856.
+\leavevmode\hypertarget{ref-jaeger2007}{}%
+Jaeger, T. F., \& Levy, R. P. (2007). Speakers optimize information density through syntactic reduction. In \emph{Advances in neural information processing systems} (pp. 849--856).
 
-\leavevmode\hypertarget{ref-kutas2011thirty}{}%
+\leavevmode\hypertarget{ref-kutas2011}{}%
 Kutas, M., \& Federmeier, K.
@@
 
-\leavevmode\hypertarget{ref-leech1992100}{}%
-Leech, G. N. (1992). \emph{100 million words of english: The british national corpus (bnc)}.
+\leavevmode\hypertarget{ref-leech1992}{}%
+Leech, G. N. (1992). 100 million words of english: The british national corpus (bnc).
 
-\leavevmode\hypertarget{ref-levy2008expectation}{}%
+\leavevmode\hypertarget{ref-levy2008}{}%
 Levy, R. (2008). Expectation-based syntactic comprehension. \emph{Cognition}, \emph{106}(3), 1126--1177.
 
-\leavevmode\hypertarget{ref-li2015dinner}{}%
+\leavevmode\hypertarget{ref-li2015}{}%
 Li, H., \& Zhou, J. (2015). \emph{Study on dinner table talk of preschool children family in shanghai} (Master's thesis). East China Normal University, Shanghai, China.
 
-\leavevmode\hypertarget{ref-macwhinney2000childes}{}%
+\leavevmode\hypertarget{ref-macwhinney2000}{}%
 MacWhinney, B. (2000). \emph{The childes project: The database} (Vol. 2). Psychology Press.
 
-\leavevmode\hypertarget{ref-mahowald2013info}{}%
+\leavevmode\hypertarget{ref-mahowald2013}{}%
 Mahowald, K., Fedorenko, E., Piantadosi, S. T., \& Gibson, E. (2013). Info/information theory: Speakers choose shorter words in predictive contexts. \emph{Cognition}, \emph{126}(2), 313--318.
 
-\leavevmode\hypertarget{ref-palasis2009syntaxe}{}%
+\leavevmode\hypertarget{ref-palasis2009}{}%
 Palasis, K. (2009). \emph{Syntaxe générative et acquisition: Le sujet dans le développement du système linguistique du jeune enfant} (PhD thesis). Nice.
 
-\leavevmode\hypertarget{ref-piantadosi2011word}{}%
+\leavevmode\hypertarget{ref-piantadosi2011}{}%
 Piantadosi, S. T., Tily, H., \& Gibson, E. (2011). Word lengths are optimized for efficient communication. \emph{Proceedings of the National Academy of Sciences}, \emph{108}(9), 3526--3529.
 
-\leavevmode\hypertarget{ref-piantadosi2012communicative}{}%
+\leavevmode\hypertarget{ref-piantadosi2012}{}%
 Piantadosi, S. T., Tily, H., \& Gibson, E. (2012). The communicative function of ambiguity in language. \emph{Cognition}, \emph{122}(3), 280--291.
 
-\leavevmode\hypertarget{ref-sanchez2019childes}{}%
+\leavevmode\hypertarget{ref-pickering2013}{}%
+Pickering, M. J., \& Garrod, S. (2013). An integrated theory of language production and comprehension. \emph{Behavioral and Brain Sciences}, \emph{36}(4), 329--347.
+
+\leavevmode\hypertarget{ref-sanchez2019}{}%
 Sanchez, A., Meylan, S.
@@
 
-\leavevmode\hypertarget{ref-shannon1948mathematical}{}%
+\leavevmode\hypertarget{ref-shannon1948}{}%
 Shannon, C. E. (1948). A mathematical theory of communication. \emph{Bell System Technical Journal}, \emph{27}(3), 379--423.
 
-\leavevmode\hypertarget{ref-shirai2001acquisition}{}%
+\leavevmode\hypertarget{ref-shirai2001}{}%
 Shirai, Y. (2001). The acquisition of causative morphology in japanese: A prototype account. \emph{East Asian Language Acquisition}.
 
-\leavevmode\hypertarget{ref-de1996estudio}{}%
+\leavevmode\hypertarget{ref-shiro1996}{}%
 Shiro, M. L. de. (1996). Un estudio de las expresiones de modalidad en hablantes de dos culturas. \emph{Boletı'n de Lingüı'stica}, (10), 43--60.
 
-\leavevmode\hypertarget{ref-smith2013effect}{}%
+\leavevmode\hypertarget{ref-smith2013}{}%
 Smith, N. J., \& Levy, R. (2013). The effect of word predictability on reading time is logarithmic. \emph{Cognition}, \emph{128}(3), 302--319.
 
-\leavevmode\hypertarget{ref-swadesh1955towards}{}%
+\leavevmode\hypertarget{ref-swadesh1955}{}%
 Swadesh, M. (1955). Towards greater accuracy in lexicostatistic dating. \emph{International Journal of American Linguistics}, \emph{21}(2), 121--137.
 
-\leavevmode\hypertarget{ref-team2013r}{}%
-Team, R. C., \& others. (2013). \emph{R: A language and environment for statistical computing}.
+\leavevmode\hypertarget{ref-tanenhaus1995}{}%
+Tanenhaus, M.
+K., Spivey-Knowlton, M.
+J., Eberhard, K.
+M., \& Sedivy, J.
+C.
+(1995).
+Integration of visual and linguistic information in spoken language comprehension.
+\emph{Science}, \emph{268}(5217), 1632--1634.
 
-\leavevmode\hypertarget{ref-wagner1985much}{}%
+\leavevmode\hypertarget{ref-team2013}{}%
+Team, R. C., \& others. (2013). R: A language and environment for statistical computing.
+
+\leavevmode\hypertarget{ref-wagner1985}{}%
 Wagner, K. R. (1985). How much do children say in a day? \emph{Journal of Child Language}, \emph{12}(2), 475--487.
 
-\leavevmode\hypertarget{ref-wichmann2016asjp}{}%
+\leavevmode\hypertarget{ref-wichmann2016}{}%
 Wichmann, S., Müller, A., Wett, A., Velupillai, V., Bischoffberger, J., Brown, C. H., \ldots{} others. (2016). The asjp database. \emph{Max Planck Institute for the Science of Human History, Jena}.
 
-\leavevmode\hypertarget{ref-yu2016distribution}{}%
+\leavevmode\hypertarget{ref-yu2016}{}%
 Yu, S., Cong, J., Liang, J., \& Liu, H. (2016). The distribution of information content in english sentences. \emph{arXiv Preprint arXiv:1609.07681}.
 
-\leavevmode\hypertarget{ref-zipf1949human}{}%
-Zipf, G. K. (1949). \emph{Human behavior and the principle of least effort.}
+\leavevmode\hypertarget{ref-zipf1949}{}%
+Zipf, G. K. (1949). Human behavior and the principle of least effort.
+
+\endgroup
 

### 2019-10-04 Paper/info-curves.Rmd
updated intro and smoothing in methods for paper

@@
 
-How these pressures shape language evolution??
+Larger parts of language, as well as languages themselves, evolve according to efficient encoding.
+Semantic categories of words across languages can evolve to be structured efficiently, maintaining a trade-off between informativeness and complexity in the semantic category, such as kinship terms [@kemp2012kinship].
+Languages more generally evolve according to principles of efficient communication: features of the world that are relevant to speakers become part of a language, while irrelevant features are disregarded [@perfors2014language] and structure in langauge evolves from a trade-off between efficient and learnable encoding on the one hand and an expressive and descriptive lexicon on the other [@kirby2015compression].
 
@@
 We replicate the results from @yu2016 with a metric tied to incremental word processing.
-We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech.
-We also apply a contextual version of the same metric, and find that in communication, listeners are able to decode information from the utterances they hear at a constant rate past the initial words.
-We expand the study of information density to the largest set of languages yet considered, and incorporate contextual and typological information.
-We find that each language has a characteristic, robust frequency-based information distribution tied to the language's morphology, phonology and syntax.
-Speakers of a language will tend to distribute information according to that language’s distribution of information in their utterances.
-As soon as a child can form multi-word utterances in a given language, they will produce utterances obeying the characteristic distribution of information in a given language.
+We find their same distribution of information based on word frequency in English speech, as well as in English parent-child conversations.
+We extend our metric to include context, and show that the addition of context for each word smoothes out language-specific distributions.
+We expand the study of information density to the largest set of languages considered so far, and incorporate contextual and typological information into our analysis.
+Speakers will tend to distribute information in a language constrained but not determined by the morphology, syntax and phonology of that language.
+As soon as a child starts speaking, they tend to distribute information in their utterances according to the characteristic distribution in their language.
 
@@
 
-To find the distribution of information within a corpus, we calculate the surprisal or contextual surprisal for each unique word or sequence of words within the corpus.
-We take all utterances of a given length, and for each word position in utterances of that length, we compute the average of the surprisals for all of the non-unique words that occur in that position, conditioned or not conditioned on context.
+We train an ngram language model on a given corpusm with order set to $1$ for the frequency-model, $2$ for a bigram model and $3$ for a trigram model.
+An ngram language model produces a probability for each continuation of each $n-1$ words it receives as input.
+Models with order greater than $3$ have issues with overfitting to the corpus.
+In our contextual models, we face another issue of overfitting: we only train our model on those utterances which occur in the corpus and test our model on the same utterances.
+This ignores possible other utterances which the speakers could have produced, e.g.
+the words "I", "saw" and "bears" are in the corpus vocabulary, which the speaker may not have produced in the corpus but could have produced.
+To combat this issue, we use modified Kneser-Ney smoothing as implemented in the KenLM toolkit [@heafield2013scalable].
+Briefly, this smoothing technique discounts all ngram counts, which reduces the impact of rare ngrams on probability calculations, and interpolates lower-order ngrams into the calcuations.
+These lower-order ngrams are weighted according to the number of distinct contexts they occur as a continuation (e.g.
+"Francisco" may be a common word in a corpus, but likely only occurs after "San" as in "San Francisco", so it receives a lower weighting).
+For a full explanation of modified Kneser-Ney smoothing, see @chen1999empirical.
+
+Once we have fitted our language model, we can compute the surprisal of a continuation by simply taking the negative log-probability of that word's ngram probability.
+To find the average information for a given position in a corpus, we take all utterances of a given length, and for each word position in utterances of that length, we compute the average of the surprisals for all of the non-unique words that occur in that position, conditioned or not conditioned on context.
 By computing these averages for each word position in an utterance, we compute a low-dimensional approximation to the average distribution of information in the corpus.
@@
 Similarly, we expect that there will be a small smoothing effect for our contextual surprisal metric.
-
-In our contextual models, we face the problem of overfitting: we only train our model on those utterances which occur in the corpus and test our model on the same utterances.
-This ignores possible other utterances which the speakers could have produced, e.g.
-the words "I", "saw" and "bears" are in the corpus vocabulary, which the speaker may not have produced in the corpus but could have produced.
-To combat this issue, we use modified Kneser-Ney smoothing [@chen1999empirical].
-
-In this smoothing technique, we filter the ngram counts from the corpus to only those above a threshold, ignoring rare ngrams.
-We then apply *absolute discounting* to all of the ngram, which reduces the counts of all observed ngrams in the corpus by a fixed number, reducing the impact of uncommon ngrams on our surprisal calculations.
-We *interpolate* lower order ngram counts when computing bigrams and trigrams: this means computing a weighted average of the ngram counts along with lower order ngram counts, in case the higher order counts for a particular ngram are rare.
-The main insight of Kneser-Ney smoothing is to weight ngram counts by how often the final word occurs as a continuation of any previous words in the corpus, downweighting common unigrams which occur only in a few contexts.
-This version of modified Kneser-Ney smoothing is in the Kenlm language modeling toolkit [@heafield2013scalable].
 

### 2019-10-04 Paper/info-curves.tex
updated intro and smoothing in methods for paper

@@
             pdftitle={Speakers of diverse languages structure their utterances for efficient communication},
-            pdfauthor={Josef Klafka~\& Daniel Yurovksy},
+            pdfauthor={Josef Klafka~\& Daniel Yurovsky},
             pdfkeywords={keywords},
@@
   \title{Speakers of diverse languages structure their utterances for efficient communication}
-    \author{Josef Klafka\textsuperscript{1}~\& Daniel Yurovksy\textsuperscript{1,2}}
+    \author{Josef Klafka\textsuperscript{1}~\& Daniel Yurovsky\textsuperscript{1,2}}
     \date{}
@@
 
-Correspondence concerning this article should be addressed to Josef Klafka, 5000 Forbes Ave. E-mail: \href{mailto:klafka@cmu.edu}{\nolinkurl{klafka@cmu.edu}}}
+Correspondence concerning this article should be addressed to Josef Klafka, 5000 Forbes Ave. E-mail: \href{mailto:jklafka@andrew.cmu.edu}{\nolinkurl{jklafka@andrew.cmu.edu}}}
 
@@
 One of the defining features of human language is its power to transmit information.
-We use language for a variety of different proximal tasks tasks: greeting friends, taking notes, signalling group identities.
+We use language for a variety of different tasks: greeting friends, taking notes, signaling group identities.
 But all share a common unifying purpose: changing the mental state of the listener or reader of our communications (Austin, 1975).
-Language can thus naturally thought of as a code that allows speakers to turn their intended meanings into a form that can be transmitted through the air (or paper, or electrons) and subsequently converted by listeners back into an approximation of the intended meaning (Shannon, 1948).
+Language can thus naturally thought of as a code that allows speakers to turn their intended meanings into a message that can be transmitted through the air (or paper, or electrons) and subsequently converted by listeners back into an approximation of the intended meaning (Shannon, 1948).
 
-Beyond its utility as a metaphor, this coding perspective on language is powerful because it affords a framework for rational analysis: If language has evolved to be an optimal code for information transmission, how should this code be structured (Anderson \& Milson, 1989)?
-The optimal code would have to optimize two competing pressures: (1) a pressure for listeners to easily and successfully decode messages sent by the speaker, and (2) a pressure for speakers to easily code their messages and transmit them with minimal effort and error.
-A fundamental constriaint on both of these processes is the serial nature of human language: sounds are produced one at a time and each is unavailable perceptually once it is no longer being produced.
+Beyond its utility as a metaphor, this coding perspective on language is powerful because it allows a framework for rational analysis: If language has evolved to be an optimal code for information transmission, how should this code be structured (Anderson \& Milson, 1989)?
+The optimal code would have to work with two competing pressures: (1) a pressure for listeners to easily and successfully decode messages sent by the speaker, and (2) a pressure for speakers to easily code their messages and transmit them with minimal effort and error.
+A fundamental constriaint on both of these processes is the linear order of spoken language: sounds are produced one at a time and each is unavailable perceptually once it is no longer being produced.
 
-The solution that languages have arrived on is incremental processing: Listeners process speech continuously as it arrives, predicting upcoming words and building expectations about the likely meaning of utterances in real-time rather than at their conclusion (Kutas \& Federmeier, 2011; Pickering \& Garrod, 2013; Tanenhaus, Spivey-Knowlton, Eberhard, \& Sedivy, 1995).
+The strategic solution people employ is \emph{incremental processing}.
+Listeners process speech continuously as it arrives, predicting upcoming words and building expectations about the likely meaning of utterances in real-time rather than at their conclusion (Kutas \& Federmeier, 2011; Pickering \& Garrod, 2013; Tanenhaus, Spivey-Knowlton, Eberhard, \& Sedivy, 1995).
 Because prediction errors can lead to severe processing costs and difficulty integrating new information, speakers should seek to minimize prediction errors.
@@
 
-Using information theory--a mathematical framework for formalizing predictability--researchers have tested and confirmed this general prediction of optimal coding across levels and contexts of language production.
-For instance, Genzel and Charniak (2002) provided a clever indirect test of this hypothesis at the level of sentences.
+Using information theory--a mathematical framework for formalizing predictability--researchers have tested and confirmed this general prediction of optimal coding across several levels and contexts of language production.
+For instance, Genzel and Charniak (2002) provided a clever indirect test of this hypothesis across sentences.
 They showed that the predictability of successive sentences in a discourse--analyzed in isolation--decreases, as would be expected if readers use prior sentences to predict the content of future sentences and total predictability remains constant.
 
-the predicted if the amount of information is constant readers use prior sentences in a discourse to predict the content of future setences as they read.
+This prediction of optimal coding--that speakers should maintain a constant rate of predictability has been tested and confirmed in a variety of ways across levels of language.
+At the level of individual words, Mahowald, Fedorenko, Piantadosi, and Gibson (2013) showed that speakers use shorter words in more predictive contexts and prefer the shorter versions of words in more predictive contexts, maximizing the amount of information in each word while minimizing the time spent on those words.
+Over time, Piantadosi, Tily, and Gibson (2011) showed that shorter words across languages tend to be more easily predictable in context and thus less informative, over and above frequency.
 
-This prediction of optimal coding--that speakers should maintain a constant rate of predictability has been tested and confirmed in a variety of ways across levels of language.
+Efficient encoding also impacts how speakers structure phrases.
+The inclusion of complementizers in relative clauses (Jaeger \& Levy, 2007) and the use of contractions (Frank \& Jaeger, 2008) are two situations in sentence formation in which speakers use structure to communicate efficiently, spreading more information across a larger amount of linguistic material.
+
+Larger parts of language, as well as languages themselves, evolve according to efficient encoding.
+Semantic categories of words across languages can evolve to be structured efficiently, maintaining a trade-off between informativeness and complexity in the semantic category, such as kinship terms (Kemp \& Regier, 2012).
+Languages more generally evolve according to principles of efficient communication: features of the world that are relevant to speakers become part of a language, while irrelevant features are disregarded (Perfors \& Navarro, 2014) and structure in langauge evolves from a trade-off between efficient and learnable encoding on the one hand and an expressive and descriptive lexicon on the other (Kirby, Tamariz, Cornish, \& Smith, 2015).
+
+However, despite this literature using the predictive coding model of language, one level has not yet been studied in depth: how speakers structure individual utterances.
+This level may show the strongest effects of variation between languages, as specific languages have properties that constrain how speakers may form utterances in those languages, such as canonical word order.
+These properties vary widely from language to language.
+
+Yu, Cong, Liang, and Liu (2016) studied the utterance level in written English sentences using a contextless entropy model based on word frequency.
+They found a distinctive three-step distribution regardless of sentence length, with little information in the first words of sentences and the most information in the final words.
+This was surprising, as the distribution they found was robustly different from the linearly increasing trend in sentences from Genzel and Charniak (2002), and also did not resemble the uniform distribution of information that one might expect from a communicative efficiency account, in which each word has approximately equal information close to the channel capacity.
+
+In this paper, we expand on this body of prior work in a number of novel ways.
+We replicate the results from Yu et al.
+(2016) with a metric tied to incremental word processing.
+We find their same distribution of information based on word frequency in English speech, as well as in English parent-child conversations.
+We extend our metric to include context, and show that the addition of context for each word smoothes out language-specific distributions.
+We expand the study of information density to the largest set of languages considered so far, and incorporate contextual and typological information into our analysis.
+Speakers will tend to distribute information in a language constrained but not determined by the morphology, syntax and phonology of that language.
+As soon as a child starts speaking, they tend to distribute information in their utterances according to the characteristic distribution in their language.
 
@@
 \end{itemize}
-
-Due to their differences in distributional frequency, words vary in how much information content they contain (Shannon, 1948).
-Frequency is a meaningful predictor of word length: the more common the word, the shorter the word tends to be (Zipf, 1949).
-Over and above frequency, contextual information content predicts word length, due to speakers minimizing effort in the amount of information they encode in speech (Piantadosi, Tily, \& Gibson, 2011).
-Speakers tend to smooth out their information to a more uniform rate within utterances that allows for easier decoding on the listener side (Jaeger \& Levy, 2007).
-Across utterances, context allows speakers to produce more informative and surprising sentences later in communication.
-The later a sentence occurs in a paragraph, the the informative it is and the more unusual and surprising words it contains (Genzel \& Charniak, 2002).
-From the level of individual words to sentences in paragraphs, information content shapes and is shaped by the context it appears in.
-
-The pressure for efficient communication influences language at all timescales, from online speech production to evolution.
-Mahowald, Fedorenko, Piantadosi, and Gibson (2013) found that speakers tend to use shorter, more efficient word forms in predictive contexts, minimizing communicative cost without sacrificing intelligibility.
-Predictability and ambiguity shape the distribution of color terms in languages (Gibson et al., 2017) as well as word lengths (Piantadosi et al., 2011) across languages, and speakers are able to deterministically decode words despite ambiguity in word forms (Piantadosi, Tily, \& Gibson, 2012).
-
-How do frequency and predictive context determine how we distribution information at the level of utterances?
-We do not speak only in phrases or paragraphs, but conversationally we speak in utterances and in writing we produce sentences as a coherent unit of a complete thought.
-Languages possess very different grammatical and morphological structures--the same thought can be encoded in at least as many ways as there are different languages.
-The average rate of information speakers transfer in syllables differs from language to language (Aylett \& Turk, 2004).
-How do the different structures seen in different languages, with different rates of information encoding, affect how speakers structure information in production?
-
-Yu, Cong, Liang, and Liu (2016) proposed a novel method for quantifying the amount of information typically found at each part of English utterances.
-Treating each word position by itself, Yu et al.
-(2016) calculated the Shannon entropy (a measure of uncertainty) over the words of each position, producing an estimate for the amount of uncertainty reduced once the listener hears the words used in a given position, related to the Shannon concept of information as a reduction in uncertainty (Shannon, 1948).
-They found a distinctive three-step shape to the distribution of information in English written sentences.
-
-In this paper, we expand on this body of prior work in a number of novel ways.
-We replicate the results from Yu et al.
-(2016) with a different metric, one intimately tied to incremental word processing.
-We find the same distribution of information based on word frequency in English adult speech as in English adult writing, as well as in English child speech.
-We also apply a contextual version of the same metric, and find that in communication, listeners are able to decode information from the utterances they hear at a constant rate past the initial words.
-We expand the study of information density to the largest set of languages yet considered, and incorporate contextual and typological information.
-We find that each language has a characteristic, robust frequency-based information distribution tied to the language's morphology, phonology and syntax.
-Speakers of a language will tend to distribute information according to that language's distribution of information in their utterances.
-As soon as a child can form multi-word utterances in a given language, they will produce utterances obeying the characteristic distribution of information in a given language.
 
@@
 
-To find the distribution of information within a corpus, we calculate the surprisal or contextual surprisal for each unique word or sequence of words within the corpus.
-We take all utterances of a given length, and for each word position in utterances of that length, we compute the average of the surprisals for all of the non-unique words that occur in that position, conditioned or not conditioned on context.
+We train an ngram language model on a given corpusm with order set to \(1\) for the frequency-model, \(2\) for a bigram model and \(3\) for a trigram model.
+An ngram language model produces a probability for each continuation of each \(n-1\) words it receives as input.
+Models with order greater than \(3\) have issues with overfitting to the corpus.
+In our contextual models, we face another issue of overfitting: we only train our model on those utterances which occur in the corpus and test our model on the same utterances.
+This ignores possible other utterances which the speakers could have produced, e.g.~the words \enquote{I}, \enquote{saw} and \enquote{bears} are in the corpus vocabulary, which the speaker may not have produced in the corpus but could have produced.
+To combat this issue, we use modified Kneser-Ney smoothing as implemented in the KenLM toolkit (Heafield, Pouzyrevsky, Clark, \& Koehn, 2013).
+Briefly, this smoothing technique discounts all ngram counts, which reduces the impact of rare ngrams on probability calculations, and interpolates lower-order ngrams into the calcuations.
+These lower-order ngrams are weighted according to the number of distinct contexts they occur as a continuation (e.g.
+\enquote{Francisco} may be a common word in a corpus, but likely only occurs after \enquote{San} as in \enquote{San Francisco}, so it receives a lower weighting).
+For a full explanation of modified Kneser-Ney smoothing, see Chen and Goodman (1999).
+
+Once we have fitted our language model, we can compute the surprisal of a continuation by simply taking the negative log-probability of that word's ngram probability.
+To find the average information for a given position in a corpus, we take all utterances of a given length, and for each word position in utterances of that length, we compute the average of the surprisals for all of the non-unique words that occur in that position, conditioned or not conditioned on context.
 By computing these averages for each word position in an utterance, we compute a low-dimensional approximation to the average distribution of information in the corpus.
@@
 
+\leavevmode\hypertarget{ref-chen1999empirical}{}%
+Chen, S. F., \& Goodman, J. (1999). An empirical study of smoothing techniques for language modeling. \emph{Computer Speech \& Language}, \emph{13}(4), 359--394.
+
 \leavevmode\hypertarget{ref-wals}{}%
@@
 Evans, K. E., \& Demuth, K. (2012). Individual differences in pronoun reversal: Evidence from two longitudinal case studies. \emph{Journal of Child Language}, \emph{39}(1), 162--191.
+
+\leavevmode\hypertarget{ref-frank2008}{}%
+Frank, A.
+F., \& Jaeger, T.
+F.
+(2008).
+Speaking rationally: Uniform information density as an optimal strategy for language production.
+In \emph{Proceedings of the annual meeting of the cognitive science society} (Vol.
+30).
 
@@
 
-\leavevmode\hypertarget{ref-gibson2017}{}%
-Gibson, E., Futrell, R., Jara-Ettinger, J., Mahowald, K., Bergen, L., Ratnasingam, S., \ldots{} Conway, B.
-R.
-(2017).
-Color naming across languages reflects color use.
-\emph{Proceedings of the National Academy of Sciences}, \emph{114}(40), 10785--10790.
+\leavevmode\hypertarget{ref-heafield2013scalable}{}%
+Heafield, K., Pouzyrevsky, I., Clark, J.
+H., \& Koehn, P.
+(2013).
+Scalable modified kneser-ney language model estimation.
+In \emph{Proceedings of the 51st annual meeting of the association for computational linguistics (volume 2: Short papers)} (pp.
+690--696).
 
@@
 Jaeger, T. F., \& Levy, R. P. (2007). Speakers optimize information density through syntactic reduction. In \emph{Advances in neural information processing systems} (pp. 849--856).
+
+\leavevmode\hypertarget{ref-kemp2012kinship}{}%
+Kemp, C., \& Regier, T. (2012). Kinship categories across languages reflect general communicative principles. \emph{Science}, \emph{336}(6084), 1049--1054.
+
+\leavevmode\hypertarget{ref-kirby2015compression}{}%
+Kirby, S., Tamariz, M., Cornish, H., \& Smith, K. (2015). Compression and communication in the cultural evolution of linguistic structure. \emph{Cognition}, \emph{141}, 87--102.
 
@@
 
+\leavevmode\hypertarget{ref-perfors2014language}{}%
+Perfors, A., \& Navarro, D. J. (2014). Language evolution can be shaped by the structure of the world. \emph{Cognitive Science}, \emph{38}(4), 775--793.
+
 \leavevmode\hypertarget{ref-piantadosi2011}{}%
 Piantadosi, S. T., Tily, H., \& Gibson, E. (2011). Word lengths are optimized for efficient communication. \emph{Proceedings of the National Academy of Sciences}, \emph{108}(9), 3526--3529.
-
-\leavevmode\hypertarget{ref-piantadosi2012}{}%
-Piantadosi, S. T., Tily, H., \& Gibson, E. (2012). The communicative function of ambiguity in language. \emph{Cognition}, \emph{122}(3), 280--291.
 
@@
 
-\leavevmode\hypertarget{ref-zipf1949}{}%
-Zipf, G. K. (1949). Human behavior and the principle of least effort.
-
 \endgroup

### 2019-10-28 Wikipedia/langvardata.Rmd
organized entropy branch

---
title: Wikipedia Text Mining and Entropy Analysis
author: Josef Klafka
date: `Sys.date()`
output: 
  html_document:
    toc: false
    number_sections: false
    theme: lumen
    toc_float: false
    code_folding: show 
---

```{r setup, include=FALSE}
# load packages
library(DBI)
library(knitr)
library(tidyverse)
library(directlabels)
library(dplyr)
library(lingtypology)
library(tidytext)

knitr::opts_chunk$set(echo = TRUE)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
               error = FALSE, cache = FALSE, tidy = FALSE)

theme_set(theme_classic(base_size = 16))
```

```{r typology}
data <- read.csv("langvardata.csv")
data[c("language", "perimeter_log")]
```

### 2019-10-28 entropy/Wikipedia/get_wiki_df.py

@@ -1,96 +1,0 @@
-import regex as re
-import numpy as np
-import pandas as pd
-import csv
-import sys
-import json
-import subprocess
-import urllib.request
-
-
-with open("language_dict.json", 'r') as f:
-	file = f.read()
-LANG_DICT = json.loads(file) #maps languages to their wikipedia url prefixes
-
-
-def extract_texts(upper_bound):
-	'''
-	Given an upper_bound on file numbers, extracts the raw text from the files
-	in the text/AA folder. That folder is given as output by the
-	cirrus_extract program.
-	'''
-	rs = ""
-	for i in range(upper_bound):
-		if i < 10:
-			number = "0" + str(i)
-		else:
-			number = str(i)
-		f = open("text/AA/wiki_" + number, 'r')
-		text = f.read()
-		rs += text
-	return rs
-
-
-def get_sen_dict(text):
-	'''
-	Given the output of extract_texts as a string, constructs a
-	dictionary mapping each sentence in the string to its length in words.
-	'''
-	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:–'&*#/—“»]|\d+|\(|\)|\[|\]", \
-		"", text)
-	text = re.sub('"', '', text)
-	#just the words from the articles
-	text = text.replace('\n', " ")
-	sens = re.split("[.!?]", text.lower())
-	sen_dict = {}
-	for sen in sens:
-		length = len(re.findall("\w+", sen))
-		if length > 0:
-			sen_dict[sen] = length
-	return sen_dict
-
-
-def get_sen_df(text):
-	'''
-	Given the output of extract_texts as a string, returns a Pandas dataframe
-	of the output of get_sen_dict.
-	'''
-	sd = get_sen_dict(text)
-	s = pd.Series(sd, name = "length")
-	del sd
-	s.index.name = "words"
-	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence
-					 	  #and the sentence length respectively
-	return df
-
-
-def main(lang_name):
-	'''
-	Given the English name of a language, constructs a dataframe with the
-	sentences of a random subset of articles in that language from Wikipedia,
-	and stores the dataframe in a csv in the working directory.
-	'''
-	lang_prefix = LANG_DICT[lang_name]
-	url = "https://dumps.wikimedia.org/other/cirrussearch/20190318/" \
-		+ lang_prefix + "wiki-20190318-cirrussearch-content.json.gz"
[…]

### 2019-10-28 entropy/Wikipedia/wiki_relative.R

@@ -1,98 +1,0 @@
-library(knitr)
-library(tidyverse)
-library(directlabels)
-library(dplyr)
-library(tidytext)
-library(entropy)
-library(tidyboot)
-library(tokenizers)
-library(googlesheets)
-
-### Begin Functions
-
-extract_wiki_tokens <- function(min_length, max_length, utterances) {
-
-  tokens <- utterances %>%
-    mutate(gloss = str_trim(gloss)) %>%
-    mutate(length = str_count(gloss, pattern = " +") + 1) %>%
-    filter(length >= min_length) %>%
-    filter(length <= max_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    group_by(length, word_order, word) %>%
-    summarise(n = n())
-}
-
-get_quantiles <- function(df, num_sections) {
-  quantile(df$word_order, probs = seq(0, 1, 1/num_sections)) %>%
-    round() %>%
-    as_data_frame() %>%
-    rename(word_order = value)
-}
-
-relative_slopes <- function(df, pos_list) {
-
-  get_slope <- function(x) {
-
-    selected_pos <- pos_list %>%
-      group_by(length) %>%
-      slice(x:(x+1)) %>%
-      group_by(length) %>%
-      summarise(min = min(word_order),
-                max = max(word_order)) %>%
-      split(.$length) %>%
-      map(~seq(.x$min, .x$max, 1) %>% as_data_frame) %>%
-      bind_rows(.id = "length") %>%
-      rename(word_order = value) %>%
-      mutate(length = as.numeric(length))
-
-    df %>%
-      inner_join(selected_pos, by = c("length", "word_order")) %>%
-      group_by(length, word_order) %>%
-      summarise(entropy = entropy(n, unit = "log2")) %>%
-      lm(entropy ~ length + I(length^2) + word_order, data  = .) %>%
-      tidy() %>%
-      filter(term == "word_order")
-  }
-
-  divides <- pos_list %>%
-    group_by(length) %>%
-    summarise(n = n()) %>%
-    summarise(n = mean(n)) %>%
-    pull()
-
[…]

### 2019-10-28 entropy/Wikipedia/wiki_script.sh

@@ -1,42 +1,0 @@
-#!/bin/bash
-
-declare -a arr=("Basque" "Dutch" "Alemannic" "Wolof" "Bosnian" "Hindi" "Danish"
-	"Russian" "Urdu" "German" "Swedish" "Afrikaans" "Anglo-Saxon" "Icelandic"
-	"Belarusian" "Bulgarian" "Croatian" "Macedonian" "Serbo-Croatian" "Ukrainian"
-	"Vietnamese" "Pangasinan" "Kapampangan" "Khmer" "Japanese" "Javanese"
-	"Indonesian" "Polish" "Malay" "Sundanese" "Hakka" "Turkish" "Kazakh"
-	"Kirghiz" "Turkmen" "Hebrew" "Persian" "Arabic" "Tajik" "Latin" "French"
-	"Galician" "Romanian" "Norman" "Corsican" "Maori" "Nepali" "Mongolian"
-	"Estonian" "Finnish" "Hungarian" "Bengali" "Waray" "Welsh" "Armenian"
-	"Korean" "Greek" "Catalan" "Walloon" "Yiddish" "Maltese" "Romansh"
-	"Banyumasan" "Võro" "Samogitian" "Chavacano")
-
-declare -a newarr=("Basque" "Dutch" "Alemannic" "Wolof" "Bosnian" "Hindi"
-	"Danish" "Russian" "Urdu" "German" "Swedish" "Afrikaans" "Anglo-Saxon"
-	"Icelandic" "Belarusian" "Bulgarian" "Croatian" "Macedonian" "Serbo-Croatian"
-	"Ukrainian" "Vietnamese" "Pangasinan" "Kapampangan" "Khmer" "Japanese"
-	"Javanese" "Indonesian" "Polish" "Malay" "Sundanese" "Hakka" "Turkish"
-	"Kazakh" "Kirghiz" "Turkmen" "Hebrew" "Persian" "Arabic" "Tajik" "Latin"
-	"French" "Galician" "Romanian" "Norman" "Corsican" "Maori" "Nepali" "Mongolian"
-	"Estonian" "Finnish" "Hungarian" "Bengali" "Waray" "Welsh" "Armenian" "Korean"
-	"Greek" "Catalan" "Walloon" "Yiddish" "Maltese" "Romansh" "Bavarian" "Faroese"
[…]

### 2019-10-28 entropy/ngrams/wiki_extract/cirrus_extract.py

@@ -1,245 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-#
-# =============================================================================
-#  Version: 1.00 (December 15, 2015)
-#  Author: Giuseppe Attardi (attardi@di.unipi.it), University of Pisa
-#
-# =============================================================================
-#  Copyright (c) 2015. Giuseppe Attardi (attardi@di.unipi.it).
-# =============================================================================
-#  This file is part of Tanl.
-#
-#  Tanl is free software; you can redistribute it and/or modify it
-#  under the terms of the GNU General Public License, version 3,
-#  as published by the Free Software Foundation.
-#
-#  Tanl is distributed in the hope that it will be useful,
-#  but WITHOUT ANY WARRANTY; without even the implied warranty of
-#  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-#  GNU General Public License for more details.
-#
-#  You should have received a copy of the GNU General Public License
-#  along with this program.  If not, see <http://www.gnu.org/licenses/>.
-# =============================================================================
-
-"""Wikipedia Cirrus Extractor:
-Extracts and cleans text from a Wikipedia Cirrus dump and stores output in a
-number of files of similar size in a given directory.
-Each file will contain several documents in the format:
-
-    <doc id="" url="" title="">
-        ...
-        </doc>
-
-"""
-
-import sys, os.path, time
-import re
-import json
-import argparse
-import bz2
-import gzip
-import logging
-
-# Program version
-version = '1.00'
-
-urlbase = 'http://it.wikipedia.org/'
-
-# ----------------------------------------------------------------------
-
-class NextFile(object):
-    """
-    Synchronous generation of next available file name.
-    """
-
-    filesPerDir = 100
-
-    def __init__(self, path_name):
-        self.path_name = path_name
-        self.dir_index = -1
-        self.file_index = -1
-
-    def next(self):
-        self.file_index = (self.file_index + 1) % NextFile.filesPerDir
-        if self.file_index == 0:
-            self.dir_index += 1
-        dirname = self._dirname()
-        if not os.path.isdir(dirname):
-            os.makedirs(dirname)
-        return self._filepath()
-
-    def _dirname(self):
-        char1 = self.dir_index % 26
-        char2 = self.dir_index / 26 % 26
-        return os.path.join(self.path_name, '%c%c' % (ord('A') + int(char2), ord('A') + int(char1)))
-
-    def _filepath(self):
[…]

### 2019-10-28 ngrams/get_wiki_df.py

@@ -1,4 +1,3 @@
-# -*- coding: utf-8 -*-
 import regex as re
 import numpy as np
 import pandas as pd

### 2019-10-28 previous/cogsci/scripts/plot_slopes.R

@@ -1,32 +1,0 @@
-library(tidyverse)
-library(here)
-
-data <- read_csv(here("../wiki_entropy/Wikipedia_relative5.csv")) %>%
-  gather(position, slope, Slope1:Slope5) %>%
-  mutate(position = gsub("[^0-9]","", position),
-         position = as.numeric(position))
-
-languages <- distinct(data, Language)
-
-sub_data <- data %>%
-  filter(Language %in% (slice(languages, 1:10) %>% pull(Language))) %>%
-  spread(position, slope) %>%
-  mutate(value0 = 0,
-         value1 = value0 + `1`,
-         value2 = value1 + `2`,
-         value3 = value2 + `3`, 
-         value4 = value3 + `4`,
-         value5 = value4 + `5`) %>%
-  select(-`1`:-`5`) %>%
-  gather(position, value, value0:value5) %>%
-  mutate(position = gsub("[^0-9]","", position),
-         position = as.numeric(position)) %>%
-  group_by(Language) %>%
-  mutate(mid_value = value[3]) %>%
-  mutate(centered_value = value - mid_value)
-
-
-ggplot(sub_data,
-       aes(x = position, y = centered_value, color = Language)) + 
-  geom_line() + 
-  theme_classic()

### 2019-10-28 surprisal/Tiny Corpus/tiny_corpus.R

@@ -1,79 +1,0 @@
-library(tidyverse)
-library(tidytext)
-
-# a1 <- c("a", "cup", "broke")
-# a2 <- c("the", "cup", "broke")
-# a3 <- c("a", "cup", "smashed")
-# a4 <- c("the", "glass", "broke")
-# a5 <- c("a", "glass", "broke")
-# tiny_corpus <- as.data.frame(rbind(a1, a2, a3, a4, a5))
-# row.names(tiny_corpus) <- 1:5
-# names(tiny_corpus) <- c("first", "second", "third")
-
-a1 <- c("a cup broke")
-a2 <- c("the cup broke")
-a3 <- c("a cup smashed")
-a4 <- c("the glass broke")
-a5 <- c("a glass broke")
-tiny_corpus <- as.data.frame(rbind(a1, a2, a3, a4, a5))
-names(tiny_corpus) <- c("text")
-tiny_corpus$text <- as.character(tiny_corpus$text)
-row.names(tiny_corpus) <- 1:5
-tiny_corpus <- tiny_corpus %>%
-  unnest_tokens(word, text, drop=F)
-
-tiny_surprisals <- tiny_corpus %>% 
-  group_by(word) %>% 
-  count() %>% 
-  ungroup() %>% 
-  mutate(p = n / sum(n)) %>% 
-  mutate(s = -log(p)) 
-
-tiny_means <- tiny_corpus %>% 
-  group_by(text) %>%
-  mutate(word_order = 1:n()) %>%
-  ungroup() %>%
-  left_join(tiny_surprisals) %>% 
-  group_by(word_order) %>%
-  summarise(mean_s = mean(s)) 
-
-tiny_means %>% 
-  ggplot(aes(x = word_order, y = mean_s)) + 
-    geom_point() + 
-    geom_line() +
-    xlab("Word order") + 
-    ylab("Mean surprisal") + 
-    ylim(0, 5)
-
-
-
-tiny_bigrams_prep <- tiny_corpus %>% 
-  group_by(text) %>%
-  mutate(lag_word = lag(word)) %>%
-  group_by(lag_word, word) %>%
-  count() %>%
-  filter(!is.na(lag_word)) %>%
-  ungroup() %>%
-  mutate(p = n/sum(n)) %>%
-  left_join(tiny_surprisals, by = c("lag_word" = "word")) %>%
-  mutate(cond_p = n.x / n.y) %>%
[…]

### 2019-10-28 surprisal/Wikipedia/test_wiki.R

@@ -1,111 +1,0 @@
-library(tidyverse)
-library(tidytext)
-library(directlabels)
-# library(googlesheets)
-
-extract_wiki_tokens <- function(min_length, max_length, utterances) {
-  
-  tokens <- utterances %>%
-    mutate(gloss = str_trim(gloss)) %>%
-    mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
-    filter(length >= min_length) %>%
-    filter(length <= max_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    group_by(length, word_order, word) %>%
-    summarise(n = n()) %>%
-    ungroup()
-  
-  sups <- tokens %>%
-    mutate(p = n/sum(n)) %>%
-    mutate(s = -log(p)) %>%
-    select(word, s)
-  
-  tokens %>%
-    left_join(sups)
-}
-
-get_quantiles <- function(df, num_sections) {
-  quantile(df$word_order, probs = seq(0, 1, 1/num_sections)) %>%
-    round() %>%
-    as_data_frame() %>%
-    rename(word_order = value)
-}
-
-relative_slopes <- function(df_tokens, pos_list) {
-  
-  get_slope <- function(x) {
-    
-    selected_pos <- pos_list %>%
-      group_by(length) %>%
-      slice(x:(x+1)) %>%
-      group_by(length) %>%
-      summarise(min = min(word_order),
-                max = max(word_order)) %>%
-      split(.$length) %>%
-      map(~seq(.x$min, .x$max, 1) %>% as_data_frame) %>%
-      bind_rows(.id = "length") %>%
-      rename(word_order = value) %>%
-      mutate(length = as.numeric(length))
-    
-    df_tokens %>%
-      inner_join(selected_pos, by = c("length", "word_order")) %>%
-      group_by(length, word_order) %>%
-      summarise(s = mean(s)) %>%
-      lm(s ~ length + I(length^2) + word_order, data  = .) %>%
-      tidy() %>%
-      filter(term == "word_order")
-  }
-  
-  divides <- pos_list %>%
[…]

### 2019-10-28 surprisal/Wikipedia/wiki_extract/cirrus_extract.py

@@ -1,246 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-#
-# =============================================================================
-#  Version: 1.00 (December 15, 2015)
-#  Author: Giuseppe Attardi (attardi@di.unipi.it), University of Pisa
-#
-# =============================================================================
-#  Copyright (c) 2015. Giuseppe Attardi (attardi@di.unipi.it).
-# =============================================================================
-#  This file is part of Tanl.
-#
-#  Tanl is free software; you can redistribute it and/or modify it
-#  under the terms of the GNU General Public License, version 3,
-#  as published by the Free Software Foundation.
-#
-#  Tanl is distributed in the hope that it will be useful,
-#  but WITHOUT ANY WARRANTY; without even the implied warranty of
-#  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-#  GNU General Public License for more details.
-#
-#  You should have received a copy of the GNU General Public License
-#  along with this program.  If not, see <http://www.gnu.org/licenses/>.
-# =============================================================================
-
-"""Wikipedia Cirrus Extractor:
-Extracts and cleans text from a Wikipedia Cirrus dump and stores output in a
-number of files of similar size in a given directory.
-Each file will contain several documents in the format:
-
-    <doc id="" url="" title="">
-        ...
-        </doc>
-
-"""
-
-import sys, os.path, time
-import re
-import json
-import argparse
-import bz2
-import gzip
-import logging
-
-# Program version
-version = '1.00'
-
-urlbase = 'http://it.wikipedia.org/'
-
-# ----------------------------------------------------------------------
-
-class NextFile(object):
-    """
-    Synchronous generation of next available file name.
-    """
-
-    filesPerDir = 100
-
-    def __init__(self, path_name):
-        self.path_name = path_name
-        self.dir_index = -1
-        self.file_index = -1
-
-    def next(self):
-        self.file_index = (self.file_index + 1) % NextFile.filesPerDir
-        if self.file_index == 0:
-            self.dir_index += 1
-        dirname = self._dirname()
-        if not os.path.isdir(dirname):
-            os.makedirs(dirname)
-        return self._filepath()
-
-    def _dirname(self):
-        char1 = self.dir_index % 26
-        char2 = self.dir_index / 26 % 26
-        return os.path.join(self.path_name, '%c%c' % (ord('A') + int(char2), ord('A') + int(char1)))
-
-    def _filepath(self):
[…]

### 2019-10-28 surprisal/Wikipedia/wiki_relative_surprisal.R

@@ -1,128 +1,0 @@
-library(tidyverse)
-library(tidytext)
-library(directlabels)
-# library(googlesheets)
-
-extract_wiki_tokens <- function(min_length, max_length, utterances) {
-
-  tokens <- utterances %>%
-    mutate(gloss = str_trim(gloss)) %>%
-    mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
-    filter(length >= min_length) %>%
-    filter(length <= max_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    group_by(length, word_order, word) %>%
-    summarise(n = n()) %>%
-    ungroup() %>%
-    mutate(word = as.numeric(as.factor(word)))
-
-  sups <- tokens %>%
-    count(word) %>% 
-    mutate(p = n/sum(n)) %>%
-    mutate(s = -log(p)) %>%
-    select(word, s)
-
-  tokens %>%
-    left_join(sups)
-}
-
-get_quantiles <- function(df, num_sections) {
-  quantile(df$word_order, probs = seq(0, 1, 1/num_sections)) %>%
-    round() %>%
-    as_data_frame() %>%
-    rename(word_order = value)
-}
-
-relative_slopes <- function(df_tokens, pos_list) {
-
-  get_slope <- function(x) {
-
-    selected_pos <- pos_list %>%
-      group_by(length) %>%
-      slice(x:(x+1)) %>%
-      group_by(length) %>%
-      summarise(min = min(word_order),
-                max = max(word_order)) %>%
-      split(.$length) %>%
-      map(~seq(.x$min, .x$max, 1) %>% as_data_frame) %>%
-      bind_rows(.id = "length") %>%
-      rename(word_order = value) %>%
-      mutate(length = as.numeric(length))
-
-    df_tokens %>%
-      inner_join(selected_pos, by = c("length", "word_order")) %>%
-      group_by(length, word_order) %>%
-      summarise(s = mean(s)) %>%
-      lm(s ~ length + I(length^2) + word_order, data  = .) %>%
-      tidy() %>%
[…]

### 2019-10-28 surprisal/Wikipedia/wiki_script.sh

@@ -1,39 +1,0 @@
-#!/bin/bash
-
-declare -a arr=("Amharic" "Armenian" "Bashkir" "Bhojpuri" "Breton" "Burmese"
-	"Catalan" "Chuvash" "Czech" "Danish" "Estonian" "Finnish" "French" "Georgian"
-	"German" "Gujarati" "Hausa" "Hungarian" "Indonesian" "Irish" "Japanese" "Kabyle"
-	"Kannada" "Karachay-Balkar" "Korean" "Lao" "Latvian" "Lezgian" "Lithuanian"
-	"Maithili" "Malayalam" "Maori" "Marathi" "Ossetian" "Polish" "Romanian"
-	"Romansh" "Sakha" "Samoan" "Somali" "Sundanese" "Swahili" "Tajik" "Tamil"
-	"Tatar" "Telugu" "Tetum" "Thai" "Tulu" "Turkish" "Turkmen" "Welsh" "Yoruba")
-
-declare -a fullarr=("Basque" "Dutch" "Alemannic" "Wolof" "Bosnian" "Hindi"
-	"Danish" "Russian" "Urdu" "German" "Swedish" "Afrikaans" "Anglo-Saxon"
-	"Icelandic" "Belarusian" "Bulgarian" "Croatian" "Macedonian" "Serbo-Croatian"
-	"Ukrainian" "Vietnamese" "Pangasinan" "Kapampangan" "Khmer" "Japanese"
-	"Javanese" "Indonesian" "Polish" "Malay" "Sundanese" "Hakka" "Turkish"
-	"Kazakh" "Kirghiz" "Turkmen" "Hebrew" "Persian" "Arabic" "Tajik" "Latin"
-	"French" "Galician" "Romanian" "Norman" "Corsican" "Maori" "Nepali" "Mongolian"
-	"Estonian" "Finnish" "Hungarian" "Bengali" "Waray" "Welsh" "Armenian" "Korean"
-	"Greek" "Catalan" "Walloon" "Yiddish" "Maltese" "Romansh" "Bavarian" "Faroese"
-	"Ripuarian" "Luxembourgish" "Limburgish" "Scots" "Zeelandic" "Aragonese"
-	"Asturian" "Emilian-Romagnol" "Extremaduran" "Franco-Provençal" "Friulian"
[…]

### 2019-10-28 old/cogsci/position-info.Rmd
reorganized master branch

--- 2019-02-02 cogsci/position-info.Rmd
@@
     mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, stem, token = stringr::str_split, pattern = " +") %>%
+    unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
     group_by(utterance_id) %>%
@@
     mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
+    unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
     group_by(utterance_id) %>%

### 2019-10-28 cogsci/scripts/process_bnc_switchboard.R

@@ -1,44 +1,0 @@
-library(here)
-library(feather)
-library(tidyverse)
-library(tidyboot)
-library(entropy)
-library(stringr)
-library(tidytext)
-
-sb_tokens <- read_feather(here("../switchboard/switchboard.feather")) %>%
-  mutate(length = str_count(value, pattern = " +") + 1) %>% 
-  mutate(utterance_id = 1:n()) %>%
-  unnest_tokens(word, value, token = stringr::str_split, pattern = " +") %>%
-  group_by(utterance_id) %>%
-  mutate(word_order = 1:n()) 
-
-bnc10_tokens <- read_feather(here("../bnc/bnc10.feather")) %>%
-  mutate(word = as.numeric(as.factor(word)))
-bnc20_tokens <- read_feather(here("../bnc/bnc20.feather")) %>%
-  mutate(word = as.numeric(as.factor(word)))
-bnc30_tokens <- read_feather(here("../bnc/bnc30.feather")) %>%
-  mutate(word = as.numeric(as.factor(word)))
-
-unigram_entropy <- function(sen_length, tokens) {
-  tokens %>%
-    filter(length == sen_length) %>%
-    group_by(word_order, word) %>%
-    summarise(n = n()) %>%
-    tidyboot(summary_function = function(x) x %>% 
-               summarise(entropy = entropy(n, unit = "log2")),
-             statistics_functions = function(x) x %>%
-               summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
-    mutate(length = sen_length)
-}
-
-switchboard_entropies <- map_df(2:10, ~unigram_entropy(.x, sb_tokens)) 
-
-bnc10_entropies <- unigram_entropy(10, bnc10_tokens)
-bnc20_entropies <- unigram_entropy(20, bnc20_tokens)
-bnc30_entropies <- unigram_entropy(30, bnc30_tokens)
-
-bnc_entropies <- bind_rows(bnc10_entropies, bnc20_entropies, bnc30_entropies)
-
-write_csv(bnc_entropies, here("data/bnc_entropies.csv"))
-write_csv(switchboard_entropies, here("data/switchboard_entropies.csv"))

### 2019-10-28 entropy/Wikipedia/wiki_absolute.R

@@ -1,170 +1,0 @@
-library(knitr)
-library(tidyverse)
-library(directlabels)
-library(dplyr)
-library(tidytext)
-library(entropy)
-library(tidyboot)
-library(tokenizers)
-library(googlesheets)
-
-### BEGIN FUNCTIONS
-
-extract_wiki_tokens <- function(min_length, max_length, utterances) {
-
-  tokens <- utterances %>%
-    mutate(gloss = str_trim(gloss)) %>%
-    mutate(length = str_count(gloss, pattern = " +") + 1) %>%
-    filter(length >= min_length) %>%
-    filter(length <= max_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(word, gloss, token = stringr::str_split, pattern = " +") %>%
-    group_by(utterance_id) %>%
-    mutate(word_order = 1:n()) %>%
-    group_by(length, word_order, word) %>%
-    summarise(n = n())
-}
-
-start_slope <- function(df) {
-
- df %>%
-    filter(word_order %in% c(1,2)) %>%
-    sample_frac(1, replace = T) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    spread(word_order, entropy) %>%
-    ungroup() %>%
-    mutate(diff = `2` - `1`) %>%
-    summarise(diff = mean(diff)) %>%
-    pull(diff)
-
-}
-
-second_slope <- function(df) {
-
- df %>%
-    ungroup() %>%
-    mutate(word_order %in% c(2, 3)) %>%
-    filter(word_order > 0) %>%
-    sample_frac(1, replace = T) %>%
-    group_by(length, word_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    spread(word_order, entropy) %>%
-    ungroup() %>%
-    mutate(diff = `3` - `2`) %>%
-    summarise(diff = mean(diff)) %>%
-    pull(diff)
-
-}
-
-#note this measurement only makes sense if utterance length is at least 6
-mid_slope <- function(df) {
-
- df %>%
-    ungroup() %>%
-    mutate(word_order = if_else(word_order == length-2, 2,
-                                if_else(word_order == 3, 1,
-                                        0))) %>%
-    filter(word_order > 0) %>%
-    group_by(length, word_order) %>%
-    sample_frac(1, replace = T) %>%
[…]

### 2019-10-28 entropy/ngrams/wiki_absolute_ngrams.R

@@ -1,182 +1,0 @@
-library(tidyverse)
-library(directlabels)
-library(tidytext)
-library(entropy)
-library(tidyboot)
-library(tokenizers)
-library(googlesheets)
-
-### BEGIN FUNCTIONS
-
-extract_wiki_tokens <- function(min_length, max_length, utterances, N) {
-
-  tokens <- utterances %>%
-    mutate(length = str_count(words, pattern = "[ +]+") + 1) %>%
-    filter(length >= min_length) %>%
-    filter(length <= max_length) %>%
-    mutate(utterance_id = 1:n()) %>%
-    unnest_tokens(gram, words, token = stringr::str_split, pattern = "[ +]+") %>%
-    group_by(utterance_id) %>%
-    mutate(gram_order = 1:n())
-
-
-  lags <- expand.grid(lag_n = 1:(N-1),
-                      utterance_id = unique(tokens$utterance_id))
-
-  tokens %>%
-    left_join(lags, by = "utterance_id") %>%
-    arrange(utterance_id, lag_n, gram_order) %>%
-    mutate(lag = lag(gram, n = first(lag_n))) %>%
-    drop_na() %>%
-    mutate(gram = str_c(lag, gram, sep = " ")) %>%
-    rename(gram_order = gram_order) %>%
-    group_by(length, gram_order, gram) %>%
-    summarise(n = n())
-
-}
-
-start_slope <- function(df) {
-
- df %>%
-    filter(gram_order %in% c(2,3)) %>%
-    sample_frac(1, replace = T) %>%
-    group_by(length, gram_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    spread(gram_order, entropy) %>%
-    ungroup() %>%
-    mutate(diff = `3` - `2`) %>%
-    summarise(diff = mean(diff)) %>%
-    pull(diff)
-
-}
-
-second_slope <- function(df) {
-
- df %>%
-    ungroup() %>%
-    mutate(gram_order %in% c(3, 4)) %>%
-    filter(gram_order > 0) %>%
-    sample_frac(1, replace = T) %>%
-    group_by(length, gram_order) %>%
-    summarise(entropy = entropy(n, unit = "log2")) %>%
-    spread(gram_order, entropy) %>%
-    ungroup() %>%
-    mutate(diff = `4` - `3`) %>%
-    summarise(diff = mean(diff)) %>%
-    pull(diff)
-
-}
-
[…]

### 2019-10-28 entropy/plot_slopes.R

@@ -1,31 +1,0 @@
-library(tidyverse)
-
-data <- read_csv("slopes.csv") %>%
-  gather(position, slope, Slope1:Slope5) %>%
-  mutate(position = gsub("[^0-9]","", position),
-         position = as.numeric(position))
-
-languages <- distinct(data, Language)
-
-sub_data <- data %>%
-  filter(Language %in% (slice(languages, 1:10) %>% pull(Language))) %>%
-  spread(position, slope) %>%
-  mutate(value0 = 0,
-         value1 = value0 + `1`,
-         value2 = value1 + `2`,
-         value3 = value2 + `3`, 
-         value4 = value3 + `4`,
-         value5 = value4 + `5`) %>%
-  select(-`1`:-`5`) %>%
-  gather(position, value, value0:value5) %>%
-  mutate(position = gsub("[^0-9]","", position),
-         position = as.numeric(position)) %>%
-  group_by(Language) %>%
-  mutate(mid_value = value[3]) %>%
-  mutate(centered_value = value - mid_value)
-
-
-ggplot(sub_data,
-       aes(x = position, y = centered_value, color = Language)) + 
-  geom_line() + 
-  theme_classic()

### 2019-10-28 ngrams/get_wiki_df.py

@@ -1,97 +1,0 @@
-# -*- coding: utf-8 -*-
-import regex as re
-import numpy as np
-import pandas as pd
-import csv
-import sys
-import json
-import subprocess
-import urllib.request
-
-
-with open("language_dict.json", 'r') as f:
-	file = f.read()
-LANG_DICT = json.loads(file) #maps languages to their wikipedia url prefixes
-
-
-def extract_texts(upper_bound):
-	'''
-	Given an upper_bound on file numbers, extracts the raw text from the files
-	in the text/AA folder. That folder is given as output by the
-	cirrus_extract program.
-	'''
-	rs = ""
-	for i in range(upper_bound):
-		if i < 10:
-			number = "0" + str(i)
-		else:
-			number = str(i)
-		f = open("text/AA/wiki_" + number, 'r')
-		text = f.read()
-		rs += text
-	return rs
-
-
-def get_sen_dict(text):
-	'''
-	Given the output of extract_texts as a string, constructs a
-	dictionary mapping each sentence in the string to its length in words.
-	'''
-	text = re.sub("<doc.+>|</doc>|http:\S+|[-%,;:–'&*#/—“»]|\d+|\(|\)|\[|\]", \
-		"", text)
-	text = re.sub('"', '', text)
-	#just the words from the articles
-	text = text.replace('\n', " ")
-	sens = re.split("[.!?]", text.lower())
-	sen_dict = {}
-	for sen in sens:
-		length = len(re.findall("\w+", sen))
-		if length > 0:
-			sen_dict[sen] = length
-	return sen_dict
-
-
-def get_sen_df(text):
-	'''
-	Given the output of extract_texts as a string, returns a Pandas dataframe
-	of the output of get_sen_dict.
-	'''
-	sd = get_sen_dict(text)
-	s = pd.Series(sd, name = "length")
-	del sd
-	s.index.name = "words"
-	df = s.reset_index() #gives a dataframe with columns "sentence" and "length" containing the sentence
-					 	  #and the sentence length respectively
-	return df
-
-
-def main(lang_name):
-	'''
-	Given the English name of a language, constructs a dataframe with the
-	sentences of a random subset of articles in that language from Wikipedia,
-	and stores the dataframe in a csv in the working directory.
-	'''
-	lang_prefix = LANG_DICT[lang_name]
-	url = "https://dumps.wikimedia.org/other/cirrussearch/20190318/" \
[…]

### 2019-11-05 Paper/info-curves.Rmd
edits to bnc & switchboard in paper

@@
 
-We train an ngram language model on a given corpusm with order set to $1$ for the frequency-model, $2$ for a bigram model and $3$ for a trigram model.
+We train an ngram language model on a given corpus with order set to $1$ for the frequency-model, $2$ for a bigram model and $3$ for a trigram model.
 An ngram language model produces a probability for each continuation of each $n-1$ words it receives as input.
-Models with order greater than $3$ have issues with overfitting to the corpus.
+Models with order greater than $3$ have issues with overfitting to the corpus, while bigram models generally underperform trigram models.
 In our contextual models, we face another issue of overfitting: we only train our model on those utterances which occur in the corpus and test our model on the same utterances.
@@
 
-We first replicate the @yu2016 result using the surprisal metric in place of the entropy metric.
+We replicate the @yu2016 result using the surprisal metric in place of the entropy metric.
 We use the frequency-based or “contextless” surprisal metric, which determines the average distribution of information based on word frequencies in a corpus.
@@
 
-See in Figure \@ref(fig:bnc_unigrams) a qualitative comparison of the information trajectory they found and the information trajectory we found.
-It is easily observable that the two information trajectories are identical, and we have found the same distribution of information in the BNC as @yu2016 did.
-The first words in written English sentences tend to contain little information, while the final words tend to contain a great deal of information.
+We find the same frequency-based information trajectory as Yu et al. with little information in the first words of utterances and the most information in the final word, see \@ref(fig:bnc_unigrams).
 
@@
 
-However, this only characterizes written English sentences.
-If this information distribution is characteristic more generally of English, then we will find the same distribution for spoken English.
-We use the Switchboard corpus of spoken English telephone conversations between adults to compute the distribution of information across spoken utterances in English.
-The sentences in the BNC, compared to the conversations in Switchboard, feature much longer sentences with larger vocabularies and more complex sentence structures.
-We split the conversations in Switchboard not by sentences, as in the BNC, but by turns.
+However, we have only found an information distribution for written English sentences in a single British English corpus.
+We predict that this frequency-based distribution will also hold for sentences spoken English.
+To show this, we use the Switchboard corpus of spoken American English telephone conversations.
+As a collection of written texts, our subset of the BNC features much longer sentences with a larger vocabulary per text and overall more complex sentence structures than Switchboard.
+We split the conversations in Switchboard not by sentences, as in the BNC, but by conversational turns.
 
-For Switchboard we find the same information trajectory as we did for the BNC, and we find the same distribution of information regardless of utterance length.
-The longer the utterance, the more pronounced the information distribution shape becomes.
-A qualitative comparison of the two trajectories is in Figure \@ref(fig:switchboard_unigrams).
+For Switchboard we find the same frequency-based information distribution as the BNC, as seen in \@ref(fig:switchboard_unigrams).
+Notice in Figure \@ref(fig:switchboard_unigrams) that the longer the utterance length for Switchboard, the more clearly the information distribution for that utterance length characterizes
 

### 2019-11-06 Paper/info-curves.Rmd
methods edits on paper

@@
 @shannon1948 defined information as "the reduction in uncertainty about one variable given that the value of another variable is known".
-We use a metric proposed by Shannon and applied by @levy2008: lexical surprisal, which defines the information in word based on the ratio of possible continuations of the sentence after to before the word is seen.
+We use a metric proposed for the study of information transmission more generally by Shannon and applied to words specifically by @levy2008: lexical surprisal.
+This measure defines the information in word based on the ratio of possible continuations of the sentence after to before the word is seen.
 Equivalently, we can compute surprisal with the predictability of the word, as in the formula below.
@@
 
-$$surprisal(word) = -log P(word)$$
+$$\text{surprisal}(\text{word}) = -\log P(\text{word})$$
 
@@
 However, when reading or listening, people don’t just consider each word as an isolated linguistic signal.
-Instead, listeners use the words they have already heard to predict and decode the word they are currently processing.
+Instead, listeners use the words they have already heard to predict and decode the current word.
 Following this incremental processing paradigm, we can also condition the surprisal of a word in its context.
-In the formula below, widenotes the word currently being read or heard, while wi-1denotes the first word before the current word, wi-2denotes the second word before the current word, and so on.
+In the formula below, widenotes the word currently being read or heard, while $w_{i-1}$ denotes the first word before the current word, $w_{i-2}$ denotes the second word before the current word, and so on.
 
-$$surprisal(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...)$$
+$$\text{surprisal}(w_i|w_{i-1}w_{i-2}...) = -log P(w_i|w_{i-1}w_{i-2}...)$$
 $$= -log \frac{P(w_i,w_{i-1}w_{i-2},...)}{P(w_{i-1}w_{i-2}...)}$$
@@
 
-We train an ngram language model on a given corpus with order set to $1$ for the frequency-model, $2$ for a bigram model and $3$ for a trigram model.
-An ngram language model produces a probability for each continuation of each $n-1$ words it receives as input.
-Models with order greater than $3$ have issues with overfitting to the corpus, while bigram models generally underperform trigram models.
+Ideally, we would like to predict each word using all of the information available.
+For example, in an utterance of twenty words, we would be able to use the previous $19$ words of context to predict the twentieth word.
+However, we would need to train on a corpus of many trillion word tokens to predict with this amount of context, and we want to directly compare how predictable each word is regardless of its position in an utterance.
+Therefore, we use the *Markov assumption*, conditioning our language models' predictions on only a few words of preceding context at most.
+We train ngram language model on each corpus.
+
+$$\text{surprisal}(w_i|w_{i-1}w_{i-2}...) \approx \text{surprisal}(w_i|w_{i-1}w_{i-2})$$
+
+For our frequency-based model, we don't incorporate context and so each ngram has an order of $1$: unigrams.
+To incorporate context into our models, we train bigram and trigram language models, which incorporate one and two words of context for each processed word, respectively.
+Although these models may seem to use an inconsequential amount of context when predicting the next word, bigram and trigram models introduce a great deal of improvement over unigram models across tasks [@chen1999empirical].
+Models with order greater than $3$ have issues with overfitting to the corpus and only predicting observed sequences, often generalizing poorly.
+
 In our contextual models, we face another issue of overfitting: we only train our model on those utterances which occur in the corpus and test our model on the same utterances.
@@
 "Francisco" may be a common word in a corpus, but likely only occurs after "San" as in "San Francisco", so it receives a lower weighting).
-For a full explanation of modified Kneser-Ney smoothing, see @chen1999empirical.
+For a more complete explanation of modified Kneser-Ney smoothing, see @chen1999empirical.
 
@@
 The first word tended to contain little information.
-While the middle words of sentences each had more information than the first word, information was flat and non-increasing across the middle of sentences.
-The final word, they found, contained the most information out of any in the sentence, with a noticeable spike in information.
-They found the same distribution across sentence lengths, from sentences with 15 words to sentences with 45 words.
+While the middle words of sentences each had more information than the first word, they found a flat and non-increasing rate of information transmission across the middle of sentences.
+The final word contained the most, though not most, of the information out of any in the sentence, with a noticeable spike in information.
+They found the same distribution across sentence lengths, from sentences with $15$ words to sentences with $45$ words.
 
